### 2023-01-09 old.qmd
test

I hope that this manuscript will bring you some new insight on our global ocean and th

European governments are no better and the far right is unfortunately getting stronger.

I could not care less about putting some mellow sentences said by some famous people in this work

With this work, I hope to bring something really useful for us be

According to the @United_Nations_Environment_Programme2019-li, we would need to globally reduce carbon emissions by 7.6% in the coming decade to keep global warming below 1.5°C above pre-industrial levels. Funnily enough, we had to wait for a global pandemic to observe the first decrease in carbon dioxide (CO<sub>2</sub>) emissions in decades without achieving more than a 6.4% decrease [@Tollefson2021-bd].

So far, we have been lucky.

We have been lucky that the global ocean acts as a net sink for anthropogenic CO<sub>2</sub>, currently absorbing \~25% of industrial CO<sub>2</sub> emissions each year [@Friedlingstein2021-kc].\
\
How the global ocean is going to respond to climate change and how long his mitigating effect on warming temperatures is going to last are key questions for our generation.

The goal of this work is to contribute to the fundamental knowledge on the marine carbon cycle and also to build new tools to study it. It thus takes fully part in the United Nations Decade of Ocean Science for Sustainable Development that finally shed a light on the global ocean, which represents only 70% of the Earth's surface in the end.

\-\-\-\--

Among those instruments, the Optical Sediment trap (OST) initially developed by @bishop2004 and further improved by [@Estapa2013-gw; @Estapa2017-jr] can be used to quantify the downward flux of particles, both the continuous flux from small particles and the stochastic flux of larger particles [@Claustre2021-tx]. However, the OST can intercept particles from the BGP but also particles carried by the other pumps. In order to better characterize each pump, there is a need to deconvolute the downward carbon flux into its pump component using additional and independent measurements [@Claustre2021-tx].

BBP (that can characterize sinking particles [@Briggs2020-gz]

the nature of particle (large vs small, labile vs refractoryà is of paramount importance to asess the fate of POC transferred to depth by the BCPs (bbp can say of it is refractory, labile, fast-sinking), UVP for the size

floats : BGP : OST + UVP; Diel Migration Pump : UVP + high-frequency sonar (acoustic)

Yamamoto et al.(2018) suggested

that circulation change and lower nutrient supply to the surface ocean will reduce the efficiency of the BCP and be the main driver for reduced future oceanic CO

2 uptake

Therefore, only the excess density $\Delta\rho$ (ie. $\rho -\rho_{0}$), the particle diameter and the porosity control the settling speed if one assumes that changes in dynamic viscosity are negligeable compared to changes from the two other parameters. This assumption is reasonable based on the fact that $\nu$ can roughly vary by a factor 2 [@Sharqawy2010-ig] while $\Delta_{\rho}$ can vary by more than a factor 2 and $d$ is quadratic.

Such ecosystems store carbon in their living biomass and non-living biomass (e.g. litter) where the latter will be stored in sediment for millenia [@Mcleod2011-iu]. I

Food web relationships and their impacts on carbon pumps are illustrated for the EZ in panel (a). Deep food web relationships and their impacts on the remineralization of DIC are depicted in panel (b). Sequestration times for DIC are set depending on the depth at which POC and DOC are exported. Figure taken from \@Siegel2022-wd. Note: This illustration does not take into account the MCP nor the fact that DOC and POC can directly be injected in the mesopelagic zone (see Table 2 in \@Claustre2021-tx). CONSIDER CHANGING IT TO THE ONE FROM IVERSEN 2022 REVIEW

For instance, the production of transparent exopolymers particles (TEP) formed by the exudation from phytoplankton and bacteria [@Turner2015-af] is thought to enhance the sinking velocities by causing the aggregation of slow sinking particles into larger, faster-sinking aggregates as well as reducing their porosity, therefore the drag [@De_La_Rocha2007-yi]. In other cases such as mineral ballasting, the main driver is still debated [@De_La_Rocha2007-yi]. Mineral ballasting consists in the inclusion of biominerals such as opal (i.e. silicate) and calcite or lithogenic material such as dust that are thought to increase $\Delta\rho$ on the one hand and protect them from microbial degradation on the other hand [@Armstrong2001-rz; @Francois2002-uc; @Klaas2002-kn]. Nevertheless, there is not a consensus on that matter because [@Ploug2008-vx; @Ploug2008-qk] did not find significant differences in microbial degradation of non ballasted vs. ballasted material while @Passow2004-re stipulates that biominerals could instead be collected (i.e. scavenged) and transported downwards by sinking aggregates and that their integration could fragment aggregates into smaller but denser particles [@Passow2006-hw] which could either decrease (smaller size) or increase (denser particles) their sinking speed.

```{=html}
<!--#https://github.com/quarto-dev/quarto-cli/discussions/1042-->
<style type="text/css">
caption, .table-caption {
  text-align: left;
}
</style>
```
---
tbl-cap-location: bottom
---

| Pump                   | \% of export | Sequestration potential (% of BGP) |
|------------------------|--------------|------------------------------------|
| Gravitational          | 50 (70)      |                                    |
| Mixed layer            | 1-6 ()       |                                    |
| Eddy subduction        | 1-22 (N/A)   |                                    |
| Large-scale subduction | 1 ()         |                                    |
| Diel migration         | 10-40 ()     |                                    |
| Seasonal migration     | 3-11 (N/A)   |                                    |

: Percentage of export flux and sequestration potential for the BGP and the PIPs. Percentage of export flux are based on models from @Boyd2019-ez and @Nowicki2022-no in brackets, with, respectively, an estimated export flux of \~9 and \~10 Pg C year$^{-1}$. Potentials for carbon sequestration potential are based on the model of @Boyd2019-ez and is quantified based on the product of their carbon-injection rate and the time until remineralized DIC is returned to the surface. {#tbl-PumpSeqPotential}

), knowledge gained be used to reduce uncertainties in contemporary and future estimates of the export and fates of global ocean NPP. The development of a predictive understanding of the export and fate of global ocean primary production remains among the hardest problems in all of the Earth Sciences, as it requires a synthesis of ocean ecological, biogeochemical, physical, and optical oceanographic processes over an extensive range of time and space scales. . Conducting a field program like EXPORT will accelerate our knowledge of the role of the oceanic food web in the global carbon cycle and provide new models for understanding contemporary and future states of the ocean's carbon cycle and its influences on climate. These results will have tangible societal relevance, leading to advancements in our understanding of our changing planet and reductions in our uncertainties for monitoring its present conditions and for predicting its future state, it will improve our understanding of global ocean carbon dynamics and reduce uncertainties in our ability to monitor and predict carbon export and its sequestration within the ocean's interior.

Question 1 : UVP with Paper of boyd : For all PIPs, the most pressing research issue---needed to address double-accounting issues and improve estimates of carbon sequestra- tion---is to better understand the mechanisms of particle transforma- tions17,65--68

(Fig. 1) within a 4D framework. Specifically, the fate of exported particles between their injection depth and the permanent pycnocline remains poorly constrained. A first step will be improved particle characterization, in particular the ability to distinguish zoo- plankton from other particle types, and to construct aggregate particle size distribution (PSD) profiles through the development and appli- cation of new sensors (Supplementary Table 2). Future developments in acoustic and imaging technologies80

must be deployed on a range of platforms from ships (that is, to provide data for sensor calibra- tion purposes) to an array of long-lived (that is, years), geographically diverse BGC-Argo floats. These developments towards the improve- ment of particle characterization will reduce the likelihood of dou- ble-accounting.

Question 2 : Tjrs boyd (et cadre de REFINE): The way forward in refining estimates of the contribution of PIPs in closing the ocean carbon budget15--17 with the paper of Legendre -\> note que ce n'est peut-être pas forcément ça, mais plus de la séquestration, et pas de l'export; claustre 2021 : This paper proposes an integrated community

observational framework to investigate the BCPs in the World Ocean to better quantify global estimates of carbon storage in the oceans and the underlying drivers of each BCP.

problème du double accounting (boyd) d'où la nécessité de nouveaux instruments (comme l'UVP?)

Understanding the BCP is important for the evaluation of the efficiency of CO2 CDR strategies and to quantify the ocean's biological pump (and reevaluate its potential for sequestration and carbon removal) [@Siegel2022-wd]

OST and upward-looking cameras have been deployed that enable the quantification of carbon export by sinking particles as well as the composition of the sinking flux (durking 2021, 2022, estap 2017) [@Siegel2022-wd]

For how long the exported carbon will remain removed from the atmosphere remains a central question @Passow2012-zh as well as changes in magnitude and sequestration fluxes to constrain the long-term carbon storage and efficiency of the biological pump. (tjrs Passow 2012à

From steinberg 2017 (2& 3) : 2. Zooplankton roles and food-web controls on remineralization and organic matter degra- dationinthemesopelagiczone,whicharefundamentaltodefiningcarbonfluxattenuation and export variability, need to be better measured and understood.

3\. Significant challenges remain in measuring key carbon flows through zooplankton, such as a standard approach for assessing secondary production, or zooplankton-mediated carbon export not measured by traditional sediment traps (e.g., jelly falls and active transport via vertical migration).

improving the resolution of zooplankton migration dynamics as well as

higher-level trophic interactions in the marine food web. Improvements in observational datasets would help to better constrain uncertain aspects of the model. In particular, observations of phytoplankton and zooplankton biomass, as well as direct measurements of carbon flux due to the different biological pump pathways in a range of environments, would provide improved constraints to this model and enhance the accuracy of our estimates. --\> faire ça est équivalent à dire que si on peut avoir des meilleures mesures, on peut mieux contraindre les modèles (logique..) et du coup améliorer notre compréhension de la BCP.

3 research questions in https://www.nature.com/articles/d41586-020-00915-7

aller chercher de l'inspi sur des questions de recherche ici ? https://www.frontiersin.org/articles/10.3389/fmars.2016.00022/full (siegel 2016)

For instance, the importance of the types, transport pathways, particles associated communities, and fates of sinking particles have become apparent thanks to technological breakthroughs, and thus need further research. \[Le Moigne 2019\]

method to clean the signal from non-related flux events (e.g. spikes from transient particles not landing on the transmissometer window) and divide the cleaned attenuation signal into a continuous component, that is the continuous accumulation of small particles on the transmissometer window and a discontinuous signal attributed to

\-\--

The first BGC-Argo floats equipped with transmissometers and UVP6 with automatic classification of particles were deployed in the Labrador Sea in late May 2022. This is the beginning of the deployment phase of the REFINE project (<a>https://erc-refine.eu</a>) which intend to deploy floats in nine oceanic regions representative of the diversity of environmental conditions.

\-\--

### 2023-01-22 article3.qmd
start project

# Case study in the Labrador Sea {#case-study-labrador-sea}

## Introduction

introduire la notion de cycle de flotteurs BGC-Argo etc (+ parking depth)

## Methods

### Deployment and floats configuration

Two couples of BGC-Argo floats were deployed in the Labrador Sea in late May 2022. The first couple (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second (WMO 6904241 and 1902578) was deployed on the 30$^{th}$ at coordinates (-52.4°E, 56.8°N).

All 4 floats used in this study are PROVOR Jumbo floats (NKE Instrumentation, France) equipped with a CTD SBE41-CP (Sea-Bird Scientific, USA), a UVP6-LP (Hydroptic, France), a c-Rover 660 nm beam transmissometer (WETLabs, USA) and a ECO FLBBCD (WETLabs, USA) that includes a chlorophyll-a (Chla) fluorometer (excitation at 470 nm, emission at 695 nm), a colored dissolved organic matter (CDOM) fluorometer (excitation at 370 nm, emission at 460 nm) and a 700 nm particle backscatteringmeter ($b_{bp}$). All floats but one (1902578) were equipped with a OCR-504 Multispectral Radiometer to measure the photosynthetic available radiation.

All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux. During these drift phases, measurements of particles abundance (between 0.6 mm - 2.5 mm) are taken every 30 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths. This configuration can be adjusted during the mission of the float via two-way Iridium communication.

```{r, init function}
library(akima)
library(broom)
library(tidyverse)
library(cmocean)
library(lubridate)
library(scales)
library(patchwork)
library(viridis)
library(ggridges)
library(ggvis)
library(pracma)
library(gbm)
library(MLmetrics)
library(purrr)
library(arrow)
library(plotly)
library(oce)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(RColorBrewer)
library(vroom)
library(castr)
```

```{r, map of 4 floats}
#| label: fig-float-position
#| fig-cap: Route of each float since their deployment (black triangles). Dots represent ascending profiles. The profile that concludes the time series for each float is highlighted by a black contour.
#library(tidyverse)

# read floats data
float_6904240 <- vroom('/home/flo/dox/ArgoShine/6904240_FromNetCDF.csv')
float_6904241 <- vroom('/home/flo/dox/ArgoShine/6904241_FromNetCDF.csv')
float_4903634 <- vroom('/home/flo/dox/ArgoShine/4903634_FromNetCDF.csv')
float_1902578 <- vroom('/home/flo/dox/ArgoShine/1902578_FromNetCDF.csv')

all_floats <- rbind(float_6904240, float_6904241, float_4903634, float_1902578)

argo_map <- all_floats |> select(WMO = wmo, cycle, lat, lon) |> drop_na(lat) |> dplyr::group_by(WMO, cycle) |> mutate(WMO = factor(WMO)) |>
  dplyr::summarise(lat = mean(lat), lon = mean(lon))

argo_map_first <- argo_map |> filter(cycle == 1)
argo_map_last <- argo_map |> filter(cycle == max(cycle))

world_data <- map_data("world")
w <- list(geom_polygon(aes(long, lat, group=group), data=world_data, fill="grey70"), coord_quickmap())

ggplot(argo_map) + w + geom_path(aes(lon, lat, colour = WMO), size = 1) +
  geom_point(aes(lon, lat, colour = WMO), size = 1.5) +
  scale_color_brewer(palette = 'Dark2') +
  theme_bw() +
  #scale_colour_manual(values = c('#a6cee3', '#33a02c', '#b2df8a', '#1f78b4')) +
  coord_cartesian(xlim = c(-55, -40), ylim = c(55,65)) + geom_point(data = argo_map_first, aes(lon, lat), shape = 17, size = 3) +
  geom_point(data = argo_map_last, aes(lon, lat, colour = WMO), size = 3) +
  geom_point(data = argo_map_last, aes(lon, lat, fill = WMO), size = 3, shape = 1) +
  labs(x = 'longitude', y = 'latitude') +
  #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
  guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
                                                  shape = c(NA, NA, NA, NA))))

```

### Data processing

Data were downloaded from the Coriolis data center (biogeochemical data at ftp://ftp.ifremer.fr/ifremer/argo/dac/coriolis/ and UVP6 data at ftp://ftp.ifremer.fr/ifremer/argo/aux/coriolis/, last access: MONTH DAY, 2023). Chla data were quality controlled for Non-Photochemical Quenching (NPQ), a photoprotective mechanism that decreases the apparent fluorescence per unit of Chla following the method implemented by @Terrats2020 who modified the method of @Xing2018 for shallow-mixing cases. $c_{p}$ data were converted from counts to physical units (m$^{-1}$) using the following relationship

$$
c_{p}~[m^{-1}] = -\frac{1}{x}~log \left(~ \frac{c_{p}~[counts]-CSC_{dark}}{CSC_{cal}-CSC_{dark}} \right)
$$

where $x$, $CSC_{dark}$ and $CSC_{cal}$ are respectively the transmissometer pathlength (0.25 m), the sensor output when the beam is blocked (i.e. offset, 0 for each c-Rover of this study) and the sensor output with clean water in the path (12998, 13598, 13115 and 12855 for, respectively, float 4903634, 6904240, 6904241 and 1902578). The latter were obtained from the calibration sheets of each c-Rover.

UVP6 data and in particular the averaged number of particles per size class (i.e. particle size distribution hereafter referred to as PSD) was divided by the number of images at each depth and the water volume imaged by the UVP (0.7L) to obtain particle concentrations (units: #/L). Note that in an impending update by Coriolis, it will not be necessary to divide by the number of images anymore.

All data were divided into vertical profiles and drifting data using time stamps for drift. Finally, data were checked for instrumental anomalies. As a result, $c_{p}$ values from float 6904241 cannot be used from its 20$^{th}$ cycle onwards (\> November 17) due to a likely failure of the pump cleaning the optical window of the transmissometer.

### Derivation of carbon fluxes from OST

We applied a similar approach to @Estapa2013-gw and @Estapa2017-jr to derive the continuous component (i.e. continuous deposition of small particles on the upward-facing window of the transmissometer) and discontinuous (i.e. deposition of stochastic big particles) components of the downward POC flux.

First, the drifting $c_{p}$ signal was cleaned by removing spikes (hypothesized as transient particles not landing on the window or active swimmers [@Estapa2013-gw]) using the despike function (oce package in R, see https://dankelley.github.io/oce) where a spike is defined here as any value higher than 4 times the standard deviation of the difference between raw $c_{p}$ data and a smoothed (i.e. reference) $c_{p}$ signal, obtained with a 7-point running median filter. Despiked $c_{p}$ data were then smoothed with a 3-point running median filter to ease the detection of jumps, defined as a continuous increase or decrease of $c_{p}$ (@fig-estapa-method_article).

```{r, fig fit}
#| label: fig-estapa-method_article
#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component and a discontinuous (positive jumps, red) component. Slopes for each group are depicted in light blue. A negative jump was also detected (yellow). Data from BGC-ARGO float WMO 6904240, drifting at 1000 m during its 21$^{st}$ cycle.
clean_cp_data <- function(data, moving_median_order){

  if(nrow(data) == 0){ # no cp data
    return(data)
  }else{

    # # Remove cp > 2 m-1 (don't really know why but eh)
    # if (any(data$cp > 2)){
    #   data <- data[-which(data$cp > 2),]
    # }

    # despike cp data
    data$cp <- oce::despike(data$cp, reference = 'median', k = 7, replace = 'NA')

    # median average
    data$cp <- pastecs::decmedian(data$cp, order = moving_median_order, ends = "fill") |> pastecs::extract(component = "filtered")
  }
  return(data)

}

tmp <- all_floats |> filter(wmo == 6904240, park_depth == '1000 m', cycle == 21, PhaseName == 'PAR') |> drop_na(cp) |> select(dates = juld, everything())

threshold <- 0.08

clean_data <- clean_cp_data(tmp, 3)

# save cleaned Cp signal from spikes to a temporary variable
tmp2 <- clean_data
tmp2$jump <- NA
# compute slope between two adjacent points (except first and last point of drifting time series)
for (i in 2:(nrow(tmp2)-1)){
  #print(i)
  delta_x <- as.numeric(difftime(tmp2$dates[i], tmp2$dates[i-1], units = 'days'))
  delta_y <- tmp2$cp[i]-tmp2$cp[i-1]
  tmp2$jump[i] <- delta_y/delta_x
}

# assign a colour for the plot
tmp2$colour <- 'base signal'
tmp2$colour[which(abs(tmp2$jump)>threshold)] <- 'jump' # abs is needed for 'negative jumps'

# add group to compute slope (for each subgroups of points, separated from a jump)
tmp2$group <- NA

# compute where the jumps are (index value of the array)
jump_index <- which(tmp2$colour == 'jump')

# assign group 
for (i in jump_index){
  for (j in 1:nrow(tmp2)){
    if ((j < i) & (is.na(tmp2$group[j]))){
      tmp2$group[j] <- paste0('group_',i)
    }
  }
}

# add last group (because it is AFTER the last index so it was not computed before)
tmp2$group[which(is.na(tmp2$group))] <- 'last_group'

# compute slope for each subgroups
slope_df <- tmp2 |> filter(colour == 'base signal', jump != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),delta_y = (last_cp-first_cp)*0.25, slope = delta_y/delta_x)

# remove negative slope from the mean slope (no physical meaning)
slope_df <- slope_df |> filter(slope > 0)

# remove if only one point (cannot fit a slope with one point)
slope_df <- slope_df |> filter(nb_points > 1)

# compute mean slope 
#mean_slope <- mean(slope_df$slope)
mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)

# estapa relationship (see POSTER)
#poc_flux <- 633*((mean_slope*0.25)**0.77) # *0.25 because ATN = cp*0.25
poc_flux <- 633*((mean_slope)**0.77) # /!\ slope for computed for ATN on y axis (delta_y *0.25 because ATN  = cp*0.25) -> should be OK

# build dataframe to plot each subgroup
part1 <- slope_df |> select(group, time = min_time, cp = first_cp)
part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
part_slope <- rbind(part1, part2)

# spot negative jump
tmp2$colour[which((tmp2$colour == 'jump') & (tmp2$jump < 0))]  <- 'negative jump'

# add big particles flux
rows_to_keep <- c(jump_index, jump_index-1) 
tmp3 <- tmp2[rows_to_keep,] |> select(dates, cp, jump, colour, group) |> dplyr::arrange(dates)

# remove negative jump, if any
check_colour <- unique(tmp3$colour)
if(length(check_colour) == 3){
  index_neg_jump <- which(tmp3$colour == 'negative jump')
  tmp3 <- tmp3[-c(index_neg_jump, index_neg_jump+1),]
  tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp)) 
  even_indexes <- seq(2,nrow(tmp3),2)
  tmp3 <- tmp3[even_indexes,]
}else if(length(check_colour) == 2){
  tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp)) 
  even_indexes <- seq(2,nrow(tmp3),2)
  tmp3 <- tmp3[even_indexes,]
}else{ # NO jump
  tmp3 <- NULL
  }

# big particles flux
if(is.null(tmp3)){ # no jumps
  big_part_poc_flux <- 0
}else{
  tmp4 <- tmp3 |> filter(diff_jump > 0)
  if(nrow(tmp4) == 0){ # no positive jumps
    big_part_poc_flux <- 0
  }else{
    delta_y <- sum(tmp4$diff_jump) *0.25 # to get ATN (= cp*0.25)
    max_time <- max(tmp2$dates)
    min_time <- min(tmp2$dates)
    delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
    slope_big_part <- delta_y/delta_x
    big_part_poc_flux <- 633*(slope_big_part**0.77)
  }
}

# compute total drifting time
max_time <- max(tmp2$dates)
min_time <- min(tmp2$dates)
drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))

check <- tmp2 |> select(dates, cp, colour)
check <- check |> filter(colour %in% c('jump', 'negative jump'))

# ggplot(clean_data, aes(x = dates, y = cp)) + geom_point(size= 3) + geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour), size = 2) + 
#     scale_colour_manual(values = c('#003366','#E31B23', '#FFC325')) + 
#     scale_shape_manual(values = c(1,1,4)) +
#     geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', size = 1) + theme_bw() + 
#     labs(x = 'Time', y = 'Cp (1/m)') +
#   geom_point(data = check, aes(x = dates, y = cp, colour = colour))

ggplot(tmp2, aes(x = dates, y = cp, colour = colour)) + geom_point(size = 3) +
    scale_colour_manual(values = c('#003366','#E31B23', '#FFC325')) + 
    #scale_shape_manual(values = c(1,1,4)) +
    geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', size = 2) + theme_bw() + 
    labs(x = 'Time', y = TeX('$c_{p}~(m^{-1})$')) +
  geom_point(data = check, aes(x = dates, y = cp, colour = colour), size = 3) +
  geom_point(data = check, aes(x = dates, y = cp), shape = 1, size = 3, colour = 'black')  +
  #https://r-graph-gallery.com/239-custom-layout-legend-ggplot2.html
  # theme(legend.position = c(0.02,0.99), legend.justification = c('left', 'top'), 
  #       legend.title = element_blank())
    theme(legend.position = 'none')
```

Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach to detect jumps. For this, we applied a 7-point running slope on drifting $c_{p}$ data at each drifting depth and we then computed the 90$^{th}$ quantile ($Q_{90}$) of each slope distribution with a 10-day increasing dataset since the deployment (Fig-XX) to determine the most appropriate slope thresholds to detect jumps. The latter are however subject to a tradeoff between the likely underestimation of $F_{small}$ and the detection of false jumps (without great impact on the value of $F_{big}$ because false jumps have low $c_{p}$ values) if the threshold is too small and the likely overestimation of $F_{small}$ to the detriment of $F_{big}$ with missed jumps if the threshold is too high. As shown in Fig-XX, the value of $Q_{90}$ decreases over time for each drifting depth. This is most likely explained by the vanishing of big particles (> 0.5 mm in this study) over time (Fig-YY). Therefore, we assumed that past a given time, the great majority of jumps have been seen by the transmissometer and that the optimal values for each slope threshold should be found in the period following the bloom up to this given time. We thus have found that on a weekly basis, particles in the 0.5 - 1.5 mm range at 1000 m were starting to steadily decrease (assumed to be slowly removed without upwards production) around week 40, that is the beginning of October. (maybe find another criteria .. thresholds as they currently are give good fit.)

```{r, slope distribution}

# remove bad data from float 6904241
float_6904241_slope <- float_6904241 |> filter(cycle < 20)

# create dataframe for all drifting data (for all floats)
all_floats_slope <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle)

# compute slope for every float, with every size window
wmo <- c(4903634, 6904240, 6904241, 1902578)
park_depth <- c('200 m', '500 m', '1000 m')
#window_size <- c(3,5,7)
window_size <- 7

# function to clean cp data
clean_cp_data <- function(data, moving_median_order){
  if(nrow(data) == 0){ # no cp data
    return(data)
  }else{
    # despike cp data
    data$cp <- oce::despike(data$cp, reference = 'median', k = 7, replace = 'NA')
    # median average
    data$cp <- pastecs::decmedian(data$cp, order = moving_median_order, ends = "fill") |> pastecs::extract(component = "filtered")
  }
  return(data)
}

# compute cp slope
delta_y <- function(x, window_size){
  delta_y <- x[window_size]-x[1]
  return(delta_y)
}

delta_x <- function(x, window_size){
  delta_x <- (x[window_size]-x[1])/86400 # units in days
  return(delta_x)
}

compute_slope <- function(data, window_size){
  tmp <- data |> drop_na(cp)
  despiked_data <- clean_cp_data(tmp, 3) 
  if(window_size == 3){
    despiked_data$slope <- slide(despiked_data$cp, 1, delta_y, window_size = 3)/slide(despiked_data$dates, 1, delta_x, window_size = 3)
  }else if(window_size == 5){
    despiked_data$slope <- slide(despiked_data$cp, 2, delta_y, window_size = 5)/slide(despiked_data$dates, 2, delta_x, window_size = 5)
  }else{
    despiked_data$slope <- slide(despiked_data$cp, 3, delta_y, window_size = 7)/slide(despiked_data$dates, 3, delta_x, window_size = 7)
  }
  return(despiked_data)
}

res <- data.frame()
for (i in wmo){
  #print(i)
  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
  for (j in park_depth){
    for (k in seq(1:max_cycle)){
      #print(k)
      tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
      if(nrow(tmp) == 0){
        next
      }else if(nrow(tmp) < 3){ # case where there is not enough data
        next
      }else{
        cleaned_signal <- clean_cp_data(tmp, 3)
        if(nrow(cleaned_signal) == 0){
          next
        }else{
          for (l in window_size){
            test <- compute_slope(cleaned_signal, l)
            test <- test |> select(dates, wmo, cycle, park_depth, slope)
            # add time_drifting
            max_time <- max(test$dates)
            min_time <- min(test$dates)
            drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))
            test$drifting_time <- drifting_time
            test$window_size <- l
            res <- rbind(res, test)
          } 
        }
      }
    }
  }
}

fixed_dates <- c('2022-06-01', '2022-06-10', '2022-06-20', 
                 '2022-07-01', '2022-07-10', '2022-07-20', 
                 '2022-08-01', '2022-08-10', '2022-08-20', 
                 '2022-09-01', '2022-09-10', '2022-09-20', 
                 '2022-10-01', '2022-10-10', '2022-10-20', 
                 '2022-11-01', '2022-11-10', '2022-11-20', 
                 '2022-12-01', '2022-12-10', '2022-12-20',
                 '2023-01-01')

res2 <- data.frame()
for (i in fixed_dates){
  tmp <- filter(res, dates < i)
  tmp2 <- tmp |> group_by(park_depth, window_size) |> dplyr::summarise(threshold = quantile(slope, 0.90, na.rm = T), nb_points = n())
  tmp2$group <- i
  tmp2$quantile <- 90
  res2 <- rbind(tmp2, res2)
}

# add levels
res2$park_depth <- factor(res2$park_depth, levels = c('200 m', '500 m', '1000 m'))

# plot result
# res3 <- res2 |> mutate(juld = as_date(group)) |> ungroup()
# ggplot(res3, aes(x = juld, y = threshold, colour = factor(window_size), group = factor(window_size))) + geom_point() + geom_path() + 
#   facet_wrap(~park_depth, scales = 'free') + theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   scale_color_brewer(palette = 'Dark2') +
#   labs(x = 'Month', y = 'Slope')
res3 <- res2 |> mutate(juld = as_date(group)) |> ungroup()
ggplot(res3, aes(x = juld, y = threshold)) + geom_point() + geom_path() + 
  facet_wrap(~park_depth, scales = 'free') + theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  geom_vline(xintercept = as_date('2022-10-01'), colour = "red", linetype = 'dashed') +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Month', y = 'Slope')

```

As a result, we have chosen to set the slope thresholds to detect jumps at week 40 for all three parking depths at, in increasing order with depth, 0.23, 0.12 and 0.10. Using those thresholds, we computed the slope between each $c_{p}$ measurement of each drifting periods and identified as jump any point above the selected thresholds. This method allowed us to separate the continuous component of the downward POC flux into groups, between each jump. We then computed the slope for each group (light blue straight lines in @fig-estapa-method_article). Any negative slope was disregarded assuming that either the so-called fit failed (ineffective threshold leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window. Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each group in order to derive the small particles flux ($F_{small}$) using the relationship of @Estapa2022

$$
F_{small} = 633(0.25~slope_{w})^{0.77}
$$ {#eq-estapa}

where $slope_{w}$ is multiplied by the transmissometer pathlength because @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.

The discontinuous component of the downward POC flux or big particle flux ($F_{big}$) was derived using @eq-estapa by replacing $slope_{w}$ as the sum of positive jumps (negative jumps were disgarded) divided by the drifting time.

All profiles analyzed with this method are available at https://fricour.shinyapps.io/CARBON_REFINE/.

## Results

### OST-derived carbon fluxes

Using the selected slope thresholds determined in the previous section, we computed $F_{small}$ and $F_{big}$ at each drifting depth for all floats (@fig-OST-flux). At 200 m, $F_{small}$ ranges between 5.1 and 26.9 mg C m$^{-2}$ day$^{-1}$ with an average of 14.1 $\pm$ 3.9 mg C m$^{-2}$ day$^{-1}$ (standard deviation). At 500 m, $F_{small}$ is comprised between 6.9 and 19.8 mg C m$^{-2}$ day$^{-1}$ with an average of 9.5 $\pm$ 2.1 mg C m$^{2}$ day$^{-1}$. At this depth, we observe a global decrease of $F_{small}$ with time though float 6904240 measured a relatively constant downward $F_{small}$ at 8.6 mg C m$^{-2}$ day$^{-1}$ on average. At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid-July followed by a steady decrease towards December, with $F_{small}$ ranging between 4.0 and 11.8 mg C m$^{-2}$ day$^{-1}$ and 7.0 $\pm$ 1.4 mg C m$^{-2}$ day$^{-1}$ on average.

```{r, function plot jump}
# plot_jumps <- function(clean_data, threshold){
#   
#   # save bbp data for later
#   #bbp_df <- clean_data |> select(dates, bbp) |> drop_na(bbp)
#   
#   # save cleaned Cp signal from spikes to a temporary variable
#   tmp2 <- clean_data |> arrange(dates)
#   #print(colnames(tmp2))
#   # init jump values to NA
#   tmp2$jump <- NA
#   # compute slope between two adjacent points (except first and last point of drifting time series)
#   for (i in 2:(nrow(tmp2)-1)){
#     #print(i)
#     delta_x <- as.numeric(difftime(tmp2$dates[i], tmp2$dates[i-1], units = 'days'))
#     delta_y <- tmp2$cp[i]-tmp2$cp[i-1]
#     tmp2$jump[i] <- delta_y/delta_x
#   }
#   
#   # assign a colour for the plot
#   tmp2$colour <- 'base signal'
#   tmp2$colour[which(abs(tmp2$jump)>threshold)] <- 'jump' # abs is needed for 'negative jumps'
#   
#   # add group to compute slope (for each subgroups of points, separated from a jump)
#   tmp2$group <- NA
#   
#   # compute where the jumps are (index value of the array)
#   jump_index <- which(tmp2$colour == 'jump')
#   
#   # assign group 
#   for (i in jump_index){
#     for (j in 1:nrow(tmp2)){
#       if ((j < i) & (is.na(tmp2$group[j]))){
#         tmp2$group[j] <- paste0('group_',i)
#       }
#     }
#   }
#   
#   # add last group (because it is AFTER the last index so it was not computed before)
#   tmp2$group[which(is.na(tmp2$group))] <- 'last_group'
#   
#   # compute slope for each subgroups
#   slope_df <- tmp2 |> filter(colour == 'base signal', jump != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), 
#                                                                                                                   nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],
#                                                                                                                   delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),
#                                                                                                                   delta_y = (last_cp-first_cp)*0.25, slope = delta_y/delta_x)
#   
#   # remove negative slope from the mean slope (no physical meaning)
#   slope_df <- slope_df |> filter(slope > 0)
#   
#   # remove if only one point (cannot fit a slope with one point)
#   slope_df <- slope_df |> filter(nb_points > 1)
#   
#   # compute mean slope 
#   #mean_slope <- mean(slope_df$slope)
#   mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)
#   
#   # estapa relationship (see POSTER)
#   #poc_flux <- 633*((mean_slope*0.25)**0.77) # *0.25 because ATN = cp*0.25
#   poc_flux <- 633*((mean_slope)**0.77) # /!\ slope for computed for ATN on y axis (delta_y *0.25 because ATN  = cp*0.25) -> should be OK
#   
#   # build dataframe to plot each subgroup
#   part1 <- slope_df |> select(group, time = min_time, cp = first_cp)
#   part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
#   part_slope <- rbind(part1, part2)
#   
#   # spot negative jump
#   tmp2$colour[which((tmp2$colour == 'jump') & (tmp2$jump < 0))]  <- 'negative jump'
#   
#   # add big particles flux
#   rows_to_keep <- c(jump_index, jump_index-1) 
#   tmp3 <- tmp2[rows_to_keep,] |> select(dates, cp, jump, colour, group) |> dplyr::arrange(dates)
#   
#   # remove negative jump, if any
#   check_colour <- unique(tmp3$colour)
#   if(length(check_colour) == 3){
#     index_neg_jump <- which(tmp3$colour == 'negative jump')
#     tmp3 <- tmp3[-c(index_neg_jump, index_neg_jump+1),]
#     tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp)) 
#     even_indexes <- seq(2,nrow(tmp3),2)
#     tmp3 <- tmp3[even_indexes,]
#   }else if(length(check_colour) == 2){
#     tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp)) 
#     even_indexes <- seq(2,nrow(tmp3),2)
#     tmp3 <- tmp3[even_indexes,]
#   }else{ # NO jump
#     tmp3 <- NULL
#   }
#   
#   # big particles flux
#   if(is.null(tmp3)){ # no jumps
#     big_part_poc_flux <- 0
#   }else{
#     tmp4 <- tmp3 |> filter(diff_jump > 0)
#     if(nrow(tmp4) == 0){ # no positive jumps
#       big_part_poc_flux <- 0
#     }else{
#       delta_y <- sum(tmp4$diff_jump) *0.25 # to get ATN (= cp*0.25)
#       max_time <- max(tmp2$dates)
#       min_time <- min(tmp2$dates)
#       delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
#       slope_big_part <- delta_y/delta_x
#       big_part_poc_flux <- 633*(slope_big_part**0.77)
#     }
#   }
#   
#   # compute total drifting time
#   max_time <- max(tmp2$dates)
#   min_time <- min(tmp2$dates)
#   drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))
#   
#   jump_plot <- ggplot(clean_data, aes(x = dates, y = cp)) + geom_point(size= 3) + geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour), size = 3) + 
#     scale_colour_manual(values = c('#003366','#E31B23', '#FFC325')) + 
#     scale_shape_manual(values = c(1,1,4)) +
#     geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', size = 2) + theme_bw() + 
#     labs(x = 'Time', y = 'Cp (1/m)', title = paste0('Drifting time: ', round(drifting_time,3), ' days\n',
#                                                     'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n', 
#                                                     'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
#                                                     'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1')) #+
#   #geom_point(data = bbp_df, aes(x = dates, y = bbp), colour = "pink")
#   
#   
#   # adapt script to return big and small flux
#   df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = clean_data$park_depth[1], wmo = clean_data$wmo[1],
#                cycle = clean_data$cycle[1])
#   
#   #return(jump_plot)
#   #return(list('jump_plot' = jump_plot, 'jump_table' = tmp3))
#   return(df)
#   
# }

plot_jumps <- function(data){

  # make sure that the cp signal is in chronological order
  tmp <- data |> arrange(dates)

  # despike cp data with a 7-point moving window
  tmp$cp <- despike(tmp$cp, k = 7)

  # smooth cp data with a 3-point moving median, n time(s)
  tmp$cp <- slide(tmp$cp, fun = median, k = 3, n = 1, na.rm=T)

  # compute slope between two adjacent points (except first point) # we could start after 1h to let the float stabilize
  delta_x <- as.numeric(tmp$dates - lag(tmp$dates), units = 'days')
  delta_y <- tmp$cp - lag(tmp$cp)
  tmp$slope <- delta_y/delta_x

  # compute a Z score (assuming a normal distribution of the slopes) on the slopes
  tmp <- tmp |> mutate(zscore = (slope - mean(slope, na.rm = T))/sd(slope, na.rm = T))

  # spot outliers on the Z score signal
  # interquartile range between Q25 and Q75
  IQR <- quantile(tmp$zscore, probs = 0.75, na.rm=T) - quantile(tmp$zscore, probs = 0.25, na.rm=T)
  # outliers ('spikes' in the Z score signal)
  spikes_down <- tmp$zscore < quantile(tmp$zscore, 0.25, na.rm=T) - 1.5 *IQR
  spikes_up <-  tmp$zscore > quantile(tmp$zscore, 0.75, na.rm=T) + 1.5 *IQR
  spikes <- as.logical(spikes_down + spikes_up)

  # assign spikes
  tmp$spikes <- spikes

  # assign colour code to cp signal
  tmp$colour <- 'base signal' # base signal = smoothed despiked cp signal
  tmp[which(tmp$spikes == TRUE),]$colour <- 'jump'

  # add group to compute the slope of each group of points, separated by a jump
  tmp$group <- NA

  # index of jumps in the array
  jump_index <- which(tmp$colour == 'jump')

  # assign group identity to each group of points, separated by a jump (= subgroup)
  for (i in jump_index){
    for (j in 1:nrow(tmp)){
      if ((j < i) & (is.na(tmp$group[j]))){
        tmp$group[j] <- paste0('group_',i)
      }
    }
  }
  tmp$group[which(is.na(tmp$group))] <- 'last_group'

  # compute slope for each subgroup
  slope_df <- tmp |> filter(colour == 'base signal', slope != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), 
                                                                                                                  nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],
                                                                                                                  delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),
                                                                                                                  delta_y = (last_cp-first_cp)*0.25, slope = delta_y/delta_x) # *0.25 to convert cp to ATN

  # remove negative slope from the mean slope (no physical meaning)
  slope_df <- slope_df |> filter(slope > 0)

  # remove if only one point (cannot fit a slope with one point)
  slope_df <- slope_df |> filter(nb_points > 1)

  # compute weighted average slope (to take into account the fact that some subgroup might have 2 points and a high slope vs. large group of points with a small slope)
  mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)

  # convert cp to POC using Estapa's relationship 
  poc_flux <- 633*((mean_slope)**0.77) # /!\ slope computed for ATN on y axis (delta_y *0.25 because ATN = cp*0.25) -> should be OK

  # build dataframe to plot each subgroup
  part1 <- slope_df |> select(group, time = min_time, cp = first_cp)
  part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
  part_slope <- rbind(part1, part2)

  # spot negative jump
  tmp$colour[which((tmp$colour == 'jump') & (tmp$slope < 0))]  <- 'negative jump'

  # add big particles flux to the party
  rows_to_keep <- c(jump_index, jump_index-1)
  tmp2 <- tmp[rows_to_keep,] |> select(dates, cp, slope, colour, group) |> arrange(dates)

  # remove negative jump, if any
  check_colour <- unique(tmp2$colour)
  if(length(check_colour) == 3){ # there is a negative jump
    index_neg_jump <- which(tmp2$colour == 'negative jump')
    tmp2 <- tmp2[-c(index_neg_jump, index_neg_jump+1),]
    tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
    even_indexes <- seq(2,nrow(tmp2),2)
    tmp2 <- tmp2[even_indexes,]
  }else if(length(check_colour) == 2){ # only positive jump
    tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
    even_indexes <- seq(2,nrow(tmp2),2)
    tmp2 <- tmp2[even_indexes,]
  }else{ # No jump
    tmp2 <- NULL
  }

  if(is.null(tmp2)){ # no jumps
    big_part_poc_flux <- 0
  }else{
    tmp3 <- tmp2 |> filter(diff_jump > 0)
    if(nrow(tmp3) == 0){ # no positive jumps
      big_part_poc_flux <- 0
    }else{
      delta_y <- sum(tmp3$diff_jump) *0.25 # to get ATN (= cp*0.25)
      max_time <- max(tmp$dates)
      min_time <- min(tmp$dates)
      delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
      slope_big_part <- delta_y/delta_x
      big_part_poc_flux <- 633*(slope_big_part**0.77)
    }
  }

  # compute total drifting time
  max_time <- max(tmp$dates)
  min_time <- min(tmp$dates)
  drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))

  # to plot subgroups
  part_slope_tmp <- part_slope |> mutate(dates = time, colour = 'slope')

  jump_plot <- plot_ly(tmp, x = ~dates, y = ~cp, type = 'scatter', mode = 'markers', color = ~colour, colors = c('#003366','#E31B23', '#FFC325')) |>
    add_lines(data= part_slope_tmp, x = ~dates, y = ~cp, split = ~group, color = I('#DCEEF3'), showlegend = F) |>
    layout(title= paste0('Drifting time: ', round(drifting_time,3), ' days\n',
                         'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n',
                         'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
                         'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1'), yaxis = list(title = 'Cp (1/m)'), xaxis = list(title = 'Time'))

  #return(jump_plot)
  #return(list('jump_plot' = jump_plot, 'jump_table' = tmp3))

    # adapt script to return big and small flux
  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = clean_data$park_depth[1], wmo = clean_data$wmo[1],
               cycle = clean_data$cycle[1])

  #return(jump_plot)
  #return(list('jump_plot' = jump_plot, 'jump_table' = tmp3))
  return(df)

}
```

```{r}
#| label: fig-OST-flux
#| fig-cap: OST-derived small (top) and big (bottom) particles flux for all floats at each drifting depth. Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.

# start loop for all data
wmo <- c(4903634, 6904240, 6904241, 1902578)
park_depth <- c('200 m', '500 m', '1000 m')

res4 <- data.frame()
for (i in wmo){
  #print(i)
  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
  for (j in park_depth){
    for (k in seq(1:max_cycle)){
      #print(k)
      tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
      if(nrow(tmp) == 0){
        next
      }else if(nrow(tmp) < 3){ # case where there is not enough data
        next
      }else{
        cleaned_signal <- clean_cp_data(tmp, 3)
        if(nrow(cleaned_signal) == 0){
          next
        }else{
          #for (l in window_size){
            #print(l)
            #for (m in quartile){
              #print(m)
          if(j == '200 m'){
            slope_threshold <- 0.23
          }else if(j == '500 m'){
            slope_threshold <- 0.12
          }else{
            slope_threshold <- 0.1
          }
          #selected_threshold <- threshold_table |> filter(park_depth == j, window_size == l, quartile == m)
          #slope_threshold <- selected_threshold$slope_threshold
          output <- plot_jumps(cleaned_signal, slope_threshold)
          #print(output)
          # append result
          res4 <- rbind(res4, output)
            #}
          #} 
        }
      }
    }
  }
}

# keep the good fluxes
cflux <- res4 |> mutate(drifting_time = difftime(max_time, min_time, units = 'days'), WMO = factor(wmo))

info_table <- cflux |> dplyr::group_by(park_depth) |> summarize(min_smallf = min(small_flux, na.rm=T), max_smallf = max(small_flux, na.rm=T),
                                                                mean_smallf = mean(small_flux, na.rm=T), median_smallf = median(small_flux, na.rm=T),
                                                                std_smallf = sd(small_flux, na.rm=T),
                                                                min_bigf = min(big_flux, na.rm=T), max_bigf = max(big_flux, na.rm=T),
                                                                mean_bigf = mean(big_flux, na.rm=T), median_bigf = median(big_flux, na.rm=T),
                                                                std_bigf = sd(big_flux, na.rm=T))

cflux$date <- as_date(cflux$min_time)

# add levels
cflux$park_depth <- factor(cflux$park_depth, levels = c('200 m', '500 m', '1000 m'))

###################################
# MOST RECENT VERSION OF THE PLOT #
###################################

up <- cflux |> ggplot(aes(x = date, y = small_flux)) +
  geom_smooth(data = cflux, aes(x = date, y = small_flux), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + #facet_grid(~park_depth)
  #scale_x_date(labels = date_format("%b"), date_breaks = '1 month') +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  #theme(axis.text.x = element_text(angle = 45, vjust = 0, hjust=0)) +
  guides(fill=guide_legend(title="WMO")) +
  #geom_pointrange(data = info_table2, aes(x = month, y = mean_small_f, ymin = mean_small_f - sd_small_f, ymax = mean_small_f + sd_small_f))
  #theme_bw() +
  geom_point(data = cflux, aes(x = date, y = small_flux, colour = WMO, shape = WMO)) +
  facet_wrap(~park_depth, scales = 'free') +
  labs(x = '', y = TeX('$F_{small}$ (mg C m$^{-2}$ day$^{-1}$)')) +
  theme_bw() #+
  # theme(axis.ticks.x = element_blank(),
  # axis.text.x = element_blank(), legend.position = 'top')
  #theme(legend.position = '')

down <- cflux |> ggplot(aes(x = date, y = big_flux)) +
  geom_smooth(data = cflux, aes(x = date, y = big_flux), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + #facet_grid(~park_depth)
  #scale_x_date(labels = date_format("%b"), date_breaks = '1 month') +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  #theme(axis.text.x = element_text(angle = 45, vjust = 0, hjust=0)) +
  guides(fill=guide_legend(title="WMO")) +
  #geom_pointrange(data = info_table2, aes(x = month, y = mean_small_f, ymin = mean_small_f - sd_small_f, ymax = mean_small_f + sd_small_f))
  #theme_bw() +
  geom_point(data = cflux, aes(x = date, y = big_flux, colour = WMO, shape = WMO)) +
  facet_wrap(~park_depth, scales = 'free') +
  labs(x = 'Month', y = TeX('$F_{big}$ (mg C m$^{-2}$ day$^{-1}$)')) +
  theme_bw() +
  theme(legend.position = 'none')

#plot_grid(up, down, labels = c('A', 'B'), label_size = 12, nrow = 2)
up / down

```

At 200 m, we estimated $F_{big}$ between 0 (no deposition of big particles on the transmissometer window) and 285.0 mg C m$^{-2}$ day$^{-1}$ with an average of 30.7 $\pm$ 54.4 mg C m$^{-2}$ day$^{-1}$ and a median of 11.8 mg C m$^{-2}$ day$^{-1}$. High fluxes (\> 100 mg C m$^{-2}$ day$^{-1}$) were measured by 3 out of 4 floats at the end of May before a sharp decrease towards July. At 500 m, $F_{big}$ ranges between 1.5 mg C m$^{-2}$ day$^{-1}$ and 150 mg C m$^{-2}$ day$^{-1}$ with an average of 19.5 $\pm$ 23.7 mg C m$^{-2}$ day$^{-1}$ and a median of 11.4 mg C m$^{-2}$ day$^{-1}$, slightly decreasing from June to December. At 1000 m, $F_{big}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to August, followed by a decrease towards December, with values ranging from 0 to 47.0 mg C m$^{-2}$ day$^{-1}$ with an average of 12.0 $\pm$ 11.1 mg C m$^{-2}$ day$^{-1}$ and a median of 8.2 mg C m$^{-2}$ day$^{-1}$.

--> ici mettre les données en variables car ça risque de changer souvent... les valeurs de flux etc and test.

```{r, smoothed plots}

#cflux |> ggplot(aes(x = date, y = small_flux, colour = park_depth, group = park_depth)) + geom_smooth() + theme_bw()

```

### UVP6 particle abundances

In this study, we focus on 12 size classes, from 102 $\mu$m to 1.63 mm. In @fig-uvp-1000m, we show the daily averaged concentrations of particles in each class size for each float at 1000 m. For the first three classes (i.e. 102-203 $\mu$m), we observe a similar pattern with concentrations starting at 4 particles/L at the end of May that reached 8 particles/L in mid-July and then decreased back to 4 particles/L towards December. For particles in the 203-406 $\mu$m range, the pattern can still be observed with a delay of 15 to 30 days between floats 6904240/4903634 and floats 6904241/4902578 which were deployed conjointly. Observed concentrations were very low at the end of May (\< 1 particle/L) then they increased by a factor of \~3-4 between mid-June and mid-July to decrease back to their initial concentrations towards December. For the remaining classes, concentrations are always below 1 particle/L with increasing uncertainty (rarer particles) in the signal. At 200 m (@fig-uvp-200), there is no clear peak following the bloom event in the smaller part of the size spectrum (\< 406 $\mu$m) whereas both floats 6904240 and 4903634 (i.e. conjointly deployed) observed the same high concentrations of particles above 406 $\mu$m at the end of May. At 500 m (@fig-uvp-500), the signal shows a slight increase in the abundance of particles in the 102-256 $\mu$m range from the end of May to the beginning of July or mid-July, depending of the float.

```{r, uvp data at 1000 m}
#| label: fig-uvp-1000m
#| fig-cap: Daily averaged concentration of particles abundance in 12 size classes measured by the UVP6 for each float at 1000 m. Shaded areas depict a 95% confidence interval around the mean.
# lpm_class <- c("NP_Size_102", "NP_Size_128", "NP_Size_161",
#                          "NP_Size_203", "NP_Size_256", "NP_Size_323",
#                          "NP_Size_406", "NP_Size_512", "NP_Size_645",
#                          "NP_Size_813", "NP_Size_1020", "NP_Size_1290")

# lpm_class <- c("NP_Size_102", "NP_Size_128", "NP_Size_161",
#                          "NP_Size_203", "NP_Size_256", "NP_Size_323",
#                          "NP_Size_406", "NP_Size_512", "NP_Size_645",
#                          "NP_Size_813", "NP_Size_1020", "NP_Size_1290",
#                         "NP_Size_1630", "NP_Size_2050")

lpm_class <- c("NP_Size_64","NP_Size_80.6", "NP_Size_102", "NP_Size_128", "NP_Size_161",
                         "NP_Size_203", "NP_Size_256", "NP_Size_323",
                         "NP_Size_406", "NP_Size_512", "NP_Size_645",
                         "NP_Size_813", "NP_Size_1020", "NP_Size_1290",
                        "NP_Size_1630", "NP_Size_2050")

uvp_data <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date, cycle, wmo, park_depth, all_of(lpm_class)) |> 
  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))

uvp_data <- uvp_data |> pivot_longer(cols = lpm_class, names_to = 'size_class', values_to = 'conc')

info2 <- uvp_data |> dplyr::group_by(week, park_depth, wmo, size_class) |> summarize(mean_conc = mean(conc))

# add level for order
uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))

p1000m <- uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = conc, colour = WMO)) + stat_smooth() + 
  scale_color_brewer(palette = 'Dark2') + 
  geom_vline(xintercept = as_date(2022-10-01), linetype = 'dashed') +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  facet_wrap(~size_class, scales = 'free_y')

p1000m

tmp1000m <- ggplot_build(p1000m)
tmp1000m <- tmp1000m$data[[1]]

### maybe we can just show the real values here, let's see with a mean first then add some uncertainties
# info3 <- uvp_data |> dplyr::group_by(juld, park_depth, wmo, size_class) |> summarize(mean_conc = mean(conc))
# 
# info3$size_class <- factor(info3$size_class, levels = lpm_class)
# info3$park_depth <- factor(info3$park_depth, levels = c('200 m', '500 m', '1000 m'))
# 
# info3 |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = mean_conc, colour = as.factor(wmo))) + geom_smooth() +
#   scale_color_brewer(palette = 'Dark2') + 
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
#   facet_wrap(~size_class, scales = 'free_y') 
# 
# info3 |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = mean_conc, colour = as.factor(wmo))) + geom_path() +
#   scale_color_brewer(palette = 'Dark2') + 
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
#   facet_wrap(~size_class, scales = 'free_y')

```

```{r, uvp data at 200 m}
#| label: fig-uvp-200
#| fig-cap: Daily averaged concentration of particles abundance in 12 size classes measured by the UVP6 for each float at 200 m. Shaded areas depict a 95% confidence interval around the mean.

# ggplot(uvp_data, aes(x = juld, y = conc, colour = park_depth)) + geom_smooth() + 
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
#   scale_color_brewer(palette = 'Dark2') + theme_bw() + 
#   facet_wrap(~size_class, scales = 'free_y')  #+ geom_point() + ylim(c(0,50))

p200m <- uvp_data |> filter(park_depth == '200 m') |> ggplot(aes(x = juld, y = conc, colour = WMO)) + stat_smooth() + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
  scale_color_brewer(palette = 'Dark2') + theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  facet_wrap(~size_class, scales = 'free_y')

p200m

tmp200m <- ggplot_build(p200m)
tmp200m <- tmp200m$data[[1]]

# uvp_data |> filter(park_depth == '500 m') |> ggplot(aes(x = week, y = conc, colour = factor(wmo))) + geom_smooth() + 
#   #scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
#   scale_color_brewer(palette = 'Dark2') + theme_bw() + 
#   facet_wrap(~size_class, scales = 'free_y')

# uvp_data |> filter(park_depth == '500 m') |> ggplot(aes(x = juld, y = conc, colour = WMO)) + geom_smooth() + 
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
#   scale_color_brewer(palette = 'Dark2') + theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
#   facet_wrap(~size_class, scales = 'free_y')

# uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = week, y = conc, colour = factor(wmo))) + geom_smooth() + 
#   #scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
#   scale_color_brewer(palette = 'Dark2') + theme_bw() + 
#   facet_wrap(~size_class, scales = 'free_y')

# uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = DOY, y = conc, colour = factor(wmo))) + geom_smooth() + 
#   scale_color_brewer(palette = 'Dark2') + 
#   theme_bw() + 
#   facet_wrap(~size_class, scales = 'free_y')

# uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = conc, colour = factor(wmo))) + geom_smooth() + 
#   scale_color_brewer(palette = 'Dark2') + 
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   theme_bw() + 
#   facet_wrap(~size_class, scales = 'free_y')

# uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = conc, colour = factor(wmo))) + geom_smooth(size = 1.5) + 
#   #scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
#   scale_color_brewer(palette = 'Dark2') + theme_bw() + 
#   facet_wrap(~size_class, scales = 'free_y')

# ggplot(info2, aes(x = juld, y = mean_conc, colour = park_depth, shape = factor(wmo))) + geom_point() + 
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
#   facet_wrap(~size_class, scales = 'free_y')

```

```{r, uvp data at 500 m}
#| label: fig-uvp-500
#| fig-cap: Daily averaged concentration of particles abundance in 12 size classes measured by the UVP6 for each float at 500 m. Shaded areas depict a 95% confidence interval around the mean.
p500m <- uvp_data |> filter(park_depth == '500 m') |> ggplot(aes(x = juld, y = conc, colour = WMO)) + stat_smooth() + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
  scale_color_brewer(palette = 'Dark2') + theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  facet_wrap(~size_class, scales = 'free_y')

p500m

tmp500m <- ggplot_build(p500m)
tmp500m <- tmp500m$data[[1]]

```

### Where is the bloom?

```{r, sat data}

#chla_sat <- vroom::vroom('DATA/SATELLITE/chla_sat.csv')
chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_sat.feather') |> mutate(DOY = yday(time))
sat_data_bbox <- filter(chla_sat, lat >= 55.5, lat <= 60, lon >= -55, lon <= -45) |> group_by(DOY) |> drop_na(CHL)

# sat data
test <- sat_data_bbox |> group_by(DOY) |> summarize(npixels = n(), nabove5 = length(which(CHL > 5)), nabove2 = length(which(CHL > 2)), r5 = nabove5/npixels)
ggplot(test, aes(DOY, r5)) + geom_point() + geom_path()

# argo data
npq_data <- vroom('DATA/ARGO/npq_corrected_data.csv')  
npq_filtered <- npq_data |> filter(pres > 0, pres <= 20, juld < '2022-12-31') |> mutate(DOY = yday(juld)) |> drop_na(chla)

npq_argo <- npq_filtered |> group_by(wmo, cycle) |> summarize(DOY = unique(DOY), mean_chla = mean(chla_npq, na.rm=T), median_chla = median(chla_npq, na.rm=T))

ggplot(npq_argo, aes(x = DOY, y = mean_chla, colour = factor(wmo))) + geom_point() + geom_path() + facet_wrap(~wmo)

# # ggplot(sat_data_bbox, aes(DOY, mean_chla)) + geom_point() + geom_path()
# ggplot(sat_data_bbox, aes(DOY, q50)) + geom_point() + geom_path()

# # check float position before july
# float_position <- filter(all_floats, juld < '2022-06-15') |> drop_na(lat)
# 
# # get lat/lon range for each float for that period
# float_position |> dplyr::group_by(wmo) |> dplyr::summarise(min_lat = min(lat), max_lat = max(lat),
#                                                            min_lon = min(lon), max_lon = max(lon))
# 
# chla_sat_couple1 <- filter(chla_sat, lat > 56.5, lat < 56.85, lon > -53, lon < -52.4) |> drop_na(CHL)
# 
# tmp <- chla_sat_couple1 |> dplyr::group_by(time) |> summarise(mean_chla = mean(CHL))
# 
# ggplot(tmp, aes(x = time, y = mean_chla)) + geom_point() + geom_path()
# 
# # at deployment
# chla_sat_couple1 <- filter(chla_sat, lat > 56.7, lat < 56.9, lon > -52.5, lon < -52.3) |> drop_na(CHL)
# tmp <- chla_sat_couple1 |> dplyr::group_by(time) |> summarise(mean_chla = mean(CHL))
# ggplot(tmp, aes(x = time, y = mean_chla)) + geom_point() + geom_path()
# 
# # couple 2
# chla_sat_couple2 <- filter(chla_sat, lat > 58.9, lat < 59.1, lon > -49.4, lon < -49.2) |> drop_na(CHL)
# tmp <- chla_sat_couple2 |> dplyr::group_by(time) |> summarise(mean_chla = mean(CHL))
# ggplot(tmp, aes(x = time, y = mean_chla)) + geom_point() + geom_path()
# 
# ggplot(argo_map) + w + geom_path(aes(lon, lat, colour = WMO), size = 1) +
#   geom_point(aes(lon, lat, colour = WMO), size = 1.5) +
#   scale_color_brewer(palette = 'Dark2') +
#   theme_bw() +
#   #scale_colour_manual(values = c('#a6cee3', '#33a02c', '#b2df8a', '#1f78b4')) +
#   coord_cartesian(xlim = c(-55, -40), ylim = c(55,65)) + geom_point(data = argo_map_first, aes(lon, lat), shape = 17, size = 3) +
#   geom_point(data = argo_map_last, aes(lon, lat, colour = WMO), size = 3) +
#   geom_point(data = argo_map_last, aes(lon, lat, fill = WMO), size = 3, shape = 1) +
#   labs(x = 'longitude', y = 'latitude') +
#   #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
#   guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
#                                                   shape = c(NA, NA, NA, NA)))) +
#   geom_point(data = chla_sat_couple1, aes(x = lon, y = lat), colour = "red") +
#   geom_point(data = chla_sat_couple2, aes(x = lon, y = lat), colour = "red")
#

```

```{r}

# all_floats_chla <- filter(all_floats, PhaseName == 'NPAR', pres <= 50, pres > 2) |> drop_na(chla) |> select(juld, pres, chla, wmo) |> mutate(datetime = as_date(juld)) #|> filter(datetime < '2022-08-01')
# 
# chla_summary <- all_floats_chla |> dplyr::group_by(wmo, datetime) |> summarize(chla_mean = mean(chla, na.rm=T))
# 
# ggplot(chla_summary, aes(x = datetime, y = chla_mean)) + geom_point() + geom_path() + facet_wrap(~wmo)

```

### Combining OST and UVP data

For the first time, BGC-Argo floats equipped with both a transmissometer and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach \[@Claustre2021-tx\]. The quantification of the POC flux using the OST method \[@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr\] is based on an empirical POC:ATN ratio \[REF estapa 2022\] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:

$$
F = \sum_{i}^{N}C_{i}Ad_{i}^{b}
$$ {#eq-guidi-article3}

where $C_{i}$, A and b represent, respectively, the concentration of particles (number/L) for the $i^{th}$ class of mean diameter $d_{i}$ (in mm) and 2 constants with $b$ being related to the fractal dimension ($D$) of an aggregate by $D = (b +1)/2$ [@Guidi2008-vp].

The use of @eq-guidi-article3 to derive carbon fluxes from PSD was firstly implemented by @Guidi2008-vp who used a minimization procedure between sediment traps data and estimated fluxes using PSD from several previous versions of the UVP to find the optimal value of $A$ and $b$ (i.e. 12.5 and 3.81). @Iversen2010-an and @Fender2019-ou followed the same procedure and concluded that the couple of parameters found by [@Guidi2008-vp] could not be used globally at the expense of greatly over- or underestimating POC fluxes, depending on the sampling region. In addition, the formulation of such a model implies that $A$ and $b$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt]. @eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2022-qj]. Due to those reasons, we have chosen to firstly compare $F_{small}$ and $F_{big}$ with UVP6 particles concentrations. For this, we computed the particles density distribution maximum using a smoothing gaussian kernel (see R package ggvis) for each class size at each parking depth and at each cycle. Then, we computed the distance correlation between OST-derived carbon fluxes and UVP size classes maxima. The distance correlation \[REF\] was preferred to the well known Pearson's correlation due to uncertainties in the linear relationship of, respectively, small classes and $F_{big}$, and big classes and $F_{small}$.

```{r, essai sur ridgeplot}
library(ggridges)
tmp <- uvp_data |> filter(park_depth == '1000 m', wmo == 6904240, conc < 20)
#tmp <- uvp_data |> filter(park_depth == '1000 m', wmo == 6904240, size_class %in% c('NP_Size_102', 'NP_Size_128', 'NP_Size_161', 'NP_Size_203', 'NP_Size_256', 'NP_Size_323'))

# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_density_ridges(scale = 5) + stat_density_ridges(quantile_lines=T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
#   theme_bw() + scale_color_brewer(palette = 'Dark2')
# 
# # ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_density_ridges() + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) 
# #   
# # ggplot(tmp, aes(y = factor(cycle), x = conc)) + geom_density_ridges_gradient() + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
# #    scale_fill_viridis_c(name = "Temp. [F]", option = "C")
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + stat_density_ridges(quantile_lines=T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
#   theme_bw() 
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc)) + stat_density_ridges(quantile_lines=T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
#   theme_bw() + facet_wrap(~size_class) + xlim(c(0, 10)) 
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc)) + stat_density_ridges(quantile_lines=T, quantiles = 2, stat="identity") + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
#   theme_bw() + facet_wrap(~size_class, scales = "free")
# 
ggplot(tmp, aes(y = factor(cycle), x = conc, height = ..density..)) + geom_density_ridges(scale = 5, stat = "density", rel_min_height = 0.01) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
  theme_bw() + facet_wrap(~size_class, scales = "free")
# 
# # test 
# tmp2 <- uvp_data |> dplyr::group_by(cycle, size_class, park_depth, wmo) |> 
#   group_modify(~ ggplot2:::compute_density(.x$conc, NULL)) |>
#   rename(conc = x)
# 
# ggplot(tmp2, aes(x = conc, y = cycle, height = density)) + 
#   geom_density_ridges(stat = "identity") + facet_wrap(~size_class, scales = "free")

# 
# ggplot(tmp, aes(y = factor(cycle), x = conc, height = ..density..)) + stat_density_ridges(scale = 5) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
#   theme_bw() + facet_wrap(~size_class, scales = "free")
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc, height = ..density..)) + stat_density_ridges(scale = 5, quantile_lines = T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
#   theme_bw() + facet_wrap(~size_class, scales = "free")
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_ridgeline() + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
#   theme_bw() + scale_color_brewer(palette = 'Dark2')

# ggplot(tmp, aes(x = factor(cycle), y = conc, fill = size_class)) + geom_boxplot() + labs(y = 'abundance distribution (#/L)', x = 'cycle')  +
#   theme_bw() + scale_color_brewer(palette = 'Dark2') + ylim(c(0,15)) + geom_jitter()
# 
# ggplot(tmp, aes(x = factor(cycle), y = conc, fill = size_class)) + geom_violin() + labs(y = 'abundance distribution (#/L)', x = 'cycle')  +
#   theme_bw() + scale_color_brewer(palette = 'Dark2') + ylim(c(0,15))

# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_density_ridges(alpha=1, stat="binline", bins=50)  + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
#   theme_bw() + scale_color_brewer(palette = 'Dark2')

```

```{r, check with particles abundance}

particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))

particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size_class', values_to = 'concentration')
particles_bis$size_class <- factor(particles_bis$size_class, levels = lpm_class)

ggplot(particles_bis, aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance') +
  facet_wrap(~size_class) + theme_bw()

ggplot(particles_bis |> filter(wmo == '6904240'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 6904240') +
  facet_wrap(~size_class) + theme_bw()

# https://stackoverflow.com/questions/48424682/how-do-i-limit-the-range-of-the-viridis-colour-scale
ggplot(particles_bis |> filter(wmo == '6904240'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_colour_viridis(option = 'H', limits = c(-1,1), oob = scales::squish, na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 6904240') +
  facet_wrap(~size_class) + theme_bw()

ggplot(particles_bis |> filter(wmo == '6904241'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 6904241') +
  facet_wrap(~size_class) + theme_bw()

ggplot(particles_bis |> filter(wmo == '4903634'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 4903634') +
  facet_wrap(~size_class) + theme_bw()

ggplot(particles_bis |> filter(wmo == '1902578'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 1902578') +
  facet_wrap(~size_class) + theme_bw()

# # sum 3 first class
# particles_tmp <- particles |> dplyr::group_by(wmo, cycle, depth) |> summarize(juld = unique(juld), sum_NP = sum(NP_Size_102, NP_Size_128, NP_Size_161)) |> filter(depth > 100)
# 
# ggplot(particles_tmp, aes(x = juld, y = depth, colour = log(sum_NP))) + geom_point() + scale_color_viridis(option = 'H') +
#   scale_y_reverse(limits = c(2000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

# ggplot(particles, aes(x = juld, y = depth, colour = log(NP_Size_102))) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(2000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux_without_depth)) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(2000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

```

### SPECTRAL SLOPES

```{r, function compute slope}

# mid_DSE <- c(0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
# sizebin <- c(0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
#              0.532318310)

mid_DSE <- c(0.0911143073, 0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
sizebin <- c(0.0209587201, 0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
             0.532318310)

vertical_profiles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)
drifting_profiles <- all_floats |> filter(PhaseName == 'PAR') |> select(juld, cycle, depth = pres, park_depth, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)

compute_slope <- function(i, data_spectra){

  spectrum <- data_spectra[i,]

  spectrum_norm <- spectrum/sizebin

  # prepare data for linear regression
  Y <- log(spectrum_norm)
  X <- log(mid_DSE)

  # check for finite value
  h <- is.finite(Y)
  Y <- Y[h]
  X <- X[h]

  data_slope <- tibble(X=X, Y=Y)

  model <- lm(formula = Y ~ X, data = data_slope)
  slope <- model$coefficients[2]

  return(slope)
}

# compute slope on drifting profiles
drifting_spectra <- as.matrix(data.frame(drifting_profiles[,6:20]))
index <- seq(from = 1, to = nrow(drifting_spectra), by = 1)
slopes <- map_dbl(index, compute_slope, data_spectra = drifting_spectra)
drifting_profiles$spectral_slope <- slopes
ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
  scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')

# compute slope on vertical profiles
vertical_spectra <- as.matrix(data.frame(vertical_profiles[,5:19]))
index <- seq(from = 1, to = nrow(vertical_spectra), by = 1)
slopes <- map_dbl(index, compute_slope, data_spectra = vertical_spectra)
vertical_profiles$spectral_slope <- slopes

```

```{r, add slopes}

# vertical_profiles$slope <- slopes

ggplot(vertical_profiles, aes(x = juld, y = depth, colour = spectral_slope)) + geom_point() + scale_colour_viridis(option = 'H') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = TeX('slope')) + facet_wrap(~wmo)

```

### DRIFT DATA

```{r, check drift per class}

tmp <- drifting_profiles |> pivot_longer(cols = all_of(lpm_class[2:16]), names_to = "size_class")
tmp$park_depth <- factor(tmp$park_depth, levels = c('200 m', '500 m', '1000 m'))
tmp$size_class <- factor(tmp$size_class, levels = lpm_class)

tmp2 <- tmp |> group_by(park_depth, wmo, cycle, size_class) |> summarize(median_conc = median(value), min_juld = min(juld))

ggplot(tmp2 |> filter(wmo == 6904241, park_depth != '200 m'), aes(x = min_juld, y = median_conc, colour = factor(park_depth))) + geom_point() + geom_path() + facet_wrap(~size_class, scales = 'free') + theme_bw() + labs(title = 'WMO 6904241')

```
```{r spectral slope + rati min/max flux}

tmp <- cflux |> mutate(ratioF = small_flux/(small_flux+big_flux))

# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF, colour = park_depth)) + geom_point() + facet_wrap(~wmo)
# 
# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_point() + facet_wrap(~park_depth)
# 
# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)

ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_point() + geom_smooth() + facet_wrap(~park_depth)

ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
  scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')

tmp2 <- drifting_profiles |> group_by(wmo, cycle, park_depth) |> summarize(median_spectral_slope = median(spectral_slope))

# merge data
check <- merge(tmp, tmp2) 
check2 <- merge(tmp, tmp2) |> filter(ratioF < 1)

ggplot(check, aes(ratioF, median_spectral_slope)) + geom_point()

ggplot(check2, aes(ratioF, median_spectral_slope)) + geom_point()

# ggplot(tmp, aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)
```

```{r, check pulse export}

# bbp_chla <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, chla, bbp) |> drop_na(chla) |> mutate(juld = as_date(juld))
# cp_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, cp) |> drop_na(cp) |> mutate(juld = as_date(juld))
# 
# ggplot(bbp_chla, aes(x = juld, y = depth, colour = log10(chla))) + geom_point() + scale_color_viridis(option = 'H') +
#   scale_y_reverse(limits = c(300,0)) + 
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
# 
# # ggplot(bbp_chla, aes(x = juld, y = depth, colour = log(bbp))) + geom_point() + scale_color_viridis(option = 'H') +
# #   scale_y_reverse(limits = c(2000,0)) +
# #   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
# #   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

```

### TRY: building a model of flux based on spectra (w/wo depth)

```{r, same as above but without depth}

  # get spectra
  uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |> 
  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))

  # get OST fluxes
  cflux2 <- cflux |> drop_na(small_flux)

  # # check 'outliers' -> justification: bad fluxes, see shiny (on fit des pentes sur 4 points, not enough, tout le reste ce sont des jumps)
  # if(parking_depth == '500 m'){
  #   cflux2 <- cflux2 |> filter(small_flux < 15)
  # }

  # merge data for training
  data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)

  # training data
  tmp <- data_model |> select(small_flux, all_of(lpm_class))

  set.seed(101)
  # model
  mg2 <- gbm(small_flux ~ ., data=tmp, distribution="gaussian",
  cv.folds=5,
  n.trees=3000, shrinkage=0.005, interaction.depth=6,
  bag.fraction=0.5,
  n.cores=1
  )

  # get best model and thus best round number
  best <- gbm.perf(mg2)
  print(best)

  #summary(mg2, n.trees=best)
  tt <- summary(mg2, n.trees=best)
  print(tibble(var = tt$var, importance = tt$rel.inf))

    # predictions on mean values
  uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = unique(small_flux), 
                                                                                   NP_Size_64 = mean(NP_Size_64),
                                                                                   NP_Size_80.6 = mean(NP_Size_80.6),
                                                                                   NP_Size_102 = mean(NP_Size_102),
                                                                                   NP_Size_128 = mean(NP_Size_128),
                                                                                   NP_Size_161 = mean(NP_Size_161),
                                                                                   NP_Size_203 = mean(NP_Size_203),
                                                                                   NP_Size_256 = mean(NP_Size_256),
                                                                                   NP_Size_323 = mean(NP_Size_323),
                                                                                   NP_Size_406 = mean(NP_Size_406),
                                                                                   NP_Size_512 = mean(NP_Size_512),
                                                                                   NP_Size_645 = mean(NP_Size_645),
                                                                                   NP_Size_813 = mean(NP_Size_813),
                                                                                   NP_Size_1020 = mean(NP_Size_1020),
                                                                                   NP_Size_1290 = mean(NP_Size_1290),
                                                                                   NP_Size_1630 = mean(NP_Size_1630),
                                                                                   NP_Size_2050 = mean(NP_Size_2050)) |> ungroup()

  # prediction at 1000 m
  spectra <- uvp_spectra_average |> select(true_flux, all_of(lpm_class[]))

  # prediction 
  pred_flux <- predict(mg2, newdata = spectra, n.trees = best)

  plot(x=spectra$true_flux, y=pred_flux, asp=1)
  print(R2_Score(pred_flux, spectra$true_flux))

```

### Plot results

Let's apply both models to all data at non parking depth (ie. vertical profiles) and let's plot a time serie to compare. Starting at 200 m car le modele devient n'a pas appris entre 0-200 et dans cette tranche, les concentrations de particules peuvent être huge huge -> note que cela pourrait dire qu'il faudrait faire des mesures (autre instrument? pour estimer un flux..)

```{r, apply model to vertical profile}

vertical_profiles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))

predicted_flux_without_depth <- predict(mg2, newdata = vertical_profiles, n.trees = best)

vertical_profiles$pred_flux_without_depth <- predicted_flux_without_depth

ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux_without_depth)) + geom_point() + scale_color_viridis(option = 'H') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log(pred_flux))) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(1000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(pred_flux_without_depth))) + geom_point() + scale_color_viridis(option = 'H') +
#   scale_y_reverse(limits = c(2000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(pred_flux_without_depth))) + geom_point() + scale_color_viridis(option = 'H') +
  scale_y_reverse(limits = c(500,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) + facet_wrap(~wmo)

ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(pred_flux_without_depth))) + geom_point() + scale_color_viridis(option = 'H') +
  scale_y_reverse(limits = c(2000,200)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) + facet_wrap(~wmo)

# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(pred_flux_without_depth))) + geom_point() + scale_color_viridis(option = 'H') +
#   scale_y_reverse(limits = c(500,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) + facet_wrap(~wmo)

# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log(pred_flux_without_depth))) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(2000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) + facet_wrap(~wmo)

# cflux_200m$depth <- 200
# cflux_200m$juld <- as_date(cflux_200m$min_time)
# 
# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux)) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(1000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) +
#   geom_point(data = cflux_200m, aes(x = juld, y = depth, colour = 'black', size = small_flux))

# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log(pred_flux))) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(1000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux)) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(1000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) +
#   facet_wrap(~wmo)
# 
# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log(pred_flux))) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(1000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) +
#   facet_wrap(~wmo)

# cflux_200m$depth <- 200
# cflux_200m$juld <- as_date(cflux_200m$min_time)
# 
# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux)) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(1000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) +
#   geom_point(data = cflux_200m, aes(x = juld, y = depth, colour = small_flux), shape = 17, size = 2)

# ggplot(cflux_200m, aes(x = juld, y = depth, colour = small_flux)) + geom_point() + scale_color_viridis()

```

Modèle appris en discontinu et prédictions en continu --> BOF BOF et ça se voit. L'autre modèle a une bien plus belle gueule et ne prend pas la profondeur en compte ce qui est encore mieux selon moi, le small flux est expliqué principalement par 102, 128 et 161, ce qui est cohérent. On voit potentiellement un pulse d'export --> faudrait le vérifier avec le bbp/cp? On voit que vers la fin de l'année, le flux en profondeur est plus faible -> cohérent vu que le bloom date de bien avant. On aurait besoin d'une time series plus longue, ici j'ai bloqué à 6 mois (car le drift à 200 m va etre mis en pause avec l'hiver et la MLD + le fait que l'on perd un flotteur avec un transmissio qui merde). Ajouter MLD, iso-sigma?

```{r, check sthg}

# particles_grouped <- particles |> mutate(NP_Size_102_256 = rowSums(across(c(NP_Size_102,NP_Size_128,NP_Size_161,NP_Size_203))),
#                                          NP_Size_256_512 = rowSums(across(c(NP_Size_256,NP_Size_323,NP_Size_406))),
#                                          NP_Size_512_1020 = rowSums(across(c(NP_Size_512,NP_Size_645,NP_Size_813))),
#                                          NP_Size_1020_2580 = rowSums(across(c(NP_Size_1020,NP_Size_1290, NP_Size_1630,NP_Size_2050))))
# 
# particles_ter <- particles_grouped |> pivot_longer(cols = c(NP_Size_102_256, NP_Size_256_512, NP_Size_512_1020, NP_Size_1020_2580), names_to = 'size_group', values_to = 'concentration')
# #particles_ter$size_group <- factor(particles_ter$size_class, levels = lpm_class)
# 
# ggplot(particles_ter |> filter(wmo == '6904240'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
#   scale_y_reverse(limits = c(2000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'), title = 'wmo 6904240') +
#   facet_wrap(~size_group) + theme_bw()

```

## Discussion

## Conclusion

### 2023-01-22 definitions.qmd
start project

# Definitions {.unnumbered}

Sequestration: Process of isolating and storing respired CO<sub>2</sub> within the ocean interior and away from atmospheric exchanges

Export flux:

respiration

photosynthesis

BCP strength

BCP efficiency

Net Primary Production: Net conversion rate of inorganic carbon to organic carbon by phytoplankton photosynthesis [@Siegel2022-wd]

Euphotic Zone: Depth range in the near-surface ocean with enough sunlight for phytoplankton net primary production to occur [@Siegel2022-wd]

Twiligth Zone = Mesopélagic zone

Mixed Layer

Remineralization: Conversion of organic matter back to inorganic forms by heterotrophic microbes and animals [@Siegel2022-wd]

Particle size distribution (or particle size spectrum)

Carbon dioxide removal

### 2023-01-22 index.qmd
start project

# Foreword {.unnumbered}

```{=html}
<!--  As of 2022, it has now become obvious to the general public that the Earth's climate is changing. Massive flood events, repeated heatwaves, never-ending droughts, disappearing glaciers, biodiversity collapse, … Reports on weather anomalies were already published in the 1960's, the Intergovernmental Panel on Climate Change (IPCC) was created in 1988, the Kyoto Protocol was adopted in 1997, An Inconvenient Truth was released in 2006, the Paris Agreement was signed in 2015 and still, what has really changed?

As I am writing this part, the COP27 (sponsored by Coca-Cola, no joke) is taking place with an acknowledged record number of fossil fuels lobbyists, bloody stadiums with outside air conditioning are about to be used in the desert for a month and it seems like 'leaders' of the Free World have been having hearing issues. See for yourself:

*"I am drowning"* said Pakistan

*"I'm sorry, what?"* said the Free World

*"I am starving"* said Ethiopia

*"Can't hear you, louder please!"* said the Free World

*"We are suffocating"* said India

*"Of course, I would love building new pipelines in your garden"* said the Free World

While I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured, there are no fake news in this work.

I hope you will enjoy it and no, I won't tweet about it. -->
```

### 2023-01-22 intro.qmd
start project

# Introduction

## The global carbon cycle {#sec-global_carbon_cycle}

The global carbon cycle (@fig-global_carbon_cycle) consists of a series of exchanges of carbon in various organic and inorganic forms among four major reservoirs: the lithosphere, the hydrosphere, the atmosphere and the biosphere. These exchanges occur at different timescales, such that the carbon cycle can be divided into a slow and a fast component. The former operates on geological timescales where carbon is hidden from the surface for millions of years (e.g. in hydrocarbon deposits such as petroleum or natural gas) before being released through volcanic activity while the latter operates on shorter timescales, from minutes to years (e.g. photosynthesis, air-sea fluxes) up to decades and millenia (e.g. ocean ventilation, changing ocean buffering capacity) [@Bertrand2021-nh; @Canadell2021-ds].

99% of the Earth's carbon is stored in rocks, holding about 100,000 times as much carbon as the atmosphere while the ocean contains about 60 times as much carbon as the atmosphere [@DeVries2022-of; @Bertrand2021-nh; @Sarmiento2006-ec] with roughly 37,000 Pg (i.e. 1 Pg = 10$^{15}$ g = 1 Gt) of dissolved inorganic carbon (DIC) and 700 Pg of dissolved organic carbon (DOC) [@Friedlingstein2021-kc].

![Illustration of the global carbon cycle (numbers were taken for the year 2020). Arrows and boxes depict, respectively, carbon fluxes in Pg C year$^{-1}$ and carbon content in Pg C. Figure taken from @DeVries2022-of.](image/intro/carbon_cycle_devries2022.jpeg){#fig-global_carbon_cycle}

The global carbon cycle had been at close equilibrium for millions of years before the industrial revolution started in the 18th century. Since then, human have been burning fossil fuels at an unprecedented rate, releasing huge amounts of carbon dioxide ($CO_{2}$) in the atmosphere that should have been discharged slowly through the slow carbon cycle [@Bertrand2021-nh]. These emissions, resulting from fossil emissions but also from land use (e.g. deforestation, land management) have been estimated around 700 Pg C since 1750 [@Canadell2021-ds].

Thankfully, more than half of this anthropogenic $CO_{2}$ has been absorbed by terrestrial ecosystems and by the ocean [@Friedlingstein2021-kc], forming major carbon sinks.

## How is the ocean a $CO_{2}$ sink? {#sec-ocean_co2_sink}

The ocean is known to play a key role in the fast carbon cycle due to its size and rapid exchange of $CO_{2}$ with the atmosphere, controlling atmospheric $CO_{2}$ levels on timescales of decades to millennia [@DeVries2022-of; @Sarmiento2006-ec]. Currently, the ocean absorbs \~2.5 Gt C year$^{-1}$, which represents about 25% of annual anthropogenic $CO_{2}$ emissions [@Canadell2021-ds; @Friedlingstein2021-kc]. In order to understand how the ocean absorbs the latter, one must take a look at seawater carbonate chemistry.

### Inorganic carbonate chemistry {#sec-inorganic_carbonate_chemistry}

The intensity of the exchange of $CO_{2}$ at the air-sea interface is driven by the difference in the partial pressure of $CO_{2}$ in the atmosphere ($pCO_{2}^{atm}$) and in the surface ocean ($pCO_{2}^{surface}$).

$$
\Delta pCO_{2} = pCO_{2}^{surface} - pCO_{2}^{atm}
$$ {#eq-deltaCO2}

When $\Delta pCO_{2} > 0$, surface waters are supersaturated in $CO_{2}$ resulting in a net flux of $CO_{2}$ from the ocean to the atmosphere. On the contrary, surface waters are undersaturated with a net $CO_{2}$ flux in the opposite direction [@Sarmiento2006-ec].

When a molecule of $CO_{2~(gas)}$ dissolves in seawater, it first gets hydrated to form $CO_{2~(aq)}$ then it reacts with water to form carbonic acid ($H_{2}CO_{3}$).

$$
CO_{2~(gas)} + H_2{O} \rightleftharpoons CO_{2~(aq)} + H_{2}O \rightleftharpoons H_{2}CO_{3}
$$ {#eq-carbonic_acid}

$H_{2}CO_{3}$ is a strong acid that quickly dissociates in two steps to form bicarbonate ($HCO_{3}^{-}$) and carbonate ($CO_{3}^{2-}$) where each reaction releases one free proton ($H^{+}$).

$$
H_{2}CO_{3} \rightleftarrows H^{+} + HCO_{3}^{-}
$$ {#eq-bicarbonate}

$$
HCO_{3}^{-} \rightleftarrows H^{+} + CO_{3}^{2-}
$$ {#eq-carbonate}

The abundance of each carbonate species is governed by the equilibrium between those three relationships [@Sarmiento2006-ec]. According to @Gattuso2011-di, at a typical surface pH of 8.2, the relative abundances of $CO_{2}$ (encompassing $CO_{2~(aq)}$ and $H_{2}CO_{3}$ because it is difficult to distinguish them analytically), $HCO_{3}^{-}$ and $CO_{3}^{2-}$ are, respectively, 0.5%, 89% and 10.5%. For convenience, the sum of all dissolved carbonate species will be hereafter referred to as DIC

$$
DIC = [CO_{2}] + [HCO_{3}^{-}] + [CO_{3}^{2-}]
$$ {#eq-DIC}

where brackets denote concentrations.

### Distribution of DIC in the water column {#sec-DIC_distribution}

It can be shown [e.g. @DeVries2022-of] that if we were to consider the ocean as a homogeneous reservoir of inorganic carbon, $pCO_{2}^{atm}$ would be much higher than it is now (roughly 416 parts per million (ppm) as of October 2022 according to the Mauna Loa Observatory). This observation is explained by the presence of a strong vertical gradient of DIC in the ocean, with surface DIC concentrations more than 200 $\mu$mol kg$^{-1}$ lower than at 1000 m as it can be seen in @fig-DIC_plot.

```{r, DIC plot}
#| label: fig-DIC_plot
#| fig-cap: Mean DIC distribution in the global ocean. Data taken from the Merged Master File from GLODAPv2, available at https://www.glodap.info/index.php/merged-and-adjusted-data-product-v2-2022/, last accessed on 26/10/2022.

library(ggplot2)
library(tidyverse)
library(ggtext)
library(latex2exp)

# CODE BELOW IS *OKAY* but takes time to compile so I am just loading the last working dataframe to plot the DIC distribution.

# DIC_data <- vroom::vroom('/home/flo/Downloads/GLODAPv2_2022_all/GLODAPv2.2022_Merged_Master_File.csv')
# DIC_data <- DIC_data |> select(depth= G2depth, DIC = G2tco2, G2tco2f, G2tco2qc)
# # clean a bit
# DIC_data <- DIC_data |> filter(G2tco2qc == 1, G2tco2f == 2)
# DIC_data <- DIC_data |> filter(depth < 4500)
# # bin it
# # brks <- c(0,10,20,30,40,50,60,70,80,90,100,150,200,250,300,350,400,450,500,600,700,800,900,1000,1500,2000,2500,3000,3500,4000,4500)
# # big bin
# brks <- c(0,20,40,60,80,100,150,200,250,300,350,400,450,500,600,700,800,900,1000,1500,2000,2500,3000,4500)
# # smaller bins
# brks <- c(0,20,40,60,80,100,150,200,250,300,350,400,450,500,600,700,800,900,1250,1500,1750, 2000,2250, 2500, 2750,3000,3250,3500, 3750, 4000, 4250, 4500)
# ints <- findInterval(DIC_data$depth, brks, all.inside = T)
# mids <- (brks[ints] + brks[ints + 1]) / 2
# DIC_data$mids <- mids
# DIC_data <- DIC_data |> dplyr::group_by(mids) |> dplyr::summarise(mean_DIC = mean(DIC, na.rm=T), nb_points = n())

# save data
#vroom::vroom_write(DIC_data, '/home/flo/dox/THESIS/DATA/GLODAP/DIC_distribution.csv')

# load data 
DIC_data <- vroom::vroom('./DATA/GLODAP/DIC_distribution.csv')

#plot it
# ggplot(DIC_data, aes(y = mids, x = mean_DIC)) + geom_path(size = 1) + scale_y_reverse() + labs(y = 'Depth [m]', x = "DIC [&mu;mol kg<sup>-1</sup>]") + theme_bw() + theme(
#     axis.title.x = element_markdown(),
#     axis.title.y = element_markdown()
#   )

# https://stackoverflow.com/questions/12514612/how-to-annotate-ggplot-with-latex
ggplot(DIC_data, aes(y = mids, x = mean_DIC)) + geom_path(size = 1) + scale_y_reverse() + ylab('Depth (m)') + xlab(TeX('DIC ($\\mu$mol kg$^{-1}$)')) + theme_bw()

```

This gradient is mostly due to processes often referred to as carbon pumps, operating against continuous ocean mixing processes that want to homogenize DIC in the water column [@Sarmiento2006-ec; @Claustre2021-tx]. They were originally defined by @Volk1985-oc as "processes that deplete the ocean surface of DIC relative to the deep-water DIC".

### Several carbon pumps at work {#sec-carbon_pumps}

#### Solubility pump {#sec-solubility-pump}

From @eq-carbonic_acid, one can derive that p$CO_{2}$<sup>water parcel</sup> is related to the dissolved $CO_{2}$ concentration by the following relation

$$
pCO_{2}^{water~parcel} = \frac{[CO_{2}]}{K_{0}}  
$$ {#eq-henry_law}

where $K_{0}$ is the solubility coefficient of $CO_{2}$ in seawater.

$K_{0}$ is a function of temperature and salinity with a strong dependence on the former [@Sarmiento2006-ec] such that the solubility of $CO_{2}$ increases with decreasing temperatures. Consequently, waters that cool as they are transported to higher latitudes by the thermohaline circulation will be undersaturated in $CO_{2}$ once they reach the surface. Thus, there will be a net exchange of carbon dioxide from the atmosphere to the ocean, enriching surface waters in DIC before they dive back to the abyss at deep water formation zones in North Atlantic polar regions (e.g. the Labrador Sea) and along the Antarctic shelf in the Southern Ocean [@DeVries2022-of].

The solubility pump is responsible for the great majority of all the anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed DIC gradient according to [@Sarmiento2006-ec] or 35% according to @Toggweiler2003-ye. Both values are used in the literature (e.g. 10% in @Boyd2019-ez and @Iversen2022-qj and 35% in @DeVries2022-of, @Passow2012-zh and @Riebesell2009-qj).

#### Carbonate pump {#sec-carbonate-pump}

Calcifying organisms such as coccolithophorids precipitate biogenic $CaCO_{3}$ (i.e. particulate inorganic carbon) either in the form of calcite or aragonite [@Sarmiento2006-ec] as shown in the equation below

$$
Ca^{2+} + CO_{3}^{2-} \rightleftharpoons CaCO_{3}
$$ {#eq-calcium_carbonate}

Combining @eq-calcium_carbonate, @eq-bicarbonate and @eq-carbonate, one can derive another formulation of calcification

$$
Ca^{2+} + 2HCO_{3}^{-} \rightleftharpoons CaCO_{3} + CO_{2} + H_{2}O
$$ {#eq-calcification}

This calcification occurs in surface waters, releasing $CO_{2}$ in the surrounding waters and subsequently to the atmosphere. That is why this pump is sometimes also referred to as the carbonate counter pump [@Zeebe2012-zk]. Once calcifying organisms die, their shells sink to depth and finally dissolve. Therefore, @eq-calcification shows that the carbonate pump also contributes to the observed DIC gradient, with an estimated contribution of 20% [@Sarmiento2006-ec].

#### Biological pump {#sec-biological-pump}

Also known as the soft-tissue pump [@Volk1985-oc], the biological carbon pump (BCP), is defined by @Ipcc2019-qc as "a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis (@eq-photosynthesis) in sunlit surface water and then transported to the ocean interior, and possibly the sediment, resulting in the storage of carbon".

$$
CO_{2} + H_{2}O \rightarrow CH_{2}O + O_{2} 
$$ {#eq-photosynthesis}

The BCP would thus be responsible for the remaining 70% of the observed DIC gradient according to @Sarmiento2006-ec. However, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess $CO_{2}$, released by anthropogenic activities. First, as explained in @Sarmiento2006-ec, atmospheric $CO_{2}$ levels were stable for several thousand years before the industrial revolution started, therefore it is assumed that the BCP was operating at steady state. Second, phytoplankton organisms are mostly controlled by light and nutrients availability, not by $CO_{2}$.

The BCP is a natural-based solution for carbon sequestration by definition (@eq-photosynthesis). However, natural-based solutions for carbon sequestration are generally aligned with the concept of 'blue carbon' defined by @Ipcc2019-qc as "all biologically-driven carbon fluxes and storage in marine systems that are amenable to management" which usually encompasses only coastal ecosystems [@Ipcc2019-qc; @Lovelock2019-dr].

Coastal blue carbon ecosystems (e.g. mangroves, tidal marshes, seagrass meadows) are thought to store in their living biomass (years to decades) and sediment (millenia) more than 30 Pg C with an estimated absorption of \~0.3 Pg C year$^{-1}$ [@Macreadie2021-rb] though those numbers should be taken with caution because there are major uncertainties in the distribution of seagrass meadows and tidal marshes [@Macreadie2021-rb] as well as in carbon burial rates [@Williamson2022-ud]. In spite of that, the current expectation of additional carbon removal by coastal blue carbon by 2100 should be around 25-30 Pg C [@Williamson2022-ud; @Gattuso2021-cr], that is the equivalent of 2.5-3 years of current anthropogenic emissions [@Friedlingstein2021-kc].

In comparison, the recent work of @Nowicki2022-no suggests that the BCP exports on average 10.2 Pg C year$^{-1}$, sequesters \~1,300 Pg C with a mean sequestration time of 127 years and buries 0.4 Pg C year$^{-1}$ in the sediment that will remain sequestered for millions of years. Therefore, when talking about natural-based solutions for carbon sequestration, one should keep in mind that even though the BCP does not contribute to the absorption of anthropogenic $CO_{2}$ *per se*, it remains the main process that maintains the DIC gradient.

However, climate change is most likely going to impact the biological (e.g. nutrients availability, shift in phytoplankton communities), physical (e.g. temperature) and chemical (e.g. acidification) drivers of the BCP [@Bopp2013-jx; @Passow2012-zh; @Riebesell2009-qj] whose long-term carbon sequestration capacity remains uncertain [@Passow2012-zh; @Turner2015-af; @Wilson2022-bf; @DeVries2022-of].

#### Microbial pump {#sec-microbial-pump}

Unlike the three previous carbon pumps, the microbial carbon pump (MCP) was only introduced recently by @Jiao2010-ax as "a conceptual framework to address the role of microbial generation of recalcitrant dissolved organic matter[^intro-1] and relevant carbon storage". Put simply, it is a suite of marine microbial mechanisms that produce longer-lived DOC such as semi-refractory, refractory and ultra-refractory DOC that have an average lifetime of respectively 205, 16,000 and 40,000 years [@Legendre2015-jc].

[^intro-1]: i.e. dissolved organic matter resistant to microbial utilization that can persist in the ocean interior for thousand of years.

The term 'pump' is attributed to the MCP because it maintains a gradient between DOC$_{<100~years}$ (14 Pg C) and DOC$_{>100~years}$ (non-bioavailable DOC, 636 Pg C) where 100 years has been chosen as the minimum time DOC needs to be stored away from the atmosphere to be considered as sequestered [@Ipcc2019-qc].

In contrast with the other pumps which are depth-dependent, the MCP sequesters carbon by biochemically transforming DOC to non-bioavailable DOC. This process occurs at all depths, making the MCP depth-independent [@Legendre2015-jc].

In order to fully understand the fate of DOC and DIC in the ocean, one would need to integrate both classic carbon pumps and the MCP. However, in this manuscript, I will only focus on the BCP, which was still considered as a single process until recently.

## A new paradigm for the BCP {#sec-a-new-paradigm-for-the-bcp}

The amount of carbon that is stored for more than 100 years in the ocean (i.e. sequestered) mainly depends on the strength and the efficiency of the BCP. In this manuscript, the strength of the BCP is defined as "the magnitude of the particulate organic carbon (POC) flux at a reference export depth" [@Claustre2021-tx] and its efficiency (also known as transfer efficiency) is defined as the fraction of POC flux transported to some reference depth below the export depth [@Buesseler2020-sn; @Claustre2021-tx]. However, the export depth is not unique. It is often set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where POC is produced (i.e. via photosynthesis) and where POC is degraded. This degradation occurs through the remineralization (i.e. respiration) where organic matter is converted back to inorganic forms (i.e. nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd]. The bulk of respiration occurs in the mesopelagic zone between the export depth and 1000 m [@Martin1987-um].

In the frame of the BCP, global carbon budgets have generally been imbalanced such that the exported POC was up to 2-3 orders of magnitude less abundant that it would need to be to sustain heterotrophic respiration in the mesopelagic zone and beyond [@Burd2010-hc; @Turner2015-af]. This imbalance raises one question: are we currently underestimating the export carbon flux or are we overestimating the heterotrophic respiration?

The last part of the question is out of the scope of this manuscript but it should be said that a lot of uncertainties orbit around those carbon deficits, namely parameters used to estimate the metabolic activity of deep ocean organisms as well as deep food web dynamics and unquantified carbon sources such as the contribution of slowly settling particles and better estimation of POC fluxes [@Burd2010-hc].

Regarding the latter, @Boyd2019-ez published a paper in which they review the BCP by adding new biological and physical mechanisms[^intro-2] that inject suspended and sinking particles to depth (@fig-global_carbon_pumps). A description of those mechanisms is provided in the next subsections.

[^intro-2]: More specifically, as stated by @Claustre2021-tx "the word 'biological' in BCP refers to the transfer of biogenic material to depth, mediated by either biological or physical processes."

![Illustration of the three main mechanisms through which the BCP exports carbon to the deep ocean. 1) Production of POC and transport to depth by gravitational settling. 2) Vertical migration of zooplankton that feed at the surface and metabolize the ingested carbon at depth (e.g. production of fecal pellets). 3) Subduction of DOC and suspended POC via mixing processes. Sequestration times for DIC are set depending on the depth at which POC and DOC are exported. Figure taken from @Iversen2022-qj.](image/intro/carbon_pumps_iversen2022.jpeg){#fig-global_carbon_pumps}

### Gravitational pump {#sec-BGP}

The most well known pump is the gravitational carbon pump (BGP). It consists in the gravitational settling (i.e. sinking) of POC[^intro-3] to the deep ocean and it is the main contributor (60-70%) of the export flux [@Boyd2019-ez; @Nowicki2022-no]. Historically, the downward POC flux was directly measured using sediment traps [@Martin1987-um; @Honjo2008-kn] and indirectly measured using thorium-234 ($^{234}Th$) deficits as a POC flux proxy [e.g. @Buesseler2006-og]. More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size spectra (PSS) that could then be converted to carbon fluxes [e.g. @Guidi2008-vp].

[^intro-3]: There obviously exists a gravitational settling of particulate inorganic carbon but in this manuscript, I focus on the BCP which was defined earlier as only focused on organic carbon as opposed to the carbonate pump.

The BGP was long thought to be the only component of the BCP therefore new instruments are being tested to decouple the downward POC flux signal from the contribution of other pumps (@sec-BGC-sensors).

### Mixing pump {#sec-MixingPump}

The mixing pump is a physically mediated particle injection pump (PIP). It consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez]. Contrarily to the BGP, all three pumps also transport DOC to depth where it is injected along with suspended (i.e. neutrally buoyant) and slow-sinking POC below the EZ on different spatial scales and timescales.

The mixed layer pump can be defined as a process through which carbon is exported below the EZ due to variation in the depth of the mixed layer [@DallOlmo2016-ex]. The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on timescales of days across horizontal spatial scales of 1-10 km [@Omand2015-mf; @Boyd2019-ez]. The large-scale (100-1000 km) subduction pump corresponds to the slow advective transport of carbon associated with the meridional overturning circulation and Ekman pumping [@Levy2013-ro; @Claustre2021-tx; @Siegel2022-wd]. The last two pumps are both 2D spatial processes (i.e. they transport POC and DOC vertically and laterally).

The mixing pump would represent at least 20% of the export flux according to @Nowicki2022-no although large uncertainties remain (see Figure 2C in the paper of @Boyd2019-ez).

### Migrant pump {#sec-MigrantPump}

The migrant pump, also known as the vertical migration pump (VMP), is a biologically mediated PIP [@Boyd2019-ez]. It consists on the vertical migration of zooplankton or fish [@Steinberg2017-ri; @Saba2021-ew] that directly inject POC (e.g. fecal pellets) below the EZ at the depth of their migration. This pump can act on daily and seasonal timescales. In the case of the diel (i.e. daily) migration pump, animals feed at the surface at night and then migrate to depth at sunrise to avoid predation [@Siegel2022-wd], metabolizing injected food at depth[^intro-4], by-passing the EZ directly [@Steinberg2017-ri]. In the case of the seasonal migration pump, carbon-rich lipids accumulated in the EZ during summer are catabolized by hibernating copepods at depth, between 600 m and 1,400 m, where they shunt carbon below the permanent pycnocline [@Jonasdottir2015-ob]. This pump is also referred to as the seasonal lipid pump [@Jonasdottir2015-ob].

[^intro-4]: They can also die at depth during the day (or their overwintering in the case of the seasonal migration pump) which also contributes directly to the export of carbon at depth.

Based on @Nowicki2022-no, a lower boundary for the migrant pump export flux could be approximately 10% of the total export taking into account the fact that the seasonal lipid pump was ignored in their model as well as the vertical migration of mesopelagic fish that could export a significant amount of carbon (e.g. 1.5 Pg C year$^{-1}$ according to @Saba2021-ew). If the export of carbon by the migrant pump is somehow lower than the mixing pump [@Nowicki2022-no], its sequestration potential is higher due to its deeper injection depth (@tbl-PumpDepthRange). According to @Boyd2019-ez, the diel migration pump could potentially store 60% as much carbon as the BGP due to the rapid sinking of fecal pellets that would reach depths such that subsequent remineralized DIC would be hidden from the surface for a long period (\> 100 years) of time (@tbl-PumpDepthRange).

However, there are still many uncertainties regarding the magnitude of the migrant pump [@Archibald2019-dj; @Aumont2018-wx; @Gorgues2019-qp] and its sequestration potential [@Boyd2019-ez; @Pinti2022-dc] thus its importance for a better understanding of the BCP.

## The fate of carbon {#sec-the-fate-of-carbon}

### Biological processes at play {#sec-biological-processes-at-play}

Sinking particles are part of a pool of highly diversified materials combining fecal pellets, phytodetritus from sinking phytoplankton, dead zooplankton and organic aggregates referred to as marine snow (i.e. macroscopic organic aggregate \> 0.5 mm) [@Alldredge1988-wm; @Turner2015-af; @Steinberg2017-ri]. Settling velocities for such particles range from 10-100 m day$^{-1}$ up to 1 km day$^{-1}$ for large salp fecal pellets [@Steinberg2017-ri; @Iversen2022-qj].

Marine snow is paramount for both the BCP (see @sec-a-new-paradigm-for-the-bcp) and the support of marine life below the EZ [@Burd2010-hc]. The remineralization of organic matter by heterotrophic organisms releases $CO_{2}$ in the water column that will then be released back to the atmosphere in a time frame depending both on its geographical location and depth (see @sec-will-it-be-sequestered). Therefore, understanding how biological processes interact with the downward POC flux and where the bulk of the remineralization occurs is important to better understand the efficiency of the BCP.

The Martin curve [@Martin1987-um] is probably the most well known POC attenuation curve (@eq-Martin-Curve and @fig-Martin-Curve). It is a power-law function described as follow

$$
F_{z} = F_{z_{export}} \left(\frac{z}{z_{export}}\right)^{-b}
$$ {#eq-Martin-Curve}

where $F_{z}$, $z_{export}$ and $b$ are, respectively, the estimated POC flux at a depth $z$, the depth of the export flux and the remineralization factor. The latter is adimensional and often set at 0.86 [@Martin1987-um] although it is well known that the magnitude of flux attenuation varies regionally (and temporally) due to the variability in food web structures beneath the EZ [@Passow2012-zh] thus variable $b$ have been proposed to take that into account [e.g. @Henson2012-ck; @Guidi2015-sj].

```{r, Martin Curve}
#| label: fig-Martin-Curve
#| fig-cap: Typical attenuation curve of the POC flux obtained with @eq-Martin-Curve. The colour-coded layers represent regions with carbon-specific degradation rate, decreasing with depth. Figure inspired from @Iversen2022-qj.

library(scales)

z_export <- 100
F_export <- 1
b <- 0.86

F_martin <- function(z){
  flux <- F_export * ((z/z_export)**(-b))
  return(flux)
}

# test <- F_martin(100)
depth <- seq(100,3000,10)
fluxes <- F_martin(depth)
df <- data.frame(z = depth, flux = fluxes)
opacity <- 1

ggplot(df, aes(x = fluxes, y = depth)) + #geom_path() + 
  # geom_hline(yintercept = 1000, linetype = 'dotted') + 
  # geom_hline(yintercept = 100, linetype = 'dotted') +
  #geom_hline(yintercept = 100, linetype = 'dotted', colour = 'black') +
  # annotate("rect", ymin = 100, ymax = 300, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#22577a") +
  # #geom_hline(yintercept = 300, linetype = 'dotted') +
  # annotate("rect", ymin = 300, ymax = 500, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#38a3a5") + 
  # #geom_hline(yintercept = 500, linetype = 'dotted') +
  # annotate("rect", ymin = 500, ymax = 1000, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#57cc99") +
  # #geom_hline(yintercept = 1000, linetype = 'dotted') +
  # annotate("rect", ymin = 1000, ymax = 3000, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#80ed99") +
  annotate("rect", ymin = 100, ymax = 300, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#edf8b1") +
  annotate("rect", ymin = 300, ymax = 500, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#7fcdbb") + 
  annotate("rect", ymin = 500, ymax = 1000, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#1d91c0") +
  annotate("rect", ymin = 1000, ymax = 3000, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#225ea8") +
  #scale_y_continuous(trans = c('log10', 'reverse')) +
  scale_y_continuous(trans = c('reverse')) + geom_path(size = 1.5) +
  #scale_y_reverse() + 
  theme_bw() +
  labs(x = 'POC flux (mg C m<sup>-2</sup> day<sup>-1</sup>)', y = 'Depth (m)') + theme(
    axis.title.x = element_markdown(),
    axis.title.y = element_markdown())

# viscosity <- function(T){
#   visc <- 0.00042844 + 1/(0.157*((T+64.993)**2)-91.296)
#   return(visc)
# }
# 
# viscosity(0)

```

Based on the Martin curve, @Iversen2022-qj divided the water column in 4 layers (@fig-Martin-Curve): the upper mesopelagic layer (export depth - 300 m) where there is a rapid attenuation of the downward POC flux, the mid-mesopelagic layer (300 - 500 m) and the lower mesopelagic layer (500 - 1000 m) where both layers show two levels of intermediate attenuation and the bathypelagic layer (1000 m - seafloor) where the attenuation is relatively uniform. For each layer, @Iversen2022-qj estimated a carbon-specific degradation rate in order to better understand the processes responsible of this attenuation. As a result, @Iversen2022-qj suggests that the POC attenuation in the upper mesopelagic layer is the combined action of microbial degradation and zooplankton that graze on settling aggregates, the latter being the main responsible for the attenuation whereas in the mid- and lower mesopelagic layer, the microbial degradation takes over due to the decrease in zooplankton abundance. Finally, the relatively constant degradation occurring below 1,000 m in the bathypelagic zone could be explained by the increasing pressure and decreasing temperatures on microbial communities.

The impact on the downward POC flux of zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2022-qj]: aggregate feeding, filter-feeding and direct ingestion of fecal pellets will decrease the POC flux, ingestion of slow-sinking cells to defecation of large and compact (i.e. fast-settling) fecal pellets will probably increase the POC flux while fragmentation of large aggregates or fecal pellets to small slower-sinking particles will decrease it. The remaining organic matter (\~15% according to @Turner2015-af) will then be degraded by microbial communities. In the end, only a small fraction of the exported carbon will reach the sediment where it will be stored for millions of years.

### The importance of the sinking speed {#sec-sinking-speed}

The settling speed $w$ is the most important physical parameter controlling the remineralization depth of aggregates and fecal pellets. Using Stokes' Law for spherical aggregates at low Reynold's number, one can derive [see for e.g. @Johnson1996-ys]

$$
w = (1-p)~\frac{g(\rho-\rho_{0})d^{2}}{18\nu\rho_{0}}
$$ {#eq-sinking-speed}

where $p$, $g$, $\rho$, $\rho_{0}$, $\nu$ and $d$ are, respectively, the porosity of the aggregate, the gravitational acceleration, the average density of the particles composing the aggregate, the density of the surrounding fluid (i.e. seawater), the fluid kinematic viscosity and the aggregate diameter. If one assume that changes in dynamic viscosity are negligible [@De_La_Rocha2007-yi] compared to changes from the remaining parameters (i.e. this assumption is reasonable based on the fact that $\nu$ can roughly vary by a factor 2 [@Sharqawy2010-ig]), then $w$ should mostly be proportional to the excess density $\Delta\rho$ (i.e. $\rho -\rho_{0}$), the porosity and the diameter of sinking aggregates. It is however a simplification of reality because most aggregates are not spherical and have fractal geometries [@Logan1990-gy; @Logan1995-id; @Laurenceau-Cornec2020-zc].

That being said, several processes can impact the sinking speed of POC such as aggregation (e.g. physical coagulation, incorporation of small particles into fecal pellets (i.e. repackaging)), disaggregation or fragmentation (either physically (e.g. fluid shear) or biologically by microbial degradation and zooplankton feeding), solubilization to DOC by bacteria and mineral ballasting [@Turner2015-af].

The driving factor of the sinking speed is nevertheless subject to caution. It would make sense that the quadratic dependence of $w$ to the aggregate size would favor $d$ as the main driver. However, aggregates are far from being perfect theoretical objects for @eq-sinking-speed and the excess density as the main driver is plebiscited by some authors such as @Iversen2020-kb.

To the best of my knowledge, the most emblematic case of discussions around that question is the mineral ballast hypothesis [@De_La_Rocha2007-yi]. Mineral ballasting consists in the inclusion of biominerals such as opal (i.e. silicate) and calcite or lithogenic material such as dust that are thought to increase $\Delta\rho$ on the one hand and protect them from microbial degradation on the other hand [@Armstrong2001-rz; @Francois2002-uc; @Klaas2002-kn]. Nevertheless, there is not a consensus on that matter because [@Ploug2008-vx; @Ploug2008-qk] did not find significant differences in microbial degradation of non ballasted vs. ballasted material while @Passow2004-re stipulates that biominerals could instead be collected (i.e. scavenged) and transported downwards by sinking aggregates and that their integration could fragment aggregates into smaller but denser particles [@Passow2006-hw] which could either decrease (smaller size) or increase (denser particles) their sinking speed.

The sinking speed of organic matter determines the time available for degradation in the water column therefore identifying the main drivers for the settling of aggregates is a key goal of oceanography [@Iversen2022-qj]. The size of aggregates was usually thought to be a key predictor of settling velocity [e.g. @Passow2006-hw] which is coherent with @eq-sinking-speed. However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor. To answer that question, there is a need for direct measurements (size, composition and density) to predict settling velocities [@Iversen2020-kb; @Iversen2022-qj]. This need could be partially filled with the multi-instrumentation approach described in @sec-observing-carbon-cycle.

### Will it be sequestered? {#sec-will-it-be-sequestered}

Once injected in the ocean interior (@tbl-PumpDepthRange), the POC flux originating from the five PIP pumps is merged back with the BGP. How long is this carbon going to be sequestered from the atmosphere is one of the key questions of carbon cycle studies [@Claustre2021-tx].

```{=html}
<!--#https://github.com/quarto-dev/quarto-cli/discussions/1042-->
<style type="text/css">
caption, .table-caption {
  text-align: left;
}
</style>
```
---
tbl-cap-location: top
---

| Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
|------------------------|----------------------------------|---------------------------|--------------------------------------|
| Gravitational          | export depth - sediment          | None                      | N/A (142)                            |
| Mixed layer            | 200 - 1000 m                     | 200 - 1000m               | 25 - 100 (54$^{b}$)                  |
| Eddy subduction        | 150 - 400 m                      | 150 - 400 m               | up to 150 (N/A)                      |
| Large-scale subduction | 200 - 1000 m                     | 200 - 1000 m              | 25 - 100 (54$^{b}$)                  |
| Diel migration         | 200 - 600 m                      | None                      | up to 250 (150$^{c}$)                |
| Seasonal migration     | 600 - 1400 m$^{a}$               | None                      | up to 500 (N/A)                      |

: Range of depths generally reached by POC and DOC via the six BCPs [@Claustre2021-tx] and sequestration time potential evaluated by the model of @Boyd2019-ez with an average estimation from the model of @Nowicki2022-no in brackets. None means that a pump does not transport DOC. N/A means that the pump was not evaluated in the model. (a) 1,400 m corresponds to the overwintering depth of *Calanus finmarchicus* as observed by @Jonasdottir2015-ob. (b) 54 years depicts the mean sequestration time of the mixing pump, without the contribution of the eddy subduction pump (c) Migration of mesopelagic fish was neglected in the model. {#tbl-PumpDepthRange}

In this manuscript, the term sequestration is based on the definition provided by [@Ipcc2019-qc], that is "The long-term removal of carbon dioxide or other forms of carbon from the atmosphere, with secure storage on climatically significant timescales (decadal to century)" with a period of storage of 100 years, which is commonly used in climate-relevant studies [@Lampitt2008-kq; @Passow2012-zh].

A great majority of studies related to carbon sequestration by the BCP usually assumed that long-term storage of DIC occurred if the exported carbon was transported below 1,000 m [e.g. @Lampitt2008-kq; @Passow2012-zh]. However, @Siegel2021-oz used an inverse ocean circulation model to assess the time over which injected $CO_{2}$ (direct injections at different depths) remains sequestered from the atmosphere. For this, they computed a factor $f_{\tau}$ corresponding to the fraction of $CO_{2}$ that will be stored away from the atmosphere for a given time $\tau$. Their results show that $f_{\tau}$ increases with depth and varies geographically. They also show that for a time horizon of 100 years (i.e. $\tau = 100$) and an injection depth at 1,000 m, there is a non negligible leakage of $CO_{2}$ back to the atmosphere with strong regional disparities (e.g. high leakage in the Northwest Atlantic compared to virtually no leakage in the Central Eastern Pacific).

Therefore, if one wishes to estimate the amount of DIC sequestered from the atmosphere by the BCP, one should not use a reference sequestration depth [@Siegel2022-wd]. Instead, the approach of @Siegel2021-oz can be translated to the study of the BCP given the fact that injected $CO_{2}$ can be replaced by remineralized organic matter. This approach will be discussed in CHAPTER XX.

## Research questions and objectives

The quantification of the different carbon export pathways of the BCP and the fate of the exported carbon in the mesopelagic zone are among the hardest problems in modern oceanography [@Siegel2016-kh; @Le_Moigne2019-zl] but also the most pressuring issues in the context of climate change [@Martin2020-wa]. If the BCP is not responsible for the absorption of anthropogenic $CO_{2}$, any changes in its strength or efficiency could greatly impact atmospheric $CO_{2}$ concentrations [@Kwon2009-ir]. Thus, there is a need to better understand the biological and physical drivers of each component of the BCP, estimate the amount of carbon they export to the deep ocean and their sequestration potential.

To this aim, a new observational framework has been proposed by @Claustre2021-tx to combine measurements from several observations platforms (i.e. research vessels, satellites, moorings and automated underwater vehicles (AUVs)). New sensors have been developed for AUVs (see section @sec-BGC-sensors) in order to better characterize the type of particles (e.g. zooplankton, fecal pellets, aggregates), their size and their mechanisms of transformation (e.g. fragmentation, aggregation). Combining independent measurements of specific biogeochemical variables will allow to tackle the double accounting problem, that is the issue of distinguishing particles associated with the BGP and a PIP, therefore potentially attributing the transport of a particle to more than one pump [@Boyd2019-ez]. This is of particular importance to better constrain the contribution of each PIP to carbon storage in the global ocean [@Boyd2019-ez].

In addition, the refinement of the contribution of each PIP would both help to close the ocean carbon budget [@Burd2010-hc] and improve estimates of carbon sequestration [@Boyd2019-ez; @Passow2012-zh] that will enhance the evaluation of the efficiency of carbon dioxide removal strategies [@Siegel2022-wd].

Based on that, the objectives of this work are threefold:

1.  Contribute to the development of new sensors for AUVs, namely the development of the zooplankton embedded recognition algorithm for the Underwater Vision Profiler 6 [@Picheral2021-sx], an imaging system that will be used to characterize particles \> 600 $\mu$m.
2.  Exploit the synergy of instruments on fully-equipped profiling floats in a specific region (Labrador Sea) to shed a light on mesopelagic processes that affect the size and sinking speed of aggregates as well as characterizing the downward flux of small and big particles and its variability.
3.  Reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.

### 2023-01-22 observingCarbonCycle.qmd
start project

# Observing the carbon cycle {#sec-observing-carbon-cycle}

## How it started {#sec-how-it-started}

In the 20th century, the downward POC flux was traditionally measured with sediment traps. Those instruments are composed of one or several collectors of sinking material whose design depends on its use and position in the water column (e.g. moored traps, surface tethered traps, drifting traps) [@Honjo2008-kn]. They were extensively used [@Honjo2008-kn] though their use has been steadily decreasing since 2004, reaching less than 500 measurements per year since 2010 (@fig-sediment-trap-thorium-yearly-distribution).

```{r, poc measurements per year, dpi = 300}
#| label: fig-sediment-trap-thorium-yearly-distribution
#| fig-cap: Yearly distibution of POC flux measurements by sediment traps or derived from Thorium 234 deficits. Compilation based on @tbl-poc-from-sediment-traps and @tbl-poc-from-thorium.

library(ggplot2)
library(viridis)
trap <- vroom::vroom('DATA/Trap_Thorium/TRAP_DATABASE.csv')
thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
trap$year <- lubridate::year(trap$start_date)
thorium$year <- lubridate::year(thorium$start_date)
sub_trap <- trap |> dplyr::select(year, instrument_type)
sub_trap$instrument_type <- 'Sediment trap'
sub_thorium <- thorium |> dplyr::select(year, instrument_type)
sub_thorium$instrument_type <- 'Thorium 234'
combined <- rbind(sub_trap, sub_thorium)
plot <- ggplot(combined, aes(x = year, fill = instrument_type)) + geom_bar() + scale_fill_manual(values = c('#003366', '#DCEEF3')) + # #57cc99 #80ed99
    #geom_vline(aes(xintercept = median(year)), color = "#d53e4f", linetype = "dashed") +
    labs(x = "Year", y = "N records") + 
    theme_bw() + 
    #theme(legend.position = c(0.12, 0.85), legend.background = element_rect(fill = "white", color = "white")) + 
    theme(legend.position = 'top') +
    theme(legend.title=element_blank())

# plot <- ggplot(combined, aes(x = year, fill = instrument_type)) + geom_bar() + scale_fill_manual(values = c('#E85D04', '#FFBA08')) + # #57cc99 #80ed99
#     #geom_vline(aes(xintercept = median(year)), color = "#d53e4f", linetype = "dashed") +
#     labs(x = "Year", y = "N records") + theme_bw() + guides(fill=guide_legend(title="Method"))

plot
```

```{=html}
<!-- 
![Examples of sediment traps. From left to right: (a) Neutrally buoyant, (b) Surface tethered and (c) Funnel-shaped drifting trap. Figure taken from @Baker-XX.](image/ObservingCarbonCycle/sed_traps_high_res.jpg){#fig-sed-traps}
-->
```
```{=html}
<!-- #https://github.com/quarto-dev/quarto-cli/discussions/1042 -->
<style type="text/css">
caption, .table-caption {
  text-align: left;
}
</style>
```
```{r, poc database from traps}
#| label: tbl-poc-from-sediment-traps
#| tbl-cap: Compilation of POC flux measurements from sediment traps. Data taken from public databases listed on PANGAEA or extracted from articles (with the help of Frédéric Le Moigne and Meike Vogt). The number of measurements could be different from the original publication after data cleaning and potential removal of duplicates.
#| tbl-cap-location: bottom

library(dplyr)
library(knitr)
library(kableExtra)
library(purrr)

# read data
trap <- vroom::vroom('DATA/Trap_Thorium/TRAP_DATABASE.csv')
#thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')

# clean a bit
trap[trap$doi == 'https://doi.org/10.1007/s00227-018-3376-1',]$doi <- '10.1007/s00227-018-3376-1'
trap[trap$doi == 'https://doi.org/10.1016/j.dsr.2016.08.015',]$doi <- '10.1016/j.dsr.2016.08.015'
trap[trap$doi == 'Geochimica et Cosmochimica Acta, Vol. 66, No. 3, pp. 457–465, 2002',]$doi <- '10.1016/S0016-7037(01)00787-6'

# some stats
trap_resume <- trap |> group_by(doi) |> dplyr::summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))

# BASED ON : https://stackoverflow.com/questions/48506712/kable-kableextra-cells-with-hyperlinks
# make url links
make_link <- function(doi){
  link <- paste0('https://doi.org/',doi)
  return(link)
}
url <- map_chr(trap_resume$doi, make_link)

# # put url in dataframe
# trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, 'html', link = url)) |> select(doi_url, data_points) |>
#   kable('html', escape = FALSE, col.names = c('Sources of data', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))

trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l')

# make final table
trap_resume
```

```{r, poc database from thorium}
#| label: tbl-poc-from-thorium
#| tbl-cap: Compilation of POC flux measurements derived from Thorium 234. Data taken from public databases listed on PANGAEA or extracted from articles (with the help of Frédéric Le Moigne). The number of measurements could be different from the original publication after data cleaning and potential removal of duplicates.
#| tbl-cap-location: bottom

# read data
thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')

# some stats
thorium_resume <- thorium |> group_by(doi) |> summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
thorium_resume$doi <- c('10.1594/PANGAEA.809717', '10.1029/2018GB00615', '10.1525/elementa.2020.030')

# make url links
url <- c('https://doi.org/10.1594/PANGAEA.809717', 'https://doi.org/10.1029/2018GB006158', 'https://doi.org/10.1525/elementa.2020.030')

# put url in dataframe
thorium_resume <- thorium_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))

# make final table
thorium_resume
```

To estimate the downward POC flux, one therefore needs to quantify the amount of POC collected through a given area over a known time period at a given depth. However, the apparent sinking flux measured with sediment traps can be biased due to hydrodynamic issues that could result in over- or under-collection of particles (especially at the surface), swimmers (zooplankton, small fish) that could enter sediment traps and affect the collected material and solubilization of collected material (e.g. by bacteria) [@Buesseler2007-ql].

Given the potential inaccuracy of flux measurements by sediment traps, independent approaches using radionuclide disequilibria have been used to assess the collection efficiency of the latter [@Buesseler1991-vo]. This method is based on the fact that $^{238}U$ is a conservative element in seawater (chemically inert with a half-life of 4.5 billion years) while its daughter nuclide $^{234}Th$ (half-life of 24.1 days) can be adsorbed onto particle surfaces and therefore be scavenged by sinking particles [@Savoye2006-gb]. Consequently, the deficiency of $^{234}Th$ relative to $^{238}U$ is used as a measure of removal of surface $^{234}Th$ through particle scavenging [@Buesseler1991-vo] that is then derived as a flux of exported $^{234}Th$ and subsequently a flux of exported POC by using a known (e.g. estimated from the material collected in sediment traps) POC/$^{234}Th$ ratio [@Buesseler2007-ql].

While this method seems to cope with some of the issues observed in sediment traps, it also has its caveats, particularly a POC/$^{234}Th$ ratio that varies with particle size and the fact that it only works well in the upper 100-200 m [@Buesseler2007-ql].

In all, POC fluxes derived from sediment traps and $^{234}Th$ have their own uncertainties and errors thus their estimations of downward POC fluxes should not be seen as ground truth. In addition, they are labor-intensive and require ship support. Their relative contribution to the estimation of POC fluxes is thus expected to keep decreasing (@fig-sediment-trap-thorium-yearly-distribution) with the advent of AUVs (see @sec-how-it-has-evolved).

## How it has evolved {#sec-how-it-has-evolved}

It is well known that ship-based measurements are the most precise. However, they lack spatiotemporal coverage and they are usually not cost-effective. Consequently, oceanographers have turned to new platforms and have developed new sensors to study the marine carbon cycle and the BCP in particular.

Instrumented moorings have supplemented ship-based measurements by collecting high-resolution data at fixed locations [@Chai2020-uq]. Moorings can be equipped with many sensors (e.g. sediment traps (@sec-how-it-started) or imaging systems to study the type of sinking particles (@sec-BGC-sensors)) to study carbon-related variables (see Table 1 in @Claustre2021-tx). They are therefore ideal platforms for long-term time series (e.g. Bermuda Atlantic Time-series Study (BATS) since 1988) with the disadvantage of limited spatial coverage and high maintenance costs [@Chai2020-uq].

In contrast, satellite remote sensing have a wide spatial coverage but they are limited to the ocean's surface and their measures can be affected by several factors such as the cloud cover or low solar elevation (e.g. @Marshall1994-vk). They have been used for ocean applications since the late 1970s [@Gordon1980-dm] and have been widely used to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few.

In order to fill the scarcity of data from research vessels and moorings as well as coping with surface-constrained satellite measurements, AUVs have been developed to take *in situ* measurements in all weathers and at all times, only being dependent on their battery life and ocean currents (for drifting AUVs).

Biogeochemical Argo (BGC-Argo) profiling floats (@fig-bgc-float) are probably the most well known AUVs. They are cost-effective, able to monitor biogeochemical processes at a global scale throughout the year and under all weather conditions [@Chai2020-uq]. The goal of the BGC-Argo community is to develop an array of 1000 floats that measure 6 key biogeochemical and bio-optical variables: Chlorophyll a (Chla), particle backscatter ($b_{bp}$), oxygen, nitrate, pH and downwelling irradiance in addition to temperature and salinity [@Roemmich2019-pn]. Besides this objective, the BGC-Argo community is testing new sensors (see @sec-BGC-sensors) to explore specific topics related to the BCP [@Claustre2020-jo].

Gliders are also effective AUVs to unveil the mysteries of the BCP [e.g. @Omand2015-mf]. In comparison with BGC-Argo floats (@fig-BGC-Argo-cycle), gliders mostly operate in shallow waters [@Chai2020-uq] and are able to take measurements along a predefined transect due to their ability to change their internal weight distribution [@Johnson2009-kb].

![Illustration of a typical BGC-Argo 10-day cycle. From the beginning of the cycle (1), the profiling float descends to its parking depth (2) then drifts for \~ 9 days (3) to dive at 2000 m (4) before the ascent where data are acquired (5) and transmitted to satellite while at the surface (6). Note that the profiling float can be programmed (and reprogrammed during its mission via Iridium transmission) with different configurations, depending on its utilization. For the study of the BCP, some floats could be configured to drift at multiple parking depths to study the BGP or to shorter cycle duration to study the VMP. Image taken from @Claustre2020-jo.](image/ObservingCarbonCycle/argo_float_profiling.jpeg){#fig-BGC-Argo-cycle}

The study of the BCP is a really hot topic in oceanography with new platforms (e.g. unmanned surface vehicles such as saildrones or wave gliders) and new technologies (e.g. lidar) being continuously developed in the hope of understanding all its components [@Chai2020-uq; @Claustre2021-tx].

## Specific sensors for the study of the BCP {#sec-BGC-sensors}

In the frame of the REFINE (Robots Explore plankton-driven Fluxes in the marine twIlight zoNE) project, enhanced BGC-Argo floats (@fig-bgc-float) will have two new sensors specifically designed for the study of the BCP: a transmissometer and the Underwater Vision Profiler 6 (UVP6) [@Picheral2021-sx].

![Illustration of a BGC-Argo float used in the frame of REFINE. From top to bottom: GPS & Iridium (two-way communication) antenna, Oxygen optode, CTD (salinity, temperature and depth), pH sensor, radiometer (downwelling photosynthetic available radiation and downwelling irradiance at 380, 412 and 490 nm), fluorometer (Chla and colored dissolved organic matter) and scattering meter ($b_{bp}$), UVP6 (particle size spectra and images), transmissometer (particle beam attenuation) and nitrate sensor. \@Thomas Boniface.](image/ObservingCarbonCycle/float_picture1.png){#fig-bgc-float width="500"}

A transmissometer measures the transmittance ($T_{r}$) of a light beam at a given wavelength. Given $T_{r}$, the beam attenuation coefficient ($c$) is computed as

$$
c = \frac{ln(T_{r})}{r}
$$ {#eq-bean-attenuation}

where $r$ is the transmissometer pathlength.

The value of $c$ is the result of absorption and scattering by particles, water and dissolved substances [@Zaneveld1994-ez]. The transmissometer carried by REFINE BGC-Argo floats uses a light beam at 660 nm whose attenuation coefficient can be splitted into three components: the absorption and scattering of light by particles ($c_{p}$), the absorption by water ($c_{w}$) and the absorption by dissolved organic matter ($c_{DOM}$) [@Jerlov1976-hr]. At 660 nm, $c_{w}$ is assumed constant at 0.364 m$^{-1}$ and $c_{DOM}$ is considered negligible [@Jerlov1976-hr] hence variations of $c$ will be assumed to be variations of $c_{p}$ in what follows.

The use of transmissometers to characterize POC concentrations was shown by @Bishop1999-es who found that POC concentrations were well correlated with $c_{p}$, independently of ocean environment, season or depth. Following that, the idea of using a transmissometer as an Optical Sediment Trap (OST) was firstly applied by @Bishop2004-at where the magnitude of carbon export was quantified based on the particles accumulation on the upward-facing window of vertically oriented transmissometers on autonomous platforms while they were drifting (i.e. parked at a specific depth). This method was later improved by [@Estapa2013-gw] who divided the attenuation signal into a continuous component, that is the continuous accumulation of small particles on the transmissometer window and a discontinuous component resulting in positive jumps likely attributed to the deposition of (rarer) large particles (@fig-estapa-method). @Estapa2017-jr further calibrated the OST carbon flux proxy using codeployed sediment traps and drifting transmissometers thus allowing the conversion of $c_{p}$ attenuation fluxes (units: m$^{-1}$d$^{-1}$) into carbon fluxes (units: mg C m$^{-2}$d$^{-1}$) using empirical POC-to-attenuance (ATN) ratio were ATN (unitless) is the product of $c_{p}$ and the transmissometer pathlength.

```{r, OST on BGC-Argo example with jump}
#| label: fig-estapa-method
#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component, cleaned from spikes and negative jumps (assumed as not related to carbon flux) and a discontinuous (positive jumps, light blue) component. Data from BGC-ARGO float WMO 4903634, drifting at 1000 m during its 9$^{th}$ cycle.
library(tidyverse)
library(ggplot2)
library(vroom)
#library(castr)
library(plyr)
library(scales)
library(lubridate)
library(latex2exp)

# From Louis Terrats
clean_cp_data <- function(file, moving_median_order, threshold) {

  # Store raw data 
  raw_file <- file

  #print(head(file))

  # Remove cp > 2 m-1 
  if (any(file$cp > 2)) {
    file <- file[-which(file$cp > 2),]
  }

  # Store relevant variables in vectors (for speeding up the process)
  cp <- file$cp
  diff_cp <- diff(cp)
  dates <- file$dates

  # Spike removal
  # Spike is identified as a significant increase (threshold = 0.01 m-1 h-1) followed by a return to normal values within the next 3 hours.
  tmp <- {diff(cp) / as.numeric(diff(dates), units = "hours")} %>% abs()
  spikes <- which(tmp > threshold)+1 # why +1?
  if (length(spikes) > 0) {
    # For each spike
    for (j in 1 : length(spikes)) {

      # Get the value measured 3 hours after the spike (= post_spike)
      spike <- spikes[j]
      time_diff <- difftime(dates, dates[spike], unit = "hours") 
      post_spike <- which(time_diff >= 3)[1]

      # If the difference between the spike and the post_spike cp is superior to the threshold : values get back to normal ones
      if (is.na(cp[spike])) {next 
      } else {
        if ({!is.na(post_spike)} && {(cp[spike]-cp[post_spike]) > threshold}) { 
          back_to_normal <- which({(cp - cp[post_spike]) < threshold} & {dates > dates[spike]})[1] # Locate the date of back to normal values 
          cp[(spike):(back_to_normal-1)] <- NA # And remove the spike zone
        } else { 
          if (is.na(post_spike)) { # If post spike is Na, that means spike is near the end of the segment
            cp[spike:length(cp)] <- NA # Remove the end zone (which contains a spike)
          }
        }
      }
    }
  }

  # Store the sum of spikes 
  spikes <- spikes[which(spikes %in% which(is.na(cp)))]
  spikes <- diff_cp[spikes-1] %>% abs() %>% sum(na.rm = TRUE)

  # Remove rows with NA (= remove spike zones)
  if (any(is.na(cp))) {
    file <- file[-which(is.na(cp)),]
  }

  # Moving median
  if (nrow(file) > 0) {
    file$cp <- pastecs::decmedian(file$cp, order = moving_median_order, ends = "fill") %>% pastecs::extract(component = "filtered")
  } else {return(list("data" = file, "plot" = NA, "spikes" = spikes))}

  # Plot showing results and previous data
  plot <- ggplot() + 
    geom_point(data = raw_file, aes(x=dates, y=cp, color = "black"), alpha = 0.5, size = 4) + 
    geom_point(data = file, aes(x=dates, y=cp, color = "blue"), size = 4) +
    scale_x_datetime(labels = date_format("%d-%m@%Hh")) + 
    labs(y = expression(c[p]~"["*m^-1*"]"), x = "Datetime") + 
    scale_colour_manual(name = '', values =c('black'='black','blue'='blue'), labels = c('Raw data','Filtered signal')) +
    # Theme
    theme_bw() +
    theme(text = element_text(size=16), 
          legend.text=element_text(size=16), 
          legend.position = 'right',
          axis.text = element_text(color="black", size = 16),
          plot.title = element_text(face="bold", hjust = 0.5))

  return(list("data" = file, "plot" = plot, "spikes" = spikes)) 
}

# From me
float_4903634 <- vroom::vroom('/home/flo/dox/THESIS/DATA/ARGO/4903634_OMTAB_VERSION.csv')
profile <- 'lovuse017c_016_01'
drift_depth <- '1000 m'

check <- float_4903634 |> select(dates = juld, everything()) |> filter(cycle == profile, park_depth == drift_depth) |> drop_na(cp)
despiked_data <- clean_cp_data(check, 3, 0.01)

# save cleaned Cp signal from spikes to a temporary variable
tmp2 <- despiked_data$data
# init jump values to NA
tmp2$jump <- NA
# compute slope between two adjacent points (except first and last point of drifting time series)
for (i in 2:(nrow(tmp2)-1)){
  #print(i)
  delta_x <- as.numeric(difftime(tmp2$dates[i], tmp2$dates[i-1], units = 'days'))
  delta_y <- tmp2$cp[i]-tmp2$cp[i-1]
  tmp2$jump[i] <- delta_y/delta_x
}

# assign a slope threshold depending on the drifting depth (this was estimated based on all 4 floats, could be refined later but works pretty well)
if(drift_depth == '200 m'){
  threshold <- 0.4
}else if(drift_depth == '500 m'){
  threshold <- 0.22
}else{
  threshold <- 0.177
}

# assign a colour for the plot
tmp2$colour <- 'cleaned signal'
tmp2$colour[which(abs(tmp2$jump)>threshold)] <- 'jump' # abs is needed for 'negative jumps'

# add group to compute slope (for each subgroups of points, separated from a jump)
tmp2$group <- NA

# compute where the jumps are (index value of the array)
jump_index <- which(tmp2$colour == 'jump')

# assign group 
for (i in jump_index){
  for (j in 1:nrow(tmp2)){
    if ((j < i) & (is.na(tmp2$group[j]))){
      tmp2$group[j] <- paste0('group_',i)
    }
  }
}

# add last group (because it is AFTER the last index so it was not computed before)
tmp2$group[which(is.na(tmp2$group))] <- 'last_group'

# compute slope for each subgroups
slope_df <- tmp2 |> filter(colour == 'cleaned signal', jump != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), 
                                                                                                                nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],
                                                                                                                delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),
                                                                                                                delta_y = last_cp-first_cp, slope = delta_y/delta_x)

# remove if only one point (cannot fit a slope with one point)
slope_df <- slope_df |> filter(nb_points > 1)

# compute mean slope
mean_slope <- mean(slope_df$slope)

# estapa relationship (see POSTER)
poc_flux <- 633*(mean_slope**0.77)

# build dataframe to plot each subgroup
part1 <- slope_df |> select(group, time = min_time, cp = first_cp)
part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
part_slope <- rbind(part1, part2)

# spot negative jump
tmp2$colour[which((tmp2$colour == 'jump') & (tmp2$jump < 0))]  <- 'negative jump'

#jump_plot <- ggplot(clean_data, aes(x = dates, y = cp)) + geom_point() + geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour))

# ggplot(despiked_data$data, aes(x = dates, y = cp)) + geom_step(size= 3, fill ='black') + 
#   geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour), size = 3) + 
#   scale_colour_manual(values = c('black','yellow', 'blue')) +
#   theme_bw() +
#   ylab(TeX('$c_{p}~(m^{-1})$')) +
#   xlab('Datetime') +
#   theme(legend.position = c(0.095, 0.875),
#         legend.background = element_rect(fill = "white", color = "black"))

jump <- filter(tmp2, colour == 'jump')

jump_plot <- ggplot(tmp2, aes(x = dates, y = cp, fill = colour)) + geom_point(size= 3) +
  geom_point(data = jump, aes(x = dates, y = cp), size = 3, shape = 21) +
  scale_fill_manual(values = c('#003366','#DCEEF3')) +
  theme_bw() +
  ylab(TeX('$c_{p}~(m^{-1})$')) +
  xlab('Datetime') +
  theme(legend.position = 'top', legend.title = element_blank())

# ggplot(despiked_data$data, aes(x = dates, y = cp)) + geom_point(size= 3, fill ='black') +
#   geom_point(data = jump, aes(x = dates, y = cp), size = 3, shape = 21, fill = '#DCEEF3') +
#   scale_colour_manual(values = c('black','yellow', 'blue')) +
#   theme_bw() +
#   ylab(TeX('$c_{p}~(m^{-1})$')) +
#   xlab('Datetime') +
#   theme(legend.position = 'top')
# 
# jump_plot <- ggplot(base_signal, aes(x = dates, y = cp)) + geom_point(size = 3, shape = 21, fill = '#003366') +
#   geom_point(data = jump, aes(x = dates, y = cp), size = 3, shape = 21, fill = '#DCEEF3') +
#   theme_bw() +
#   ylab(TeX('$c_{p}~(m^{-1})$')) +
#   xlab('Datetime')

#geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = 'red', size = 2) + theme_bw() + 
#labs(x = 'Time', y = 'Cp (1/m)', title = paste0('Mean slope (red): ', round(mean_slope,3), ' m-1 day-1')) # \n', 'POC flux (small particles): ', poc_flux, 'mg'))

jump_plot

```

The UVP6 (@fig-UVP6) is a miniaturized version of the UVP5 [@Picheral2021-sx], an imaging system developed to characterize particles (e.g. size, gray level, taxonomic group). Two versions of the UVP6 have been commercialized: the UVP6-HF (high frequency) and the UVP6-LP (low power) which has been specifically designed for AUVs that have strict energy requirements (e.g. autonomous floats).

The UVP6-LP (hereafter referred to as UVP6) is thus both a particle counter (0.1 mm - 2 mm) and a particle classifier (\> 0.6 mm). The former counts particles in different class sizes in order to compute the particle size distribution (PSD) - also referred to as PSS (@sec-BGP) - while the latter extracts pixel-related and geometric features from each image to classify them (i.e. assign them a taxonomic group) directly on board using artificial intelligence (CHAPTER-XX) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].

![Illustration of the basic functioning of the UVP6. A light beam illuminates a region in front of the camera (imaged volume: 0.7L) which takes pictures at a given temporal resolution and subsequently characterizes and classifies relevant objects. \@Thomas Boniface.](image/ObservingCarbonCycle/UVP6_b.png){#fig-UVP6}

While the gathering of data from automatic on board particles classification will have to increase to study related applications, the use of PSD in carbon flux studies has already been thoroughly explored [@Guidi2008-vp; @Iversen2010-an; @Jouandet2011-bt; @McDonnell2012-ic; @Roullier2014-wf; @Guidi2015-sj; @Guidi2016-nj; @Ramondenc2016-nu; @Kiko2017-qy; @Fender2019-ou; @Giering2020-mc; @Cael2021-qn; @Bisson2022-bt].

Using the PSD, one can compute the number spectrum $n_{i}$ [@Stemmann2004-jv] which can be defined as the normalized concentration of particles in a given size bin $i$. Knowing the mass ($m$) and sinking speed ($w$, see @eq-sinking-speed) of aggregates, one can derive the total carbon flux ($F$)

$$
F = \int_{0}^{\infty} n(d) m(d) w(d) dd
$$ {#eq-carbon-flux}

where $d$ is the aggregate diameter.

Assuming spherical aggregates, $m$ can be written as follow

$$
m(d) = \frac{\pi\rho d^{3}}{6}
$$ {#eq-mass-aggregate}

where $\rho$ is the average density of aggregates.

As explained in @sec-sinking-speed, aggregates have fractal geometries ($D$) hence @eq-sinking-speed and @eq-mass-aggregate can be rewritten as follow

$$
m(d) = \alpha d^{D}
$$ {#eq-mass-aggregate-fractal}

$$
w(d) = \beta d^{D-1}
$$ {#eq-sinking-speed-fractal}

where $\alpha$ and $\beta$ are constant.

Aggregate mass and sinking speed are often described using a power law relationship [@Alldredge1988-lx; @Guidi2008-vp] therefore it can be derived that

$$
mw = Ad^{B}
$$

and therefore from @eq-mass-aggregate-fractal and @eq-sinking-speed-fractal that $A= \alpha\beta$ and $D = \frac{B+1}{2}$.

Due to a finite number ($N$) of size bins, @eq-carbon-flux is rewritten as

$$
F = \sum_{i=1}^{N}n_{i}Ad_{i}^{B}\Delta d_{i}
$$ {#eq-carbon-flux-finite}

where $d_{i}$ is the mid-point of bin $i$.

UVP6 PSD measurements will therefore allow the computation of carbon fluxes assuming that $A$ and $B$ are known. This was done empirically by @Guidi2008-vp who minimized the following cost function

$$
Cost = \sum_{j = 1}^{n} (log(F_{UVP,j})-log(F_{ST},j))^{2}
$$ {#eq-cost-function}

where $F_{UVP,j}$, $F_{ST,j}$ and $n$ are, respectively, the flux estimate for the $j^{th}$ sediment trap measurement, the flux measured in sediment trap $j$ and the number of paired measurements (i.e. match-ups) for trap-UVP comparisons.

@Guidi2008-vp found an optimal couple of parameters $(A, B) = (12.5, 3.81)$ while @Iversen2010-an followed the same procedure and found $(A, B) = (273, 4.27)$ whereas @Fender2019-ou found $(A, B) = (15.4, 1.05)$. @Iversen2010-an assumed that the difference in the optimal couple of parameters was due to the study site used for the minimization (productive upwelling system versus open ocean in @Guidi2008-vp) while @Fender2019-ou suggested that a value of $B < 3$ means that either the density or sinking speed of particles is inversely correlated with particles diameter thus they suspected that the flux was dominated by particles in the lower size detection limit of the UVP they were using. Other values for $A$ and $B$ are being used in the literature [e.g. @Kiko2017-qy] based on a mechanistic approach developed by @Kriest2002-oc to investigate the formation of marine aggregates.

In all, it seems that the choice for the values of $A$ and $B$ is critical if one does not want to under- or overestimate carbon fluxes derived from PSD [@Iversen2010-an; @Fender2019-ou]. Three main limitations highlighted by @Bisson2022-bt for UVP optically derived estimates of flux would be: (1) assuming that $A$ and $B$ are globally valid at all depths, for all UVP models and across all size classes, (2) the size to sinking rate uncertainty and (3) the size to carbon mass uncertainty..

The assumption that all particles of a given size class have the same carbon content and settling rate questions the formulation of the model itself (@eq-carbon-flux-finite). It has been shown that the pool of sinking particles is varied (e.g. compact fecal pellets, porous aggregates) and that the main driver of the sinking speed is currently subject to debate [@sec-sinking-speed]. Therefore, the size-dependent model of [@Guidi2008-vp] could be improved by taking into account the composition and nature of particles and by increasing the number of paired measurements between UVP PSD-derived flux estimations and other direct POC flux estimations (e.g. sediment traps, $^{234}Th$, OST).

## The future of carbon observations {#sec-what-is-next}

The BCP have been studied for many decades by a variety of instruments with their respective pros and cons. Their measurements were unfortunately often taken independently, which hindered their intercomparability due to the low number of concurrent measurements between instruments.

With the advent of fully-equipped AUVs, it is now possible to take high-resolution and independent measurements of key biogeochemical variables. Combining the latter with measurements from conventional platforms (i.e. ships, satellites, moorings) will fulfill the observational framework described by @Claustre2021-tx to investigate the BCP.

This framework will hopefully answer questions raised so far, that is better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and better quantify global estimates of carbon storage.

### 2023-01-23 article3.qmd
update article 3

@@
 All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux.
-During these drift phases, measurements of particles abundance (between 0.6 mm - 2.5 mm) are taken every 30 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths.
+During these drift phases, measurements of particles abundance (14 size classes, between 0.1 mm and 2.5 mm) are taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths.
 This configuration can be adjusted during the mission of the float via two-way Iridium communication.
 
-```{r, init function}
+```{r, load libraries}
 library(akima)
@@
 The profile that concludes the time series for each float is highlighted by a black contour.
-#library(tidyverse)
 
@@
 continuous deposition of small particles on the upward-facing window of the transmissometer) and discontinuous (i.e.
-deposition of stochastic big particles) components of the downward POC flux.
-
-First, the drifting $c_{p}$ signal was cleaned by removing spikes (hypothesized as transient particles not landing on the window or active swimmers [@Estapa2013-gw]) using the despike function (oce package in R, see https://dankelley.github.io/oce) where a spike is defined here as any value higher than 4 times the standard deviation of the difference between raw $c_{p}$ data and a smoothed (i.e.
-reference) $c_{p}$ signal, obtained with a 7-point running median filter.
-Despiked $c_{p}$ data were then smoothed with a 3-point running median filter to ease the detection of jumps, defined as a continuous increase or decrease of $c_{p}$ (@fig-estapa-method_article).
-
-```{r, fig fit}
+deposition of stochastic big particles) components of the downward POC flux, hereafter referred to as, respectively, $F_{small}$ and $F_{big}$.
+
+First, the drifting $c_{p}$ signal was cleaned from spikes (hypothesized as transient particles not landing on the window or active swimmers [@Estapa2013-gw]).
+For this, we computed a 7-point moving median and a 7-point moving median absolute deviation (MAD).
+Then, we computed the anomaly to the moving median as the difference between the drifting $c_{p}$ signal and the moving median.
+We considered that a $c_{p}$ measurement was a spike when the anomaly was above $5.2 \times MAD$ following @Leys2013.
+Despiked $c_{p}$ data were then smoothed with a 7-point moving median filter.
+
+Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach instead of fixed thresholds to detect jumps.
+As a result, we firstly computed the slope between each point, then we computed the z-score (i.e.
+standard score) of each slope assuming a normal distribution of the latter.
+The detection of jumps thus consisted in spotting outliers in the z-score distribution.
+For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles.
+Positive and negative jumps were defined as, respectively, a z-score above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$.
+We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.
+
+In order to compute $F_{small}$, we firstly computed the slope of each group of points between jumps (positive and negative, see @fig-estapa-method_article).
+Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window.
+Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the relationship of @Estapa2022
+
+$$
+F_{small} = 633(0.25~slope_{w})^{0.77}
+$$
+
+where $slope_{w}$ is multiplied by the transmissometer pathlength because @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.
+
+We used the same relationship to compute $F_{big}$ was replacing $slope_{w}$ as the sum of positive jumps (negative jumps were disgarded) divided by the drifting time.
+
+```{r, example of a cp signal analyzed for jumps (with an old method here)}
 #| label: fig-estapa-method_article
@@
 
-Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach to detect jumps.
-For this, we applied a 7-point running slope on drifting $c_{p}$ data at each drifting depth and we then computed the 90$^{th}$ quantile ($Q_{90}$) of each slope distribution with a 10-day increasing dataset since the deployment (Fig-XX) to determine the most appropriate slope thresholds to detect jumps.
-The latter are however subject to a tradeoff between the likely underestimation of $F_{small}$ and the detection of false jumps (without great impact on the value of $F_{big}$ because false jumps have low $c_{p}$ values) if the threshold is too small and the likely overestimation of $F_{small}$ to the detriment of $F_{big}$ with missed jumps if the threshold is too high.
-As shown in Fig-XX, the value of $Q_{90}$ decreases over time for each drifting depth.
-This is most likely explained by the vanishing of big particles (> 0.5 mm in this study) over time (Fig-YY).
-Therefore, we assumed that past a given time, the great majority of jumps have been seen by the transmissometer and that the optimal values for each slope threshold should be found in the period following the bloom up to this given time.
-We thus have found that on a weekly basis, particles in the 0.5 - 1.5 mm range at 1000 m were starting to steadily decrease (assumed to be slowly removed without upwards production) around week 40, that is the beginning of October.
-(maybe find another criteria ..
-thresholds as they currently are give good fit.)
-
-```{r, slope distribution}
-
-# remove bad data from float 6904241
-float_6904241_slope <- float_6904241 |> filter(cycle < 20)
-
-# create dataframe for all drifting data (for all floats)
-all_floats_slope <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
-all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle)
-
-# compute slope for every float, with every size window
-wmo <- c(4903634, 6904240, 6904241, 1902578)
-park_depth <- c('200 m', '500 m', '1000 m')
-#window_size <- c(3,5,7)
-window_size <- 7
-
-# function to clean cp data
-clean_cp_data <- function(data, moving_median_order){
-  if(nrow(data) == 0){ # no cp data
-    return(data)
-  }else{
-    # despike cp data
-    data$cp <- oce::despike(data$cp, reference = 'median', k = 7, replace = 'NA')
-    # median average
-    data$cp <- pastecs::decmedian(data$cp, order = moving_median_order, ends = "fill") |> pastecs::extract(component = "filtered")
-  }
-  return(data)
-}
-
-# compute cp slope
-delta_y <- function(x, window_size){
-  delta_y <- x[window_size]-x[1]
-  return(delta_y)
-}
-
-delta_x <- function(x, window_size){
-  delta_x <- (x[window_size]-x[1])/86400 # units in days
-  return(delta_x)
-}
-
-compute_slope <- function(data, window_size){
-  tmp <- data |> drop_na(cp)
-  despiked_data <- clean_cp_data(tmp, 3) 
-  if(window_size == 3){
-    despiked_data$slope <- slide(despiked_data$cp, 1, delta_y, window_size = 3)/slide(despiked_data$dates, 1, delta_x, window_size = 3)
-  }else if(window_size == 5){
-    despiked_data$slope <- slide(despiked_data$cp, 2, delta_y, window_size = 5)/slide(despiked_data$dates, 2, delta_x, window_size = 5)
-  }else{
-    despiked_data$slope <- slide(despiked_data$cp, 3, delta_y, window_size = 7)/slide(despiked_data$dates, 3, delta_x, window_size = 7)
-  }
-  return(despiked_data)
-}
-
-res <- data.frame()
-for (i in wmo){
-  #print(i)
-  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
-  for (j in park_depth){
-    for (k in seq(1:max_cycle)){
-      #print(k)
-      tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
-      if(nrow(tmp) == 0){
-        next
-      }else if(nrow(tmp) < 3){ # case where there is not enough data
-        next
-      }else{
-        cleaned_signal <- clean_cp_data(tmp, 3)
-        if(nrow(cleaned_signal) == 0){
-          next
-        }else{
-          for (l in window_size){
-            test <- compute_slope(cleaned_signal, l)
-            test <- test |> select(dates, wmo, cycle, park_depth, slope)
-            # add time_drifting
-            max_time <- max(test$dates)
-            min_time <- min(test$dates)
-            drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))
-            test$drifting_time <- drifting_time
-            test$window_size <- l
-            res <- rbind(res, test)
-          } 
-        }
-      }
-    }
-  }
-}
-
-fixed_dates <- c('2022-06-01', '2022-06-10', '2022-06-20', 
-                 '2022-07-01', '2022-07-10', '2022-07-20', 
-                 '2022-08-01', '2022-08-10', '2022-08-20', 
-                 '2022-09-01', '2022-09-10', '2022-09-20', 
-                 '2022-10-01', '2022-10-10', '2022-10-20', 
-                 '2022-11-01', '2022-11-10', '2022-11-20', 
-                 '2022-12-01', '2022-12-10', '2022-12-20',
-                 '2023-01-01')
-
-res2 <- data.frame()
-for (i in fixed_dates){
-  tmp <- filter(res, dates < i)
-  tmp2 <- tmp |> group_by(park_depth, window_size) |> dplyr::summarise(threshold = quantile(slope, 0.90, na.rm = T), nb_points = n())
-  tmp2$group <- i
-  tmp2$quantile <- 90
-  res2 <- rbind(tmp2, res2)
-}
-
-# add levels
-res2$park_depth <- factor(res2$park_depth, levels = c('200 m', '500 m', '1000 m'))
-
-# plot result
-# res3 <- res2 |> mutate(juld = as_date(group)) |> ungroup()
-# ggplot(res3, aes(x = juld, y = threshold, colour = factor(window_size), group = factor(window_size))) + geom_point() + geom_path() + 
-#   facet_wrap(~park_depth, scales = 'free') + theme_bw() +
-#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   scale_color_brewer(palette = 'Dark2') +
-#   labs(x = 'Month', y = 'Slope')
-res3 <- res2 |> mutate(juld = as_date(group)) |> ungroup()
-ggplot(res3, aes(x = juld, y = threshold)) + geom_point() + geom_path() + 
-  facet_wrap(~park_depth, scales = 'free') + theme_bw() +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
-  geom_vline(xintercept = as_date('2022-10-01'), colour = "red", linetype = 'dashed') +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  scale_color_brewer(palette = 'Dark2') +
-  labs(x = 'Month', y = 'Slope')
-
-```
-
-As a result, we have chosen to set the slope thresholds to detect jumps at week 40 for all three parking depths at, in increasing order with depth, 0.23, 0.12 and 0.10.
-Using those thresholds, we computed the slope between each $c_{p}$ measurement of each drifting periods and identified as jump any point above the selected thresholds.
-This method allowed us to separate the continuous component of the downward POC flux into groups, between each jump.
-We then computed the slope for each group (light blue straight lines in @fig-estapa-method_article).
-Any negative slope was disregarded assuming that either the so-called fit failed (ineffective threshold leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window.
-Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each group in order to derive the small particles flux ($F_{small}$) using the relationship of @Estapa2022
-
-$$
-F_{small} = 633(0.25~slope_{w})^{0.77}
-$$ {#eq-estapa}
-
-where $slope_{w}$ is multiplied by the transmissometer pathlength because @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.
-
-The discontinuous component of the downward POC flux or big particle flux ($F_{big}$) was derived using @eq-estapa by replacing $slope_{w}$ as the sum of positive jumps (negative jumps were disgarded) divided by the drifting time.
-
-All profiles analyzed with this method are available at https://fricour.shinyapps.io/CARBON_REFINE/.
-
 ## Results
@@
 
-Using the selected slope thresholds determined in the previous section, we computed $F_{small}$ and $F_{big}$ at each drifting depth for all floats (@fig-OST-flux).
-At 200 m, $F_{small}$ ranges between 5.1 and 26.9 mg C m$^{-2}$ day$^{-1}$ with an average of 14.1 $\pm$ 3.9 mg C m$^{-2}$ day$^{-1}$ (standard deviation).
-At 500 m, $F_{small}$ is comprised between 6.9 and 19.8 mg C m$^{-2}$ day$^{-1}$ with an average of 9.5 $\pm$ 2.1 mg C m$^{2}$ day$^{-1}$.
-At this depth, we observe a global decrease of $F_{small}$ with time though float 6904240 measured a relatively constant downward $F_{small}$ at 8.6 mg C m$^{-2}$ day$^{-1}$ on average.
-At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid-July followed by a steady decrease towards December, with $F_{small}$ ranging between 4.0 and 11.8 mg C m$^{-2}$ day$^{-1}$ and 7.0 $\pm$ 1.4 mg C m$^{-2}$ day$^{-1}$ on average.
-
-```{r, function plot jump}
-# plot_jumps <- function(clean_data, threshold){
-#   
-#   # save bbp data for later
-#   #bbp_df <- clean_data |> select(dates, bbp) |> drop_na(bbp)
-#   
-#   # save cleaned Cp signal from spikes to a temporary variable
-#   tmp2 <- clean_data |> arrange(dates)
-#   #print(colnames(tmp2))
-#   # init jump values to NA
-#   tmp2$jump <- NA
-#   # compute slope between two adjacent points (except first and last point of drifting time series)
-#   for (i in 2:(nrow(tmp2)-1)){
-#     #print(i)
-#     delta_x <- as.numeric(difftime(tmp2$dates[i], tmp2$dates[i-1], units = 'days'))
-#     delta_y <- tmp2$cp[i]-tmp2$cp[i-1]
-#     tmp2$jump[i] <- delta_y/delta_x
-#   }
-#   
-#   # assign a colour for the plot
-#   tmp2$colour <- 'base signal'
-#   tmp2$colour[which(abs(tmp2$jump)>threshold)] <- 'jump' # abs is needed for 'negative jumps'
-#   
-#   # add group to compute slope (for each subgroups of points, separated from a jump)
-#   tmp2$group <- NA
-#   
-#   # compute where the jumps are (index value of the array)
-#   jump_index <- which(tmp2$colour == 'jump')
-#   
-#   # assign group 
-#   for (i in jump_index){
-#     for (j in 1:nrow(tmp2)){
-#       if ((j < i) & (is.na(tmp2$group[j]))){
-#         tmp2$group[j] <- paste0('group_',i)
-#       }
-#     }
-#   }
-#   
-#   # add last group (because it is AFTER the last index so it was not computed before)
-#   tmp2$group[which(is.na(tmp2$group))] <- 'last_group'
-#   
-#   # compute slope for each subgroups
-#   slope_df <- tmp2 |> filter(colour == 'base signal', jump != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), 
-#                                                                                                                   nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],
-#                                                                                                                   delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),
-#                                                                                                                   delta_y = (last_cp-first_cp)*0.25, slope = delta_y/delta_x)
-#   
-#   # remove negative slope from the mean slope (no physical meaning)
-#   slope_df <- slope_df |> filter(slope > 0)
-#   
-#   # remove if only one point (cannot fit a slope with one point)
-#   slope_df <- slope_df |> filter(nb_points > 1)
-#   
-#   # compute mean slope 
-#   #mean_slope <- mean(slope_df$slope)
-#   mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)
-#   
-#   # estapa relationship (see POSTER)
-#   #poc_flux <- 633*((mean_slope*0.25)**0.77) # *0.25 because ATN = cp*0.25
-#   poc_flux <- 633*((mean_slope)**0.77) # /!\ slope for computed for ATN on y axis (delta_y *0.25 because ATN  = cp*0.25) -> should be OK
-#   
-#   # build dataframe to plot each subgroup
-#   part1 <- slope_df |> select(group, time = min_time, cp = first_cp)
-#   part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
-#   part_slope <- rbind(part1, part2)
-#   
-#   # spot negative jump
-#   tmp2$colour[which((tmp2$colour == 'jump') & (tmp2$jump < 0))]  <- 'negative jump'
-#   
-#   # add big particles flux
-#   rows_to_keep <- c(jump_index, jump_index-1) 
-#   tmp3 <- tmp2[rows_to_keep,] |> select(dates, cp, jump, colour, group) |> dplyr::arrange(dates)
-#   
-#   # remove negative jump, if any
-#   check_colour <- unique(tmp3$colour)
-#   if(length(check_colour) == 3){
-#     index_neg_jump <- which(tmp3$colour == 'negative jump')
-#     tmp3 <- tmp3[-c(index_neg_jump, index_neg_jump+1),]
-#     tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp)) 
-#     even_indexes <- seq(2,nrow(tmp3),2)
-#     tmp3 <- tmp3[even_indexes,]
-#   }else if(length(check_colour) == 2){
-#     tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp)) 
-#     even_indexes <- seq(2,nrow(tmp3),2)
-#     tmp3 <- tmp3[even_indexes,]
-#   }else{ # NO jump
-#     tmp3 <- NULL
-#   }
-#   
-#   # big particles flux
-#   if(is.null(tmp3)){ # no jumps
-#     big_part_poc_flux <- 0
-#   }else{
-#     tmp4 <- tmp3 |> filter(diff_jump > 0)
-#     if(nrow(tmp4) == 0){ # no positive jumps
-#       big_part_poc_flux <- 0
-#     }else{
-#       delta_y <- sum(tmp4$diff_jump) *0.25 # to get ATN (= cp*0.25)
-#       max_time <- max(tmp2$dates)
-#       min_time <- min(tmp2$dates)
-#       delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
-#       slope_big_part <- delta_y/delta_x
-#       big_part_poc_flux <- 633*(slope_big_part**0.77)
-#     }
-#   }
-#   
-#   # compute total drifting time
-#   max_time <- max(tmp2$dates)
-#   min_time <- min(tmp2$dates)
-#   drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))
-#   
-#   jump_plot <- ggplot(clean_data, aes(x = dates, y = cp)) + geom_point(size= 3) + geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour), size = 3) + 
-#     scale_colour_manual(values = c('#003366','#E31B23', '#FFC325')) + 
-#     scale_shape_manual(values = c(1,1,4)) +
-#     geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', size = 2) + theme_bw() + 
-#     labs(x = 'Time', y = 'Cp (1/m)', title = paste0('Drifting time: ', round(drifting_time,3), ' days\n',
-#                                                     'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n', 
-#                                                     'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
-#                                                     'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1')) #+
-#   #geom_point(data = bbp_df, aes(x = dates, y = bbp), colour = "pink")
-#   
-#   
-#   # adapt script to return big and small flux
-#   df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = clean_data$park_depth[1], wmo = clean_data$wmo[1],
-#                cycle = clean_data$cycle[1])
-#   
-#   #return(jump_plot)
-#   #return(list('jump_plot' = jump_plot, 'jump_table' = tmp3))
-#   return(df)
-#   
-# }
+```{r, function to detect and plot jumps}
 
@@
   # despike cp data with a 7-point moving window
-  tmp$cp <- despike(tmp$cp, k = 7)
+  tmp$cp <- despike(tmp$cp, k = 3)
 
@@
   # spot outliers on the Z score signal
-  # interquartile range between Q25 and Q75
+  # interquartile range between Q25 and Q75 -> had to used that and not the despike function because slopes are often close (or equal) to 0 so it can miss clear jumps.
+Q25 and Q75 are more trustworthy in this case than the despike function of Jean-Olivier (see package castr on his github: https://github.com/jiho/castr)
   IQR <- quantile(tmp$zscore, probs = 0.75, na.rm=T) - quantile(tmp$zscore, probs = 0.25, na.rm=T)
@@
 
-  # compute weighted average slope (to take into account the fact that some subgroup might have 2 points and a high slope vs. large group of points with a small slope)
+  # compute weighted average slope (to take into account the fact that some subgroups might have 2 points and a high slope vs. large group of points with a small slope)
   mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)
@@
   # convert cp to POC using Estapa's relationship 
-  poc_flux <- 633*((mean_slope)**0.77) # /!\ slope computed for ATN on y axis (delta_y *0.25 because ATN = cp*0.25) -> should be OK
+  poc_flux <- 633*(mean_slope**0.77) # /!\ slope computed for ATN on y axis (delta_y *0.25 because ATN = cp*0.25) -> should be OK
 
@@
 
-  # remove negative jump, if any
+  # remove negative jumps, if any
   check_colour <- unique(tmp2$colour)
-  if(length(check_colour) == 3){ # there is a negative jump
+  if(length(check_colour) == 3){ # there is at least a negative jump
     index_neg_jump <- which(tmp2$colour == 'negative jump')
@@
     tmp2 <- tmp2[even_indexes,]
-  }else if(length(check_colour) == 2){ # only positive jump
+  }else if(length(check_colour) == 2){ # only positive jumps
     tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
@@
 
-  if(is.null(tmp2)){ # no jumps
+  if(is.null(tmp2)){ # no jump
     big_part_poc_flux <- 0
+    tmp3 <- NULL
   }else{
@@
 
+  # plot
   jump_plot <- plot_ly(tmp, x = ~dates, y = ~cp, type = 'scatter', mode = 'markers', color = ~colour, colors = c('#003366','#E31B23', '#FFC325')) |>
@@
     # adapt script to return big and small flux
-  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = clean_data$park_depth[1], wmo = clean_data$wmo[1],
-               cycle = clean_data$cycle[1])
-
-  #return(jump_plot)
-  #return(list('jump_plot' = jump_plot, 'jump_table' = tmp3))
+  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],
+               cycle = data$cycle[1])
+
   return(df)
@@
 
-```{r}
-#| label: fig-OST-flux
-#| fig-cap: OST-derived small (top) and big (bottom) particles flux for all floats at each drifting depth.
-Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.
+```{r, compute Fsmall and F big}
+
+# remove bad data from float 6904241
+float_6904241_slope <- float_6904241 |> filter(cycle < 20)
+# create dataframe for all drifting data (for all floats)
+all_floats_slope <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
+all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle)
 
@@
 
-res4 <- data.frame()
+res <- data.frame()
 for (i in wmo){
-  #print(i)
   max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
@@
     for (k in seq(1:max_cycle)){
-      #print(k)
       tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
@@
       }else{
-        cleaned_signal <- clean_cp_data(tmp, 3)
-        if(nrow(cleaned_signal) == 0){
-          next
-        }else{
-          #for (l in window_size){
-            #print(l)
-            #for (m in quartile){
-              #print(m)
-          if(j == '200 m'){
-            slope_threshold <- 0.23
-          }else if(j == '500 m'){
-            slope_threshold <- 0.12
-          }else{
-            slope_threshold <- 0.1
-          }
-          #selected_threshold <- threshold_table |> filter(park_depth == j, window_size == l, quartile == m)
-          #slope_threshold <- selected_threshold$slope_threshold
-          output <- plot_jumps(cleaned_signal, slope_threshold)
-          #print(output)
-          # append result
-          res4 <- rbind(res4, output)
-            #}
-          #} 
-        }
+          output <- plot_jumps(tmp)
+          res <- rbind(res, output)
       }
@@
 # keep the good fluxes
-cflux <- res4 |> mutate(drifting_time = difftime(max_time, min_time, units = 'days'), WMO = factor(wmo))
-
-info_table <- cflux |> dplyr::group_by(park_depth) |> summarize(min_smallf = min(small_flux, na.rm=T), max_smallf = max(small_flux, na.rm=T),
+cflux <- res |> mutate(drifting_time = difftime(max_time, min_time, units = 'days'), WMO = factor(wmo))
+
+cflux_info_table <- cflux |> dplyr::group_by(park_depth) |> summarize(min_smallf = min(small_flux, na.rm=T), max_smallf = max(small_flux, na.rm=T),
                                                                 mean_smallf = mean(small_flux, na.rm=T), median_smallf = median(small_flux, na.rm=T),
@@
 
-###################################
-# MOST RECENT VERSION OF THE PLOT #
-###################################
+```
+
+$F_{small}$ and $F_{big}$ were computed at each drifting depth for all floats (@fig-OST-flux).
+At 200 m, $F_{small}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ (average $\pm$ standard deviation).
+At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$.
+At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid-July followed by a steady decrease towards January, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.
+
+```{r, plot Fsmall and Fbig}
+#| label: fig-OST-flux
+#| fig-cap: OST-derived small (top) and big (bottom) particles flux for all floats at each drifting depth.
+Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.
 
 up <- cflux |> ggplot(aes(x = date, y = small_flux)) +
+    geom_point(data = cflux, aes(x = date, y = small_flux, colour = WMO, shape = WMO)) +
   geom_smooth(data = cflux, aes(x = date, y = small_flux), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + #facet_grid(~park_depth)
-  #scale_x_date(labels = date_format("%b"), date_breaks = '1 month') +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  #theme(axis.text.x = element_text(angle = 45, vjust = 0, hjust=0)) +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
   guides(fill=guide_legend(title="WMO")) +
-  #geom_pointrange(data = info_table2, aes(x = month, y = mean_small_f, ymin = mean_small_f - sd_small_f, ymax = mean_small_f + sd_small_f))
-  #theme_bw() +
-  geom_point(data = cflux, aes(x = date, y = small_flux, colour = WMO, shape = WMO)) +
   facet_wrap(~park_depth, scales = 'free') +
   labs(x = '', y = TeX('$F_{small}$ (mg C m$^{-2}$ day$^{-1}$)')) +
-  theme_bw() #+
-  # theme(axis.ticks.x = element_blank(),
-  # axis.text.x = element_blank(), legend.position = 'top')
-  #theme(legend.position = '')
+  scale_y_continuous(trans = 'log10') +
+  theme_bw()
 
 down <- cflux |> ggplot(aes(x = date, y = big_flux)) +
+  geom_point(data = cflux, aes(x = date, y = big_flux, colour = WMO, shape = WMO)) +
   geom_smooth(data = cflux, aes(x = date, y = big_flux), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + #facet_grid(~park_depth)
-  #scale_x_date(labels = date_format("%b"), date_breaks = '1 month') +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  #theme(axis.text.x = element_text(angle = 45, vjust = 0, hjust=0)) +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
   guides(fill=guide_legend(title="WMO")) +
-  #geom_pointrange(data = info_table2, aes(x = month, y = mean_small_f, ymin = mean_small_f - sd_small_f, ymax = mean_small_f + sd_small_f))
-  #theme_bw() +
-  geom_point(data = cflux, aes(x = date, y = big_flux, colour = WMO, shape = WMO)) +
   facet_wrap(~park_depth, scales = 'free') +
@@
   theme_bw() +
+  scale_y_continuous(trans = 'log10') +
   theme(legend.position = 'none')
 
-#plot_grid(up, down, labels = c('A', 'B'), label_size = 12, nrow = 2)
 up / down
@@
 
-At 200 m, we estimated $F_{big}$ between 0 (no deposition of big particles on the transmissometer window) and 285.0 mg C m$^{-2}$ day$^{-1}$ with an average of 30.7 $\pm$ 54.4 mg C m$^{-2}$ day$^{-1}$ and a median of 11.8 mg C m$^{-2}$ day$^{-1}$.
+$F_{big}$ was estimated at 200 m between 0 (no deposition of big particles on the transmissometer window) and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$.
 High fluxes (\> 100 mg C m$^{-2}$ day$^{-1}$) were measured by 3 out of 4 floats at the end of May before a sharp decrease towards July.
-At 500 m, $F_{big}$ ranges between 1.5 mg C m$^{-2}$ day$^{-1}$ and 150 mg C m$^{-2}$ day$^{-1}$ with an average of 19.5 $\pm$ 23.7 mg C m$^{-2}$ day$^{-1}$ and a median of 11.4 mg C m$^{-2}$ day$^{-1}$, slightly decreasing from June to December.
-At 1000 m, $F_{big}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to August, followed by a decrease towards December, with values ranging from 0 to 47.0 mg C m$^{-2}$ day$^{-1}$ with an average of 12.0 $\pm$ 11.1 mg C m$^{-2}$ day$^{-1}$ and a median of 8.2 mg C m$^{-2}$ day$^{-1}$.
-
---> ici mettre les données en variables car ça risque de changer souvent... les valeurs de flux etc and test.
-
-```{r, smoothed plots}
-
-#cflux |> ggplot(aes(x = date, y = small_flux, colour = park_depth, group = park_depth)) + geom_smooth() + theme_bw()
-
-```
+At 500 m, $F_{big}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to January.
+At 1000 m, $F_{big}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-JulyAugust, followed by a decrease towards January, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$.
 
@@
 
-In this study, we focus on 12 size classes, from 102 $\mu$m to 1.63 mm.
+In this study, we focus on 14 size classes, from 102 $\mu$m to 2.5 mm.
 In @fig-uvp-1000m, we show the daily averaged concentrations of particles in each class size for each float at 1000 m.
@@
 
-```{r, uvp data at 1000 m}
-#| label: fig-uvp-1000m
-#| fig-cap: Daily averaged concentration of particles abundance in 12 size classes measured by the UVP6 for each float at 1000 m. Shaded areas depict a 95% confidence interval around the mean.
-# lpm_class <- c("NP_Size_102", "NP_Size_128", "NP_Size_161",
-#                          "NP_Size_203", "NP_Size_256", "NP_Size_323",
-#                          "NP_Size_406", "NP_Size_512", "NP_Size_645",
-#                          "NP_Size_813", "NP_Size_1020", "NP_Size_1290")
-
-# lpm_class <- c("NP_Size_102", "NP_Size_128", "NP_Size_161",
-#                          "NP_Size_203", "NP_Size_256", "NP_Size_323",
-#                          "NP_Size_406", "NP_Size_512", "NP_Size_645",
-#                          "NP_Size_813", "NP_Size_1020", "NP_Size_1290",
-#                         "NP_Size_1630", "NP_Size_2050")
-
-lpm_class <- c("NP_Size_64","NP_Size_80.6", "NP_Size_102", "NP_Size_128", "NP_Size_161",
+```{r, uvp data}
+
+lpm_class <- c("NP_Size_102", "NP_Size_128", "NP_Size_161",
                          "NP_Size_203", "NP_Size_256", "NP_Size_323",
@@
 
-info2 <- uvp_data |> dplyr::group_by(week, park_depth, wmo, size_class) |> summarize(mean_conc = mean(conc))
-
 # add level for order
@@
 
-p1000m <- uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = conc, colour = WMO)) + stat_smooth() + 
+# compute daily mean chla for each float at each cycle and at each drifting depth
+mean_uvp_data <- uvp_data |> group_by(cycle, WMO, size_class, park_depth, juld) |> summarize(mean_conc = mean(conc, na.rm=T))
+```
+
+```{r, uvp data at 1000 m}
+#| label: fig-uvp-1000m
+#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 1000 m.
+Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
+
+p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
   scale_color_brewer(palette = 'Dark2') + 
-  geom_vline(xintercept = as_date(2022-10-01), linetype = 'dashed') +
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
   theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
   facet_wrap(~size_class, scales = 'free_y')
@@
 p1000m
-
-tmp1000m <- ggplot_build(p1000m)
-tmp1000m <- tmp1000m$data[[1]]
-
-### maybe we can just show the real values here, let's see with a mean first then add some uncertainties
-# info3 <- uvp_data |> dplyr::group_by(juld, park_depth, wmo, size_class) |> summarize(mean_conc = mean(conc))
-# 
-# info3$size_class <- factor(info3$size_class, levels = lpm_class)
-# info3$park_depth <- factor(info3$park_depth, levels = c('200 m', '500 m', '1000 m'))
-# 
-# info3 |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = mean_conc, colour = as.factor(wmo))) + geom_smooth() +
-#   scale_color_brewer(palette = 'Dark2') + 
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
-#   facet_wrap(~size_class, scales = 'free_y') 
-# 
-# info3 |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = mean_conc, colour = as.factor(wmo))) + geom_path() +
-#   scale_color_brewer(palette = 'Dark2') + 
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
-#   facet_wrap(~size_class, scales = 'free_y')
-
 ```
@@
 #| label: fig-uvp-200
-#| fig-cap: Daily averaged concentration of particles abundance in 12 size classes measured by the UVP6 for each float at 200 m. Shaded areas depict a 95% confidence interval around the mean.
-
-# ggplot(uvp_data, aes(x = juld, y = conc, colour = park_depth)) + geom_smooth() + 
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-#   scale_color_brewer(palette = 'Dark2') + theme_bw() + 
-#   facet_wrap(~size_class, scales = 'free_y')  #+ geom_point() + ylim(c(0,50))
-
-p200m <- uvp_data |> filter(park_depth == '200 m') |> ggplot(aes(x = juld, y = conc, colour = WMO)) + stat_smooth() + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-  scale_color_brewer(palette = 'Dark2') + theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 200 m.
+Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
+
+p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
   facet_wrap(~size_class, scales = 'free_y')
@@
 p200m
-
-tmp200m <- ggplot_build(p200m)
-tmp200m <- tmp200m$data[[1]]
-
-# uvp_data |> filter(park_depth == '500 m') |> ggplot(aes(x = week, y = conc, colour = factor(wmo))) + geom_smooth() + 
-#   #scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-#   scale_color_brewer(palette = 'Dark2') + theme_bw() + 
-#   facet_wrap(~size_class, scales = 'free_y')
-
-# uvp_data |> filter(park_depth == '500 m') |> ggplot(aes(x = juld, y = conc, colour = WMO)) + geom_smooth() + 
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-#   scale_color_brewer(palette = 'Dark2') + theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
-#   facet_wrap(~size_class, scales = 'free_y')
-
-# uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = week, y = conc, colour = factor(wmo))) + geom_smooth() + 
-#   #scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-#   scale_color_brewer(palette = 'Dark2') + theme_bw() + 
-#   facet_wrap(~size_class, scales = 'free_y')
-
-# uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = DOY, y = conc, colour = factor(wmo))) + geom_smooth() + 
-#   scale_color_brewer(palette = 'Dark2') + 
-#   theme_bw() + 
-#   facet_wrap(~size_class, scales = 'free_y')
-
-# uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = conc, colour = factor(wmo))) + geom_smooth() + 
-#   scale_color_brewer(palette = 'Dark2') + 
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   theme_bw() + 
-#   facet_wrap(~size_class, scales = 'free_y')
-
-# uvp_data |> filter(park_depth == '1000 m') |> ggplot(aes(x = juld, y = conc, colour = factor(wmo))) + geom_smooth(size = 1.5) + 
-#   #scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-#   scale_color_brewer(palette = 'Dark2') + theme_bw() + 
-#   facet_wrap(~size_class, scales = 'free_y')
-
-# ggplot(info2, aes(x = juld, y = mean_conc, colour = park_depth, shape = factor(wmo))) + geom_point() + 
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-#   facet_wrap(~size_class, scales = 'free_y')
-
 ```
@@
 #| label: fig-uvp-500
-#| fig-cap: Daily averaged concentration of particles abundance in 12 size classes measured by the UVP6 for each float at 500 m. Shaded areas depict a 95% confidence interval around the mean.
-p500m <- uvp_data |> filter(park_depth == '500 m') |> ggplot(aes(x = juld, y = conc, colour = WMO)) + stat_smooth() + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-  scale_color_brewer(palette = 'Dark2') + theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 500 m.
+Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
+
+p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
   facet_wrap(~size_class, scales = 'free_y')
@@
 p500m
-
-tmp500m <- ggplot_build(p500m)
-tmp500m <- tmp500m$data[[1]]
-
 ```
@@
 
-For the first time, BGC-Argo floats equipped with both a transmissometer and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach \[@Claustre2021-tx\].
-The quantification of the POC flux using the OST method \[@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr\] is based on an empirical POC:ATN ratio \[REF estapa 2022\] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:
+For the first time, BGC-Argo floats equipped with both a transmissometer and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx].
+The quantification of the POC flux using the OST method [@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr] is based on an empirical POC:ATN ratio \[REF estapa 2022\] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:
 
@@
 ```
+
 ```{r spectral slope + rati min/max flux}
@@
 vertical profiles) and let's plot a time serie to compare.
-Starting at 200 m car le modele devient n'a pas appris entre 0-200 et dans cette tranche, les concentrations de particules peuvent être huge huge -> note que cela pourrait dire qu'il faudrait faire des mesures (autre instrument?
+Starting at 200 m car le modele devient n'a pas appris entre 0-200 et dans cette tranche, les concentrations de particules peuvent être huge huge -\> note que cela pourrait dire qu'il faudrait faire des mesures (autre instrument?
 pour estimer un flux..)
@@
 
-Modèle appris en discontinu et prédictions en continu --> BOF BOF et ça se voit.
+Modèle appris en discontinu et prédictions en continu --\> BOF BOF et ça se voit.
 L'autre modèle a une bien plus belle gueule et ne prend pas la profondeur en compte ce qui est encore mieux selon moi, le small flux est expliqué principalement par 102, 128 et 161, ce qui est cohérent.
-On voit potentiellement un pulse d'export --> faudrait le vérifier avec le bbp/cp?
-On voit que vers la fin de l'année, le flux en profondeur est plus faible -> cohérent vu que le bloom date de bien avant.
+On voit potentiellement un pulse d'export --\> faudrait le vérifier avec le bbp/cp?
+On voit que vers la fin de l'année, le flux en profondeur est plus faible -\> cohérent vu que le bloom date de bien avant.
 On aurait besoin d'une time series plus longue, ici j'ai bloqué à 6 mois (car le drift à 200 m va etre mis en pause avec l'hiver et la MLD + le fait que l'on perd un flotteur avec un transmissio qui merde).
@@
 
+PARLER DU TRADEOFF BETWEEN SMALL FLUX AND BIG PARTICLE FLUX
+
+The latter are however subject to a tradeoff between the likely underestimation of $F_{small}$ and the detection of false jumps (without great impact on the value of $F_{big}$ because false jumps have low $c_{p}$ values) if the threshold is too small and the likely overestimation of $F_{small}$ to the detriment of $F_{big}$ with missed jumps if the threshold is too high.
+
 ## Conclusion

### 2023-01-24 article3.qmd
update

@@
 As a result, we firstly computed the slope between each point, then we computed the z-score (i.e.
-standard score) of each slope assuming a normal distribution of the latter.
+standard score) of each slope, that follows a normal distribution.
 The detection of jumps thus consisted in spotting outliers in the z-score distribution.
@@
 
-We used the same relationship to compute $F_{big}$ was replacing $slope_{w}$ as the sum of positive jumps (negative jumps were disgarded) divided by the drifting time.
+We used the same relationship to compute $F_{big}$ by replacing $slope_{w}$ as the sum of positive jumps (negative jumps were discarded) divided by the drifting time.
 
@@
 At 500 m, $F_{big}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to January.
-At 1000 m, $F_{big}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-JulyAugust, followed by a decrease towards January, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$.
+At 1000 m, $F_{big}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards January, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$.
 

### 2023-01-30 article3.qmd
update article 3

@@
 library(castr)
+library(reactablefmtr)
 ```
@@
 Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach instead of fixed thresholds to detect jumps.
-As a result, we firstly computed the slope between each point, then we computed the z-score (i.e.
+Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase.
+Then, we computed the z-score (i.e.
 standard score) of each slope, that follows a normal distribution.
@@
 
-In order to compute $F_{small}$, we firstly computed the slope of each group of points between jumps (positive and negative, see @fig-estapa-method_article).
+In order to compute $F_{small}$, we firstly computed the slope of each group of observations between jumps (positive and negative, see @fig-estapa-method_article).
 Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window.
@@
 F_{small} = 633(0.25~slope_{w})^{0.77}
-$$
-
-where $slope_{w}$ is multiplied by the transmissometer pathlength because @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.
-
-We used the same relationship to compute $F_{big}$ by replacing $slope_{w}$ as the sum of positive jumps (negative jumps were discarded) divided by the drifting time.
+$$ {#eq-estapa}
+
+where $slope_{w}$ is multiplied by the transmissometer pathlength due to the fact that @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.
+
+We used the same relationship to compute $F_{big}$ by replacing $slope_{w}$ as the sum of positive jumps divided by the drifting time.
 
@@
 At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$.
+For both drifting depths, the trend shows a decrease of $F_{small}$ from June to January.
 At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid-July followed by a steady decrease towards January, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.
@@
 
-In this study, we focus on 14 size classes, from 102 $\mu$m to 2.5 mm.
-In @fig-uvp-1000m, we show the daily averaged concentrations of particles in each class size for each float at 1000 m.
-For the first three classes (i.e.
-102-203 $\mu$m), we observe a similar pattern with concentrations starting at 4 particles/L at the end of May that reached 8 particles/L in mid-July and then decreased back to 4 particles/L towards December.
-For particles in the 203-406 $\mu$m range, the pattern can still be observed with a delay of 15 to 30 days between floats 6904240/4903634 and floats 6904241/4902578 which were deployed conjointly.
-Observed concentrations were very low at the end of May (\< 1 particle/L) then they increased by a factor of \~3-4 between mid-June and mid-July to decrease back to their initial concentrations towards December.
-For the remaining classes, concentrations are always below 1 particle/L with increasing uncertainty (rarer particles) in the signal.
-At 200 m (@fig-uvp-200), there is no clear peak following the bloom event in the smaller part of the size spectrum (\< 406 $\mu$m) whereas both floats 6904240 and 4903634 (i.e.
-conjointly deployed) observed the same high concentrations of particles above 406 $\mu$m at the end of May.
-At 500 m (@fig-uvp-500), the signal shows a slight increase in the abundance of particles in the 102-256 $\mu$m range from the end of May to the beginning of July or mid-July, depending of the float.
+#### When drifting
 
@@
 
-uvp_data <- uvp_data |> pivot_longer(cols = lpm_class, names_to = 'size_class', values_to = 'conc')
+uvp_data <- uvp_data |> pivot_longer(cols = lpm_class, names_to = 'size', values_to = 'conc') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
 
 # add level for order
-uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
+#uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
 uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))
@@
 # compute daily mean chla for each float at each cycle and at each drifting depth
-mean_uvp_data <- uvp_data |> group_by(cycle, WMO, size_class, park_depth, juld) |> summarize(mean_conc = mean(conc, na.rm=T))
-```
+mean_uvp_data <- uvp_data |> group_by(cycle, WMO, size, park_depth, juld) |> summarize(mean_conc = mean(conc, na.rm=T))
+```
+
+In this study, we focus on 14 size classes, from 102 $\mu$m to 2.5 mm.
+In @fig-uvp-200m, @fig-uvp-500m and @fig-uvp-1000m, we show the daily averaged concentrations of particles in each class size for each float at their respective drifting depth.
+At 200 m, it can be observed that particles concentrations in the 102 - 512 $\mu$m range are relatively uniform between June and October before decreasing sharply towards January.
+Between 512 $\mu$m and 1.29 mm, concentrations are steadily decreasing from June.
+Above 1.29 mm, the abundance of particles is already very low (< 0.1 particle/L) in June and keeps decreasing over time.
+
+```{r, uvp data at 200 m}
+#| label: fig-uvp-200m
+#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 200 m.
+Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
+Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.
+
+p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') #+
+  #geom_vline(data = test, aes(xintercept = date_max))
+
+tmp200m <- ggplot_build(p200m)
+
+test <- tmp200m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> summarize(index_max = which.max(y), max_date = x[index_max]) |> mutate(date_max = as_date(max_date, origin = "1970-01-01"))
+test$size <- unique(uvp_data$size)
+
+p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') +
+  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
+
+p200m
+```
+
+At 500 m, particles concentrations in the 102 - 645 $\mu$ range slowly decrease from July to October with a sharp decrease in November.
+Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (< 0.5 particle/L) on the entire timeserie.
+
+```{r, uvp data at 500 m}
+#| label: fig-uvp-500m
+#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 500 m.
+Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
+Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.
+
+p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') #+
+  #geom_vline(data = test, aes(xintercept = date_max))
+
+tmp500m <- ggplot_build(p500m)
+
+test <- tmp500m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> summarize(index_max = which.max(y), max_date = x[index_max]) |> mutate(date_max = as_date(max_date, origin = "1970-01-01"))
+test$size <- unique(uvp_data$size)
+
+p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') +
+  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
+
+p500m
+```
+
+At 1000 m, all 4 floats measured a clear maximum in particles abundance in the 102 - 645 $\mu$m in the second half of July.
+Above 645 $\mu$m, the same trend can still be observed with concentrations of less than 1 particle/L.
 
@@
 Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
+Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.
 
@@
   scale_color_brewer(palette = 'Dark2') + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
   theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
   scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size_class, scales = 'free_y')
-
-p1000m
-```
-
-```{r, uvp data at 200 m}
-#| label: fig-uvp-200
-#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 200 m.
-Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
-
-p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  facet_wrap(~size, scales = 'free_y') #+
+  #geom_vline(data = test, aes(xintercept = date_max))
+
+tmp1000m <- ggplot_build(p1000m)
+
+test <- tmp1000m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> summarize(index_max = which.max(y), max_date = x[index_max]) |> mutate(date_max = as_date(max_date, origin = "1970-01-01"))
+test$size <- unique(uvp_data$size)
+
+p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
   geom_smooth(aes(juld, mean_conc), colour = 'black') +
   scale_color_brewer(palette = 'Dark2') + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
   theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
   scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size_class, scales = 'free_y')
-
-p200m
-```
-
-```{r, uvp data at 500 m}
-#| label: fig-uvp-500
-#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 500 m.
-Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
-
-p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
-  geom_smooth(aes(juld, mean_conc), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
-  scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size_class, scales = 'free_y')
-
-p500m
+  facet_wrap(~size, scales = 'free_y') +
+  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
+
+p1000m
+```
+
+#### When profiling
+
+@fig-uvp-vertical-profiles shows that all floats measure similar patterns in the vertical profiles of particles concentrations in all size classes.
+In the 102-512 $\mu$m range, all 4 floats measured a local maximum between 500 and 1500 m, likely due to both sinking and remineralization processes.
+Above 512 $\mu$m, the relative abundance is maximum along a depth track centered on August.
+It can be seen among all size classes that the water column is being drained proportionally to the class size.
+
+```{r, uvp_data on the vertical}
+#| label: fig-uvp-vertical-profiles
+#| fig-cap: Normalized concentrations of particles for all UVP6 vertical profiles.
+Plots were arranged to respect the deployment of each couple of floats (e.g.
+the first two rows of plots are related to co-deployed floats 6904241 and 1902578).
+The colorbar depicts the log$_{10}$ of normalized concentrations.
+Grey areas indicate that particles were not measured in the size range of interest.
+
+particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))
+
+particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+#particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
+
+# rel_conc
+particles_bis <- particles_bis |> group_by(wmo, size) |> mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
+
+# order floats by couple of deployment
+particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904241,1902578,6904240,4903634))
+
+# plot concentration
+# particles_bis |> filter(depth < 200) |>
+#   ggplot() + facet_grid(wmo~size) + theme_bw() +
+#   geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
+#   #scale_colour_viridis(option = 'H', limits = c(-1,1), oob = scales::squish, na.value = 'transparent') +
+#   scale_colour_viridis(option = 'H') +
+#   scale_y_reverse() +
+#   scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+#   labs(x = 'Month', y = 'Depth (m)') +
+#   geom_path(data = particles_bis, aes(juld, MLD), colour = "white")
+
+## plot absolute concentrations
+particles_bis |> filter(depth < 2000) |>
+  #group_by(size) |>
+  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
+  scale_colour_viridis(option = 'H', limits = c(-3,-1), oob = scales::squish, na.value = 'grey') +
+  #scale_colour_viridis(option = 'H') +
+  scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = 'Normalized concentrations (#/L)', x = 'Month', y = 'Depth (m)') +
+  #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
+  theme(legend.position = 'bottom') #+
+  #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
+  #guides(colour = guide_legend(title.position = "bottom"))
+
+# # interpolate to see better
+# dli <- particles_bis |> mutate(juldi = as.numeric(juld)) |>
+#   group_by(wmo, size) |>
+#   do({
+#     i <- akima::interp(x=.$juldi, y=.$depth, z=.$concentration, nx=50, ny=150, duplicate="mean") |>
+#       tidy()
+#     # ggplot(i) + geom_raster(aes(x=x, y=-y, fill=z))
+#     rename(i, juldi=x, pres=y, concentration=z)
+#   }) |>
+#   # re-compute relative concentration to see stuff better
+#   group_by(size) |>
+#   mutate(rel_conc=concentration/max(concentration, na.rm=TRUE)) |>
+#   mutate(juld = as_date(juldi, origin = "1970-01-01")) |>
+#   drop_na(rel_conc) |>
+#   ungroup()
+# 
+# dli |> 
+#   ggplot() + facet_grid(wmo~size) +
+#   geom_raster(aes(x=juldi, y=-pres, fill=rel_conc)) +
+#   #scale_y_reverse() |>
+#   scale_x_continuous(expand=c(0,0), limits=c(NA, NA)) +
+#   scale_y_continuous(expand=c(0,0), limits=c(-300,0)) +
+#   scale_fill_viridis_c(option = 'H')
+
+# test_sigma <- all_floats |> filter(PhaseName == 'NPAR', sigma > 26) |> select(juld, cycle, depth = pres, wmo, sigma, MLD) |> drop_na(sigma) |> mutate(juld = as_date(juld))
+# test_sigma |> ggplot() + geom_point(aes(x = juld, y = depth, colour = sigma)) + 
+#   scale_colour_viridis(option = 'H', oob = scales::squish, na.value = 'transparent') +
+#   scale_y_reverse(limits = c(200,0)) +
+#   scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+#   labs(x = 'Month', y = 'Depth (m)') +
+#   facet_wrap(~wmo, nrow = 4) + theme_bw() +
+#   geom_path(data = test_sigma, aes(juld, MLD), colour = "grey")
+
+```
+
+## Discussion
+
+### Remineralization processes
+
+```{r, monthly anomaly between drifting at 500 and 1000 m}
+
+tmp <- uvp_data |> filter(park_depth != '200 m', month > 5) |> group_by(wmo, size, month, park_depth) |> 
+  summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
+  # compute concentration anomaly at 500 and 1000 m
+  mutate(anomaly = monthly_mean_conc - lead(monthly_mean_conc)) |> drop_na(anomaly) |>
+  select(wmo, size, month, anomaly) |> mutate(month = month.abb[month]) |> 
+  mutate(size = as.character(size)) |>
+  #mutate(anomaly = round(anomaly, 3))
+  ungroup()
+
+tt <- tmp |> filter(wmo == 6904241) |> select(size, month, anom = anomaly)
+#tt <- tt |> pivot_wider(names_from = month, values_from = anom) |> mutate(size = as.character(size))
+
+x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')
+y <- unique(tt$size)
+data <- expand.grid(X=x, Y=y)
+data$Z <- tt$anom
+
+ggplot(data, aes(X, Y, fill= Z)) + 
+  geom_tile() + scale_fill_gradientn(colors = c('darkblue', 'blue', 'white', 'red', 'darkred'))
+
+# # https://kcuilla.github.io/reactablefmtr/articles/reactablefmtr_cookbook.html
+# tt <- tt |> pivot_wider(names_from = month, values_from = anom) |> mutate(size = as.character(size))
+# reactable(
+#   tt,
+#   pagination = FALSE,
+#   showSortIcon = FALSE,
+#   theme = void(
+#     centered = TRUE, 
+#     cell_padding = 0,
+#     header_font_color = 'black',
+#     font_color = 'black'
+#     ),
+#   defaultColDef = colDef(
+#     maxWidth = 50,
+#     align = 'center',
+#     cell = tooltip(),
+#     style = color_scales(
+#       data = tt,
+#       span = 2:8,
+#       colors = c('#002347','#003366','#003F7D','#FF8E00','#FD7702','#FF5003'),
+#       bias = 1.4,
+#       opacity = 0.9,
+#       show_text = FALSE
+#     )
+#   ),
+#   columns = list(
+#     total = colDef(
+#       maxWidth = 225,
+#       cell = data_bars(
+#         data = tt,
+#         fill_color = c('#002347','#003366','#003F7D','#FF8E00','#FD7702','#FF5003'),
+#         bias = 1.4,
+#         fill_opacity = 0.9,
+#         background = 'transparent',
+#         bar_height = 40,
+#         text_position = 'center'
+#         ),
+#       style = list(borderLeft = "2px solid #999999")
+#     )
+#   )
+# )
+
 ```
@@
 ggplot(npq_argo, aes(x = DOY, y = mean_chla, colour = factor(wmo))) + geom_point() + geom_path() + facet_wrap(~wmo)
-
-# # ggplot(sat_data_bbox, aes(DOY, mean_chla)) + geom_point() + geom_path()
-# ggplot(sat_data_bbox, aes(DOY, q50)) + geom_point() + geom_path()
-
-# # check float position before july
-# float_position <- filter(all_floats, juld < '2022-06-15') |> drop_na(lat)
-# 
-# # get lat/lon range for each float for that period
-# float_position |> dplyr::group_by(wmo) |> dplyr::summarise(min_lat = min(lat), max_lat = max(lat),
-#                                                            min_lon = min(lon), max_lon = max(lon))
-# 
-# chla_sat_couple1 <- filter(chla_sat, lat > 56.5, lat < 56.85, lon > -53, lon < -52.4) |> drop_na(CHL)
-# 
-# tmp <- chla_sat_couple1 |> dplyr::group_by(time) |> summarise(mean_chla = mean(CHL))
-# 
-# ggplot(tmp, aes(x = time, y = mean_chla)) + geom_point() + geom_path()
-# 
-# # at deployment
-# chla_sat_couple1 <- filter(chla_sat, lat > 56.7, lat < 56.9, lon > -52.5, lon < -52.3) |> drop_na(CHL)
-# tmp <- chla_sat_couple1 |> dplyr::group_by(time) |> summarise(mean_chla = mean(CHL))
-# ggplot(tmp, aes(x = time, y = mean_chla)) + geom_point() + geom_path()
-# 
-# # couple 2
-# chla_sat_couple2 <- filter(chla_sat, lat > 58.9, lat < 59.1, lon > -49.4, lon < -49.2) |> drop_na(CHL)
-# tmp <- chla_sat_couple2 |> dplyr::group_by(time) |> summarise(mean_chla = mean(CHL))
-# ggplot(tmp, aes(x = time, y = mean_chla)) + geom_point() + geom_path()
-# 
-# ggplot(argo_map) + w + geom_path(aes(lon, lat, colour = WMO), size = 1) +
-#   geom_point(aes(lon, lat, colour = WMO), size = 1.5) +
-#   scale_color_brewer(palette = 'Dark2') +
-#   theme_bw() +
-#   #scale_colour_manual(values = c('#a6cee3', '#33a02c', '#b2df8a', '#1f78b4')) +
-#   coord_cartesian(xlim = c(-55, -40), ylim = c(55,65)) + geom_point(data = argo_map_first, aes(lon, lat), shape = 17, size = 3) +
-#   geom_point(data = argo_map_last, aes(lon, lat, colour = WMO), size = 3) +
-#   geom_point(data = argo_map_last, aes(lon, lat, fill = WMO), size = 3, shape = 1) +
-#   labs(x = 'longitude', y = 'latitude') +
-#   #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
-#   guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
-#                                                   shape = c(NA, NA, NA, NA)))) +
-#   geom_point(data = chla_sat_couple1, aes(x = lon, y = lat), colour = "red") +
-#   geom_point(data = chla_sat_couple2, aes(x = lon, y = lat), colour = "red")
-#
-
-```
-
-```{r}
-
-# all_floats_chla <- filter(all_floats, PhaseName == 'NPAR', pres <= 50, pres > 2) |> drop_na(chla) |> select(juld, pres, chla, wmo) |> mutate(datetime = as_date(juld)) #|> filter(datetime < '2022-08-01')
-# 
-# chla_summary <- all_floats_chla |> dplyr::group_by(wmo, datetime) |> summarize(chla_mean = mean(chla, na.rm=T))
-# 
-# ggplot(chla_summary, aes(x = datetime, y = chla_mean)) + geom_point() + geom_path() + facet_wrap(~wmo)
 
@@
 particles_bis$size_class <- factor(particles_bis$size_class, levels = lpm_class)
+
+ggplot(particles_bis |> filter(wmo == '1902578'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
+  scale_y_reverse(limits = c(2000,0)) +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  labs(x = 'Time', y = 'abundance', title = 'wmo 1902578') +
+  facet_wrap(~size_class) + theme_bw()
 
@@
 
-## Discussion
+## Discussion2
 

### 2023-01-31 article3.qmd
update

@@
 library(reactablefmtr)
+library(gt)
 ```
@@
 
-#### When drifting
+#### On drifting mode
 
@@
 
-#### When profiling
+#### On profiling mode
 
@@
 Plots were arranged to respect the deployment of each couple of floats (e.g.
-the first two rows of plots are related to co-deployed floats 6904241 and 1902578).
+the first two rows of plots are related to co-deployed floats 6904240 and 4903634).
 The colorbar depicts the log$_{10}$ of normalized concentrations.
@@
 # order floats by couple of deployment
-particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904241,1902578,6904240,4903634))
+particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904240,4903634,6904241,1902578))
 
@@
 ## plot absolute concentrations
-particles_bis |> filter(depth < 2000) |>
+up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
   #group_by(size) |>
@@
   #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
-  scale_colour_viridis(option = 'H', limits = c(-3,-1), oob = scales::squish, na.value = 'grey') +
+  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
   #scale_colour_viridis(option = 'H') +
@@
   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-  labs(colour = 'Normalized concentrations (#/L)', x = 'Month', y = 'Depth (m)') +
+  #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
+  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
+  #theme(legend.title = element_text(angle = -90))
+  #guides(fill = guide_legend(title.theme = element_text(size = 200)))
+  #+ theme(legend.position = 'none')
   #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
-  theme(legend.position = 'bottom') #+
+  #theme(legend.position = 'none') #+
   #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
   #guides(colour = guide_legend(title.position = "bottom"))
+
+#Normalized concentrations (#/L)
+
+up_part
+
+down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
+  #group_by(size) |>
+  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
+  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) +
+  #scale_colour_viridis(option = 'H') +
+  scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
+  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
+
+down_part
 
@@
 
-### Remineralization processes
+### Processes at play
 
 ```{r, monthly anomaly between drifting at 500 and 1000 m}
-
-tmp <- uvp_data |> filter(park_depth != '200 m', month > 5) |> group_by(wmo, size, month, park_depth) |> 
+#| label: fig-processes-at-play
+#| fig-cap: Ratio of monthly mean particles concentrations during drifting (1000 m over 500 m) for size classes below 1.63 mm.
+
+tmp <- uvp_data |> filter(park_depth != '200 m') |> group_by(wmo, size, month, park_depth) |> 
   summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
   # compute concentration anomaly at 500 and 1000 m
-  mutate(anomaly = monthly_mean_conc - lead(monthly_mean_conc)) |> drop_na(anomaly) |>
-  select(wmo, size, month, anomaly) |> mutate(month = month.abb[month]) |> 
+  mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
+  select(wmo, size, month, diff_anomaly, div_anomaly) |> mutate(month = month.abb[month]) |> 
   mutate(size = as.character(size)) |>
-  #mutate(anomaly = round(anomaly, 3))
+  mutate(rev_div_anomaly = 1/div_anomaly) |> 
+  #filter(is.finite(rev_div_anomaly)) |>
+  mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1), div_anomaly = round(div_anomaly, 1)) |>
   ungroup()
 
-tt <- tmp |> filter(wmo == 6904241) |> select(size, month, anom = anomaly)
-#tt <- tt |> pivot_wider(names_from = month, values_from = anom) |> mutate(size = as.character(size))
-
-x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')
-y <- unique(tt$size)
+# test en weekly
+# tmp <- uvp_data |> filter(park_depth != '200 m', month != 5) |> group_by(wmo, size, week, park_depth) |> 
+#   summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
+#   # compute concentration anomaly at 500 and 1000 m
+#   mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
+#   select(wmo, size, week, diff_anomaly, div_anomaly) |> 
+#   mutate(size = as.character(size)) |>
+#   mutate(rev_div_anomaly = 1/div_anomaly) |> 
+#   #filter(is.finite(rev_div_anomaly)) |>
+#   mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1)) |>
+#   ungroup()
+# 
+# create_data_for_heatmap <- function(float){
+#   y <- unique(tmp$size)[1:12] # above, data is way too low
+#   d <- tmp |> filter(wmo == float, size %in% y, week > 5)
+#   x <- unique(d$week)
+#   data <- expand.grid(X=x, Y=y)
+#   #data$Z <- tt$diff_anomaly
+#   data$Z <- d$rev_div_anomaly
+#   data$wmo <- float
+#   return(data)
+# }
+
+# d1 <- create_data_for_heatmap(6904241)
+# d2 <- create_data_for_heatmap(1902578)
+# d3 <- create_data_for_heatmap(4903634)
+# d4 <- create_data_for_heatmap(6904240)
+# 
+# dd <- rbind(d1, d2, d3, d4)
+
+# ggplot(dd, aes(X, Y, fill= Z)) + 
+#   geom_tile() + scale_fill_viridis() + facet_wrap(~wmo)
+
+create_data_for_heatmap <- function(float, df){
+  y <- unique(df$size)[1:12] # above, data is way too low
+  d <- df |> filter(wmo == float, size %in% y)
+  #x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')
+  x <- unique(d$month)
   data <- expand.grid(X=x, Y=y)
-data$Z <- tt$anom
-
-ggplot(data, aes(X, Y, fill= Z)) + 
-  geom_tile() + scale_fill_gradientn(colors = c('darkblue', 'blue', 'white', 'red', 'darkred'))
+  #data$Z <- tt$diff_anomaly
+  data$Z <- d$rev_div_anomaly
+  #data$Z <- d$div_anomaly
+  data$wmo <- float
+  return(data)
+}
+
+# for some reason, there is some bug with the quarto compilation .. so I save the tibble .. and reload it 
+# d1 <- create_data_for_heatmap(6904241, tmp)
+# d2 <- create_data_for_heatmap(1902578, tmp)
+# d3 <- create_data_for_heatmap(4903634, tmp)
+# d4 <- create_data_for_heatmap(6904240, tmp)
+# dd <- rbind(d1, d2, d3, d4)
+# vroom_write(dd, 'heatmap_data.csv')
+
+# reload + redo some annoying stuff that were lost during the writing of the csv
+dd <- vroom('heatmap_data.csv') |> mutate(Y = as.character(Y)) 
+dd$X <- factor(dd$X, levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))
+dd$Y <- factor(dd$Y, levels = c('102', '128', '161', '203', '256', '323', '406', '512', '645', '813', '1020', '1290'))
+dd$Y <- factor(dd$Y, levels = c('1290', '1020', '813', '645', '512', '406', '323', '256', '203', '161', '128', '102'))
+dd$wmo <- factor(dd$wmo, levels = c(6904240,4903634,6904241,1902578))
+
+#dd <- dd |> mutate(Z_binned = cut(Z, breaks = c(0,1,1.5,2.5,max(Z))))
+dd <- dd |> mutate(Z_binned = cut(Z, breaks = c(0,.5,1,1.5,2,max(Z))))
+
+# ggplot(dd) +
+#   geom_tile(aes(X, Y, fill = Z), alpha = 0.75) + scale_fill_viridis() + theme_bw() + labs(x = 'Month', y = 'Size class') +
+#   facet_wrap(~wmo) +
+#   geom_text(data = dd, aes(X, Y, label = Z))
+
+ggplot(dd) + 
+  geom_tile(aes(X, Y, fill = Z_binned), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Ratio', x = 'Month', y = 'Size class') +
+  facet_wrap(~wmo) + scale_fill_viridis_d(option = 'A')
+
+# plot_heatmap <- function(float){
+#   y <- unique(tmp$size)[1:12]
+#   tt <- tmp |> filter(wmo == float, size %in% y) 
+#   x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')
+#   data <- expand.grid(X=x, Y=y)
+#   #data$Z <- tt$anom
+#   data$Z <- tt$diff_anomaly
+#   data$Z <- tt$rev_div_anomaly
+# 
+#   p <- ggplot(data, aes(X, Y, fill= Z)) + 
+#     geom_tile() + scale_fill_viridis() +
+#     geom_text(data = data, aes(X, Y, label = Z)) +
+#     labs(title = float)
+#   
+#   return(p)
+# }
+# 
+# p1 <- plot_heatmap(4903634)
+# p2 <- plot_heatmap(6904240)
+# p3 <- plot_heatmap(1902578)
+# p4 <- plot_heatmap(6904241)
+# 
+# (p1 | p2) / 
+# (p3 | p4)
 
@@
 
+```{r, test gt table}
+
+monthly_mean <- uvp_data |> mutate(month = month.abb[month(juld)]) |> group_by(size, park_depth, month) |> summarize(mean_conc = round(mean(conc, na.rm=T), 2)) |> ungroup()
+
+# pivot wider
+monthly_mean <- monthly_mean |> pivot_wider(names_from = month, values_from = mean_conc) |> select(size, depth = park_depth, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec, Jan)
+
+monthly_mean |> group_by(size) |> gt(rowname_col = 'depth')
+
+#monthly_mean |> gt(rowname_col = c('size')) |> tab_stubhead(label = 'Class')
+
+```
+
 ### Where is the bloom?

### 2023-02-01 article3.qmd
update

@@
 @fig-uvp-vertical-profiles shows that all floats measure similar patterns in the vertical profiles of particles concentrations in all size classes.
-In the 102-512 $\mu$m range, all 4 floats measured a local maximum between 500 and 1500 m, likely due to both sinking and remineralization processes.
+In the 102-512 $\mu$m range, all 4 floats measured a local maximum between 650 m and 1050 m.
 Above 512 $\mu$m, the relative abundance is maximum along a depth track centered on August.
@@
 particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904240,4903634,6904241,1902578))
+
+# stats on the local maximum
+# test <- particles_bis |> filter(depth > 500) |> group_by(wmo, cycle, size) |> # find depth of maximum relative abundance below 500 m between June and September
+#   summarize(index_max = which.max(rel_conc), max_depth = depth[index_max], date = unique(juld)) |> 
+#   filter(size < 512, date < '2022-10-01', date > '2022-05-31') |> # compute mean depth for each class size for each float
+#   group_by(wmo, size) |> summarize(mean_depth = mean(max_depth))
 
@@
   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
+  #scale_colour_viridis(option = 'A', limits = c(-2,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
   #scale_colour_viridis(option = 'H') +
@@
   #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
-  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
+  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') #+
+  #geom_hline(yintercept = c(600, 1100), linetype = 'dotted')
   #theme(legend.title = element_text(angle = -90))
@@
 
-### Where is the bloom?
-
-```{r, sat data}
-
-#chla_sat <- vroom::vroom('DATA/SATELLITE/chla_sat.csv')
-chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_sat.feather') |> mutate(DOY = yday(time))
-sat_data_bbox <- filter(chla_sat, lat >= 55.5, lat <= 60, lon >= -55, lon <= -45) |> group_by(DOY) |> drop_na(CHL)
-
-# sat data
-test <- sat_data_bbox |> group_by(DOY) |> summarize(npixels = n(), nabove5 = length(which(CHL > 5)), nabove2 = length(which(CHL > 2)), r5 = nabove5/npixels)
-ggplot(test, aes(DOY, r5)) + geom_point() + geom_path()
+### Chlorophyll data
+
+```{r, chla data}
 
@@
 npq_data <- vroom('DATA/ARGO/npq_corrected_data.csv')
-npq_filtered <- npq_data |> filter(pres > 0, pres <= 20, juld < '2022-12-31') |> mutate(DOY = yday(juld)) |> drop_na(chla)
-
-npq_argo <- npq_filtered |> group_by(wmo, cycle) |> summarize(DOY = unique(DOY), mean_chla = mean(chla_npq, na.rm=T), median_chla = median(chla_npq, na.rm=T))
-
-ggplot(npq_argo, aes(x = DOY, y = mean_chla, colour = factor(wmo))) + geom_point() + geom_path() + facet_wrap(~wmo)
-
+
+# clean and smoothed chla data
+d <- npq_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> mutate(date = as_date(juld))
+d <- d |> group_by(date, wmo, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
+d <- d |> mutate(smoothed_chla = smooth(mean_chla))
+
+d$wmo <- factor(d$wmo, levels = c(6904240,4903634,6904241,1902578))
+
+# plot it
+ggplot(d) + geom_point(aes(x = date, y = pres, colour = log10(smoothed_chla))) +
+  scale_y_reverse(limits = c(200,0)) +
+  #scale_colour_viridis(option = 'A') +
+  scale_colour_cmocean(name = 'algae', oob = scales::squish, limits = c(-1,1), na.value = 'grey') +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  labs(x = 'Time', y = 'NPQ-corrected Chla') + facet_wrap(~wmo) + theme_bw()
+
+# compute chla integration in the MLD and in the 0-100 m zone
+d_chla <- d |> group_by(wmo, date) |> summarize(chla_0_100m = integrate(smoothed_chla, pres, from=0, to=100),
+                                       mld = unique(mld),
+                                       chla_0_mld = integrate(smoothed_chla, pres, from = 0, to=mld))
+
+ggplot(d_chla) + geom_point(aes(date, chla_0_100m)) + facet_wrap(~wmo, scales = 'free') + theme_bw() + labs(x = 'Month', y = 'Integrated Chla 0-100 m (CHLA UNITS)')
+#ggplot(d_chla) + geom_point(aes(date, chla_0_mld)) + facet_wrap(~wmo, scales = 'free')
+
+```
+### Mixed layer depth
+
+```{r, mld data}
+ggplot(d_chla) + geom_point(aes(date, mld)) + facet_wrap(~wmo, scales = 'free') + theme_bw() + labs(x = 'Month', y = 'Mixed layer depth (m)')
 ```
@@
 Due to those reasons, we have chosen to firstly compare $F_{small}$ and $F_{big}$ with UVP6 particles concentrations.
-For this, we computed the particles density distribution maximum using a smoothing gaussian kernel (see R package ggvis) for each class size at each parking depth and at each cycle.
-Then, we computed the distance correlation between OST-derived carbon fluxes and UVP size classes maxima.
-The distance correlation \[REF\] was preferred to the well known Pearson's correlation due to uncertainties in the linear relationship of, respectively, small classes and $F_{big}$, and big classes and $F_{small}$.
-
-```{r, essai sur ridgeplot}
-library(ggridges)
-tmp <- uvp_data |> filter(park_depth == '1000 m', wmo == 6904240, conc < 20)
-#tmp <- uvp_data |> filter(park_depth == '1000 m', wmo == 6904240, size_class %in% c('NP_Size_102', 'NP_Size_128', 'NP_Size_161', 'NP_Size_203', 'NP_Size_256', 'NP_Size_323'))
-
-# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_density_ridges(scale = 5) + stat_density_ridges(quantile_lines=T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
-#   theme_bw() + scale_color_brewer(palette = 'Dark2')
-# 
-# # ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_density_ridges() + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) 
-# #   
-# # ggplot(tmp, aes(y = factor(cycle), x = conc)) + geom_density_ridges_gradient() + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
-# #    scale_fill_viridis_c(name = "Temp. [F]", option = "C")
-# 
-# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + stat_density_ridges(quantile_lines=T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
-#   theme_bw() 
-# 
-# ggplot(tmp, aes(y = factor(cycle), x = conc)) + stat_density_ridges(quantile_lines=T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
-#   theme_bw() + facet_wrap(~size_class) + xlim(c(0, 10)) 
-# 
-# ggplot(tmp, aes(y = factor(cycle), x = conc)) + stat_density_ridges(quantile_lines=T, quantiles = 2, stat="identity") + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
-#   theme_bw() + facet_wrap(~size_class, scales = "free")
-# 
-ggplot(tmp, aes(y = factor(cycle), x = conc, height = ..density..)) + geom_density_ridges(scale = 5, stat = "density", rel_min_height = 0.01) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
-  theme_bw() + facet_wrap(~size_class, scales = "free")
-# 
-# # test 
-# tmp2 <- uvp_data |> dplyr::group_by(cycle, size_class, park_depth, wmo) |> 
-#   group_modify(~ ggplot2:::compute_density(.x$conc, NULL)) |>
-#   rename(conc = x)
-# 
-# ggplot(tmp2, aes(x = conc, y = cycle, height = density)) + 
-#   geom_density_ridges(stat = "identity") + facet_wrap(~size_class, scales = "free")
-
-# 
-# ggplot(tmp, aes(y = factor(cycle), x = conc, height = ..density..)) + stat_density_ridges(scale = 5) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
-#   theme_bw() + facet_wrap(~size_class, scales = "free")
-# 
-# ggplot(tmp, aes(y = factor(cycle), x = conc, height = ..density..)) + stat_density_ridges(scale = 5, quantile_lines = T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
-#   theme_bw() + facet_wrap(~size_class, scales = "free")
-# 
-# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_ridgeline() + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
-#   theme_bw() + scale_color_brewer(palette = 'Dark2')
-
-# ggplot(tmp, aes(x = factor(cycle), y = conc, fill = size_class)) + geom_boxplot() + labs(y = 'abundance distribution (#/L)', x = 'cycle')  +
-#   theme_bw() + scale_color_brewer(palette = 'Dark2') + ylim(c(0,15)) + geom_jitter()
-# 
-# ggplot(tmp, aes(x = factor(cycle), y = conc, fill = size_class)) + geom_violin() + labs(y = 'abundance distribution (#/L)', x = 'cycle')  +
-#   theme_bw() + scale_color_brewer(palette = 'Dark2') + ylim(c(0,15))
-
-# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_density_ridges(alpha=1, stat="binline", bins=50)  + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
-#   theme_bw() + scale_color_brewer(palette = 'Dark2')
-
-```
-
-```{r, check with particles abundance}
-
-particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))
-
-particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size_class', values_to = 'concentration')
-particles_bis$size_class <- factor(particles_bis$size_class, levels = lpm_class)
-
-ggplot(particles_bis |> filter(wmo == '1902578'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
-  scale_y_reverse(limits = c(2000,0)) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = 'abundance', title = 'wmo 1902578') +
-  facet_wrap(~size_class) + theme_bw()
-
-ggplot(particles_bis, aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
-  scale_y_reverse(limits = c(2000,0)) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = 'abundance') +
-  facet_wrap(~size_class) + theme_bw()
-
-ggplot(particles_bis |> filter(wmo == '6904240'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
-  scale_y_reverse(limits = c(2000,0)) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = 'abundance', title = 'wmo 6904240') +
-  facet_wrap(~size_class) + theme_bw()
-
-# https://stackoverflow.com/questions/48424682/how-do-i-limit-the-range-of-the-viridis-colour-scale
-ggplot(particles_bis |> filter(wmo == '6904240'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_colour_viridis(option = 'H', limits = c(-1,1), oob = scales::squish, na.value = 'transparent') +
-  scale_y_reverse(limits = c(2000,0)) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = 'abundance', title = 'wmo 6904240') +
-  facet_wrap(~size_class) + theme_bw()
-
-ggplot(particles_bis |> filter(wmo == '6904241'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
-  scale_y_reverse(limits = c(2000,0)) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = 'abundance', title = 'wmo 6904241') +
-  facet_wrap(~size_class) + theme_bw()
-
-ggplot(particles_bis |> filter(wmo == '4903634'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
-  scale_y_reverse(limits = c(2000,0)) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = 'abundance', title = 'wmo 4903634') +
-  facet_wrap(~size_class) + theme_bw()
-
-ggplot(particles_bis |> filter(wmo == '1902578'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
-  scale_y_reverse(limits = c(2000,0)) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = 'abundance', title = 'wmo 1902578') +
-  facet_wrap(~size_class) + theme_bw()
-
-# # sum 3 first class
-# particles_tmp <- particles |> dplyr::group_by(wmo, cycle, depth) |> summarize(juld = unique(juld), sum_NP = sum(NP_Size_102, NP_Size_128, NP_Size_161)) |> filter(depth > 100)
-# 
-# ggplot(particles_tmp, aes(x = juld, y = depth, colour = log(sum_NP))) + geom_point() + scale_color_viridis(option = 'H') +
-#   scale_y_reverse(limits = c(2000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
-
-# ggplot(particles, aes(x = juld, y = depth, colour = log(NP_Size_102))) + geom_point() + scale_color_viridis() +
-#   scale_y_reverse(limits = c(2000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
-
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux_without_depth)) + geom_point() + scale_color_viridis() +
-#   scale_y_reverse(limits = c(2000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
-
-```
-
-### SPECTRAL SLOPES
-
-```{r, function compute slope}
-
-# mid_DSE <- c(0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
-# sizebin <- c(0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
-#              0.532318310)
-
-mid_DSE <- c(0.0911143073, 0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
-sizebin <- c(0.0209587201, 0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
-             0.532318310)
-
-vertical_profiles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)
-drifting_profiles <- all_floats |> filter(PhaseName == 'PAR') |> select(juld, cycle, depth = pres, park_depth, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)
-
-compute_slope <- function(i, data_spectra){
-
-  spectrum <- data_spectra[i,]
-
-  spectrum_norm <- spectrum/sizebin
-
-  # prepare data for linear regression
-  Y <- log(spectrum_norm)
-  X <- log(mid_DSE)
-
-  # check for finite value
-  h <- is.finite(Y)
-  Y <- Y[h]
-  X <- X[h]
-
-  data_slope <- tibble(X=X, Y=Y)
-
-  model <- lm(formula = Y ~ X, data = data_slope)
-  slope <- model$coefficients[2]
-
-  return(slope)
-}
-
-# compute slope on drifting profiles
-drifting_spectra <- as.matrix(data.frame(drifting_profiles[,6:20]))
-index <- seq(from = 1, to = nrow(drifting_spectra), by = 1)
-slopes <- map_dbl(index, compute_slope, data_spectra = drifting_spectra)
-drifting_profiles$spectral_slope <- slopes
-ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-  scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')
-
-# compute slope on vertical profiles
-vertical_spectra <- as.matrix(data.frame(vertical_profiles[,5:19]))
-index <- seq(from = 1, to = nrow(vertical_spectra), by = 1)
-slopes <- map_dbl(index, compute_slope, data_spectra = vertical_spectra)
-vertical_profiles$spectral_slope <- slopes
-
-```
-
-```{r, add slopes}
-
-# vertical_profiles$slope <- slopes
-
-ggplot(vertical_profiles, aes(x = juld, y = depth, colour = spectral_slope)) + geom_point() + scale_colour_viridis(option = 'H') +
-  scale_y_reverse(limits = c(2000,0)) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = TeX('slope')) + facet_wrap(~wmo)
-
-```
-
-### DRIFT DATA
-
-```{r, check drift per class}
-
-tmp <- drifting_profiles |> pivot_longer(cols = all_of(lpm_class[2:16]), names_to = "size_class")
-tmp$park_depth <- factor(tmp$park_depth, levels = c('200 m', '500 m', '1000 m'))
-tmp$size_class <- factor(tmp$size_class, levels = lpm_class)
-
-tmp2 <- tmp |> group_by(park_depth, wmo, cycle, size_class) |> summarize(median_conc = median(value), min_juld = min(juld))
-
-ggplot(tmp2 |> filter(wmo == 6904241, park_depth != '200 m'), aes(x = min_juld, y = median_conc, colour = factor(park_depth))) + geom_point() + geom_path() + facet_wrap(~size_class, scales = 'free') + theme_bw() + labs(title = 'WMO 6904241')
-
-```
-
-```{r spectral slope + rati min/max flux}
-
-tmp <- cflux |> mutate(ratioF = small_flux/(small_flux+big_flux))
-
-# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF, colour = park_depth)) + geom_point() + facet_wrap(~wmo)
-# 
-# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_point() + facet_wrap(~park_depth)
-# 
-# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)
-
-ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_point() + geom_smooth() + facet_wrap(~park_depth)
-
-ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-  scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')
-
-tmp2 <- drifting_profiles |> group_by(wmo, cycle, park_depth) |> summarize(median_spectral_slope = median(spectral_slope))
-
-# merge data
-check <- merge(tmp, tmp2) 
-check2 <- merge(tmp, tmp2) |> filter(ratioF < 1)
-
-ggplot(check, aes(ratioF, median_spectral_slope)) + geom_point()
-
-ggplot(check2, aes(ratioF, median_spectral_slope)) + geom_point()
-
-# ggplot(tmp, aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)
-```
-
-```{r, check pulse export}
-
-# bbp_chla <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, chla, bbp) |> drop_na(chla) |> mutate(juld = as_date(juld))
-# cp_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, cp) |> drop_na(cp) |> mutate(juld = as_date(juld))
-# 
-# ggplot(bbp_chla, aes(x = juld, y = depth, colour = log10(chla))) + geom_point() + scale_color_viridis(option = 'H') +
-#   scale_y_reverse(limits = c(300,0)) + 
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
-# 
-# # ggplot(bbp_chla, aes(x = juld, y = depth, colour = log(bbp))) + geom_point() + scale_color_viridis(option = 'H') +
-# #   scale_y_reverse(limits = c(2000,0)) +
-# #   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-# #   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
-
-```
-
-### TRY: building a model of flux based on spectra (w/wo depth)
-
-```{r, same as above but without depth}
+
+### Flux model(s)
+
+```{r, flux model, include = FALSE}
 
@@
 
-  # get OST fluxes
-  cflux2 <- cflux |> drop_na(small_flux)
-
-  # # check 'outliers' -> justification: bad fluxes, see shiny (on fit des pentes sur 4 points, not enough, tout le reste ce sont des jumps)
-  # if(parking_depth == '500 m'){
-  #   cflux2 <- cflux2 |> filter(small_flux < 15)
-  # }
+flux_model <- function(boosting_rounds, which_flux){
+
+    # select which flux to build the model
+    if(which_flux == 'small'){
+      cflux2 <- cflux |> mutate(flux = small_flux) 
+    }else if(which_flux == 'big'){
+      cflux2 <- cflux |> mutate(flux = big_flux)
+    }else{
+      cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
+    }
 
@@
   # training data
-  tmp <- data_model |> select(small_flux, all_of(lpm_class))
+  tmp <- data_model |> select(flux, all_of(lpm_class))
 
@@
   # model
-  mg2 <- gbm(small_flux ~ ., data=tmp, distribution="gaussian",
+  model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
   cv.folds=5,
-  n.trees=3000, shrinkage=0.005, interaction.depth=6,
+  n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
   bag.fraction=0.5,
@@
   # get best model and thus best round number
-  best <- gbm.perf(mg2)
-  print(best)
-
-  #summary(mg2, n.trees=best)
-  tt <- summary(mg2, n.trees=best)
-  print(tibble(var = tt$var, importance = tt$rel.inf))
-
-    # predictions on mean values
-  uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = unique(small_flux), 
-                                                                                   NP_Size_64 = mean(NP_Size_64),
-                                                                                   NP_Size_80.6 = mean(NP_Size_80.6),
+  best <- gbm.perf(model)
+  #print(best)
+
+  # get features importance
+  features_importance <- summary(model, n.trees=best, plotit = FALSE)
+
+  return(list(features_importance = features_importance, model = model, best_tree = best))
+
+}
+
+# small_model <- flux_model(3000, 'small')
+# big_model <- flux_model(3000, 'big')
+# total_model <- flux_model(3000, 'total')
+# 
+# save(small_model, big_model, total_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data')
+
+load(file = 'DATA/MODEL/model_xgboost_data')
+
+```
+
+```{r, model predictions}
+
+model_predict <- function(which_flux, mod, best_tree_number){
+
+    # get spectra
+    uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
+  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
+
+    # select which flux to build the model
+    if(which_flux == 'small'){
+      cflux2 <- cflux |> mutate(flux = small_flux) 
+    }else if(which_flux == 'big'){
+      cflux2 <- cflux |> mutate(flux = big_flux)
+    }else{
+      cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
+    }
+
+    # merge data for training
+    data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
+
+  # predictions on mean spectra values
+  uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = unique(flux),
                                                                                    NP_Size_102 = mean(NP_Size_102),
@@
   # prediction at 1000 m
-  spectra <- uvp_spectra_average |> select(true_flux, all_of(lpm_class[]))
+  #spectra <- uvp_spectra_average |> select(true_flux, all_of(lpm_class[]))
 
   # prediction 
-  pred_flux <- predict(mg2, newdata = spectra, n.trees = best)
-
-  plot(x=spectra$true_flux, y=pred_flux, asp=1)
-  print(R2_Score(pred_flux, spectra$true_flux))
-
-```
-
-### Plot results
-
-Let's apply both models to all data at non parking depth (ie.
-vertical profiles) and let's plot a time serie to compare.
-Starting at 200 m car le modele devient n'a pas appris entre 0-200 et dans cette tranche, les concentrations de particules peuvent être huge huge -\> note que cela pourrait dire qu'il faudrait faire des mesures (autre instrument?
-pour estimer un flux..)
+  pred_flux <- predict(mod, newdata = uvp_spectra_average, n.trees = best_tree_number)
+
+  uvp_spectra_average$pred_flux <- pred_flux
+
+  return(uvp_spectra_average)
+}
+
+# pred_small <- model_predict('small', small_model$model, small_model$best_tree) |> select(wmo, cycle, depth, true_small = true_flux, pred_small = pred_flux)
+# pred_big <- model_predict('big', big_model$model, big_model$best_tree) |> select(true_big = true_flux, pred_big = pred_flux)
+# pred_total <- model_predict('total', total_model$model, total_model$best_tree) |> select(true_total = true_flux, pred_total = pred_flux)
+# 
+# all_predictions <- cbind(pred_small, pred_big, pred_total) |> mutate(pred_total2 = pred_small + pred_big, anom = pred_total - pred_total2)
+# 
+# vroom::vroom_write(all_predictions, file = '/home/flo/dox/THESIS/DATA/MODEL/all_predictions.csv')
+all_predictions <- vroom('DATA/MODEL/all_predictions.csv')
+
+```
+
+#### Features importance
+
+```{r, features importance}
+
+features <- small_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'small') |>
+  bind_rows(big_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'big')) |>
+  bind_rows(total_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'total'))
+
+rownames(features) <- NULL
+
+features <- features |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+
+ggplot(features) + geom_point(aes(x = as.factor(size), y = importance, colour = model), size = 5) + labs(x = 'Size', y = 'Features importance')
+
+```
+
+#### Predictions
+
+```{r, plot predictions}
+
+p_small <- ggplot(all_predictions) + geom_point(aes(true_small, pred_small)) + xlim(c(0,25)) + ylim(c(0,25)) + theme_bw() +
+  annotate("text", x = 3, y = 23, label = paste0('R² = ' , round(R2_Score(all_predictions$true_small, all_predictions$pred_small), 2)))
+
+p_big <- ggplot(all_predictions) + geom_point(aes(true_big, pred_big)) + xlim(c(0,50)) + ylim(c(0,50)) + theme_bw() +
+  annotate("text", x = 5, y = 45, label = paste0('R² = ' , round(R2_Score(all_predictions$true_big, all_predictions$pred_big), 2)))
+
+p_total <- ggplot(all_predictions) + geom_point(aes(true_total, pred_total)) + xlim(c(0,100)) + ylim(c(0,100)) + theme_bw() +
+  annotate("text", x = 3, y = 80, label = paste0('R² = ' , round(R2_Score(all_predictions$true_total, all_predictions$pred_total), 2)))
+
+(p_small + p_big) / p_total
+
+# plot(test$true_flux, y=test$pred_flux, asp=1)
+# print(R2_Score(test$pred_flux, test$true_flux))
+
+```
+
+#### Apply the total flux model on all vertical profiles
 
@@
 
-predicted_flux_without_depth <- predict(mg2, newdata = vertical_profiles, n.trees = best)
-
-vertical_profiles$pred_flux_without_depth <- predicted_flux_without_depth
-
-ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux_without_depth)) + geom_point() + scale_color_viridis(option = 'H') +
+# make predictions
+small_predicted_flux <- predict(small_model$model, newdata = vertical_profiles, n.trees = small_model$best_tree)
+big_predicted_flux <- predict(big_model$model, newdata = vertical_profiles, n.trees = big_model$best_tree)
+total_predicted_flux <- predict(total_model$model, newdata = vertical_profiles, n.trees = total_model$best_tree)
+
+# add predictions to vertical data
+vertical_profiles <- vertical_profiles |> mutate(small_pred = small_predicted_flux, big_pred = big_predicted_flux, total_pred = total_predicted_flux)
+
+vertical_profiles$wmo <- factor(vertical_profiles$wmo, levels = c(6904240,4903634,6904241,1902578))
+
+# pv_small <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(small_pred))) + geom_point() + scale_color_viridis(option = 'H') +
+#   scale_y_reverse(limits = c(500,0)) +
+#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+#   labs(x = 'Time', y = 'Predicted Fsmall') + facet_wrap(~wmo)
+# 
+# pv_big <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(big_pred))) + geom_point() + scale_color_viridis(option = 'H') +
+#   scale_y_reverse(limits = c(500,0)) +
+#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+#   labs(x = 'Time', y = 'Predicted Fbig') + facet_wrap(~wmo)
+
+pv_total <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
+  scale_y_reverse(limits = c(2000,0)) +
+  scale_colour_viridis(option = 'A', limits = c(quantile(all_predictions$true_total, probs = 0.25), quantile(all_predictions$true_total, probs = 0.75)), oob = scales::squish, na.value = 'grey', direction = -1) +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  labs(x = 'Time', y = 'Predicted Ftotal') + facet_wrap(~wmo) + theme_bw()
+
+pv_total
+
+ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
+  scale_y_reverse(limits = c(500,0)) +
+  scale_colour_viridis(option = 'A', limits = c(20,100), oob = scales::squish, na.value = 'grey', direction = -1) +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  labs(x = 'Time', y = 'Predicted Ftotal') + facet_wrap(~wmo) + theme_bw()
+
+```
+
+```{r, flux distribution}
+
+depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
+flux_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, total_pred) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
+  mutate(month = month.abb[month(juld)]) |> ungroup() #|> filter(month != 'May') # I filtered May because I had some troubles..
+I need to fill some cases unseen by the float (eg.
+when they start profiling until 1000 m then go to 2000 -> ça fout un peu la merde..)
+
+tmp <- flux_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_flux = mean(total_pred)) |> 
+  mutate(cut_flux = cut(mean_flux, breaks = c(0,10,20,30,40,50,75,100,+Inf))) |> ungroup()
+
+create_data_for_heatmap_bis <- function(float, df){
+  # df <- tmp |> filter(wmo == float)
+  # y <- unique(df$depth_bin)
+  # if(float == 1902578){
+  #   y <- y[1:15]
+  # }
+  # x <- unique(df$month)
+  # data <- expand.grid(X=x, Y=y)
+  # data$Z <- df$cut_flux
+  # data$wmo <- float
+  # data$month <- df$month
+  df <- df |> filter(wmo == float) |> select(cut_flux, depth_bin, month) 
+  y <- unique(df$depth_bin)
+  df <- df|> pivot_wider(names_from = depth_bin, values_from = cut_flux) 
+  x <- unique(df$month)
+  data <- expand.grid(X=x, Y=y)
+  data$Z <- as.vector(as.matrix(df |> select(-month)))
+  data$wmo <- float
+  return(data)
+}
+
+d1 <- create_data_for_heatmap_bis(6904241, tmp)
+d2 <- create_data_for_heatmap_bis(1902578, tmp)
+d3 <- create_data_for_heatmap_bis(4903634, tmp)
+d4 <- create_data_for_heatmap_bis(6904240, tmp)
+
+dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
+  mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
+  mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) |>
+  mutate(Z = factor(Z, levels = levels(tmp$cut_flux)))
+
+ggplot(dd) + 
+  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Flux range', x = 'Month', y = 'Depth bin') +
+  facet_wrap(~wmo) + scale_fill_viridis_d(option = 'A')
+
+```
+
+#### same plot mais pour les classes de tailles
+
+```{r, TEST }
+
+depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
+part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_512) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
+  mutate(month = month.abb[month(juld)]) |> ungroup() #|> filter(month != 'May') # I filtered May because I had some troubles..
+I need to fill some cases unseen by the float (eg.
+when they start profiling until 1000 m then go to 2000 -> ça fout un peu la merde..)
+
+tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_512))) |> 
+  mutate(cut_part = cut(mean_part, breaks = c(seq(0,100,20),+Inf))) |> ungroup()
+
+create_data_for_heatmap_bis <- function(float, df){
+  df <- df |> filter(wmo == float) |> select(mean_part, depth_bin, month) 
+  y <- unique(df$depth_bin)
+  df <- df|> pivot_wider(names_from = depth_bin, values_from = mean_part) 
+  x <- unique(df$month)
+  data <- expand.grid(X=x, Y=y)
+  data$Z <- as.vector(as.matrix(df |> select(-month)))
+  data$wmo <- float
+  return(data)
+}
+
+d1 <- create_data_for_heatmap_bis(6904241, tmp)
+d2 <- create_data_for_heatmap_bis(1902578, tmp)
+d3 <- create_data_for_heatmap_bis(4903634, tmp)
+d4 <- create_data_for_heatmap_bis(6904240, tmp)
+
+dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
+  mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
+  mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) #|>
+  #mutate(Z = factor(Z, levels = levels(tmp$cut_part)))
+
+ggplot(dd) + 
+  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Flux range', x = 'Month', y = 'Depth bin') +
+  facet_wrap(~wmo) + scale_fill_viridis(option = 'A')
+
+```
+
+#### Compute fluxes with Guidi et al. 2008
+
+A = 12.5, b = 3.81 fitted on 250 microns - 1.5 mm.
+
+```{r, guidi 2008}
+
+mid_DSE <- c(0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289)
+
+guidi_spectra <- vertical_profiles |> select(NP_Size_203:NP_Size_1290)
+
+# Guidi 2008
+A <- 12.5
+b <- 3.81
+exp <- A * mid_DSE ** b
+estimated_flux <- rowSums(t(t(guidi_spectra)*exp))
+#estimated_flux
+
+vertical_profiles$guidi_flux <- estimated_flux
+
+ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
+  scale_y_reverse(limits = c(2000,0)) +
+  scale_colour_viridis(option = 'A', limits = c(quantile(vertical_profiles$guidi_flux, probs = 0.25), quantile(vertical_profiles$guidi_flux, probs = 0.50)), oob = scales::squish, na.value = 'grey', direction = -1) +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  labs(x = 'Time', y = 'Predicted Guidi Flux') + facet_wrap(~wmo) + theme_bw()
+```
+
+#### Guidi vs. No Guidi Ha Diguedi
+
+```{r, anomaly guidi}
+
+ggplot(vertical_profiles |> mutate(ratio_flux = guidi_flux/total_pred)) + geom_point(aes(x = juld, y = depth, colour = ratio_flux)) +
+  scale_y_reverse(limits = c(2000,0)) +
+  scale_colour_viridis(option = 'A', limits = c(1,5), oob = scales::squish, na.value = 'grey') +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  labs(x = 'Time', y = 'Ratio Guidi/Total Model') + facet_wrap(~wmo) + theme_bw()
+
+```
+
+### SPECTRAL SLOPES
+
+```{r, function compute slope}
+
+# mid_DSE <- c(0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
+# sizebin <- c(0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
+#              0.532318310)
+
+mid_DSE <- c(0.0911143073, 0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
+sizebin <- c(0.0209587201, 0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
+             0.532318310)
+
+vertical_profiles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)
+drifting_profiles <- all_floats |> filter(PhaseName == 'PAR') |> select(juld, cycle, depth = pres, park_depth, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)
+
+compute_slope <- function(i, data_spectra){
+
+  spectrum <- data_spectra[i,]
+
+  spectrum_norm <- spectrum/sizebin
+
+  # prepare data for linear regression
+  Y <- log(spectrum_norm)
+  X <- log(mid_DSE)
+
+  # check for finite value
+  h <- is.finite(Y)
+  Y <- Y[h]
+  X <- X[h]
+
+  data_slope <- tibble(X=X, Y=Y)
+
+  model <- lm(formula = Y ~ X, data = data_slope)
+  slope <- model$coefficients[2]
+
+  return(slope)
+}
+
+# compute slope on drifting profiles
+drifting_spectra <- as.matrix(data.frame(drifting_profiles[,6:20]))
+index <- seq(from = 1, to = nrow(drifting_spectra), by = 1)
+slopes <- map_dbl(index, compute_slope, data_spectra = drifting_spectra)
+drifting_profiles$spectral_slope <- slopes
+ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
+  scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')
+
+# compute slope on vertical profiles
+vertical_spectra <- as.matrix(data.frame(vertical_profiles[,5:19]))
+index <- seq(from = 1, to = nrow(vertical_spectra), by = 1)
+slopes <- map_dbl(index, compute_slope, data_spectra = vertical_spectra)
+vertical_profiles$spectral_slope <- slopes
+
+```
+
+```{r, add slopes}
+
+# vertical_profiles$slope <- slopes
+
+ggplot(vertical_profiles, aes(x = juld, y = depth, colour = spectral_slope)) + geom_point() + scale_colour_viridis(option = 'H') +
   scale_y_reverse(limits = c(2000,0)) +
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
-
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log(pred_flux))) + geom_point() + scale_color_viridis() +
-#   scale_y_reverse(limits = c(1000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
-
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(pred_flux_without_depth))) + geom_point() + scale_color_viridis(option = 'H') +
-#   scale_y_reverse(limits = c(2000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
-
-ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(pred_flux_without_depth))) + geom_point() + scale_color_viridis(option = 'H') +
-  scale_y_reverse(limits = c(500,0)) +
+  labs(x = 'Time', y = TeX('slope')) + facet_wrap(~wmo)
+
+```
+
+### DRIFT DATA
+
+```{r spectral slope + rati min/max flux}
+
+tmp <- cflux |> mutate(ratioF = small_flux/(small_flux+big_flux))
+
+# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF, colour = park_depth)) + geom_point() + facet_wrap(~wmo)
+# 
+# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_point() + facet_wrap(~park_depth)
+# 
+# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)
+
+ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_point() + geom_smooth() + facet_wrap(~park_depth)
+
+ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + 
-  labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) + facet_wrap(~wmo)
-
-ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(pred_flux_without_depth))) + geom_point() + scale_color_viridis(option = 'H') +
-  scale_y_reverse(limits = c(2000,200)) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) + facet_wrap(~wmo)
-
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(pred_flux_without_depth))) + geom_point() + scale_color_viridis(option = 'H') +
-#   scale_y_reverse(limits = c(500,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) + facet_wrap(~wmo)
-
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log(pred_flux_without_depth))) + geom_point() + scale_color_viridis() +
-#   scale_y_reverse(limits = c(2000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) + facet_wrap(~wmo)
-
-# cflux_200m$depth <- 200
-# cflux_200m$juld <- as_date(cflux_200m$min_time)
-# 
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux)) + geom_point() + scale_color_viridis() +
-#   scale_y_reverse(limits = c(1000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) +
-#   geom_point(data = cflux_200m, aes(x = juld, y = depth, colour = 'black', size = small_flux))
-
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log(pred_flux))) + geom_point() + scale_color_viridis() +
-#   scale_y_reverse(limits = c(1000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))
-
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux)) + geom_point() + scale_color_viridis() +
-#   scale_y_reverse(limits = c(1000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) +
-#   facet_wrap(~wmo)
-# 
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log(pred_flux))) + geom_point() + scale_color_viridis() +
-#   scale_y_reverse(limits = c(1000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) +
-#   facet_wrap(~wmo)
-
-# cflux_200m$depth <- 200
-# cflux_200m$juld <- as_date(cflux_200m$min_time)
-# 
-# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux)) + geom_point() + scale_color_viridis() +
-#   scale_y_reverse(limits = c(1000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$')) +
-#   geom_point(data = cflux_200m, aes(x = juld, y = depth, colour = small_flux), shape = 17, size = 2)
-
-# ggplot(cflux_200m, aes(x = juld, y = depth, colour = small_flux)) + geom_point() + scale_color_viridis()
-
-```
-
-Modèle appris en discontinu et prédictions en continu --\> BOF BOF et ça se voit.
-L'autre modèle a une bien plus belle gueule et ne prend pas la profondeur en compte ce qui est encore mieux selon moi, le small flux est expliqué principalement par 102, 128 et 161, ce qui est cohérent.
-On voit potentiellement un pulse d'export --\> faudrait le vérifier avec le bbp/cp?
-On voit que vers la fin de l'année, le flux en profondeur est plus faible -\> cohérent vu que le bloom date de bien avant.
-On aurait besoin d'une time series plus longue, ici j'ai bloqué à 6 mois (car le drift à 200 m va etre mis en pause avec l'hiver et la MLD + le fait que l'on perd un flotteur avec un transmissio qui merde).
-Ajouter MLD, iso-sigma?
-
-```{r, check sthg}
-
-# particles_grouped <- particles |> mutate(NP_Size_102_256 = rowSums(across(c(NP_Size_102,NP_Size_128,NP_Size_161,NP_Size_203))),
-#                                          NP_Size_256_512 = rowSums(across(c(NP_Size_256,NP_Size_323,NP_Size_406))),
-#                                          NP_Size_512_1020 = rowSums(across(c(NP_Size_512,NP_Size_645,NP_Size_813))),
-#                                          NP_Size_1020_2580 = rowSums(across(c(NP_Size_1020,NP_Size_1290, NP_Size_1630,NP_Size_2050))))
-# 
-# particles_ter <- particles_grouped |> pivot_longer(cols = c(NP_Size_102_256, NP_Size_256_512, NP_Size_512_1020, NP_Size_1020_2580), names_to = 'size_group', values_to = 'concentration')
-# #particles_ter$size_group <- factor(particles_ter$size_class, levels = lpm_class)
-# 
-# ggplot(particles_ter |> filter(wmo == '6904240'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
-#   scale_y_reverse(limits = c(2000,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'), title = 'wmo 6904240') +
-#   facet_wrap(~size_group) + theme_bw()
-
+  scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')
+
+tmp2 <- drifting_profiles |> group_by(wmo, cycle, park_depth) |> summarize(median_spectral_slope = median(spectral_slope))
+
+# merge data
+check <- merge(tmp, tmp2) 
+check2 <- merge(tmp, tmp2) |> filter(ratioF < 1)
+
+ggplot(check, aes(ratioF, median_spectral_slope)) + geom_point()
+
+ggplot(check2, aes(ratioF, median_spectral_slope)) + geom_point()
+
+# ggplot(tmp, aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)
 ```

### 2023-02-03 article3.qmd
update

@@
 depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
-part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_512) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
-  mutate(month = month.abb[month(juld)]) |> ungroup() #|> filter(month != 'May') # I filtered May because I had some troubles..
-I need to fill some cases unseen by the float (eg.
-when they start profiling until 1000 m then go to 2000 -> ça fout un peu la merde..)
-
-tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_512))) |> 
+part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_256) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
+  mutate(month = month.abb[month(juld)]) |> ungroup()
+
+tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_256))) |> 
   mutate(cut_part = cut(mean_part, breaks = c(seq(0,100,20),+Inf))) |> ungroup()
@@
 ggplot(dd) + 
-  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Flux range', x = 'Month', y = 'Depth bin') +
+  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Log10(abundance) #/L', x = 'Month', y = 'Depth bin', title = 'NP_Size_256') +
   facet_wrap(~wmo) + scale_fill_viridis(option = 'A')
+
+```
+
+#### same plots mais par mois
+
+Il y a une erreur lors de la standardisation -> WHY doesn't it want to group it by month???
+
+```{r, monthly heatmap}
+
+depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
+part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, all_of(lpm_class)) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |> mutate(month = month.abb[month(juld)]) |> ungroup() |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration')
+
+tmp <- part_distrib |> group_by(depth_bin, size, month) |> summarize(mean_part = mean(concentration)) |> mutate(rel_mean_conc = (mean_part-min(mean_part))/(max(mean_part) - min(mean_part))) |> ungroup()
+
+plot_heat_month <- function(m){
+  # transform data for heatmap
+  #m <- 'May'
+  tmp2 <- tmp |> filter(month == m) |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) |> select(-c(month, mean_part))
+  y <- unique(tmp2$depth_bin)
+  x <- unique(tmp2$size)
+  test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = rel_mean_conc) 
+  #test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = mean_part) 
+  data <- expand.grid(X=x, Y=y)
+  data$Z <- as.vector(as.matrix(test |> select(-size)))
+  data$month <- m
+  return(data)
+}
+
+m_data <- map_dfr(c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'), plot_heat_month)
+
+m_data <- m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)),
+                   month = factor(month, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')))
+
+ggplot(m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)))) + 
+  geom_tile(aes(as.factor(X), Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + scale_fill_viridis(option = 'A') +
+  facet_wrap(~month)
 
@@
   labs(x = 'Time', y = 'Ratio Guidi/Total Model') + facet_wrap(~wmo) + theme_bw()
+
+```
+### Modèle basé sur les pentes
+
+```{r, model based on slope}
+
+# compute biovolumes for mid size bin (ESD)
+# DSE from 102 microns to 2.05 mm (this class to goes to 2.58 mm) (values obtained from Guidi's Matlab code converted on python)
+DSE <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
+              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+              1.16170742, 1.46365963, 1.84409558, 2.32341484)
+
+# DSE obtained with Rainer's code
+DSE_Rainer <- c(0.11645117591608793,0.14636003747903648,0.18439151433041295,
+                0.23251999827981656,0.2933257316240555,0.36916500651202966,0.4650399965596331,
+                0.5860454935035602,0.738553273391974,0.9280422447589767,1.1705684285914486,
+                1.4795320531800742,1.8636618038457486,2.3449457391257185)
+
+# compute biovolume
+get_biovolume <- function(DSE){
+  r <- DSE/2
+  biovolume <- (4*pi/3)*r**3
+  return(biovolume)
+}
+
+# get biovolume for each DSE (equivalent to say for each size bin)
+biovolumes <- map_dbl(DSE, get_biovolume)  # unit = m^3
+
+# spectra at parking depth
+uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
+  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
+
+# convert abundance to biovolume
+spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes))) 
+colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')
+
+uvp_spectra <- cbind(uvp_spectra, spectra_biovolume) |> mutate()
+
+cflux <- cflux |> mutate(small_slope = (small_flux/633)**(1/0.77), big_slope = (big_flux/633)**(1/0.77))
+
+## try model for small slopes
+cflux2 <- cflux |> mutate(flux = big_slope)
+
+data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
+
+# training data
+tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))
+
+set.seed(101)
+boosting_rounds <- 3000
+# model
+model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
+cv.folds=5,
+n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
+bag.fraction=0.5,
+n.cores=1
+)
+
+# get best model and thus best round number
+best <- gbm.perf(model)
+#print(best)
+
+# get features importance
+features_importance <- summary(model, n.trees=best, plotit = FALSE)
 

### 2023-02-05 article3.qmd
update

@@
 introduire la notion de cycle de flotteurs BGC-Argo etc (+ parking depth)
+
++ dire que le but du papier est de faire une comparaison OST UVP qui sont deux instruments qui permettent de mesurer la meme chose (le flux des particules, l'abondance de particules) et donc le flux de carbone de manière indirecte
+papier GRL Lacour de 2015
 
@@
 library(gt)
+library(gtExtras)
 ```
@@
 all_floats <- rbind(float_6904240, float_6904241, float_4903634, float_1902578)
+
+all_floats <- all_floats |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
+                                   park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
 
@@
 Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach instead of fixed thresholds to detect jumps.
-Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase.
-Then, we computed the z-score (i.e.
-standard score) of each slope, that follows a normal distribution.
-The detection of jumps thus consisted in spotting outliers in the z-score distribution.
+Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase and computed the z-score (i.e.
+standard score) of each of them.
+The detection of jumps therefore consisted in spotting outliers in the z-score distribution.
 For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles.
@@
 
-# add levels
-cflux$park_depth <- factor(cflux$park_depth, levels = c('200 m', '500 m', '1000 m'))
+# # add levels
+# cflux$park_depth <- factor(cflux$park_depth, levels = c('200 m', '500 m', '1000 m'))
 
@@
 #uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
-uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))
+# uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))
 
@@
 
-At 500 m, particles concentrations in the 102 - 645 $\mu$ range slowly decrease from July to October with a sharp decrease in November.
+At 500 m, particles concentrations in the 102 - 645 $\mu$ range slowly decrease from July to September with a sharp decrease starting in October.
 Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (< 0.5 particle/L) on the entire timeserie.
@@
 
+TODO : It should be noted that all floats measure the same trend -> logique si on parle de la dynamique du bruit de fond de la zone et pas de la dynamique évènementielle qui serait différentes en fonction des flotteurs vu qu'ils dérivent chacun dans le parcours respectifs.
+Ça devrait renforcer cet aspect.
+
+## Discussion
+
+### Comparing OST and UVP data
+
+For the first time, BGC-Argo floats equipped with both a transmissometer and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx].
+The quantification of the POC flux using the OST method [@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr] is based on an empirical POC:ATN ratio \[REF estapa 2022\] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:
+
+$$
+F = \sum_{i}^{N}C_{i}Ad_{i}^{b}
+$$ {#eq-guidi-article3}
+
+where $C_{i}$, A and b represent, respectively, the concentration of particles (number/L) for the $i^{th}$ class of mean diameter $d_{i}$ (in mm) and 2 constants with $b$ being related to the fractal dimension ($D$) of an aggregate by $D = (b +1)/2$ [@Guidi2008-vp].
+
+The use of @eq-guidi-article3 to derive carbon fluxes from PSD was firstly implemented by @Guidi2008-vp who used a minimization procedure between sediment traps data and estimated fluxes using PSD from several previous versions of the UVP to find the optimal value of $A$ and $b$ (i.e.
+12.5 and 3.81).
+@Iversen2010-an and @Fender2019-ou followed the same procedure and concluded that the couple of parameters found by [@Guidi2008-vp] could not be used globally at the expense of greatly over- or underestimating POC fluxes, depending on the sampling region.
+In addition, the formulation of such a model implies that $A$ and $b$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt].
+@eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2022-qj].
+As a result, the direct comparison of carbon fluxes derived from both OST and UVP data with @eq-guidi-article3 should be focused on the trend rather than on absolute carbon flux values.
+
+```{r, compute guidi fluxes at parking depth}
+#| label: fig-comparison-ost-uvp
+#| fig-cap: Comparison of total carbon fluxes derived from the OST (black) and the UVP6 (red) where the error bars represent the 25$^{th}$ and 75^{th}$ percentiles around the median.
+
+compute_flux <- function(A, b, PSD){ # PSD = Particle Size Distribution
+  # for classes sizes between 0.102 mm and 1.5 mm
+  #mid_ESD <- c(0.1164512, 0.1463600, 0.1843915, 0.2325200, 0.2933257, 0.3691650, 0.4650400, 0.5860455, 0.7385533, 0.9280422, 1.1705684, 1.4795321) # With Rainer's code (paruvpy)
+  # mid_ESD <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
+  #              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+  #              1.16170742, 1.46365963) # With Lionel's script
+  mid_ESD <- c(0.23051195, 0.29042685,
+               0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+               1.16170742, 1.46365963) # With Lionel's script
+  exp <- A * mid_ESD ** b
+  estimated_flux <- rowSums(t(t(PSD)*exp))
+  return(estimated_flux)
+}
+
+parking_data <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
+parking_data <- parking_data |> filter(PhaseName == 'PAR') |> drop_na(NP_Size_102) |> select(park_depth, all_of(lpm_class), dates = juld, park_depth, wmo, cycle) |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
+                                   park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
+parking_spectra <- parking_data |> dplyr::select(NP_Size_203:NP_Size_1290)
+
+parking_data <- parking_data |> mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) |> group_by(park_depth, cycle, wmo) #|> summarize(median_guidi_flux = median(guidi_flux))
+
+# plot it
+comparison_flux_ost_uvp <- merge(parking_data, cflux) |> mutate(total_flux = small_flux + big_flux)
+
+# https://stackoverflow.com/questions/41077199/how-to-use-r-ggplot-stat-summary-to-plot-median-and-quartiles
+# ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = small_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
+#   geom_errorbar(aes(x = date, y = guidi_flux),
+#                       stat = 'summary',
+#                       fun.min = function(z) {quantile(z, 0.25)},
+#                       fun.max = function(z) {quantile(z, 0.75)},
+#                       fun = median, colour = 'red', width = 5) +
+#   geom_line(aes(x = date, y = guidi_flux),
+#                       stat = 'summary',
+#                       fun = median, colour = 'red') + scale_y_continuous(trans = 'log10') +
+#   geom_line(data = comparison_flux_ost_uvp, aes(x = date, y = big_flux), linetype = 'dotdash') + theme_bw()
+
+ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = total_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
+  geom_errorbar(aes(x = date, y = guidi_flux),
+                      stat = 'summary',
+                      fun.min = function(z) {quantile(z, 0.25)},
+                      fun.max = function(z) {quantile(z, 0.75)},
+                      fun = median, colour = 'red', width = 5) +
+  geom_line(aes(x = date, y = guidi_flux),
+                      stat = 'summary',
+                      fun = median, colour = 'red') + scale_y_continuous(trans = 'log10') + theme_bw()
+
+correlations <- comparison_flux_ost_uvp |> select(park_depth, wmo, cycle, guidi_flux, small_flux, big_flux, date, total_flux) |> group_by(wmo, park_depth, cycle) |> mutate(median_guidi_flux = median(guidi_flux)) |> group_by(park_depth, wmo) |> summarise(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_big = round(cor(median_guidi_flux, big_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))
+
+correlations <- correlations |> pivot_wider(names_from = park_depth, values_from = c(cor_small, cor_big, cor_total))
+gt(correlations) |> tab_spanner(label = "Small flux", columns = 2:4) |> tab_spanner(label = "Big flux", columns = 5:7) |> tab_spanner(label = 'Total flux', columns = 8:10) |>
+  cols_label(wmo = 'WMO', `cor_small_200 m` = '200 m',
+             `cor_small_500 m` = '500 m',
+             `cor_small_1000 m` = '1000 m',
+             `cor_big_200 m` = '200 m',
+             `cor_big_500 m` = '500 m',
+             `cor_big_1000 m` = '1000 m',
+             `cor_total_200 m` = '200 m',
+             `cor_total_500 m` = '500 m',
+             `cor_total_1000 m` = '1000 m')
+
+```
+
+TODO :
+
+- TS diagram at drifting depth + isopycnals
+- NPQ chla + MLD (same y axis) 
+- satellite chla + MLD -> show that conditions for bloom (stable and shallow MLD) + chla happened before deployment in the area
+- faire le modèle de pente avec les biovolues pour prédire quels types de particules explique le plus la pente -> faire un modèle pour le small flux et un modèle pour le big flux, pas un modèle pour les small + big car les pentes ne sont pas additives
+- discuter des corrélations
+- science: on voit le système latent, on a raté le bloom
+- rajouter les vitesses de sédimentations (n'importe quelle méthode fera l'affaire)
+- rajouter dans les méthodes les produits satellites qu'on utilise
+- différentes dynamiques post-bloom
+- phénologie du bloom printanier
+- MLD au cours du temps dans la zone, voir les variations
+
+### Which particles drive
+
+```{r, flux model 2, include = FALSE}
+
+# get spectra
+uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
+  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
+
+flux_model <- function(boosting_rounds, which_flux){
+
+    # select which flux to build the model
+    if(which_flux == 'small'){
+      cflux2 <- cflux |> mutate(flux = small_flux) 
+    }else if(which_flux == 'big'){
+      cflux2 <- cflux |> mutate(flux = big_flux)
+    }else{
+      cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
+    }
+
+    # merge data for training
+    data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
+
+  # training data
+  tmp <- data_model |> select(flux, all_of(lpm_class))
+
+  set.seed(101)
+  # model
+  model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
+  cv.folds=5,
+  n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
+  bag.fraction=0.5,
+  n.cores=1
+  )
+
+  # get best model and thus best round number
+  best <- gbm.perf(model)
+  #print(best)
+
+  # get features importance
+  features_importance <- summary(model, n.trees=best, plotit = FALSE)
+
+  return(list(features_importance = features_importance, model = model, best_tree = best))
+
+}
+
+# small_model <- flux_model(3000, 'small')
+# big_model <- flux_model(3000, 'big')
+# total_model <- flux_model(3000, 'total')
+# 
+# save(small_model, big_model, total_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data')
+
+#load(file = 'DATA/MODEL/model_xgboost_data')
+
+```
+
 #### On profiling mode
+
+Cette partie devrait venir dans la discussion si on part du principe que la première partie est liée à la méthodologie
 
@@
 # order floats by couple of deployment
-particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904240,4903634,6904241,1902578))
+# particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904240,4903634,6904241,1902578))
 
@@
 
-## Discussion
+### Environmental conditions
+
+Add NPQ plot avec MLD dessus, les memes échelles verticales (depth) + aller chercher les produits satellitaires pour montrer que l'on a bien raté le bloom ou que les conditions du bloom n'étaient pas réunies
 
@@
 dd$Y <- factor(dd$Y, levels = c('1290', '1020', '813', '645', '512', '406', '323', '256', '203', '161', '128', '102'))
-dd$wmo <- factor(dd$wmo, levels = c(6904240,4903634,6904241,1902578))
+# dd$wmo <- factor(dd$wmo, levels = c(6904240,4903634,6904241,1902578))
 
@@
 
-### Combining OST and UVP data
-
-For the first time, BGC-Argo floats equipped with both a transmissometer and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx].
-The quantification of the POC flux using the OST method [@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr] is based on an empirical POC:ATN ratio \[REF estapa 2022\] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:
-
-$$
-F = \sum_{i}^{N}C_{i}Ad_{i}^{b}
-$$ {#eq-guidi-article3}
-
-where $C_{i}$, A and b represent, respectively, the concentration of particles (number/L) for the $i^{th}$ class of mean diameter $d_{i}$ (in mm) and 2 constants with $b$ being related to the fractal dimension ($D$) of an aggregate by $D = (b +1)/2$ [@Guidi2008-vp].
-
-The use of @eq-guidi-article3 to derive carbon fluxes from PSD was firstly implemented by @Guidi2008-vp who used a minimization procedure between sediment traps data and estimated fluxes using PSD from several previous versions of the UVP to find the optimal value of $A$ and $b$ (i.e.
-12.5 and 3.81).
-@Iversen2010-an and @Fender2019-ou followed the same procedure and concluded that the couple of parameters found by [@Guidi2008-vp] could not be used globally at the expense of greatly over- or underestimating POC fluxes, depending on the sampling region.
-In addition, the formulation of such a model implies that $A$ and $b$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt].
-@eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2022-qj].
-Due to those reasons, we have chosen to firstly compare $F_{small}$ and $F_{big}$ with UVP6 particles concentrations.
-
 ### Flux model(s)

### 2023-02-06 article3.qmd
update

@@
 
-+ dire que le but du papier est de faire une comparaison OST UVP qui sont deux instruments qui permettent de mesurer la meme chose (le flux des particules, l'abondance de particules) et donc le flux de carbone de manière indirecte
-papier GRL Lacour de 2015
+-   dire que le but du papier est de faire une comparaison OST UVP qui sont deux instruments qui permettent de mesurer la meme chose (le flux des particules, l'abondance de particules) et donc le flux de carbone de manière indirecte papier GRL Lacour de 2015
 
@@
 Between 512 $\mu$m and 1.29 mm, concentrations are steadily decreasing from June.
-Above 1.29 mm, the abundance of particles is already very low (< 0.1 particle/L) in June and keeps decreasing over time.
+Above 1.29 mm, the abundance of particles is already very low (\< 0.1 particle/L) in June and keeps decreasing over time.
 
@@
 At 500 m, particles concentrations in the 102 - 645 $\mu$ range slowly decrease from July to September with a sharp decrease starting in October.
-Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (< 0.5 particle/L) on the entire timeserie.
+Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (\< 0.5 particle/L) on the entire timeserie.
 
@@
 Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.
+#| fig-dpi: 96
+
+#{r, uvp data at 1000 m, out.width="900px", out.height="800px"}
 
@@
 
-TODO : It should be noted that all floats measure the same trend -> logique si on parle de la dynamique du bruit de fond de la zone et pas de la dynamique évènementielle qui serait différentes en fonction des flotteurs vu qu'ils dérivent chacun dans le parcours respectifs.
+TODO : It should be noted that all floats measure the same trend -\> logique si on parle de la dynamique du bruit de fond de la zone et pas de la dynamique évènementielle qui serait différentes en fonction des flotteurs vu qu'ils dérivent chacun dans le parcours respectifs.
 Ça devrait renforcer cet aspect.
@@
 -   TS diagram at drifting depth + isopycnals
-- NPQ chla + MLD (same y axis) 
-- satellite chla + MLD -> show that conditions for bloom (stable and shallow MLD) + chla happened before deployment in the area
-- faire le modèle de pente avec les biovolues pour prédire quels types de particules explique le plus la pente -> faire un modèle pour le small flux et un modèle pour le big flux, pas un modèle pour les small + big car les pentes ne sont pas additives
+-   satellite chla + MLD -\> show that conditions for bloom (stable and shallow MLD) + chla happened before deployment in the area
+-   faire le modèle de pente avec les biovolues pour prédire quels types de particules explique le plus la pente -\> faire un modèle pour le small flux et un modèle pour le big flux, pas un modèle pour les small + big car les pentes ne sont pas additives
 -   discuter des corrélations
 -   science: on voit le système latent, on a raté le bloom
-- rajouter les vitesses de sédimentations (n'importe quelle méthode fera l'affaire)
+-   rajouter les vitesses de sédimentations (n'importe quelle méthode fera l'affaire) -> permet d'utiliser le graphique de l'UVP en vertical -> hop une justification en plus
 -   rajouter dans les méthodes les produits satellites qu'on utilise
@@
 -   phénologie du bloom printanier
-- MLD au cours du temps dans la zone, voir les variations
-
-### Which particles drive
+
+### Which particles drive the flux
 
@@
 down_part
-
-# # interpolate to see better
-# dli <- particles_bis |> mutate(juldi = as.numeric(juld)) |>
-#   group_by(wmo, size) |>
-#   do({
-#     i <- akima::interp(x=.$juldi, y=.$depth, z=.$concentration, nx=50, ny=150, duplicate="mean") |>
-#       tidy()
-#     # ggplot(i) + geom_raster(aes(x=x, y=-y, fill=z))
-#     rename(i, juldi=x, pres=y, concentration=z)
-#   }) |>
-#   # re-compute relative concentration to see stuff better
-#   group_by(size) |>
-#   mutate(rel_conc=concentration/max(concentration, na.rm=TRUE)) |>
-#   mutate(juld = as_date(juldi, origin = "1970-01-01")) |>
-#   drop_na(rel_conc) |>
-#   ungroup()
+```
+
+### Environmental conditions prior to bloom
+
+ + aller chercher les produits satellitaires pour montrer que l'on a bien raté le bloom ou que les conditions du bloom n'étaient pas réunies
+
+```{r, chla corrected for NPQ and MLD}
+
+# fetch data
+npq_data <- vroom('DATA/ARGO/npq_corrected_data.csv')
+
+# clean and smoothed chla data
+d <- npq_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> mutate(date = as_date(juld))
+d <- d |> group_by(date, wmo, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
+d <- d |> mutate(smoothed_chla = smooth(mean_chla))
+
+#d$wmo <- factor(d$wmo, levels = c(6904240,4903634,6904241,1902578))
+
+# plot it
+ggplot(d) + geom_point(aes(x = date, y = pres, colour = log10(smoothed_chla))) +
+  scale_y_reverse(limits = c(200,0)) +
+  #scale_colour_viridis(option = 'A') +
+  scale_colour_cmocean(name = 'algae', oob = scales::squish, limits = c(-1,1), na.value = 'grey') +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  labs(x = 'Time', y = 'NPQ-corrected Chla') + facet_wrap(~wmo) + theme_bw() +
+  geom_line(aes(x = date, y = mld), colour = 'black', size = 1)
+
+# compute chla integration in the MLD and in the 0-100 m zone
+d_chla <- d |> group_by(wmo, date) |> summarize(chla_0_100m = integrate(smoothed_chla, pres, from=0, to=100),
+                                       mld = unique(mld),
+                                       chla_0_mld = integrate(smoothed_chla, pres, from = 0, to=mld))
 # 
-# dli |> 
-#   ggplot() + facet_grid(wmo~size) +
-#   geom_raster(aes(x=juldi, y=-pres, fill=rel_conc)) +
-#   #scale_y_reverse() |>
-#   scale_x_continuous(expand=c(0,0), limits=c(NA, NA)) +
-#   scale_y_continuous(expand=c(0,0), limits=c(-300,0)) +
-#   scale_fill_viridis_c(option = 'H')
-
-# test_sigma <- all_floats |> filter(PhaseName == 'NPAR', sigma > 26) |> select(juld, cycle, depth = pres, wmo, sigma, MLD) |> drop_na(sigma) |> mutate(juld = as_date(juld))
-# test_sigma |> ggplot() + geom_point(aes(x = juld, y = depth, colour = sigma)) + 
-#   scale_colour_viridis(option = 'H', oob = scales::squish, na.value = 'transparent') +
-#   scale_y_reverse(limits = c(200,0)) +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
-#   labs(x = 'Month', y = 'Depth (m)') +
-#   facet_wrap(~wmo, nrow = 4) + theme_bw() +
-#   geom_path(data = test_sigma, aes(juld, MLD), colour = "grey")
-
-```
-
-### Environmental conditions
-
-Add NPQ plot avec MLD dessus, les memes échelles verticales (depth) + aller chercher les produits satellitaires pour montrer que l'on a bien raté le bloom ou que les conditions du bloom n'étaient pas réunies
+# ggplot(d_chla) + geom_point(aes(date, chla_0_100m)) + facet_wrap(~wmo, scales = 'free') + theme_bw() + labs(x = 'Month', y = 'Integrated Chla 0-100 m (CHLA UNITS)')
+#ggplot(d_chla) + geom_point(aes(date, chla_0_mld)) + facet_wrap(~wmo, scales = 'free')
+
+```
+
+```{r, chla and MLD from satellite}
+
+chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_with_clouds_sat.feather')
+#chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_without_clouds_sat.feather')
+mld_sat <- arrow::read_feather('DATA/SATELLITE/mld_sat.feather') |> mutate(lat = latitude, lon = longitude - 360) |> select(time, lat, lon, mld = mlotst)
+
+# CHLA
+x <- 1
+y <- 1
+chla_cover_deployment1 <- chla_sat |> filter(lat > 58.9 - y, lat < 58.9 + y, lon > -49.3 - x, lon < -49.3 + x) |> mutate(date = as_date(time), group = 'couple 1') |> mutate(deploy = as_date('2022-05-22'))
+chla_cover_deployment2 <- chla_sat |> filter(lat > 56.7 - y, lat < 56.7 + y, lon > -52.3 - x, lon < -52.3 + x) |> mutate(date = as_date(time), group = 'couple 2') |> mutate(deploy = as_date('2022-05-30'))
+chla_cover <- rbind(chla_cover_deployment1, chla_cover_deployment2)
+
+ggplot(chla_cover) + geom_errorbar(aes(x = date, y = CHL),
+                      stat = 'summary',
+                      fun.min = function(z) {quantile(z, 0.25)},
+                      fun.max = function(z) {quantile(z, 0.75)},
+                      fun = median, colour = 'black', width = .5) + theme_bw() +
+  labs(x = 'Time', y = 'Chla')+ facet_wrap(~group) +
+  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "red")
+
+# MLD
+mld_cover_deployment1 <- mld_sat |> filter(lat > 58.9 - y, lat < 58.9 + y, lon > -49.3 - x, lon < -49.3 + x) |> mutate(date = as_date(time), group = 'couple 1') 
+mld_cover_deployment2 <- mld_sat |> filter(lat > 56.7 - y, lat < 56.7 + y, lon > -52.3 - x, lon < -52.3 + x) |> mutate(date = as_date(time), group = 'couple 2')
+mld_cover <- rbind(mld_cover_deployment1, mld_cover_deployment2)
+
+ggplot(mld_cover) + geom_errorbar(aes(x = date, y = mld),
+                      stat = 'summary',
+                      fun.min = function(z) {quantile(z, 0.25)},
+                      fun.max = function(z) {quantile(z, 0.75)},
+                      fun = median, colour = 'black', width = 5) +
+  labs(x = 'Time', y = 'MLD (m)') + facet_wrap(~group, scales = 'free_y')
+
+```
 
@@
 ```
+
 ### Mixed layer depth
@@
 p_small <- ggplot(all_predictions) + geom_point(aes(true_small, pred_small)) + xlim(c(0,25)) + ylim(c(0,25)) + theme_bw() +
+
   annotate("text", x = 3, y = 23, label = paste0('R² = ' , round(R2_Score(all_predictions$true_small, all_predictions$pred_small), 2)))
@@
 p_big <- ggplot(all_predictions) + geom_point(aes(true_big, pred_big)) + xlim(c(0,50)) + ylim(c(0,50)) + theme_bw() +
+
   annotate("text", x = 5, y = 45, label = paste0('R² = ' , round(R2_Score(all_predictions$true_big, all_predictions$pred_big), 2)))
@@
 p_total <- ggplot(all_predictions) + geom_point(aes(true_total, pred_total)) + xlim(c(0,100)) + ylim(c(0,100)) + theme_bw() +
+
   annotate("text", x = 3, y = 80, label = paste0('R² = ' , round(R2_Score(all_predictions$true_total, all_predictions$pred_total), 2)))
@@
 # plot(test$true_flux, y=test$pred_flux, asp=1)
+
 # print(R2_Score(test$pred_flux, test$true_flux))
@@
 # make predictions
+
 small_predicted_flux <- predict(small_model$model, newdata = vertical_profiles, n.trees = small_model$best_tree)
+
 big_predicted_flux <- predict(big_model$model, newdata = vertical_profiles, n.trees = big_model$best_tree)
+
 total_predicted_flux <- predict(total_model$model, newdata = vertical_profiles, n.trees = total_model$best_tree)
@@
 # add predictions to vertical data
+
 vertical_profiles <- vertical_profiles |> mutate(small_pred = small_predicted_flux, big_pred = big_predicted_flux, total_pred = total_predicted_flux)
@@
 # pv_small <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(small_pred))) + geom_point() + scale_color_viridis(option = 'H') +
+
 #   scale_y_reverse(limits = c(500,0)) +
+
 #   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+
 #   labs(x = 'Time', y = 'Predicted Fsmall') + facet_wrap(~wmo)
+
 #
+
 # pv_big <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(big_pred))) + geom_point() + scale_color_viridis(option = 'H') +
+
 #   scale_y_reverse(limits = c(500,0)) +
+
 #   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+
 #   labs(x = 'Time', y = 'Predicted Fbig') + facet_wrap(~wmo)
@@
 pv_total <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
+
   scale_y_reverse(limits = c(2000,0)) +
+
   scale_colour_viridis(option = 'A', limits = c(quantile(all_predictions$true_total, probs = 0.25), quantile(all_predictions$true_total, probs = 0.75)), oob = scales::squish, na.value = 'grey', direction = -1) +
+
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+
   labs(x = 'Time', y = 'Predicted Ftotal') + facet_wrap(~wmo) + theme_bw()
@@
 ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
+
   scale_y_reverse(limits = c(500,0)) +
+
   scale_colour_viridis(option = 'A', limits = c(20,100), oob = scales::squish, na.value = 'grey', direction = -1) +
+
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+
   labs(x = 'Time', y = 'Predicted Ftotal') + facet_wrap(~wmo) + theme_bw()
@@
 depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
+
 flux_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, total_pred) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
+
   mutate(month = month.abb[month(juld)]) |> ungroup() #|> filter(month != 'May') # I filtered May because I had some troubles..
@@
 tmp <- flux_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_flux = mean(total_pred)) |>
+
   mutate(cut_flux = cut(mean_flux, breaks = c(0,10,20,30,40,50,75,100,+Inf))) |> ungroup()
@@
 create_data_for_heatmap_bis <- function(float, df){
+
   # df <- tmp |> filter(wmo == float)
+
   # y <- unique(df$depth_bin)
+
   # if(float == 1902578){
+
   #   y <- y[1:15]
+
   # }
+
   # x <- unique(df$month)
+
   # data <- expand.grid(X=x, Y=y)
+
   # data$Z <- df$cut_flux
+
   # data$wmo <- float
+
   # data$month <- df$month
+
   df <- df |> filter(wmo == float) |> select(cut_flux, depth_bin, month)
+
   y <- unique(df$depth_bin)
+
   df <- df|> pivot_wider(names_from = depth_bin, values_from = cut_flux)
+
   x <- unique(df$month)
+
   data <- expand.grid(X=x, Y=y)
+
   data$Z <- as.vector(as.matrix(df |> select(-month)))
+
   data$wmo <- float
+
   return(data)
+
 }
@@
 d1 <- create_data_for_heatmap_bis(6904241, tmp)
+
 d2 <- create_data_for_heatmap_bis(1902578, tmp)
+
 d3 <- create_data_for_heatmap_bis(4903634, tmp)
+
 d4 <- create_data_for_heatmap_bis(6904240, tmp)
@@
 dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
+
   mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
+
   mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) |>
+
   mutate(Z = factor(Z, levels = levels(tmp$cut_flux)))
@@
 ggplot(dd) +
+
   geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Flux range', x = 'Month', y = 'Depth bin') +
+
   facet_wrap(~wmo) + scale_fill_viridis_d(option = 'A')
@@
 depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
+
 part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_256) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
+
   mutate(month = month.abb[month(juld)]) |> ungroup()
@@
 tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_256))) |>
+
   mutate(cut_part = cut(mean_part, breaks = c(seq(0,100,20),+Inf))) |> ungroup()
@@
 create_data_for_heatmap_bis <- function(float, df){
+
   df <- df |> filter(wmo == float) |> select(mean_part, depth_bin, month)
+
   y <- unique(df$depth_bin)
+
   df <- df|> pivot_wider(names_from = depth_bin, values_from = mean_part)
+
   x <- unique(df$month)
+
   data <- expand.grid(X=x, Y=y)
+
   data$Z <- as.vector(as.matrix(df |> select(-month)))
+
   data$wmo <- float
+
   return(data)
+
 }
@@
 dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
+
   mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
+
   mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) #|>
+
   #mutate(Z = factor(Z, levels = levels(tmp$cut_part)))
@@
 ggplot(dd) +
+
   geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Log10(abundance) #/L', x = 'Month', y = 'Depth bin', title = 'NP_Size_256') +
+
   facet_wrap(~wmo) + scale_fill_viridis(option = 'A')
@@
 depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
+
 part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, all_of(lpm_class)) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |> mutate(month = month.abb[month(juld)]) |> ungroup() |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration')
@@
 plot_heat_month <- function(m){
+
   # transform data for heatmap
+
   #m <- 'May'
+
   tmp2 <- tmp |> filter(month == m) |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) |> select(-c(month, mean_part))
+
   y <- unique(tmp2$depth_bin)
+
   x <- unique(tmp2$size)
+
   test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = rel_mean_conc)
+
   #test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = mean_part)
+
   data <- expand.grid(X=x, Y=y)
+
   data$Z <- as.vector(as.matrix(test |> select(-size)))
+
   data$month <- m
+
   return(data)
+
 }
@@
 m_data <- m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)),
+
                    month = factor(month, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')))
@@
 ggplot(m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)))) +
+
   geom_tile(aes(as.factor(X), Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + scale_fill_viridis(option = 'A') +
+
   facet_wrap(~month)
@@
 # Guidi 2008
+
 A <- 12.5
+
 b <- 3.81
+
 exp <- A * mid_DSE ** b
+
 estimated_flux <- rowSums(t(t(guidi_spectra)*exp))
+
 #estimated_flux
@@
 ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
+
   scale_y_reverse(limits = c(2000,0)) +
+
   scale_colour_viridis(option = 'A', limits = c(quantile(vertical_profiles$guidi_flux, probs = 0.25), quantile(vertical_profiles$guidi_flux, probs = 0.50)), oob = scales::squish, na.value = 'grey', direction = -1) +
+
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+
   labs(x = 'Time', y = 'Predicted Guidi Flux') + facet_wrap(~wmo) + theme_bw()
+
 ```
@@
 ggplot(vertical_profiles |> mutate(ratio_flux = guidi_flux/total_pred)) + geom_point(aes(x = juld, y = depth, colour = ratio_flux)) +
+
   scale_y_reverse(limits = c(2000,0)) +
+
   scale_colour_viridis(option = 'A', limits = c(1,5), oob = scales::squish, na.value = 'grey') +
+
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+
   labs(x = 'Time', y = 'Ratio Guidi/Total Model') + facet_wrap(~wmo) + theme_bw()
@@
 ```
+
 ### Modèle basé sur les pentes
@@
 # compute biovolumes for mid size bin (ESD)
+
 # DSE from 102 microns to 2.05 mm (this class to goes to 2.58 mm) (values obtained from Guidi's Matlab code converted on python)
+
 DSE <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
+
               0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+
               1.16170742, 1.46365963, 1.84409558, 2.32341484)
@@
 # DSE obtained with Rainer's code
+
 DSE_Rainer <- c(0.11645117591608793,0.14636003747903648,0.18439151433041295,
+
                 0.23251999827981656,0.2933257316240555,0.36916500651202966,0.4650399965596331,
+
                 0.5860454935035602,0.738553273391974,0.9280422447589767,1.1705684285914486,
+
                 1.4795320531800742,1.8636618038457486,2.3449457391257185)
@@
 # compute biovolume
+
 get_biovolume <- function(DSE){
+
   r <- DSE/2
+
   biovolume <- (4*pi/3)*r**3
+
   return(biovolume)
+
 }
@@
 # get biovolume for each DSE (equivalent to say for each size bin)
+
 biovolumes <- map_dbl(DSE, get_biovolume)  # unit = m^3
@@
 # spectra at parking depth
+
 uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
+
   drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
@@
 # convert abundance to biovolume
+
 spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes)))
+
 colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')
@@
 ## try model for small slopes
+
 cflux2 <- cflux |> mutate(flux = big_slope)
@@
 # training data
+
 tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))
@@
 set.seed(101)
+
 boosting_rounds <- 3000
+
 # model
+
 model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
+
 cv.folds=5,
+
 n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
+
 bag.fraction=0.5,
+
 n.cores=1
+
 )
@@
 # get best model and thus best round number
+
 best <- gbm.perf(model)
+
 #print(best)
@@
 # get features importance
+
 features_importance <- summary(model, n.trees=best, plotit = FALSE)
@@
 # mid_DSE <- c(0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
+
 # sizebin <- c(0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
+
 #              0.532318310)
@@
 mid_DSE <- c(0.0911143073, 0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
+
 sizebin <- c(0.0209587201, 0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
+
              0.532318310)
@@
 vertical_profiles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)
+
 drifting_profiles <- all_floats |> filter(PhaseName == 'PAR') |> select(juld, cycle, depth = pres, park_depth, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)
@@
   # prepare data for linear regression
+
   Y <- log(spectrum_norm)
+
   X <- log(mid_DSE)
@@
   # check for finite value
+
   h <- is.finite(Y)
+
   Y <- Y[h]
+
   X <- X[h]
@@
   model <- lm(formula = Y ~ X, data = data_slope)
+
   slope <- model$coefficients[2]
@@
   return(slope)
+
 }
@@
 # compute slope on drifting profiles
+
 drifting_spectra <- as.matrix(data.frame(drifting_profiles[,6:20]))
+
 index <- seq(from = 1, to = nrow(drifting_spectra), by = 1)
+
 slopes <- map_dbl(index, compute_slope, data_spectra = drifting_spectra)
+
 drifting_profiles$spectral_slope <- slopes
+
 ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
+
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+
   scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')
@@
 # compute slope on vertical profiles
+
 vertical_spectra <- as.matrix(data.frame(vertical_profiles[,5:19]))
+
 index <- seq(from = 1, to = nrow(vertical_spectra), by = 1)
+
 slopes <- map_dbl(index, compute_slope, data_spectra = vertical_spectra)
+
 vertical_profiles$spectral_slope <- slopes
@@
 ggplot(vertical_profiles, aes(x = juld, y = depth, colour = spectral_slope)) + geom_point() + scale_colour_viridis(option = 'H') +
+
   scale_y_reverse(limits = c(2000,0)) +
+
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+
   labs(x = 'Time', y = TeX('slope')) + facet_wrap(~wmo)
@@
 # ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF, colour = park_depth)) + geom_point() + facet_wrap(~wmo)
+
 #
+
 # ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_point() + facet_wrap(~park_depth)
+
 #
+
 # ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)
@@
 ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
+
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+
   scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')
@@
 # merge data
+
 check <- merge(tmp, tmp2)
+
 check2 <- merge(tmp, tmp2) |> filter(ratioF < 1)
@@
 # ggplot(tmp, aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)
+
 ```

### 2023-02-07 article3.qmd
update

@@
 library(gtExtras)
+library(spatialrisk)
 ```
@@
 #| label: fig-comparison-ost-uvp
-#| fig-cap: Comparison of total carbon fluxes derived from the OST (black) and the UVP6 (red) where the error bars represent the 25$^{th}$ and 75^{th}$ percentiles around the median.
+#| fig-cap: Comparison of total carbon fluxes derived from the OST (black) and the UVP6 (red) where the error bars represent the 25$^{th}$ and 75$^{th}$ percentiles around the median.
 
@@
                1.16170742, 1.46365963) # With Lionel's script
+    # mid_ESD <- c(0.29042685,
+    #            0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+    #            1.16170742, 1.46365963) # With Lionel's script
   exp <- A * mid_ESD ** b
@@
 
--   TS diagram at drifting depth + isopycnals
--   satellite chla + MLD -\> show that conditions for bloom (stable and shallow MLD) + chla happened before deployment in the area
 -   faire le modèle de pente avec les biovolues pour prédire quels types de particules explique le plus la pente -\> faire un modèle pour le small flux et un modèle pour le big flux, pas un modèle pour les small + big car les pentes ne sont pas additives
@@
 
-particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))
-
-particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
-#particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
-
-# rel_conc
-particles_bis <- particles_bis |> group_by(wmo, size) |> mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
+# particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))
+# 
+# particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+# #particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
+# 
+# # rel_conc
+# particles_bis <- particles_bis |> group_by(wmo, size) |> mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
 
@@
 
-## plot absolute concentrations
-up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
-  #group_by(size) |>
-  ggplot() + facet_grid(wmo~size) + theme_bw() +
-  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
-  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
-  #scale_colour_viridis(option = 'A', limits = c(-2,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
-  #scale_colour_viridis(option = 'H') +
-  scale_y_reverse() +
-  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-  #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
-  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') #+
-  #geom_hline(yintercept = c(600, 1100), linetype = 'dotted')
-  #theme(legend.title = element_text(angle = -90))
-  #guides(fill = guide_legend(title.theme = element_text(size = 200)))
-  #+ theme(legend.position = 'none')
-  #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
-  #theme(legend.position = 'none') #+
-  #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
-  #guides(colour = guide_legend(title.position = "bottom"))
-
-#Normalized concentrations (#/L)
-
-up_part
-
-down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
-  #group_by(size) |>
-  ggplot() + facet_grid(wmo~size) + theme_bw() +
-  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
-  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) +
-  #scale_colour_viridis(option = 'H') +
-  scale_y_reverse() +
-  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-  #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
-  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
-
-down_part
+# ## plot absolute concentrations
+# up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
+#   #group_by(size) |>
+#   ggplot() + facet_grid(wmo~size) + theme_bw() +
+#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+#   #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
+#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
+#   #scale_colour_viridis(option = 'A', limits = c(-2,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
+#   #scale_colour_viridis(option = 'H') +
+#   scale_y_reverse() +
+#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+#   #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
+#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)') #+
+#   #geom_hline(yintercept = c(600, 1100), linetype = 'dotted')
+#   #theme(legend.title = element_text(angle = -90))
+#   #guides(fill = guide_legend(title.theme = element_text(size = 200)))
+#   #+ theme(legend.position = 'none')
+#   #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
+#   #theme(legend.position = 'none') #+
+#   #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
+#   #guides(colour = guide_legend(title.position = "bottom"))
+# 
+# #Normalized concentrations (#/L)
+# 
+# up_part
+# 
+# down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
+#   #group_by(size) |>
+#   ggplot() + facet_grid(wmo~size) + theme_bw() +
+#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+#   #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
+#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) +
+#   #scale_colour_viridis(option = 'H') +
+#   scale_y_reverse() +
+#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+#   #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
+#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
+# 
+# down_part
 ```
@@
 d <- d |> group_by(date, wmo, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
-d <- d |> mutate(smoothed_chla = smooth(mean_chla))
+d <- d |> mutate(smoothed_chla = smooth(mean_chla)/2.15, smoothed_chla_up = smooth(mean_chla)/(2.15-0.37), smoothed_chla_down = smooth(mean_chla)/(2.15+0.37)) # 2.15 = Roesler factor in the Labrador Sea
 
@@
                                        mld = unique(mld),
-                                       chla_0_mld = integrate(smoothed_chla, pres, from = 0, to=mld))
+                                       chla_0_mld = integrate(smoothed_chla, pres, from = 0, to=mld),
+                                       median_chla_0_20m = median(smoothed_chla[pres<=20], na.rm=T),
+                                       median_chla_0_20m_up = median(smoothed_chla_up[pres<=20], na.rm=T),
+                                       median_chla_0_20m_down = median(smoothed_chla_down[pres<=20], na.rm=T))
 # 
@@
 
+# add couple info
+d_chla <- d_chla |> mutate(group = if_else(wmo %in% c(6904240,4903634), '6904240 - 4903634', '6904241 - 1902578'))
+
 ```
@@
 
-chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_with_clouds_sat.feather')
-#chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_without_clouds_sat.feather')
-mld_sat <- arrow::read_feather('DATA/SATELLITE/mld_sat.feather') |> mutate(lat = latitude, lon = longitude - 360) |> select(time, lat, lon, mld = mlotst)
+#chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_with_clouds_sat.feather') |> select(time, lat, lon, chla = CHL)
+chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_without_clouds_sat.feather') |> select(time, lat, lon, chla = CHL)
+mld_sat <- arrow::read_feather('DATA/SATELLITE/mld_sat.feather') |> mutate(longitude = longitude - 360) |> select(time, lat = latitude, lon = longitude, mld = mlotst)
 
 # CHLA
-x <- 1
-y <- 1
-chla_cover_deployment1 <- chla_sat |> filter(lat > 58.9 - y, lat < 58.9 + y, lon > -49.3 - x, lon < -49.3 + x) |> mutate(date = as_date(time), group = 'couple 1') |> mutate(deploy = as_date('2022-05-22'))
-chla_cover_deployment2 <- chla_sat |> filter(lat > 56.7 - y, lat < 56.7 + y, lon > -52.3 - x, lon < -52.3 + x) |> mutate(date = as_date(time), group = 'couple 2') |> mutate(deploy = as_date('2022-05-30'))
+chla_cover_deployment1 <- points_in_circle(chla_sat, lon_center = -49.2, lat_center = 58.9, radius = 100000) # 50 km around center
+chla_cover_deployment1 <- chla_cover_deployment1 |> mutate(date = as_date(time), group = '6904240 - 4903634') |> mutate(deploy = as_date('2022-05-22'))
+
+chla_cover_deployment2 <- points_in_circle(chla_sat, lon_center = -52.3, lat_center = 56.7, radius = 10000) # 50 km around center
+chla_cover_deployment2 <- chla_cover_deployment2 |> mutate(date = as_date(time), group = '6904241 - 1902578') |> mutate(deploy = as_date('2022-05-30'))
+
+# all deploy
 chla_cover <- rbind(chla_cover_deployment1, chla_cover_deployment2)
 
-ggplot(chla_cover) + geom_errorbar(aes(x = date, y = CHL),
+chla_plot <- ggplot(chla_cover |> filter(date > '2022-04-01')) + geom_errorbar(aes(x = date, y = chla),
                       stat = 'summary',
@@
   labs(x = 'Time', y = 'Chla')+ facet_wrap(~group) +
-  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "red")
+  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
+  geom_pointrange(data = d_chla |> filter(date < '2022-08-01'), aes(x = date, ymin = median_chla_0_20m_down, ymax = median_chla_0_20m_up, y = median_chla_0_20m), colour = "red", size = .25)
 
 # MLD
-mld_cover_deployment1 <- mld_sat |> filter(lat > 58.9 - y, lat < 58.9 + y, lon > -49.3 - x, lon < -49.3 + x) |> mutate(date = as_date(time), group = 'couple 1') 
-mld_cover_deployment2 <- mld_sat |> filter(lat > 56.7 - y, lat < 56.7 + y, lon > -52.3 - x, lon < -52.3 + x) |> mutate(date = as_date(time), group = 'couple 2')
+mld_cover_deployment1 <- points_in_circle(mld_sat, lon_center = -49.2, lat_center = 58.9, radius = 100000) # 50 km around center
+mld_cover_deployment1 <- mld_cover_deployment1 |> mutate(date = as_date(time), group = '6904240 - 4903634') |> mutate(deploy = as_date('2022-05-22'))
+
+mld_cover_deployment2 <- points_in_circle(mld_sat, lon_center = -52.3, lat_center = 56.7, radius = 100000) # 50 km around center
+mld_cover_deployment2 <- mld_cover_deployment2 |> mutate(date = as_date(time), group = '6904241 - 1902578') |> mutate(deploy = as_date('2022-05-30'))
+
 mld_cover <- rbind(mld_cover_deployment1, mld_cover_deployment2)
 
-ggplot(mld_cover) + geom_errorbar(aes(x = date, y = mld),
+mld_plot <- ggplot(mld_cover |> filter(date > '2022-04-01')) + geom_errorbar(aes(x = date, y = mld),
                       stat = 'summary',
@@
                       fun = median, colour = 'black', width = 5) +
-  labs(x = 'Time', y = 'MLD (m)') + facet_wrap(~group, scales = 'free_y')
+  labs(x = 'Time', y = 'MLD (m)') + facet_wrap(~group) + theme_bw() +
+  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
+  geom_point(data = d_chla |> filter(date < '2022-08-01') |> mutate(week = week(date), DOY = yday(date)) |> group_by(week, group) |> summarize(median_mld = median(mld), mean_date = as_date(mean(DOY), origin = '2022-01-01')), aes(mean_date, median_mld), colour = "red")
+
+chla_plot / mld_plot
+
+```
+
+with a radius of 100 km around both deployment positions
+
+### Sinking speed
+
+- Montrer que les sinking speed observées ne sont pas compatibles avec des valeurs de bloom (càd rapides pour les grosses particules à minima)
+
+```{r, sinking speed plot}
+
+sinking_fit <- vroom('DATA/ARGO/gaussian_fit_sinking_speed_full.csv') |> mutate(size = size_class |> str_remove('NP_Size_') |> as.numeric(), date = as_date(date))
+
+estimate_sinking_speed <- function(siz, data, threshold){
+
+  # above 500 m
+  tmp <- data |> filter(size == siz) |> filter(depth <= threshold)
+  model <- lm(formula = depth ~ time_max_abundance, data = tmp)
+  tmp$sinking_speed_above_threshold <- model$coefficients[2] # equivalent to slope of the fit
+  tmp2 <- tmp
+
+  # below 500 m
+  tmp <- data |> filter(size == siz) |> filter(depth > threshold)
+  model <- lm(formula = depth ~ time_max_abundance, data = tmp)
+  tmp2$sinking_speed_below_threshold <- model$coefficients[2] # equivalent to slope of the fit
+
+  return(tmp2)
+}
+
+# sinking speed for each float
+sinking_6904240 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 6904240), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
+sinking_6904241 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 6904241), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
+sinking_1902578 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 1902578), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
+sinking_4903634 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 4903634), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
+all_sinking_speed <- rbind(sinking_1902578, sinking_4903634, sinking_6904240, sinking_6904241)
+
+```
+
+```{r, vertical plots + sinking speed}
+
+particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))
+
+particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+#particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
+
+# rel_conc
+particles_bis <- particles_bis |> group_by(wmo, size) |> mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
+
+up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
+  #group_by(size) |>
+  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+  scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
+  geom_pointrange(data = sinking_fit |> filter(size < 512), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'white')
+
+up_part
+
+down_part <- particles_bis |> filter(depth < 2000, size > 512) |>
+  #group_by(size) |>
+  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+  scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
+  geom_pointrange(data = sinking_fit |> filter(size > 512), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'white')
+
+down_part
+
+```
+
+### TS diagram
+
+- Pour le drift à 200, 500 et 1000 m
+
+```{r, TS diagram}
+
+#ts_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth) |> drop_na(temp) |> filter(sigma > 26) |> mutate(date = as_date(date))
+
+ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth) |> drop_na(temp) |> filter(sigma > 26, park_depth != '200 m') |> mutate(date = as_date(date))
+#ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth) |> drop_na(temp) |> filter(sigma > 26) |> mutate(date = as_date(date))
+ggplot(ts_data) + geom_point(aes(psal, temp, colour = factor(park_depth))) + theme_bw()
+
+# ggplot(ts_data) + geom_point(aes(psal, temp, colour = wmo))
+
+ts_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth) |> drop_na(temp) |> filter(sigma > 26) |> mutate(date = as_date(date))
+
+# add contour
+# breaks_sigma <- seq(min(round(ts_data$sigma*2)/2), max(round(ts_data$sigma*2)/2), by = 1)
+ggplot(ts_data) + geom_point(aes(psal, temp, colour = depth)) + scale_color_viridis_c(option = 'A') + theme_bw()
 

### 2023-02-08 article3.qmd
update

@@
 
+# combine all data
 all_floats <- rbind(float_6904240, float_6904241, float_4903634, float_1902578)
 
+# add levels for some field
 all_floats <- all_floats |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
@@
 
+# summarize localization of floats
 argo_map <- all_floats |> select(WMO = wmo, cycle, lat, lon) |> drop_na(lat) |> dplyr::group_by(WMO, cycle) |> mutate(WMO = factor(WMO)) |>
@@
 
+# map positions of floats
 world_data <- map_data("world")
@@
 all_floats_slope <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
-all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle)
+all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle) |>
+   mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
+          park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
 
@@
 cflux$date <- as_date(cflux$min_time)
-
-# # add levels
-# cflux$park_depth <- factor(cflux$park_depth, levels = c('200 m', '500 m', '1000 m'))
-
 ```
@@
 
-# add level for order
-#uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
-# uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))
-
 # compute daily mean chla for each float at each cycle and at each drifting depth
@@
   #              1.16170742, 1.46365963) # With Lionel's script
-  mid_ESD <- c(0.23051195, 0.29042685,
+  # mid_ESD <- c(0.23051195, 0.29042685,
+  #              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+  #              1.16170742, 1.46365963) # With Lionel's script
+  #mid_ESD <- c(0.287555212,0.362129811,0.455929819,0.57466512,0.724144323,0.91063714,1.147083258,1.450068964)
+  mid_ESD <- c(0.29042685,
                0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
                1.16170742, 1.46365963) # With Lionel's script
-    # mid_ESD <- c(0.29042685,
-    #            0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
-    #            1.16170742, 1.46365963) # With Lionel's script
   exp <- A * mid_ESD ** b
@@
                                    park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
-parking_spectra <- parking_data |> dplyr::select(NP_Size_203:NP_Size_1290)
-
-parking_data <- parking_data |> mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) |> group_by(park_depth, cycle, wmo) #|> summarize(median_guidi_flux = median(guidi_flux))
+parking_spectra <- parking_data |> dplyr::select(NP_Size_256:NP_Size_1290)
+
+parking_data <- parking_data |> mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) |> group_by(park_depth, cycle, wmo)
 
@@
 # https://stackoverflow.com/questions/41077199/how-to-use-r-ggplot-stat-summary-to-plot-median-and-quartiles
-# ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = small_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
-#   geom_errorbar(aes(x = date, y = guidi_flux),
-#                       stat = 'summary',
-#                       fun.min = function(z) {quantile(z, 0.25)},
-#                       fun.max = function(z) {quantile(z, 0.75)},
-#                       fun = median, colour = 'red', width = 5) +
-#   geom_line(aes(x = date, y = guidi_flux),
-#                       stat = 'summary',
-#                       fun = median, colour = 'red') + scale_y_continuous(trans = 'log10') +
-#   geom_line(data = comparison_flux_ost_uvp, aes(x = date, y = big_flux), linetype = 'dotdash') + theme_bw()
-
 ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = total_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
@@
 
+```
+
+```{r, correlations between OST and UVP carbon fluxes}
+#| label: tbl-correlation
+#| tbl-cap: TODO
 correlations <- comparison_flux_ost_uvp |> select(park_depth, wmo, cycle, guidi_flux, small_flux, big_flux, date, total_flux) |> group_by(wmo, park_depth, cycle) |> mutate(median_guidi_flux = median(guidi_flux)) |> group_by(park_depth, wmo) |> summarise(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_big = round(cor(median_guidi_flux, big_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))
@@
 correlations <- correlations |> pivot_wider(names_from = park_depth, values_from = c(cor_small, cor_big, cor_total))
+
 gt(correlations) |> tab_spanner(label = "Small flux", columns = 2:4) |> tab_spanner(label = "Big flux", columns = 5:7) |> tab_spanner(label = 'Total flux', columns = 8:10) |>
@@
 
-TODO :
-
--   faire le modèle de pente avec les biovolues pour prédire quels types de particules explique le plus la pente -\> faire un modèle pour le small flux et un modèle pour le big flux, pas un modèle pour les small + big car les pentes ne sont pas additives
--   discuter des corrélations
--   science: on voit le système latent, on a raté le bloom
--   rajouter les vitesses de sédimentations (n'importe quelle méthode fera l'affaire) -> permet d'utiliser le graphique de l'UVP en vertical -> hop une justification en plus
--   rajouter dans les méthodes les produits satellites qu'on utilise
--   différentes dynamiques post-bloom
--   phénologie du bloom printanier
-
-### Which particles drive the flux
-
-```{r, flux model 2, include = FALSE}
-
-# get spectra
+Figure @fig-comparison-ost-uvp shows a comparison of carbon fluxes estimated with the OST ($F_{total} = F_{small} + F_{big}$) and the UVP6 (median of all UVP measurements while the float was drifting) at all drifting depths.
+Overall, carbon fluxes estimated with the coefficients ($A$, $b$) of @Guidi2008-vp, are lower than the ones measured by the OST.
+Without additional measurements (e.g.
+co-deployed sediment traps), it is however impossible to decipher which measurements is closest to the truth.
+Nonetheless, the trends measured by both instruments are very well correlated at all drifting depths for floats 6904240 and 6904241 with correlations comprised between `r min(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])` and `r max(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])`.
+@tbl-correlation also shows that the correlation between UVP-derived carbon fluxes and $F_{small}$ and $F_{big}$ is `r mean(as.vector(as.matrix(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,2:7])))` on average.
+We calculated lower correlations with the other two floats, especially the float 1902578 at 200 m, where correlations between the UVP-derived carbon fluxes and all fluxes (small, big and total) are close to 0.
+
+Floats equipped with both a transmissometer and a UVP6 represent an opportunity to look at the influence of parts of the size spectrum on $F_{small}$ and $F_{big}$.
+For this, we decided to build a regression model using XGBoost, a gradient boosting method [REF], using biovolumes measured by the UVP to predict the slope of $F_{small}$ and $F_{big}$ measured by the OST.
+
+```{r, model based on slope}
+
+# compute biovolumes for mid size bin (ESD)
+# DSE from 102 microns to 2.05 mm (this class to goes to 2.58 mm) (values obtained from Guidi's Matlab code converted on python)
+
+# DSE <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
+#               0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+#               1.16170742, 1.46365963, 1.84409558, 2.32341484)
+# 
+# # DSE obtained with Rainer's code
+# DSE_Rainer <- c(0.11645117591608793,0.14636003747903648,0.18439151433041295,
+#                 0.23251999827981656,0.2933257316240555,0.36916500651202966,0.4650399965596331,
+#                 0.5860454935035602,0.738553273391974,0.9280422447589767,1.1705684285914486,
+#                 1.4795320531800742,1.8636618038457486,2.3449457391257185)
+
+DSE <- c(0.29042685,0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779, 1.16170742, 1.46365963)
+
+# compute biovolume
+get_biovolume <- function(DSE){
+  r <- DSE/2
+  biovolume <- (4*pi/3)*r**3
+  return(biovolume)
+}
+
+# get biovolume for each DSE (equivalent to say for each size bin)
+biovolumes <- map_dbl(DSE, get_biovolume)  # unit = m^3
+
+# spectra at parking depth
 uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
@@
 
-flux_model <- function(boosting_rounds, which_flux){
+# convert abundance to biovolume
+spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes)))
+colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')
+
+# add data to initial abundance spectra
+uvp_spectra <- cbind(uvp_spectra, spectra_biovolume)
+
+# recompute slope based on Estapa's formula
+cflux <- cflux |> mutate(small_slope = (small_flux/633)**(1/0.77), big_slope = (big_flux/633)**(1/0.77))
+
+## build model
+flux_model <- function(boosting_rounds, which_slope){
 
   # select which flux to build the model
-    if(which_flux == 'small'){
-      cflux2 <- cflux |> mutate(flux = small_flux) 
-    }else if(which_flux == 'big'){
-      cflux2 <- cflux |> mutate(flux = big_flux)
+  if(which_slope == 'small'){
+    cflux2 <- cflux |> mutate(flux = log(small_slope)) |> filter(is.finite(flux))
   }else{
-      cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
+    cflux2 <- cflux |> mutate(flux = log(big_slope)) |> filter(is.finite(flux)) # because log(0) creates non finite value -> bug
   }
@@
   # training data
-  tmp <- data_model |> select(flux, all_of(lpm_class))
+  tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))
 
@@
   return(list(features_importance = features_importance, model = model, best_tree = best))
-
 }
@@
 # big_model <- flux_model(3000, 'big')
-# total_model <- flux_model(3000, 'total')
-# 
-# save(small_model, big_model, total_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data')
-
-#load(file = 'DATA/MODEL/model_xgboost_data')
-
-```
-
-#### On profiling mode
-
-Cette partie devrait venir dans la discussion si on part du principe que la première partie est liée à la méthodologie
-
-@fig-uvp-vertical-profiles shows that all floats measure similar patterns in the vertical profiles of particles concentrations in all size classes.
-In the 102-512 $\mu$m range, all 4 floats measured a local maximum between 650 m and 1050 m.
-Above 512 $\mu$m, the relative abundance is maximum along a depth track centered on August.
-It can be seen among all size classes that the water column is being drained proportionally to the class size.
-
-```{r, uvp_data on the vertical}
-#| label: fig-uvp-vertical-profiles
-#| fig-cap: Normalized concentrations of particles for all UVP6 vertical profiles.
-Plots were arranged to respect the deployment of each couple of floats (e.g.
-the first two rows of plots are related to co-deployed floats 6904240 and 4903634).
-The colorbar depicts the log$_{10}$ of normalized concentrations.
-Grey areas indicate that particles were not measured in the size range of interest.
-
-# particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))
-# 
-# particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
-# #particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
-# 
-# # rel_conc
-# particles_bis <- particles_bis |> group_by(wmo, size) |> mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
-
-# order floats by couple of deployment
-# particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904240,4903634,6904241,1902578))
-
-# stats on the local maximum
-# test <- particles_bis |> filter(depth > 500) |> group_by(wmo, cycle, size) |> # find depth of maximum relative abundance below 500 m between June and September
-#   summarize(index_max = which.max(rel_conc), max_depth = depth[index_max], date = unique(juld)) |> 
-#   filter(size < 512, date < '2022-10-01', date > '2022-05-31') |> # compute mean depth for each class size for each float
-#   group_by(wmo, size) |> summarize(mean_depth = mean(max_depth))
-
-# plot concentration
-# particles_bis |> filter(depth < 200) |>
-#   ggplot() + facet_grid(wmo~size) + theme_bw() +
-#   geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
-#   #scale_colour_viridis(option = 'H', limits = c(-1,1), oob = scales::squish, na.value = 'transparent') +
-#   scale_colour_viridis(option = 'H') +
-#   scale_y_reverse() +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
-#   labs(x = 'Month', y = 'Depth (m)') +
-#   geom_path(data = particles_bis, aes(juld, MLD), colour = "white")
-
-# ## plot absolute concentrations
-# up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
-#   #group_by(size) |>
-#   ggplot() + facet_grid(wmo~size) + theme_bw() +
-#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-#   #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
-#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
-#   #scale_colour_viridis(option = 'A', limits = c(-2,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
-#   #scale_colour_viridis(option = 'H') +
-#   scale_y_reverse() +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-#   #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
-#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)') #+
-#   #geom_hline(yintercept = c(600, 1100), linetype = 'dotted')
-#   #theme(legend.title = element_text(angle = -90))
-#   #guides(fill = guide_legend(title.theme = element_text(size = 200)))
-#   #+ theme(legend.position = 'none')
-#   #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
-#   #theme(legend.position = 'none') #+
-#   #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
-#   #guides(colour = guide_legend(title.position = "bottom"))
-# 
-# #Normalized concentrations (#/L)
-# 
-# up_part
-# 
-# down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
-#   #group_by(size) |>
-#   ggplot() + facet_grid(wmo~size) + theme_bw() +
-#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-#   #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
-#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) +
-#   #scale_colour_viridis(option = 'H') +
-#   scale_y_reverse() +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-#   #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
-#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
-# 
-# down_part
-```
+# save(small_model, big_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data_slope')
+
+load(file = 'DATA/MODEL/model_xgboost_data_slope')
+
+# # get spectra
+# uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
+#   drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
+# 
+# flux_model <- function(boosting_rounds, which_flux){
+#     
+#     # select which flux to build the model
+#     if(which_flux == 'small'){
+#       cflux2 <- cflux |> mutate(flux = small_flux) 
+#     }else if(which_flux == 'big'){
+#       cflux2 <- cflux |> mutate(flux = big_flux)
+#     }else{
+#       cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
+#     }
+#     
+#     # merge data for training
+#     data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
+#     
+#   # training data
+#   tmp <- data_model |> select(flux, all_of(lpm_class))
+#   
+#   set.seed(101)
+#   # model
+#   model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
+#   cv.folds=5,
+#   n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
+#   bag.fraction=0.5,
+#   n.cores=1
+#   )
+#    
+#   # get best model and thus best round number
+#   best <- gbm.perf(model)
+#   #print(best)
+#   
+#   # get features importance
+#   features_importance <- summary(model, n.trees=best, plotit = FALSE)
+#   
+#   
+#   return(list(features_importance = features_importance, model = model, best_tree = best))
+#    
+# }
+#   
+# # small_model <- flux_model(3000, 'small')
+# # big_model <- flux_model(3000, 'big')
+# # total_model <- flux_model(3000, 'total')
+# # 
+# # save(small_model, big_model, total_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data') 
+#   
+# load(file = 'DATA/MODEL/model_xgboost_data')
+
+```
+
+```{r, model predictions}
+
+model_predict <- function(which_slope, mod, best_tree_number){
+
+    # get spectra
+    uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
+  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
+
+    # convert abundance to biovolume
+    spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes)))
+    colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')
+
+    # add data to initial abundance spectra
+    uvp_spectra <- cbind(uvp_spectra, spectra_biovolume)
+
+    # select which flux to build the model
+    if(which_slope == 'small'){
+      cflux2 <- cflux |> mutate(flux = small_slope)
+    }else{
+      cflux2 <- cflux |> mutate(flux = big_slope)
+    }
+
+    # merge data for training
+    data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
+
+    # # training data
+    # tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))
+
+    # predictions on mean spectra values
+    uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = log(unique(flux)),
+                                                                                   NP_Biovolume_102 = mean(NP_Biovolume_102),
+                                                                                   NP_Biovolume_128 = mean(NP_Biovolume_128),
+                                                                                   NP_Biovolume_161 = mean(NP_Biovolume_161),
+                                                                                   NP_Biovolume_203 = mean(NP_Biovolume_203),
+                                                                                   NP_Biovolume_256 = mean(NP_Biovolume_256),
+                                                                                   NP_Biovolume_323 = mean(NP_Biovolume_323),
+                                                                                   NP_Biovolume_406 = mean(NP_Biovolume_406),
+                                                                                   NP_Biovolume_512 = mean(NP_Biovolume_512),
+                                                                                   NP_Biovolume_645 = mean(NP_Biovolume_645),
+                                                                                   NP_Biovolume_813 = mean(NP_Biovolume_813),
+                                                                                   NP_Biovolume_1020 = mean(NP_Biovolume_1020),
+                                                                                   NP_Biovolume_1290 = mean(NP_Biovolume_1290),
+                                                                                   NP_Biovolume_1630 = mean(NP_Biovolume_1630),
+                                                                                   NP_Biovolume_2050 = mean(NP_Biovolume_2050)) |> ungroup()
+
+  # prediction
+  pred_flux <- predict(mod, newdata = uvp_spectra_average, n.trees = best_tree_number)
+
+  uvp_spectra_average$pred_flux <- pred_flux
+
+  return(uvp_spectra_average)
+}
+
+# predictions
+# pred_small <- model_predict('small', small_model$model, small_model$best_tree) |> select(wmo, cycle, depth, true_small = true_flux, pred_small = pred_flux)
+# pred_big <- model_predict('big', big_model$model, big_model$best_tree) |> select(true_big = true_flux, pred_big = pred_flux) 
+# all_predictions <- cbind(pred_small, pred_big)
+# vroom::vroom_write(all_predictions, file = '/home/flo/dox/THESIS/DATA/MODEL/all_predictions_slopes.csv')
+all_predictions <- vroom('DATA/MODEL/all_predictions_slopes.csv')
+
+# model_predict <- function(which_flux, mod, best_tree_number){
+# 
+#     # get spectra
+#     uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
+#   drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
+# 
+#     # select which flux to build the model
+#     if(which_flux == 'small'){
+#       cflux2 <- cflux |> mutate(flux = small_flux)
+#     }else if(which_flux == 'big'){
+#       cflux2 <- cflux |> mutate(flux = big_flux)
+#     }else{
+#       cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
+#     }
+# 
+#     # merge data for training
+#     data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
+# 
+#   # predictions on mean spectra values
+#   uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = unique(flux),
+#                                                                                    NP_Size_102 = mean(NP_Size_102),
+#                                                                                    NP_Size_128 = mean(NP_Size_128),
+#                                                                                    NP_Size_161 = mean(NP_Size_161),
+#                                                                                    NP_Size_203 = mean(NP_Size_203),
+#                                                                                    NP_Size_256 = mean(NP_Size_256),
+#                                                                                    NP_Size_323 = mean(NP_Size_323),
+#                                                                                    NP_Size_406 = mean(NP_Size_406),
+#                                                                                    NP_Size_512 = mean(NP_Size_512),
+#                                                                                    NP_Size_645 = mean(NP_Size_645),
+#                                                                                    NP_Size_813 = mean(NP_Size_813),
+#                                                                                    NP_Size_1020 = mean(NP_Size_1020),
+#                                                                                    NP_Size_1290 = mean(NP_Size_1290),
+#                                                                                    NP_Size_1630 = mean(NP_Size_1630),
+#                                                                                    NP_Size_2050 = mean(NP_Size_2050)) |> ungroup()
+# 
+# 
+#   # prediction at 1000 m
+#   #spectra <- uvp_spectra_average |> select(true_flux, all_of(lpm_class[]))
+# 
+#   # prediction
+#   pred_flux <- predict(mod, newdata = uvp_spectra_average, n.trees = best_tree_number)
+# 
+#   uvp_spectra_average$pred_flux <- pred_flux
+# 
+#   return(uvp_spectra_average)
+# }
+# 
+# # pred_small <- model_predict('small', small_model$model, small_model$best_tree) |> select(wmo, cycle, depth, true_small = true_flux, pred_small = pred_flux)
+# # pred_big <- model_predict('big', big_model$model, big_model$best_tree) |> select(true_big = true_flux, pred_big = pred_flux)
+# # pred_total <- model_predict('total', total_model$model, total_model$best_tree) |> select(true_total = true_flux, pred_total = pred_flux)
+# #
+# # all_predictions <- cbind(pred_small, pred_big, pred_total) |> mutate(pred_total2 = pred_small + pred_big, anom = pred_total - pred_total2)
+# #
+# # vroom::vroom_write(all_predictions, file = '/home/flo/dox/THESIS/DATA/MODEL/all_predictions.csv')
+# all_predictions <- vroom('DATA/MODEL/all_predictions.csv')
+
+```
+
+```{r, plot model predictions}
+
+p_small <- ggplot(all_predictions) + geom_point(aes(true_small, pred_small)) + theme_bw() 
+round(R2_Score(y_true = all_predictions$true_small, y_pred = all_predictions$pred_small), 2)
+
+p_big <- ggplot(all_predictions |> filter(is.finite(true_big))) + geom_point(aes(true_big, pred_big)) + theme_bw()
+all_predictions |> filter(is.finite(true_big)) |> summarize(round(R2_Score(true_big, pred_big)), 2)
+
+p_small / p_big
+
+```
+
+TODO :
+
+-   science: on voit le système latent, on a raté le bloom
+-   rajouter les vitesses de sédimentations (n'importe quelle méthode fera l'affaire) -\> permet d'utiliser le graphique de l'UVP en vertical -\> hop une justification en plus
+-   rajouter dans les méthodes les produits satellites qu'on utilise
+-   différentes dynamiques post-bloom
+-   phénologie du bloom printanier
+-
 
@@
 
- + aller chercher les produits satellitaires pour montrer que l'on a bien raté le bloom ou que les conditions du bloom n'étaient pas réunies
+-   aller chercher les produits satellitaires pour montrer que l'on a bien raté le bloom ou que les conditions du bloom n'étaient pas réunies
 
@@
 ```
+```{r, sinking speed values}
+
+gt(all_sinking_speed)
+
+```
 
@@
 
+J'ai enlevé le truc du spectre inversé à 500 et 1000 m car difficile de conclure -> mais comme on est dans la meme masse d'eau, on pourrait le garder, why not
+
 ```{r, monthly anomaly between drifting at 500 and 1000 m}
@@
 
-tmp <- uvp_data |> filter(park_depth != '200 m') |> group_by(wmo, size, month, park_depth) |> 
-  summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
-  # compute concentration anomaly at 500 and 1000 m
-  mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
-  select(wmo, size, month, diff_anomaly, div_anomaly) |> mutate(month = month.abb[month]) |> 
-  mutate(size = as.character(size)) |>
-  mutate(rev_div_anomaly = 1/div_anomaly) |> 
-  #filter(is.finite(rev_div_anomaly)) |>
-  mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1), div_anomaly = round(div_anomaly, 1)) |>
-  ungroup()
+# tmp <- uvp_data |> filter(park_depth != '200 m') |> group_by(wmo, size, month, park_depth) |> 
+#   summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
+#   # compute concentration anomaly at 500 and 1000 m
+#   mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
+#   select(wmo, size, month, diff_anomaly, div_anomaly) |> mutate(month = month.abb[month]) |> 
+#   mutate(size = as.character(size)) |>
+#   mutate(rev_div_anomaly = 1/div_anomaly) |> 
+#   #filter(is.finite(rev_div_anomaly)) |>
+#   mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1), div_anomaly = round(div_anomaly, 1)) |>
+#   ungroup()
 
@@
 
-# plot_heatmap <- function(float){
-#   y <- unique(tmp$size)[1:12]
-#   tt <- tmp |> filter(wmo == float, size %in% y) 
-#   x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')
+```
+
+```{r, test gt table}
+
+monthly_mean <- uvp_data |> mutate(month = month.abb[month(juld)]) |> group_by(size, park_depth, month) |> summarize(mean_conc = round(mean(conc, na.rm=T), 2)) |> ungroup()
+
+# pivot wider
+monthly_mean <- monthly_mean |> pivot_wider(names_from = month, values_from = mean_conc) |> select(size, depth = park_depth, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec, Jan, Feb)
+
+monthly_mean |> group_by(size) |> gt(rowname_col = 'depth')
+
+#monthly_mean |> gt(rowname_col = c('size')) |> tab_stubhead(label = 'Class')
+
+```
+
+#### Features importance
+
+```{r, features importance}
+# 
+# features <- small_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'small') |>
+#   bind_rows(big_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'big')) |>
+#   bind_rows(total_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'total'))
+# 
+# rownames(features) <- NULL
+# 
+# features <- features |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+# 
+# ggplot(features) + geom_point(aes(x = as.factor(size), y = importance, colour = model), size = 5) + labs(x = 'Size', y = 'Features importance')
+
+```
+
+#### Predictions
+
+```{r, plot predictions}
+
+# p_small <- ggplot(all_predictions) + geom_point(aes(true_small, pred_small)) + xlim(c(0,25)) + ylim(c(0,25)) + theme_bw() +
+# 
+#   annotate("text", x = 3, y = 23, label = paste0('R² = ' , round(R2_Score(all_predictions$true_small, all_predictions$pred_small), 2)))
+# 
+# p_big <- ggplot(all_predictions) + geom_point(aes(true_big, pred_big)) + xlim(c(0,50)) + ylim(c(0,50)) + theme_bw() +
+# 
+#   annotate("text", x = 5, y = 45, label = paste0('R² = ' , round(R2_Score(all_predictions$true_big, all_predictions$pred_big), 2)))
+# 
+# p_total <- ggplot(all_predictions) + geom_point(aes(true_total, pred_total)) + xlim(c(0,100)) + ylim(c(0,100)) + theme_bw() +
+# 
+#   annotate("text", x = 3, y = 80, label = paste0('R² = ' , round(R2_Score(all_predictions$true_total, all_predictions$pred_total), 2)))
+# 
+# (p_small + p_big) / p_total
+
+# plot(test$true_flux, y=test$pred_flux, asp=1)
+
+# print(R2_Score(test$pred_flux, test$true_flux))
+
+```
+
+#### same plot mais pour les classes de tailles
+
+```{r, TEST }
+
+# depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
+# 
+# part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_256) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
+# 
+#   mutate(month = month.abb[month(juld)]) |> ungroup()
+# 
+# tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_256))) |>
+# 
+#   mutate(cut_part = cut(mean_part, breaks = c(seq(0,100,20),+Inf))) |> ungroup()
+# 
+# create_data_for_heatmap_bis <- function(float, df){
+# 
+#   df <- df |> filter(wmo == float) |> select(mean_part, depth_bin, month)
+# 
+#   y <- unique(df$depth_bin)
+# 
+#   df <- df|> pivot_wider(names_from = depth_bin, values_from = mean_part)
+# 
+#   x <- unique(df$month)
+# 
 #   data <- expand.grid(X=x, Y=y)
-#   #data$Z <- tt$anom
-#   data$Z <- tt$diff_anomaly
-#   data$Z <- tt$rev_div_anomaly
-# 
-#   p <- ggplot(data, aes(X, Y, fill= Z)) + 
-#     geom_tile() + scale_fill_viridis() +
-#     geom_text(data = data, aes(X, Y, label = Z)) +
-#     labs(title = float)
-#   
-#   return(p)
+# 
+#   data$Z <- as.vector(as.matrix(df |> select(-month)))
+# 
+#   data$wmo <- float
+# 
+#   return(data)
+# 
 # }
 # 
-# p1 <- plot_heatmap(4903634)
-# p2 <- plot_heatmap(6904240)
-# p3 <- plot_heatmap(1902578)
-# p4 <- plot_heatmap(6904241)
-# 
-# (p1 | p2) / 
-# (p3 | p4)
-
-# # https://kcuilla.github.io/reactablefmtr/articles/reactablefmtr_cookbook.html
-# tt <- tt |> pivot_wider(names_from = month, values_from = anom) |> mutate(size = as.character(size))
-# reactable(
-#   tt,
-#   pagination = FALSE,
-#   showSortIcon = FALSE,
-#   theme = void(
-#     centered = TRUE, 
-#     cell_padding = 0,
-#     header_font_color = 'black',
-#     font_color = 'black'
-#     ),
-#   defaultColDef = colDef(
-#     maxWidth = 50,
-#     align = 'center',
-#     cell = tooltip(),
-#     style = color_scales(
-#       data = tt,
-#       span = 2:8,
-#       colors = c('#002347','#003366','#003F7D','#FF8E00','#FD7702','#FF5003'),
-#       bias = 1.4,
-#       opacity = 0.9,
-#       show_text = FALSE
-#     )
-#   ),
-#   columns = list(
-#     total = colDef(
-#       maxWidth = 225,
-#       cell = data_bars(
-#         data = tt,
-#         fill_color = c('#002347','#003366','#003F7D','#FF8E00','#FD7702','#FF5003'),
-#         bias = 1.4,
-#         fill_opacity = 0.9,
-#         background = 'transparent',
-#         bar_height = 40,
-#         text_position = 'center'
-#         ),
-#       style = list(borderLeft = "2px solid #999999")
-#     )
-#   )
-# )
-
-```
-
-```{r, test gt table}
-
-monthly_mean <- uvp_data |> mutate(month = month.abb[month(juld)]) |> group_by(size, park_depth, month) |> summarize(mean_conc = round(mean(conc, na.rm=T), 2)) |> ungroup()
-
-# pivot wider
-monthly_mean <- monthly_mean |> pivot_wider(names_from = month, values_from = mean_conc) |> select(size, depth = park_depth, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec, Jan)
-
-monthly_mean |> group_by(size) |> gt(rowname_col = 'depth')
-
-#monthly_mean |> gt(rowname_col = c('size')) |> tab_stubhead(label = 'Class')
-
-```
-
-### Chlorophyll data
-
-```{r, chla data}
-
-# argo data
-npq_data <- vroom('DATA/ARGO/npq_corrected_data.csv')
-
-# clean and smoothed chla data
-d <- npq_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> mutate(date = as_date(juld))
-d <- d |> group_by(date, wmo, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
-d <- d |> mutate(smoothed_chla = smooth(mean_chla))
-
-d$wmo <- factor(d$wmo, levels = c(6904240,4903634,6904241,1902578))
-
-# plot it
-ggplot(d) + geom_point(aes(x = date, y = pres, colour = log10(smoothed_chla))) +
-  scale_y_reverse(limits = c(200,0)) +
-  #scale_colour_viridis(option = 'A') +
-  scale_colour_cmocean(name = 'algae', oob = scales::squish, limits = c(-1,1), na.value = 'grey') +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = 'NPQ-corrected Chla') + facet_wrap(~wmo) + theme_bw()
-
-# compute chla integration in the MLD and in the 0-100 m zone
-d_chla <- d |> group_by(wmo, date) |> summarize(chla_0_100m = integrate(smoothed_chla, pres, from=0, to=100),
-                                       mld = unique(mld),
-                                       chla_0_mld = integrate(smoothed_chla, pres, from = 0, to=mld))
-
-ggplot(d_chla) + geom_point(aes(date, chla_0_100m)) + facet_wrap(~wmo, scales = 'free') + theme_bw() + labs(x = 'Month', y = 'Integrated Chla 0-100 m (CHLA UNITS)')
-#ggplot(d_chla) + geom_point(aes(date, chla_0_mld)) + facet_wrap(~wmo, scales = 'free')
-
-```
-
-### Mixed layer depth
-
-```{r, mld data}
-ggplot(d_chla) + geom_point(aes(date, mld)) + facet_wrap(~wmo, scales = 'free') + theme_bw() + labs(x = 'Month', y = 'Mixed layer depth (m)')
-```
-
-### Flux model(s)
-
-```{r, flux model, include = FALSE}
-
-# get spectra
-uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
-  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
-
-flux_model <- function(boosting_rounds, which_flux){
-
-    # select which flux to build the model
-    if(which_flux == 'small'){
-      cflux2 <- cflux |> mutate(flux = small_flux) 
-    }else if(which_flux == 'big'){
-      cflux2 <- cflux |> mutate(flux = big_flux)
-    }else{
-      cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
-    }
-
-    # merge data for training
-    data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
-
-  # training data
-  tmp <- data_model |> select(flux, all_of(lpm_class))
-
-  set.seed(101)
-  # model
-  model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
-  cv.folds=5,
-  n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
-  bag.fraction=0.5,
-  n.cores=1
-  )
-
-  # get best model and thus best round number
-  best <- gbm.perf(model)
-  #print(best)
-
-  # get features importance
-  features_importance <- summary(model, n.trees=best, plotit = FALSE)
-
-  return(list(features_importance = features_importance, model = model, best_tree = best))
-
-}
-
-# small_model <- flux_model(3000, 'small')
-# big_model <- flux_model(3000, 'big')
-# total_model <- flux_model(3000, 'total')
-# 
-# save(small_model, big_model, total_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data')
-
-load(file = 'DATA/MODEL/model_xgboost_data')
-
-```
-
-```{r, model predictions}
-
-model_predict <- function(which_flux, mod, best_tree_number){
-
-    # get spectra
-    uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
-  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
-
-    # select which flux to build the model
-    if(which_flux == 'small'){
-      cflux2 <- cflux |> mutate(flux = small_flux) 
-    }else if(which_flux == 'big'){
-      cflux2 <- cflux |> mutate(flux = big_flux)
-    }else{
-      cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
-    }
-
-    # merge data for training
-    data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
-
-  # predictions on mean spectra values
-  uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = unique(flux),
-                                                                                   NP_Size_102 = mean(NP_Size_102),
-                                                                                   NP_Size_128 = mean(NP_Size_128),
-                                                                                   NP_Size_161 = mean(NP_Size_161),
-                                                                                   NP_Size_203 = mean(NP_Size_203),
-                                                                                   NP_Size_256 = mean(NP_Size_256),
-                                                                                   NP_Size_323 = mean(NP_Size_323),
-                                                                                   NP_Size_406 = mean(NP_Size_406),
-                                                                                   NP_Size_512 = mean(NP_Size_512),
-                                                                                   NP_Size_645 = mean(NP_Size_645),
-                                                                                   NP_Size_813 = mean(NP_Size_813),
-                                                                                   NP_Size_1020 = mean(NP_Size_1020),
-                                                                                   NP_Size_1290 = mean(NP_Size_1290),
-                                                                                   NP_Size_1630 = mean(NP_Size_1630),
-                                                                                   NP_Size_2050 = mean(NP_Size_2050)) |> ungroup()
-
-  # prediction at 1000 m
-  #spectra <- uvp_spectra_average |> select(true_flux, all_of(lpm_class[]))
-
-  # prediction 
-  pred_flux <- predict(mod, newdata = uvp_spectra_average, n.trees = best_tree_number)
-
-  uvp_spectra_average$pred_flux <- pred_flux
-
-  return(uvp_spectra_average)
-}
-
-# pred_small <- model_predict('small', small_model$model, small_model$best_tree) |> select(wmo, cycle, depth, true_small = true_flux, pred_small = pred_flux)
-# pred_big <- model_predict('big', big_model$model, big_model$best_tree) |> select(true_big = true_flux, pred_big = pred_flux)
-# pred_total <- model_predict('total', total_model$model, total_model$best_tree) |> select(true_total = true_flux, pred_total = pred_flux)
-# 
-# all_predictions <- cbind(pred_small, pred_big, pred_total) |> mutate(pred_total2 = pred_small + pred_big, anom = pred_total - pred_total2)
-# 
-# vroom::vroom_write(all_predictions, file = '/home/flo/dox/THESIS/DATA/MODEL/all_predictions.csv')
-all_predictions <- vroom('DATA/MODEL/all_predictions.csv')
-
-```
-
-#### Features importance
-
-```{r, features importance}
-
-features <- small_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'small') |>
-  bind_rows(big_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'big')) |>
-  bind_rows(total_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'total'))
-
-rownames(features) <- NULL
-
-features <- features |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
-
-ggplot(features) + geom_point(aes(x = as.factor(size), y = importance, colour = model), size = 5) + labs(x = 'Size', y = 'Features importance')
-
-```
-
-#### Predictions
-
-```{r, plot predictions}
-
-p_small <- ggplot(all_predictions) + geom_point(aes(true_small, pred_small)) + xlim(c(0,25)) + ylim(c(0,25)) + theme_bw() +
-
-  annotate("text", x = 3, y = 23, label = paste0('R² = ' , round(R2_Score(all_predictions$true_small, all_predictions$pred_small), 2)))
-
-p_big <- ggplot(all_predictions) + geom_point(aes(true_big, pred_big)) + xlim(c(0,50)) + ylim(c(0,50)) + theme_bw() +
-
-  annotate("text", x = 5, y = 45, label = paste0('R² = ' , round(R2_Score(all_predictions$true_big, all_predictions$pred_big), 2)))
-
-p_total <- ggplot(all_predictions) + geom_point(aes(true_total, pred_total)) + xlim(c(0,100)) + ylim(c(0,100)) + theme_bw() +
-
-  annotate("text", x = 3, y = 80, label = paste0('R² = ' , round(R2_Score(all_predictions$true_total, all_predictions$pred_total), 2)))
-
-(p_small + p_big) / p_total
-
-# plot(test$true_flux, y=test$pred_flux, asp=1)
-
-# print(R2_Score(test$pred_flux, test$true_flux))
-
-```
-
-#### Apply the total flux model on all vertical profiles
-
-```{r, apply model to vertical profile}
-
-vertical_profiles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))
-
-# make predictions
-
-small_predicted_flux <- predict(small_model$model, newdata = vertical_profiles, n.trees = small_model$best_tree)
-
-big_predicted_flux <- predict(big_model$model, newdata = vertical_profiles, n.trees = big_model$best_tree)
-
-total_predicted_flux <- predict(total_model$model, newdata = vertical_profiles, n.trees = total_model$best_tree)
-
-# add predictions to vertical data
-
-vertical_profiles <- vertical_profiles |> mutate(small_pred = small_predicted_flux, big_pred = big_predicted_flux, total_pred = total_predicted_flux)
-
-vertical_profiles$wmo <- factor(vertical_profiles$wmo, levels = c(6904240,4903634,6904241,1902578))
-
-# pv_small <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(small_pred))) + geom_point() + scale_color_viridis(option = 'H') +
-
-#   scale_y_reverse(limits = c(500,0)) +
-
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-
-#   labs(x = 'Time', y = 'Predicted Fsmall') + facet_wrap(~wmo)
-
-#
-
-# pv_big <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(big_pred))) + geom_point() + scale_color_viridis(option = 'H') +
-
-#   scale_y_reverse(limits = c(500,0)) +
-
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-
-#   labs(x = 'Time', y = 'Predicted Fbig') + facet_wrap(~wmo)
-
-pv_total <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
-
-  scale_y_reverse(limits = c(2000,0)) +
-
-  scale_colour_viridis(option = 'A', limits = c(quantile(all_predictions$true_total, probs = 0.25), quantile(all_predictions$true_total, probs = 0.75)), oob = scales::squish, na.value = 'grey', direction = -1) +
-
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-
-  labs(x = 'Time', y = 'Predicted Ftotal') + facet_wrap(~wmo) + theme_bw()
-
-pv_total
-
-ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
-
-  scale_y_reverse(limits = c(500,0)) +
-
-  scale_colour_viridis(option = 'A', limits = c(20,100), oob = scales::squish, na.value = 'grey', direction = -1) +
-
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-
-  labs(x = 'Time', y = 'Predicted Ftotal') + facet_wrap(~wmo) + theme_bw()
-
-```
-
-```{r, flux distribution}
-
-depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
-
-flux_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, total_pred) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
-
-  mutate(month = month.abb[month(juld)]) |> ungroup() #|> filter(month != 'May') # I filtered May because I had some troubles..
-I need to fill some cases unseen by the float (eg.
-when they start profiling until 1000 m then go to 2000 -> ça fout un peu la merde..)
-
-tmp <- flux_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_flux = mean(total_pred)) |>
-
-  mutate(cut_flux = cut(mean_flux, breaks = c(0,10,20,30,40,50,75,100,+Inf))) |> ungroup()
-
-create_data_for_heatmap_bis <- function(float, df){
-
-  # df <- tmp |> filter(wmo == float)
-
-  # y <- unique(df$depth_bin)
-
-  # if(float == 1902578){
-
-  #   y <- y[1:15]
-
+# d1 <- create_data_for_heatmap_bis(6904241, tmp)
+# 
+# d2 <- create_data_for_heatmap_bis(1902578, tmp)
+# 
+# d3 <- create_data_for_heatmap_bis(4903634, tmp)
+# 
+# d4 <- create_data_for_heatmap_bis(6904240, tmp)
+# 
+# dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
+# 
+#   mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
+# 
+#   mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) #|>
+# 
+#   #mutate(Z = factor(Z, levels = levels(tmp$cut_part)))
+# 
+# ggplot(dd) +
+# 
+#   geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Log10(abundance) #/L', x = 'Month', y = 'Depth bin', title = 'NP_Size_256') +
+# 
+#   facet_wrap(~wmo) + scale_fill_viridis(option = 'A')
+
+```
+
+#### same plots mais par mois
+
+```{r, monthly heatmap}
+
+# depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
+# 
+# part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, all_of(lpm_class)) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |> mutate(month = month.abb[month(juld)]) |> ungroup() |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration')
+# 
+# tmp <- part_distrib |> group_by(depth_bin, size, month) |> summarize(mean_part = mean(concentration)) |> mutate(rel_mean_conc = (mean_part-min(mean_part))/(max(mean_part) - min(mean_part))) |> ungroup()
+# 
+# plot_heat_month <- function(m){
+# 
+#   # transform data for heatmap
+# 
+#   #m <- 'May'
+# 
+#   tmp2 <- tmp |> filter(month == m) |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) |> select(-c(month, mean_part))
+# 
+#   y <- unique(tmp2$depth_bin)
+# 
+#   x <- unique(tmp2$size)
+# 
+#   test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = rel_mean_conc)
+# 
+#   #test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = mean_part)
+# 
+#   data <- expand.grid(X=x, Y=y)
+# 
+#   data$Z <- as.vector(as.matrix(test |> select(-size)))
+# 
+#   data$month <- m
+# 
+#   return(data)
+# 
 # }
-
-  # x <- unique(df$month)
-
-  # data <- expand.grid(X=x, Y=y)
-
-  # data$Z <- df$cut_flux
-
-  # data$wmo <- float
-
-  # data$month <- df$month
-
-  df <- df |> filter(wmo == float) |> select(cut_flux, depth_bin, month)
-
-  y <- unique(df$depth_bin)
-
-  df <- df|> pivot_wider(names_from = depth_bin, values_from = cut_flux)
-
-  x <- unique(df$month)
-
-  data <- expand.grid(X=x, Y=y)
-
-  data$Z <- as.vector(as.matrix(df |> select(-month)))
-
-  data$wmo <- float
-
-  return(data)
-
-}
-
-d1 <- create_data_for_heatmap_bis(6904241, tmp)
-
-d2 <- create_data_for_heatmap_bis(1902578, tmp)
-
-d3 <- create_data_for_heatmap_bis(4903634, tmp)
-
-d4 <- create_data_for_heatmap_bis(6904240, tmp)
-
-dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
-
-  mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
-
-  mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) |>
-
-  mutate(Z = factor(Z, levels = levels(tmp$cut_flux)))
-
-ggplot(dd) +
-
-  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Flux range', x = 'Month', y = 'Depth bin') +
-
-  facet_wrap(~wmo) + scale_fill_viridis_d(option = 'A')
-
-```
-
-#### same plot mais pour les classes de tailles
-
-```{r, TEST }
-
-depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
-
-part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_256) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
-
-  mutate(month = month.abb[month(juld)]) |> ungroup()
-
-tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_256))) |>
-
-  mutate(cut_part = cut(mean_part, breaks = c(seq(0,100,20),+Inf))) |> ungroup()
-
-create_data_for_heatmap_bis <- function(float, df){
-
-  df <- df |> filter(wmo == float) |> select(mean_part, depth_bin, month)
-
-  y <- unique(df$depth_bin)
-
-  df <- df|> pivot_wider(names_from = depth_bin, values_from = mean_part)
-
-  x <- unique(df$month)
-
-  data <- expand.grid(X=x, Y=y)
-
-  data$Z <- as.vector(as.matrix(df |> select(-month)))
-
-  data$wmo <- float
-
-  return(data)
-
-}
-
-d1 <- create_data_for_heatmap_bis(6904241, tmp)
-
-d2 <- create_data_for_heatmap_bis(1902578, tmp)
-
-d3 <- create_data_for_heatmap_bis(4903634, tmp)
-
-d4 <- create_data_for_heatmap_bis(6904240, tmp)
-
-dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
-
-  mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
-
-  mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) #|>
-
-  #mutate(Z = factor(Z, levels = levels(tmp$cut_part)))
-
-ggplot(dd) +
-
-  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Log10(abundance) #/L', x = 'Month', y = 'Depth bin', title = 'NP_Size_256') +
-
-  facet_wrap(~wmo) + scale_fill_viridis(option = 'A')
-
-```
-
-#### same plots mais par mois
-
-Il y a une erreur lors de la standardisation -> WHY doesn't it want to group it by month???
-
-```{r, monthly heatmap}
-
-depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
-
-part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, all_of(lpm_class)) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |> mutate(month = month.abb[month(juld)]) |> ungroup() |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration')
-
-tmp <- part_distrib |> group_by(depth_bin, size, month) |> summarize(mean_part = mean(concentration)) |> mutate(rel_mean_conc = (mean_part-min(mean_part))/(max(mean_part) - min(mean_part))) |> ungroup()
-
-plot_heat_month <- function(m){
-
-  # transform data for heatmap
-
-  #m <- 'May'
-
-  tmp2 <- tmp |> filter(month == m) |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) |> select(-c(month, mean_part))
-
-  y <- unique(tmp2$depth_bin)
-
-  x <- unique(tmp2$size)
-
-  test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = rel_mean_conc)
-
-  #test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = mean_part)
-
-  data <- expand.grid(X=x, Y=y)
-
-  data$Z <- as.vector(as.matrix(test |> select(-size)))
-
-  data$month <- m
-
-  return(data)
-
-}
-
-m_data <- map_dfr(c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'), plot_heat_month)
-
-m_data <- m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)),
-
-                   month = factor(month, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')))
-
-ggplot(m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)))) +
-
-  geom_tile(aes(as.factor(X), Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + scale_fill_viridis(option = 'A') +
-
-  facet_wrap(~month)
-
-```
-
-#### Compute fluxes with Guidi et al. 2008
-
-A = 12.5, b = 3.81 fitted on 250 microns - 1.5 mm.
-
-```{r, guidi 2008}
-
-mid_DSE <- c(0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289)
-
-guidi_spectra <- vertical_profiles |> select(NP_Size_203:NP_Size_1290)
-
-# Guidi 2008
-
-A <- 12.5
-
-b <- 3.81
-
-exp <- A * mid_DSE ** b
-
-estimated_flux <- rowSums(t(t(guidi_spectra)*exp))
-
-#estimated_flux
-
-vertical_profiles$guidi_flux <- estimated_flux
-
-ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
-
-  scale_y_reverse(limits = c(2000,0)) +
-
-  scale_colour_viridis(option = 'A', limits = c(quantile(vertical_profiles$guidi_flux, probs = 0.25), quantile(vertical_profiles$guidi_flux, probs = 0.50)), oob = scales::squish, na.value = 'grey', direction = -1) +
-
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-
-  labs(x = 'Time', y = 'Predicted Guidi Flux') + facet_wrap(~wmo) + theme_bw()
-
-```
-
-#### Guidi vs. No Guidi Ha Diguedi
-
-```{r, anomaly guidi}
-
-ggplot(vertical_profiles |> mutate(ratio_flux = guidi_flux/total_pred)) + geom_point(aes(x = juld, y = depth, colour = ratio_flux)) +
-
-  scale_y_reverse(limits = c(2000,0)) +
-
-  scale_colour_viridis(option = 'A', limits = c(1,5), oob = scales::squish, na.value = 'grey') +
-
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-
-  labs(x = 'Time', y = 'Ratio Guidi/Total Model') + facet_wrap(~wmo) + theme_bw()
-
-```
-
-### Modèle basé sur les pentes
-
-```{r, model based on slope}
-
-# compute biovolumes for mid size bin (ESD)
-
-# DSE from 102 microns to 2.05 mm (this class to goes to 2.58 mm) (values obtained from Guidi's Matlab code converted on python)
-
-DSE <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
-
-              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
-
-              1.16170742, 1.46365963, 1.84409558, 2.32341484)
-
-# DSE obtained with Rainer's code
-
-DSE_Rainer <- c(0.11645117591608793,0.14636003747903648,0.18439151433041295,
-
-                0.23251999827981656,0.2933257316240555,0.36916500651202966,0.4650399965596331,
-
-                0.5860454935035602,0.738553273391974,0.9280422447589767,1.1705684285914486,
-
-                1.4795320531800742,1.8636618038457486,2.3449457391257185)
-
-# compute biovolume
-
-get_biovolume <- function(DSE){
-
-  r <- DSE/2
-
-  biovolume <- (4*pi/3)*r**3
-
-  return(biovolume)
-
-}
-
-# get biovolume for each DSE (equivalent to say for each size bin)
-
-biovolumes <- map_dbl(DSE, get_biovolume)  # unit = m^3
-
-# spectra at parking depth
-
-uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
-
-  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
-
-# convert abundance to biovolume
-
-spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes)))
-
-colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')
-
-uvp_spectra <- cbind(uvp_spectra, spectra_biovolume) |> mutate()
-
-cflux <- cflux |> mutate(small_slope = (small_flux/633)**(1/0.77), big_slope = (big_flux/633)**(1/0.77))
-
-## try model for small slopes
-
-cflux2 <- cflux |> mutate(flux = big_slope)
-
-data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
-
-# training data
-
-tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))
-
-set.seed(101)
-
-boosting_rounds <- 3000
-
-# model
-
-model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
-
-cv.folds=5,
-
-n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
-
-bag.fraction=0.5,
-
-n.cores=1
-
-)
-
-# get best model and thus best round number
-
-best <- gbm.perf(model)
-
-#print(best)
-
-# get features importance
-
-features_importance <- summary(model, n.trees=best, plotit = FALSE)
-
-```
-
-### SPECTRAL SLOPES
-
-```{r, function compute slope}
-
-# mid_DSE <- c(0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
-
-# sizebin <- c(0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
-
-#              0.532318310)
-
-mid_DSE <- c(0.0911143073, 0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
-
-sizebin <- c(0.0209587201, 0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,
-
-             0.532318310)
-
-vertical_profiles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)
-
-drifting_profiles <- all_floats |> filter(PhaseName == 'PAR') |> select(juld, cycle, depth = pres, park_depth, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |> select(-NP_Size_64)
-
-compute_slope <- function(i, data_spectra){
-
-  spectrum <- data_spectra[i,]
-
-  spectrum_norm <- spectrum/sizebin
-
-  # prepare data for linear regression
-
-  Y <- log(spectrum_norm)
-
-  X <- log(mid_DSE)
-
-  # check for finite value
-
-  h <- is.finite(Y)
-
-  Y <- Y[h]
-
-  X <- X[h]
-
-  data_slope <- tibble(X=X, Y=Y)
-
-  model <- lm(formula = Y ~ X, data = data_slope)
-
-  slope <- model$coefficients[2]
-
-  return(slope)
-
-}
-
-# compute slope on drifting profiles
-
-drifting_spectra <- as.matrix(data.frame(drifting_profiles[,6:20]))
-
-index <- seq(from = 1, to = nrow(drifting_spectra), by = 1)
-
-slopes <- map_dbl(index, compute_slope, data_spectra = drifting_spectra)
-
-drifting_profiles$spectral_slope <- slopes
-
-ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
-
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-
-  scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')
-
-# compute slope on vertical profiles
-
-vertical_spectra <- as.matrix(data.frame(vertical_profiles[,5:19]))
-
-index <- seq(from = 1, to = nrow(vertical_spectra), by = 1)
-
-slopes <- map_dbl(index, compute_slope, data_spectra = vertical_spectra)
-
-vertical_profiles$spectral_slope <- slopes
-
-```
-
-```{r, add slopes}
-
-# vertical_profiles$slope <- slopes
-
-ggplot(vertical_profiles, aes(x = juld, y = depth, colour = spectral_slope)) + geom_point() + scale_colour_viridis(option = 'H') +
-
-  scale_y_reverse(limits = c(2000,0)) +
-
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-
-  labs(x = 'Time', y = TeX('slope')) + facet_wrap(~wmo)
-
-```
-
-### DRIFT DATA
-
-```{r spectral slope + rati min/max flux}
-
-tmp <- cflux |> mutate(ratioF = small_flux/(small_flux+big_flux))
-
-# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF, colour = park_depth)) + geom_point() + facet_wrap(~wmo)
-
-#
-
-# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_point() + facet_wrap(~park_depth)
-
-#
-
-# ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)
-
-ggplot(tmp |> filter(big_flux>0), aes(x = min_time, y = ratioF)) + geom_point() + geom_smooth() + facet_wrap(~park_depth)
-
-ggplot(drifting_profiles, aes(x = juld, y = spectral_slope, colour = factor(park_depth))) + geom_smooth() + facet_wrap(~wmo) + theme_bw() +
-
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-
-  scale_color_brewer(palette = 'Dark2') + labs(x = 'Month', y = 'Spectral slope')
-
-tmp2 <- drifting_profiles |> group_by(wmo, cycle, park_depth) |> summarize(median_spectral_slope = median(spectral_slope))
-
-# merge data
-
-check <- merge(tmp, tmp2)
-
-check2 <- merge(tmp, tmp2) |> filter(ratioF < 1)
-
-ggplot(check, aes(ratioF, median_spectral_slope)) + geom_point()
-
-ggplot(check2, aes(ratioF, median_spectral_slope)) + geom_point()
-
-# ggplot(tmp, aes(x = min_time, y = ratioF)) + geom_smooth() + facet_wrap(~park_depth)
+# 
+# m_data <- map_dfr(c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'), plot_heat_month)
+# 
+# m_data <- m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)),
+# 
+#                    month = factor(month, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')))
+# 
+# ggplot(m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)))) +
+# 
+#   geom_tile(aes(as.factor(X), Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + scale_fill_viridis(option = 'A') +
+# 
+#   facet_wrap(~month)
 

### 2023-02-09 article3.qmd
update

@@
   check_colour <- unique(tmp2$colour)
-  if(length(check_colour) == 3){ # there is at least a negative jump
-    index_neg_jump <- which(tmp2$colour == 'negative jump')
-    tmp2 <- tmp2[-c(index_neg_jump, index_neg_jump+1),]
-    tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
-    even_indexes <- seq(2,nrow(tmp2),2)
-    tmp2 <- tmp2[even_indexes,]
-  }else if(length(check_colour) == 2){ # only positive jumps
+  # if(length(check_colour) == 3){ # there is at least a negative jump
+  #   index_neg_jump <- which(tmp2$colour == 'negative jump')
+  #   tmp2 <- tmp2[-c(index_neg_jump, index_neg_jump+1),]
+  #   tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
+  #   even_indexes <- seq(2,nrow(tmp2),2)
+  #   tmp2 <- tmp2[even_indexes,]
+  # }else if(length(check_colour) == 2){ # only positive jumps
+  #   tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
+  #   even_indexes <- seq(2,nrow(tmp2),2)
+  #   tmp2 <- tmp2[even_indexes,]
+  # }else{ # No jump
+  #   tmp2 <- NULL
+  # }
+  if(length(check_colour) >= 2){ # there is a least a jump (positive or negative)
     tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
@@
 
+```{r, get jump info}
+
+get_jump_info <- function(data){
+
+  # make sure that the cp signal is in chronological order
+  tmp <- data |> arrange(dates)
+
+  # despike cp data with a 7-point moving window
+  tmp$cp <- despike(tmp$cp, k = 3)
+
+  # smooth cp data with a 3-point moving median, n time(s)
+  tmp$cp <- slide(tmp$cp, fun = median, k = 3, n = 1, na.rm=T)
+
+  # compute slope between two adjacent points (except first point) # we could start after 1h to let the float stabilize
+  delta_x <- as.numeric(tmp$dates - lag(tmp$dates), units = 'days')
+  delta_y <- tmp$cp - lag(tmp$cp)
+  tmp$slope <- delta_y/delta_x
+
+  # compute a Z score (assuming a normal distribution of the slopes) on the slopes
+  tmp <- tmp |> mutate(zscore = (slope - mean(slope, na.rm = T))/sd(slope, na.rm = T))
+
+  # spot outliers on the Z score signal
+  # interquartile range between Q25 and Q75 -> had to used that and not the despike function because slopes are often close (or equal) to 0 so it can miss clear jumps.
+Q25 and Q75 are more trustworthy in this case than the despike function of Jean-Olivier (see package castr on his github: https://github.com/jiho/castr)
+  IQR <- quantile(tmp$zscore, probs = 0.75, na.rm=T) - quantile(tmp$zscore, probs = 0.25, na.rm=T)
+  # outliers ('spikes' in the Z score signal)
+  spikes_down <- tmp$zscore < quantile(tmp$zscore, 0.25, na.rm=T) - 1.5 *IQR
+  spikes_up <-  tmp$zscore > quantile(tmp$zscore, 0.75, na.rm=T) + 1.5 *IQR
+  spikes <- as.logical(spikes_down + spikes_up)
+
+  # assign spikes
+  tmp$spikes <- spikes
+
+  # assign colour code to cp signal
+  tmp$colour <- 'base signal' # base signal = smoothed despiked cp signal
+  tmp[which(tmp$spikes == TRUE),]$colour <- 'jump'
+
+  # add group to compute the slope of each group of points, separated by a jump
+  tmp$group <- NA
+
+  # index of jumps in the array
+  jump_index <- which(tmp$colour == 'jump')
+
+  # assign group identity to each group of points, separated by a jump (= subgroup)
+  for (i in jump_index){
+    for (j in 1:nrow(tmp)){
+      if ((j < i) & (is.na(tmp$group[j]))){
+        tmp$group[j] <- paste0('group_',i)
+      }
+    }
+  }
+  tmp$group[which(is.na(tmp$group))] <- 'last_group'
+
+  # compute slope for each subgroup
+  slope_df <- tmp |> filter(colour == 'base signal', slope != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), 
+                                                                                                                  nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],
+                                                                                                                  delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),
+                                                                                                                  delta_y = (last_cp-first_cp)*0.25, slope = delta_y/delta_x) # *0.25 to convert cp to ATN
+
+  # remove negative slope from the mean slope (no physical meaning)
+  slope_df <- slope_df |> filter(slope > 0)
+
+  # remove if only one point (cannot fit a slope with one point)
+  slope_df <- slope_df |> filter(nb_points > 1)
+
+  # compute weighted average slope (to take into account the fact that some subgroups might have 2 points and a high slope vs. large group of points with a small slope)
+  mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)
+
+  # convert cp to POC using Estapa's relationship 
+  poc_flux <- 633*(mean_slope**0.77) # /!\ slope computed for ATN on y axis (delta_y *0.25 because ATN = cp*0.25) -> should be OK
+
+  # build dataframe to plot each subgroup
+  part1 <- slope_df |> select(group, time = min_time, cp = first_cp)
+  part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
+  part_slope <- rbind(part1, part2)
+
+  # spot negative jump
+  tmp$colour[which((tmp$colour == 'jump') & (tmp$slope < 0))]  <- 'negative jump'
+
+  # add big particles flux to the party
+  rows_to_keep <- c(jump_index, jump_index-1)
+  tmp2 <- tmp[rows_to_keep,] |> select(dates, cp, slope, colour, group) |> arrange(dates)
+
+  # remove negative jumps, if any
+  check_colour <- unique(tmp2$colour)
+  if(length(check_colour) >= 2){ # there is a least a jump (positive or negative)
+    tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
+    even_indexes <- seq(2,nrow(tmp2),2)
+    tmp2 <- tmp2[even_indexes,]
+  }else{ # No jump
+    tmp2 <- NULL
+  }
+
+  if(is.null(tmp2)){ # no jump
+    big_part_poc_flux <- 0
+    tmp3 <- NULL
+  }else{
+    tmp3 <- tmp2 |> filter(diff_jump > 0) # remove negative
+    if(nrow(tmp3) == 0){ # no positive jumps
+      big_part_poc_flux <- 0
+    }else{
+      delta_y <- sum(tmp3$diff_jump) *0.25 # to get ATN (= cp*0.25)
+      max_time <- max(tmp$dates)
+      min_time <- min(tmp$dates)
+      delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
+      slope_big_part <- delta_y/delta_x
+      big_part_poc_flux <- 633*(slope_big_part**0.77)
+    }
+  }
+
+  # compute total drifting time
+  max_time <- max(tmp$dates)
+  min_time <- min(tmp$dates)
+  drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))
+
+  # to plot subgroups
+  part_slope_tmp <- part_slope |> mutate(dates = time, colour = 'slope')
+
+  # plot
+  # jump_plot <- plot_ly(tmp, x = ~dates, y = ~cp, type = 'scatter', mode = 'markers', color = ~colour, colors = c('#003366','#E31B23', '#FFC325')) |>
+  #   add_lines(data= part_slope_tmp, x = ~dates, y = ~cp, split = ~group, color = I('#DCEEF3'), showlegend = F) |>
+  #   layout(title= paste0('Drifting time: ', round(drifting_time,3), ' days\n',
+  #                        'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n',
+  #                        'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
+  #                        'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1'), yaxis = list(title = 'Cp (1/m)'), xaxis = list(title = 'Time'))
+
+  # adapt script to return big and small flux
+  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],
+               cycle = data$cycle[1])
+
+  # clean data to send
+  slope_df <- slope_df |> arrange(min_time) |> mutate(min_drifting_time = df$min_time, max_drifting_time = df$max_time, small_flux = df$small_flux, big_flux = df$big_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1])
+  if(is.null(tmp3)){
+    big_jump <- tmp3
+  }else{
+     big_jump <- tmp3 |> mutate(park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1]) 
+  }
+
+  return(list(slope_info = slope_df, jump_info = big_jump))
+}
+```
+
+TO DO HERE
+
+```{r, compute slope and jump info}
+
+# start loop for all data
+wmo <- c(4903634, 6904240, 6904241, 1902578)
+park_depth <- c('200 m', '500 m', '1000 m')
+
+# i <- 4903634
+# j <- '1000 m'
+# k <- 15
+# tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
+# 
+# data <- tmp
+
+res_slope <- data.frame()
+res_jumps <- data.frame()
+for (i in wmo){
+  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
+  for (j in park_depth){
+    for (k in seq(1:max_cycle)){
+      tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
+      if(nrow(tmp) == 0){
+        next
+      }else if(nrow(tmp) < 3){ # case where there is not enough data
+        next
+      }else{
+          output <- get_jump_info(tmp)
+          res_slope <- rbind(res_slope, output$slope_info)
+          res_jumps <- rbind(res_jumps, output$jump_info)
+      }
+    }
+  }
+}
+
+```
+
 ```{r, model based on slope}
@@
 
+```{r, check gaussian profiles}
+
+# depth layers for Gaussian fit
+depthL <- c(seq(40,200,20), c(250,300,400,500,600,800,1000,1250,1500,1750,2000))
+
+# Gaussian formula #> but z is not the depth but the TIME
+fgauss <- function(z, peak_conc, peak_doy, dz){
+  peak_conc*exp(-((z-peak_doy)**2)/(dz**2))
+}
+
+return_fitted <- function(d){
+
+  #timeframe near the bloom so to say
+  ztime <- seq(160,240,1)
+  # gaussian fit
+  tib <- map_dfr(1:nrow(d), function(i){tibble(ztime = ztime, fgauss = fgauss(ztime, d$max_abundance[i], d$time_max_abundance[i], d$standard_dev[i]), depth = d$depth[i])})
+  return(tib)
+}
+
+# test
+tmp <- sinking_fit |> filter(wmo == 6904240, size == 645)
+
+test <- return_fitted(tmp)
+
+ggplot(test, aes(x = ztime, y = fgauss, colour = factor(depth))) + geom_point() + theme_bw() #+ ylim(c(0,.2))
+
+```
+
 ### TS diagram

### 2023-02-13 article3.qmd
update

@@
 
-  # plot
-  # jump_plot <- plot_ly(tmp, x = ~dates, y = ~cp, type = 'scatter', mode = 'markers', color = ~colour, colors = c('#003366','#E31B23', '#FFC325')) |>
-  #   add_lines(data= part_slope_tmp, x = ~dates, y = ~cp, split = ~group, color = I('#DCEEF3'), showlegend = F) |>
-  #   layout(title= paste0('Drifting time: ', round(drifting_time,3), ' days\n',
-  #                        'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n',
-  #                        'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
-  #                        'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1'), yaxis = list(title = 'Cp (1/m)'), xaxis = list(title = 'Time'))
-
   # adapt script to return big and small flux
@@
 
+```{r, check total biovolume with slope value}
+
+uvp_drifting <- all_floats |> filter(PhaseName == 'PAR') |> select(juld, cycle, wmo, park_depth, all_of(lpm_class)) |> 
+  drop_na(NP_Size_102) |> mutate(date = as_date(juld))
+
+# DSE for computing biovolumes
+#DSE <- c(0.29042685,0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779, 1.16170742, 1.46365963)
+DSE_long <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685, 0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+                1.16170742, 1.46365963, 1.84409558, 2.32341484)
+
+# compute biovolume
+get_biovolume <- function(DSE){
+  r <- DSE/2
+  biovolume <- (4*pi/3)*r**3
+  return(biovolume)
+}
+
+# get biovolume for each DSE (equivalent to say for each size bin)
+biovolumes <- map_dbl(DSE_long, get_biovolume)  # unit = mm^3
+
+# compute biovolume and add them to the dataframe
+spectra_biovolume <- tibble(as.data.frame(t(t(uvp_drifting[,all_of(lpm_class)])*biovolumes)))
+colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('NP_Size', 'Biovolume')
+uvp_drifting <- cbind(uvp_drifting, spectra_biovolume) |> rowwise() |> mutate(total_biovolume = sum(across(all_of(biovolumes_names))))
+biovolumes_names <- colnames(spectra_biovolume)
+
+assign_time_group <- function(d_uvp, d_ost){
+  d_uvp$group <- NA
+  for (i in 1:nrow(d_ost)){
+    low_time <- d_ost$min_time[i]
+    high_time <- d_ost$max_time[i]
+    #print(c(low_time, high_time))
+    group <- d_ost$group[i]
+    time_index <- which(d_uvp$juld >= low_time & d_uvp$juld <= high_time)
+    d_uvp$group[time_index] <- group
+  }
+  return(d_uvp)
+}
+
+# loop for it
+# start loop for all data
+wmo <- c(4903634, 6904240, 6904241, 1902578)
+park_depth <- c('200 m', '500 m', '1000 m')
+
+test <- data.frame()
+for (i in wmo){
+  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
+  for (j in park_depth){
+    for (k in seq(1:max_cycle)){
+      tmp_uvp <- filter(uvp_drifting, wmo == i, cycle == k, park_depth == j) |> arrange(juld) |> select(juld, cycle, wmo, park_depth, total_biovolume) #, biovolumes_names)
+      tmp_ost <- res_slope |> filter(wmo == i, cycle == k, park_depth == j)
+      tmp <- assign_time_group(tmp_uvp, tmp_ost)
+      #tmp2 <- tmp |> filter(!is.na(group)) |> group_by(group) |> mutate(uvp_points = n(), min_date = min(juld), max_date = max(juld), mean_date = mean(floor_date(juld))) # https://stackoverflow.com/questions/44814435/calculate-average-daily-value-from-large-data-set-with-r-standard-format-date-ti
+      tmp2 <- tmp |> filter(!is.na(group)) |> group_by(group) |> mutate(uvp_points = n(), mean_date = mean(floor_date(juld))) 
+      tmp3 <- merge(tmp2, tmp_ost) |> group_by(group, cycle, wmo, park_depth) |> summarize(mean_date = unique(mean_date), uvp_points = unique(uvp_points), ost_points = unique(nb_points), group_total_biovolume = sum(total_biovolume), slope = unique(slope), weighted_group_biovolume = group_total_biovolume/uvp_points)
+      test <- rbind(test, tmp3)
+    }
+  }
+}
+
+ggplot(test) + geom_line(aes(x = mean_date, y = weighted_group_biovolume)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4)
+ggplot(test) + geom_line(aes(x = mean_date, y = slope)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4)
+
+# check
+test2 <- test |> group_by(wmo, cycle) |> mutate(mean_mean_date = mean(floor_date(mean_date)))
+
+ggplot(test2) +
+  geom_pointrange(aes(x = mean_mean_date, y = weighted_group_biovolume),
+                      stat = 'summary',
+                      fun.min = function(z) {quantile(z, 0.25)},
+                      fun.max = function(z) {quantile(z, 0.75)},
+                      fun = median, colour = 'black', size = .1) +
+    facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
+
+ggplot(test2) +
+  geom_pointrange(aes(x = mean_mean_date, y = slope),
+                      stat = 'summary',
+                      fun.min = function(z) {quantile(z, 0.25)},
+                      fun.max = function(z) {quantile(z, 0.75)},
+                      fun = median, colour = 'black', size = .1) +
+    facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4)
+
+#ggplot(test) + geom_point(aes(x = log(slope), y = log(weighted_group_biovolume))) + facet_wrap(~park_depth)
+
+# what about jump values with uvp measurements at jump
+uvp_drifting <- uvp_drifting |> mutate(num_date = as.numeric(juld))
+res_jumps <- res_jumps |> mutate(num_date = as.numeric(dates))
+
+assign_uvp_to_jump <- function(d_uvp, d_jump){
+  d_jump$biovolume <- NA
+  for (i in 1:nrow(d_jump)){
+    index_jump_for_uvp <- which.min(abs(tmp_uvp$num_date - tmp_ost_jump$num_date[i]))
+    d_jump$biovolume[i] <- tmp_uvp$total_biovolume[index_jump_for_uvp]
+  }
+  return(d_jump)
+}
+
+test_jump <- data.frame()
+for (i in wmo){
+  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
+  for (j in park_depth){
+    for (k in seq(1:max_cycle)){
+      tmp_uvp <- filter(uvp_drifting, wmo == i, cycle == k, park_depth == j) |> arrange(juld) |> select(juld, cycle, wmo, park_depth, total_biovolume, num_date) 
+      tmp_ost_jump <- res_jumps |> filter(wmo == i, cycle == k, park_depth == j)
+      if(nrow(tmp_ost_jump) == 0 | nrow(tmp_uvp) == 0){
+        next
+      }else{
+        tmp <- assign_uvp_to_jump(tmp_uvp, tmp_ost_jump)
+      }
+      test_jump <- rbind(test_jump, tmp)
+    }
+  }
+}
+
+ggplot(test_jump) + geom_point(aes(diff_jump, biovolume))
+ggplot(test_jump) + geom_point(aes(log(diff_jump), log(biovolume)))
+```
+
 ```{r, model based on slope}
@@
 
-DSE <- c(0.29042685,0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779, 1.16170742, 1.46365963)
-
-# compute biovolume
-get_biovolume <- function(DSE){
-  r <- DSE/2
-  biovolume <- (4*pi/3)*r**3
-  return(biovolume)
-}
-
-# get biovolume for each DSE (equivalent to say for each size bin)
-biovolumes <- map_dbl(DSE, get_biovolume)  # unit = m^3
+# DSE <- c(0.29042685,0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779, 1.16170742, 1.46365963)
+# 
+# # compute biovolume
+# get_biovolume <- function(DSE){
+#   r <- DSE/2
+#   biovolume <- (4*pi/3)*r**3
+#   return(biovolume)
+# }
+
+# # get biovolume for each DSE (equivalent to say for each size bin)
+# biovolumes <- map_dbl(DSE, get_biovolume)  # unit = mm^3
 

### 2023-02-13 intro.qmd
update

@@
 The global carbon cycle (@fig-global_carbon_cycle) consists of a series of exchanges of carbon in various organic and inorganic forms among four major reservoirs: the lithosphere, the hydrosphere, the atmosphere and the biosphere.
-These exchanges occur at different timescales, such that the carbon cycle can be divided into a slow and a fast component.
+These exchanges occur at different timescales, such that the carbon cycle can be roughly divided into a slow and a fast component.
 The former operates on geological timescales where carbon is hidden from the surface for millions of years (e.g.
-in hydrocarbon deposits such as petroleum or natural gas) before being released through volcanic activity while the latter operates on shorter timescales, from minutes to years (e.g.
+in hydrocarbon deposits such as petroleum or natural gas) before a part of it is released through volcanic activity while the latter operates on shorter timescales, from minutes to years (e.g.
 photosynthesis, air-sea fluxes) up to decades and millenia (e.g.
@@
 
-99% of the Earth's carbon is stored in rocks, holding about 100,000 times as much carbon as the atmosphere while the ocean contains about 60 times as much carbon as the atmosphere [@DeVries2022-of; @Bertrand2021-nh; @Sarmiento2006-ec] with roughly 37,000 Pg (i.e.
-1 Pg = 10$^{15}$ g = 1 Gt) of dissolved inorganic carbon (DIC) and 700 Pg of dissolved organic carbon (DOC) [@Friedlingstein2021-kc].
+99% of the Earth's carbon is stored in rocks, holding about 100,000 times as much carbon as the atmosphere while the ocean contains about 60 times as much carbon as the atmosphere [@Sarmiento2006-ec; @Bertrand2021-nh; @DeVries2022-of] with roughly 37,000 Pg (i.e.
+1 Pg = 10$^{15}$ g = 1 Gt) of dissolved inorganic carbon (DIC) and 700 Pg of dissolved organic carbon (DOC) [@Friedlingstein2022-th].
 
-![Illustration of the global carbon cycle (numbers were taken for the year 2020).
-Arrows and boxes depict, respectively, carbon fluxes in Pg C year$^{-1}$ and carbon content in Pg C.
-Figure taken from @DeVries2022-of.](image/intro/carbon_cycle_devries2022.jpeg){#fig-global_carbon_cycle}
+![Illustration of the global carbon cycle (numbers represent a global average for the decade 2012-2021).
+Arrows and circles depict, respectively, carbon fluxes in Pg C year$^{-1}$ and carbon content in Pg C.
+Figure taken from @Friedlingstein2022-XX.](image/intro/global_carbon_cycle.png){#fig-global_carbon_cycle}
 
@@
 
-Thankfully, more than half of this anthropogenic $CO_{2}$ has been absorbed by terrestrial ecosystems and by the ocean [@Friedlingstein2021-kc], forming major carbon sinks.
+Thankfully, more than half of this anthropogenic $CO_{2}$ has been absorbed by terrestrial ecosystems and by the ocean [@Friedlingstein2022-XX], forming major carbon sinks.
 
@@
 
-The ocean is known to play a key role in the fast carbon cycle due to its size and rapid exchange of $CO_{2}$ with the atmosphere, controlling atmospheric $CO_{2}$ levels on timescales of decades to millennia [@DeVries2022-of; @Sarmiento2006-ec].
-Currently, the ocean absorbs \~2.5 Gt C year$^{-1}$, which represents about 25% of annual anthropogenic $CO_{2}$ emissions [@Canadell2021-ds; @Friedlingstein2021-kc].
+The ocean is known to play a key role in the fast carbon cycle due to its size and rapid exchange of $CO_{2}$ with the atmosphere, controlling atmospheric $CO_{2}$ levels on timescales of decades to millennia [@Sarmiento2006-ec; @DeVries2022-of].
+Currently, the ocean absorbs \~2.5 Pg C year$^{-1}$, which represents about 25% of annual anthropogenic $CO_{2}$ emissions [@Canadell2021-ds; @Friedlingstein2022-XX].
 In order to understand how the ocean absorbs the latter, one must take a look at seawater carbonate chemistry.
@@
 
-The intensity of the exchange of $CO_{2}$ at the air-sea interface is driven by the difference in the partial pressure of $CO_{2}$ in the atmosphere ($pCO_{2}^{atm}$) and in the surface ocean ($pCO_{2}^{surface}$).
+The intensity of the exchange of $CO_{2}$ at the air-sea interface is driven by the difference between the partial pressure of $CO_{2}$ in the atmosphere ($pCO_{2}^{atm}$) and in the surface ocean ($pCO_{2}^{surface}$).
 
@@
 
-From @eq-carbonic_acid, one can derive that p$CO_{2}$<sup>water parcel</sup> is related to the dissolved $CO_{2}$ concentration by the following relation
+From @eq-carbonic_acid, one can derive that p$CO_{2}^{water~parcel}$ is related to the dissolved $CO_{2}$ concentration by the following relation
 
@@
 
-The solubility pump is responsible for the great majority of all the anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed DIC gradient according to [@Sarmiento2006-ec] or 35% according to @Toggweiler2003-ye.
+The solubility pump is responsible for the great majority of all the anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed DIC gradient according to @Sarmiento2006-ec or 35% according to @Toggweiler2003-ye.
 Both values are used in the literature (e.g.
-10% in @Boyd2019-ez and @Iversen2022-qj and 35% in @DeVries2022-of, @Passow2012-zh and @Riebesell2009-qj).
+10% in @Boyd2019-ez and @Iversen2022-qj and 35% in @Riebesell2009-qj, @Passow2012-zh and @DeVries2022-of).
 
@@
 
-The BCP would thus be responsible for the remaining 70% of the observed DIC gradient according to @Sarmiento2006-ec.
-However, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess $CO_{2}$, released by anthropogenic activities.
-First, as explained in @Sarmiento2006-ec, atmospheric $CO_{2}$ levels were stable for several thousand years before the industrial revolution started, therefore it is assumed that the BCP was operating at steady state.
-Second, phytoplankton organisms are mostly controlled by light and nutrients availability, not by $CO_{2}$.
+This sequestration (storage) of carbon is usually defined as the long-term removal of $CO_{2}$ (or other forms of carbon) from the atmosphere on climatically significant time scales [e.g.
+@Boyd2019-ez; @Ipcc2019-qc].
+While this definition seems logical, it does not specify precisely what is meant by "climatically significant time scales".
+In what follows, I will use the 100-year time horizon to refer to carbon sequestration (unless specified otherwise) as used by other BCP studies [e.g.
+@Lampitt2008-kq; @Passow2012-zh].
 
-The BCP is a natural-based solution for carbon sequestration by definition (@eq-photosynthesis).
-However, natural-based solutions for carbon sequestration are generally aligned with the concept of 'blue carbon' defined by @Ipcc2019-qc as "all biologically-driven carbon fluxes and storage in marine systems that are amenable to management" which usually encompasses only coastal ecosystems [@Ipcc2019-qc; @Lovelock2019-dr].
-
-Coastal blue carbon ecosystems (e.g.
-mangroves, tidal marshes, seagrass meadows) are thought to store in their living biomass (years to decades) and sediment (millenia) more than 30 Pg C with an estimated absorption of \~0.3 Pg C year$^{-1}$ [@Macreadie2021-rb] though those numbers should be taken with caution because there are major uncertainties in the distribution of seagrass meadows and tidal marshes [@Macreadie2021-rb] as well as in carbon burial rates [@Williamson2022-ud].
-In spite of that, the current expectation of additional carbon removal by coastal blue carbon by 2100 should be around 25-30 Pg C [@Williamson2022-ud; @Gattuso2021-cr], that is the equivalent of 2.5-3 years of current anthropogenic emissions [@Friedlingstein2021-kc].
+The amount of carbon that is sequestered by the BCP mainly depends its strength and its efficiency.
+In this manuscript, the strength of the BCP is defined as "the magnitude of the particulate organic carbon (POC) flux at a reference export depth" [@Claustre2021-tx] and its efficiency (also known as transfer efficiency) is defined as the fraction of POC flux transported to some reference depth below the export depth [@Buesseler2020-sn; @Claustre2021-tx].
+The latter is often set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where POC is produced (i.e.
+via photosynthesis) and where POC is remineralized (or respired), that is organic matter converted back to inorganic forms (i.e.
+nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd].
 
@@
 
+::: {.blackbox data-latex=""}
+::: {.center data-latex=""}
+BCP and blue carbon
+:::
+
+The BCP is a natural-based solution for carbon sequestration by definition (@eq-photosynthesis).
+However, natural-based solutions for carbon sequestration are generally aligned with the concept of 'blue carbon' defined by @Ipcc2019-qc as "all biologically-driven carbon fluxes and storage in marine systems that are amenable to management" which usually encompasses only coastal ecosystems [@Ipcc2019-qc; @Lovelock2019-dr].
+Coastal blue carbon ecosystems (e.g.
+mangroves, tidal marshes, seagrass meadows) are thought to store in their living biomass (years to decades) and sediment (millenia) more than 30 Pg C with an estimated absorption of \~0.3 Pg C year$^{-1}$ [@Macreadie2021-rb] though those numbers should be taken with caution because there are major uncertainties in the distribution of seagrass meadows and tidal marshes [@Macreadie2021-rb] as well as in carbon burial rates [@Williamson2022-ud].
+In spite of that, the current expectation of additional carbon removal by coastal blue carbon by 2100 should be around 25-30 Pg C [@Gattuso2021-cr; @Williamson2022-ud], that is the equivalent of 2.5-3 years of current anthropogenic emissions [@Friedlingstein2022-XX].
+:::
+
+According to @Sarmiento2006-ec, the BCP would thus be responsible for the remaining 70% of the observed DIC gradient.
+However, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess atmospheric $CO_{2}$, released by anthropogenic activities.
+First, as explained in @Sarmiento2006-ec, atmospheric $CO_{2}$ levels were stable for several thousand years before the industrial revolution started, therefore it is assumed that the BCP was operating at steady state.
+Second, photosynthesis is mostly controlled by light and nutrients availability, not by $CO_{2}$.
+
 However, climate change is most likely going to impact the biological (e.g.
-nutrients availability, shift in phytoplankton communities), physical (e.g.
+shift in phytoplankton communities), physical (e.g.
 temperature) and chemical (e.g.
-acidification) drivers of the BCP [@Bopp2013-jx; @Passow2012-zh; @Riebesell2009-qj] whose long-term carbon sequestration capacity remains uncertain [@Passow2012-zh; @Turner2015-af; @Wilson2022-bf; @DeVries2022-of].
+nutrients availability, acidification) drivers of the BCP [@Riebesell2009-qj; @Passow2012-zh; @Bopp2013-jx] whose long-term carbon sequestration capacity remains uncertain [@Passow2012-zh; @Turner2015-af; @DeVries2022-of; @Wilson2022-bf].
 
@@
 
-[^intro-1]: i.e. dissolved organic matter resistant to microbial utilization that can persist in the ocean interior for thousand of years.
+[^intro-1]: i.e. dissolved organic matter resistant to microbial utilization that can persist in the ocean interior for thousands of years.
 
@@
 ## A new paradigm for the BCP {#sec-a-new-paradigm-for-the-bcp}
-
-The amount of carbon that is stored for more than 100 years in the ocean (i.e.
-sequestered) mainly depends on the strength and the efficiency of the BCP.
-In this manuscript, the strength of the BCP is defined as "the magnitude of the particulate organic carbon (POC) flux at a reference export depth" [@Claustre2021-tx] and its efficiency (also known as transfer efficiency) is defined as the fraction of POC flux transported to some reference depth below the export depth [@Buesseler2020-sn; @Claustre2021-tx].
-However, the export depth is not unique.
-It is often set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where POC is produced (i.e.
-via photosynthesis) and where POC is degraded.
-This degradation occurs through the remineralization (i.e.
-respiration) where organic matter is converted back to inorganic forms (i.e.
-nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd].
-The bulk of respiration occurs in the mesopelagic zone between the export depth and 1000 m [@Martin1987-um].
 
@@
 | Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
-|------------------------|----------------------------------|---------------------------|--------------------------------------|
+|------------------|------------------|------------------|-------------------|
 | Gravitational          | export depth - sediment          | None                      | N/A (142)                            |

### 2023-02-14 intro.qmd
update

@@
 Arrows and circles depict, respectively, carbon fluxes in Pg C year$^{-1}$ and carbon content in Pg C.
-Figure taken from @Friedlingstein2022-XX.](image/intro/global_carbon_cycle.png){#fig-global_carbon_cycle}
+Figure taken from @Friedlingstein2022-th.](image/intro/global_carbon_cycle.png){#fig-global_carbon_cycle}
 
@@
 
-Thankfully, more than half of this anthropogenic $CO_{2}$ has been absorbed by terrestrial ecosystems and by the ocean [@Friedlingstein2022-XX], forming major carbon sinks.
+Thankfully, more than half of this anthropogenic $CO_{2}$ has been absorbed by terrestrial ecosystems and by the ocean [@Friedlingstein2022-th], forming major carbon sinks.
 
@@
 The ocean is known to play a key role in the fast carbon cycle due to its size and rapid exchange of $CO_{2}$ with the atmosphere, controlling atmospheric $CO_{2}$ levels on timescales of decades to millennia [@Sarmiento2006-ec; @DeVries2022-of].
-Currently, the ocean absorbs \~2.5 Pg C year$^{-1}$, which represents about 25% of annual anthropogenic $CO_{2}$ emissions [@Canadell2021-ds; @Friedlingstein2022-XX].
+Currently, the ocean absorbs \~2.5 Pg C year$^{-1}$, which represents about 25% of annual anthropogenic $CO_{2}$ emissions [@Canadell2021-ds; @Friedlingstein2022-th].
 In order to understand how the ocean absorbs the latter, one must take a look at seawater carbonate chemistry.
@@
 
-Also known as the soft-tissue pump [@Volk1985-oc], the biological carbon pump (BCP), is defined by @Ipcc2019-qc as "a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis (@eq-photosynthesis) in sunlit surface water and then transported to the ocean interior, and possibly the sediment, resulting in the storage of carbon".
+Also known as the soft-tissue pump [@Volk1985-oc], the biological carbon pump (BCP), is defined by @Ipcc2019-qc as "a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis (@eq-photosynthesis) in sunlit surface water and then transported to the ocean interior (i.e.
+exported flux), and possibly the sediment, resulting in the storage of carbon".
 
@@
 
-This sequestration (storage) of carbon is usually defined as the long-term removal of $CO_{2}$ (or other forms of carbon) from the atmosphere on climatically significant time scales [e.g.
-@Boyd2019-ez; @Ipcc2019-qc].
+In order to estimate the carbon export flux, one needs to set a depth horizon [@Siegel2022-wd].
+Historically, the latter has often been set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should now use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where particulate organic carbon (POC) is produced (i.e.
+via photosynthesis) and where POC is consumed (i.e.
+remineralized).
+This remineralization (or respiration) consists in the conversion of organic matter back to inorganic forms (i.e.
+nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd].
+
+According to @Sarmiento2006-ec, the BCP would thus be responsible for the remaining 70% of the observed DIC gradient.
+Without it, there would be no production of organic matter in the EZ (consuming DIC) and no remineralization in the ocean interior (producing DIC) resulting in a weaker DIC gradient.
+In fact, @DeVries2022-of calculated that such a decrease in the DIC gradient would reduce the ocean carbon inventory by roughly 10% and increase atmospheric $CO_{2}$ levels by a factor 2.
+Similarly, @Kwon2009-ir showed that a deepening of the remineralization depth (i.e.
+where 63% of sinking carbon is respired) would decrease atmospheric $CO_{2}$ concentrations due to the redistribution of respired carbon from intermediate waters to bottom waters, enhancing the DIC gradient.
+
+The sinking carbon that is not remineralized is said to be sequestered.
+The sequestration (storage) of carbon is usually defined as the long-term removal of $CO_{2}$ (or other forms of carbon) from the atmosphere on climatically significant time scales [@Boyd2019-ez; @Ipcc2019-qc; @Siegel2022-wd].
 While this definition seems logical, it does not specify precisely what is meant by "climatically significant time scales".
@@
 
-The amount of carbon that is sequestered by the BCP mainly depends its strength and its efficiency.
-In this manuscript, the strength of the BCP is defined as "the magnitude of the particulate organic carbon (POC) flux at a reference export depth" [@Claustre2021-tx] and its efficiency (also known as transfer efficiency) is defined as the fraction of POC flux transported to some reference depth below the export depth [@Buesseler2020-sn; @Claustre2021-tx].
-The latter is often set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where POC is produced (i.e.
-via photosynthesis) and where POC is remineralized (or respired), that is organic matter converted back to inorganic forms (i.e.
-nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd].
-
-In comparison, the recent work of @Nowicki2022-no suggests that the BCP exports on average 10.2 Pg C year$^{-1}$, sequesters \~1,300 Pg C with a mean sequestration time of 127 years and buries 0.4 Pg C year$^{-1}$ in the sediment that will remain sequestered for millions of years.
-Therefore, when talking about natural-based solutions for carbon sequestration, one should keep in mind that even though the BCP does not contribute to the absorption of anthropogenic $CO_{2}$ *per se*, it remains the main process that maintains the DIC gradient.
+The amount of carbon that is sequestered by the BCP mainly depends on its strength and its efficiency.
+In this manuscript, the strength of the BCP is defined as "the magnitude of the POC flux at a reference export depth" [@Claustre2021-tx] and its efficiency (also known as transfer efficiency) is defined as the fraction of POC flux transported to some reference depth below the export depth [@Buesseler2020-sn; @Claustre2021-tx].
 
@@
 mangroves, tidal marshes, seagrass meadows) are thought to store in their living biomass (years to decades) and sediment (millenia) more than 30 Pg C with an estimated absorption of \~0.3 Pg C year$^{-1}$ [@Macreadie2021-rb] though those numbers should be taken with caution because there are major uncertainties in the distribution of seagrass meadows and tidal marshes [@Macreadie2021-rb] as well as in carbon burial rates [@Williamson2022-ud].
-In spite of that, the current expectation of additional carbon removal by coastal blue carbon by 2100 should be around 25-30 Pg C [@Gattuso2021-cr; @Williamson2022-ud], that is the equivalent of 2.5-3 years of current anthropogenic emissions [@Friedlingstein2022-XX].
+In spite of that, the current expectation of additional carbon removal by coastal blue carbon by 2100 should be around 25-30 Pg C [@Gattuso2021-cr; @Williamson2022-ud], that is the equivalent of 2.5-3 years of current anthropogenic emissions [@Friedlingstein2022-th].
 :::
 
-According to @Sarmiento2006-ec, the BCP would thus be responsible for the remaining 70% of the observed DIC gradient.
-However, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess atmospheric $CO_{2}$, released by anthropogenic activities.
+Estimates of global carbon export range between 5 and 12 Pg C year$^{-1}$ [@Dunne2007-mi; @Henson2011-mt; @Siegel2014-wg; @DeVries2017-wu; @Nowicki2022-no] with only 0.02-0.2 Pg C year$^{-1}$ reaching the sediment [@Dunne2007-mi; @Hayes2021-gg] where carbon will be sequestered for millions of years.
+Almost all of the exported organic carbon will therefore be remineralized and sequestered away from the atmosphere on time scales ranging from months to millennia [@Siegel2022-wd].
+
+Nonetheless, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess atmospheric $CO_{2}$, released by anthropogenic activities.
 First, as explained in @Sarmiento2006-ec, atmospheric $CO_{2}$ levels were stable for several thousand years before the industrial revolution started, therefore it is assumed that the BCP was operating at steady state.
@@
 
+<!-- he recent work of @Nowicki2022-no suggests that the BCP exports on average 10.2 Pg C year$^{-1}$, sequesters ~1,300 Pg C with a mean sequestration time of 127 years and buries 0.4 Pg C year$^{-1}$ in the sediment that will remain sequestered for millions of years.
+Therefore, when talking about natural-based solutions for carbon sequestration, one should keep in mind that even though the BCP does not contribute to the absorption of anthropogenic $CO_{2}$ per se, it remains the main process that maintains the DIC gradient -->
+
 In the frame of the BCP, global carbon budgets have generally been imbalanced such that the exported POC was up to 2-3 orders of magnitude less abundant that it would need to be to sustain heterotrophic respiration in the mesopelagic zone and beyond [@Burd2010-hc; @Turner2015-af].
@@
 Regarding the latter, @Boyd2019-ez published a paper in which they review the BCP by adding new biological and physical mechanisms[^intro-2] that inject suspended and sinking particles to depth (@fig-global_carbon_pumps).
-A description of those mechanisms is provided in the next subsections.
+A better understanding of those mechanisms (see next subsections) could help to close the carbon budget.
 
@@
 
-### Gravitational pump {#sec-BGP}
+### Biological gravitational pump  {#sec-BGP}
 
@@
 @Buesseler2006-og].
-More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size spectra (PSS) that could then be converted to carbon fluxes [e.g.
+More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size distributions (PSD) that could then be converted to carbon fluxes [e.g.
 @Guidi2008-vp].
@@
 
-The BGP was long thought to be the only component of the BCP therefore new instruments are being tested to decouple the downward POC flux signal from the contribution of other pumps (@sec-BGC-sensors).
+### Physically mediated particle injection pump {#sec-MixingPump}
 
-### Mixing pump {#sec-MixingPump}
-
-The mixing pump is a physically mediated particle injection pump (PIP).
-It consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez].
+The physically mediated particle injection pump consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez].
 Contrarily to the BGP, all three pumps also transport DOC to depth where it is injected along with suspended (i.e.
@@
 
-The mixed layer pump can be defined as a process through which carbon is exported below the EZ due to variation in the depth of the mixed layer [@DallOlmo2016-ex].
-The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on timescales of days across horizontal spatial scales of 1-10 km [@Omand2015-mf; @Boyd2019-ez].
-The large-scale (100-1000 km) subduction pump corresponds to the slow advective transport of carbon associated with the meridional overturning circulation and Ekman pumping [@Levy2013-ro; @Claustre2021-tx; @Siegel2022-wd].
+The mixed layer pump can be defined as a 1D process through which organic carbon is exported below the EZ when there is a deepening of the mixed layer followed by the formation of a new shallower mixed layer, isolating the exported carbon from the surface [@DallOlmo2016-ex].
+This pump can occur at different time scales but is mainly related to the building of the summer stratification in mid and high latitude regions where the seasonal variations of the mixed layer are the strongest [@DallOlmo2016-ex; @Boyd2019-ez].
+
+The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on timescales of days across horizontal spatial scales of 1-10 km \[@Omand2015-mf; @Resplandy2019-nx\].
+The large-scale (100-1000 km) subduction pump corresponds to the slow advective transport of carbon associated with the meridional overturning circulation and Ekman pumping \[@Levy2013-ro; @Claustre2021-tx; @Siegel2022-wd\].
 The last two pumps are both 2D spatial processes (i.e.
@@
 
-The mixing pump would represent at least 20% of the export flux according to @Nowicki2022-no although large uncertainties remain (see Figure 2C in the paper of @Boyd2019-ez).
+The physically mediated particle injection pump would represent at least 20% of the export flux according to @Nowicki2022-no although large uncertainties still remain (see Figure 2C in the paper of @Boyd2019-ez).
 
-### Migrant pump {#sec-MigrantPump}
+### Biologically mediated particle injection pump {#sec-MigrantPump}
 
-The migrant pump, also known as the vertical migration pump (VMP), is a biologically mediated PIP [@Boyd2019-ez].
+The biologically mediated particle injection pump is usually referred to as the vertical migration pump (VMP) [@Boyd2019-ez].
 It consists on the vertical migration of zooplankton or fish [@Steinberg2017-ri; @Saba2021-ew] that directly inject POC (e.g.
 fecal pellets) below the EZ at the depth of their migration.
-This pump can act on daily and seasonal timescales.
+This pump can act on daily as well as on seasonal timescales.
 In the case of the diel (i.e.
-daily) migration pump, animals feed at the surface at night and then migrate to depth at sunrise to avoid predation [@Siegel2022-wd], metabolizing injected food at depth[^intro-4], by-passing the EZ directly [@Steinberg2017-ri].
+daily) migration pump, animals feed at the surface at night and then migrate to depth during daytime to avoid visual predation [@Siegel2022-wd], metabolizing injected food at depth[^intro-4], by-passing the EZ directly [@Steinberg2017-ri].
 In the case of the seasonal migration pump, carbon-rich lipids accumulated in the EZ during summer are catabolized by hibernating copepods at depth, between 600 m and 1,400 m, where they shunt carbon below the permanent pycnocline [@Jonasdottir2015-ob].

### 2023-02-15 annexes.qmd
update

--- 2023-01-22 observingCarbonCycle.qmd
@@
-# Observing the carbon cycle {#sec-observing-carbon-cycle}
-
-## How it started {#sec-how-it-started}
-
-In the 20th century, the downward POC flux was traditionally measured with sediment traps.
-Those instruments are composed of one or several collectors of sinking material whose design depends on its use and position in the water column (e.g.
-moored traps, surface tethered traps, drifting traps) [@Honjo2008-kn].
-They were extensively used [@Honjo2008-kn] though their use has been steadily decreasing since 2004, reaching less than 500 measurements per year since 2010 (@fig-sediment-trap-thorium-yearly-distribution).
-
-```{r, poc measurements per year, dpi = 300}
-#| label: fig-sediment-trap-thorium-yearly-distribution
-#| fig-cap: Yearly distibution of POC flux measurements by sediment traps or derived from Thorium 234 deficits. Compilation based on @tbl-poc-from-sediment-traps and @tbl-poc-from-thorium.
-
-library(ggplot2)
-library(viridis)
-trap <- vroom::vroom('DATA/Trap_Thorium/TRAP_DATABASE.csv')
-thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
-trap$year <- lubridate::year(trap$start_date)
-thorium$year <- lubridate::year(thorium$start_date)
-sub_trap <- trap |> dplyr::select(year, instrument_type)
-sub_trap$instrument_type <- 'Sediment trap'
-sub_thorium <- thorium |> dplyr::select(year, instrument_type)
-sub_thorium$instrument_type <- 'Thorium 234'
-combined <- rbind(sub_trap, sub_thorium)
-plot <- ggplot(combined, aes(x = year, fill = instrument_type)) + geom_bar() + scale_fill_manual(values = c('#003366', '#DCEEF3')) + # #57cc99 #80ed99
-    #geom_vline(aes(xintercept = median(year)), color = "#d53e4f", linetype = "dashed") +
-    labs(x = "Year", y = "N records") + 
-    theme_bw() + 
-    #theme(legend.position = c(0.12, 0.85), legend.background = element_rect(fill = "white", color = "white")) + 
-    theme(legend.position = 'top') +
-    theme(legend.title=element_blank())
-
-# plot <- ggplot(combined, aes(x = year, fill = instrument_type)) + geom_bar() + scale_fill_manual(values = c('#E85D04', '#FFBA08')) + # #57cc99 #80ed99
-#     #geom_vline(aes(xintercept = median(year)), color = "#d53e4f", linetype = "dashed") +
-#     labs(x = "Year", y = "N records") + theme_bw() + guides(fill=guide_legend(title="Method"))
-
-plot
-```
+# Annexes
 
 ```{=html}
-<!-- 
-![Examples of sediment traps.
-From left to right: (a) Neutrally buoyant, (b) Surface tethered and (c) Funnel-shaped drifting trap.
-Figure taken from @Baker-XX.](image/ObservingCarbonCycle/sed_traps_high_res.jpg){#fig-sed-traps}
--->
-```
-```{=html}
-<!-- #https://github.com/quarto-dev/quarto-cli/discussions/1042 -->
+<!--#https://github.com/quarto-dev/quarto-cli/discussions/1042-->
 <style type="text/css">
@@
 ```
+
 ```{r, poc database from traps}
@@
 ```
-
-To estimate the downward POC flux, one therefore needs to quantify the amount of POC collected through a given area over a known time period at a given depth.
-However, the apparent sinking flux measured with sediment traps can be biased due to hydrodynamic issues that could result in over- or under-collection of particles (especially at the surface), swimmers (zooplankton, small fish) that could enter sediment traps and affect the collected material and solubilization of collected material (e.g.
-by bacteria) [@Buesseler2007-ql].
-
-Given the potential inaccuracy of flux measurements by sediment traps, independent approaches using radionuclide disequilibria have been used to assess the collection efficiency of the latter [@Buesseler1991-vo].
-This method is based on the fact that $^{238}U$ is a conservative element in seawater (chemically inert with a half-life of 4.5 billion years) while its daughter nuclide $^{234}Th$ (half-life of 24.1 days) can be adsorbed onto particle surfaces and therefore be scavenged by sinking particles [@Savoye2006-gb].
-Consequently, the deficiency of $^{234}Th$ relative to $^{238}U$ is used as a measure of removal of surface $^{234}Th$ through particle scavenging [@Buesseler1991-vo] that is then derived as a flux of exported $^{234}Th$ and subsequently a flux of exported POC by using a known (e.g.
-estimated from the material collected in sediment traps) POC/$^{234}Th$ ratio [@Buesseler2007-ql].
-
-While this method seems to cope with some of the issues observed in sediment traps, it also has its caveats, particularly a POC/$^{234}Th$ ratio that varies with particle size and the fact that it only works well in the upper 100-200 m [@Buesseler2007-ql].
-
-In all, POC fluxes derived from sediment traps and $^{234}Th$ have their own uncertainties and errors thus their estimations of downward POC fluxes should not be seen as ground truth.
-In addition, they are labor-intensive and require ship support.
-Their relative contribution to the estimation of POC fluxes is thus expected to keep decreasing (@fig-sediment-trap-thorium-yearly-distribution) with the advent of AUVs (see @sec-how-it-has-evolved).
-
-## How it has evolved {#sec-how-it-has-evolved}
-
-It is well known that ship-based measurements are the most precise.
-However, they lack spatiotemporal coverage and they are usually not cost-effective.
-Consequently, oceanographers have turned to new platforms and have developed new sensors to study the marine carbon cycle and the BCP in particular.
-
-Instrumented moorings have supplemented ship-based measurements by collecting high-resolution data at fixed locations [@Chai2020-uq].
-Moorings can be equipped with many sensors (e.g.
-sediment traps (@sec-how-it-started) or imaging systems to study the type of sinking particles (@sec-BGC-sensors)) to study carbon-related variables (see Table 1 in @Claustre2021-tx).
-They are therefore ideal platforms for long-term time series (e.g.
-Bermuda Atlantic Time-series Study (BATS) since 1988) with the disadvantage of limited spatial coverage and high maintenance costs [@Chai2020-uq].
-
-In contrast, satellite remote sensing have a wide spatial coverage but they are limited to the ocean's surface and their measures can be affected by several factors such as the cloud cover or low solar elevation (e.g.
-@Marshall1994-vk).
-They have been used for ocean applications since the late 1970s [@Gordon1980-dm] and have been widely used to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few.
-
-In order to fill the scarcity of data from research vessels and moorings as well as coping with surface-constrained satellite measurements, AUVs have been developed to take *in situ* measurements in all weathers and at all times, only being dependent on their battery life and ocean currents (for drifting AUVs).
-
-Biogeochemical Argo (BGC-Argo) profiling floats (@fig-bgc-float) are probably the most well known AUVs.
-They are cost-effective, able to monitor biogeochemical processes at a global scale throughout the year and under all weather conditions [@Chai2020-uq].
-The goal of the BGC-Argo community is to develop an array of 1000 floats that measure 6 key biogeochemical and bio-optical variables: Chlorophyll a (Chla), particle backscatter ($b_{bp}$), oxygen, nitrate, pH and downwelling irradiance in addition to temperature and salinity [@Roemmich2019-pn].
-Besides this objective, the BGC-Argo community is testing new sensors (see @sec-BGC-sensors) to explore specific topics related to the BCP [@Claustre2020-jo].
-
-Gliders are also effective AUVs to unveil the mysteries of the BCP [e.g.
-@Omand2015-mf].
-In comparison with BGC-Argo floats (@fig-BGC-Argo-cycle), gliders mostly operate in shallow waters [@Chai2020-uq] and are able to take measurements along a predefined transect due to their ability to change their internal weight distribution [@Johnson2009-kb].
-
-![Illustration of a typical BGC-Argo 10-day cycle.
-From the beginning of the cycle (1), the profiling float descends to its parking depth (2) then drifts for \~ 9 days (3) to dive at 2000 m (4) before the ascent where data are acquired (5) and transmitted to satellite while at the surface (6).
-Note that the profiling float can be programmed (and reprogrammed during its mission via Iridium transmission) with different configurations, depending on its utilization.
-For the study of the BCP, some floats could be configured to drift at multiple parking depths to study the BGP or to shorter cycle duration to study the VMP.
-Image taken from @Claustre2020-jo.](image/ObservingCarbonCycle/argo_float_profiling.jpeg){#fig-BGC-Argo-cycle}
-
-The study of the BCP is a really hot topic in oceanography with new platforms (e.g.
-unmanned surface vehicles such as saildrones or wave gliders) and new technologies (e.g.
-lidar) being continuously developed in the hope of understanding all its components [@Chai2020-uq; @Claustre2021-tx].
-
-## Specific sensors for the study of the BCP {#sec-BGC-sensors}
-
-In the frame of the REFINE (Robots Explore plankton-driven Fluxes in the marine twIlight zoNE) project, enhanced BGC-Argo floats (@fig-bgc-float) will have two new sensors specifically designed for the study of the BCP: a transmissometer and the Underwater Vision Profiler 6 (UVP6) [@Picheral2021-sx].
-
-![Illustration of a BGC-Argo float used in the frame of REFINE.
-From top to bottom: GPS & Iridium (two-way communication) antenna, Oxygen optode, CTD (salinity, temperature and depth), pH sensor, radiometer (downwelling photosynthetic available radiation and downwelling irradiance at 380, 412 and 490 nm), fluorometer (Chla and colored dissolved organic matter) and scattering meter ($b_{bp}$), UVP6 (particle size spectra and images), transmissometer (particle beam attenuation) and nitrate sensor.
-\@Thomas Boniface.](image/ObservingCarbonCycle/float_picture1.png){#fig-bgc-float width="500"}
-
-A transmissometer measures the transmittance ($T_{r}$) of a light beam at a given wavelength. Given $T_{r}$, the beam attenuation coefficient ($c$) is computed as
-
-$$
-c = \frac{ln(T_{r})}{r}
-$$ {#eq-bean-attenuation}
-
-where $r$ is the transmissometer pathlength.
-
-The value of $c$ is the result of absorption and scattering by particles, water and dissolved substances [@Zaneveld1994-ez].
-The transmissometer carried by REFINE BGC-Argo floats uses a light beam at 660 nm whose attenuation coefficient can be splitted into three components: the absorption and scattering of light by particles ($c_{p}$), the absorption by water ($c_{w}$) and the absorption by dissolved organic matter ($c_{DOM}$) [@Jerlov1976-hr].
-At 660 nm, $c_{w}$ is assumed constant at 0.364 m$^{-1}$ and $c_{DOM}$ is considered negligible [@Jerlov1976-hr] hence variations of $c$ will be assumed to be variations of $c_{p}$ in what follows.
-
-The use of transmissometers to characterize POC concentrations was shown by @Bishop1999-es who found that POC concentrations were well correlated with $c_{p}$, independently of ocean environment, season or depth.
-Following that, the idea of using a transmissometer as an Optical Sediment Trap (OST) was firstly applied by @Bishop2004-at where the magnitude of carbon export was quantified based on the particles accumulation on the upward-facing window of vertically oriented transmissometers on autonomous platforms while they were drifting (i.e.
-parked at a specific depth).
-This method was later improved by [@Estapa2013-gw] who divided the attenuation signal into a continuous component, that is the continuous accumulation of small particles on the transmissometer window and a discontinuous component resulting in positive jumps likely attributed to the deposition of (rarer) large particles (@fig-estapa-method).
-@Estapa2017-jr further calibrated the OST carbon flux proxy using codeployed sediment traps and drifting transmissometers thus allowing the conversion of $c_{p}$ attenuation fluxes (units: m$^{-1}$d$^{-1}$) into carbon fluxes (units: mg C m$^{-2}$d$^{-1}$) using empirical POC-to-attenuance (ATN) ratio were ATN (unitless) is the product of $c_{p}$ and the transmissometer pathlength.
-
-```{r, OST on BGC-Argo example with jump}
-#| label: fig-estapa-method
-#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component, cleaned from spikes and negative jumps (assumed as not related to carbon flux) and a discontinuous (positive jumps, light blue) component.
-Data from BGC-ARGO float WMO 4903634, drifting at 1000 m during its 9$^{th}$ cycle.
-library(tidyverse)
-library(ggplot2)
-library(vroom)
-#library(castr)
-library(plyr)
-library(scales)
-library(lubridate)
-library(latex2exp)
-
-# From Louis Terrats
-clean_cp_data <- function(file, moving_median_order, threshold) {
-
-  # Store raw data 
-  raw_file <- file
-
-  #print(head(file))
-
-  # Remove cp > 2 m-1 
-  if (any(file$cp > 2)) {
-    file <- file[-which(file$cp > 2),]
-  }
-
-  # Store relevant variables in vectors (for speeding up the process)
-  cp <- file$cp
-  diff_cp <- diff(cp)
-  dates <- file$dates
-
-  # Spike removal
-  # Spike is identified as a significant increase (threshold = 0.01 m-1 h-1) followed by a return to normal values within the next 3 hours.
-  tmp <- {diff(cp) / as.numeric(diff(dates), units = "hours")} %>% abs()
-  spikes <- which(tmp > threshold)+1 # why +1?
-  if (length(spikes) > 0) {
-    # For each spike
-    for (j in 1 : length(spikes)) {
-
-      # Get the value measured 3 hours after the spike (= post_spike)
-      spike <- spikes[j]
-      time_diff <- difftime(dates, dates[spike], unit = "hours") 
-      post_spike <- which(time_diff >= 3)[1]
-
-      # If the difference between the spike and the post_spike cp is superior to the threshold : values get back to normal ones
-      if (is.na(cp[spike])) {next 
-      } else {
-        if ({!is.na(post_spike)} && {(cp[spike]-cp[post_spike]) > threshold}) { 
-          back_to_normal <- which({(cp - cp[post_spike]) < threshold} & {dates > dates[spike]})[1] # Locate the date of back to normal values 
-          cp[(spike):(back_to_normal-1)] <- NA # And remove the spike zone
-        } else { 
-          if (is.na(post_spike)) { # If post spike is Na, that means spike is near the end of the segment
-            cp[spike:length(cp)] <- NA # Remove the end zone (which contains a spike)
-          }
-        }
-      }
-    }
-  }
-
-  # Store the sum of spikes 
-  spikes <- spikes[which(spikes %in% which(is.na(cp)))]
-  spikes <- diff_cp[spikes-1] %>% abs() %>% sum(na.rm = TRUE)
-
-  # Remove rows with NA (= remove spike zones)
-  if (any(is.na(cp))) {
-    file <- file[-which(is.na(cp)),]
-  }
-
-  # Moving median
-  if (nrow(file) > 0) {
-    file$cp <- pastecs::decmedian(file$cp, order = moving_median_order, ends = "fill") %>% pastecs::extract(component = "filtered")
-  } else {return(list("data" = file, "plot" = NA, "spikes" = spikes))}
-
-  # Plot showing results and previous data
-  plot <- ggplot() + 
-    geom_point(data = raw_file, aes(x=dates, y=cp, color = "black"), alpha = 0.5, size = 4) + 
-    geom_point(data = file, aes(x=dates, y=cp, color = "blue"), size = 4) +
-    scale_x_datetime(labels = date_format("%d-%m@%Hh")) + 
-    labs(y = expression(c[p]~"["*m^-1*"]"), x = "Datetime") + 
-    scale_colour_manual(name = '', values =c('black'='black','blue'='blue'), labels = c('Raw data','Filtered signal')) +
-    # Theme
-    theme_bw() +
-    theme(text = element_text(size=16), 
-          legend.text=element_text(size=16), 
-          legend.position = 'right',
-          axis.text = element_text(color="black", size = 16),
-          plot.title = element_text(face="bold", hjust = 0.5))
-
-  return(list("data" = file, "plot" = plot, "spikes" = spikes)) 
-}
-
-# From me
-float_4903634 <- vroom::vroom('/home/flo/dox/THESIS/DATA/ARGO/4903634_OMTAB_VERSION.csv')
-profile <- 'lovuse017c_016_01'
-drift_depth <- '1000 m'
-
-check <- float_4903634 |> select(dates = juld, everything()) |> filter(cycle == profile, park_depth == drift_depth) |> drop_na(cp)
-despiked_data <- clean_cp_data(check, 3, 0.01)
-
-# save cleaned Cp signal from spikes to a temporary variable
-tmp2 <- despiked_data$data
-# init jump values to NA
-tmp2$jump <- NA
-# compute slope between two adjacent points (except first and last point of drifting time series)
-for (i in 2:(nrow(tmp2)-1)){
-  #print(i)
-  delta_x <- as.numeric(difftime(tmp2$dates[i], tmp2$dates[i-1], units = 'days'))
-  delta_y <- tmp2$cp[i]-tmp2$cp[i-1]
-  tmp2$jump[i] <- delta_y/delta_x
-}
-
-# assign a slope threshold depending on the drifting depth (this was estimated based on all 4 floats, could be refined later but works pretty well)
-if(drift_depth == '200 m'){
-  threshold <- 0.4
-}else if(drift_depth == '500 m'){
-  threshold <- 0.22
-}else{
-  threshold <- 0.177
-}
-
-# assign a colour for the plot
-tmp2$colour <- 'cleaned signal'
-tmp2$colour[which(abs(tmp2$jump)>threshold)] <- 'jump' # abs is needed for 'negative jumps'
-
-# add group to compute slope (for each subgroups of points, separated from a jump)
-tmp2$group <- NA
-
-# compute where the jumps are (index value of the array)
-jump_index <- which(tmp2$colour == 'jump')
-
-# assign group 
-for (i in jump_index){
-  for (j in 1:nrow(tmp2)){
-    if ((j < i) & (is.na(tmp2$group[j]))){
-      tmp2$group[j] <- paste0('group_',i)
-    }
-  }
-}
-
-# add last group (because it is AFTER the last index so it was not computed before)
-tmp2$group[which(is.na(tmp2$group))] <- 'last_group'
-
-# compute slope for each subgroups
-slope_df <- tmp2 |> filter(colour == 'cleaned signal', jump != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), 
-                                                                                                                nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],
-                                                                                                                delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),
-                                                                                                                delta_y = last_cp-first_cp, slope = delta_y/delta_x)
-
-# remove if only one point (cannot fit a slope with one point)
-slope_df <- slope_df |> filter(nb_points > 1)
-
-# compute mean slope
-mean_slope <- mean(slope_df$slope)
-
-# estapa relationship (see POSTER)
-poc_flux <- 633*(mean_slope**0.77)
-
-# build dataframe to plot each subgroup
-part1 <- slope_df |> select(group, time = min_time, cp = first_cp)
-part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
-part_slope <- rbind(part1, part2)
-
-# spot negative jump
-tmp2$colour[which((tmp2$colour == 'jump') & (tmp2$jump < 0))]  <- 'negative jump'
-
-#jump_plot <- ggplot(clean_data, aes(x = dates, y = cp)) + geom_point() + geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour))
-
-# ggplot(despiked_data$data, aes(x = dates, y = cp)) + geom_step(size= 3, fill ='black') + 
-#   geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour), size = 3) + 
-#   scale_colour_manual(values = c('black','yellow', 'blue')) +
-#   theme_bw() +
-#   ylab(TeX('$c_{p}~(m^{-1})$')) +
-#   xlab('Datetime') +
-#   theme(legend.position = c(0.095, 0.875),
-#         legend.background = element_rect(fill = "white", color = "black"))
-
-jump <- filter(tmp2, colour == 'jump')
-
-jump_plot <- ggplot(tmp2, aes(x = dates, y = cp, fill = colour)) + geom_point(size= 3) +
-  geom_point(data = jump, aes(x = dates, y = cp), size = 3, shape = 21) +
-  scale_fill_manual(values = c('#003366','#DCEEF3')) +
-  theme_bw() +
-  ylab(TeX('$c_{p}~(m^{-1})$')) +
-  xlab('Datetime') +
-  theme(legend.position = 'top', legend.title = element_blank())
-
-# ggplot(despiked_data$data, aes(x = dates, y = cp)) + geom_point(size= 3, fill ='black') +
-#   geom_point(data = jump, aes(x = dates, y = cp), size = 3, shape = 21, fill = '#DCEEF3') +
-#   scale_colour_manual(values = c('black','yellow', 'blue')) +
-#   theme_bw() +
-#   ylab(TeX('$c_{p}~(m^{-1})$')) +
-#   xlab('Datetime') +
-#   theme(legend.position = 'top')
-# 
-# jump_plot <- ggplot(base_signal, aes(x = dates, y = cp)) + geom_point(size = 3, shape = 21, fill = '#003366') +
-#   geom_point(data = jump, aes(x = dates, y = cp), size = 3, shape = 21, fill = '#DCEEF3') +
-#   theme_bw() +
-#   ylab(TeX('$c_{p}~(m^{-1})$')) +
-#   xlab('Datetime')
-
-#geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = 'red', size = 2) + theme_bw() + 
-#labs(x = 'Time', y = 'Cp (1/m)', title = paste0('Mean slope (red): ', round(mean_slope,3), ' m-1 day-1')) # \n', 'POC flux (small particles): ', poc_flux, 'mg'))
-
-jump_plot
-
-```
-
-The UVP6 (@fig-UVP6) is a miniaturized version of the UVP5 [@Picheral2021-sx], an imaging system developed to characterize particles (e.g.
-size, gray level, taxonomic group).
-Two versions of the UVP6 have been commercialized: the UVP6-HF (high frequency) and the UVP6-LP (low power) which has been specifically designed for AUVs that have strict energy requirements (e.g.
-autonomous floats).
-
-The UVP6-LP (hereafter referred to as UVP6) is thus both a particle counter (0.1 mm - 2 mm) and a particle classifier (\> 0.6 mm).
-The former counts particles in different class sizes in order to compute the particle size distribution (PSD) - also referred to as PSS (@sec-BGP) - while the latter extracts pixel-related and geometric features from each image to classify them (i.e.
-assign them a taxonomic group) directly on board using artificial intelligence (CHAPTER-XX) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].
-
-![Illustration of the basic functioning of the UVP6.
-A light beam illuminates a region in front of the camera (imaged volume: 0.7L) which takes pictures at a given temporal resolution and subsequently characterizes and classifies relevant objects.
-\@Thomas Boniface.](image/ObservingCarbonCycle/UVP6_b.png){#fig-UVP6}
-
-While the gathering of data from automatic on board particles classification will have to increase to study related applications, the use of PSD in carbon flux studies has already been thoroughly explored [@Guidi2008-vp; @Iversen2010-an; @Jouandet2011-bt; @McDonnell2012-ic; @Roullier2014-wf; @Guidi2015-sj; @Guidi2016-nj; @Ramondenc2016-nu; @Kiko2017-qy; @Fender2019-ou; @Giering2020-mc; @Cael2021-qn; @Bisson2022-bt].
-
-Using the PSD, one can compute the number spectrum $n_{i}$ [@Stemmann2004-jv] which can be defined as the normalized concentration of particles in a given size bin $i$.
-Knowing the mass ($m$) and sinking speed ($w$, see @eq-sinking-speed) of aggregates, one can derive the total carbon flux ($F$)
-
-$$
-F = \int_{0}^{\infty} n(d) m(d) w(d) dd
-$$ {#eq-carbon-flux}
-
-where $d$ is the aggregate diameter.
-
-Assuming spherical aggregates, $m$ can be written as follow
-
-$$
-m(d) = \frac{\pi\rho d^{3}}{6}
-$$ {#eq-mass-aggregate}
-
-where $\rho$ is the average density of aggregates.
-
-As explained in @sec-sinking-speed, aggregates have fractal geometries ($D$) hence @eq-sinking-speed and @eq-mass-aggregate can be rewritten as follow
-
-$$
-m(d) = \alpha d^{D}
-$$ {#eq-mass-aggregate-fractal}
-
-$$
-w(d) = \beta d^{D-1}
-$$ {#eq-sinking-speed-fractal}
-
-where $\alpha$ and $\beta$ are constant.
-
-Aggregate mass and sinking speed are often described using a power law relationship [@Alldredge1988-lx; @Guidi2008-vp] therefore it can be derived that
-
-$$
-mw = Ad^{B}
-$$
-
-and therefore from @eq-mass-aggregate-fractal and @eq-sinking-speed-fractal that $A= \alpha\beta$ and $D = \frac{B+1}{2}$.
-
-Due to a finite number ($N$) of size bins, @eq-carbon-flux is rewritten as
-
-$$
-F = \sum_{i=1}^{N}n_{i}Ad_{i}^{B}\Delta d_{i}
-$$ {#eq-carbon-flux-finite}
-
-where $d_{i}$ is the mid-point of bin $i$.
-
-UVP6 PSD measurements will therefore allow the computation of carbon fluxes assuming that $A$ and $B$ are known. This was done empirically by @Guidi2008-vp who minimized the following cost function
-
-$$
-Cost = \sum_{j = 1}^{n} (log(F_{UVP,j})-log(F_{ST},j))^{2}
-$$ {#eq-cost-function}
-
-where $F_{UVP,j}$, $F_{ST,j}$ and $n$ are, respectively, the flux estimate for the $j^{th}$ sediment trap measurement, the flux measured in sediment trap $j$ and the number of paired measurements (i.e.
-match-ups) for trap-UVP comparisons.
-
-@Guidi2008-vp found an optimal couple of parameters $(A, B) = (12.5, 3.81)$ while @Iversen2010-an followed the same procedure and found $(A, B) = (273, 4.27)$ whereas @Fender2019-ou found $(A, B) = (15.4, 1.05)$.
-@Iversen2010-an assumed that the difference in the optimal couple of parameters was due to the study site used for the minimization (productive upwelling system versus open ocean in @Guidi2008-vp) while @Fender2019-ou suggested that a value of $B < 3$ means that either the density or sinking speed of particles is inversely correlated with particles diameter thus they suspected that the flux was dominated by particles in the lower size detection limit of the UVP they were using.
-Other values for $A$ and $B$ are being used in the literature [e.g.
-@Kiko2017-qy] based on a mechanistic approach developed by @Kriest2002-oc to investigate the formation of marine aggregates.
-
-In all, it seems that the choice for the values of $A$ and $B$ is critical if one does not want to under- or overestimate carbon fluxes derived from PSD [@Iversen2010-an; @Fender2019-ou].
-Three main limitations highlighted by @Bisson2022-bt for UVP optically derived estimates of flux would be: (1) assuming that $A$ and $B$ are globally valid at all depths, for all UVP models and across all size classes, (2) the size to sinking rate uncertainty and (3) the size to carbon mass uncertainty..
-
-The assumption that all particles of a given size class have the same carbon content and settling rate questions the formulation of the model itself (@eq-carbon-flux-finite).
-It has been shown that the pool of sinking particles is varied (e.g.
-compact fecal pellets, porous aggregates) and that the main driver of the sinking speed is currently subject to debate [@sec-sinking-speed].
-Therefore, the size-dependent model of [@Guidi2008-vp] could be improved by taking into account the composition and nature of particles and by increasing the number of paired measurements between UVP PSD-derived flux estimations and other direct POC flux estimations (e.g.
-sediment traps, $^{234}Th$, OST).
-
-## The future of carbon observations {#sec-what-is-next}
-
-The BCP have been studied for many decades by a variety of instruments with their respective pros and cons.
-Their measurements were unfortunately often taken independently, which hindered their intercomparability due to the low number of concurrent measurements between instruments.
-
-With the advent of fully-equipped AUVs, it is now possible to take high-resolution and independent measurements of key biogeochemical variables.
-Combining the latter with measurements from conventional platforms (i.e.
-ships, satellites, moorings) will fulfill the observational framework described by @Claustre2021-tx to investigate the BCP.
-
-This framework will hopefully answer questions raised so far, that is better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and better quantify global estimates of carbon storage.

### 2023-02-15 intro.qmd
update

@@
 @DeVries2022-of] that if we were to consider the ocean as a homogeneous reservoir of inorganic carbon, $pCO_{2}^{atm}$ would be much higher than it is now (roughly 416 parts per million (ppm) as of October 2022 according to the Mauna Loa Observatory).
-This observation is explained by the presence of a strong vertical gradient of DIC in the ocean, with surface DIC concentrations more than 200 $\mu$mol kg$^{-1}$ lower than at 1000 m as it can be seen in @fig-DIC_plot.
+This observation is explained by the presence of a strong vertical gradient of DIC in the ocean, with surface DIC concentrations more than 200 $\mu$mol kg$^{-1}$ lower than at 1,000 m as it can be seen in @fig-DIC_plot.
 
@@
 
-Also known as the soft-tissue pump [@Volk1985-oc], the biological carbon pump (BCP), is defined by @Ipcc2019-qc as "a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis (@eq-photosynthesis) in sunlit surface water and then transported to the ocean interior (i.e.
-exported flux), and possibly the sediment, resulting in the storage of carbon".
+Also known as the soft-tissue pump [@Volk1985-oc], the biological carbon pump (BCP), is defined by @Ipcc2019-qc as "a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis (@eq-photosynthesis) in sunlit surface water and then transported to the ocean interior, and possibly the sediment, resulting in the storage of carbon".
 
@@
 
-In order to estimate the carbon export flux, one needs to set a depth horizon [@Siegel2022-wd].
-Historically, the latter has often been set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should now use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where particulate organic carbon (POC) is produced (i.e.
+The carbon flux leaving the surface layer for the ocean interior is referred to as the export flux [@Siegel2022-wd].
+Historically, the export depth was usually set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should now use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where particulate organic carbon (POC) is produced (i.e.
 via photosynthesis) and where POC is consumed (i.e.
@@
 This remineralization (or respiration) consists in the conversion of organic matter back to inorganic forms (i.e.
-nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd].
+nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd] such as zooplankton or fish.
 
-According to @Sarmiento2006-ec, the BCP would thus be responsible for the remaining 70% of the observed DIC gradient.
+Those two biological activities significantly enhance the DIC gradient such that the BCP would thus be responsible for the remaining 70% of the observed DIC gradient [@Sarmiento2006-ec].
 Without it, there would be no production of organic matter in the EZ (consuming DIC) and no remineralization in the ocean interior (producing DIC) resulting in a weaker DIC gradient.
@@
 Similarly, @Kwon2009-ir showed that a deepening of the remineralization depth (i.e.
-where 63% of sinking carbon is respired) would decrease atmospheric $CO_{2}$ concentrations due to the redistribution of respired carbon from intermediate waters to bottom waters, enhancing the DIC gradient.
+where 63% of sinking carbon is respired) would decrease atmospheric $CO_{2}$ concentrations due to the redistribution of respired carbon from intermediate waters to bottom waters (i.e.
+increased DIC gradient).
 
@@
 
-Estimates of global carbon export range between 5 and 12 Pg C year$^{-1}$ [@Dunne2007-mi; @Henson2011-mt; @Siegel2014-wg; @DeVries2017-wu; @Nowicki2022-no] with only 0.02-0.2 Pg C year$^{-1}$ reaching the sediment [@Dunne2007-mi; @Hayes2021-gg] where carbon will be sequestered for millions of years.
-Almost all of the exported organic carbon will therefore be remineralized and sequestered away from the atmosphere on time scales ranging from months to millennia [@Siegel2022-wd].
+Estimates of global carbon export range between 5 and 12 Pg C year$^{-1}$ [@Dunne2007-mi; @Henson2011-mt; @Siegel2014-wg; @DeVries2017-wu; @Nowicki2022-no] with only 0.02-0.4 Pg C year$^{-1}$ reaching the sediment [@Dunne2007-mi; @Hayes2021-gg; @Nowicki2022-no] where carbon will be sequestered for millions of years.
+Almost all of the exported organic carbon will therefore be remineralized and sequestered away from the atmosphere on time scales ranging from months to millennia.
+[@Siegel2022-wd].
 
@@
 
-The physically mediated particle injection pump consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez].
+The physically mediated particle injection pump (also referred to as the mixing pump) consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez].
 Contrarily to the BGP, all three pumps also transport DOC to depth where it is injected along with suspended (i.e.
@@
 
-The mixed layer pump can be defined as a 1D process through which organic carbon is exported below the EZ when there is a deepening of the mixed layer followed by the formation of a new shallower mixed layer, isolating the exported carbon from the surface [@DallOlmo2016-ex].
-This pump can occur at different time scales but is mainly related to the building of the summer stratification in mid and high latitude regions where the seasonal variations of the mixed layer are the strongest [@DallOlmo2016-ex; @Boyd2019-ez].
+The mixed layer pump can be defined as a 1D mechanism through which organic carbon is exported below the EZ when there is a deepening of the mixed layer followed by the formation of a new shallower mixed layer, isolating the exported carbon from the surface [@DallOlmo2016-ex].
+This pump can occur at different time scales but it is mainly related to the building of the summer stratification in mid and high latitude regions where seasonal variations of the mixed layer are the strongest [@DallOlmo2016-ex; @Boyd2019-ez].
 
-The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on timescales of days across horizontal spatial scales of 1-10 km \[@Omand2015-mf; @Resplandy2019-nx\].
-The large-scale (100-1000 km) subduction pump corresponds to the slow advective transport of carbon associated with the meridional overturning circulation and Ekman pumping \[@Levy2013-ro; @Claustre2021-tx; @Siegel2022-wd\].
-The last two pumps are both 2D spatial processes (i.e.
+The two other pumps are both 2D spatial mechanisms (i.e.
 they transport POC and DOC vertically and laterally).
+The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on timescales of days across horizontal spatial scales of 1-10 km [@Omand2015-mf; @Resplandy2019-nx].
+The large-scale (100-1000 km) subduction pump corresponds to the slow advective transport of carbon associated with the meridional overturning circulation and Ekman pumping [@Levy2013-ro; @Claustre2021-tx; @Siegel2022-wd].
 
-The physically mediated particle injection pump would represent at least 20% of the export flux according to @Nowicki2022-no although large uncertainties still remain (see Figure 2C in the paper of @Boyd2019-ez).
+In all, the mixing pump would represent at least 20% of the export flux according to @Nowicki2022-no although large uncertainties still remain (see Figure 2C in the paper of @Boyd2019-ez).
 
@@
 
-Based on @Nowicki2022-no, a lower boundary for the migrant pump export flux could be approximately 10% of the total export taking into account the fact that the seasonal lipid pump was ignored in their model as well as the vertical migration of mesopelagic fish that could export a significant amount of carbon (e.g.
-1.5 Pg C year$^{-1}$ according to @Saba2021-ew).
-If the export of carbon by the migrant pump is somehow lower than the mixing pump [@Nowicki2022-no], its sequestration potential is higher due to its deeper injection depth (@tbl-PumpDepthRange).
-According to @Boyd2019-ez, the diel migration pump could potentially store 60% as much carbon as the BGP due to the rapid sinking of fecal pellets that would reach depths such that subsequent remineralized DIC would be hidden from the surface for a long period (\> 100 years) of time (@tbl-PumpDepthRange).
+In his model, @Nowicki2022-no estimates that the VMP accounts for approximately 10% of the total export flux, albeit large uncertainties remain on the number of migrating animals and their migration depth.
+The VMP would thus export less organic carbon than the mixing pump but its sequestration potential would be higher (@tbl-PumpDepthRange).
+In contrast to the mixing pump that mostly acts on DOC, migrating animals release rapid sinking fecal pellets that would reach depths such that subsequent remineralized DIC would be hidden from the surface for a longer period of time (@tbl-PumpDepthRange).
+As a result, it is estimated that the diel migration pump could potentially sequester 60% as much carbon as the BGP [@Boyd2019-ez].
 
-However, there are still many uncertainties regarding the magnitude of the migrant pump [@Archibald2019-dj; @Aumont2018-wx; @Gorgues2019-qp] and its sequestration potential [@Boyd2019-ez; @Pinti2022-dc] thus its importance for a better understanding of the BCP.
+However, there are still many uncertainties regarding the magnitude of VMP [@Archibald2019-dj; @Aumont2018-wx; @Gorgues2019-qp] and its sequestration potential [@Boyd2019-ez; @Pinti2022-dc] thus its importance for a better understanding of the BCP.
 
@@
 
-Marine snow is paramount for both the BCP (see @sec-a-new-paradigm-for-the-bcp) and the support of marine life below the EZ [@Burd2010-hc].
+Marine snow is paramount for both the BCP (see @sec-a-new-paradigm-for-the-bcp) and the support of marine life within the mesopelagic domain [@Burd2010-hc].
 The remineralization of organic matter by heterotrophic organisms releases $CO_{2}$ in the water column that will then be released back to the atmosphere in a time frame depending both on its geographical location and depth (see @sec-will-it-be-sequestered).
@@
 
-The Martin curve [@Martin1987-um] is probably the most well known POC attenuation curve (@eq-Martin-Curve and @fig-Martin-Curve). It is a power-law function described as follow
+The Martin curve [@Martin1987-um] is probably the most well known POC flux attenuation curve (@eq-Martin-Curve and @fig-Martin-Curve). It is a power-law function described as follow
 
@@
 where $F_{z}$, $z_{export}$ and $b$ are, respectively, the estimated POC flux at a depth $z$, the depth of the export flux and the remineralization factor.
-The latter is adimensional and often set at 0.86 [@Martin1987-um] although it is well known that the magnitude of flux attenuation varies regionally (and temporally) due to the variability in food web structures beneath the EZ [@Passow2012-zh] thus variable $b$ have been proposed to take that into account [e.g.
-@Henson2012-ck; @Guidi2015-sj].
+The latter is adimensional and often set at 0.86 [@Martin1987-um] although it is well known that the magnitude of flux attenuation varies regionally (and temporally) due to the variability in food web structures beneath the EZ [@Passow2012-zh].
+Geographically variable [@Henson2012-ck] and regionalized [@Guidi2015-sj] $b$ values have been proposed to take that into account.
 
@@
 
-Based on the Martin curve, @Iversen2022-qj divided the water column in 4 layers (@fig-Martin-Curve): the upper mesopelagic layer (export depth - 300 m) where there is a rapid attenuation of the downward POC flux, the mid-mesopelagic layer (300 - 500 m) and the lower mesopelagic layer (500 - 1000 m) where both layers show two levels of intermediate attenuation and the bathypelagic layer (1000 m - seafloor) where the attenuation is relatively uniform.
+Based on the Martin curve, @Iversen2022-qj divided the water column in 4 layers (@fig-Martin-Curve): the upper mesopelagic layer (export depth - 300 m) where there is a rapid attenuation of the downward POC flux, the mid-mesopelagic layer (300 - 500 m) and the lower mesopelagic layer (500 - 1,000 m) where both layers show two levels of intermediate attenuation and the bathypelagic layer (1,000 m - seafloor) where the attenuation is relatively uniform.
 For each layer, @Iversen2022-qj estimated a carbon-specific degradation rate in order to better understand the processes responsible of this attenuation.
-As a result, @Iversen2022-qj suggests that the POC attenuation in the upper mesopelagic layer is the combined action of microbial degradation and zooplankton that graze on settling aggregates, the latter being the main responsible for the attenuation whereas in the mid- and lower mesopelagic layer, the microbial degradation takes over due to the decrease in zooplankton abundance.
+As a result, @Iversen2022-qj suggests that the POC attenuation in the upper mesopelagic layer is the combined action of microbial degradation and zooplankton that graze on settling aggregates.
+In the upper mesopelagic, zooplankton are the main responsible for the attenuation whereas in the mid- and lower mesopelagic layer, the microbial degradation takes over due to the decrease in zooplankton abundance.
 Finally, the relatively constant degradation occurring below 1,000 m in the bathypelagic zone could be explained by the increasing pressure and decreasing temperatures on microbial communities.
 
-The impact on the downward POC flux of zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2022-qj]: aggregate feeding, filter-feeding and direct ingestion of fecal pellets will decrease the POC flux, ingestion of slow-sinking cells to defecation of large and compact (i.e.
-fast-settling) fecal pellets will probably increase the POC flux while fragmentation of large aggregates or fecal pellets to small slower-sinking particles will decrease it.
+The impact on the downward POC flux of zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2022-qj].
+On the one hand, aggregate feeding, filter-feeding, direct ingestion of fecal pellets (i.e.
+coprophagy) and fragmentation of large aggregates to small slower-sinking particles will likely decrease the downward POC flux.
+On the other hand, the ingestion of slow-sinking cells to defecation of large and compact (i.e.
+fast-settling) fecal pellets will increase it.
 The remaining organic matter (\~15% according to @Turner2015-af) will then be degraded by microbial communities.
@@
 
-That being said, several processes can impact the sinking speed of POC such as aggregation (e.g.
+Taking this into consideration, several processes can impact the sinking speed of POC such as aggregation (e.g.
 physical coagulation, incorporation of small particles into fecal pellets (i.e.
@@
 
-The driving factor of the sinking speed is nevertheless subject to caution.
+The driving factors of the sinking speed are nevertheless subject to caution.
 It would make sense that the quadratic dependence of $w$ to the aggregate size would favor $d$ as the main driver.
@@
 
-To the best of my knowledge, the most emblematic case of discussions around that question is the mineral ballast hypothesis [@De_La_Rocha2007-yi].
+To the best of my knowledge, the most emblematic case of discussions around this question is the mineral ballast hypothesis [@De_La_Rocha2007-yi].
 Mineral ballasting consists in the inclusion of biominerals such as opal (i.e.
 silicate) and calcite or lithogenic material such as dust that are thought to increase $\Delta\rho$ on the one hand and protect them from microbial degradation on the other hand [@Armstrong2001-rz; @Francois2002-uc; @Klaas2002-kn].
-Nevertheless, there is not a consensus on that matter because [@Ploug2008-vx; @Ploug2008-qk] did not find significant differences in microbial degradation of non ballasted vs.
-ballasted material while @Passow2004-re stipulates that biominerals could instead be collected (i.e.
-scavenged) and transported downwards by sinking aggregates and that their integration could fragment aggregates into smaller but denser particles [@Passow2006-hw] which could either decrease (smaller size) or increase (denser particles) their sinking speed.
+However, several studies argue otherwise.
+@Passow2004-re stipulates that biominerals could instead be collected (i.e.
+scavenged) and transported downwards by sinking aggregates.
+@Ploug2008-qk did not measure significant differences in microbial degradation of non ballasted vs.
+ballasted material.
+Finally, @Passow2006-hw developed an experiment to investigate the relationship between the incorporation of biominerals into aggregates and their sinking speed.
+They showed that this relationship was not straigthforward due to the fact that the inclusion of biominerals fragmented aggregates into smaller but denser particles which could either decrease (smaller size) or increase (denser particles) their sinking speed.
 
-The sinking speed of organic matter determines the time available for degradation in the water column therefore identifying the main drivers for the settling of aggregates is a key goal of oceanography [@Iversen2022-qj].
 The size of aggregates was usually thought to be a key predictor of settling velocity [e.g.
@@
 However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor.
-To answer that question, there is a need for direct measurements (size, composition and density) to predict settling velocities [@Iversen2020-kb; @Iversen2022-qj].
-This need could be partially filled with the multi-instrumentation approach described in @sec-observing-carbon-cycle.
+With all things considered, the sinking speed of particulate organic matter determines the time available for degradation in the water column.
+Thus, there is a need for direct measurements of the morphological characteristics (size, composition and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Iversen2022-qj].
+<!-- This need could be partially filled with the multi-instrumentation approach described in @sec-observing-carbon-cycle.
+-->
 
@@
 
-Once injected in the ocean interior (@tbl-PumpDepthRange), the POC flux originating from the five PIP pumps is merged back with the BGP.
+Once injected in the ocean interior (@tbl-PumpDepthRange), the POC flux originating from the five particle injection pumps is merged back with the BGP.
 How long is this carbon going to be sequestered from the atmosphere is one of the key questions of carbon cycle studies [@Claustre2021-tx].
@@
 | Gravitational          | export depth - sediment          | None                      | N/A (142)                            |
-| Mixed layer            | 200 - 1000 m                     | 200 - 1000m               | 25 - 100 (54$^{b}$)                  |
+| Mixed layer            | 200 - 1,000 m                    | 200 - 1,000 m             | 25 - 100 (54$^{b}$)                  |
 | Eddy subduction        | 150 - 400 m                      | 150 - 400 m               | up to 150 (N/A)                      |
-| Large-scale subduction | 200 - 1000 m                     | 200 - 1000 m              | 25 - 100 (54$^{b}$)                  |
+| Large-scale subduction | 200 - 1,000 m                    | 200 - 1,000 m             | 25 - 100 (54$^{b}$)                  |
 | Diel migration         | 200 - 600 m                      | None                      | up to 250 (150$^{c}$)                |
-| Seasonal migration     | 600 - 1400 m$^{a}$               | None                      | up to 500 (N/A)                      |
+| Seasonal migration     | 600 - 1,400 m$^{a}$              | None                      | up to 500 (N/A)                      |
 
@@
 
-In this manuscript, the term sequestration is based on the definition provided by [@Ipcc2019-qc], that is "The long-term removal of carbon dioxide or other forms of carbon from the atmosphere, with secure storage on climatically significant timescales (decadal to century)" with a period of storage of 100 years, which is commonly used in climate-relevant studies [@Lampitt2008-kq; @Passow2012-zh].
+```{r, test table}
+# THERE IS NO CURRENT SOLUTION TO THIS ISSUE WITH PDF RENDERING UNFORTUNATELY, SEE https://github.com/rstudio/gt/issues/119
+# library(tibble)
+# library(gt)
+# library(gtsummary)
+# sequestration_table <- tibble('Pump' = c('Gravitational', 'Mixed layer', 'Eddy subduction', 'Large-scale subduction', 'Diel migration', 'Seasonal migration'),
+#                               'POC export/injection depth range' = c('export depth - sediment', '200 - 1000 m', '150 - 400 m', '200 - 1000 m', '200 - 600 m', '600 - 1400 m*'),
+#                               'DOC injection depth range' = c('None', '200 - 1000 m', '150 - 400 m', '200 - 1000 m', 'None', 'None'),
+#                               'Sequestration time potential (years)' = c('N/A (142)', '25 - 100 (54)**', 'up to 150 (N/A)', '25 - 100 (54)**', 'up to 250 (150)***', 'up to 500 (N/A)'))
+# 
+# sequestration_table |> tbl_summary() |> tab_source_note(source_note = md('1,400 m corresponds to the overwintering depth of *Calanus finmarchicus* as observed by (@Jonasdottir2015-ob)')) |> 
+#   #cols_width(starts_with('Pump') ~ px(100), starts_with('seq') ~ px(100))
+#   as_kable_extra()
 
-A great majority of studies related to carbon sequestration by the BCP usually assumed that long-term storage of DIC occurred if the exported carbon was transported below 1,000 m [e.g.
+```
+
+A great majority of studies related to carbon sequestration by the BCP usually assumed that the part of the exported carbon reaching 1,000 m was sequestered [e.g.
 @Lampitt2008-kq; @Passow2012-zh].
-However, @Siegel2021-oz used an inverse ocean circulation model to assess the time over which injected $CO_{2}$ (direct injections at different depths) remains sequestered from the atmosphere.
+However, @Siegel2021-oz revises this common assumption by proposing that carbon sequestration can occur on the entire water column, with varying efficiencies (i.e.
+sequestration times).
+To come to this conclusion, @Siegel2021-oz used an inverse ocean circulation model to assess the time over which injected $CO_{2}$ (direct injections at different depths) remains sequestered from the atmosphere.
 For this, they computed a factor $f_{\tau}$ corresponding to the fraction of $CO_{2}$ that will be stored away from the atmosphere for a given time $\tau$.
@@
 
-Therefore, if one wishes to estimate the amount of DIC sequestered from the atmosphere by the BCP, one should not use a reference sequestration depth [@Siegel2022-wd].
-Instead, the approach of @Siegel2021-oz can be translated to the study of the BCP given the fact that injected $CO_{2}$ can be replaced by remineralized organic matter.
-This approach will be discussed in CHAPTER XX.
+So far, we have seen that significant incertitudes remain in the estimations of the carbon export as well as the sequestration time by the different export pathways composing the BCP.
+To address this issue, biogeochemical models and observation tools are continuously being improved to further our understanding of the BCP (see @sec-modelling-observing-BCP).
 
@@
 
-The quantification of the different carbon export pathways of the BCP and the fate of the exported carbon in the mesopelagic zone are among the hardest problems in modern oceanography [@Siegel2016-kh; @Le_Moigne2019-zl] but also the most pressuring issues in the context of climate change [@Martin2020-wa].
+<!-- The quantification of the different carbon export pathways of the BCP and the fate of the exported carbon in the mesopelagic zone are among the hardest problems in modern oceanography [@Siegel2016-kh; @Le_Moigne2019-zl] but also the most pressuring issues in the context of climate change [@Martin2020-wa].
 If the BCP is not responsible for the absorption of anthropogenic $CO_{2}$, any changes in its strength or efficiency could greatly impact atmospheric $CO_{2}$ concentrations [@Kwon2009-ir].
 Thus, there is a need to better understand the biological and physical drivers of each component of the BCP, estimate the amount of carbon they export to the deep ocean and their sequestration potential.
+-->
 
-To this aim, a new observational framework has been proposed by @Claustre2021-tx to combine measurements from several observations platforms (i.e.
+<!-- To this aim, a new observational framework has been proposed by @Claustre2021-tx to combine measurements from several observations platforms (i.e.
 research vessels, satellites, moorings and automated underwater vehicles (AUVs)).
@@
 This is of particular importance to better constrain the contribution of each PIP to carbon storage in the global ocean [@Boyd2019-ez].
+-->
 
-In addition, the refinement of the contribution of each PIP would both help to close the ocean carbon budget [@Burd2010-hc] and improve estimates of carbon sequestration [@Boyd2019-ez; @Passow2012-zh] that will enhance the evaluation of the efficiency of carbon dioxide removal strategies [@Siegel2022-wd].
+<!-- In addition, the refinement of the contribution of each PIP would both help to close the ocean carbon budget [@Burd2010-hc] and improve estimates of carbon sequestration [@Boyd2019-ez; @Passow2012-zh] that will enhance the evaluation of the efficiency of carbon dioxide removal strategies [@Siegel2022-wd].
+-->
 
-Based on that, the objectives of this work are threefold:
+<!-- Based on that, the objectives of this work are threefold: -->
 
 1.
-Contribute to the development of new sensors for AUVs, namely the development of the zooplankton embedded recognition algorithm for the Underwater Vision Profiler 6 [@Picheral2021-sx], an imaging system that will be used to characterize particles \> 600 $\mu$m.
+<!-- Contribute to the development of new sensors for AUVs, namely the development of the zooplankton embedded recognition algorithm for the Underwater Vision Profiler 6 [@Picheral2021-sx], an imaging system that will be used to characterize particles > 600 $\mu$m.
+-->
+
 2.
-Exploit the synergy of instruments on fully-equipped profiling floats in a specific region (Labrador Sea) to shed a light on mesopelagic processes that affect the size and sinking speed of aggregates as well as characterizing the downward flux of small and big particles and its variability.
+<!-- Exploit the synergy of instruments on fully-equipped profiling floats in a specific region (Labrador Sea) to shed a light on mesopelagic processes that affect the size and sinking speed of aggregates as well as characterizing the downward flux of small and big particles and its variability.
+-->
+
 3.
-Reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
+<!-- Reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
+-->

### 2023-02-15 observingCarbonCycle.qmd
update

@@
-# Observing the carbon cycle {#sec-observing-carbon-cycle}
+# Modelling and observing the BCP {#sec-modelling-observing-BCP}
 
-## How it started {#sec-how-it-started}
+## Models
+
+## Observation tools
+
+### Historical measurements
 
@@
 combined <- rbind(sub_trap, sub_thorium)
-plot <- ggplot(combined, aes(x = year, fill = instrument_type)) + geom_bar() + scale_fill_manual(values = c('#003366', '#DCEEF3')) + # #57cc99 #80ed99
+plot <- ggplot(combined, aes(x = year, fill = instrument_type)) + geom_bar(colour = 'black') + scale_fill_manual(values = c('#003366', '#DCEEF3')) + # #57cc99 #80ed99
     #geom_vline(aes(xintercept = median(year)), color = "#d53e4f", linetype = "dashed") +
@@
 
-```{=html}
-<!-- 
-![Examples of sediment traps.
-From left to right: (a) Neutrally buoyant, (b) Surface tethered and (c) Funnel-shaped drifting trap.
-Figure taken from @Baker-XX.](image/ObservingCarbonCycle/sed_traps_high_res.jpg){#fig-sed-traps}
--->
-```
-```{=html}
-<!-- #https://github.com/quarto-dev/quarto-cli/discussions/1042 -->
-<style type="text/css">
-caption, .table-caption {
-  text-align: left;
-}
-</style>
-```
-```{r, poc database from traps}
-#| label: tbl-poc-from-sediment-traps
-#| tbl-cap: Compilation of POC flux measurements from sediment traps.
-Data taken from public databases listed on PANGAEA or extracted from articles (with the help of Frédéric Le Moigne and Meike Vogt).
-The number of measurements could be different from the original publication after data cleaning and potential removal of duplicates.
-#| tbl-cap-location: bottom
-
-library(dplyr)
-library(knitr)
-library(kableExtra)
-library(purrr)
-
-# read data
-trap <- vroom::vroom('DATA/Trap_Thorium/TRAP_DATABASE.csv')
-#thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
-
-# clean a bit
-trap[trap$doi == 'https://doi.org/10.1007/s00227-018-3376-1',]$doi <- '10.1007/s00227-018-3376-1'
-trap[trap$doi == 'https://doi.org/10.1016/j.dsr.2016.08.015',]$doi <- '10.1016/j.dsr.2016.08.015'
-trap[trap$doi == 'Geochimica et Cosmochimica Acta, Vol. 66, No. 3, pp. 457–465, 2002',]$doi <- '10.1016/S0016-7037(01)00787-6'
-
-# some stats
-trap_resume <- trap |> group_by(doi) |> dplyr::summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
-  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
-
-# BASED ON : https://stackoverflow.com/questions/48506712/kable-kableextra-cells-with-hyperlinks
-# make url links
-make_link <- function(doi){
-  link <- paste0('https://doi.org/',doi)
-  return(link)
-}
-url <- map_chr(trap_resume$doi, make_link)
-
-# # put url in dataframe
-# trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, 'html', link = url)) |> select(doi_url, data_points) |>
-#   kable('html', escape = FALSE, col.names = c('Sources of data', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))
-
-trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
-  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l')
-
-# make final table
-trap_resume
-```
-
-```{r, poc database from thorium}
-#| label: tbl-poc-from-thorium
-#| tbl-cap: Compilation of POC flux measurements derived from Thorium 234.
-Data taken from public databases listed on PANGAEA or extracted from articles (with the help of Frédéric Le Moigne).
-The number of measurements could be different from the original publication after data cleaning and potential removal of duplicates.
-#| tbl-cap-location: bottom
-
-# read data
-thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
-
-# some stats
-thorium_resume <- thorium |> group_by(doi) |> summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
-  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
-thorium_resume$doi <- c('10.1594/PANGAEA.809717', '10.1029/2018GB00615', '10.1525/elementa.2020.030')
-
-# make url links
-url <- c('https://doi.org/10.1594/PANGAEA.809717', 'https://doi.org/10.1029/2018GB006158', 'https://doi.org/10.1525/elementa.2020.030')
-
-# put url in dataframe
-thorium_resume <- thorium_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
-  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))
-
-# make final table
-thorium_resume
-```
-
 To estimate the downward POC flux, one therefore needs to quantify the amount of POC collected through a given area over a known time period at a given depth.
@@
 
-Given the potential inaccuracy of flux measurements by sediment traps, independent approaches using radionuclide disequilibria have been used to assess the collection efficiency of the latter [@Buesseler1991-vo].
+Given some potential bias in the carbon fluxes estimated with sediment traps, independent approaches using radionuclide disequilibria have been used to assess the collection efficiency of the latter [@Buesseler1991-vo].
 This method is based on the fact that $^{238}U$ is a conservative element in seawater (chemically inert with a half-life of 4.5 billion years) while its daughter nuclide $^{234}Th$ (half-life of 24.1 days) can be adsorbed onto particle surfaces and therefore be scavenged by sinking particles [@Savoye2006-gb].
-Consequently, the deficiency of $^{234}Th$ relative to $^{238}U$ is used as a measure of removal of surface $^{234}Th$ through particle scavenging [@Buesseler1991-vo] that is then derived as a flux of exported $^{234}Th$ and subsequently a flux of exported POC by using a known (e.g.
+Consequently, the deficiency of $^{234}Th$ relative to $^{238}U$ is used as a measure of removal of surface $^{234}Th$ through particle scavenging [@Buesseler1991-vo] that is then translated into a flux of exported $^{234}Th$ and subsequently a flux of exported POC by using a known (e.g.
 estimated from the material collected in sediment traps) POC/$^{234}Th$ ratio [@Buesseler2007-ql].
@@
 In addition, they are labor-intensive and require ship support.
-Their relative contribution to the estimation of POC fluxes is thus expected to keep decreasing (@fig-sediment-trap-thorium-yearly-distribution) with the advent of AUVs (see @sec-how-it-has-evolved).
+Their relative contribution to the estimation of POC fluxes is thus expected to keep decreasing (@fig-sediment-trap-thorium-yearly-distribution) with the advent of alternative new methods promoted by an increasing use of automated underwater vehicles (AUVs).
 

### 2023-02-16 article3.qmd
update

@@
 
-parking_data <- parking_data |> mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) |> group_by(park_depth, cycle, wmo)
+parking_data <- parking_data |> mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) #|> group_by(park_depth, cycle, wmo)
+
+# add rainer computation
+compute_flux_rainer <- function(A, b, PSD){
+
+  # Rainer uses an ESD in cm ..
+  mid_ESD <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
+       0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+       1.16170742, 1.46365963, 1.84409558, 2.32341484) / 10
+  exp <- A * mid_ESD ** b # because Rainer uses the units of particles/m3 and not particles/L
+  estimated_flux <- rowSums(t(t(PSD)*exp))
+  return(estimated_flux)
+}
+
+parking_spectra_rainer <- parking_data |> dplyr::select(lpm_class)
+parking_data <- parking_data |> mutate(rainer_flux = compute_flux_rainer(2.8649, 2.24, parking_spectra_rainer)) 
+parking_data <- parking_data |> mutate(iversen_flux = compute_flux(273, 4.27, parking_spectra))
 
@@
 colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('NP_Size', 'Biovolume')
+biovolumes_names <- colnames(spectra_biovolume)
 uvp_drifting <- cbind(uvp_drifting, spectra_biovolume) |> rowwise() |> mutate(total_biovolume = sum(across(all_of(biovolumes_names))))
-biovolumes_names <- colnames(spectra_biovolume)
 
@@
 
-ggplot(test) + geom_line(aes(x = mean_date, y = weighted_group_biovolume)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4)
+ggplot(test) + geom_line(aes(x = mean_date, y = weighted_group_biovolume)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
 ggplot(test) + geom_line(aes(x = mean_date, y = slope)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4)
@@
 
-#ggplot(test) + geom_point(aes(x = log(slope), y = log(weighted_group_biovolume))) + facet_wrap(~park_depth)
+test2 |> group_by(wmo, park_depth) |> summarize(correl = cor(slope, weighted_group_biovolume))
+
+ggplot(test) + geom_point(aes(x = log(slope), y = log(weighted_group_biovolume))) + facet_wrap(~park_depth)
+#ggplot(test) + geom_point(aes(x = slope, y = weighted_group_biovolume)) + facet_wrap(~park_depth)
 
@@
 
-ggplot(test_jump) + geom_point(aes(diff_jump, biovolume))
+#ggplot(test_jump) + geom_point(aes(diff_jump, biovolume))
 ggplot(test_jump) + geom_point(aes(log(diff_jump), log(biovolume)))

### 2023-02-16 intro.qmd
update

@@
 To address this issue, biogeochemical models and observation tools are continuously being improved to further our understanding of the BCP (see @sec-modelling-observing-BCP).
-
-## Research questions and objectives
-
-<!-- The quantification of the different carbon export pathways of the BCP and the fate of the exported carbon in the mesopelagic zone are among the hardest problems in modern oceanography [@Siegel2016-kh; @Le_Moigne2019-zl] but also the most pressuring issues in the context of climate change [@Martin2020-wa].
-If the BCP is not responsible for the absorption of anthropogenic $CO_{2}$, any changes in its strength or efficiency could greatly impact atmospheric $CO_{2}$ concentrations [@Kwon2009-ir].
-Thus, there is a need to better understand the biological and physical drivers of each component of the BCP, estimate the amount of carbon they export to the deep ocean and their sequestration potential.
--->
-
-<!-- To this aim, a new observational framework has been proposed by @Claustre2021-tx to combine measurements from several observations platforms (i.e.
-research vessels, satellites, moorings and automated underwater vehicles (AUVs)).
-New sensors have been developed for AUVs (see section @sec-BGC-sensors) in order to better characterize the type of particles (e.g.
-zooplankton, fecal pellets, aggregates), their size and their mechanisms of transformation (e.g.
-fragmentation, aggregation).
-Combining independent measurements of specific biogeochemical variables will allow to tackle the double accounting problem, that is the issue of distinguishing particles associated with the BGP and a PIP, therefore potentially attributing the transport of a particle to more than one pump [@Boyd2019-ez].
-This is of particular importance to better constrain the contribution of each PIP to carbon storage in the global ocean [@Boyd2019-ez].
--->
-
-<!-- In addition, the refinement of the contribution of each PIP would both help to close the ocean carbon budget [@Burd2010-hc] and improve estimates of carbon sequestration [@Boyd2019-ez; @Passow2012-zh] that will enhance the evaluation of the efficiency of carbon dioxide removal strategies [@Siegel2022-wd].
--->
-
-<!-- Based on that, the objectives of this work are threefold: -->
-
-1.
-<!-- Contribute to the development of new sensors for AUVs, namely the development of the zooplankton embedded recognition algorithm for the Underwater Vision Profiler 6 [@Picheral2021-sx], an imaging system that will be used to characterize particles > 600 $\mu$m.
--->
-
-2.
-<!-- Exploit the synergy of instruments on fully-equipped profiling floats in a specific region (Labrador Sea) to shed a light on mesopelagic processes that affect the size and sinking speed of aggregates as well as characterizing the downward flux of small and big particles and its variability.
--->
-
-3.
-<!-- Reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
--->

### 2023-02-16 observingCarbonCycle.qmd
update

@@
 
-## How it has evolved {#sec-how-it-has-evolved}
+## The potential of autonomous platforms {#sec-potential-autonomous-platforms}
 
-It is well known that ship-based measurements are the most precise.
+It is admitted that ship-based measurements are potentially the most precise.
 However, they lack spatiotemporal coverage and they are usually not cost-effective.
@@
 
-Instrumented moorings have supplemented ship-based measurements by collecting high-resolution data at fixed locations [@Chai2020-uq].
+Remote sensing data acquired with ocean color satellites have been used for a long time [@Gordon1980-dm] to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few.
+While having a wide spatial coverage, these platforms are limited to the ocean's surface and their measures can be affected by several factors such as the cloud cover or low solar elevation (e.g.
+@Marshall1994-vk).
+
+Instrumented moorings have supplemented ship-based and remote sensing measurements by collecting high-resolution data at fixed locations [@Chai2020-uq].
 Moorings can be equipped with many sensors (e.g.
-sediment traps (@sec-how-it-started) or imaging systems to study the type of sinking particles (@sec-BGC-sensors)) to study carbon-related variables (see Table 1 in @Claustre2021-tx).
+sediment traps or imaging systems to study the type of sinking particles (@sec-BGC-sensors)) to study carbon-related variables (see Table 1 in @Claustre2021-tx).
 They are therefore ideal platforms for long-term time series (e.g.
@@
 
-In contrast, satellite remote sensing have a wide spatial coverage but they are limited to the ocean's surface and their measures can be affected by several factors such as the cloud cover or low solar elevation (e.g.
-@Marshall1994-vk).
-They have been used for ocean applications since the late 1970s [@Gordon1980-dm] and have been widely used to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few.
-
 In order to fill the scarcity of data from research vessels and moorings as well as coping with surface-constrained satellite measurements, AUVs have been developed to take *in situ* measurements in all weathers and at all times, only being dependent on their battery life and ocean currents (for drifting AUVs).
 
-Biogeochemical Argo (BGC-Argo) profiling floats (@fig-bgc-float) are probably the most well known AUVs.
+Biogeochemical Argo (BGC-Argo) profiling floats (@fig-bgc-float) are probably the most well known to address the BCP.
 They are cost-effective, able to monitor biogeochemical processes at a global scale throughout the year and under all weather conditions [@Chai2020-uq].
@@
 
-Gliders are also effective AUVs to unveil the mysteries of the BCP [e.g.
+Gliders are also effective AUVs to study the BCP [e.g.
 @Omand2015-mf].
@@
 
-The study of the BCP is a really hot topic in oceanography with new platforms (e.g.
+In allThe study of the BCP is a really hot topic in oceanography with new platforms (e.g.
 unmanned surface vehicles such as saildrones or wave gliders) and new technologies (e.g.
@@
 This framework will hopefully answer questions raised so far, that is better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and better quantify global estimates of carbon storage.
+
+## Research questions and objectives
+
+<!-- The quantification of the different carbon export pathways of the BCP and the fate of the exported carbon in the mesopelagic zone are among the hardest problems in modern oceanography [@Siegel2016-kh; @Le_Moigne2019-zl] but also the most pressuring issues in the context of climate change [@Martin2020-wa].
+If the BCP is not responsible for the absorption of anthropogenic $CO_{2}$, any changes in its strength or efficiency could greatly impact atmospheric $CO_{2}$ concentrations [@Kwon2009-ir].
+Thus, there is a need to better understand the biological and physical drivers of each component of the BCP, estimate the amount of carbon they export to the deep ocean and their sequestration potential.
+-->
+
+<!-- To this aim, a new observational framework has been proposed by @Claustre2021-tx to combine measurements from several observations platforms (i.e.
+research vessels, satellites, moorings and automated underwater vehicles (AUVs)).
+New sensors have been developed for AUVs (see section @sec-BGC-sensors) in order to better characterize the type of particles (e.g.
+zooplankton, fecal pellets, aggregates), their size and their mechanisms of transformation (e.g.
+fragmentation, aggregation).
+Combining independent measurements of specific biogeochemical variables will allow to tackle the double accounting problem, that is the issue of distinguishing particles associated with the BGP and a PIP, therefore potentially attributing the transport of a particle to more than one pump [@Boyd2019-ez].
+This is of particular importance to better constrain the contribution of each PIP to carbon storage in the global ocean [@Boyd2019-ez].
+-->
+
+<!-- In addition, the refinement of the contribution of each PIP would both help to close the ocean carbon budget [@Burd2010-hc] and improve estimates of carbon sequestration [@Boyd2019-ez; @Passow2012-zh] that will enhance the evaluation of the efficiency of carbon dioxide removal strategies [@Siegel2022-wd].
+-->
+
+<!-- Based on that, the objectives of this work are threefold: -->
+
+1.
+<!-- Contribute to the development of new sensors for AUVs, namely the development of the zooplankton embedded recognition algorithm for the Underwater Vision Profiler 6 [@Picheral2021-sx], an imaging system that will be used to characterize particles > 600 $\mu$m.
+-->
+
+2.
+<!-- Exploit the synergy of instruments on fully-equipped profiling floats in a specific region (Labrador Sea) to shed a light on mesopelagic processes that affect the size and sinking speed of aggregates as well as characterizing the downward flux of small and big particles and its variability.
+-->
+
+3.
+<!-- Reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
+-->

### 2023-02-20 intro.qmd
update

@@
 The carbon flux leaving the surface layer for the ocean interior is referred to as the export flux [@Siegel2022-wd].
-Historically, the export depth was usually set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should now use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where particulate organic carbon (POC) is produced (i.e.
+Historically, the export depth was usually set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should now use the base of the euphotic zone (EZ, where there is enough light for net primary production (NPP)) as the export depth to clearly differentiate the layer where particulate organic carbon (POC) is produced (i.e.
 via photosynthesis) and where POC is consumed (i.e.
@@
 
-Estimates of global carbon export range between 5 and 12 Pg C year$^{-1}$ [@Dunne2007-mi; @Henson2011-mt; @Siegel2014-wg; @DeVries2017-wu; @Nowicki2022-no] with only 0.02-0.4 Pg C year$^{-1}$ reaching the sediment [@Dunne2007-mi; @Hayes2021-gg; @Nowicki2022-no] where carbon will be sequestered for millions of years.
+Estimates of global carbon export range between 5 and 12 Pg C year$^{-1}$ [@Laws2000-XX; @Dunne2007-mi; @Henson2011-mt; @Siegel2014-wg; @DeVries2017-wu; @Nowicki2022-no] with only 0.02-0.4 Pg C year$^{-1}$ reaching the sediment [@Dunne2007-mi; @Hayes2021-gg; @Nowicki2022-no] where carbon will be sequestered for millions of years.
 Almost all of the exported organic carbon will therefore be remineralized and sequestered away from the atmosphere on time scales ranging from months to millennia.

### 2023-02-20 observingCarbonCycle.qmd
update

@@
-# Modelling and observing the BCP {#sec-modelling-observing-BCP}
+# Modeling and observing the BCP {#sec-modelling-observing-BCP}
 
-## Models
+## Biogeochemical models
 
-## Observation tools
+The study of the BCP is of paramount importance to answer the question of how much carbon is exported and how much of the exported carbon will be sequestered?
+In order to achieve this goal, ocean modelers have developed different kinds of biogeochemical models to better quantify the different export pathways of the BCP.
+
+![Top of the atmosphere (TOA) radiance is converted into remote sensing reflectance $R_{rs}(\lambda)$, which is the ratio of the solar radiation reflected from the surface of the ocean that penetrates the sea surface at a discrete wavelength $\lambda$.
+Using TOA and $R_{rs}(\lambda)$, primary products (yellow circles) can be derived (e.g.
+POC, inherent optical properties (IOP), chlorophyll (Chl) or sea surface temperature (SST).
+Those primary products can be used in numerical models (e.g.
+empirical relationships) to compute secondary products (orange and red circles) such as the phytoplankton commmunity composition (PCC), particle size spectra (PSS), phytoplankton carbon concentrations (C$_{phyto})$ and the NPP.
+Models (light blue circles) therefore incorporate primary and secondary products as well as ancillary data (red squares) either from satellite (e.g.
+SST), in situ data or model-based data (ocean circulation) to estimate carbon export and sequestration (dark blue circles).
+Figure taken from @Siegel2022-wd.](image/ObservingCarbonCycle/carbon_models.jpeg){#fig-bgc-models width="750" height="580"}
+
+Because measuring directly the POC export flux would be challenging, it is common to estimate it as the fraction of the total NPP that is exported (i.e. export efficiency) [@Henson2011-mt].
+
+::: {.blackbox data-latex=""}
+::: {#box-npp .center data-latex=""}
+A deeper look into net primary production
+:::
+
+Total net primary production includes new production (NP, mainly fueled from the upwelling of nitrogen from deeper water to the EZ) and regenerated production (RP, fueled with nitrogen recycled through the food web within the EZ) \[@Dugdale1967\].
+
+@Eppley1979 introduced the f-ratio as follows
+
+$$
+f-ratio = \frac{NP}{NP+RP}
+$$
+
+In a steady state system, new production equals export production therefore the $f-ratio$ should be equal to the ratio of export production (EP) to total NPP, referred to as the $e-ratio$ by @laws2000-XX.
+
+$$
+e-ratio = \frac{EP}{NPP}
+$$
+
+define steady state?
+:::
+
+@box-npp
+
+Empirical models likely represent the most straightforward option to estimate carbon export.
+
+combining satellite observations and in situ measurements
+
+Data-assimilated models
+
+These models can either target the estimations of carbon export
+
+These models (i.e. food-web, empirical and inverse models) incorporate satellite data (@sec-observation-tools), that can be defined as primary data product
+
+numerical models based on simple empirical relationships to derive primary and secondary data products
+
+As illustrated in @fig-bgc-models,
+
+One of the key variable to estimate carbon export is the NPP. It has been used to compute the f-ratio, e-ratio -\> rentrer un peu là-dedans
+
+The former
+
+Henson2010
+
+-   common approach = estimate fraction of total primary production that is exported (with f-ratio, Th234-ratio
+
+-   encadré sur le f-ratio? en lien avec laws2000 ry Dugdale and Goering 1967
+
+-   satellite data
+
+Dunne2007
+
+-   NPZD models (non exhaustive)
+
+-   empirical models: predictor variables
+
+-   empirical algorithm
+
+Henson2011
+
+-   Martin b : -0.86 (6 stations, in the NOrtheast Pacific), lack of data so b is generally assumed to be spatially constant and this b value has been applied in several biogeochemical models to describe the rate of particle reminrelaiation in the water column
+
+-   data-derived algorithms
+
+Siegel 2014
+
+-   food-web model (e.g. NPZD)
+
+-   estimation og floval exports by empirical extrapolations of field measurements or by using global coupled carboncyckle-pcean circualtion models but estimations of global export remain widely divergent
+
+-   mechanisticlly, it is the flux of sinking carbon leaving the EZ that is relevant
+
+-   e-ratio ? different du f-ratio?
+
+-   satellite-based approach for assessing carbon export (thank you kostadinov 2009 and 2010) -\> slope of the particle size spectrum
+
+-   mechanistic approach for estimating global carbon export using satellite observations and a food-web model
+
+-   a complete assessment of carbon export from the surfce ocean must accout for this and other additional pathways
+
+-   the BCP is more than just the export of carbon from the EZ. Understanding it, also requires the quantification of the vertical attenuation of the export flux below the EZ as well.
+
+-   aller absoilument checker siegel 2022
+
+Devries2017
+
+Note perso
+
+Ajouter Guidi 2015 pour une estimation de b regionale
+
+## Observation tools {#sec-observation-tools}
 
@@
 
-In order to fill the scarcity of data from research vessels and moorings as well as coping with surface-constrained satellite measurements, AUVs have been developed to take *in situ* measurements in all weathers and at all times, only being dependent on their battery life and ocean currents (for drifting AUVs).
+In order to fill the scarcity of data from research vessels and moorings as well as coping with surface-constrained satellite measurements, AUVs have been developed to take in situ measurements in all weathers and at all times, only being dependent on their battery life and ocean currents (for drifting AUVs).
 

### 2023-02-21 intro.qmd
update

@@
 ::: {.blackbox data-latex=""}
-::: {.center data-latex=""}
-BCP and blue carbon
+::: {.flushleft data-latex=""}
+**Information Box 1.** **BCP and blue carbon**
 :::
@@
 
-Estimates of global carbon export range between 5 and 12 Pg C year$^{-1}$ [@Laws2000-XX; @Dunne2007-mi; @Henson2011-mt; @Siegel2014-wg; @DeVries2017-wu; @Nowicki2022-no] with only 0.02-0.4 Pg C year$^{-1}$ reaching the sediment [@Dunne2007-mi; @Hayes2021-gg; @Nowicki2022-no] where carbon will be sequestered for millions of years.
+Estimates of global carbon export range between 5 and 12 Pg C year$^{-1}$ [@Laws2000-jl; @Dunne2007-mi; @Henson2011-mt; @Siegel2014-wg; @DeVries2017-wu; @Nowicki2022-no] with only 0.02-0.4 Pg C year$^{-1}$ reaching the sediment [@Dunne2007-mi; @Hayes2021-gg; @Nowicki2022-no] where carbon will be sequestered for millions of years.
 Almost all of the exported organic carbon will therefore be remineralized and sequestered away from the atmosphere on time scales ranging from months to millennia.
@@
 @Buesseler2006-og].
-More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size distributions (PSD) that could then be converted to carbon fluxes [e.g.
+More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size distributions (PSD, also referred to in the literature as particle size spectra (PSS)) that could then be converted to carbon fluxes [e.g.
 @Guidi2008-vp].

### 2023-02-21 observingCarbonCycle.qmd
update

--- 2023-02-16 observingCarbonCycle.qmd
@@
-# Modelling and observing the BCP {#sec-modelling-observing-BCP}
+# Modeling and observing the BCP {#sec-modelling-observing-BCP}
 
-## Models
+## Biogeochemical models {#sec-biogeochem-models}
 
-## Observation tools
+The study of the BCP is of paramount importance to answer the question of how much carbon is exported and how much of the exported carbon will be sequestered.
+In order to achieve this goal, ocean modelers have developed different kinds of biogeochemical models to better quantify the different export pathways of the BCP.
 
-### Historical measurements
+Because measuring directly the POC export flux would be challenging, it is common to estimate the fraction of the total NPP that is exported (i.e.
+export efficiency) [@Henson2011-mt].
+This fraction is usually referred to as the $e-ratio$ (see Information Box 2 and @fig-export-ratios).
+
+![$e-ratio$ estimations obtained with different types of models: empirical models (a, b), food-web models (c, d) and data-assimilated models (e, f).
+Note that the export value computed by @Laws2011-nw (9 Pg C y$^{-1}$) was recomputed by @Siegel2022-wd (6.9 Pg C y$^{-1}$), therefore the $e-ratio$ map presented here might not be representative of the one we would have got with with the initial export value.
+Figure taken from @Siegel2022-wd.](image/ObservingCarbonCycle/export_ratio.jpeg){#fig-export-ratios}
+
+::: {.blackbox data-latex=""}
+::: {.flushleft data-latex=""}
+**Information Box 2.** **Net primary production and export ratio**
+:::
+
+Total net primary production includes new production (NP, mainly fueled from the upwelling of nitrogen from deeper water to the EZ) and regenerated production (RP, fueled with nitrogen recycled through the food web within the EZ) [@Dugdale1967-ut].
+The ratio of new production over total production is referred to as the $f-ratio$ [@Eppley1979-pj].
+
+$$
+f-ratio = \frac{NP}{NP+RP}
+$$
+
+In a steady state system, new production equals export production therefore the $f-ratio$ should be equal to the ratio of export production (EP) to total NPP, referred to as the $e-ratio$ by @Laws2000-jl.
+
+$$
+e-ratio = \frac{EP}{NPP}
+$$
+:::
+
+I will hereafter briefly describe a non exhaustive list of biogeochemical models to illustrate how the export production can be estimated using the $e-ratio$.
+
+@Henson2011-mt and @Laws2011-nw (panels a and b in @fig-export-ratios) are both empirical models.
+@Henson2011-mt parameterized the $e-ratio$ as a decreasing function of SST (derived from satellite observations, see @fig-bgc-models) tuned on a compilation of $^{234}Th$-derived export measurement (see @sec-historical-measurements).
+They estimated a carbon export of 5 Pg C y$^{-1}$.
+@Laws2011-nw modelled the $e-ratio$ as a decreasing linear function of SST and a power-law function of NPP whose parameters were determined using particle export estimates based on estimates of new production (nitrogen-based) and $^{234}Th$ and sediment trap export flux data compiled by @Dunne2005-an.
+They estimated a carbon export of 9 Pg C y$^{-1}$.
+
+@Siegel2014-wg and @Archibald2019-dj (panels c and d in @fig-export-ratios) are both food-web models (see Information Box 3).
+@Siegel2014-wg used a mechanistic approach using satellite observations (NPP, PSS and phytoplankton carbon concentrations, see also @fig-bgc-models) to estimate the export of the BGP.
+They estimated a carbon export of 6 Pg C y$^{-1}$.
+@Archibald2019-dj added a new module to the model of @Siegel2014-wg in order to incorporate the diel vertical migration of zooplankton for a total export of 6.5 Pg C y$^{-1}$.
+
+::: {.blackbox data-latex=""}
+::: {.flushleft data-latex=""}
+**Information Box 3.** **Food-web model**
+:::
+
+A simple food-web model is the NPZD model, which consists of four state variables: Nutrient (DIN), Phytoplankton (PHYTO), Zooplankton (ZOO) and DETRITUS.
+The arrows represent the flows (or interactions) between the state variables (e.g.
+f1 represents the DIN uptake by PHYTO, f2 represents the grazing of PHYTO by ZOO, etc.), describing the sources and sinks of a given state variable \[@Soetaert2008-er\].
+Forcing functions (or external variables) are external factors (i.e.
+not computed by the model) imposed (i.e.
+forced) to the model because they drive or regulate the system.
+Examples of forcing functions (among many others) in ecological models are light intensity (represented here by the solar radiation), temperature or wind.
+Finally, a model usually contains output variables (represented here by the Chlorophyll) that allow the comparison of the model to reality @Soetaert2008-er.
+
+::: {.center data-latex=""}
+![](image/ObservingCarbonCycle/npzd_no_background.png)
+:::
+
+Figure taken from [@Soetaert2008-er].
+:::
+
+@DeVries2017-wu and @Nowicki2022-no (panels e and f in @fig-export-ratios) are both data-assimilated models.
+These models operate by nudging model solutions to the observed fields (e.g.
+satellite and/or in situ data) or by optimizing model parameters to the observed fields \[@Siegel2022-wd\].
+@DeVries2017-wu combined both satellite-derived data (NPP and PSS) and in situ observations (POC fluxes, DOC concentrations and dissolved $O_{2}$ distributions) to assess the carbon export and its transfer in the mesopelagic zone using a steady state ocean circulation inverse model (OCIM, the circulation component of their model).
+They estimated a carbon export between 7.3 and 8.6 Pg C y$^{-1}$.
+@Nowicki2022-no added a two-phytoplankton/two-zooplankton food-web model with a migration zooplankton component to the model of @DeVries2017-wu in order to estimate the magnitude of each export pathway of the BCP.
+In all, @Nowicki2022-no estimated a carbon export of 10 Pg C y$^{-1}$.
+
+### Model uncertainties {#sec-model-uncertainties}
+
+When we compare the $e-ratio$ maps (@fig-export-ratios) or the carbon export estimations of just six biogeochemical models (see @sec-biogeochem-models), we already observe strong differences.
+In my opinion, it means at least two things.
+On the one hand, we are seeing an increase in model complexity from empirical models to data-assimilated inverse models.
+At the same time, we also gathered more observational data as well as a deeper understanding of the BCP (@sec-a-new-paradigm-for-the-bcp) that allow us to better represent its underlying processes.
+On the other hand, each degree of complexity propagates more and more uncertainty (@fig-bgc-models) that we need to account for in the results.
+
+![Top of the atmosphere (TOA) radiance is converted into remote sensing reflectance $R_{rs}(\lambda)$, which is the ratio of the solar radiation reflected from the surface of the ocean that penetrates the sea surface at a discrete wavelength $\lambda$.
+Using TOA and $R_{rs}(\lambda)$, primary products (yellow circles) can be derived (e.g.
+POC, inherent optical properties (IOP), chlorophyll (Chl) or sea surface temperature (SST).
+Those primary products can be used in numerical models (e.g.
+empirical relationships) to compute secondary products (orange and red circles) such as the phytoplankton commmunity composition (PCC), particle size spectra (PSS), phytoplankton carbon concentrations (C$_{phyto})$ and the NPP.
+Models (light blue circles) therefore incorporate primary and secondary products as well as ancillary data (red squares) either from satellite (e.g.
+SST), in situ data or model-based data (ocean circulation) to estimate carbon export and sequestration (dark blue circles).
+Figure taken from @Siegel2022-wd.](image/ObservingCarbonCycle/carbon_models.jpeg){#fig-bgc-models width="750" height="580"}
+
+## Observation tools {#sec-observation-tools}
+
+### Historical measurements {#sec-historical-measurements}
 
@@
 
-Remote sensing data acquired with ocean color satellites have been used for a long time [@Gordon1980-dm] to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few.
+Remote sensing data acquired with ocean color satellites have been used for a long time [@Gordon1980-dm] to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few (see also @sec-biogeochem-models.
 While having a wide spatial coverage, these platforms are limited to the ocean's surface and their measures can be affected by several factors such as the cloud cover or low solar elevation (e.g.
@@
 
-In allThe study of the BCP is a really hot topic in oceanography with new platforms (e.g.
-unmanned surface vehicles such as saildrones or wave gliders) and new technologies (e.g.
-lidar) being continuously developed in the hope of understanding all its components [@Chai2020-uq; @Claustre2021-tx].
+In the future, new types of platforms (e.g.
+unmanned surface vehicles such as saildrones or wave gliders), new sensors (see @sec-BGC-sensors) and new technologies (e.g.
+lidar) will be continuously deployed, improved and developed in the hope of better understanding the BCP and its components [@Chai2020-uq; @Claustre2021-tx].
 
-## Specific sensors for the study of the BCP {#sec-BGC-sensors}
+## Specific sensors on autonomous platforms for the study of the BCP {#sec-BGC-sensors}
+
+The BGC-Argo array (@sec-potential-autonomous-platforms) has already demonstrated its ability to understand some aspects of the BCP [e.g.
+@Lacour2017-ro; @Lacour2019-id; @Briggs2020-gz; @Yang2021-vg; @Llort2018-pa; @Wang2022-hn].
+However, the current array may lack some key data to further our understanding of the BCP.
+Therefore, other types of sensors could be added to BGC-Argo floats to help fill this gap.
+
+décrire les 2 capteurs (pour lesquels je vais m'intéresser) et finir par le projet refine, dire que ça va étudier tel et telle zone, comme cela, ça me fait une rampe de lancement pour le chapitre sur le labrador.
 

### 2023-02-22 article3.qmd
update

@@
-# Case study in the Labrador Sea {#case-study-labrador-sea}
+# Case study in the Labrador Sea {#sec-case-study-labrador-sea}
 
@@
 Floats equipped with both a transmissometer and a UVP6 represent an opportunity to look at the influence of parts of the size spectrum on $F_{small}$ and $F_{big}$.
-For this, we decided to build a regression model using XGBoost, a gradient boosting method [REF], using biovolumes measured by the UVP to predict the slope of $F_{small}$ and $F_{big}$ measured by the OST.
+For this, we decided to build a regression model using XGBoost, a gradient boosting method \[REF\], using biovolumes measured by the UVP to predict the slope of $F_{small}$ and $F_{big}$ measured by the OST.
 
@@
 ```
+
 ```{r, sinking speed values}
@@
 
-J'ai enlevé le truc du spectre inversé à 500 et 1000 m car difficile de conclure -> mais comme on est dans la meme masse d'eau, on pourrait le garder, why not
+J'ai enlevé le truc du spectre inversé à 500 et 1000 m car difficile de conclure -\> mais comme on est dans la meme masse d'eau, on pourrait le garder, why not
 

### 2023-02-22 intro.qmd
update

@@
 
-So far, we have seen that significant incertitudes remain in the estimations of the carbon export as well as the sequestration time by the different export pathways composing the BCP.
+So far, we have seen that significant incertitudes remain in the estimations of carbon export as well as the sequestration time by the different export pathways composing the BCP.
 To address this issue, biogeochemical models and observation tools are continuously being improved to further our understanding of the BCP (see @sec-modelling-observing-BCP).

### 2023-02-22 observingCarbonCycle.qmd
update

@@
 @Henson2011-mt and @Laws2011-nw (panels a and b in @fig-export-ratios) are both empirical models.
-@Henson2011-mt parameterized the $e-ratio$ as a decreasing function of SST (derived from satellite observations, see @fig-bgc-models) tuned on a compilation of $^{234}Th$-derived export measurement (see @sec-historical-measurements).
-They estimated a carbon export of 5 Pg C y$^{-1}$.
+@Henson2011-mt parameterized the $e-ratio$ as a decreasing function of SST (derived from satellite observations, see @fig-bgc-models) tuned on a compilation of $^{234}Th$-derived export measurement (see @sec-historical-measurements) for a total export estimated at 5 Pg C y$^{-1}$.
 @Laws2011-nw modelled the $e-ratio$ as a decreasing linear function of SST and a power-law function of NPP whose parameters were determined using particle export estimates based on estimates of new production (nitrogen-based) and $^{234}Th$ and sediment trap export flux data compiled by @Dunne2005-an.
-They estimated a carbon export of 9 Pg C y$^{-1}$.
+They estimated a total export of 9 Pg C y$^{-1}$.
 
 @Siegel2014-wg and @Archibald2019-dj (panels c and d in @fig-export-ratios) are both food-web models (see Information Box 3).
-@Siegel2014-wg used a mechanistic approach using satellite observations (NPP, PSS and phytoplankton carbon concentrations, see also @fig-bgc-models) to estimate the export of the BGP.
-They estimated a carbon export of 6 Pg C y$^{-1}$.
+@Siegel2014-wg used a mechanistic approach using satellite observations (NPP, PSS and phytoplankton carbon concentrations, see also @fig-bgc-models) to estimate the export of the BGP (6 Pg C y$^{-1}$).
 @Archibald2019-dj added a new module to the model of @Siegel2014-wg in order to incorporate the diel vertical migration of zooplankton for a total export of 6.5 Pg C y$^{-1}$.
@@
 The arrows represent the flows (or interactions) between the state variables (e.g.
-f1 represents the DIN uptake by PHYTO, f2 represents the grazing of PHYTO by ZOO, etc.), describing the sources and sinks of a given state variable \[@Soetaert2008-er\].
+f1 represents the DIN uptake by PHYTO, f2 represents the grazing of PHYTO by ZOO, etc.), describing the sources and sinks of a given state variable [@Soetaert2008-er].
 Forcing functions (or external variables) are external factors (i.e.
@@
 These models operate by nudging model solutions to the observed fields (e.g.
-satellite and/or in situ data) or by optimizing model parameters to the observed fields \[@Siegel2022-wd\].
+satellite and/or in situ data) or by optimizing model parameters to the observed fields [@Siegel2022-wd].
 @DeVries2017-wu combined both satellite-derived data (NPP and PSS) and in situ observations (POC fluxes, DOC concentrations and dissolved $O_{2}$ distributions) to assess the carbon export and its transfer in the mesopelagic zone using a steady state ocean circulation inverse model (OCIM, the circulation component of their model).
-They estimated a carbon export between 7.3 and 8.6 Pg C y$^{-1}$.
+Their carbon export estimate is between 7.3 and 8.6 Pg C y$^{-1}$.
 @Nowicki2022-no added a two-phytoplankton/two-zooplankton food-web model with a migration zooplankton component to the model of @DeVries2017-wu in order to estimate the magnitude of each export pathway of the BCP.
@@
 
-Remote sensing data acquired with ocean color satellites have been used for a long time [@Gordon1980-dm] to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few (see also @sec-biogeochem-models.
+Remote sensing data acquired with ocean color satellites have been used for a long time [@Gordon1980-dm] to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few (see also @sec-biogeochem-models).
 While having a wide spatial coverage, these platforms are limited to the ocean's surface and their measures can be affected by several factors such as the cloud cover or low solar elevation (e.g.
@@
 However, the current array may lack some key data to further our understanding of the BCP.
-Therefore, other types of sensors could be added to BGC-Argo floats to help fill this gap.
+Therefore, other types of sensors could be added to help fill this gap.
 
-décrire les 2 capteurs (pour lesquels je vais m'intéresser) et finir par le projet refine, dire que ça va étudier tel et telle zone, comme cela, ça me fait une rampe de lancement pour le chapitre sur le labrador.
+In this work, I will focus on two sensors: a transmissometer used as an optical sediment trap (OST) and the Underwater Vision Profiler 6 (UVP6) [@Picheral2021-sx], a miniaturized version of the UVP5 (@fig-bgc-float).
 
-In the frame of the REFINE (Robots Explore plankton-driven Fluxes in the marine twIlight zoNE) project, enhanced BGC-Argo floats (@fig-bgc-float) will have two new sensors specifically designed for the study of the BCP: a transmissometer and the Underwater Vision Profiler 6 (UVP6) [@Picheral2021-sx].
-
-![Illustration of a BGC-Argo float used in the frame of REFINE.
+![Illustration of an augmented BGC-Argo float.
 From top to bottom: GPS & Iridium (two-way communication) antenna, Oxygen optode, CTD (salinity, temperature and depth), pH sensor, radiometer (downwelling photosynthetic available radiation and downwelling irradiance at 380, 412 and 490 nm), fluorometer (Chla and colored dissolved organic matter) and scattering meter ($b_{bp}$), UVP6 (particle size spectra and images), transmissometer (particle beam attenuation) and nitrate sensor.
-\@Thomas Boniface.](image/ObservingCarbonCycle/float_picture1.png){#fig-bgc-float width="500"}
+\@Thomas Boniface.](image/ObservingCarbonCycle/float_picture1.png){#fig-bgc-float width="400"}
 
@@
 The value of $c$ is the result of absorption and scattering by particles, water and dissolved substances [@Zaneveld1994-ez].
-The transmissometer carried by REFINE BGC-Argo floats uses a light beam at 660 nm whose attenuation coefficient can be splitted into three components: the absorption and scattering of light by particles ($c_{p}$), the absorption by water ($c_{w}$) and the absorption by dissolved organic matter ($c_{DOM}$) [@Jerlov1976-hr].
-At 660 nm, $c_{w}$ is assumed constant at 0.364 m$^{-1}$ and $c_{DOM}$ is considered negligible [@Jerlov1976-hr] hence variations of $c$ will be assumed to be variations of $c_{p}$ in what follows.
+The transmissometer uses a light beam at 660 nm whose attenuation coefficient can be splitted into three components: the absorption and scattering of light by particles ($c_{p}$), the absorption by water ($c_{w}$) and the absorption by dissolved organic matter ($c_{DOM}$) [@Jerlov1976-hr].
+At 660 nm, $c_{w}$ is assumed constant at 0.364 m$^{-1}$ and $c_{DOM}$ is considered negligible [@Jerlov1976-hr] hence variations of $c$ essentially represent variations of $c_{p}$.
 
 The use of transmissometers to characterize POC concentrations was shown by @Bishop1999-es who found that POC concentrations were well correlated with $c_{p}$, independently of ocean environment, season or depth.
-Following that, the idea of using a transmissometer as an Optical Sediment Trap (OST) was firstly applied by @Bishop2004-at where the magnitude of carbon export was quantified based on the particles accumulation on the upward-facing window of vertically oriented transmissometers on autonomous platforms while they were drifting (i.e.
+Subsequently, the idea of using a transmissometer as an OST was firstly applied by @Bishop2004-at where the magnitude of carbon export was quantified based on the particles accumulation on the upward-facing window of vertically oriented transmissometers on autonomous floats while they were drifting (i.e.
 parked at a specific depth).
-This method was later improved by [@Estapa2013-gw] who divided the attenuation signal into a continuous component, that is the continuous accumulation of small particles on the transmissometer window and a discontinuous component resulting in positive jumps likely attributed to the deposition of (rarer) large particles (@fig-estapa-method).
+This method was later improved by [@Estapa2013-gw] who divided the attenuation signal into a continuous component, that is the regular accumulation of small particles on the transmissometer window and a discontinuous component resulting in positive jumps likely attributed to the deposition of (rarer) large particles (see @fig-estapa-method and @sec-case-study-labrador-sea).
 @Estapa2017-jr further calibrated the OST carbon flux proxy using codeployed sediment traps and drifting transmissometers thus allowing the conversion of $c_{p}$ attenuation fluxes (units: m$^{-1}$d$^{-1}$) into carbon fluxes (units: mg C m$^{-2}$d$^{-1}$) using empirical POC-to-attenuance (ATN) ratio were ATN (unitless) is the product of $c_{p}$ and the transmissometer pathlength.
@@
 
-#jump_plot <- ggplot(clean_data, aes(x = dates, y = cp)) + geom_point() + geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour))
+# change names
+tmp2 <- tmp2 |> mutate(colour = if_else(colour == 'cleaned signal', 'Small particle deposition', 'Big particle deposition (jump)'))
 
-# ggplot(despiked_data$data, aes(x = dates, y = cp)) + geom_step(size= 3, fill ='black') + 
-#   geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour), size = 3) + 
-#   scale_colour_manual(values = c('black','yellow', 'blue')) +
-#   theme_bw() +
-#   ylab(TeX('$c_{p}~(m^{-1})$')) +
-#   xlab('Datetime') +
-#   theme(legend.position = c(0.095, 0.875),
-#         legend.background = element_rect(fill = "white", color = "black"))
-
-jump <- filter(tmp2, colour == 'jump')
+jump <- filter(tmp2, colour == 'Big particle deposition (jump)')
 
@@
   geom_point(data = jump, aes(x = dates, y = cp), size = 3, shape = 21) +
-  scale_fill_manual(values = c('#003366','#DCEEF3')) +
+  #scale_fill_manual(values = c('#003366','#DCEEF3')) +
+  scale_fill_manual(values = c('#DCEEF3', '#003366')) +
   theme_bw() +
@@
   theme(legend.position = 'top', legend.title = element_blank())
-
-# ggplot(despiked_data$data, aes(x = dates, y = cp)) + geom_point(size= 3, fill ='black') +
-#   geom_point(data = jump, aes(x = dates, y = cp), size = 3, shape = 21, fill = '#DCEEF3') +
-#   scale_colour_manual(values = c('black','yellow', 'blue')) +
-#   theme_bw() +
-#   ylab(TeX('$c_{p}~(m^{-1})$')) +
-#   xlab('Datetime') +
-#   theme(legend.position = 'top')
-# 
-# jump_plot <- ggplot(base_signal, aes(x = dates, y = cp)) + geom_point(size = 3, shape = 21, fill = '#003366') +
-#   geom_point(data = jump, aes(x = dates, y = cp), size = 3, shape = 21, fill = '#DCEEF3') +
-#   theme_bw() +
-#   ylab(TeX('$c_{p}~(m^{-1})$')) +
-#   xlab('Datetime')
-
-#geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = 'red', size = 2) + theme_bw() + 
-#labs(x = 'Time', y = 'Cp (1/m)', title = paste0('Mean slope (red): ', round(mean_slope,3), ' m-1 day-1')) # \n', 'POC flux (small particles): ', poc_flux, 'mg'))
 
@@
 size, gray level, taxonomic group).
-Two versions of the UVP6 have been commercialized: the UVP6-HF (high frequency) and the UVP6-LP (low power) which has been specifically designed for AUVs that have strict energy requirements (e.g.
+Two versions of the UVP6 are available: the UVP6-HF (high frequency) and the UVP6-LP (low power) which has been specifically designed for AUVs that have strict energy requirements (e.g.
 autonomous floats).
@@
 The UVP6-LP (hereafter referred to as UVP6) is thus both a particle counter (0.1 mm - 2 mm) and a particle classifier (\> 0.6 mm).
-The former counts particles in different class sizes in order to compute the particle size distribution (PSD) - also referred to as PSS (@sec-BGP) - while the latter extracts pixel-related and geometric features from each image to classify them (i.e.
-assign them a taxonomic group) directly on board using artificial intelligence (CHAPTER-XX) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].
+The former counts particles in different class sizes in order to compute the PSD while the latter extracts pixel-related and geometric features from each image to classify them (i.e.
+assign them a taxonomic group) directly on board using artificial intelligence (@sec-uvp-ia) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].
 
@@
 
-While the gathering of data from automatic on board particles classification will have to increase to study related applications, the use of PSD in carbon flux studies has already been thoroughly explored [@Guidi2008-vp; @Iversen2010-an; @Jouandet2011-bt; @McDonnell2012-ic; @Roullier2014-wf; @Guidi2015-sj; @Guidi2016-nj; @Ramondenc2016-nu; @Kiko2017-qy; @Fender2019-ou; @Giering2020-mc; @Cael2021-qn; @Bisson2022-bt].
+The deployment of autonomous platforms equipped with a UVP6 and the zooplankton recognition algorithm for the study of the BCP is just beginning (first deployment in May 2022 in the Labrador Sea, see @sec-case-study-labrador-sea).
+On the contrary, the use of PSD in carbon flux studies has already been thoroughly explored [@Guidi2008-vp; @Iversen2010-an; @Jouandet2011-bt; @McDonnell2012-ic; @Roullier2014-wf; @Guidi2015-sj; @Guidi2016-nj; @Ramondenc2016-nu; @Kiko2017-qy; @Fender2019-ou; @Giering2020-mc; @Cael2021-qn; @Bisson2022-bt].
 
 Using the PSD, one can compute the number spectrum $n_{i}$ [@Stemmann2004-jv] which can be defined as the normalized concentration of particles in a given size bin $i$.
-Knowing the mass ($m$) and sinking speed ($w$, see @eq-sinking-speed) of aggregates, one can derive the total carbon flux ($F$)
+Knowing the mass ($m$) and sinking speed ($w$, see @eq-sinking-speed) of aggregates, one can derive the total mass flux ($F$)
 
@@
 
-and therefore from @eq-mass-aggregate-fractal and @eq-sinking-speed-fractal that $A= \alpha\beta$ and $D = \frac{B+1}{2}$.
+and therefore from @eq-mass-aggregate-fractal and @eq-sinking-speed-fractal that $A= \alpha\beta$ and $B = 2D-1$.
 
@@
 
-UVP6 PSD measurements will therefore allow the computation of carbon fluxes assuming that $A$ and $B$ are known. This was done empirically by @Guidi2008-vp who minimized the following cost function
+@Guidi2008-vp used POC fluxes derived from sediment traps and UVP PSD data to estimate the $A$ and $b$ values of @eq-carbon-flux-finite by minimizing the following cost function
 
@@
 
-In all, it seems that the choice for the values of $A$ and $B$ is critical if one does not want to under- or overestimate carbon fluxes derived from PSD [@Iversen2010-an; @Fender2019-ou].
+In all, it seems that the choice for the values of $A$ and $B$ is critical for deriving carbon fluxes derived from PSD [@Iversen2010-an; @Fender2019-ou].
 Three main limitations highlighted by @Bisson2022-bt for UVP optically derived estimates of flux would be: (1) assuming that $A$ and $B$ are globally valid at all depths, for all UVP models and across all size classes, (2) the size to sinking rate uncertainty and (3) the size to carbon mass uncertainty..
@@
 
-## The future of carbon observations {#sec-what-is-next}
-
-The BCP have been studied for many decades by a variety of instruments with their respective pros and cons.
-Their measurements were unfortunately often taken independently, which hindered their intercomparability due to the low number of concurrent measurements between instruments.
-
-With the advent of fully-equipped AUVs, it is now possible to take high-resolution and independent measurements of key biogeochemical variables.
-Combining the latter with measurements from conventional platforms (i.e.
-ships, satellites, moorings) will fulfill the observational framework described by @Claustre2021-tx to investigate the BCP.
-
-This framework will hopefully answer questions raised so far, that is better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and better quantify global estimates of carbon storage.
-
 ## Research questions and objectives
 
-<!-- The quantification of the different carbon export pathways of the BCP and the fate of the exported carbon in the mesopelagic zone are among the hardest problems in modern oceanography [@Siegel2016-kh; @Le_Moigne2019-zl] but also the most pressuring issues in the context of climate change [@Martin2020-wa].
-If the BCP is not responsible for the absorption of anthropogenic $CO_{2}$, any changes in its strength or efficiency could greatly impact atmospheric $CO_{2}$ concentrations [@Kwon2009-ir].
-Thus, there is a need to better understand the biological and physical drivers of each component of the BCP, estimate the amount of carbon they export to the deep ocean and their sequestration potential.
--->
+In the context of the pressing issue of climate change, better understanding the BCP is of paramount importance [@Martin2020-wa].
+Although the BCP is not responsible for the absorption of anthropogenic $CO_{2}$, any change in its strength or efficiency could have a significant impact on atmospheric $CO_{2}$ concentrations [@Kwon2009-ir].
+Therefore, there is a need to better understand the biological and physical drivers of each component of the BCP to reduce uncertainties in the amount of carbon exported and its sequestration potential.
 
-<!-- To this aim, a new observational framework has been proposed by @Claustre2021-tx to combine measurements from several observations platforms (i.e.
-research vessels, satellites, moorings and automated underwater vehicles (AUVs)).
-New sensors have been developed for AUVs (see section @sec-BGC-sensors) in order to better characterize the type of particles (e.g.
-zooplankton, fecal pellets, aggregates), their size and their mechanisms of transformation (e.g.
+To this end, a new observational framework has been proposed by @Claustre2021-tx to combine measurements from several platforms (i.e.
+research vessels, satellites, moorings and AUVs) to investigate the BCP.
+Within this framework, AUVs are being deployed with newly developed sensors to characterize the type of sinking particles (e.g.
+zooplankton, fecal pellets, aggregates), their size and the transformation mechanisms affecting them (e.g.
 fragmentation, aggregation).
-Combining independent measurements of specific biogeochemical variables will allow to tackle the double accounting problem, that is the issue of distinguishing particles associated with the BGP and a PIP, therefore potentially attributing the transport of a particle to more than one pump [@Boyd2019-ez].
-This is of particular importance to better constrain the contribution of each PIP to carbon storage in the global ocean [@Boyd2019-ez].
--->
 
-<!-- In addition, the refinement of the contribution of each PIP would both help to close the ocean carbon budget [@Burd2010-hc] and improve estimates of carbon sequestration [@Boyd2019-ez; @Passow2012-zh] that will enhance the evaluation of the efficiency of carbon dioxide removal strategies [@Siegel2022-wd].
--->
+This framework will hopefully address the questions raised so far, i.e., better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and estimates of carbon sequestration \[@Boyd2019-ez; @Passow2012-zh\] as well as close the ocean carbon budget \[@Burd2010-hc\].
 
-<!-- Based on that, the objectives of this work are threefold: -->
+This work will therefore address three mains topics related to the study of the BCP:
 
 1.
-<!-- Contribute to the development of new sensors for AUVs, namely the development of the zooplankton embedded recognition algorithm for the Underwater Vision Profiler 6 [@Picheral2021-sx], an imaging system that will be used to characterize particles > 600 $\mu$m.
--->
-
+BGC-Argo floats equipped with a UVP6 and the zooplankton recognition algorithm are able to classify particles larger than 0.6 mm.
+However, these AUVs are limited in terms of their battery life.
+They are typically deployed for several years, which requires special attention to the energy consumption of each sensor.
+In @sec-uvp-ia I will therefore describe the methodology used to develop the zooplankton classification model for the UVP6 accounting for energy limitations.
 2.
-<!-- Exploit the synergy of instruments on fully-equipped profiling floats in a specific region (Labrador Sea) to shed a light on mesopelagic processes that affect the size and sinking speed of aggregates as well as characterizing the downward flux of small and big particles and its variability.
--->
-
+4 BGC-Argo floats equipped with both a transmissometer and a UVP6 were deployed for the first time in the Labrador Sea in late May 2022 to investigate the post-spring bloom and its associated POC fluxes.
+In @sec-case-study-labrador-sea, I will show how transmissometers used as OSTs compare with PSDs measured by UVP6s as well as investigate the link between small and big particle fluxes derived by both instruments.
 3.
-<!-- Reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
--->
+Many studies have assumed that carbon sequestration only occurs in the deep ocean (typically 1000 m).
+In @sec-to-do, I reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.

### 2023-02-27 intro.qmd
update

@@
 
-Calcifying organisms such as coccolithophorids precipitate biogenic $CaCO_{3}$ (i.e.
-particulate inorganic carbon) either in the form of calcite or aragonite [@Sarmiento2006-ec] as shown in the equation below
+Calcifying organisms (mostly coccolithophores, foraminifera and pteropods) produce particulate inorganic carbon (PIC) by precipitating biogenic $CaCO_{3}$ [@Sarmiento2006-ec]
 
@@
 $$ {#eq-calcium_carbonate}
+
+to form calcite structures such as calcite platelets (coccoliths), calcite shells or spines [@Neukermans2023-vl].
 
@@
 
-This calcification occurs in surface waters, releasing $CO_{2}$ in the surrounding waters and subsequently to the atmosphere.
-That is why this pump is sometimes also referred to as the carbonate counter pump [@Zeebe2012-zk].
-Once calcifying organisms die, their shells sink to depth and finally dissolve.
-Therefore, @eq-calcification shows that the carbonate pump also contributes to the observed DIC gradient, with an estimated contribution of 20% [@Sarmiento2006-ec].
+This calcification occurs in surface waters, releasing $CO_{2}$ in the surrounding waters and subsequently to the atmosphere (0.26 - 1.19 Pg C year$^{-1}$ following @Neukermans2023-vl).
+That is why this pump is sometimes also referred to as the carbonate counter pump \[@Zeebe2012-zk\].
+However, the dissolution of $CaCO_{3}$ mainly occurs at depth and thus increases the DIC concentration (@eq-calcification).
+In all, @Sarmiento2006-ec estimated that the carbonate pump contributes to 20% of the observed DIC gradient.
 
@@
 temperature) and chemical (e.g.
-nutrients availability, acidification) drivers of the BCP [@Riebesell2009-qj; @Passow2012-zh; @Bopp2013-jx] whose long-term carbon sequestration capacity remains uncertain [@Passow2012-zh; @Turner2015-af; @DeVries2022-of; @Wilson2022-bf].
+nutrients availability, acidification) drivers of the BCP [@Riebesell2009-qj; @Passow2012-zh; @Bopp2013-jx] whose long-term carbon sequestration capacity remains uncertain [@Passow2012-zh; @Turner2015-af; @DeVries2022-of; @Henson2022-hg ; @Wilson2022-bf].
 
@@
 They showed that this relationship was not straigthforward due to the fact that the inclusion of biominerals fragmented aggregates into smaller but denser particles which could either decrease (smaller size) or increase (denser particles) their sinking speed.
+Other studies on the ballasting effect have been published since but even 20 years after its proposition, no global ballasting mechanism has been found [@Neukermans2023-vl].
 

### 2023-02-27 observingCarbonCycle.qmd
update

@@
 
-I will hereafter briefly describe a non exhaustive list of biogeochemical models to illustrate how the export production can be estimated using the $e-ratio$.
+I will hereafter briefly describe a non exhaustive list of biogeochemical model types to illustrate how the export production can be estimated using the $e-ratio$.
 
@@
 
-When we compare the $e-ratio$ maps (@fig-export-ratios) or the carbon export estimations of just six biogeochemical models (see @sec-biogeochem-models), we already observe strong differences.
-In my opinion, it means at least two things.
+When we compare the $e-ratio$ maps (@fig-export-ratios) or the carbon export estimations of only six biogeochemical models (see @sec-biogeochem-models), we already observe strong differences.
 On the one hand, we are seeing an increase in model complexity from empirical models to data-assimilated inverse models.
-At the same time, we also gathered more observational data as well as a deeper understanding of the BCP (@sec-a-new-paradigm-for-the-bcp) that allow us to better represent its underlying processes.
-On the other hand, each degree of complexity propagates more and more uncertainty (@fig-bgc-models) that we need to account for in the results.
+At the same time, we also gathered more observational data as well as a deeper understanding of the BCP (@sec-a-new-paradigm-for-the-bcp) that allows us to better represent its underlying processes.
+On the other hand, each degree of complexity inevitably accumulates more uncertainty due to the increasing number of processing and modeling steps (@fig-bgc-models).
 
@@
 Figure taken from @Siegel2022-wd.](image/ObservingCarbonCycle/carbon_models.jpeg){#fig-bgc-models width="750" height="580"}
+
+From the data we feed models with to uncertainties caused by missing or poorly misunderstood processes, current BCP models still disagree on the estimation of carbon export both currently and in response to climate change [@Henson2022-hg].
+However, the ongoing increase in deployments of autonomous platforms and the development of new sensors is providing new opportunities to enhance the parametrization of key biogeochemical processes to better estimate carbon export and sequestration.
 
@@
 
-This framework will hopefully address the questions raised so far, i.e., better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and estimates of carbon sequestration \[@Boyd2019-ez; @Passow2012-zh\] as well as close the ocean carbon budget \[@Burd2010-hc\].
+This framework will hopefully address the questions raised so far, i.e., better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and estimates of carbon sequestration [@Boyd2019-ez; @Passow2012-zh] as well as close the ocean carbon budget [@Burd2010-hc].
 
@@
 Many studies have assumed that carbon sequestration only occurs in the deep ocean (typically 1000 m).
-In @sec-to-do, I reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
+In @sec-to-do, we reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.

### 2023-02-28 intro.qmd
update

@@
 The global carbon cycle (@fig-global_carbon_cycle) consists of a series of exchanges of carbon in various organic and inorganic forms among four major reservoirs: the lithosphere, the hydrosphere, the atmosphere and the biosphere.
-These exchanges occur at different timescales, such that the carbon cycle can be roughly divided into a slow and a fast component.
-The former operates on geological timescales where carbon is hidden from the surface for millions of years (e.g.
-in hydrocarbon deposits such as petroleum or natural gas) before a part of it is released through volcanic activity while the latter operates on shorter timescales, from minutes to years (e.g.
+These exchanges occur at different time scales, such that the carbon cycle can be roughly divided into a slow and a fast component.
+The former operates on geological time scales where carbon is hidden from the surface for millions of years (e.g.
+in hydrocarbon deposits such as petroleum or natural gas) before a part of it is released through volcanic activity while the latter operates on shorter time scales, from minutes to years (e.g.
 photosynthesis, air-sea fluxes) up to decades and millenia (e.g.
@@
 
-The ocean is known to play a key role in the fast carbon cycle due to its size and rapid exchange of $CO_{2}$ with the atmosphere, controlling atmospheric $CO_{2}$ levels on timescales of decades to millennia [@Sarmiento2006-ec; @DeVries2022-of].
+The ocean is known to play a key role in the fast carbon cycle due to its size and rapid exchange of $CO_{2}$ with the atmosphere, controlling atmospheric $CO_{2}$ levels on time scales of decades to millennia [@Sarmiento2006-ec; @DeVries2022-of].
 Currently, the ocean absorbs \~2.5 Pg C year$^{-1}$, which represents about 25% of annual anthropogenic $CO_{2}$ emissions [@Canadell2021-ds; @Friedlingstein2022-th].
@@
 
-#plot it
-# ggplot(DIC_data, aes(y = mids, x = mean_DIC)) + geom_path(size = 1) + scale_y_reverse() + labs(y = 'Depth [m]', x = "DIC [&mu;mol kg<sup>-1</sup>]") + theme_bw() + theme(
-#     axis.title.x = element_markdown(),
-#     axis.title.y = element_markdown()
-#   )
-
 # https://stackoverflow.com/questions/12514612/how-to-annotate-ggplot-with-latex
@@
 
-This calcification occurs in surface waters, releasing $CO_{2}$ in the surrounding waters and subsequently to the atmosphere (0.26 - 1.19 Pg C year$^{-1}$ following @Neukermans2023-vl).
-That is why this pump is sometimes also referred to as the carbonate counter pump \[@Zeebe2012-zk\].
-However, the dissolution of $CaCO_{3}$ mainly occurs at depth and thus increases the DIC concentration (@eq-calcification).
+This calcification occurs in surface waters, releasing $CO_{2}$ in the surrounding waters and subsequently to the atmosphere (0.26-1.19 Pg C year$^{-1}$ according to @Neukermans2023-vl).
+That is why this pump is sometimes also referred to as the carbonate counter pump [@Zeebe2012-zk].
+However, the dissolution of $CaCO_{3}$ that mainly occurs at depth increases the DIC concentration in contrast with the calcification (see @eq-calcification).
 In all, @Sarmiento2006-ec estimated that the carbonate pump contributes to 20% of the observed DIC gradient.
@@
 Unlike the three previous carbon pumps, the microbial carbon pump (MCP) was only introduced recently by @Jiao2010-ax as "a conceptual framework to address the role of microbial generation of recalcitrant dissolved organic matter[^intro-1] and relevant carbon storage".
-Put simply, it is a suite of marine microbial mechanisms that produce longer-lived DOC such as semi-refractory, refractory and ultra-refractory DOC that have an average lifetime of respectively 205, 16,000 and 40,000 years [@Legendre2015-jc].
+In simple terms, it is a suite of marine microbial mechanisms that produce longer-lived DOC such as semi-refractory, refractory and ultra-refractory DOC that have an average lifetime of respectively 205, 16,000 and 40,000 years [@Legendre2015-jc].
 
@@
 
-The term 'pump' is attributed to the MCP because it maintains a gradient between DOC$_{<100~years}$ (14 Pg C) and DOC$_{>100~years}$ (non-bioavailable DOC, 636 Pg C) where 100 years has been chosen as the minimum time DOC needs to be stored away from the atmosphere to be considered as sequestered [@Ipcc2019-qc].
+The term 'pump' is attributed to the MCP because it maintains a gradient between DOC$_{<100~years}$ (14 Pg C) and DOC$_{>100~years}$ (non-bioavailable DOC, 636 Pg C) where 100 years corresponds to the 100-year time horizon that I use to refer to carbon sequestration.
 
@@
 
-<!-- he recent work of @Nowicki2022-no suggests that the BCP exports on average 10.2 Pg C year$^{-1}$, sequesters ~1,300 Pg C with a mean sequestration time of 127 years and buries 0.4 Pg C year$^{-1}$ in the sediment that will remain sequestered for millions of years.
-Therefore, when talking about natural-based solutions for carbon sequestration, one should keep in mind that even though the BCP does not contribute to the absorption of anthropogenic $CO_{2}$ per se, it remains the main process that maintains the DIC gradient -->
-
 In the frame of the BCP, global carbon budgets have generally been imbalanced such that the exported POC was up to 2-3 orders of magnitude less abundant that it would need to be to sustain heterotrophic respiration in the mesopelagic zone and beyond [@Burd2010-hc; @Turner2015-af].
@@
 
-The last part of the question is out of the scope of this manuscript but it should be said that a lot of uncertainties orbit around those carbon deficits, namely parameters used to estimate the metabolic activity of deep ocean organisms as well as deep food web dynamics and unquantified carbon sources such as the contribution of slowly settling particles and better estimation of POC fluxes [@Burd2010-hc].
+The last part of the question is out of the scope of this manuscript but it should be said that a lot of uncertainties orbit around those carbon deficits, namely parameters used to estimate the metabolic activity of deep ocean organisms as well as deep food web dynamics and unquantified carbon sources such as the contribution of slowly settling particles and better estimation of POC fluxes \[@Burd2010-hc; @Martin2020-wa\].
 
@@
 
-The physically mediated particle injection pump (also referred to as the mixing pump) consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez].
+The physically mediated particle injection pump (also referred to as the mixing pump in the literature) consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez].
 Contrarily to the BGP, all three pumps also transport DOC to depth where it is injected along with suspended (i.e.
-neutrally buoyant) and slow-sinking POC below the EZ on different spatial scales and timescales.
+neutrally buoyant) and slow-sinking POC below the EZ on different spatial and time scales.
 
@@
 
-The two other pumps are both 2D spatial mechanisms (i.e.
+The other two pumps are both 2D spatial mechanisms (i.e.
 they transport POC and DOC vertically and laterally).
-The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on timescales of days across horizontal spatial scales of 1-10 km [@Omand2015-mf; @Resplandy2019-nx].
+The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on time scales of days across horizontal spatial scales of 1-10 km [@Omand2015-mf; @Resplandy2019-nx].
 The large-scale (100-1000 km) subduction pump corresponds to the slow advective transport of carbon associated with the meridional overturning circulation and Ekman pumping [@Levy2013-ro; @Claustre2021-tx; @Siegel2022-wd].
@@
 fecal pellets) below the EZ at the depth of their migration.
-This pump can act on daily as well as on seasonal timescales.
+This pump can act on daily as well as on seasonal time scales.
 In the case of the diel (i.e.
@@
 The VMP would thus export less organic carbon than the mixing pump but its sequestration potential would be higher (@tbl-PumpDepthRange).
-In contrast to the mixing pump that mostly acts on DOC, migrating animals release rapid sinking fecal pellets that would reach depths such that subsequent remineralized DIC would be hidden from the surface for a longer period of time (@tbl-PumpDepthRange).
+In contrast to the mixing pump that mostly acts on DOC, migrating animals release rapid sinking fecal pellets that would reach depths such that subsequent remineralized DIC would be hidden from the surface for longer time periods (@tbl-PumpDepthRange).
 As a result, it is estimated that the diel migration pump could potentially sequester 60% as much carbon as the BGP [@Boyd2019-ez].
@@
 
-### Biological processes at play {#sec-biological-processes-at-play}
+### Biological processes impacting the downward POC flux {#sec-biological-processes-at-play}
 
@@
 Marine snow is paramount for both the BCP (see @sec-a-new-paradigm-for-the-bcp) and the support of marine life within the mesopelagic domain [@Burd2010-hc].
-The remineralization of organic matter by heterotrophic organisms releases $CO_{2}$ in the water column that will then be released back to the atmosphere in a time frame depending both on its geographical location and depth (see @sec-will-it-be-sequestered).
+The remineralization of organic matter by heterotrophic organisms releases $CO_{2}$ in the water column that will be released back to the atmosphere in a time frame depending both on its geographical location and depth (see @sec-will-it-be-sequestered).
 Therefore, understanding how biological processes interact with the downward POC flux and where the bulk of the remineralization occurs is important to better understand the efficiency of the BCP.
@@
 The latter is adimensional and often set at 0.86 [@Martin1987-um] although it is well known that the magnitude of flux attenuation varies regionally (and temporally) due to the variability in food web structures beneath the EZ [@Passow2012-zh].
-Geographically variable [@Henson2012-ck] and regionalized [@Guidi2015-sj] $b$ values have been proposed to take that into account.
+Geographically variable [@Henson2012-ck] and regionalized [@Guidi2015-sj] estimations of $b$ were proposed to account for non-uniform remineralization at global scale.
 
@@
 #| fig-cap: Typical attenuation curve of the POC flux obtained with @eq-Martin-Curve.
-The colour-coded layers represent regions with carbon-specific degradation rate, decreasing with depth.
+The color-coded layers represent regions with carbon-specific degradation rate, decreasing with depth.
 Figure inspired from @Iversen2022-qj.
@@
 
-ggplot(df, aes(x = fluxes, y = depth)) + #geom_path() + 
-  # geom_hline(yintercept = 1000, linetype = 'dotted') + 
-  # geom_hline(yintercept = 100, linetype = 'dotted') +
-  #geom_hline(yintercept = 100, linetype = 'dotted', colour = 'black') +
-  # annotate("rect", ymin = 100, ymax = 300, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#22577a") +
-  # #geom_hline(yintercept = 300, linetype = 'dotted') +
-  # annotate("rect", ymin = 300, ymax = 500, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#38a3a5") + 
-  # #geom_hline(yintercept = 500, linetype = 'dotted') +
-  # annotate("rect", ymin = 500, ymax = 1000, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#57cc99") +
-  # #geom_hline(yintercept = 1000, linetype = 'dotted') +
-  # annotate("rect", ymin = 1000, ymax = 3000, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#80ed99") +
+ggplot(df, aes(x = fluxes, y = depth)) + 
   annotate("rect", ymin = 100, ymax = 300, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#edf8b1") +
@@
   annotate("rect", ymin = 1000, ymax = 3000, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#225ea8") +
-  #scale_y_continuous(trans = c('log10', 'reverse')) +
   scale_y_continuous(trans = c('reverse')) + geom_path(size = 1.5) +
-  #scale_y_reverse() + 
   theme_bw() +
@@
     axis.title.y = element_markdown()) 
-
-# viscosity <- function(T){
-#   visc <- 0.00042844 + 1/(0.157*((T+64.993)**2)-91.296)
-#   return(visc)
-# }
-# 
-# viscosity(0)
-
 ```
@@
 As a result, @Iversen2022-qj suggests that the POC attenuation in the upper mesopelagic layer is the combined action of microbial degradation and zooplankton that graze on settling aggregates.
-In the upper mesopelagic, zooplankton are the main responsible for the attenuation whereas in the mid- and lower mesopelagic layer, the microbial degradation takes over due to the decrease in zooplankton abundance.
+Here, zooplankton are the main responsible for the attenuation whereas in the middle and lower mesopelagic layers, microbial degradation takes over due to the decrease in zooplankton abundance.
 Finally, the relatively constant degradation occurring below 1,000 m in the bathypelagic zone could be explained by the increasing pressure and decreasing temperatures on microbial communities.
@@
 
-### The importance of the sinking speed {#sec-sinking-speed}
+### Physical properties influencing the downward POC flux {#sec-sinking-speed}
 
-The settling speed $w$ is the most important physical parameter controlling the remineralization depth of aggregates and fecal pellets.
-Using Stokes' Law for spherical aggregates at low Reynold's number, one can derive [see for e.g.
-@Johnson1996-ys]
+The settling speed $w$ is the most important physical parameter controlling the remineralization depth (@fig-sinking-speed) of aggregates and fecal pellets [@Laurenceau-Cornec2015-pm].
+Using Stokes' Law for a spherical particle at low Reynolds number, the sinking speed can be expressed as
 
 $$
-w = (1-p)~\frac{g(\rho-\rho_{0})d^{2}}{18\nu\rho_{0}}
+w = \frac{g}{18\mu} (\rho_{p}-\rho_{w})d^{2}
 $$ {#eq-sinking-speed}
 
-where $p$, $g$, $\rho$, $\rho_{0}$, $\nu$ and $d$ are, respectively, the porosity of the aggregate, the gravitational acceleration, the average density of the particles composing the aggregate, the density of the surrounding fluid (i.e.
-seawater), the fluid kinematic viscosity and the aggregate diameter.
-If one assume that changes in dynamic viscosity are negligible [@De_La_Rocha2007-yi] compared to changes from the remaining parameters (i.e.
-this assumption is reasonable based on the fact that $\nu$ can roughly vary by a factor 2 [@Sharqawy2010-ig]), then $w$ should mostly be proportional to the excess density $\Delta\rho$ (i.e.
-$\rho -\rho_{0}$), the porosity and the diameter of sinking aggregates.
+where $g$ is the gravitational acceleration, $\mu$ is the dynamic viscosity of seawater, $\rho_{p}$ and $\rho_{w}$ are the particle and seawater densities and $d$ is the particle diameter.
+If one assumes that changes in dynamic viscosity are negligible [@De_La_Rocha2007-yi] compared to changes from the remaining parameters (i.e.
+this assumption is reasonable based on the fact that $\mu$ can roughly vary by a factor 2 [@Sharqawy2010-ig]), then $w$ should mostly be proportional to the excess density $\Delta\rho$ (i.e.
+$\rho_{p} -\rho_{w}$) and $d$.
 It is however a simplification of reality because most aggregates are not spherical and have fractal geometries [@Logan1990-gy; @Logan1995-id; @Laurenceau-Cornec2020-zc].
@@
 silicate) and calcite or lithogenic material such as dust that are thought to increase $\Delta\rho$ on the one hand and protect them from microbial degradation on the other hand [@Armstrong2001-rz; @Francois2002-uc; @Klaas2002-kn].
-However, several studies argue otherwise.
-@Passow2004-re stipulates that biominerals could instead be collected (i.e.
+However, several studies argued otherwise.
+@Passow2004-re stipulated that biominerals could instead be collected (i.e.
 scavenged) and transported downwards by sinking aggregates.
@@
 
-The size of aggregates was usually thought to be a key predictor of settling velocity [e.g.
-@Passow2006-hw] which is coherent with @eq-sinking-speed.
+The size of aggregates was usually thought to be a key predictor of settling velocity \[e.g.
+@Passow2006-hw\] which is coherent with @eq-sinking-speed.
 However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor.
-With all things considered, the sinking speed of particulate organic matter determines the time available for degradation in the water column.
-Thus, there is a need for direct measurements of the morphological characteristics (size, composition and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Iversen2022-qj].
-<!-- This need could be partially filled with the multi-instrumentation approach described in @sec-observing-carbon-cycle.
--->
+Because the sinking speed determines the time available for degradation in the water column, it also directly impacts the potential for carbon sequestration.
+With all things considered, there is a thus need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities \[@Iversen2020-kb; @Giering2020-zk; @Iversen2022-qj; @Williams2022-xx\].
 
-### Will it be sequestered? {#sec-will-it-be-sequestered}
+![Illustration of the main factors influencing the sinking velocity following Stokes' Law and hypothetical attenuation fluxes (in % of export flux) with depth (vertical axis) for slow and fast sinking particles.
+The latter shows that slow sinking particles experience higher rates of flux attenuation due to increased duration of exposure to remineralization hence contributing less to carbon sequestration.
+Figure taken from @Williams2022-xx.](image/intro/sinking_speed.jpeg){#fig-sinking-speed}
+
+### Sequestration potential {#sec-will-it-be-sequestered}
 
@@
 | Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
-|------------------|------------------|------------------|-------------------|
+|-----------------|------------------|-----------------|--------------------|
 | Gravitational          | export depth - sediment          | None                      | N/A (142)                            |
@@
 sequestration times).
-To come to this conclusion, @Siegel2021-oz used an inverse ocean circulation model to assess the time over which injected $CO_{2}$ (direct injections at different depths) remains sequestered from the atmosphere.
-For this, they computed a factor $f_{\tau}$ corresponding to the fraction of $CO_{2}$ that will be stored away from the atmosphere for a given time $\tau$.
+To come to this conclusion, @Siegel2021-oz used an ocean inverse circulation model to assess the time over which injected $CO_{2}$ (direct injections at different depths) remains sequestered from the atmosphere.
+For this, they computed a factor $f_{\tau}$ corresponding to the fraction of injected $CO_{2}$ that will be stored away from the atmosphere for a given time $\tau$.
 Their results show that $f_{\tau}$ increases with depth and varies geographically.
@@
 
-So far, we have seen that significant incertitudes remain in the estimations of carbon export as well as the sequestration time by the different export pathways composing the BCP.
+So far, we have seen that significant incertitudes remain in the estimations of carbon export as well as the sequestration time by the different export pathways of the BCP.
 To address this issue, biogeochemical models and observation tools are continuously being improved to further our understanding of the BCP (see @sec-modelling-observing-BCP).

### 2023-02-28 observingCarbonCycle.qmd
update

@@
 ![$e-ratio$ estimations obtained with different types of models: empirical models (a, b), food-web models (c, d) and data-assimilated models (e, f).
-Note that the export value computed by @Laws2011-nw (9 Pg C y$^{-1}$) was recomputed by @Siegel2022-wd (6.9 Pg C y$^{-1}$), therefore the $e-ratio$ map presented here might not be representative of the one we would have got with with the initial export value.
+Note that the export value computed by @Laws2011-nw (9 Pg C year$^{-1}$) was recomputed by @Siegel2022-wd (6.9 Pg C year$^{-1}$), therefore the $e-ratio$ map presented here might not be representative of the one we would have got with the initial export value.
 Figure taken from @Siegel2022-wd.](image/ObservingCarbonCycle/export_ratio.jpeg){#fig-export-ratios}
@@
 @Henson2011-mt and @Laws2011-nw (panels a and b in @fig-export-ratios) are both empirical models.
-@Henson2011-mt parameterized the $e-ratio$ as a decreasing function of SST (derived from satellite observations, see @fig-bgc-models) tuned on a compilation of $^{234}Th$-derived export measurement (see @sec-historical-measurements) for a total export estimated at 5 Pg C y$^{-1}$.
+@Henson2011-mt parameterized the $e-ratio$ as a decreasing function of sea surface temperature (SST, derived from satellite observations (see @fig-bgc-models)) tuned on a compilation of $^{234}Th$-derived export measurement (see @sec-historical-measurements) for a total export estimated at 5 Pg C year$^{-1}$.
 @Laws2011-nw modelled the $e-ratio$ as a decreasing linear function of SST and a power-law function of NPP whose parameters were determined using particle export estimates based on estimates of new production (nitrogen-based) and $^{234}Th$ and sediment trap export flux data compiled by @Dunne2005-an.
-They estimated a total export of 9 Pg C y$^{-1}$.
+They estimated a total export of 9 Pg C year$^{-1}$.
 
 @Siegel2014-wg and @Archibald2019-dj (panels c and d in @fig-export-ratios) are both food-web models (see Information Box 3).
-@Siegel2014-wg used a mechanistic approach using satellite observations (NPP, PSS and phytoplankton carbon concentrations, see also @fig-bgc-models) to estimate the export of the BGP (6 Pg C y$^{-1}$).
-@Archibald2019-dj added a new module to the model of @Siegel2014-wg in order to incorporate the diel vertical migration of zooplankton for a total export of 6.5 Pg C y$^{-1}$.
+@Siegel2014-wg used a mechanistic approach using satellite observations (NPP, PSS and phytoplankton carbon concentrations, see also @fig-bgc-models) to estimate the export of the BGP (6 Pg C year$^{-1}$).
+@Archibald2019-dj added a new module to the model of @Siegel2014-wg in order to incorporate the diel vertical migration of zooplankton for a total export of 6.5 Pg C year$^{-1}$.
 
@@
 Examples of forcing functions (among many others) in ecological models are light intensity (represented here by the solar radiation), temperature or wind.
-Finally, a model usually contains output variables (represented here by the Chlorophyll) that allow the comparison of the model to reality @Soetaert2008-er.
+Finally, a model usually contains output variables (represented here by the Chlorophyll) that allow the comparison of the model to reality [@Soetaert2008-er].
 
@@
 
-Figure taken from [@Soetaert2008-er].
+Figure taken from @Soetaert2008-er.
 :::
@@
 satellite and/or in situ data) or by optimizing model parameters to the observed fields [@Siegel2022-wd].
-@DeVries2017-wu combined both satellite-derived data (NPP and PSS) and in situ observations (POC fluxes, DOC concentrations and dissolved $O_{2}$ distributions) to assess the carbon export and its transfer in the mesopelagic zone using a steady state ocean circulation inverse model (OCIM, the circulation component of their model).
-Their carbon export estimate is between 7.3 and 8.6 Pg C y$^{-1}$.
+@DeVries2017-wu combined both satellite-derived data (NPP and PSS) and in situ observations (POC fluxes, DOC concentrations and dissolved $O_{2}$ distributions) to assess the carbon export and its transfer in the mesopelagic zone using a steady state ocean circulation inverse model (i.e.
+the circulation component of their model).
+Their carbon export estimate range between 7.3 and 8.6 Pg C year$^{-1}$, depending on the selected export depth (i.e.
+respectively, 100 m and the euphotic depth).
 @Nowicki2022-no added a two-phytoplankton/two-zooplankton food-web model with a migration zooplankton component to the model of @DeVries2017-wu in order to estimate the magnitude of each export pathway of the BCP.
-In all, @Nowicki2022-no estimated a carbon export of 10 Pg C y$^{-1}$.
+In all, @Nowicki2022-no estimated a carbon export of 10 Pg C year$^{-1}$.
 
@@
 
-When we compare the $e-ratio$ maps (@fig-export-ratios) or the carbon export estimations of only six biogeochemical models (see @sec-biogeochem-models), we already observe strong differences.
-On the one hand, we are seeing an increase in model complexity from empirical models to data-assimilated inverse models.
-At the same time, we also gathered more observational data as well as a deeper understanding of the BCP (@sec-a-new-paradigm-for-the-bcp) that allows us to better represent its underlying processes.
+When we compare the $e-ratio$ maps (@fig-export-ratios) or the carbon export estimations of only six biogeochemical models (see @sec-biogeochem-models), we already observe significant differences.
+On the one hand, we see an increase in model complexity from empirical models to data-assimilated inverse models thanks to the increase in the quantity and quality of observations as well as well as a deeper understanding of the BCP (@sec-a-new-paradigm-for-the-bcp).
 On the other hand, each degree of complexity inevitably accumulates more uncertainty due to the increasing number of processing and modeling steps (@fig-bgc-models).
@@
 
-From the data we feed models with to uncertainties caused by missing or poorly misunderstood processes, current BCP models still disagree on the estimation of carbon export both currently and in response to climate change [@Henson2022-hg].
+From the data we feed models with to uncertainties caused by missing or poorly understood processes, current BCP models still disagree on the estimation of carbon export both currently and in response to climate change [@Henson2022-hg].
 However, the ongoing increase in deployments of autonomous platforms and the development of new sensors is providing new opportunities to enhance the parametrization of key biogeochemical processes to better estimate carbon export and sequestration.
@@
 
-## The potential of autonomous platforms {#sec-potential-autonomous-platforms}
+### The potential of autonomous platforms
 
@@
 
-Remote sensing data acquired with ocean color satellites have been used for a long time [@Gordon1980-dm] to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few (see also @sec-biogeochem-models).
+Remote sensing data acquired with ocean color satellites have been used for a long time [@Gordon1980-dm] to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few (see also @sec-biogeochem-models and @fig-bgc-models).
 While having a wide spatial coverage, these platforms are limited to the ocean's surface and their measures can be affected by several factors such as the cloud cover or low solar elevation (e.g.
@@
 
-## Specific sensors on autonomous platforms for the study of the BCP {#sec-BGC-sensors}
+### Specific sensors on autonomous platforms for the study of the BCP {#sec-BGC-sensors}
 
-The BGC-Argo array (@sec-potential-autonomous-platforms) has already demonstrated its ability to understand some aspects of the BCP [e.g.
+The BGC-Argo array has already demonstrated its ability to understand some aspects of the BCP [e.g.
 @Lacour2017-ro; @Lacour2019-id; @Briggs2020-gz; @Yang2021-vg; @Llort2018-pa; @Wang2022-hn].
@@
 ![Illustration of an augmented BGC-Argo float.
-From top to bottom: GPS & Iridium (two-way communication) antenna, Oxygen optode, CTD (salinity, temperature and depth), pH sensor, radiometer (downwelling photosynthetic available radiation and downwelling irradiance at 380, 412 and 490 nm), fluorometer (Chla and colored dissolved organic matter) and scattering meter ($b_{bp}$), UVP6 (particle size spectra and images), transmissometer (particle beam attenuation) and nitrate sensor.
+From top to bottom: GPS & Iridium (two-way communication) antenna, Oxygen optode, CTD (salinity, temperature and depth), pH sensor, radiometer (downwelling photosynthetic available radiation and downwelling irradiance at 380, 412 and 490 nm), fluorometer (Chla and colored dissolved organic matter) and scattering meter ($b_{bp}$), underwater camera UVP6 (particle size spectra and images), transmissometer (particle beam attenuation) and nitrate sensor.
 \@Thomas Boniface.](image/ObservingCarbonCycle/float_picture1.png){#fig-bgc-float width="400"}
@@
 $$
-c = \frac{ln(T_{r})}{r}
+c = -\frac{ln(T_{r})}{r}
 $$ {#eq-bean-attenuation}
@@
 This method was later improved by [@Estapa2013-gw] who divided the attenuation signal into a continuous component, that is the regular accumulation of small particles on the transmissometer window and a discontinuous component resulting in positive jumps likely attributed to the deposition of (rarer) large particles (see @fig-estapa-method and @sec-case-study-labrador-sea).
-@Estapa2017-jr further calibrated the OST carbon flux proxy using codeployed sediment traps and drifting transmissometers thus allowing the conversion of $c_{p}$ attenuation fluxes (units: m$^{-1}$d$^{-1}$) into carbon fluxes (units: mg C m$^{-2}$d$^{-1}$) using empirical POC-to-attenuance (ATN) ratio were ATN (unitless) is the product of $c_{p}$ and the transmissometer pathlength.
+@Estapa2017-jr [-@Estapa2023-oz] further calibrated the OST carbon flux proxy using codeployed sediment traps and drifting transmissometers thus allowing the conversion of $c_{p}$ attenuation fluxes (units: m$^{-1}$d$^{-1}$) into carbon fluxes (units: mg C m$^{-2}$d$^{-1}$) using empirical POC-to-attenuance (ATN) ratio were ATN (unitless) is the product of $c_{p}$ and the transmissometer pathlength.
 
@@
   ylab(TeX('$c_{p}~(m^{-1})$')) +
-  xlab('Datetime') +
+  xlab('Date') +
   theme(legend.position = 'top', legend.title = element_blank())
@@
 @Guidi2008-vp found an optimal couple of parameters $(A, B) = (12.5, 3.81)$ while @Iversen2010-an followed the same procedure and found $(A, B) = (273, 4.27)$ whereas @Fender2019-ou found $(A, B) = (15.4, 1.05)$.
-@Iversen2010-an assumed that the difference in the optimal couple of parameters was due to the study site used for the minimization (productive upwelling system versus open ocean in @Guidi2008-vp) while @Fender2019-ou suggested that a value of $B < 3$ means that either the density or sinking speed of particles is inversely correlated with particles diameter thus they suspected that the flux was dominated by particles in the lower size detection limit of the UVP they were using.
-Other values for $A$ and $B$ are being used in the literature [e.g.
+@Iversen2010-an assumed that the difference in the optimal couple of parameters was due to the study site used for the minimization (productive upwelling system versus open ocean in @Guidi2008-vp).
+@Fender2019-ou suggested that a value of $B < 3$ means that either the density or sinking speed of particles is inversely correlated with particles diameter thus they suspected that the flux was dominated by particles in the lower size detection limit of the UVP they were using.
+Other values for $A$ and $B$ are used in the literature [e.g.
 @Kiko2017-qy] based on a mechanistic approach developed by @Kriest2002-oc to investigate the formation of marine aggregates.
 
-In all, it seems that the choice for the values of $A$ and $B$ is critical for deriving carbon fluxes derived from PSD [@Iversen2010-an; @Fender2019-ou].
-Three main limitations highlighted by @Bisson2022-bt for UVP optically derived estimates of flux would be: (1) assuming that $A$ and $B$ are globally valid at all depths, for all UVP models and across all size classes, (2) the size to sinking rate uncertainty and (3) the size to carbon mass uncertainty..
+It seems that the choice for the values of $A$ and $B$ is critical for deriving realistic carbon fluxes estimations from PSD [@Iversen2010-an; @Fender2019-ou].
+Three main limitations highlighted by @Bisson2022-bt for UVP optically derived estimates of flux would be: (1) assuming that $A$ and $B$ are globally valid at all depths, for all UVP models and across all size classes, (2) the size to sinking rate uncertainty and (3) the size to carbon mass uncertainty.
 
@@
 
-## Research questions and objectives
+## Research questions and objectives of this work
 
@@
 
-This framework will hopefully address the questions raised so far, i.e., better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and estimates of carbon sequestration [@Boyd2019-ez; @Passow2012-zh] as well as close the ocean carbon budget [@Burd2010-hc].
+This framework will hopefully address the key points raised so far, i.e., better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and refine carbon export and sequestration estimates [@Boyd2019-ez; @Siegel2022-wd].
 
@@
 They are typically deployed for several years, which requires special attention to the energy consumption of each sensor.
-In @sec-uvp-ia I will therefore describe the methodology used to develop the zooplankton classification model for the UVP6 accounting for energy limitations.
+In @sec-uvp-ia, I will therefore describe the methodology used to develop the zooplankton classification model for the UVP6 accounting for energy limitations.
 2.
-4 BGC-Argo floats equipped with both a transmissometer and a UVP6 were deployed for the first time in the Labrador Sea in late May 2022 to investigate the post-spring bloom and its associated POC fluxes.
-In @sec-case-study-labrador-sea, I will show how transmissometers used as OSTs compare with PSDs measured by UVP6s as well as investigate the link between small and big particle fluxes derived by both instruments.
+Four BGC-Argo floats equipped with both a transmissometer and a UVP6 were deployed for the first time in the Labrador Sea in late May 2022 with the aim of observing POC export events due to the annual spring bloom.
+Unfortunately, it appears that the deployment occurred at the end of the spring bloom therefore we likely missed the major export events.
+In @sec-case-study-labrador-sea, I will therefore study the aftermaths of the spring bloom in terms of carbon fluxes estimated by both instruments.
+I will also compare the small (continuous) and big (stochastic) flux of particles as seen by transmissometers used as OSTs with PSDs measurements by UVP6s at three different drifting depths to investigate the relationship between the two instruments.
 3.

### 2023-03-02 intro.qmd
update

@@
 
-The last part of the question is out of the scope of this manuscript but it should be said that a lot of uncertainties orbit around those carbon deficits, namely parameters used to estimate the metabolic activity of deep ocean organisms as well as deep food web dynamics and unquantified carbon sources such as the contribution of slowly settling particles and better estimation of POC fluxes \[@Burd2010-hc; @Martin2020-wa\].
+The last part of the question is out of the scope of this manuscript but it should be said that a lot of uncertainties orbit around those carbon deficits, namely parameters used to estimate the metabolic activity of deep ocean organisms as well as deep food web dynamics and unquantified carbon sources such as the contribution of slowly settling particles and better estimation of POC fluxes [@Burd2010-hc; @Martin2020-wa].
 
@@
 
-The size of aggregates was usually thought to be a key predictor of settling velocity \[e.g.
-@Passow2006-hw\] which is coherent with @eq-sinking-speed.
+The size of aggregates was usually thought to be a key predictor of settling velocity [e.g.
+@Passow2006-hw] which is coherent with @eq-sinking-speed.
 However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor.
 Because the sinking speed determines the time available for degradation in the water column, it also directly impacts the potential for carbon sequestration.
-With all things considered, there is a thus need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities \[@Iversen2020-kb; @Giering2020-zk; @Iversen2022-qj; @Williams2022-xx\].
+With all things considered, there is a thus need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Giering2020-zk; @Iversen2022-qj; @Williams2022-xx].
 
@@
 | Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
-|-----------------|------------------|-----------------|--------------------|
+|-----------------|-------------------|-----------------|---------------------|
 | Gravitational          | export depth - sediment          | None                      | N/A (142)                            |

### 2023-03-02 observingCarbonCycle.qmd
update

@@
 This method was later improved by [@Estapa2013-gw] who divided the attenuation signal into a continuous component, that is the regular accumulation of small particles on the transmissometer window and a discontinuous component resulting in positive jumps likely attributed to the deposition of (rarer) large particles (see @fig-estapa-method and @sec-case-study-labrador-sea).
-@Estapa2017-jr [-@Estapa2023-oz] further calibrated the OST carbon flux proxy using codeployed sediment traps and drifting transmissometers thus allowing the conversion of $c_{p}$ attenuation fluxes (units: m$^{-1}$d$^{-1}$) into carbon fluxes (units: mg C m$^{-2}$d$^{-1}$) using empirical POC-to-attenuance (ATN) ratio were ATN (unitless) is the product of $c_{p}$ and the transmissometer pathlength.
+[@Estapa2017-jr; -@Estapa2023-oz] further calibrated the OST carbon flux proxy using codeployed sediment traps and drifting transmissometers thus allowing the conversion of $c_{p}$ attenuation fluxes (units: m$^{-1}$d$^{-1}$) into carbon fluxes (units: mg C m$^{-2}$d$^{-1}$) using empirical POC-to-attenuance (ATN) ratio were ATN (unitless) is the product of $c_{p}$ and the transmissometer pathlength.
 

### 2023-03-02 uvp6_part.qmd
update

# UVP6 embedded classification algorithm

Introduire le chapitre et dire que le papier est en cours somehow ...

## Embedded recognition of zooplankton using the Underwater Vision Profiler 6 for use on autonomous platforms

Florian Ricour, Marc Picheral, Camille Catalano, Edouard Leymarie, Hervé Claustre, Léo Lacour, Lionel Guidi, Jean-Olivier Irisson

### Abstract

Millions of images of oceanic particles have been acquired with ship-based imaging instruments. The task of labeling images is time consuming but necessary to train classification models using artificial intelligence algorithms. Autonomous Underwater Vehicles (AUVs) such as Biogeochemical-Argo floats (BGC-Argo) have revolutionized our understanding of oceanic systems but they are flawed by their battery life and by the time they can spend at the ocean surface to transmit their data. In this paper, we show how we developed the first zooplankton classification model for AUVs using the Underwater Vision Profiler 6 (UVP6). To this end, we used a features-based classifier called eXtreme Gradient Boosting (XGBoost) adapted to specific energy constraints imposed by profiling floats and we compared its performance with a state-of-the-art convolutional neural network as if we were not energy-limited. In addition, we show how we dealt with a highly imbalanced data set towards marine snow (\> 80% of detritus) and what were the compromises to improve the classification scores of biological classes. Finally, we show the classification report (precision and recall for each label) of the final model and use it to analyse some preliminary results from BGC-Argo floats deployed in the Labrador Sea.

### Introduction

Inside the ocean, a cohort of diverse organisms known as zooplankton thrive. Those tiny beasts are so important for our oceans [@Steinberg2017-ri] that we decided to take pictures of them [@Picheral2010-to; @Gorsky2010-ca; @Cowen2008-wq; @Benfield2007-lt; @Lombard2019-ld]. Though, the amount of images taken by those imaging instruments has become so huge that it is now a massive work to classify them by hand. Fortunately, powerful machine learning techniques have been implemented to tackle the issue of automatic plankton image classification [@Luo2018-oo].

However, it remains a fierce challenge. Within a single taxon, one can find organisms with very different sizes and shapes, organisms with significant morphological differences acquired during their ontogenetic development, partial organisms (e.g. siphonophores) or organisms with various orientations, hindering the automatic classification process [@Benfield2007-lt]. In addition to this global heterogeneity, the great majority of images are not composed of zooplankton. Instead, non-living particles (mostly marine snow) steal the show, producing a highly imbalanced data set towards the latter.

The discussion around automatic plankton classification is a key topic in oceanography and the development of convolutional neural networks (CNNs) within that frame has already lead to promising results [@Luo2018-oo; @Cui2018-nt; @Dai2017-ex; @Campbell2020-cq; @Schroder2019-um; @Bochinski2019-af; @Lee2016-jl; @Cheng2019-ar; @Guo2021-xi; @Ellen2019-nh; @Orenstein2017-ya]. Nevertheless, real-time image processing done at sea is far less mentioned to our knowledge. Usually, images are brought back to the lab and classified with CNNs which, in comparison with features-based classifiers, that is machine learning algorithms using a set of geometric features extracted from the image (e.g. area, eccentricity, grey levels) have outperformed the latter for plankton recognition [@Ellen2019-nh]. Though, it comes at a cost. CNNs are energy consuming hence their implementation for embedded applications can be limited by power consumption [@Biswas2018-uw; @Khabbazan2019-oi; @Oh2017-kj].

This is of particular importance for autonomous underwater vehicles (AUVs) and especially for the next generation of Biogeochemical-Argo (BGC-Argo) floats that will be equipped with the Underwater Vision Profiler 6 (UVP6), the portable version of the UVP5 [@Picheral2010-to] to better understand the carbon cycle [@Xing2018-zo]. The lifetime of floats is limited by their batteries hence by the power consumed by its sensors. Therefore, using an embedded CNN to classify plankton images or sending images through iridium transmission to be classified in the lab would not be possible for economical (cost of iridium transmission) and technical reasons. Typically, a BGC-Argo float samples the water column between the surface and 1000 m, it would therefore have stay at the surface for a time long enough to transmit all images collected after each profile, consuming energy proportionally to the size of the image set while being subject to biofouling or unpredictable hazards. For all these reasons, the development of an embedded plankton recognition algorithm on board the UVP6 is only possible with a features-based classifier at the moment.

In the literature, such algorithms (e.g. Decision Tree (DT), Random Forest (RF), Support Vector Machine (SVM)) have already been used for plankton recognition (see Table 1 of @Gonzalez2017-tr for an exhaustive list). For instance, @Blaschko2005-ep obtained an accuracy of 71% using a SVM on a set of 982 images among 13 classes using a 10-fold cross-validation (CV), @Sosik2007-cj obtained a global accuracy of 88% using a SVM on a test set of 3300 images covering 22 classes, @Bell2008-yo obtained 82% of accuracy using a 10-fold CV RF algorithm on 53 classes with 10 to 30 images per class. However, it should be noted that this accuracy dropped to 63% after the removal of non-biological classes such as detritus and fiber. For bigger training sets, which will be the case in this study, @Orenstein2017-ya\]trained a RF model on approximately 650.000 images across 70 classes for an accuracy of 91% on the test set (2.75 million images). Though, one class in that test set contained 1.6 million images and if the classifier had put every object in that class, the accuracy would have been around 60% so caution must be applied when interpreting performance metrics. @Faillettaz2016-wr trained a RF model on 6000 images across 14 classes, reaching an accuracy of 56% on a test set of 1.5 million objects, discarding those below a certain probability threshold. This allowed to improve the global accuracy by 16% but resulted in the loss of 72% of objects in the predicted data set.

In fine, it is difficult to compare all methods due to the diversity of imaging instruments hence the variety of image formats (e.g. resolution), the number of classes (tradeoff between the model performance and the number of classes as illustrated in @Fernandes2009-hw, the training set (e.g. number of images per class, diversity, balance), the scientific objective (e.g. precision-recall tradeoff), the type and number of features and the algorithm used. In many cases, RF seemed to be the best classifier [@Gorsky2010-ca; @Faillettaz2016-wr] but in embedded applications, energy constraints need to be taken into account. Therefore, the model needs to be small and the classification should be easy and fast. Gradient Tree Boosting (GTB) seems to offer this compromise when compared to RF (model size) and SVM (computation complexity). For these reasons, we chose XGBoost [@Chen2016-xt], an optimized implementation of GTB, also known to be one of the best assets to win Kaggle's machine learning competitions.

In this paper, we present the results of an embedded zooplankton recognition algorithm using XGBoost on UVP5 data. We show how we built the embedded recognition (ER) model under strict technical constraints, what were our compromises and what techniques we used to improve the classification. We also compare the performances of our features-based algorithm with a state-of-the-art neural network as if we were not limited by energy constraints. Finally, we apply the method and follow the recommendations learnt from the preliminary UVP5 model to the UVP6 data set to build an ER model that will be included in the UVP6 on future BGC-Argo floats.

### Material and method

#### Instruments

The UVP6 is the miniaturized version of the UVP5 [@Picheral2021-sx]. It is smaller, it uses less energy and it has been specifically developed to be mounted on AUVs such as BGC-Argo floats with the aim of providing data of similar quality than from the UVP5. There are 2 versions: the LP (low power) and the HF (high frequency). The former uses less energy (lower image acquisition rate) and will be deployed on AUVs while the latter is designed to be used on CTD rosettes and cruising AUVs [@Picheral2021-sx].

![UVP5 and UVP6 profile coverage used in this study. UVP5 profiles used in the test set are depicted by green crosses.](image/uvp6/map_of_profiles.png){#fig-map-of-profiles}

UVP5 Standard Definition (SD) profiles have been collected since 2008 at a global scale (@fig-map-of-profiles) whereas the UVP6 has been deployed very recently in specific locations and thus lacks spatial coverage to be fully representative. For this reason, UVP5 data will be used to create a preliminary ER model and explore ways of improvement. Ultimately, we will train the ER model for the UVP6 by applying observations made on the preliminary UVP5 model.

#### Data preparation

The full UVP5 SD data set comprises 3.578.901 validated images in 2273 profiles. The full UVP6 LP data set comprises 634.459 validated images in XX profiles (all UVP6 profiles were not entirely validated). Each image was segmented to find the main object (i.e. the biggest connected region in pixels). From it, we extracted 55 features describing this object in terms of shape (e.g. pixel area, eccentricity), grey level and mostly statistical moments (all features are available in **Fig. SXX**). For technical reasons, all UVP6 images with a pixel area below 79 pixels were removed (7% of images).

Validated objects were manually classified into specific taxa belonging to well known biological groups such as Rhizaria, Copepoda or Cnidaria to name a few. However, the maximum number of classes (also referred to as labels or taxa) for the ER was fixed to 40 in order to limit the size of the ER model and the volume of data that needs to be transmitted. We chose to aggregate classes iteratively to find the best number of classes for the ER. The first level (ie. all labels (\> 40) are retained) served as a baseline for a prior model, without making any hypotheses. The next level was used as a first aggregation step based on the number of objects in each class and the purity of the class. Below a given threshold, rare classes were merged with their most look-alike taxa and parents on the phylogenetic tree were removed when their children were abundant (i.e. parents could not be classified to a deeper taxonomic level, likely causing confusion with their children during training). The second level of aggregation was based on statistical metrics post training (see hereafter how a model is trained), meaning that we looked at the confusion matrix and the classification report (e.g. **Fig. XX and XX**, respectively) to see where classes did not perform well such that they had to be merged to improve the classification score. This last step was repeated until an agreement was found. Thus, classes were merged only if they were confused amongst the same biological group on the phylogenetic tree or if they had a similar ecological role. This involves that some classes end up with a variety of organisms that do not necessarily show similar morphological traits (e.g. other\<living). It is definitely not ideal but we were bound to keep as many biological classes as possible to counterbalance non-living classes.

In machine learning tasks, it is common to split the data into a training set, a validation set and a test set. The validation set is usually used to assess the quality of the model by tuning hyperparameters while avoiding overfitting (stopping the training before the model is too acquainted with the data). Ideally, the test set is not to be used during any step except at the very end where it is used to assess the global quality of the final model. In our case, both datasets (UVP5 and UVP6) are very imbalanced towards few non-biological classes (mostly detritus). As a result, in order for the model to see as much biological data as possible during training, we did not use a validation set but instead we did a threefold cross-validation.

One objective of the ER on board the UVP6 is to be used on AUVs such as BGC-Argo floats to study the vertical migration pump [@Claustre2021-tx]. Hence, we built the test set with entire UVP5 profiles going at least to 1000 m depth, the typical parking depth of BGC-Argo floats. Those profiles were also chosen to be representative of the global ocean (**see Fig. XX**). For UVP6 data, we did not select entire profiles because we were limited by the number of images in comparison with the UVP5 to build the biggest training set possible (80% of detritus in the UVP6 data set). Therefore, we splitted the data set in a 10-90 % stratified test-train fashion regardless of specific profiles.

Following the previous steps, the UVP5 test and training sets are composed, respectively, of 315.277 and 3.023.506 images while the UVP6 has 59.054 and 531.483 images for its test and training sets.

### Gradient boosting and model training

GTB generates a sequence of trees, each one trying to cope with the errors made by their predecessors. XGBoost is designed to be an optimized and regularized version of GTB. To learn from the data and derive a model, XBoost first minimizes a regularized objective function $L$

$$
L = \sum_{i}l(\hat{y}_{i},y_{i}) + \sum_{k}\Omega(f_{k})
$$ {#eq-objective-function-xgboost}

with

$$
\Omega(f) = \gamma T + \frac{1}{2}\lambda||w||^{2} 
$$

where $l$ is a loss function that measures the difference between the prediction $\hat{y}_{i}$ and the true value $y_{i}$ and $\Omega$ is the regularization term, for each $k$ decision tree $f_{k}$. $\Omega$ is then splitted into two terms where $\gamma T$ penalizes the complexity of a tree proportionally to the number of leaves ($T$) and where the L2 norm penalizes the weights ($w$) of the leaves to avoid over-fitting (both $\gamma$ and $\lambda$ are hyperparameters that can be tuned by the user). By doing so, XGBoost encourages both a simple and predictive model. From @eq-objective-function-xgboost, a scoring function that measures the quality of a tree structure can be derived (see Eq. 6 in @Chen2016-xt for details). However, in practice, it is not possible to test all suitable trees because there are infinite possibilities. Instead, the algorithm starts from a single leaf and iteratively adds branches to the tree, searching for the best split options.

To build our ER model, we first did a threefold cross-validation with XGBoost to find the optimal number of boosting rounds (i.e. $k$) before the model overfits. For this, we looked at the minimum of the multi-class log loss curve (**Fig. S3**), then we trained the ER model with the entire training set at this optimal number with specific hyperparameters (ie. learning rate ($\eta$), maximum depth of a tree ($D_{max}$) and weight sensitivity ($\omega$). The optimal couple ($\eta$, $D_{max}$) was determined using a customary grid search (**see Fig. 3**) while $\omega$ was chosen empirically (see hereafter).

To cope with our unbalanced dataset (more than 70% of detritus in the UVP5 dataset), several options were available. We could artificially increase the number of biological data with the well known SMOTE oversampling technique [@Chawla2002-kg; @Fernandez2018-ft], decrease the number of detritus (random undersampling, keep only detritus close to unsupervised cluster centroids, etc.) or add a weight to each label during training to nudge the model towards under-represented classes (i.e. biological classes). In order to compare a features-based classifier with a state-of-the-art CNN, we chose the latter.

$$
class~weight = \left(\frac{size~max}{class~size}\right)^{\omega}
$$ {#eq-class-weight}

where $size~max$ and $class~size$ are, respectively, the size of the biggest class (i.e. detritus) and the size of each class. Hence, the weight of detritus is 1 and all other classes have a weight \> 1.

We decided to select the best $\omega$ among 0 (no weights), 0.25, 0.50 and 0.75 (model very biased towards biological classes). In order to decipher between all options, we compared the predicted concentrations of classes for each $\omega$ with the real concentration of each class in different regions of the global ocean. As always in machine learning, we had a tradeoff between precision and recall but because the UVP6 is more focused on the estimation of carbon fluxes (high precision) than on the detection of rare objects (high recall with a low precision), we selected $\omega$ = 0.25 to be a satisfactory compromise.

In order to compare our features-based model with a state-of-the-art neural network not limited by energy constraints, we used a MobileNetV2 (depth multiplier 1.40) pre-trained on ImageNet. This is known as transfer learning and it was already applied with success for zooplankton classification in @Orenstein2017-ya. Input images have a size of \[224, 224, 3\], the size of the features extractor vector is 1792 followed by a fully connected layer (600 nodes) with a dropout of 0.4 to avoid overfitting. The classification layer at the end of the network also has a dropout of 0.2. The CNN was trained with a decaying learning rate,  a Categorical Cross Entropy loss function and weights were defined with the ω = 0.25. The CNN was trained with a sufficient number of epochs such that it was not overfitting and it was not learning much anymore (validation plateau, see Fig. S5).

![Scheme of the UVP6 pipeline with the processing done in the lab to create the ER model and the processing time (in ms) of the main steps when the UVP6 is deployed onboard an AUV. kpx means kilo pixel.](image/uvp6/fig_pipeline_UVP6.jpg){#fig-uvp-pipeline}

### Results

It is well known that the task of labelling images in supervised machine learning tasks is time consuming and most probably painful when you have more than 3 million images in your data set. It is also very frustrating to see that one class (largely) dominates when you are interested in the small ones and that some of them are so scarce that you need to merge them with their siblings (at the very best) on the phylogenetic tree. All in all, we found - after multiple trials and errors - 11 classes fulfilling a first compromise between classification performances and ecological importance: 7 biological classes (*Acantharea*, *Aulacanthidae*, *Aulosphaeridae*, *Copepoda*, *Trichodesmium*, darksphere and other\<living) and 4 non-living classes (detritus, bubble, fiber and artefact).

Once the number of classes is known, hyperparameters can be tuned to find the optimal couple for the classification results (ie. grid search results shown in Fig. 3). In our case, the optimal model hyperparameters were 0.2 for η and 5 for Dmax (also with only 1 class of detritus, see Discussion) to have both a good weighted living (or biological) precision and recall and a loading time (ie. loading of the ER model in memory) below 700 ms for energetic reasons. In that configuration, the model has an accuracy of 81%, a macro (ie. label imbalance not taken into account) precision and recall of, respectively, 64% and 69% and a weighted living precision and recall of, respectively, 60% and 65%. CV found an optimal number of 850 boosting rounds before overfitting and the expected loading time is 382 ms.

INSERT FIGURE 3

INSERT FIGURE 4

INSERT FIGURE 5 -\> SWITCH FOR A TABLE !!

The classification report (CR) gives the precision, the recall, the f1-score (harmonic mean of the precision and recall) and global metrics on the test set. As it can be seen in Fig. 5, detritus have a recall of 86% thus because detritus are abundant (more than 230.000 in the test set), it results in more than 30.000 detritus wrongly labelled and therefore disseminated in all other classes. The confusion matrix (CM) helps to see where those wrongly labelled images are. Fig. 4 shows that a lot of Copepoda are confused with detritus as well as fiber. The worst class in the CR is other\<living, which is not surprising because this class is composed of organisms that are not all morphologically and ecologically related. The best biological class is Aulosphaeridae with a f1-score of 82%.

In comparison with a CNN trained on UVP5 images, the CR (Fig. S2) shows that the global accuracy is 1% higher (82%) with a macro precision and recall of, respectively, 65% and 74% and a weighted living precision and recall of, respectively, 59% and 78%. *Copepoda* are better detected (higher recall) but are still often (56% precision) misclassified. The *Aulosphaeridae* class has improved its f1-score to 88%. The class other\<living has also greatly improved even though this class cannot be interpreted further in terms of ecology. Overall, the CNN performs slightly better on biological (living) classes than XGBoost if we consider the f1-score but this 4% increase is mostly due to an increase in global recall and not in precision. For instance, the 14% increase in recall for *Trichodesmium* is followed by a 12% decrease in precision so there is still a lot of confusion at this stage for the UVP5 for some classes, even with a CNN that is not limited by any energy constraints.

### Discussion

#### Ways to improve the classification

The on board classification of plankton is definitely not an easy task. In order to improve the classification scores of biological classes, we tried several adjustments such as quantifying planktonic samples, splitting the detritus class into several subgroups and assigning weights to classes proportional to their size, which was the only option that enhanced the recognition of scarce classes (see Methods).

Quantification is an interesting concept that focuses on the sample instead of the individual. It was already shown in @Solow2001-jc, using plankton images from the Video Plankton Recorder, that the estimation of taxonomic abundances could be improved by computing the confusion matrix $\pmb{P}$ where $p_{i,j}$ is the probability that a taxon $i$ is classified as taxon \$j\$. If the classifier is perfect, $p_{i,j} = 1$ where $i = j$, the quantification resumes to a simple Classify and Count method (CC, see @Forman2005-eu). However, a perfect (especially in plankton ecology) classifier does not exist. Therefore, quantification is a tool to better estimate the abundance of planktonic groups while perfectly knowing that the classifier is biassed. This quantification method was successively applied in the paper of @Sosik2007-cj on a 2 month data set although it is very likely that P needs to be adapted to the sampling conditions. This reevaluation of $\pmb{P}$ is due to what is defined as a data set shift where part of the unseen data follows a different distribution than the one from the training data [@Orenstein2020-jl; @Moreno-Torres2012-oq]). It is especially true for ER systems deployed on AUVs where planktonic communities change throughout the year. Thus, we tested different quantification methods (CC, adjusted count (AC), threshold count (TC), probabilistic adjusted count (PAC) and scaled probability adjusted count (SPAC), see @Gonzalez2017-tr for a review on those quantification methods) assuming that the estimation of $\pmb{P}$ was accurate enough (\> 3M images in our training set) and that the global coverage of UVP5 profiles in the training set was sufficient to cover various planktonic distributions from the test set.

We compared the abundance estimations of all quantification methods on the entire test set, all profiles included (\> 300k images). The SPAC showed the best estimations whereas TC drastically underestimated them (85% of data were lost using a probability threshold of 0.9). However, the aim of the ER is to be deployed on AUVs sampling vertical profiles in the water column therefore we applied the SPAC on 10.000 profile-like subsets and we summed all taxonomic abundances. In that configuration, the SPAC overestimated the real total abundance by 1 to 15% depending on the projected (high or low) number of objects in each profile. These percentages were literally exploding for small classes in both projections (also with PAC) therefore quantification was discarded for use on autonomous profilers at this stage.

Increasing the number of detritus classes was also considered in the hope of clustering them in the features space. As shown in Fig. SXX, detritus display a variety of size, shape and grey levels so it must be possible to cluster them and therefore improve the classification score. For this, we applied a classic K-Means algorithm [@Hartigan1979-oz] on detritus cleaned from likely outliers using an Isolation Forest Algorithm [@Liu2012-xz] and we executed a grid search as described earlier. Fig. 3 clearly shows that using more than 1 class of detritus drastically decreases the global accuracy of the model, decreases the weighted living recall and only increases by 1% the weighted living precision. In addition, the confusion matrix (Fig. S4) shows that it does not improve the ER model because there is a high confusion between subgroups of detritus. Essentially, it seems that detritus form a continuum in the features space and that more research should be done on the choice of features, clustering algorithms and related methods to better differentiate them in separate entities, which is not the purpose of this study.

#### Application to the UVP6 data set

The aim of building an ER model was intended towards its implementation on the UVP6 for future deployments on BGC-Argo floats (or other profiling AUVs). Based on what was done with the UVP5 data set, we applied the same method for UVP6 data. Thus, we started with the data aggregation step (choices of classes) where, in contrast with UVP5 data, some classes were kept (see CR in Fig. 7) because of their ecological role (e.g. small\<*Cnidaria* or tuff, an elongated cyanobacteria belonging to *Trichodesmium*) even though their classification metrics (precision and recall) were not particularly convincing. In the end, we reached an agreement on 20 classes. Among them, there are 16 biological classes (*Acantharia*, *Actinopterygii*, *Appendicularia*, *Aulacanthidae*, *Calanoida*, *Chaetognatha*, *Collodaria*, *Creseis*, *Foraminifera*, *Rhizaria*, *Salpida*, other\<living, tuff and puff \[another *Trichodesmium*\], small\<*Cnidaria*, solitaryglobule \[part of the *Rhizaria* group\] and 4 non-living classes (detritus, fiber, artefact, crystal). One will notice that we have the *Rhizaria* class as well as *Acantharia*, *Collodaria* and solitaryglobule. Those children of *Rhizaria* on the phylogenetic tree showed satisfying classification metrics compared to other *Rhizaria* subgroups (e.g. *Aulosphaeridae*, *Phaeodaria*, *Coelodendridae*, etc) hence we decided to separate them from their parent. Therefore, we assume that those classes are well recognized by human classifiers so that their parent *Rhizaria* is virtually free from them. Then, we executed a grid search to find the optimal couple $\eta$ and $D_{max}$ ($\omega$ and the number of classes of detritus were fixed based on UVP5 results) and trained the final ER model whose CM and CR are reported hereafter for further interpretation.

The specifications of the final UVP6 ER model are the following: \$\\eta\$: 0.2, $D_{max}$: 3, $\omega$: 0.25, number of boosting rounds: 400 for a total loading time of 73 ms. Note that the CV initially found an optimal number of boosting rounds of 555 (loading time: 101 ms) but the log loss curve had already reached a plateau so we were able to decrease the size of the model without losing much (\< 1%) in biological precision or recall, allowing to decrease a bit more the energy consumption of the UVP6 on board a BGC-Argo float for an extended battery life. For the interested reader, results from a CNN trained on UVP6 data are available in Fig. S6 and S7.

INSERT FIGURE 6

INSERT FIGURE 7

#### How to use the CR?

4 BGC-Argo floats equipped with the ER were deployed in late May 2022 in the southern part of the Labrador Sea at the time of the spring bloom. As it can be seen in Fig. 8, *Calanoida* are more abundant around the deep chlorophyll maximum (DCM) and they would be observed in very low concentrations even below 1000 m. The CR for *Calanoida* shows that the model has a precision of 66% and a recall of 69% for that class. It implies that caution must be applied when interpreting the presence or absence of such copepods. In that case, we theoretically miss 31% of them (ie. 31% of *Calanoida* are not detected as such, they are classified in another class) and out of the 69% that we detect as *Calanoida*, we are wrong 34% of the time. As a result, even though *Calanoida* have been shown to overwinter in deep waters (\> 1000 m) [@Krumhansl2018-dr], an isolated signal at depth does not mean a copepod was really there. Similarly, *Trichodesmium* should not be observed outside the tropical and subtropical regions @Westberry2006-bt therefore depending on the sampling region, some results can directly be rejected. Here, tuff are predicted in higher concentrations than puff which are only predicted at one depth. Again, if one looks at the CR for these 2 classes, one can see that the precision and recall for tuff are, respectively, 18% and 15% below those of puff. Finally, the interpretation of the detritus class is more straightforward because they have excellent metrics in the CR. For instance, it can be seen that detritus nicely follow the DCM curve. In contrast with biological classes where we are looking for patterns of presence or absence, we know that detritus are ubiquitous but their successive abundances could be interesting for flux studies.

INSERT FIGURE 8

#### Prospects

The ER of zooplankton on board AUVs is just at its beginning. As expected, the CNN trained on UVP6 data (Fig. S7) outperforms the results of our features-based classifier. However, it should be noted that not all biological classes are better classified (e.g. *Actinopeterygii*, *Appendicularia*, *Chaetognatha*, *Creseis*). One can also observe that the difference between XGBoost and the CNN trained on UVP5 data on biological classes is less significant than for UVP6 data. This is likely explained by the fact that we were able to keep more (small) classes for the UVP6. Nevertheless, previous tests on UVP5 data with more classes increased the gap between XGBoost and CNN weighted results therefore the latter recognises better small classes. If energy and hardware constraints limit the performances of our current ER model, it is not forbidden to think about light-weight CNNs for future applications on AUVs to better understand the dynamics and the behaviour of zooplankton.

### Conclusion

The objective of this paper was to develop a zooplankton classification model under strict energy constraints for the UVP6 to be deployed on BGC-Argo floats. It was clear at the beginning that the use of powerful energy consuming CNNs had to be discarded and that we needed to build a satisfactory features-based model instead. To this aim, we firstly used XGBoost on UVP5 data, a more spatially and temporally representative data set than the recent UVP6 data set to create the data processing pipeline (classes aggregation, hyperparameter tuning, model training) and investigate ways to improve classification scores (quantification, detritus subgroups, label weights). In order to see how the model could be improved, we compared the XGBoost model with results from a CNN not limited by energy. Afterwards, we applied the data processing pipeline and the recommendations from the UVP5 data set analysis on the UVP6 data set to create the final ER model. This model contains 20 labels, 16 biological and 4 non-living classes, with a total loading time of 73 ms. The initial upper limit for the loading time was 700 ms therefore we are saving precious energy to extend the lifetime of BGC-Argo floats. Finally, we showed the CM and CR of the final model and how to use it to interpret some UVP6 taxonomic profiles observed by the float WMO 4903634 during the spring bloom in the southern part of the Labrador Sea.

### Supplementary material

### 2023-03-05 merci.qmd
remaster uvp6 article

remerciement

thomas boniface pour les figures

anais

alexbarth pour son bouquin sur le machine car j'ai passé mes 3 première semaines à faire ça vu que j'y connaissais absolument rien

faire des stats du vélo (et meme mettre une carte si possible rapidement)

### 2023-03-05 uvp6_part.qmd
remaster uvp6 article

@@
-# UVP6 embedded classification algorithm
+# UVP6 embedded classification algorithm {#sec-uvp6}
 
@@
 In this paper, we show how we developed the first zooplankton classification model for AUVs using the Underwater Vision Profiler 6 (UVP6).
-To this end, we used a features-based classifier called eXtreme Gradient Boosting (XGBoost) adapted to specific energy constraints imposed by profiling floats and we compared its performance with a state-of-the-art convolutional neural network as if we were not energy-limited.
+To this end, we used a features-based classifier called eXtreme Gradient Boosting (XGBoost) adapted to specific energy constraints imposed by profiling floats and we compared its performance with a state-of-the-art convolutional neural network as if energy was not a limiting factor.
 In addition, we show how we dealt with a highly imbalanced data set towards marine snow (\> 80% of detritus) and what were the compromises to improve the classification scores of biological classes.
-Finally, we show the classification report (precision and recall for each label) of the final model and use it to analyse some preliminary results from BGC-Argo floats deployed in the Labrador Sea.
+Finally, we show the classification report (precision and recall for each label) of the final classification model and use it to analyse some preliminary results from BGC-Argo floats deployed in the Labrador Sea.
 
@@
 Those tiny beasts are so important for our oceans [@Steinberg2017-ri] that we decided to take pictures of them [@Picheral2010-to; @Gorsky2010-ca; @Cowen2008-wq; @Benfield2007-lt; @Lombard2019-ld].
-Though, the amount of images taken by those imaging instruments has become so huge that it is now a massive work to classify them by hand.
+Nonetheless, the amount of images taken by those imaging instruments has become so huge that it is now a massive work to classify them by hand.
 Fortunately, powerful machine learning techniques have been implemented to tackle the issue of automatic plankton image classification [@Luo2018-oo].
@@
 Nevertheless, real-time image processing done at sea is far less mentioned to our knowledge.
-Usually, images are brought back to the lab and classified with CNNs which, in comparison with features-based classifiers, that is machine learning algorithms using a set of geometric features extracted from the image (e.g.
+Usually, images are brought back to the lab and classified with CNNs which, in comparison with features-based classifiers, that is machine learning algorithms using a set of features extracted from the image (e.g.
 area, eccentricity, grey levels) have outperformed the latter for plankton recognition [@Ellen2019-nh].
@@
 
-This is of particular importance for autonomous underwater vehicles (AUVs) and especially for the next generation of Biogeochemical-Argo (BGC-Argo) floats that will be equipped with the Underwater Vision Profiler 6 (UVP6), the portable version of the UVP5 [@Picheral2010-to] to better understand the carbon cycle [@Xing2018-zo].
+This is of particular importance for autonomous underwater vehicles (AUVs) and especially for a new version of Biogeochemical-Argo (BGC-Argo) floats that will be equipped with the Underwater Vision Profiler 6 (UVP6), the portable version of the UVP5 [@Picheral2010-to] to better understand the carbon cycle [@Xing2018-zo].
 The lifetime of floats is limited by their batteries hence by the power consumed by its sensors.
 Therefore, using an embedded CNN to classify plankton images or sending images through iridium transmission to be classified in the lab would not be possible for economical (cost of iridium transmission) and technical reasons.
-Typically, a BGC-Argo float samples the water column between the surface and 1000 m, it would therefore have stay at the surface for a time long enough to transmit all images collected after each profile, consuming energy proportionally to the size of the image set while being subject to biofouling or unpredictable hazards.
-For all these reasons, the development of an embedded plankton recognition algorithm on board the UVP6 is only possible with a features-based classifier at the moment.
+Typically, a BGC-Argo float samples the water column between the surface and 1000 m, it would therefore have to stay at the surface for a time long enough to transmit all images collected after each profile, consuming energy proportionally to the size of the image set while being subject to biofouling or unpredictable hazards.
+For all these reasons, it was decided to use a features-based classifier for the development of the embedded plankton recognition algorithm on board the UVP6.
 
@@
 However, it should be noted that this accuracy dropped to 63% after the removal of non-biological classes such as detritus and fiber.
-For bigger training sets, which will be the case in this study, @Orenstein2017-ya\]trained a RF model on approximately 650.000 images across 70 classes for an accuracy of 91% on the test set (2.75 million images).
+For bigger training sets, which will be the case in this study, @Orenstein2017-ya trained a RF model on approximately 650.000 images across 70 classes for an accuracy of 91% on the test set (2.75 million images).
 Though, one class in that test set contained 1.6 million images and if the classifier had put every object in that class, the accuracy would have been around 60% so caution must be applied when interpreting performance metrics.
@@
 
-In fine, it is difficult to compare all methods due to the diversity of imaging instruments hence the variety of image formats (e.g.
-resolution), the number of classes (tradeoff between the model performance and the number of classes as illustrated in @Fernandes2009-hw, the training set (e.g.
+It is quite difficult to compare all methods due to the diversity of imaging instruments hence the variety of image formats (e.g.
+resolution), the number of classes (trade off between the model performance and the number of classes as illustrated in @Fernandes2009-hw, the training set (e.g.
 number of images per class, diversity, balance), the scientific objective (e.g.
-precision-recall tradeoff), the type and number of features and the algorithm used.
+precision-recall trade off), the type and number of features and the algorithm used.
 In many cases, RF seemed to be the best classifier [@Gorsky2010-ca; @Faillettaz2016-wr] but in embedded applications, energy constraints need to be taken into account.
@@
 We show how we built the embedded recognition (ER) model under strict technical constraints, what were our compromises and what techniques we used to improve the classification.
-We also compare the performances of our features-based algorithm with a state-of-the-art neural network as if we were not limited by energy constraints.
-Finally, we apply the method and follow the recommendations learnt from the preliminary UVP5 model to the UVP6 data set to build an ER model that will be included in the UVP6 on future BGC-Argo floats.
+We also compare the performances of our features-based algorithm with a state-of-the-art neural network as if energy was not a limiting factor.
+Finally, we apply the method and follow the recommendations learnt from the preliminary UVP5 model to the UVP6 data set to build the ER model that will be included in the UVP6 for AUVs.
 
@@
 The UVP6 is the miniaturized version of the UVP5 [@Picheral2021-sx].
-It is smaller, it uses less energy and it has been specifically developed to be mounted on AUVs such as BGC-Argo floats with the aim of providing data of similar quality than from the UVP5.
+It is smaller, it uses less energy and it has been specifically developed to be mounted on AUVs such as BGC-Argo floats with the aim of providing data of similar quality than from the UVP5 (@fig-uvp-pipeline).
 There are 2 versions: the LP (low power) and the HF (high frequency).
@@
 
-![UVP5 and UVP6 profile coverage used in this study. UVP5 profiles used in the test set are depicted by green crosses.](image/uvp6/map_of_profiles.png){#fig-map-of-profiles}
+![Scheme of the UVP6 pipeline with the processing done in the lab to create the ER model and the processing time (in ms) of the main steps when the UVP6 is deployed on board an AUV.
+kpx means kilo pixel.](image/uvp6/fig_pipeline_UVP6.jpg){#fig-uvp-pipeline}
 
@@
 
+![UVP5 and UVP6 profile coverage used in this study. UVP5 profiles used in the test set are depicted by green crosses.](image/uvp6/map_of_profiles.png){#fig-map-of-profiles}
+
 #### Data preparation
 
-The full UVP5 SD data set comprises 3.578.901 validated images in 2273 profiles.
-The full UVP6 LP data set comprises 634.459 validated images in XX profiles (all UVP6 profiles were not entirely validated).
+The full UVP5 SD data set is composed of 3.578.901 validated images whereas the full UVP6 LP data set has 634.459 validated images.
 Each image was segmented to find the main object (i.e.
 the biggest connected region in pixels).
-From it, we extracted 55 features describing this object in terms of shape (e.g.
-pixel area, eccentricity), grey level and mostly statistical moments (all features are available in **Fig.
-SXX**).
+From it, 55 features were extracted to describe this object in terms of shape (e.g.
+pixel area, eccentricity), gray level and mostly statistical moments (see @tbl-features).
 For technical reasons, all UVP6 images with a pixel area below 79 pixels were removed (7% of images).
+
+```{r, features, dpi = 300}
+#| label: tbl-features
+#| tbl-cap: List of 55 features (abbreviations used in our python package UVPec (https://github.com/ecotaxa/uvpec) and their descriptions) extracted from each object based on the object geometrical characteristics, the pixel gray values histogram and statistical moments.
+
+# library(gt)
+# library(gtExtras)
+#library(flextable)
+library(kableExtra)
+library(tibble)
+
+# object_attributes <- tibble(attr = c('area', 'width', 'height', 'x','y','xm','ym','major','minor', 'angle', 'eccentricity', 'esd', 'bbox_area', 'extent'), comment = c('number of pixels', 'bounding box width', 'bounding box height', 'local centroid (x direction)', 'local centroid (y direction)', 'weighted local centroid (x direction)', 'weighted local centroid (y direction)', 'major axis', 'minor axis', 'orientation angle','-', 'equivalent spherical diameter', 'bounding box area', 'pixel ratio between object and bounding box'))
+# 
+# hist_attributes <- tibble(attr = c('mean', 'stddev', 'mode', 'min', 'max', 'intden', 'median','histcum1', 'histcum3', 'vrange', 'meanpos', 'cv', 'sr'), comment = c('-', 'standard deviation', '-', '-', '-', 'sum of all pixel values', '-', 'first quartile', 'third quartile', 'vmax - vmin', '(vmax - vmin)/range', '100*(stddev/mean)', '100*(stddev/vrange)'))
+# 
+# stat_moment <- tibble(attr = c('hu_moment-i (i: 1 to 7)', 'gray_hu_moment-i (i: 1 to 7)',
+#                                'central_moment-2-0',
+#                                'central_moment-1-1',
+#                                'central_moment-0-2',
+#                                'central_moment-3-0',
+#                                'central_moment-2-1',
+#                                'central_moment-1-2',
+#                                'central_moment-0-3',
+#                                'gray_central_moment-2-0',
+#                                'gray_central_moment-1-1',
+#                                'gray_central_moment-0-2',
+#                                'gray_central_moment-3-0',
+#                                'gray_central_moment-2-1',
+#                                'gray_central_moment-1-2',
+#                                'gray_central_moment-0-3'),
+#                       comment = c('Hu moments (translation, scale and rotation invariant)', 'weighted Hu moments', 'central moments (translation invariant)', '-', '-', '-', '-', '-', '-', 'weighted central moments', '-', '-', '-', '-', '-', '-'))
+
+object_attributes <- tibble(attr = c('area', 'width', 'height','major','minor', 'angle', 'eccentricity', 'esd', 'bbox_area', 'extent'), comment = c('number of pixels', 'bounding box width', 'bounding box height', 'major axis', 'minor axis', 'orientation angle','eccentricity', 'equivalent spherical diameter', 'bounding box area', 'pixel ratio between object and bounding box'))
+
+hist_attributes <- tibble(attr = c('mean', 'stddev', 'mode', 'min', 'max', 'x','y','xm','ym','intden', 'median','histcum1', 'histcum3', 'vrange', 'meanpos', 'cv', 'sr'), comment = c('mean pixel intensity', 'standard deviation of pixel intensity', 'pixel intensity mode', 'min pixel intensity', 'max pixel intensity', 'local centroid (x direction)', 'local centroid (y direction)', 'weighted local centroid (x direction)', 'weighted local centroid (y direction)', 'sum of all pixel values', 'median pixel intensity', 'pixel intensity first quartile', 'pixel intensity third quartile', 'vmax - vmin', '(vmax - vmin)/range', '100*(stddev/mean)', '100*(stddev/vrange)'))
+
+stat_moment <- tibble(attr = c('hu_moment-i (i: 1 to 7)', 'gray_hu_moment-i (i: 1 to 7)',
+                               'central_moment-i-j (2-0;1-1;0-2;3-0;2-1;1-2;0-3)',
+                               'gray_central_moment-i-j (2-0;1-1;0-2;3-0;2-1;1-2;0-3)'),
+                      comment = c('Hu moment', 'weighted Hu moment', 'central moment', 'weighted central moment'))
+
+all_features <- rbind(object_attributes, hist_attributes, stat_moment)
+colnames(all_features) <- c('Features name', 'Description')
+
+# gt(all_features) |> gt_highlight_rows(rows = 1:13,
+#                                       fill = 'lightgrey')
+#all_features |> flextable() |> color(i = 1:13, color = 'lightgrey')
+# kable(head(iris, 5), align = 'c', booktabs = TRUE) %>%
+#   row_spec(1, bold = TRUE, italic = TRUE) %>% 
+#   row_spec(2:3, color = 'white', background = 'black') %>%
+#   row_spec(4, underline = TRUE, monospace = TRUE) %>% 
+#   row_spec(5, angle = 45) %>% 
+#   column_spec(5, strikeout = TRUE)
+
+# kable(all_features) |> row_spec(1:14, background = '#F2F2F2') |> 
+#   row_spec(28:43, background = "#F2F2F2") |> row_spec(0, bold=T, align = 'c')
+
+kable(all_features) |> row_spec(0, bold=T, align = 'c')
+
+```
 
@@
 We chose to aggregate classes iteratively to find the best number of classes for the ER.
-The first level (ie.
+The first level (i.e.
 all labels (\> 40) are retained) served as a baseline for a prior model, without making any hypotheses.
@@
 parents could not be classified to a deeper taxonomic level, likely causing confusion with their children during training).
-The second level of aggregation was based on statistical metrics post training (see hereafter how a model is trained), meaning that we looked at the confusion matrix and the classification report (e.g.
-**Fig.
-XX and XX**, respectively) to see where classes did not perform well such that they had to be merged to improve the classification score.
+The second level of aggregation was based on statistical metrics post training (see hereafter how a model is trained), meaning that we looked at the confusion matrix and the classification report (@fig-confusion-matrix-uvp5 and @tbl-classification-report-uvp5-xgboost-cnn, respectively) to find which classes did not perform well such that they had to be merged to improve the classification score.
 This last step was repeated until an agreement was found.
@@
 
+<!-- ![Confusion matrix (normalized on true labels) of the best UVP5 XGBoost model.](image/uvp6/confusion_matrix_best_model_uvp5.jpg){#fig-confusion-matrix-uvp5-xgboost} -->
+
+```{r, confusion matrix CNN UVP5, fig.height = 6}
+#| label: fig-confusion-matrix-uvp5
+#| fig-cap: Confusion matrix (normalized on true labels) for the best UVP5 model trained with XGBoost and a CNN.
+
+library(viridis)
+library(gridExtra)
+
+# define x and y axes
+x <- c('Acantharia', 'Aulacanthidae', 'Aulosphaeridae', 'Copepoda', 'Trichodesmium', 
+            'artefact', 'bubble', 'darksphere', 'detritus', 'fiber', 'other<living')
+
+y <- factor(x, levels = rev(x))
+
+# create grid
+data <- expand.grid(X=x, Y=y, stringsAsFactors = T)
+
+# gray levels
+gray_cnn_uvp5 <- c(38,252,254,254,242,254,255,255,251,253,250,
+           249,61,252,255,241,253,255,255,247,255,251,
+           251,250,17,255,255,253,255,255,250,255,252,
+           255,255,255,61,255,255,255,255,214,253,252,
+           254,255,255,255,0,255,255,255,245,253,255,
+           255,255,255,255,255,92,255,255,187,254,254,
+           255,255,255,255,253,253,0,255,246,255,255,
+           251,248,253,255,242,254,252,109,222,255,252,
+           255,255,255,253,251,248,255,255,17,251,254,
+           255,255,255,255,248,253,255,255,189,108,254,
+           250,252,255,247,253,247,255,255,230,249,99)
+
+# reverse grayscale so that 0 becomes 255 and 255 becomes 0 and normalize to 1
+gray_cnn_uvp5 <- abs(gray_cnn_uvp5 - 255)/255
+
+# add gray values in the grid
+data$gray_level <- gray_cnn_uvp5
+
+# add group for facet plot
+data <- data |> mutate(group = 'CNN')
+
+right <- ggplot(data, aes(X, Y, fill= gray_level)) + geom_tile() + scale_fill_viridis(option = "F", direction = -1) +
+  labs(x = "Predicted labels", y = "True labels") + theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# do the same for the xgboost confusion matrix
+gray_xgboost_uvp5 <- c(66,252,249,255,250,255,255,255,233,249,250,
+                        253,62,248,255,248,244,255,255,244,255,249,
+                        254,252,4,255,255,255,255,255,243,255,251,
+                        255,255,255,126,255,253,255,255,147,253,252,
+                        254,255,255,255,35,255,255,253,226,248,255,
+                        255,255,255,255,255,35,255,255,213,255,255,
+                        255,255,255,255,253,250,17,255,235,255,255,
+                        254,253,253,255,251,254,255,65,215,255,252,
+                        255,255,255,253,253,246,255,255,0,250,253,
+                        255,255,255,254,250,252,255,255,154,120,255,
+                        250,252,253,246,252,241,255,253,221,245,135)
+
+gray_xgboost_uvp5 <- abs(gray_xgboost_uvp5 - 255)/255
+data2 <- expand.grid(X=x, Y=y, stringsAsFactors = T)
+data2 <- data2 |> mutate(gray_level = gray_xgboost_uvp5, group = 'XGBoost')
+
+left <- ggplot(data2, aes(X, Y, fill= gray_level)) + geom_tile() + scale_fill_viridis(option = "F", direction = -1) +
+  labs(x = "Predicted labels", y = "True labels") + theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# left / right
+
+# plot facet
+data <- rbind(data, data2)
+data <- data |> mutate(group = factor(group, levels = c('XGBoost', 'CNN'))) 
+ggplot(data, aes(X, Y, fill= gray_level)) + geom_tile() + scale_fill_viridis(option = "F", direction = -1) +
+  labs(x = "Predicted labels", y = "True labels", fill = '') + theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + facet_wrap(~group, nrow = 2)
+
+```
+
+```{r, classification report XGBoost UVP5}
+#| label: tbl-classification-report-uvp5-xgboost-cnn
+#| tbl-cap: Precision, recall and f1-score for each class and globally (lightgray background) in the UVP5 model (XGBoost and CNN). avg average. avg stands for average.
+
+labels <- c('Acantharea', 'Aulacanthidae', 'Aulosphaeridae', 'Copepoda', 'Trichodesmium', 
+            'artefact', 'bubble', 'darsphere', 'detritus', 'fiber', 'other<living', 'accuracy',
+            'macro avg', 'weighted avg', 'living macro avg', 'living weighted avg')
+
+precision <- c(0.68,0.63,0.80,0.54,0.68,0.6,0.77,0.57,0.90,0.50,0.36,0.81,0.64,0.82,0.61,0.60)
+recall <- c(0.68,0.69,0.84,0.50,0.75,0.75,0.81,0.68,0.86,0.52,0.48,0.81,0.69,0.81,0.66,0.65)
+f1_score <- c(0.68,0.66,0.82,0.52,0.71,0.67,0.79,0.62,0.88,0.51,0.41,0.81,0.66,0.82,0.63,0.62)
+
+precision_cnn <- c(0.60,0.61,0.89,0.56,0.56,0.63,0.76,.59,0.91,0.56,0.52,0.82,0.65,0.83,0.62,0.59)
+recall_cnn <- c(0.79,0.73,0.87,0.72,0.89,0.63,0.89,0.58,0.87,0.58,0.61,0.82,0.74,0.82,0.74,0.78)
+f1_score_cnn <- c(0.68,0.66,0.88,0.63,0.69,0.63,0.82,0.58,0.89,0.57,0.56,0.82,0.69,0.82,0.67,0.67)
+
+cf_uvp5_xgboost_cnn <- tibble(labels, precision, recall, f1_score, precision_cnn, recall_cnn, f1_score_cnn)
+colnames(cf_uvp5_xgboost_cnn) <- c('Labels', 'Precision', 'Recall', 'F1-score', 'Precision', 'Recall', 'F1-score')
+
+kable(cf_uvp5_xgboost_cnn) |> row_spec(0, bold=T, align = 'c') |> #row_spec(1:16, align = 'c') |>
+  row_spec(row = 12:16, background = '#F2F2F2') |> add_header_above(c(" " = 1, "XGBoost" = 3, "CNN" = 3))
+  #column_spec(2, background = spec_color(cf_uvp5_xgboost$precision))
+  #row_spec(12:16, background = '#F2F2F2')
+```
+
 In machine learning tasks, it is common to split the data into a training set, a validation set and a test set.
-The validation set is usually used to assess the quality of the model by tuning hyperparameters while avoiding overfitting (stopping the training before the model is too acquainted with the data).
+The validation set is usually used to assess the quality of a model by tuning hyperparameters while avoiding overfitting (stopping the training before the model is too acquainted with the data).
 Ideally, the test set is not to be used during any step except at the very end where it is used to assess the global quality of the final model.
-In our case, both datasets (UVP5 and UVP6) are very imbalanced towards few non-biological classes (mostly detritus).
+In our case, both data sets (UVP5 and UVP6) are very imbalanced towards few non-biological classes (mostly detritus).
 As a result, in order for the model to see as much biological data as possible during training, we did not use a validation set but instead we did a threefold cross-validation.
@@
 Hence, we built the test set with entire UVP5 profiles going at least to 1000 m depth, the typical parking depth of BGC-Argo floats.
-Those profiles were also chosen to be representative of the global ocean (**see Fig.
-XX**).
+Those profiles were also chosen to be representative of the global ocean (@fig-map-of-profiles).
 For UVP6 data, we did not select entire profiles because we were limited by the number of images in comparison with the UVP5 to build the biggest training set possible (80% of detritus in the UVP6 data set).
@@
 
-Following the previous steps, the UVP5 test and training sets are composed, respectively, of 315.277 and 3.023.506 images while the UVP6 has 59.054 and 531.483 images for its test and training sets.
+Following the previous steps, the UVP5 test and training sets are composed of, respectively, 315.277 and 3.023.506 images while the UVP6 has 59.054 and 531.483 images for its test and training sets.
 
@@
 XGBoost is designed to be an optimized and regularized version of GTB.
-To learn from the data and derive a model, XBoost first minimizes a regularized objective function $L$
+To learn from the data and derive a model, XGBoost first minimizes a regularized objective function $L$
 
@@
 where $l$ is a loss function that measures the difference between the prediction $\hat{y}_{i}$ and the true value $y_{i}$ and $\Omega$ is the regularization term, for each $k$ decision tree $f_{k}$.
-$\Omega$ is then splitted into two terms where $\gamma T$ penalizes the complexity of a tree proportionally to the number of leaves ($T$) and where the L2 norm penalizes the weights ($w$) of the leaves to avoid over-fitting (both $\gamma$ and $\lambda$ are hyperparameters that can be tuned by the user).
+$\Omega$ is then splitted into two terms where $\gamma T$ penalizes the complexity of a tree proportionally to the number of leaves ($T$) and where the L2 norm penalizes the weights ($w$) of the leaves to avoid overfitting (both $\gamma$ and $\lambda$ are hyperparameters that can be tuned by the user).
 By doing so, XGBoost encourages both a simple and predictive model.
@@
 
-To build our ER model, we first did a threefold cross-validation with XGBoost to find the optimal number of boosting rounds (i.e.
+To build the ER model, the first step was a threefold cross-validation using XGBoost parameterized with a set of given hyperparameters (i.e.
+learning rate ($\eta$), maximum depth of a tree ($D_{max}$) and weight sensitivity ($\omega$)).
+The second step consisted in the search of the optimal number of boosting rounds (i.e.
 $k$) before the model overfits.
-For this, we looked at the minimum of the multi-class log loss curve (**Fig.
-S3**), then we trained the ER model with the entire training set at this optimal number with specific hyperparameters (ie.
-learning rate ($\eta$), maximum depth of a tree ($D_{max}$) and weight sensitivity ($\omega$).
-The optimal couple ($\eta$, $D_{max}$) was determined using a customary grid search (**see Fig.
-3**) while $\omega$ was chosen empirically (see hereafter).
+For this, we searched for the minimum value of the multi class log loss(also known as logistic loss or cross-entropy loss) evaluated at each boosting round.
+Using that optimal $k$ value, we then trained the ER model with the entire training set.
+Therefore, finding the best model consisted in finding the couple ($\eta$, $D_{max}$) that respected the energy constraints and outputed the best global metrics.
+The latters were determined using a customary grid search (@fig-grid-search) while $\omega$ was chosen empirically (see hereafter).
 
-To cope with our unbalanced dataset (more than 70% of detritus in the UVP5 dataset), several options were available.
+```{r, figure for grid search, fig.width=10, fig.height=6}
+#| label: fig-grid-search
+#| fig-cap: Grid search results with several classes of detritus (1, 3 or 5 classes). D$_{max}$ is depicted by the colour of the border of each shape (red or black).
+
+library(tidyverse)
+library(ggplot2)
+library(stringr)
+library(gghighlight)
+
+# read all data
+d <- bind_rows(
+  read_csv('DATA/UVP6/results_grid_search_UVP5.csv') %>%
+    filter(str_detect(folder_name, "_1000")) %>%
+    mutate(learning_rate = learning_rate/10) %>%
+    filter(learning_rate < 0.5, max_depth < 7, n_detritus < 6)
+) %>%
+  # remove metrics with less meaning
+  #select(-ends_with("bio"), -mylogloss) %>%
+  # convert to tall format
+  gather(key="metric", val="val", accuracy:living_recall_w) %>%
+  # reformat hyperparameters
+  rename(n_det=n_detritus) %>%
+  mutate(
+    eta=factor(learning_rate),
+    Dmax=factor(max_depth),
+    n_det=factor(n_det),
+  )
+
+#https://stackoverflow.com/questions/64271350/changing-column-names-in-a-scatterplot-with-facet-wrap?noredirect=1&lq=1
+panel_names <- c(`accuracy`="global accuracy",
+                    `living_precision_w`="weighted living precision",
+                    `living_recall_w`="weighted living recall",
+                    `precision_macro`="macro precision",
+                    `recall_macro`="macro recall")
+
+d %>%
+  ggplot() +
+  geom_point(aes(x=loading_time, y=val,
+                 shape=eta, fill=eta, colour=Dmax, size=n_det)) +
+  facet_wrap(~metric, scales="free_y", labeller=as_labeller(panel_names), nrow = 3) +
+  scale_shape_manual(values=c(21:25), guide=guide_legend(expression(eta))) +
+  scale_fill_manual(values=c('#ffffcc','#a1dab4','#41b6c4','#225ea8'), guide=guide_legend(expression(eta))) + 
+  #scale_fill_manual(values=c('#ffffb2', '#fecc5c', '#fd8d3c', '#e31a1c'), guide=guide_legend(expression(eta))) + 
+  scale_size_manual(values=c(3:5), guide=guide_legend("detritus")) +
+  scale_colour_manual(values=c("red", "black"), guide=guide_legend(expression(D[max]))) +
+  xlab('Model loading time (ms)') +
+  ylab('Metrics') + theme_bw()
+```
+
+To cope with our unbalanced data set (more than 70% of detritus in the UVP5 data set), several options were available.
 We could artificially increase the number of biological data with the well known SMOTE oversampling technique [@Chawla2002-kg; @Fernandez2018-ft], decrease the number of detritus (random undersampling, keep only detritus close to unsupervised cluster centroids, etc.) or add a weight to each label during training to nudge the model towards under-represented classes (i.e.
 biological classes).
-In order to compare a features-based classifier with a state-of-the-art CNN, we chose the latter.
+In order to compare a features-based classifier with a state-of-the-art CNN, we chose the latter and we defined the weight of a class as follows:
 
@@
 In order to decipher between all options, we compared the predicted concentrations of classes for each $\omega$ with the real concentration of each class in different regions of the global ocean.
-As always in machine learning, we had a tradeoff between precision and recall but because the UVP6 is more focused on the estimation of carbon fluxes (high precision) than on the detection of rare objects (high recall with a low precision), we selected $\omega$ = 0.25 to be a satisfactory compromise.
+As always in machine learning, we encountered a trade off between precision and recall.
+However, given that the UVP6 is more focused on carbon flux estimation (high precision) than on the detection of rare objects (high recall with low precision), a weight sensitivity of 0.25 seemed to be the best compromise.
 
@@
 This is known as transfer learning and it was already applied with success for zooplankton classification in @Orenstein2017-ya.
-Input images have a size of \[224, 224, 3\], the size of the features extractor vector is 1792 followed by a fully connected layer (600 nodes) with a dropout of 0.4 to avoid overfitting.
+Input images have a size of \[224, 224, 3\], the size of the features extractor vector is 1792 followed by a fully connected layer (600 nodes) with a dropout of 0.4 to avoid over-fitting.
 The classification layer at the end of the network also has a dropout of 0.2.
-The CNN was trained with a decaying learning rate,  a Categorical Cross Entropy loss function and weights were defined with the ω = 0.25.
-The CNN was trained with a sufficient number of epochs such that it was not overfitting and it was not learning much anymore (validation plateau, see Fig.
-S5).
-
-![Scheme of the UVP6 pipeline with the processing done in the lab to create the ER model and the processing time (in ms) of the main steps when the UVP6 is deployed onboard an AUV.
-kpx means kilo pixel.](image/uvp6/fig_pipeline_UVP6.jpg){#fig-uvp-pipeline}
+The CNN was trained with a decaying learning rate, a categorical cross entropy loss function and weights were defined with the same weight sensitivity of 0.25.
+The CNN was trained with a sufficient number of epochs such that its training reached a validation plateau (i.e.
+the model was not learning much with additional epochs).
 
@@
 
-It is well known that the task of labelling images in supervised machine learning tasks is time consuming and most probably painful when you have more than 3 million images in your data set.
+It is well known that the task of labeling images in supervised machine learning tasks is time consuming and most probably painful when you have millions of images to label.
 It is also very frustrating to see that one class (largely) dominates when you are interested in the small ones and that some of them are so scarce that you need to merge them with their siblings (at the very best) on the phylogenetic tree.
@@
 
-Once the number of classes is known, hyperparameters can be tuned to find the optimal couple for the classification results (ie.
-grid search results shown in Fig.
-3).
-In our case, the optimal model hyperparameters were 0.2 for η and 5 for Dmax (also with only 1 class of detritus, see Discussion) to have both a good weighted living (or biological) precision and recall and a loading time (ie.
+Once the number of classes is known, hyperparameters can be tuned to find the optimal couple for the classification results (i.e.
+grid search results shown in @fig-grid-search).
+In our case, the optimal model hyperparameters were 0.2 for $\eta$ and 5 for D$_{max}$ (also with only 1 class of detritus, see Discussion) to have both a good weighted living (or biological) precision and recall and a loading time (i.e.
 loading of the ER model in memory) below 700 ms for energetic reasons.
-In that configuration, the model has an accuracy of 81%, a macro (ie.
+In that configuration, the UVP5 model had an accuracy of 81%, a macro (i.e.
 label imbalance not taken into account) precision and recall of, respectively, 64% and 69% and a weighted living precision and recall of, respectively, 60% and 65%.
-CV found an optimal number of 850 boosting rounds before overfitting and the expected loading time is 382 ms.
-
-INSERT FIGURE 3
-
-INSERT FIGURE 4
-
-INSERT FIGURE 5 -\> SWITCH FOR A TABLE !!
+CV found an optimal number of 850 boosting rounds before over-fitting and the expected loading time is 382 ms.
 
 The classification report (CR) gives the precision, the recall, the f1-score (harmonic mean of the precision and recall) and global metrics on the test set.
-As it can be seen in Fig.
-5, detritus have a recall of 86% thus because detritus are abundant (more than 230.000 in the test set), it results in more than 30.000 detritus wrongly labelled and therefore disseminated in all other classes.
+As it can be seen in @tbl-classification-report-uvp5-xgboost-cnn, detritus have a recall of 86% thus because detritus are abundant (more than 230.000 in the test set), it results in more than 30.000 detritus wrongly labelled and therefore disseminated in all other classes.
 The confusion matrix (CM) helps to see where those wrongly labelled images are.
-Fig.
-4 shows that a lot of Copepoda are confused with detritus as well as fiber.
+@fig-confusion-matrix-uvp5 shows that a lot of Copepoda are confused with detritus as well as fiber.
 The worst class in the CR is other\<living, which is not surprising because this class is composed of organisms that are not all morphologically and ecologically related.
@@
 
-In comparison with a CNN trained on UVP5 images, the CR (Fig.
-S2) shows that the global accuracy is 1% higher (82%) with a macro precision and recall of, respectively, 65% and 74% and a weighted living precision and recall of, respectively, 59% and 78%.
+In comparison with a CNN trained on UVP5 images, the CR (@tbl-classification-report-uvp5-xgboost-cnn) shows that the global accuracy is 1% higher (82%) with a macro precision and recall of, respectively, 65% and 74% and a weighted living precision and recall of, respectively, 59% and 78%.
 *Copepoda* are better detected (higher recall) but are still often (56% precision) misclassified.
@@
 For instance, the 14% increase in recall for *Trichodesmium* is followed by a 12% decrease in precision so there is still a lot of confusion at this stage for the UVP5 for some classes, even with a CNN that is not limited by any energy constraints.
+
+<!-- ![Confusion matrix (normalized on true labels) of the UVP5 CNN model.](image/uvp6/confusion_matrix_CNN_uvp5.jpeg){#fig-confusion-matrix-uvp5-cnn} -->
 
@@
 Quantification is an interesting concept that focuses on the sample instead of the individual.
-It was already shown in @Solow2001-jc, using plankton images from the Video Plankton Recorder, that the estimation of taxonomic abundances could be improved by computing the confusion matrix $\pmb{P}$ where $p_{i,j}$ is the probability that a taxon $i$ is classified as taxon \$j\$.
+It was already shown in @Solow2001-jc, using plankton images from the Video Plankton Recorder, that the estimation of taxonomic abundances could be improved by computing the confusion matrix $\pmb{P}$ where $p_{i,j}$ is the probability that a taxon $i$ is classified as taxon $j$.
 If the classifier is perfect, $p_{i,j} = 1$ where $i = j$, the quantification resumes to a simple Classify and Count method (CC, see @Forman2005-eu).
 However, a perfect (especially in plankton ecology) classifier does not exist.
-Therefore, quantification is a tool to better estimate the abundance of planktonic groups while perfectly knowing that the classifier is biassed.
+Therefore, quantification is a tool to better estimate the abundance of planktonic groups while perfectly knowing that the classifier is biased.
 This quantification method was successively applied in the paper of @Sosik2007-cj on a 2 month data set although it is very likely that P needs to be adapted to the sampling conditions.
@@
 Increasing the number of detritus classes was also considered in the hope of clustering them in the features space.
-As shown in Fig.
-SXX, detritus display a variety of size, shape and grey levels so it must be possible to cluster them and therefore improve the classification score.
+Detritus display a variety of size, shape and grey levels so it must be possible to cluster them and therefore improve the classification score.
 For this, we applied a classic K-Means algorithm [@Hartigan1979-oz] on detritus cleaned from likely outliers using an Isolation Forest Algorithm [@Liu2012-xz] and we executed a grid search as described earlier.
-Fig.
-3 clearly shows that using more than 1 class of detritus drastically decreases the global accuracy of the model, decreases the weighted living recall and only increases by 1% the weighted living precision.
-In addition, the confusion matrix (Fig.
-S4) shows that it does not improve the ER model because there is a high confusion between subgroups of detritus.
+@fig-grid-search clearly shows that using more than 1 class of detritus drastically decreases the global accuracy of the model, decreases the weighted living recall and only increases by 1% the weighted living precision.
+In addition, the confusion matrix (@fig-confusion-matrix-uvp6-ndetritus) shows that it does not improve the ER model because there is a high confusion between subgroups of detritus.
 Essentially, it seems that detritus form a continuum in the features space and that more research should be done on the choice of features, clustering algorithms and related methods to better differentiate them in separate entities, which is not the purpose of this study.
 
+<!-- ![Confusion matrix (normalized on true labels) of the a UVP5 XGBoost model with 5 classes of detritus.](image/uvp6/Muvpec_367314a7_confusion_matrix.jpeg){#fig-confusion-matrix-uvp6-ndetritus} -->
+```{r, confusion matrix N detritus}
+#| label: fig-confusion-matrix-uvp6-ndetritus
+#| fig-cap: Confusion matrix (normalized on true labels) of a UVP5 XGBoost model with 5 classes of detritus.
+
+x <- c('Acantharia', 'Aulacanthidae', 'Aulosphaeridae', 'Copepoda', 'Trichodesmium', 
+            'artefact', 'bubble', 'darksphere', 'detritus_0', 'detritus_1', 'detritus_2', 'detritus_3', 'detritus_4', 'fiber', 'other<living')
+
+y <- factor(x, levels = rev(x))
+
+data <- expand.grid(X=x, Y=y, stringsAsFactors = T)
+
+gray_levels_ndetritus <- c(64,252,249,255,250,255,255,255,255,240,253,247,254,249,251,
+                           253,55,247,255,247,246,255,255,250,253,255,252,251,255,249,
+                           254,252,0,255,255,255,255,255,250,250,255,250,254,255,250,
+                           255,255,255,137,254,254,255,255,247,231,253,229,230,255,251,
+                           254,255,255,255,33,255,255,253,254,251,255,238,246,249,255,
+                           255,255,255,255,255,51,255,255,252,245,253,219,255,255,255,
+                           255,255,255,255,253,250,11,255,253,251,255,247,245,255,255,
+                           254,252,253,255,250,254,254,76,253,254,255,252,214,255,251,
+                           255,255,255,255,255,181,255,255,253,192,247,239,255,226,251,
+                           255,255,255,255,254,193,255,255,248,189,249,212,255,246,251,
+                           255,255,255,254,255,212,255,255,253,194,221,242,255,229,251,
+                           255,255,255,255,255,181,255,255,250,161,251,229,255,247,252,
+                           255,255,255,253,253,249,255,255,246,239,255,173,159,252,255,
+                           255,255,255,254,250,253,255,255,253,229,249,213,243,138,255,
+                           251,252,253,246,252,242,254,253,247,246,253,247,249,245,134)
+
+gray_levels_ndetritus <- abs(gray_levels_ndetritus - 255)/255
+
+data$gray_xgboost_ndetritus <- gray_levels_ndetritus
+
+ggplot(data, aes(X, Y, fill= gray_xgboost_ndetritus)) + geom_tile() + scale_fill_viridis(option = "F", direction = -1) +
+  labs(x = "Predicted labels", y = "True labels", fill = '') + theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
 #### Application to the UVP6 data set
@@
 The aim of building an ER model was intended towards its implementation on the UVP6 for future deployments on BGC-Argo floats (or other profiling AUVs).
-Based on what was done with the UVP5 data set, we applied the same method for UVP6 data.
-Thus, we started with the data aggregation step (choices of classes) where, in contrast with UVP5 data, some classes were kept (see CR in Fig.
-7) because of their ecological role (e.g.
+Following the work done with the UVP5 data set, we applied the same method for UVP6 data.
+Thus, we started with the data aggregation step (choices of classes) where, in contrast with UVP5 data, some classes were kept (see @tbl-classification-report-uvp6-xgboost-cnn) because of their ecological role (e.g.
 small\<*Cnidaria* or tuff, an elongated cyanobacteria belonging to *Trichodesmium*) even though their classification metrics (precision and recall) were not particularly convincing.
@@
 Therefore, we assume that those classes are well recognized by human classifiers so that their parent *Rhizaria* is virtually free from them.
-Then, we executed a grid search to find the optimal couple $\eta$ and $D_{max}$ ($\omega$ and the number of classes of detritus were fixed based on UVP5 results) and trained the final ER model whose CM and CR are reported hereafter for further interpretation.
+Afterwards, we executed a grid search to find the optimal couple $\eta$ and $D_{max}$ ($\omega$ and the number of classes of detritus were fixed based on UVP5 results) and trained the final ER model whose CM and CR are reported in @fig-confusion-matrix-final-model and @tbl-classification-report-uvp6-xgboost-cnn.
 
-The specifications of the final UVP6 ER model are the following: \$\\eta\$: 0.2, $D_{max}$: 3, $\omega$: 0.25, number of boosting rounds: 400 for a total loading time of 73 ms.
-Note that the CV initially found an optimal number of boosting rounds of 555 (loading time: 101 ms) but the log loss curve had already reached a plateau so we were able to decrease the size of the model without losing much (\< 1%) in biological precision or recall, allowing to decrease a bit more the energy consumption of the UVP6 on board a BGC-Argo float for an extended battery life.
-For the interested reader, results from a CNN trained on UVP6 data are available in Fig.
-S6 and S7.
+```{r, confusion matrix N detritus}
+#| label: fig-confusion-matrix-final-model
+#| fig-cap: Confusion matrix (normalized on true labels) of the final UVP6 model trained with XGBoost.
 
-INSERT FIGURE 6
+library(RcppCNPy)
+gray_levels_uvp6 <- array(t(npyLoad('DATA/UVP6/cm_uvp6.npy')))
 
-INSERT FIGURE 7
+x <- c('Acantharia', 'Actinopterygii', 'Appendicularia', 'Aulacanthidae',
+       'Calanoida', 'Chaetognatha', 'Collodaria', 'Creseis',
+       'Foraminifera', 'Rhizaria', 'Salpida', 'artefact', 'crystal',
+       'detritus', 'fiber', 'other<living', 'puff', 'small<Cnidaria',
+       'solitaryglobule', 'tuff')
 
-#### How to use the CR?
+y <- factor(x, levels = rev(x))
+data <- expand.grid(X=x, Y=y, stringsAsFactors = T)
+data$gray_levels <- gray_levels_uvp6
+
+ggplot(data, aes(X, Y, fill= gray_levels)) + geom_tile() + scale_fill_viridis(option = "F", direction = -1) +
+  labs(x = "Predicted labels", y = "True labels", fill = '') + theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+The specifications of the final UVP6 ER model are the following: $\eta$: 0.2, $D_{max}$: 3, $\omega$: 0.25, number of boosting rounds: 400 for a total loading time of 73 ms.
+Note that the CV initially found an optimal number of boosting rounds of 555 for a loading time of 101 ms).
+However, we observed that the log loss curve had already reached a plateau hence we were able to decrease the size of the model without losing much (\< 1%) in biological precision or recall, decreasing a bit more the energy consumption of the UVP6 on board AUVs.
+
+```{r, classification report XGBoost and CNN uvp6}
+#| label: tbl-classification-report-uvp6-xgboost-cnn
+#| tbl-cap: Precision, recall and f1-score for each class and globally (lightgray background) of the best UVP5 model trained with XGBoost and a CNN. The abbreviation avg stands for average.
+
+labels <- c('Acantharia', 'Actinopterygii', 'Appendicularia', 'Aulacanthidae', 'Calanoida', 'Chaetognatha', 'Collodaria', 'Creseis', 'Foraminifera', 'Rhizaria', 'Salpida', 'artefact', 'crystal', 'detritus', 'fiber','other<living', 'puff', 'small<Cnidaria', 'solitaryglobule', 'tuff', 'accuracy','macro avg', 'weighted avg', 'living macro avg', 'living weighted avg')
+
+precision <- c(0.89,1,0.89,0.82,0.66,0.80,.73,0.81,0.71,0.69,0.63,0.91,0.89,0.95,0.73,0.44,0.73,0.55,0.83,0.55,0.91,0.76,0.91,0.73,0.61)
+recall <- c(0.69,0.40,0.44,0.89,0.69,0.62,0.62,1,0.48,0.75,0.61,0.92,0.89,0.95,0.83,0.24,0.78,0.47,0.71,0.63,0.91,0.68,0.91,0.63,0.56)
+f1_score <- c(0.78,0.57,0.59,0.85,0.68,0.70,0.67,0.89,0.57,0.72,0.62,0.91,0.89,0.95,0.78,0.31,0.75,0.51,0.77,0.59,0.91,0.71,0.91,0.66,0.57)
+
+precision_cnn <- c(0.98,0.47,0.38,0.86,0.82,0.52,0.74,0.42,0.76,0.71,0.74,0.94,0.69,0.98,0.77,0.55,0.84,0.72,0.81,0.68,0.94,0.72,0.94,0.69,0.71)
+recall_cnn <- c(1,0.73,0.64,0.86,0.89,0.88,0.88,1,0.89,0.92,0.89,0.92,0.97,0.96,0.85,0.62,0.94,0.86,0.88,0.93,0.94,0.88,0.94,0.86,0.82)
+f1_score_cnn <- c(0.99,0.57,0.47,0.86,0.86,0.65,0.81,0.59,0.82,0.80,0.81,0.93,0.81,0.97,0.80,0.59,0.89,0.79,0.84,0.79,0.94,0.78,0.94,0.76,0.76)
+
+cf_uvp6_xgboost_cnn <- tibble(labels, precision, recall, f1_score, precision_cnn, recall_cnn, f1_score_cnn)
+colnames(cf_uvp6_xgboost_cnn) <- c('Labels', 'Precision', 'Recall', 'F1-score', 'Precision', 'Recall', 'F1-score')
+
+kable(cf_uvp6_xgboost_cnn) |> row_spec(0, bold=T, align = 'c') |> #row_spec(1:16, align = 'c') |>
+  row_spec(row = 21:25, background = '#F2F2F2') |> add_header_above(c(" " = 1, "XGBoost" = 3, "CNN" = 3))
+```
+
+#### How to use the CR in real applications?
 
 4 BGC-Argo floats equipped with the ER were deployed in late May 2022 in the southern part of the Labrador Sea at the time of the spring bloom.
-As it can be seen in Fig.
-8, *Calanoida* are more abundant around the deep chlorophyll maximum (DCM) and they would be observed in very low concentrations even below 1000 m.
+As it can be seen in @fig-bgc-application-uvp6, *Calanoida* are more abundant around the deep chlorophyll maximum (DCM) and they would be observed in very low concentrations even below 1000 m.
 The CR for *Calanoida* shows that the model has a precision of 66% and a recall of 69% for that class.
 It implies that caution must be applied when interpreting the presence or absence of such copepods.
-In that case, we theoretically miss 31% of them (ie.
+In that case, we theoretically miss 31% of them (i.e.
 31% of *Calanoida* are not detected as such, they are classified in another class) and out of the 69% that we detect as *Calanoida*, we are wrong 34% of the time.
@@
 
-INSERT FIGURE 8
+```{r, uvp6 BGC application}
+#| label: fig-bgc-application-uvp6
+#| fig-cap : Concentrations of 4 classes (3 biological and detritus) estimated with data acquired from the UVP6 on board the BGC-Argo float (WMO 4903634) on the 15$^{th}$ of June 2022 in the Labrador Sea at latitude 59.27° N and longitude -50.56° E.
+Red points represent positive concentrations.
+Dotted lines represent the depth of the deep chlorophyll maximum.
+
+library(ggplot2)
+library(tidyverse)
+library(gridExtra)
+library(cowplot)
+library(patchwork)
+
+# extract bgc-argo data
+#data <- read.csv('bgc_data/lovuse017c_010_01.csv', sep = '\t')
+load("DATA/UVP6/lovuse017c_015_01.RData")
+
+# for the discussion of the uvp paper
+data <- dataprofile$data$uvp6_txo
+ecodata <- dataprofile$data$eco
+
+# plot chlorophyll for DCM
+ecodata <- ecodata %>% select(pres = Pressure_dbar, chla = `chlorophyll-a_ug/l`)
+tmp <- ecodata %>% group_by(pres) %>% summarize(n_sample = n(), total_conc = sum(chla), mean_conc = total_conc/n_sample) %>% filter(mean_conc > 0)
+chla <- ggplot(tmp, aes(x = pres, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip() +
+  xlab('Pressure (dbar)') + ylab(expression('Chlorophyll a (mu/L)')) + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) 
+DCM <- tmp$pres[which.max(tmp$mean_conc)]
+
+# object number 5 (Calanoida)
+test <- data %>% select(Pressure_dbar, Nimages, ObjectNbr5)
+test$ObjectConc5 <- test$ObjectNbr5/(test$Nimages*0.6)
+test <- test %>% group_by(Pressure_dbar) %>% summarize(n_sample = n(), total_conc = sum(ObjectConc5), mean_conc = total_conc/n_sample)
+cal <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip()+
+  xlab('Pressure (dbar)') + ylab('Calanoida (#/L)') + geom_point(data = test[test$mean_conc>0,], 
+                                                                  aes(x = Pressure_dbar, y = mean_conc), colour = "red") +
+  geom_vline(xintercept = DCM, lty = 'dotted') + theme_bw()
+
+# object number 14 (detritus)
+test <- data %>% select(Pressure_dbar, Nimages, ObjectNbr14)
+test$ObjectConc14 <- test$ObjectNbr14/(test$Nimages*0.6)
+test <- test %>% group_by(Pressure_dbar) %>% summarize(n_sample = n(), total_conc = sum(ObjectConc14), mean_conc = total_conc/n_sample)
+det <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip() +
+  geom_point(data = test[test$mean_conc>0,], 
+             aes(x = Pressure_dbar, y = mean_conc), colour = "red") +
+  xlab('Pressure (dbar)') + ylab('detritus (#/L)') + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
+  geom_vline(xintercept = DCM, lty = 'dotted') + theme_bw()
+
+test <- data %>% select(Pressure_dbar, Nimages, ObjectNbr17)
+test$ObjectConc17 <- test$ObjectNbr17/(test$Nimages*0.6)
+test <- test %>% group_by(Pressure_dbar) %>% summarize(n_sample = n(), total_conc = sum(ObjectConc17), mean_conc = total_conc/n_sample)
+puf <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip()+
+  xlab('Pressure (dbar)') + ylab('puff (#/L)') + geom_point(data = test[test$mean_conc>0,], 
+                                                                          aes(x = Pressure_dbar, y = mean_conc), colour = "red") +
+  geom_vline(xintercept = DCM, lty = 'dotted') + theme_bw()
+
+test <- data %>% select(Pressure_dbar, Nimages, ObjectNbr20)
+test$ObjectConc20 <- test$ObjectNbr20/(test$Nimages*0.6)
+test <- test %>% group_by(Pressure_dbar) %>% summarize(n_sample = n(), total_conc = sum(ObjectConc20), mean_conc = total_conc/n_sample)
+tuf <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip()+
+  xlab('Pressure (dbar)') + ylab('tuff (#/L)') + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
+  geom_point(data = test[test$mean_conc>0,], 
+             aes(x = Pressure_dbar, y = mean_conc), colour = "red") +
+  geom_vline(xintercept = DCM, lty = 'dotted') + theme_bw()
+
+#grid.arrange(cal, det, puf, tuf)
+plot_grid(cal, det, puf, tuf, labels = "", nrow = 2, ncol = 2)
+
+```
 
@@
 The ER of zooplankton on board AUVs is just at its beginning.
-As expected, the CNN trained on UVP6 data (Fig.
-S7) outperforms the results of our features-based classifier.
+As expected, the CNN trained on UVP6 data (see @tbl-classification-report-uvp6-xgboost-cnn) outperforms the results of our features-based classifier.
 However, it should be noted that not all biological classes are better classified (e.g.
@@
 This is likely explained by the fact that we were able to keep more (small) classes for the UVP6.
-Nevertheless, previous tests on UVP5 data with more classes increased the gap between XGBoost and CNN weighted results therefore the latter recognises better small classes.
-If energy and hardware constraints limit the performances of our current ER model, it is not forbidden to think about light-weight CNNs for future applications on AUVs to better understand the dynamics and the behaviour of zooplankton.
+Nevertheless, previous tests on UVP5 data with more classes increased the gap between XGBoost and CNN weighted results therefore the latter recognizes better small classes.
+If energy and hardware constraints limit the performances of our current ER model, it is not forbidden to think about light-weight CNNs for future applications on AUVs to better understand the dynamics and the behavior of zooplankton.
 
@@
 Finally, we showed the CM and CR of the final model and how to use it to interpret some UVP6 taxonomic profiles observed by the float WMO 4903634 during the spring bloom in the southern part of the Labrador Sea.
-
-### Supplementary material

### 2023-03-06 article3.qmd
update

@@
 ```{r, load libraries}
+library(tidyverse)
 library(akima)
 library(broom)
-library(tidyverse)
 library(cmocean)
@@
 library(gt)
-library(gtExtras)
+#library(gtExtras)
 library(spatialrisk)
+library(ggOceanMaps)
 ```
@@
 # add levels for some field
-all_floats <- all_floats |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
+all_floats <- all_floats |> dplyr::mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
                                    park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
@@
 # summarize localization of floats
-argo_map <- all_floats |> select(WMO = wmo, cycle, lat, lon) |> drop_na(lat) |> dplyr::group_by(WMO, cycle) |> mutate(WMO = factor(WMO)) |>
+argo_map <- all_floats |> select(WMO = wmo, cycle, lat, lon) |> drop_na(lat) |> dplyr::group_by(WMO, cycle) |> dplyr::mutate(WMO = factor(WMO)) |>
   dplyr::summarise(lat = mean(lat), lon = mean(lon))
@@
 
-# map positions of floats
-world_data <- map_data("world")
-w <- list(geom_polygon(aes(long, lat, group=group), data=world_data, fill="grey70"), coord_quickmap())
-
-ggplot(argo_map) + w + geom_path(aes(lon, lat, colour = WMO), size = 1) +
-  geom_point(aes(lon, lat, colour = WMO), size = 1.5) +
-  scale_color_brewer(palette = 'Dark2') +
-  theme_bw() +
-  #scale_colour_manual(values = c('#a6cee3', '#33a02c', '#b2df8a', '#1f78b4')) +
-  coord_cartesian(xlim = c(-55, -40), ylim = c(55,65)) + geom_point(data = argo_map_first, aes(lon, lat), shape = 17, size = 3) +
-  geom_point(data = argo_map_last, aes(lon, lat, colour = WMO), size = 3) +
-  geom_point(data = argo_map_last, aes(lon, lat, fill = WMO), size = 3, shape = 1) +
-  labs(x = 'longitude', y = 'latitude') +
+# # map positions of floats
+# world_data <- map_data("world")
+# w <- list(geom_polygon(aes(long, lat, group=group), data=world_data, fill="grey70"), coord_quickmap())
+# 
+# ggplot(argo_map) + w + geom_path(aes(lon, lat, colour = WMO), size = 1) +
+#   geom_point(aes(lon, lat, colour = WMO), size = 1.5) +
+#   scale_color_brewer(palette = 'Dark2') +
+#   theme_bw() +
+#   #scale_colour_manual(values = c('#a6cee3', '#33a02c', '#b2df8a', '#1f78b4')) +
+#   coord_cartesian(xlim = c(-55, -40), ylim = c(55,65)) + geom_point(data = argo_map_first, aes(lon, lat), shape = 17, size = 3) +
+#   geom_point(data = argo_map_last, aes(lon, lat, colour = WMO), size = 3) +
+#   geom_point(data = argo_map_last, aes(lon, lat, fill = WMO), size = 3, shape = 1) +
+#   labs(x = 'longitude', y = 'latitude') +
+#   #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
+#   guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
+#                                                   shape = c(NA, NA, NA, NA))))
+
+# prepare data for geomapping, see https://mikkovihtakari.github.io/ggOceanMaps/
+argo_map_geo <- transform_coord(argo_map) |> dplyr::mutate(cycle = argo_map$cycle, WMO = as.factor(argo_map$WMO))
+argo_map_last_geo <- transform_coord(argo_map_last) |> mutate(WMO = as.factor(argo_map_last$WMO))
+
+basemap(limits = c(-55,-40,55,65), bathymetry = T) +
+  geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), size = 1.5) + 
+  geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 3) +
+  scale_color_brewer(palette = 'Dark2') + theme_bw() +
+  geom_point(data = transform_coord(argo_map_first), aes(lon, lat), shape = 17, size = 3) +
+  geom_point(data = argo_map_last_geo, aes(lon, lat, colour = WMO), size = 6) +
+  geom_point(data = argo_map_last_geo, aes(lon, lat), fill = 'black', size = 6, shape = 1) +
   #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
@@
 Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window.
-Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the relationship of @Estapa2022
+Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).
 
@@
 clean_cp_data <- function(data, moving_median_order){
+
+  data <- data |> arrange(dates)
 
@@
   tmp3 <- tmp3[-c(index_neg_jump, index_neg_jump+1),]
-  tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp))
+  tmp3 <- tmp3 |> dplyr::mutate(diff_jump = cp - lag(cp))
   even_indexes <- seq(2,nrow(tmp3),2)
@@
 }else if(length(check_colour) == 2){
-  tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp))
+  tmp3 <- tmp3 |> dplyr::mutate(diff_jump = cp - lag(cp))
   even_indexes <- seq(2,nrow(tmp3),2)
@@
   # compute a Z score (assuming a normal distribution of the slopes) on the slopes
-  tmp <- tmp |> mutate(zscore = (slope - mean(slope, na.rm = T))/sd(slope, na.rm = T))
+  tmp <- tmp |> dplyr::mutate(zscore = (slope - mean(slope, na.rm = T))/sd(slope, na.rm = T))
 
@@
   #   tmp2 <- tmp2[-c(index_neg_jump, index_neg_jump+1),]
-  #   tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
+  #   tmp2 <- tmp2 |> dplyr::mutate(diff_jump = cp - lag(cp)) 
   #   even_indexes <- seq(2,nrow(tmp2),2)
@@
   # }else if(length(check_colour) == 2){ # only positive jumps
-  #   tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
+  #   tmp2 <- tmp2 |> dplyr::mutate(diff_jump = cp - lag(cp)) 
   #   even_indexes <- seq(2,nrow(tmp2),2)
@@
   if(length(check_colour) >= 2){ # there is a least a jump (positive or negative)
-    tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
+    tmp2 <- tmp2 |> dplyr::mutate(diff_jump = cp - lag(cp)) 
     even_indexes <- seq(2,nrow(tmp2),2)
@@
   # to plot subgroups
-  part_slope_tmp <- part_slope |> mutate(dates = time, colour = 'slope')
+  part_slope_tmp <- part_slope |> dplyr::mutate(dates = time, colour = 'slope')
 
@@
   return(df)
-
-}
+  #return(jump_plot)
+
+}
+
+# tmp <- all_floats_slope |> filter(wmo == 6904240, park_depth == '1000 m', cycle == 21)
+# plot_jumps(tmp)
 ```
@@
 all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle) |>
-   mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
+   dplyr::mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
           park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
@@
 # keep the good fluxes
-cflux <- res |> mutate(drifting_time = difftime(max_time, min_time, units = 'days'), WMO = factor(wmo))
+cflux <- res |> dplyr::mutate(drifting_time = difftime(max_time, min_time, units = 'days'), WMO = factor(wmo))
 
@@
 
-uvp_data <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date, cycle, wmo, park_depth, all_of(lpm_class)) |>
-  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
-
-uvp_data <- uvp_data |> pivot_longer(cols = lpm_class, names_to = 'size', values_to = 'conc') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+uvp_data <- all_floats |> filter(PhaseName == 'PAR') |> dplyr::mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date, cycle, wmo, park_depth, all_of(lpm_class)) |>
+  drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld), WMO = factor(wmo))
+
+uvp_data <- uvp_data |> pivot_longer(cols = lpm_class, names_to = 'size', values_to = 'conc') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
 
 # compute daily mean chla for each float at each cycle and at each drifting depth
-mean_uvp_data <- uvp_data |> group_by(cycle, WMO, size, park_depth, juld) |> summarize(mean_conc = mean(conc, na.rm=T))
+mean_uvp_data <- uvp_data |> dplyr::group_by(cycle, WMO, size, park_depth, juld) |> dplyr::summarize(mean_conc = mean(conc, na.rm=T))
 ```
@@
 
-test <- tmp200m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> summarize(index_max = which.max(y), max_date = x[index_max]) |> mutate(date_max = as_date(max_date, origin = "1970-01-01"))
+test <- tmp200m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> dplyr::summarize(index_max = which.max(y), max_date = x[index_max]) |> dplyr::mutate(date_max = as_date(max_date, origin = "1970-01-01"))
 test$size <- unique(uvp_data$size)
@@
 
-test <- tmp500m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> summarize(index_max = which.max(y), max_date = x[index_max]) |> mutate(date_max = as_date(max_date, origin = "1970-01-01"))
+test <- tmp500m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> dplyr::summarize(index_max = which.max(y), max_date = x[index_max]) |> dplyr::mutate(date_max = as_date(max_date, origin = "1970-01-01"))
 test$size <- unique(uvp_data$size)
@@
 
-test <- tmp1000m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> summarize(index_max = which.max(y), max_date = x[index_max]) |> mutate(date_max = as_date(max_date, origin = "1970-01-01"))
+test <- tmp1000m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> dplyr::summarize(index_max = which.max(y), max_date = x[index_max]) |> dplyr::mutate(date_max = as_date(max_date, origin = "1970-01-01"))
 test$size <- unique(uvp_data$size)
@@
 parking_data <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
-parking_data <- parking_data |> filter(PhaseName == 'PAR') |> drop_na(NP_Size_102) |> select(park_depth, all_of(lpm_class), dates = juld, park_depth, wmo, cycle) |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
+parking_data <- parking_data |> filter(PhaseName == 'PAR') |> drop_na(NP_Size_102) |> select(park_depth, all_of(lpm_class), dates = juld, park_depth, wmo, cycle) |> dplyr::mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
                                    park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
@@
 
-parking_data <- parking_data |> mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) #|> group_by(park_depth, cycle, wmo)
+parking_data <- parking_data |> dplyr::mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) #|> group_by(park_depth, cycle, wmo)
 
@@
 parking_spectra_rainer <- parking_data |> dplyr::select(lpm_class)
-parking_data <- parking_data |> mutate(rainer_flux = compute_flux_rainer(2.8649, 2.24, parking_spectra_rainer)) 
-parking_data <- parking_data |> mutate(iversen_flux = compute_flux(273, 4.27, parking_spectra))
+parking_data <- parking_data |> dplyr::mutate(rainer_flux = compute_flux_rainer(2.8649, 2.24, parking_spectra_rainer)) 
+parking_data <- parking_data |> dplyr::mutate(iversen_flux = compute_flux(273, 4.27, parking_spectra))
 
 # plot it
-comparison_flux_ost_uvp <- merge(parking_data, cflux) |> mutate(total_flux = small_flux + big_flux)
+comparison_flux_ost_uvp <- merge(parking_data, cflux) |> dplyr::mutate(total_flux = small_flux + big_flux)
+
+# add guidi flux obtained with fminsearch
+# mean_d <- c(0.29042685, 0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779, 1.16170742, 1.46365963)
+# comparison_flux_ost_uvp <-  comparison_flux_ost_uvp[-c(bad_index),]
+# ost_flux <- comparison_flux_ost_uvp$total_flux
+# spectra <- comparison_flux_ost_uvp |> select(NP_Size_256:NP_Size_1290)
+# # some spectra gives 0 for all classes - makes a buggy fminsearch -> remove those lines
+# bad_index <- which((rowSums(t(t(spectra)))) == 0)
+# spectra <- spectra[-c(bad_index),]
+# compute_flux <- function(x) sum((log10(rowSums(t(t(spectra)*x[1]*mean_d ** x[2]))) - log10(ost_flux))**2)
+# x0 <- c(12.5, 2.2)
+# fminsearch(compute_flux, x0 = x0)
+# fmincon(x0 = x0, compute_flux, lb = c(1, 1.5), ub = c(1000, 3))
+
+# add guidi flux using fminsearch using the optimize function of scipy in python (see a paper of someoe who did it, see Giering 2020 ou willimaoson et giering 2022)
+comparison_flux_ost_uvp <- comparison_flux_ost_uvp |> dplyr::mutate(guidi_flux_fminsearch = compute_flux(21.98, 1.18, parking_spectra))
 
@@
                       fun.max = function(z) {quantile(z, 0.75)},
-                      fun = median, colour = 'red', width = 5) +
+                      fun = median, colour = '#E31B23', width = 5) +
     geom_line(aes(x = date, y = guidi_flux),
                       stat = 'summary',
-                      fun = median, colour = 'red') + scale_y_continuous(trans = 'log10') + theme_bw()
-
-```
+                      fun = median, colour = '#E31B23') +
+    geom_errorbar(aes(x = date, y = iversen_flux),
+                      stat = 'summary',
+                      fun.min = function(z) {quantile(z, 0.25)},
+                      fun.max = function(z) {quantile(z, 0.75)},
+                      fun = median, colour = '#003366', width = 5) +
+    geom_line(aes(x = date, y = iversen_flux),
+                      stat = 'summary',
+                      fun = median, colour = '#003366') + 
+  scale_y_continuous(trans = 'log10') + theme_bw() + labs(x = 'Month', y = TeX('$F_{total}$ (mg C m$^{-2}$ day$^{-1}$)'))
+
+```
+
+ici je compare finalement un système ultra productif (iversen) et un système pas produtif (lionel) -> on se situerait entre les deux.
+Après je peux fournir els valeurs que j'ai trouvées avec le fminsearch avec la fonctiono optimize depython (21.98 et 1.18.
+Mais bon, cela assumerait que 'lon pourrait prendre l'OST comme référence, ce qui est un peu too much..)
 
@@
 #| tbl-cap: TODO
-correlations <- comparison_flux_ost_uvp |> select(park_depth, wmo, cycle, guidi_flux, small_flux, big_flux, date, total_flux) |> group_by(wmo, park_depth, cycle) |> mutate(median_guidi_flux = median(guidi_flux)) |> group_by(park_depth, wmo) |> summarise(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_big = round(cor(median_guidi_flux, big_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))
+correlations <- comparison_flux_ost_uvp |> dplyr::select(park_depth, wmo, cycle, guidi_flux, small_flux, big_flux, date, total_flux) |> dplyr::group_by(wmo, park_depth, cycle) |> dplyr::mutate(median_guidi_flux = median(guidi_flux)) |> dplyr::group_by(park_depth, wmo) |> dplyr::summarize(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_big = round(cor(median_guidi_flux, big_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))
 
@@
   # build dataframe to plot each subgroup
-  part1 <- slope_df |> select(group, time = min_time, cp = first_cp)
+  part1 <- slope_df |> dplyr::select(group, time = min_time, cp = first_cp)
   part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
@@
   # clean data to send
-  slope_df <- slope_df |> arrange(min_time) |> mutate(min_drifting_time = df$min_time, max_drifting_time = df$max_time, small_flux = df$small_flux, big_flux = df$big_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1])
+  slope_df <- slope_df |> arrange(min_time) |> dplyr::mutate(min_drifting_time = df$min_time, max_drifting_time = df$max_time, small_flux = df$small_flux, big_flux = df$big_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1])
   if(is.null(tmp3)){
@@
   }else{
-     big_jump <- tmp3 |> mutate(park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1]) 
+     big_jump <- tmp3 |> dplyr::mutate(park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1]) 
   }
@@
 uvp_drifting <- all_floats |> filter(PhaseName == 'PAR') |> select(juld, cycle, wmo, park_depth, all_of(lpm_class)) |> 
-  drop_na(NP_Size_102) |> mutate(date = as_date(juld))
+  drop_na(NP_Size_102) |> dplyr::mutate(date = as_date(juld))
 
@@
 biovolumes_names <- colnames(spectra_biovolume)
-uvp_drifting <- cbind(uvp_drifting, spectra_biovolume) |> rowwise() |> mutate(total_biovolume = sum(across(all_of(biovolumes_names))))
+uvp_drifting <- cbind(uvp_drifting, spectra_biovolume) |> rowwise() |> dplyr::mutate(total_biovolume = sum(across(all_of(biovolumes_names))))
 
@@
 
-test <- data.frame()
-for (i in wmo){
-  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
-  for (j in park_depth){
-    for (k in seq(1:max_cycle)){
-      tmp_uvp <- filter(uvp_drifting, wmo == i, cycle == k, park_depth == j) |> arrange(juld) |> select(juld, cycle, wmo, park_depth, total_biovolume) #, biovolumes_names)
-      tmp_ost <- res_slope |> filter(wmo == i, cycle == k, park_depth == j)
-      tmp <- assign_time_group(tmp_uvp, tmp_ost)
-      #tmp2 <- tmp |> filter(!is.na(group)) |> group_by(group) |> mutate(uvp_points = n(), min_date = min(juld), max_date = max(juld), mean_date = mean(floor_date(juld))) # https://stackoverflow.com/questions/44814435/calculate-average-daily-value-from-large-data-set-with-r-standard-format-date-ti
-      tmp2 <- tmp |> filter(!is.na(group)) |> group_by(group) |> mutate(uvp_points = n(), mean_date = mean(floor_date(juld))) 
-      tmp3 <- merge(tmp2, tmp_ost) |> group_by(group, cycle, wmo, park_depth) |> summarize(mean_date = unique(mean_date), uvp_points = unique(uvp_points), ost_points = unique(nb_points), group_total_biovolume = sum(total_biovolume), slope = unique(slope), weighted_group_biovolume = group_total_biovolume/uvp_points)
-      test <- rbind(test, tmp3)
-    }
-  }
-}
-
-ggplot(test) + geom_line(aes(x = mean_date, y = weighted_group_biovolume)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
-ggplot(test) + geom_line(aes(x = mean_date, y = slope)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4)
+# test <- data.frame()
+# for (i in wmo){
+#   max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
+#   for (j in park_depth){
+#     for (k in seq(1:max_cycle)){
+#       tmp_uvp <- filter(uvp_drifting, wmo == i, cycle == k, park_depth == j) |> arrange(juld) |> select(juld, cycle, wmo, park_depth, total_biovolume) 
+#       tmp_ost <- res_slope |> filter(wmo == i, cycle == k, park_depth == j)
+#       if(nrow(tmp_uvp) == 0 | nrow(tmp_ost) == 0){
+#         next
+#       }else{
+#       tmp <- assign_time_group(tmp_uvp, tmp_ost)
+#       #tmp2 <- tmp |> filter(!is.na(group)) |> dplyr::group_by(group) |> dplyr::mutate(uvp_points = n(), min_date = min(juld), max_date = max(juld), mean_date = mean(floor_date(juld))) # https://stackoverflow.com/questions/44814435/calculate-average-daily-value-from-large-data-set-with-r-standard-format-date-ti
+#       tmp2 <- tmp |> filter(!is.na(group)) |> group_by(group) |> dplyr::mutate(uvp_points = n(), mean_date = mean(floor_date(juld)))
+#       tmp3 <- merge(tmp2, tmp_ost) |> dplyr::group_by(group, cycle, wmo, park_depth) |> dplyr::summarize(mean_date = unique(mean_date), uvp_points = unique(uvp_points), ost_points = unique(nb_points), group_total_biovolume = sum(total_biovolume), slope = unique(slope), weighted_group_biovolume = group_total_biovolume/uvp_points)
+#       test <- rbind(test, tmp3)
+#       }
+#     }
+#   }
+# }
+
+# write data because it's long
+# vroom_write(test, file = '/home/flo/dox/THESIS/DATA/ARGO/slope_vs_biovolume.csv')
+
+# load data computed with the above commented code
+slope_vs_biovolume <- vroom::vroom('DATA/ARGO/slope_vs_biovolume.csv')
+slope_vs_biovolume <- slope_vs_biovolume |> mutate(park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
+
+# ggplot(test) + geom_line(aes(x = mean_date, y = weighted_group_biovolume)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
+# ggplot(test) + geom_line(aes(x = mean_date, y = slope)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4)
 
 # check
-test2 <- test |> group_by(wmo, cycle) |> mutate(mean_mean_date = mean(floor_date(mean_date)))
-
-ggplot(test2) +
-  geom_pointrange(aes(x = mean_mean_date, y = weighted_group_biovolume),
-                      stat = 'summary',
-                      fun.min = function(z) {quantile(z, 0.25)},
-                      fun.max = function(z) {quantile(z, 0.75)},
-                      fun = median, colour = 'black', size = .1) +
-    facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
-
-ggplot(test2) +
-  geom_pointrange(aes(x = mean_mean_date, y = slope),
-                      stat = 'summary',
-                      fun.min = function(z) {quantile(z, 0.25)},
-                      fun.max = function(z) {quantile(z, 0.75)},
-                      fun = median, colour = 'black', size = .1) +
-    facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4)
-
-test2 |> group_by(wmo, park_depth) |> summarize(correl = cor(slope, weighted_group_biovolume))
-
-ggplot(test) + geom_point(aes(x = log(slope), y = log(weighted_group_biovolume))) + facet_wrap(~park_depth)
+# test2 <- test |> group_by(wmo, cycle) |> dplyr::mutate(mean_mean_date = mean(floor_date(mean_date))) 
+# 
+# ggplot(test2) +
+#   geom_pointrange(aes(x = mean_mean_date, y = weighted_group_biovolume),
+#                       stat = 'summary',
+#                       fun.min = function(z) {quantile(z, 0.25)},
+#                       fun.max = function(z) {quantile(z, 0.75)},
+#                       fun = median, colour = 'black', size = .1) +
+#     facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
+#   
+# ggplot(test2) +
+#   geom_pointrange(aes(x = mean_mean_date, y = slope),
+#                       stat = 'summary',
+#                       fun.min = function(z) {quantile(z, 0.25)},
+#                       fun.max = function(z) {quantile(z, 0.75)},
+#                       fun = median, colour = 'black', size = .1) +
+#     facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) 
+# 
+# test2 |> group_by(wmo, park_depth) |> summarize(correl = cor(slope, weighted_group_biovolume))
+
+ggplot(slope_vs_biovolume) + geom_point(aes(x = log(slope), y = log(weighted_group_biovolume))) + facet_wrap(~park_depth) + labs(x = 'OST slope [UNITS]', y = 'UVP biovolume [UNITS]') #+ theme(axis.title.x = element_blank())
 #ggplot(test) + geom_point(aes(x = slope, y = weighted_group_biovolume)) + facet_wrap(~park_depth)
@@
 # what about jump values with uvp measurements at jump
-uvp_drifting <- uvp_drifting |> mutate(num_date = as.numeric(juld))
-res_jumps <- res_jumps |> mutate(num_date = as.numeric(dates))
+uvp_drifting <- uvp_drifting |> dplyr::mutate(num_date = as.numeric(juld))
+res_jumps <- res_jumps |> dplyr::mutate(num_date = as.numeric(dates))
 
@@
 
-test_jump <- data.frame()
-for (i in wmo){
-  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
-  for (j in park_depth){
-    for (k in seq(1:max_cycle)){
-      tmp_uvp <- filter(uvp_drifting, wmo == i, cycle == k, park_depth == j) |> arrange(juld) |> select(juld, cycle, wmo, park_depth, total_biovolume, num_date) 
-      tmp_ost_jump <- res_jumps |> filter(wmo == i, cycle == k, park_depth == j)
-      if(nrow(tmp_ost_jump) == 0 | nrow(tmp_uvp) == 0){
-        next
-      }else{
-        tmp <- assign_uvp_to_jump(tmp_uvp, tmp_ost_jump)
-      }
-      test_jump <- rbind(test_jump, tmp)
-    }
-  }
-}
+# test_jump <- data.frame()
+# for (i in wmo){
+#   max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
+#   for (j in park_depth){
+#     for (k in seq(1:max_cycle)){
+#       tmp_uvp <- filter(uvp_drifting, wmo == i, cycle == k, park_depth == j) |> arrange(juld) |> select(juld, cycle, wmo, park_depth, total_biovolume, num_date) 
+#       tmp_ost_jump <- res_jumps |> filter(wmo == i, cycle == k, park_depth == j)
+#       if(nrow(tmp_ost_jump) == 0 | nrow(tmp_uvp) == 0){
+#         next
+#       }else{
+#         tmp <- assign_uvp_to_jump(tmp_uvp, tmp_ost_jump)
+#       }
+#       test_jump <- rbind(test_jump, tmp)
+#     }
+#   }
+# }
+
+#vroom_write(test_jump, file = '/home/flo/dox/THESIS/DATA/ARGO/diff_jump_vs_biovolume.csv')
+
+# load data computed with the above commented code
+diff_jump_vs_biovolume <- vroom::vroom('DATA/ARGO/diff_jump_vs_biovolume.csv')
+diff_jump_vs_biovolume <- diff_jump_vs_biovolume|> mutate(park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
 
 #ggplot(test_jump) + geom_point(aes(diff_jump, biovolume))
-ggplot(test_jump) + geom_point(aes(log(diff_jump), log(biovolume)))
-```
-
-```{r, model based on slope}
-
-# compute biovolumes for mid size bin (ESD)
-# DSE from 102 microns to 2.05 mm (this class to goes to 2.58 mm) (values obtained from Guidi's Matlab code converted on python)
-
-# DSE <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
-#               0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
-#               1.16170742, 1.46365963, 1.84409558, 2.32341484)
-# 
-# # DSE obtained with Rainer's code
-# DSE_Rainer <- c(0.11645117591608793,0.14636003747903648,0.18439151433041295,
-#                 0.23251999827981656,0.2933257316240555,0.36916500651202966,0.4650399965596331,
-#                 0.5860454935035602,0.738553273391974,0.9280422447589767,1.1705684285914486,
-#                 1.4795320531800742,1.8636618038457486,2.3449457391257185)
-
-# DSE <- c(0.29042685,0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779, 1.16170742, 1.46365963)
-# 
-# # compute biovolume
-# get_biovolume <- function(DSE){
-#   r <- DSE/2
-#   biovolume <- (4*pi/3)*r**3
-#   return(biovolume)
-# }
-
-# # get biovolume for each DSE (equivalent to say for each size bin)
-# biovolumes <- map_dbl(DSE, get_biovolume)  # unit = mm^3
-
-# spectra at parking depth
-uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
-  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
-
-# convert abundance to biovolume
-spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes)))
-colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')
-
-# add data to initial abundance spectra
-uvp_spectra <- cbind(uvp_spectra, spectra_biovolume)
-
-# recompute slope based on Estapa's formula
-cflux <- cflux |> mutate(small_slope = (small_flux/633)**(1/0.77), big_slope = (big_flux/633)**(1/0.77))
-
-## build model
-flux_model <- function(boosting_rounds, which_slope){
-
-  # select which flux to build the model
-  if(which_slope == 'small'){
-    cflux2 <- cflux |> mutate(flux = log(small_slope)) |> filter(is.finite(flux))
-  }else{
-    cflux2 <- cflux |> mutate(flux = log(big_slope)) |> filter(is.finite(flux)) # because log(0) creates non finite value -> bug
-  }
-
-  # merge data for training
-  data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
-
-  # training data
-  tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))
-
-  set.seed(101)
-  # model
-  model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
-  cv.folds=5,
-  n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
-  bag.fraction=0.5,
-  n.cores=1
-  )
-
-  # get best model and thus best round number
-  best <- gbm.perf(model)
-  #print(best)
-
-  # get features importance
-  features_importance <- summary(model, n.trees=best, plotit = FALSE)
-
-  return(list(features_importance = features_importance, model = model, best_tree = best))
-}
-
-# small_model <- flux_model(3000, 'small')
-# big_model <- flux_model(3000, 'big')
-# save(small_model, big_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data_slope')
-
-load(file = 'DATA/MODEL/model_xgboost_data_slope')
-
-# # get spectra
-# uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
-#   drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
-# 
-# flux_model <- function(boosting_rounds, which_flux){
-#     
-#     # select which flux to build the model
-#     if(which_flux == 'small'){
-#       cflux2 <- cflux |> mutate(flux = small_flux) 
-#     }else if(which_flux == 'big'){
-#       cflux2 <- cflux |> mutate(flux = big_flux)
-#     }else{
-#       cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
-#     }
-#     
-#     # merge data for training
-#     data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
-#     
-#   # training data
-#   tmp <- data_model |> select(flux, all_of(lpm_class))
-#   
-#   set.seed(101)
-#   # model
-#   model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
-#   cv.folds=5,
-#   n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
-#   bag.fraction=0.5,
-#   n.cores=1
-#   )
-#    
-#   # get best model and thus best round number
-#   best <- gbm.perf(model)
-#   #print(best)
-#   
-#   # get features importance
-#   features_importance <- summary(model, n.trees=best, plotit = FALSE)
-#   
-#   
-#   return(list(features_importance = features_importance, model = model, best_tree = best))
-#    
-# }
-#   
-# # small_model <- flux_model(3000, 'small')
-# # big_model <- flux_model(3000, 'big')
-# # total_model <- flux_model(3000, 'total')
-# # 
-# # save(small_model, big_model, total_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data') 
-#   
-# load(file = 'DATA/MODEL/model_xgboost_data')
-
-```
-
-```{r, model predictions}
-
-model_predict <- function(which_slope, mod, best_tree_number){
-
-    # get spectra
-    uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
-  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
-
-    # convert abundance to biovolume
-    spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes)))
-    colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')
-
-    # add data to initial abundance spectra
-    uvp_spectra <- cbind(uvp_spectra, spectra_biovolume)
-
-    # select which flux to build the model
-    if(which_slope == 'small'){
-      cflux2 <- cflux |> mutate(flux = small_slope)
-    }else{
-      cflux2 <- cflux |> mutate(flux = big_slope)
-    }
-
-    # merge data for training
-    data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
-
-    # # training data
-    # tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))
-
-    # predictions on mean spectra values
-    uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = log(unique(flux)),
-                                                                                   NP_Biovolume_102 = mean(NP_Biovolume_102),
-                                                                                   NP_Biovolume_128 = mean(NP_Biovolume_128),
-                                                                                   NP_Biovolume_161 = mean(NP_Biovolume_161),
-                                                                                   NP_Biovolume_203 = mean(NP_Biovolume_203),
-                                                                                   NP_Biovolume_256 = mean(NP_Biovolume_256),
-                                                                                   NP_Biovolume_323 = mean(NP_Biovolume_323),
-                                                                                   NP_Biovolume_406 = mean(NP_Biovolume_406),
-                                                                                   NP_Biovolume_512 = mean(NP_Biovolume_512),
-                                                                                   NP_Biovolume_645 = mean(NP_Biovolume_645),
-                                                                                   NP_Biovolume_813 = mean(NP_Biovolume_813),
-                                                                                   NP_Biovolume_1020 = mean(NP_Biovolume_1020),
-                                                                                   NP_Biovolume_1290 = mean(NP_Biovolume_1290),
-                                                                                   NP_Biovolume_1630 = mean(NP_Biovolume_1630),
-                                                                                   NP_Biovolume_2050 = mean(NP_Biovolume_2050)) |> ungroup()
-
-  # prediction
-  pred_flux <- predict(mod, newdata = uvp_spectra_average, n.trees = best_tree_number)
-
-  uvp_spectra_average$pred_flux <- pred_flux
-
-  return(uvp_spectra_average)
-}
-
-# predictions
-# pred_small <- model_predict('small', small_model$model, small_model$best_tree) |> select(wmo, cycle, depth, true_small = true_flux, pred_small = pred_flux)
-# pred_big <- model_predict('big', big_model$model, big_model$best_tree) |> select(true_big = true_flux, pred_big = pred_flux) 
-# all_predictions <- cbind(pred_small, pred_big)
-# vroom::vroom_write(all_predictions, file = '/home/flo/dox/THESIS/DATA/MODEL/all_predictions_slopes.csv')
-all_predictions <- vroom('DATA/MODEL/all_predictions_slopes.csv')
-
-# model_predict <- function(which_flux, mod, best_tree_number){
-# 
-#     # get spectra
-#     uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
-#   drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
-# 
-#     # select which flux to build the model
-#     if(which_flux == 'small'){
-#       cflux2 <- cflux |> mutate(flux = small_flux)
-#     }else if(which_flux == 'big'){
-#       cflux2 <- cflux |> mutate(flux = big_flux)
-#     }else{
-#       cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
-#     }
-# 
-#     # merge data for training
-#     data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
-# 
-#   # predictions on mean spectra values
-#   uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = unique(flux),
-#                                                                                    NP_Size_102 = mean(NP_Size_102),
-#                                                                                    NP_Size_128 = mean(NP_Size_128),
-#                                                                                    NP_Size_161 = mean(NP_Size_161),
-#                                                                                    NP_Size_203 = mean(NP_Size_203),
-#                                                                                    NP_Size_256 = mean(NP_Size_256),
-#                                                                                    NP_Size_323 = mean(NP_Size_323),
-#                                                                                    NP_Size_406 = mean(NP_Size_406),
-#                                                                                    NP_Size_512 = mean(NP_Size_512),
-#                                                                                    NP_Size_645 = mean(NP_Size_645),
-#                                                                                    NP_Size_813 = mean(NP_Size_813),
-#                                                                                    NP_Size_1020 = mean(NP_Size_1020),
-#                                                                                    NP_Size_1290 = mean(NP_Size_1290),
-#                                                                                    NP_Size_1630 = mean(NP_Size_1630),
-#                                                                                    NP_Size_2050 = mean(NP_Size_2050)) |> ungroup()
-# 
-# 
-#   # prediction at 1000 m
-#   #spectra <- uvp_spectra_average |> select(true_flux, all_of(lpm_class[]))
-# 
-#   # prediction
-#   pred_flux <- predict(mod, newdata = uvp_spectra_average, n.trees = best_tree_number)
-# 
-#   uvp_spectra_average$pred_flux <- pred_flux
-# 
-#   return(uvp_spectra_average)
-# }
-# 
-# # pred_small <- model_predict('small', small_model$model, small_model$best_tree) |> select(wmo, cycle, depth, true_small = true_flux, pred_small = pred_flux)
-# # pred_big <- model_predict('big', big_model$model, big_model$best_tree) |> select(true_big = true_flux, pred_big = pred_flux)
-# # pred_total <- model_predict('total', total_model$model, total_model$best_tree) |> select(true_total = true_flux, pred_total = pred_flux)
-# #
-# # all_predictions <- cbind(pred_small, pred_big, pred_total) |> mutate(pred_total2 = pred_small + pred_big, anom = pred_total - pred_total2)
-# #
-# # vroom::vroom_write(all_predictions, file = '/home/flo/dox/THESIS/DATA/MODEL/all_predictions.csv')
-# all_predictions <- vroom('DATA/MODEL/all_predictions.csv')
-
-```
-
-```{r, plot model predictions}
-
-p_small <- ggplot(all_predictions) + geom_point(aes(true_small, pred_small)) + theme_bw() 
-round(R2_Score(y_true = all_predictions$true_small, y_pred = all_predictions$pred_small), 2)
-
-p_big <- ggplot(all_predictions |> filter(is.finite(true_big))) + geom_point(aes(true_big, pred_big)) + theme_bw()
-all_predictions |> filter(is.finite(true_big)) |> summarize(round(R2_Score(true_big, pred_big)), 2)
-
-p_small / p_big
+ggplot(diff_jump_vs_biovolume) + geom_point(aes(log(diff_jump), log(biovolume))) + facet_wrap(~park_depth) + labs(x = 'OST jump difference [UNITS]', y = 'UVP biovolume [UNITS]')
 
@@
 -   phénologie du bloom printanier
--
+-   qinking speed -> trouver un truc à dire et estimer l'incertitude (en lien avec le graphique ou le DZ augmente)   + expliquer la méthode vite fait (Léo Lacour, personal communication)
+- on se trouve dans un sytème producitve entre guidi 2008 et iversen (très productif)
 
@@
 # clean and smoothed chla data
-d <- npq_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> mutate(date = as_date(juld))
+d <- npq_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> dplyr::mutate(date = as_date(juld))
 d <- d |> group_by(date, wmo, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
-d <- d |> mutate(smoothed_chla = smooth(mean_chla)/2.15, smoothed_chla_up = smooth(mean_chla)/(2.15-0.37), smoothed_chla_down = smooth(mean_chla)/(2.15+0.37)) # 2.15 = Roesler factor in the Labrador Sea
+d <- d |> dplyr::mutate(smoothed_chla = smooth(mean_chla)/2.15, smoothed_chla_up = smooth(mean_chla)/(2.15-0.37), smoothed_chla_down = smooth(mean_chla)/(2.15+0.37)) # 2.15 = Roesler factor in the Labrador Sea
 
@@
 # add couple info
-d_chla <- d_chla |> mutate(group = if_else(wmo %in% c(6904240,4903634), '6904240 - 4903634', '6904241 - 1902578'))
+d_chla <- d_chla |> dplyr::mutate(group = if_else(wmo %in% c(6904240,4903634), '6904240 - 4903634', '6904241 - 1902578'))
 
@@
 chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_without_clouds_sat.feather') |> select(time, lat, lon, chla = CHL)
-mld_sat <- arrow::read_feather('DATA/SATELLITE/mld_sat.feather') |> mutate(longitude = longitude - 360) |> select(time, lat = latitude, lon = longitude, mld = mlotst)
+mld_sat <- arrow::read_feather('DATA/SATELLITE/mld_sat.feather') |> dplyr::mutate(longitude = longitude - 360) |> select(time, lat = latitude, lon = longitude, mld = mlotst)
 
@@
 chla_cover_deployment1 <- points_in_circle(chla_sat, lon_center = -49.2, lat_center = 58.9, radius = 100000) # 50 km around center
-chla_cover_deployment1 <- chla_cover_deployment1 |> mutate(date = as_date(time), group = '6904240 - 4903634') |> mutate(deploy = as_date('2022-05-22'))
+chla_cover_deployment1 <- chla_cover_deployment1 |> dplyr::mutate(date = as_date(time), group = '6904240 - 4903634') |> dplyr::mutate(deploy = as_date('2022-05-22'))
 
 chla_cover_deployment2 <- points_in_circle(chla_sat, lon_center = -52.3, lat_center = 56.7, radius = 10000) # 50 km around center
-chla_cover_deployment2 <- chla_cover_deployment2 |> mutate(date = as_date(time), group = '6904241 - 1902578') |> mutate(deploy = as_date('2022-05-30'))
+chla_cover_deployment2 <- chla_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-30'))
 
@@
 mld_cover_deployment1 <- points_in_circle(mld_sat, lon_center = -49.2, lat_center = 58.9, radius = 100000) # 50 km around center
-mld_cover_deployment1 <- mld_cover_deployment1 |> mutate(date = as_date(time), group = '6904240 - 4903634') |> mutate(deploy = as_date('2022-05-22'))
+mld_cover_deployment1 <- mld_cover_deployment1 |> dplyr::mutate(date = as_date(time), group = '6904240 - 4903634') |> dplyr::mutate(deploy = as_date('2022-05-22'))
 
 mld_cover_deployment2 <- points_in_circle(mld_sat, lon_center = -52.3, lat_center = 56.7, radius = 100000) # 50 km around center
-mld_cover_deployment2 <- mld_cover_deployment2 |> mutate(date = as_date(time), group = '6904241 - 1902578') |> mutate(deploy = as_date('2022-05-30'))
+mld_cover_deployment2 <- mld_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-30'))
 
@@
   geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
-  geom_point(data = d_chla |> filter(date < '2022-08-01') |> mutate(week = week(date), DOY = yday(date)) |> group_by(week, group) |> summarize(median_mld = median(mld), mean_date = as_date(mean(DOY), origin = '2022-01-01')), aes(mean_date, median_mld), colour = "red")
+  geom_point(data = d_chla |> filter(date < '2022-08-01') |> dplyr::mutate(week = week(date), DOY = yday(date)) |> group_by(week, group) |> summarize(median_mld = median(mld), mean_date = as_date(mean(DOY), origin = '2022-01-01')), aes(mean_date, median_mld), colour = "red")
 
@@
 -   Montrer que les sinking speed observées ne sont pas compatibles avec des valeurs de bloom (càd rapides pour les grosses particules à minima)
+- rajouter dans les méthodes le calcul de cette sinking speed.
 
@@
 
-sinking_fit <- vroom('DATA/ARGO/gaussian_fit_sinking_speed_full.csv') |> mutate(size = size_class |> str_remove('NP_Size_') |> as.numeric(), date = as_date(date))
+# NOTE : pas besoin d'updater ici vu que l'on regarde par rapport au 'bloom' de fin mai/juin
+sinking_fit <- vroom('DATA/ARGO/gaussian_fit_sinking_speed_full_before_06032023.csv') |> dplyr::mutate(size = size_class |> str_remove('NP_Size_') |> as.numeric(), date = as_date(date))
 
@@
 
-particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))
-
-particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld))
+
+particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) 
 #particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
 
+# remove buggy data from float 4903634 (negative depth for some reason)
+particles_bis <- particles_bis |> filter(depth >= 0)
+
 # rel_conc
-particles_bis <- particles_bis |> group_by(wmo, size) |> mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
+particles_bis <- particles_bis |> dplyr::group_by(wmo, size) |> dplyr::mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
 
@@
 
+all_sinking_speed |> ggplot() + geom_point(aes(y = sinking_speed_below_threshold, x = size))
+
 ```
@@
 
-ggplot(test, aes(x = ztime, y = fgauss, colour = factor(depth))) + geom_point() + theme_bw() #+ ylim(c(0,.2))
+ggplot(test, aes(x = ztime, y = fgauss, colour = factor(depth))) + geom_point() + 
+  scale_color_viridis_d() +
+  theme_bw()
 
@@
 
--   Pour le drift à 200, 500 et 1000 m
+-   Pour le drift à 200, 500 et 1000 m -> interpète ce graph -> va checker les masses d'eau de la region...
+attention on est en PARKING DEPTH -> mettre la couleur de la parking depth dans le truc, ça permettrait de voir si on peut comparer les processus (style ratio 500/1000m) dans les couches qui sont dans la meme masse d'eau.
 
 ```{r, TS diagram}
-
-#ts_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth) |> drop_na(temp) |> filter(sigma > 26) |> mutate(date = as_date(date))
-
-ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth) |> drop_na(temp) |> filter(sigma > 26, park_depth != '200 m') |> mutate(date = as_date(date))
-#ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth) |> drop_na(temp) |> filter(sigma > 26) |> mutate(date = as_date(date))
-ggplot(ts_data) + geom_point(aes(psal, temp, colour = factor(park_depth))) + theme_bw()
-
-# ggplot(ts_data) + geom_point(aes(psal, temp, colour = wmo))
-
-ts_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth) |> drop_na(temp) |> filter(sigma > 26) |> mutate(date = as_date(date))
-
-# add contour
-# breaks_sigma <- seq(min(round(ts_data$sigma*2)/2), max(round(ts_data$sigma*2)/2), by = 1)
-ggplot(ts_data) + geom_point(aes(psal, temp, colour = depth)) + scale_color_viridis_c(option = 'A') + theme_bw()
+#ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26, park_depth != '200 m') |> dplyr::mutate(date = as_date(date))
+ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
+  dplyr::mutate(date = as_date(date), colour = if_else(park_depth == '200 m', '#E31B23', if_else(park_depth == '500 m', '#FFC325', 'black')))
+
+ts_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
+  dplyr::mutate(date = as_date(date))
+
+# https://dankelley.github.io/oce/reference/plotTS.html
+# https://dankelley.github.io/oce/reference/argo-class.html
+test <- as.argo(time = ts_data$date, longitude = ts_data$lon, latitude = ts_data$lat, salinity = ts_data$psal, temperature = ts_data$temp, pressure = ts_data$depth,
+        id = ts_data$wmo)
+
+plotTS(test, inSitu = T, type = 'p', xlab = 'Salinity [PSU]')
 
@@
 
-# tmp <- uvp_data |> filter(park_depth != '200 m') |> group_by(wmo, size, month, park_depth) |> 
-#   summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
-#   # compute concentration anomaly at 500 and 1000 m
-#   mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
-#   select(wmo, size, month, diff_anomaly, div_anomaly) |> mutate(month = month.abb[month]) |> 
-#   mutate(size = as.character(size)) |>
-#   mutate(rev_div_anomaly = 1/div_anomaly) |> 
-#   #filter(is.finite(rev_div_anomaly)) |>
-#   mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1), div_anomaly = round(div_anomaly, 1)) |>
-#   ungroup()
+tmp <- uvp_data |> filter(park_depth != '200 m') |> group_by(wmo, size, month, park_depth) |>
+  summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
+  # compute concentration anomaly at 500 and 1000 m
+  dplyr::mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
+  select(wmo, size, month, diff_anomaly, div_anomaly) |> dplyr::mutate(month = month.abb[month]) |>
+  dplyr::mutate(size = as.character(size)) |>
+  dplyr::mutate(rev_div_anomaly = 1/div_anomaly) |>
+  #filter(is.finite(rev_div_anomaly)) |>
+  dplyr::mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1), div_anomaly = round(div_anomaly, 1)) |>
+  ungroup()
 
@@
 #   # compute concentration anomaly at 500 and 1000 m
-#   mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
+#   dplyr::mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
 #   select(wmo, size, week, diff_anomaly, div_anomaly) |> 
-#   mutate(size = as.character(size)) |>
-#   mutate(rev_div_anomaly = 1/div_anomaly) |> 
+#   dplyr::mutate(size = as.character(size)) |>
+#   dplyr::mutate(rev_div_anomaly = 1/div_anomaly) |> 
 #   #filter(is.finite(rev_div_anomaly)) |>
-#   mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1)) |>
+#   dplyr::mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1)) |>
 #   ungroup()
@@
 # reload + redo some annoying stuff that were lost during the writing of the csv
-dd <- vroom('heatmap_data.csv') |> mutate(Y = as.character(Y)) 
-dd$X <- factor(dd$X, levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))
+dd <- vroom('heatmap_data.csv') |> dplyr::mutate(Y = as.character(Y)) 
+dd$X <- factor(dd$X, levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan', 'Feb', 'Mar'))
 dd$Y <- factor(dd$Y, levels = c('102', '128', '161', '203', '256', '323', '406', '512', '645', '813', '1020', '1290'))
@@
 
-#dd <- dd |> mutate(Z_binned = cut(Z, breaks = c(0,1,1.5,2.5,max(Z))))
-dd <- dd |> mutate(Z_binned = cut(Z, breaks = c(0,.5,1,1.5,2,max(Z))))
+#dd <- dd |> dplyr::mutate(Z_binned = cut(Z, breaks = c(0,1,1.5,2.5,max(Z))))
+dd <- dd |> dplyr::mutate(Z_binned = cut(Z, breaks = c(0,.5,1,1.5,2,max(Z))))
 
@@
 
-```{r, test gt table}
-
-monthly_mean <- uvp_data |> mutate(month = month.abb[month(juld)]) |> group_by(size, park_depth, month) |> summarize(mean_conc = round(mean(conc, na.rm=T), 2)) |> ungroup()
-
-# pivot wider
-monthly_mean <- monthly_mean |> pivot_wider(names_from = month, values_from = mean_conc) |> select(size, depth = park_depth, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec, Jan, Feb)
-
-monthly_mean |> group_by(size) |> gt(rowname_col = 'depth')
-
-#monthly_mean |> gt(rowname_col = c('size')) |> tab_stubhead(label = 'Class')
-
-```
-
-#### Features importance
-
-```{r, features importance}
-# 
-# features <- small_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'small') |>
-#   bind_rows(big_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'big')) |>
-#   bind_rows(total_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'total'))
-# 
-# rownames(features) <- NULL
-# 
-# features <- features |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
-# 
-# ggplot(features) + geom_point(aes(x = as.factor(size), y = importance, colour = model), size = 5) + labs(x = 'Size', y = 'Features importance')
-
-```
-
-#### Predictions
-
-```{r, plot predictions}
-
-# p_small <- ggplot(all_predictions) + geom_point(aes(true_small, pred_small)) + xlim(c(0,25)) + ylim(c(0,25)) + theme_bw() +
-# 
-#   annotate("text", x = 3, y = 23, label = paste0('R² = ' , round(R2_Score(all_predictions$true_small, all_predictions$pred_small), 2)))
-# 
-# p_big <- ggplot(all_predictions) + geom_point(aes(true_big, pred_big)) + xlim(c(0,50)) + ylim(c(0,50)) + theme_bw() +
-# 
-#   annotate("text", x = 5, y = 45, label = paste0('R² = ' , round(R2_Score(all_predictions$true_big, all_predictions$pred_big), 2)))
-# 
-# p_total <- ggplot(all_predictions) + geom_point(aes(true_total, pred_total)) + xlim(c(0,100)) + ylim(c(0,100)) + theme_bw() +
-# 
-#   annotate("text", x = 3, y = 80, label = paste0('R² = ' , round(R2_Score(all_predictions$true_total, all_predictions$pred_total), 2)))
-# 
-# (p_small + p_big) / p_total
-
-# plot(test$true_flux, y=test$pred_flux, asp=1)
-
-# print(R2_Score(test$pred_flux, test$true_flux))
-
-```
-
-#### same plot mais pour les classes de tailles
-
-```{r, TEST }
-
-# depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
-# 
-# part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_256) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
-# 
-#   mutate(month = month.abb[month(juld)]) |> ungroup()
-# 
-# tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_256))) |>
-# 
-#   mutate(cut_part = cut(mean_part, breaks = c(seq(0,100,20),+Inf))) |> ungroup()
-# 
-# create_data_for_heatmap_bis <- function(float, df){
-# 
-#   df <- df |> filter(wmo == float) |> select(mean_part, depth_bin, month)
-# 
-#   y <- unique(df$depth_bin)
-# 
-#   df <- df|> pivot_wider(names_from = depth_bin, values_from = mean_part)
-# 
-#   x <- unique(df$month)
-# 
-#   data <- expand.grid(X=x, Y=y)
-# 
-#   data$Z <- as.vector(as.matrix(df |> select(-month)))
-# 
-#   data$wmo <- float
-# 
-#   return(data)
-# 
-# }
-# 
-# d1 <- create_data_for_heatmap_bis(6904241, tmp)
-# 
-# d2 <- create_data_for_heatmap_bis(1902578, tmp)
-# 
-# d3 <- create_data_for_heatmap_bis(4903634, tmp)
-# 
-# d4 <- create_data_for_heatmap_bis(6904240, tmp)
-# 
-# dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
-# 
-#   mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
-# 
-#   mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) #|>
-# 
-#   #mutate(Z = factor(Z, levels = levels(tmp$cut_part)))
-# 
-# ggplot(dd) +
-# 
-#   geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Log10(abundance) #/L', x = 'Month', y = 'Depth bin', title = 'NP_Size_256') +
-# 
-#   facet_wrap(~wmo) + scale_fill_viridis(option = 'A')
-
-```
-
-#### same plots mais par mois
-
-```{r, monthly heatmap}
-
-# depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
-# 
-# part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, all_of(lpm_class)) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |> mutate(month = month.abb[month(juld)]) |> ungroup() |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration')
-# 
-# tmp <- part_distrib |> group_by(depth_bin, size, month) |> summarize(mean_part = mean(concentration)) |> mutate(rel_mean_conc = (mean_part-min(mean_part))/(max(mean_part) - min(mean_part))) |> ungroup()
-# 
-# plot_heat_month <- function(m){
-# 
-#   # transform data for heatmap
-# 
-#   #m <- 'May'
-# 
-#   tmp2 <- tmp |> filter(month == m) |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) |> select(-c(month, mean_part))
-# 
-#   y <- unique(tmp2$depth_bin)
-# 
-#   x <- unique(tmp2$size)
-# 
-#   test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = rel_mean_conc)
-# 
-#   #test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = mean_part)
-# 
-#   data <- expand.grid(X=x, Y=y)
-# 
-#   data$Z <- as.vector(as.matrix(test |> select(-size)))
-# 
-#   data$month <- m
-# 
-#   return(data)
-# 
-# }
-# 
-# m_data <- map_dfr(c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'), plot_heat_month)
-# 
-# m_data <- m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)),
-# 
-#                    month = factor(month, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')))
-# 
-# ggplot(m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)))) +
-# 
-#   geom_tile(aes(as.factor(X), Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + scale_fill_viridis(option = 'A') +
-# 
-#   facet_wrap(~month)
-
-```
-
 ## Discussion2

### 2023-03-06 discussion.qmd
update

# General discussion and perspectives

## UVP6 classification model: From start to finish

The development of the UVP6 classification model started without a data set of labeled UVP6 images. Therefore, a significant part of my work was to test and prepare a fully functional data preparation and model training pipeline for the upcoming UVP6 data set using UVP5 images. In the following sections, I will describe a brief historic of what has been done prior to the acquisition and human labeling of UVP6 images.

### Classification metric

<!-- https://datascience.stackexchange.com/questions/61761/how-does-xgboost-use-softmax-as-an-objective-function -->

In multiclass classification task, the goal is to predict the label (i.e. class) of a given input among a defined number of labels. In @sec-uvp6, I explained that XGBoost minimizes a regularized objective function $L$ (see @eq-objective-function-xgboost) during model training. The XGBoost library proposes two choices of objective functions for multiclass classification problems: `multi:softmax` and `multi:softprob` (https://xgboost.readthedocs.io/en/latest/parameter.html, last accessed: 03/03/2023). Both functions are based on the softmax normalization which normalizes the classification score of each class using an exponential function (@eq-softmax). `multi:softmax` returns the class of highest probability whereas `multi:softprob` outputs a class probability array, that allows the computation of custom objective functions.

$$
\sigma(x_{i}) = \frac{e^{x_{i}}}{\sum_{j=1}^{N}e^{x_{j}}}, i = 1,...,N
$$ {#eq-softmax}

where $x_{i}$ and $N$ are, respectively, the classification score of the $i^{th}$ class and the number of classes.

By definition, `multi:softmax` amplifies the difference between classification scores. Here is a simple illustration of this effect. Let A, B and C, three classes of zooplankton whose classification scores are $(1,7,8)$. Using @eq-softmax, the output probabilities are $(0.0006, 0.26, 0.73)$. Therefore, a unit difference in the classification scores of class B and class C results in a probability three times higher for class C than class B.

This lead us to define what we called a trust index, that is a metric whose score indicates the degree of confidence in the classification score. In other words, we looked for the best error estimator (@fig-trust-index). To this end, we trained a first classification model with classes from the UVP5 data set using a `multi:softprob` objective function. Then, we used the model to make a prediction on a test set that returned a matrix $(M \times N)$ of $M$ images and $N$ classes. Afterwards, we computed four metrics on each row of the probability matrix: the maximum probability ($max$, equivalent to `multi:softmax`), the softmax of the probability array ($softmax$, equivalent to a double `multi:softmax`), the difference between the highest and the second highest probability ($diff_{max}$) and the difference between the highest and second highest probability of the softmax array ($diff_{softmax}$). We compared those metrics for each class of the test set in @fig-trust-index which clearly shows that all metrics were very similar and that they captured the errors at the same speed. Thus, it meant that we could keep using XGBoost as we initially did, that is using `multi:softprob` as the objective function. @fig-trust-index also shows that some classes kept accumulating errors in the prediction with increasing rank (e.g. *Copepoda*, darksphere and fiber\<detritus). In their case, an increasing rank is not a trustworthy indicator of increasing confidence.

```{r, trust index, dpi = 300}
#| label: fig-trust-index
#| fig-cap: Percentage of error captured as a function of rank for four metrics (max, softmax, diff$_{max}$ and diff$_{softmax}$) on six UVP5 classes (4 biological, 2 non biological). The number of objects predicted in each class is provided in the title of each subplot.

library(gridExtra)
library(grid)
library(jpeg)

path <- paste0(getwd(), '/image/uvp6/trust_index')

grid.arrange(
  rasterGrob(as.raster(readJPEG(paste0(path, "/Aulacanthidae.jpeg")))),
  rasterGrob(as.raster(readJPEG(paste0(path, "/Aulosphaeridae.jpeg")))),
  rasterGrob(as.raster(readJPEG(paste0(path, "/Copepoda.jpeg")))),
  rasterGrob(as.raster(readJPEG(paste0(path, "/darksphere.jpeg")))),
  rasterGrob(as.raster(readJPEG(paste0(path, "/detritus.jpeg")))),
  rasterGrob(as.raster(readJPEG(paste0(path, "/fiber_detritus.jpeg")))),
  ncol = 3
)
```

### Technical constraints on the classification model

Because the UVP6-LP is designed to work on AUVs, several constraints were imposed on the model to fulfill technical and energy requirements. The UVP6 data treatment at sea shown in @fig-uvp-pipeline briefly describes the steps from the image acquisition to the prediction. The two most energy consuming steps are the image segmentation (median execution time of 650 ms) and the loading of the classification model whose execution time was capped at 700 ms. This limitation imposed a direct constraint on the prediction quality of the model because it implied a limitation on the choice of values for hyperparameters of the model (learning rate, tree depth) and the number of boosting rounds.

The second critical aspect of the model was its limitation at 40 classes. In the UVP5 SD data set, the most detailed classification at a taxonomic level by a human was roughly 150 classes (referred to as $level_{0}$). It was therefore necessary to merge classes based on morphological and ecological attributes to decrease that number in order to get to $level_{1}$ that had 38 classes. From $level_{1}$, we thus experimented a lot of different prediction models with a hyperparameters gridsearch in order to find both the right set of parameters and the right number of classes. There were indeed some classes that had very low precision and recall because they were confused with some specific taxonomic groups. They were merged accordingly with those groups. This trial-error approach was necessary to create the final aggregation level ($level_{2}$) used for the final UVP5 model. The same work had to be done later with the UVP6 data because the taxonomic groups were not identical.

The last technical limitation was the number of features (see table XX). Before I started my PhD, 55 features (basically geometry and gray levels) were defined. I was not allowed (though, I asked!) to do some feature engineering, that is either doing feature selection (selecting the most useful features) either doing feature combination (combining existing features to create new ones) or creating new features by collecting new data [@Geron2019-is]. Why was this proposition refused? Again, those 55 selected features had been tested earlier and they were compliant with the energy balance. Therefore, adding more features was not a possibility (we would have needed to redo tests and energy budget that were not possible at the time) and because they were computed quickly, they were not consuming too much energy hence there was no need to remove some of them. At worse, some features would have been redundant, without impacting the classification quality.

Nonetheless, when we compare the results from a featured-based model and a CNN (see SECTION XX), perhaps the results would have been closer if we would have done some feature engineering to add other kinds of features.

```{r, mean and median abundances in a typical vertical profile}
#| label: tbl-stats-uvp6
#| tbl-cap: Mean and median concentrations and number of segmented objects whose size is above 0.6 mm(referred to as "vignettes" in french) expected from a standard vertical profile taken by a BGC-Argo float equipped with a UVP6 in different parts of the water column.

library(gt)
library(gtsummary)
library(gtExtras)
library(tibble)
library(tidyverse)
library(latex2exp)

stats_uvp6 <- tibble(mean_conc = c(119, 457, 110, 76)/1000, 
                     median_conc = c(85, 332, 59, 38)/1000, 
                     mean_nb =  c(426,146,143,137),
                     median_nb = c(305,106,77,75),
                     layer = c('0-2000', '0-100', '100-500', '500-2000'),
                     water_volume = c(3.57, 0.32, 1.3, 1.95)*1000)

stats_uvp6 |> gt() |> cols_label(mean_conc = 'Mean',
                                 median_conc = 'Median',
                                 mean_nb = 'Mean',
                                 median_nb = 'Median',
                                 layer = 'Depth range (m)',
                                 water_volume = 'Sampled volume (L)') |>
  tab_spanner(label = 'Concentration (#/L)',
              columns = c(mean_conc, median_conc)) |>
   tab_spanner(label = 'Number',
              columns = c(mean_nb, median_nb))
```

feature engineering, citer le bouquin là

###

### Features

### Bilan énergique

### detritus case + imbalanced

### future embedded low energy models

### time series, get the UVP6 data back (intéret de les mettre sur les flotteurs)

volume échantillonnée vraiment faible, les trucs petits sont importants donc il faudrait l'uVP6 micro qui est on the way

### 2023-03-07 article3.qmd
update

@@
 -   dire que le but du papier est de faire une comparaison OST UVP qui sont deux instruments qui permettent de mesurer la meme chose (le flux des particules, l'abondance de particules) et donc le flux de carbone de manière indirecte papier GRL Lacour de 2015
+-   introduire le POC ici et introduire la BCP vite fait (classique)
+-   dire que Chla = proxy pour phyto biomass
+-   BGC-Argo floats are pseudo-lagrangian platforms
+-   si on peut prouver que le diagramme ST, les massses d'eau sont les memes, alors on peut peut-être parler des ratios entre particlues à 500 et 1000 m
+-   définir la méthode de calcul de la MLD
+-   plotter la MLD sur les profils verticaux de
+
+Two pairs of Biogeochemical-Argo (BGC) floats were planned to be deployed at the time of the annual spring bloom in the Labrador Sea (LS).
+The latter is characterized by two bioregions with distinct phytoplankton bloom phelonogies on both side of the 60°N parallel [@Lacour2015-da].
+The northern part is characterized by a short spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June [@Lacour2015-da].
+In this study, two pairs of floats were deployed south of the 60°N parallel with a first couple (WMO 6904240 and 4903634) deployed closely to it (58.9°N) in late May 2022.
 
@@
 
-```{r, map of 4 floats}
+```{r, map of 4 floats, fig.height=6, fig.width=6}
 #| label: fig-float-position
@@
 
-basemap(limits = c(-55,-40,55,65), bathymetry = T) +
-  geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), size = 1.5) + 
-  geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 3) +
+basemap(limits = c(-55,-40,55,65), bathymetry = T, lat.interval = 2.5, bathy.style = 'poly_blues',
+        grid.col = 'black') +
+  geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), linewidth = 1) + 
+  geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 2) +
   scale_color_brewer(palette = 'Dark2') + theme_bw() +
   geom_point(data = transform_coord(argo_map_first), aes(lon, lat), shape = 17, size = 3) +
-  geom_point(data = argo_map_last_geo, aes(lon, lat, colour = WMO), size = 6) +
-  geom_point(data = argo_map_last_geo, aes(lon, lat), fill = 'black', size = 6, shape = 1) +
+  geom_point(data = argo_map_last_geo, aes(lon, lat, colour = WMO), size = 4) +
+  geom_point(data = argo_map_last_geo, aes(lon, lat), fill = 'black', size = 4, shape = 1) +
   #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
   guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
-                                                  shape = c(NA, NA, NA, NA))))
+                                                  shape = c(NA, NA, NA, NA)))) #+
+  # basemap(limits = c(-55,-40,55,65), bathymetry = F, lat.interval = 2.5,
+  #       grid.col = 'black')
 
@@
 
-Data were downloaded from the Coriolis data center (biogeochemical data at ftp://ftp.ifremer.fr/ifremer/argo/dac/coriolis/ and UVP6 data at ftp://ftp.ifremer.fr/ifremer/argo/aux/coriolis/, last access: MONTH DAY, 2023).
-Chla data were quality controlled for Non-Photochemical Quenching (NPQ), a photoprotective mechanism that decreases the apparent fluorescence per unit of Chla following the method implemented by @Terrats2020 who modified the method of @Xing2018 for shallow-mixing cases.
+Data were downloaded from the Coriolis data center (*ftp://ftp.ifremer.fr/ifremer/argo/*, with biogeochemical data at *dac/coriolis/* and UVP6 data at *aux/coriolis/*, last accessed: March 6, 2023).
+Chla data were quality controlled for Non-Photochemical Quenching (NPQ), a photoprotective mechanism that decreases the apparent fluorescence per unit of Chla following the method implemented by @Terrats2020-ef who modified the method of @Xing2018-pn for shallow-mixing cases.
+Chla data were also divided by a slope factor of 2.15 following the recommendation of @Roesler2017-me for obtaining unbiased Chla estimates from in situ WETLabs ECO fluorometer.
 $c_{p}$ data were converted from counts to physical units (m$^{-1}$) using the following relationship
@@
 particle size distribution hereafter referred to as PSD) was divided by the number of images at each depth and the water volume imaged by the UVP (0.7L) to obtain particle concentrations (units: #/L).
-Note that in an impending update by Coriolis, it will not be necessary to divide by the number of images anymore.
-
-All data were divided into vertical profiles and drifting data using time stamps for drift.
+Note that in an impending update by the Coriolis data center, it should not be necessary to divide by the number of images anymore.
+
+All data were divided into vertical profiles and drifting data using time stamps for profiling and drifting acquisition periods.
 Finally, data were checked for instrumental anomalies.
 As a result, $c_{p}$ values from float 6904241 cannot be used from its 20$^{th}$ cycle onwards (\> November 17) due to a likely failure of the pump cleaning the optical window of the transmissometer.
+In total, the data coverage of this study spans a period of 9 months (end of May 2022 to early March 2023).
 
@@
 Then, we computed the anomaly to the moving median as the difference between the drifting $c_{p}$ signal and the moving median.
-We considered that a $c_{p}$ measurement was a spike when the anomaly was above $5.2 \times MAD$ following @Leys2013.
+We considered that a $c_{p}$ measurement was a spike when the anomaly was above $5.2 \times MAD$ following @Leys2013-eu.
 Despiked $c_{p}$ data were then smoothed with a 7-point moving median filter.
 
-Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach instead of fixed thresholds to detect jumps.
+Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr; @Estapa2023-oz], we chose an empirical statistical approach instead of fixed thresholds to detect jumps.
 Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase and computed the z-score (i.e.
@@
 In order to compute $F_{small}$, we firstly computed the slope of each group of observations between jumps (positive and negative, see @fig-estapa-method_article).
-Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window.
+Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window (e.g.
+turbulence ou zooplankton consumption of accumulated material).
 Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).
@@
 ```
+
+### Satellite data
+
+Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: [10.48670/moi-00281](https://doi.org/10.48670/moi-00281)) with a spatial resolution of 4 km.
+Mixed layer depth (MLD) derived from satellite observations were downloaded from near-real-time weekly data from the ARMOR3D product (doi: [10.48670/moi-00052](https://doi.org/10.48670/moi-00052)).
 
@@
 At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$.
-For both drifting depths, the trend shows a decrease of $F_{small}$ from June to January.
-At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid-July followed by a steady decrease towards January, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.
-
-```{r, plot Fsmall and Fbig}
+For both drifting depths, the trend shows a decrease of $F_{small}$ from June to, respectively, January and February.
+Note that the drifting configuration of all 4 floats was adapted to the deepening of the MLD (Fig.
+XX) hence drifting at 200 m and 500 m was stopped at this period to avoid measurements bias due to turbulence.
+At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid/end of July followed by a steady decrease towards March, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.
+
+```{r, plot Fsmall and Fbig, fig.width=8, fig.height=6}
 #| label: fig-OST-flux
-#| fig-cap: OST-derived small (top) and big (bottom) particles flux for all floats at each drifting depth.
+#| fig-cap: OST-derived small (top) and big (bottom) particles carbon flux for all floats at each drifting depth.
 Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.
@@
 High fluxes (\> 100 mg C m$^{-2}$ day$^{-1}$) were measured by 3 out of 4 floats at the end of May before a sharp decrease towards July.
-At 500 m, $F_{big}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to January.
-At 1000 m, $F_{big}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards January, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$.
-
-### UVP6 particle abundances
+At 500 m, $F_{big}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to February.
+At 1000 m, $F_{big}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards March, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$.
+
+### UVP6 particle abundance
 
@@
 In this study, we focus on 14 size classes, from 102 $\mu$m to 2.5 mm.
-In @fig-uvp-200m, @fig-uvp-500m and @fig-uvp-1000m, we show the daily averaged concentrations of particles in each class size for each float at their respective drifting depth.
-At 200 m, it can be observed that particles concentrations in the 102 - 512 $\mu$m range are relatively uniform between June and October before decreasing sharply towards January.
-Between 512 $\mu$m and 1.29 mm, concentrations are steadily decreasing from June.
-Above 1.29 mm, the abundance of particles is already very low (\< 0.1 particle/L) in June and keeps decreasing over time.
-
-```{r, uvp data at 200 m}
+In @fig-uvp-200m, @fig-uvp-500m and @fig-uvp-1000m, we show the daily averaged concentrations of particles in each size class for each float at their respective drifting depth.
+At 200 m, it can be observed that particle concentration in the 102 - 512 $\mu$m range are relatively uniform between June and October before decreasing sharply towards January.
+Between 512 $\mu$m and 1.63 mm, concentrations are steadily decreasing from June.
+Above 1.63 mm, the particle abundance is already very low (\< 0.1 particle/L) in June but keeps decreasing until September when they are practically absent (\< 0.01 particle/L).
+
+```{r, uvp data at 200 m, fig.height=8, fig.width=8}
 #| label: fig-uvp-200m
-#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 200 m.
-Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
-Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.
+#| fig-cap: Daily averaged concentrations of particle abundance in 14 size classes measured by the UVP6 for each float at 200 m.
+Black lines and shaded areas represent, respectively, the daily mean particle abundance and the 95% confidence interval around the mean.
+Dashed vertical lines indicate the time at which the daily mean particle abundance is maximum.
 
@@
 
-At 500 m, particles concentrations in the 102 - 645 $\mu$ range slowly decrease from July to September with a sharp decrease starting in October.
+At 500 m, particle concentrations in the 102 - 645 $\mu$m range slowly decrease from July to September with a sharp decrease starting in October.
 Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (\< 0.5 particle/L) on the entire timeserie.
 
-```{r, uvp data at 500 m}
+```{r, uvp data at 500 m, fig.width=8, fig.height=8}
 #| label: fig-uvp-500m
-#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 500 m.
-Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
-Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.
+#| fig-cap: Daily averaged concentrations of particle abundance in 14 size classes measured by the UVP6 for each float at 500 m.
+Black lines and shaded areas represent, respectively, the daily mean particle abundance and the 95% confidence interval around the mean.
+Dashed vertical lines indicate the time at which the daily mean particle abundance is maximum.
 
@@
 
-At 1000 m, all 4 floats measured a clear maximum in particles abundance in the 102 - 645 $\mu$m in the second half of July.
+At 1000 m, all 4 floats measured a clear maximum in particle abundance in the 102 - 645 $\mu$m in the second half of July.
 Above 645 $\mu$m, the same trend can still be observed with concentrations of less than 1 particle/L.
 
-```{r, uvp data at 1000 m}
+```{r, uvp data at 1000 m, fig.width=8, fig.height=8}
 #| label: fig-uvp-1000m
-#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 1000 m.
-Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.
-Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.
+#| fig-cap: Daily averaged concentrations of particle abundance in 14 size classes measured by the UVP6 for each float at 1000 m.
+Black lines and shaded areas represent, respectively, the daily mean particle abundance and the 95% confidence interval around the mean.
+Dashed vertical lines indicate the time at which the daily mean particle abundance is maximum.
 #| fig-dpi: 96
@@
 
-TODO : It should be noted that all floats measure the same trend -\> logique si on parle de la dynamique du bruit de fond de la zone et pas de la dynamique évènementielle qui serait différentes en fonction des flotteurs vu qu'ils dérivent chacun dans le parcours respectifs.
-Ça devrait renforcer cet aspect.
+#### On profiling mode
+
+Vertical profiles of particle concentration in the lower (102 - 512 $\mu$m) and upper (0.512 - 2.5 mm) size range of the UVP6 shows similar patterns across all floats.
+In the lower part of the size spectrum, @fig-uvp6-profiling-102-406 shows the rapid emptying of the water column in November.
+This process occurred on the entire water column (e.g.
+size class 406, float 4903634) but also started with more intensity from below the surface (e.g.
+size class 102, float 6904240).
+All floats also measured a local particle abundance maximum around 800 - 1000 m whose origin could be due to a previous particle export or remineralization of bigger particles.
+
+```{r, UVP6 profiling mode without sinking speed small classes, fig.height=8, fig.width=8}
+#| label: fig-uvp6-profiling-102-406
+#| fig-cap: Vertical profiles of UVP6 particle concentration for size classes in the 102 - 512 $\mu$m size range.
+Concentrations were normalized for each size class.
+The colorbar is in log$_{10}$ scale.
+
+particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld))
+
+particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+
+# remove buggy data from float 4903634 (negative depth for some reason)
+particles_bis <- particles_bis |> filter(depth >= 0)
+
+# rel_conc
+particles_bis <- particles_bis |> dplyr::group_by(wmo, size) |> dplyr::mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
+
+up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
+  #group_by(size) |>
+  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+  scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
+
+up_part
+```
+
+In the upper part of the size range, @fig-uvp6-profiling-512-2500 shows a rapid emptying of the entire water column starting in November.
+In contrast with the small classes, some particles, especially above 1.5 mm, have already virtually disappeared from the water column by September.
+However, the rapid increase and decrease of particles on the entire water column around July-August for floats 6904241 and 1902578 (still in the same zone at that time) could suggest an export event.
+
+```{r, UVP6 profiling mode without sinking speed big classes, fig.height=8, fig.width=8}
+#| label: fig-uvp6-profiling-512-2500
+#| fig-cap: Vertical profiles of UVP6 particle concentration for size classes in the 0.512 - 2.5 mm size range.
+Concentrations were normalized for each size class.
+The colorbar is in log$_{10}$ scale.
+The gray background depicts the absence of particle in the specified size class.
+
+down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
+  #group_by(size) |>
+  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+  scale_y_reverse() +
+  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
+  #geom_vline(xintercept = as_date('2022-01-01')) +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month')
+
+down_part
+```
 
 ## Discussion
+
+### Environmental conditions before and shortly after deployment
+
+All floats measured similar particle concentrations at each drifting depth and on the vertical.
+In addition, except perhaps for the upper part of the UVP6 size spectrum for floats 6904241 and 1902578, the results do not indicate a clear hint of particle export signature.
+Therefore, we suggest that the 4 BGC-Argo floats mostly measured the dynamics of the seasonal, continuous, particle signal but that they missed the main transient particle export events associated with the LS spring bloom.
+
+To estimate the time of the bloom in the deployment area, we compared MLD and surface (0 - 20 m) Chla concentrations from both satellite and in situ data (@fig-chla-mld-satellite) in a 100 km radius around each deployment.
+It can be seen that these estimates are in agreement, therefore we hypothesize that we can trust satellite data prior to the floats deployment.
+Taking this into consideration, it seems that the pair 6904240-4903634 missed a bloom that occurred in early May whereas the pair 6904241-1902578 missed a first Chla increase in late May days before their deployment.
+However, a second equivalent increase (median of 2 mg m$^{-3}$) happened in late June - early July, which could be a contributing factor to the apparent particle track observed around July and August.
+
+```{r, chla and MLD from satellite, fig.height=6, fig.width=6}
+#|: label: fig-chla-mld-satellite
+#| fig-cap: Comparison of MLD and Chla concentrations measured from satellites (black) and BGC-Argo floats (red) following their deployment.
+Satellite error bars represent the first and third quartiles around the median in a 50 km radius around each deployment (each pair of floats) coordinates.
+In situ Chla represent the Chla median in the upper 20 m of the water column with the upper and lower incertitude in the correction factor of @Roesler2017-me.
+
+# fetch data
+npq_data <- vroom('/home/flo/dox/ArgoShine/npq_corrected_data.csv')
+
+# clean and smoothed chla data
+d <- npq_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> dplyr::mutate(date = as_date(juld))
+d <- d |> group_by(date, wmo, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
+d <- d |> dplyr::mutate(smoothed_chla = smooth(mean_chla)/2.15, smoothed_chla_up = smooth(mean_chla)/(2.15-0.37), smoothed_chla_down = smooth(mean_chla)/(2.15+0.37)) # 2.15 = Roesler factor in the Labrador Sea
+
+d$wmo <- factor(d$wmo, levels = c(6904240,4903634,6904241,1902578))
+
+# compute chla integration in the MLD and in the 0-100 m zone
+d_chla <- d |> group_by(wmo, date) |> summarize(chla_0_100m = integrate(smoothed_chla, pres, from=0, to=100),
+                                       mld = unique(mld),
+                                       chla_0_mld = integrate(smoothed_chla, pres, from = 0, to=mld),
+                                       median_chla_0_20m = median(smoothed_chla[pres<=20], na.rm=T),
+                                       median_chla_0_20m_up = median(smoothed_chla_up[pres<=20], na.rm=T),
+                                       median_chla_0_20m_down = median(smoothed_chla_down[pres<=20], na.rm=T))
+#
+# ggplot(d_chla) + geom_point(aes(date, chla_0_100m)) + facet_wrap(~wmo, scales = 'free') + theme_bw() + labs(x = 'Month', y = 'Integrated Chla 0-100 m (CHLA UNITS)')
+#ggplot(d_chla) + geom_point(aes(date, chla_0_mld)) + facet_wrap(~wmo, scales = 'free')
+
+# add couple info
+d_chla <- d_chla |> dplyr::mutate(group = if_else(wmo %in% c(6904240,4903634), '6904240 - 4903634', '6904241 - 1902578'))
+
+#chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_with_clouds_sat.feather') |> select(time, lat, lon, chla = CHL)
+chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_without_clouds_sat.feather') |> select(time, lat, lon, chla = CHL)
+mld_sat <- arrow::read_feather('DATA/SATELLITE/mld_sat.feather') |> dplyr::mutate(longitude = longitude - 360) |> select(time, lat = latitude, lon = longitude, mld = mlotst)
+
+# CHLA
+chla_cover_deployment1 <- points_in_circle(chla_sat, lon_center = -49.2, lat_center = 58.9, radius = 100000) # 50 km around center
+chla_cover_deployment1 <- chla_cover_deployment1 |> dplyr::mutate(date = as_date(time), group = '6904240 - 4903634') |> dplyr::mutate(deploy = as_date('2022-05-22'))
+
+chla_cover_deployment2 <- points_in_circle(chla_sat, lon_center = -52.3, lat_center = 56.7, radius = 10000) # 50 km around center
+chla_cover_deployment2 <- chla_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-30'))
+
+# all deploy
+chla_cover <- rbind(chla_cover_deployment1, chla_cover_deployment2)
+
+chla_plot <- ggplot(chla_cover |> filter(date > '2022-03-01')) + geom_errorbar(aes(x = date, y = chla),
+                      stat = 'summary',
+                      fun.min = function(z) {quantile(z, 0.25)},
+                      fun.max = function(z) {quantile(z, 0.75)},
+                      fun = median, colour = 'black', width = .5) + theme_bw() +
+  labs(x = 'Month', y = TeX('Chla (mg m$^{-3}$)'))+ facet_wrap(~group) +
+  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
+  geom_pointrange(data = d_chla |> filter(date < '2022-08-01'), aes(x = date, ymin = median_chla_0_20m_down, ymax = median_chla_0_20m_up, y = median_chla_0_20m), colour = "#E31B23", size = .25) +
+   scale_x_date(labels = date_format("%m"), date_breaks = '1 month')
+
+# MLD
+mld_cover_deployment1 <- points_in_circle(mld_sat, lon_center = -49.2, lat_center = 58.9, radius = 100000) # 50 km around center
+mld_cover_deployment1 <- mld_cover_deployment1 |> dplyr::mutate(date = as_date(time), group = '6904240 - 4903634') |> dplyr::mutate(deploy = as_date('2022-05-22'))
+
+mld_cover_deployment2 <- points_in_circle(mld_sat, lon_center = -52.3, lat_center = 56.7, radius = 100000) # 50 km around center
+mld_cover_deployment2 <- mld_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-30'))
+
+mld_cover <- rbind(mld_cover_deployment1, mld_cover_deployment2)
+
+mld_plot <- ggplot(mld_cover |> filter(date > '2022-03-01')) + geom_errorbar(aes(x = date, y = mld),
+                      stat = 'summary',
+                      fun.min = function(z) {quantile(z, 0.25)},
+                      fun.max = function(z) {quantile(z, 0.75)},
+                      fun = median, colour = 'black', width = 5) +
+  labs(x = 'Month', y = 'MLD (m)') + facet_wrap(~group) + theme_bw() +
+  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
+  geom_point(data = d_chla |> filter(date < '2022-08-01') |> dplyr::mutate(week = week(date), DOY = yday(date)) |> group_by(week, group) |> summarize(median_mld = median(mld), mean_date = as_date(mean(DOY), origin = '2022-01-01')), aes(mean_date, median_mld), colour = "#E31B23") +
+   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + scale_y_continuous(trans = 'log10')
+
+chla_plot / mld_plot
+```
+
+The Chla satellite pattern for the pair 6904240-4903634 typically describes the phenology of the spring bloom in the northern part of the LS \[@Lacour2015-da\].
+This is not surprising given that this pair was deployed one degree southward.
+The other pair was deployed in the bioregion where the phytoplankton biomass slowly develops during May to reach its maximum in June \[@Lacour2015-da\].
+The Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD and that it started again in mid-June when the water column was well stratified.
+However, the Chla satellite concentration for that pair was three times lower than for the first one.
+Hence, it is possible that the bloom intensity was not sufficient to produce clear particle export signals, at least for BGC-Argo floats in their 10-day vertical sampling configuration.
+
+```{r, chla corrected for NPQ and MLD}
+# #|: label: fig-chla-bgc
+# #|: fig-cap : Chla concentrations and MLD measured by all floats in the upper 250 m for the period following the deployment.
+# 
+# 
+# # plot it
+# ggplot(d |> filter(date < '2022-08-01')) + geom_point(aes(x = date, y = pres, colour = log10(smoothed_chla))) +
+#   scale_y_reverse(limits = c(250,0)) +
+#   #scale_y_continuous(trans = 'rev_sqrt') +
+#   #scale_colour_viridis(option = 'A') +
+#   scale_colour_cmocean(name = 'algae', oob = scales::squish, limits = c(-1,1), na.value = 'grey') +
+#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+#   labs(x = 'Time', y = 'Depth (m)', colour = TeX('Chla (mg m$^{-3}$)')) + facet_wrap(~wmo) + theme_bw() +
+#   geom_line(aes(x = date, y = mld), colour = 'black', size = 1)
+```
+
+#### Just because it needs to go somewhere
+
+```{r, mld}
+
+ggplot(d) + geom_line(aes(x = date, y = mld), colour = 'black') +
+  facet_wrap(~wmo) + labs(x = 'Month', y = 'MLD (m)') + scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month')
+
+```
 
@@
 
-```{r, compute guidi fluxes at parking depth}
+```{r, compute guidi fluxes at parking depth, fig.height=6, fig.width=6}
 #| label: fig-comparison-ost-uvp
@@
 
-ici je compare finalement un système ultra productif (iversen) et un système pas produtif (lionel) -> on se situerait entre les deux.
+ici je compare finalement un système ultra productif (iversen) et un système pas produtif (lionel) -\> on se situerait entre les deux.
 Après je peux fournir els valeurs que j'ai trouvées avec le fminsearch avec la fonctiono optimize depython (21.98 et 1.18.
@@
 -   phénologie du bloom printanier
--   qinking speed -> trouver un truc à dire et estimer l'incertitude (en lien avec le graphique ou le DZ augmente)   + expliquer la méthode vite fait (Léo Lacour, personal communication)
+-   qinking speed -\> trouver un truc à dire et estimer l'incertitude (en lien avec le graphique ou le DZ augmente) + expliquer la méthode vite fait (Léo Lacour, personal communication)
 -   on se trouve dans un sytème producitve entre guidi 2008 et iversen (très productif)
-
-### Environmental conditions prior to bloom
-
--   aller chercher les produits satellitaires pour montrer que l'on a bien raté le bloom ou que les conditions du bloom n'étaient pas réunies
-
-```{r, chla corrected for NPQ and MLD}
-
-# fetch data
-npq_data <- vroom('DATA/ARGO/npq_corrected_data.csv')
-
-# clean and smoothed chla data
-d <- npq_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> dplyr::mutate(date = as_date(juld))
-d <- d |> group_by(date, wmo, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
-d <- d |> dplyr::mutate(smoothed_chla = smooth(mean_chla)/2.15, smoothed_chla_up = smooth(mean_chla)/(2.15-0.37), smoothed_chla_down = smooth(mean_chla)/(2.15+0.37)) # 2.15 = Roesler factor in the Labrador Sea
-
-#d$wmo <- factor(d$wmo, levels = c(6904240,4903634,6904241,1902578))
-
-# plot it
-ggplot(d) + geom_point(aes(x = date, y = pres, colour = log10(smoothed_chla))) +
-  scale_y_reverse(limits = c(200,0)) +
-  #scale_colour_viridis(option = 'A') +
-  scale_colour_cmocean(name = 'algae', oob = scales::squish, limits = c(-1,1), na.value = 'grey') +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  labs(x = 'Time', y = 'NPQ-corrected Chla') + facet_wrap(~wmo) + theme_bw() +
-  geom_line(aes(x = date, y = mld), colour = 'black', size = 1)
-
-# compute chla integration in the MLD and in the 0-100 m zone
-d_chla <- d |> group_by(wmo, date) |> summarize(chla_0_100m = integrate(smoothed_chla, pres, from=0, to=100),
-                                       mld = unique(mld),
-                                       chla_0_mld = integrate(smoothed_chla, pres, from = 0, to=mld),
-                                       median_chla_0_20m = median(smoothed_chla[pres<=20], na.rm=T),
-                                       median_chla_0_20m_up = median(smoothed_chla_up[pres<=20], na.rm=T),
-                                       median_chla_0_20m_down = median(smoothed_chla_down[pres<=20], na.rm=T))
-# 
-# ggplot(d_chla) + geom_point(aes(date, chla_0_100m)) + facet_wrap(~wmo, scales = 'free') + theme_bw() + labs(x = 'Month', y = 'Integrated Chla 0-100 m (CHLA UNITS)')
-#ggplot(d_chla) + geom_point(aes(date, chla_0_mld)) + facet_wrap(~wmo, scales = 'free')
-
-# add couple info
-d_chla <- d_chla |> dplyr::mutate(group = if_else(wmo %in% c(6904240,4903634), '6904240 - 4903634', '6904241 - 1902578'))
-
-```
-
-```{r, chla and MLD from satellite}
-
-#chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_with_clouds_sat.feather') |> select(time, lat, lon, chla = CHL)
-chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_without_clouds_sat.feather') |> select(time, lat, lon, chla = CHL)
-mld_sat <- arrow::read_feather('DATA/SATELLITE/mld_sat.feather') |> dplyr::mutate(longitude = longitude - 360) |> select(time, lat = latitude, lon = longitude, mld = mlotst)
-
-# CHLA
-chla_cover_deployment1 <- points_in_circle(chla_sat, lon_center = -49.2, lat_center = 58.9, radius = 100000) # 50 km around center
-chla_cover_deployment1 <- chla_cover_deployment1 |> dplyr::mutate(date = as_date(time), group = '6904240 - 4903634') |> dplyr::mutate(deploy = as_date('2022-05-22'))
-
-chla_cover_deployment2 <- points_in_circle(chla_sat, lon_center = -52.3, lat_center = 56.7, radius = 10000) # 50 km around center
-chla_cover_deployment2 <- chla_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-30'))
-
-# all deploy
-chla_cover <- rbind(chla_cover_deployment1, chla_cover_deployment2)
-
-chla_plot <- ggplot(chla_cover |> filter(date > '2022-04-01')) + geom_errorbar(aes(x = date, y = chla),
-                      stat = 'summary',
-                      fun.min = function(z) {quantile(z, 0.25)},
-                      fun.max = function(z) {quantile(z, 0.75)},
-                      fun = median, colour = 'black', width = .5) + theme_bw() +
-  labs(x = 'Time', y = 'Chla')+ facet_wrap(~group) +
-  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
-  geom_pointrange(data = d_chla |> filter(date < '2022-08-01'), aes(x = date, ymin = median_chla_0_20m_down, ymax = median_chla_0_20m_up, y = median_chla_0_20m), colour = "red", size = .25)
-
-# MLD
-mld_cover_deployment1 <- points_in_circle(mld_sat, lon_center = -49.2, lat_center = 58.9, radius = 100000) # 50 km around center
-mld_cover_deployment1 <- mld_cover_deployment1 |> dplyr::mutate(date = as_date(time), group = '6904240 - 4903634') |> dplyr::mutate(deploy = as_date('2022-05-22'))
-
-mld_cover_deployment2 <- points_in_circle(mld_sat, lon_center = -52.3, lat_center = 56.7, radius = 100000) # 50 km around center
-mld_cover_deployment2 <- mld_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-30'))
-
-mld_cover <- rbind(mld_cover_deployment1, mld_cover_deployment2)
-
-mld_plot <- ggplot(mld_cover |> filter(date > '2022-04-01')) + geom_errorbar(aes(x = date, y = mld),
-                      stat = 'summary',
-                      fun.min = function(z) {quantile(z, 0.25)},
-                      fun.max = function(z) {quantile(z, 0.75)},
-                      fun = median, colour = 'black', width = 5) +
-  labs(x = 'Time', y = 'MLD (m)') + facet_wrap(~group) + theme_bw() +
-  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
-  geom_point(data = d_chla |> filter(date < '2022-08-01') |> dplyr::mutate(week = week(date), DOY = yday(date)) |> group_by(week, group) |> summarize(median_mld = median(mld), mean_date = as_date(mean(DOY), origin = '2022-01-01')), aes(mean_date, median_mld), colour = "red")
-
-chla_plot / mld_plot
-
-```
-
-with a radius of 100 km around both deployment positions
 
@@
 
-```{r, vertical plots + sinking speed}
+```{r, vertical plots + sinking speed bis, fig.height=8, fig.width=8}
 
@@
 
--   Pour le drift à 200, 500 et 1000 m -> interpète ce graph -> va checker les masses d'eau de la region...
-attention on est en PARKING DEPTH -> mettre la couleur de la parking depth dans le truc, ça permettrait de voir si on peut comparer les processus (style ratio 500/1000m) dans les couches qui sont dans la meme masse d'eau.
+-   Pour le drift à 200, 500 et 1000 m -\> interpète ce graph -\> va checker les masses d'eau de la region...
+attention on est en PARKING DEPTH -\> mettre la couleur de la parking depth dans le truc, ça permettrait de voir si on peut comparer les processus (style ratio 500/1000m) dans les couches qui sont dans la meme masse d'eau.
 

### 2023-03-07 intro.qmd
update

@@
 | Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
-|-----------------|-------------------|-----------------|---------------------|
+|------------------------|----------------------------------|---------------------------|--------------------------------------|
 | Gravitational          | export depth - sediment          | None                      | N/A (142)                            |

### 2023-03-07 observingCarbonCycle.qmd
update

@@
 When we compare the $e-ratio$ maps (@fig-export-ratios) or the carbon export estimations of only six biogeochemical models (see @sec-biogeochem-models), we already observe significant differences.
-On the one hand, we see an increase in model complexity from empirical models to data-assimilated inverse models thanks to the increase in the quantity and quality of observations as well as well as a deeper understanding of the BCP (@sec-a-new-paradigm-for-the-bcp).
+On the one hand, we see an increase in model complexity from empirical models to data-assimilated inverse models thanks to the increase in the quantity and quality of observations as well as a deeper understanding of the BCP (@sec-a-new-paradigm-for-the-bcp).
 On the other hand, each degree of complexity inevitably accumulates more uncertainty due to the increasing number of processing and modeling steps (@fig-bgc-models).

### 2023-03-09 article3.qmd
update

@@
 
-basemap(limits = c(-55,-40,55,65), bathymetry = T, lat.interval = 2.5, bathy.style = 'poly_blues',
-        grid.col = 'black') +
+p <- basemap(limits = c(-55,-40,55,65), bathymetry = T, lat.interval = 2.5, bathy.style = 'poly_blues')
+
+# Make the graticules:
+lims <- attributes(p)$limits 
+graticule <- sf::st_graticule(
+  c(lims[1], lims[3], lims[2], lims[4]), 
+  crs = attributes(p)$proj,
+  lon = attributes(p)$map.grid$lon.breaks, 
+  lat = attributes(p)$map.grid$lat.breaks
+)
+
+p <- p + 
+  geom_sf(data = graticule, color = "grey", size = LS(1)) + # graticules
+  coord_sf(xlim = lims[1:2], ylim = lims[3:4], # redefine limits
+           crs = attributes(p)$proj)
+
+p +
   geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), linewidth = 1) + 
@@
   guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
-                                                  shape = c(NA, NA, NA, NA)))) #+
-  # basemap(limits = c(-55,-40,55,65), bathymetry = F, lat.interval = 2.5,
-  #       grid.col = 'black')
-
+                                                  shape = c(NA, NA, NA, NA)))) +
+  annotation_scale(location = "br") + 
+  annotation_north_arrow(location = "tr", which_north = "true") 
 ```
@@
 
-The Chla satellite pattern for the pair 6904240-4903634 typically describes the phenology of the spring bloom in the northern part of the LS \[@Lacour2015-da\].
+The Chla satellite pattern for the pair 6904240-4903634 typically describes the phenology of the spring bloom in the northern part of the LS [@Lacour2015-da].
 This is not surprising given that this pair was deployed one degree southward.
-The other pair was deployed in the bioregion where the phytoplankton biomass slowly develops during May to reach its maximum in June \[@Lacour2015-da\].
+The other pair was deployed in the bioregion where the phytoplankton biomass slowly develops during May to reach its maximum in June [@Lacour2015-da].
 The Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD and that it started again in mid-June when the water column was well stratified.

### 2023-03-09 uvp6_part.qmd
update

@@
 Usually, images are brought back to the lab and classified with CNNs which, in comparison with features-based classifiers, that is machine learning algorithms using a set of features extracted from the image (e.g.
-area, eccentricity, grey levels) have outperformed the latter for plankton recognition [@Ellen2019-nh].
+area, eccentricity, gray levels) have outperformed the latter for plankton recognition [@Ellen2019-nh].
 Though, it comes at a cost.
@@
 The UVP6 is the miniaturized version of the UVP5 [@Picheral2021-sx].
-It is smaller, it uses less energy and it has been specifically developed to be mounted on AUVs such as BGC-Argo floats with the aim of providing data of similar quality than from the UVP5 (@fig-uvp-pipeline).
+It is smaller, it uses less energy and it has been specifically developed to be mounted on AUVs such as BGC-Argo floats with the aim of providing data of similar quality than from the UVP5.
 There are 2 versions: the LP (low power) and the HF (high frequency).
@@
 
-![Scheme of the UVP6 pipeline with the processing done in the lab to create the ER model and the processing time (in ms) of the main steps when the UVP6 is deployed on board an AUV.
-kpx means kilo pixel.](image/uvp6/fig_pipeline_UVP6.jpg){#fig-uvp-pipeline}
-
 UVP5 Standard Definition (SD) profiles have been collected since 2008 at a global scale (@fig-map-of-profiles) whereas the UVP6 has been deployed very recently in specific locations and thus lacks spatial coverage to be fully representative.
 For this reason, UVP5 data will be used to create a preliminary ER model and explore ways of improvement.
-Ultimately, we will train the ER model for the UVP6 by applying observations made on the preliminary UVP5 model.
+Ultimately, we will train the ER model for the UVP6 by applying the lessons learnt from the preliminary UVP5 model.
 
@@
 #| label: tbl-classification-report-uvp5-xgboost-cnn
-#| tbl-cap: Precision, recall and f1-score for each class and globally (lightgray background) in the UVP5 model (XGBoost and CNN). avg average. avg stands for average.
+#| tbl-cap: Precision, recall and f1-score for each class and globally (lightgray background) in the UVP5 model (XGBoost and CNN). The abbreviation avg stands for average.
 
@@
 $k$) before the model overfits.
-For this, we searched for the minimum value of the multi class log loss(also known as logistic loss or cross-entropy loss) evaluated at each boosting round.
+For this, we searched for the minimum value of the multi class log loss (also known as logistic loss or cross-entropy loss) evaluated at each boosting round.
 Using that optimal $k$ value, we then trained the ER model with the entire training set.
@@
 It is also very frustrating to see that one class (largely) dominates when you are interested in the small ones and that some of them are so scarce that you need to merge them with their siblings (at the very best) on the phylogenetic tree.
-All in all, we found - after multiple trials and errors - 11 classes fulfilling a first compromise between classification performances and ecological importance: 7 biological classes (*Acantharea*, *Aulacanthidae*, *Aulosphaeridae*, *Copepoda*, *Trichodesmium*, darksphere and other\<living) and 4 non-living classes (detritus, bubble, fiber and artefact).
+All in all, we found - after multiple trials and errors - 11 classes fulfilling a first compromise between classification performances and ecological importance: 7 biological classes (*Acantharia*, *Aulacanthidae*, *Aulosphaeridae*, *Copepoda*, *Trichodesmium*, darksphere and other\<living) and 4 non-living classes (detritus, bubble, fiber and artefact).
 
@@
 label imbalance not taken into account) precision and recall of, respectively, 64% and 69% and a weighted living precision and recall of, respectively, 60% and 65%.
-CV found an optimal number of 850 boosting rounds before over-fitting and the expected loading time is 382 ms.
+The CV found an optimal number of 850 boosting rounds before over-fitting and the expected loading time is 382 ms.
 
@@
 
-Quantification is an interesting concept that focuses on the sample instead of the individual.
+Quantification is a concept that focuses on the sample instead of the individual.
 It was already shown in @Solow2001-jc, using plankton images from the Video Plankton Recorder, that the estimation of taxonomic abundances could be improved by computing the confusion matrix $\pmb{P}$ where $p_{i,j}$ is the probability that a taxon $i$ is classified as taxon $j$.
@@
 Increasing the number of detritus classes was also considered in the hope of clustering them in the features space.
-Detritus display a variety of size, shape and grey levels so it must be possible to cluster them and therefore improve the classification score.
+Detritus display a variety of size, shape and gray levels so it must be possible to cluster them and therefore improve the classification score.
 For this, we applied a classic K-Means algorithm [@Hartigan1979-oz] on detritus cleaned from likely outliers using an Isolation Forest Algorithm [@Liu2012-xz] and we executed a grid search as described earlier.
@@
 <!-- ![Confusion matrix (normalized on true labels) of the a UVP5 XGBoost model with 5 classes of detritus.](image/uvp6/Muvpec_367314a7_confusion_matrix.jpeg){#fig-confusion-matrix-uvp6-ndetritus} -->
+
 ```{r, confusion matrix N detritus}
@@
 ```
+
 #### Application to the UVP6 data set
@@
 Note that the CV initially found an optimal number of boosting rounds of 555 for a loading time of 101 ms).
-However, we observed that the log loss curve had already reached a plateau hence we were able to decrease the size of the model without losing much (\< 1%) in biological precision or recall, decreasing a bit more the energy consumption of the UVP6 on board AUVs.
+However, we observed that the log loss curve had already reached a plateau hence we were able to decrease the size of the model without losing much (\< 1%) in biological precision or recall, decreasing a bit more the energy consumption of the UVP6.
 
@@
 
-```{r, uvp6 BGC application}
+```{r, uvp6 BGC application, fig.width=6, fig.height=5}
 #| label: fig-bgc-application-uvp6
@@
 
+![Scheme of the UVP6 pipeline with the processing done in the lab to create the ER model and the processing time (in ms) of the main steps when the UVP6 is deployed on board an AUV.
+kpx means kilo pixel.](image/uvp6/fig_pipeline_UVP6.jpg){#fig-uvp-pipeline}
+
 ### Conclusion
@@
 The objective of this paper was to develop a zooplankton classification model under strict energy constraints for the UVP6 to be deployed on BGC-Argo floats.
-It was clear at the beginning that the use of powerful energy consuming CNNs had to be discarded and that we needed to build a satisfactory features-based model instead.
-To this aim, we firstly used XGBoost on UVP5 data, a more spatially and temporally representative data set than the recent UVP6 data set to create the data processing pipeline (classes aggregation, hyperparameter tuning, model training) and investigate ways to improve classification scores (quantification, detritus subgroups, label weights).
+It was clear at the beginning that the use of powerful energy consuming CNNs had to be disregarded and that we needed to build a satisfactory features-based model instead.
+To this aim, we firstly used XGBoost on UVP5 data, a more spatially and temporally representative data set than the recent UVP6 data set to create the data processing pipeline (classes aggregation, hyperparameter tuning, model training, see also @fig-uvp-pipeline) and investigate ways to improve classification scores (quantification, detritus subgroups, label weights).
 In order to see how the model could be improved, we compared the XGBoost model with results from a CNN not limited by energy.

### 2023-03-10 article3.qmd
update

@@
 ## Introduction
+
+The biological carbon pump (BCP) refers to a series of processes that transport biogenic carbon produced in sunlit surface waters to the ocean interior [@Claustre2021-tx].
+It is composed of three main pathways [@Claustre2021-tx; @Boyd2019-ez; @Siegel2022-wd]: the biological gravitational pump that is the gravitational settling of particulate organic carbon (POC); the physically mediated particle injection pump that represents physical mechanisms that transport dissolved organic carbon (DOC) to depth via mixing processes; and the biologically mediated particle injection pump that characterizes the transport of POC (and DOC) by diel and seasonal migration of animals in the water column.
+Historically, the BCP was mainly studied with sediment traps [@Honjo2008-kn] and Thorium-234 radionuclide disequilibria [@Buesseler2007-ql] focused on the gravitational pump hence the contribution of DOC and migrating animals to the BCP was likely under-estimated.
+However, the advent of autonomous underwater vehicles (AUVs) equipped with a variety of physical and biogeochemical sensors have demonstrated their ability to better understand the latter [e.g.
+@Omand2015-mf; @DallOlmo2016-ex; @Llort2018-pa; @Lacour2019-id; @Haentjens2020-dc; @Stukel2017-lc].
+Still, we are missing processes to close the carbon budget [@Giering2014-tj; @Boyd2019-ez] and unkowns remain in the under-sampled mesopelagic (i.e.
+twilight) zone [@Martin2020-wa].
+In order to fill the observational gaps, the Biogeochemical-Argo (BGC-Argo) network aims to develop an operational global array of 1000 profiling floats measuring six core variables from the surface to 2000 m [@Claustre2020-jo]: oxygen, nitrate, pH, chlorophyll a (Chla), suspended particles ($b_{bp}$) and downwelling irradiance.
+In addition to the six core variables, new sensors have been developped to target specific BCP-related questions [@Claustre2020-jo; @Claustre2021-tx; @Chai2020-uq].
+In this study, we will focus on BGC-Argo floats equipped with two new sensors: the Underwater Vision Profiler (UVP6) [@Picheral2021-sx] and a transmissometer used as an optical sediment trap (OST).
+
+The UVP6 is a low power miniaturized version of the UVP5 [@Picheral2010-to].
+It consists of an underwater camera that counts particles and images objects in, respectively, the 0.1 - 2.5 cm and 0.8 - 2.5 cm size range.
+The latest version of the UVP6 also embarks an embedded image classification algorithm that transmits the abundance of 20 classes in near-real time.
+The OST consists of a vertically mounted transmissometer whose upward-facing optical window intercepts sinking particles.
+The latter have already been used on AUVs [@Bishop2004-at; @Bishop2009-ka; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk].
+
+Two pairs of BGC-Argo floats equipped with both the UVP6 and an OST were deployed in the southern part of the Labrador Sea in May 2022 to investigate seasonal and bloom-related particle and carbon fluxes.
+The Labrador and the northern part the atlantic are charateterid by blablabla -\> lire le truc et faire la micro revie sur l'atlantique nord, son bloom et la mer du labrador.
+
+but it is the first time that BGC-Argo floats are deployed with both a UVP6 and an OST.
+
+In this work, we will
+
+-   Pour l'intro je dois lire briggs 2011-2020
+
+-   projet remocean Hervé
+
+-   miwing convection virginie thierry?
+
+-   atlantic nord, le bloom de l'atlantic nord
+
+-   léo lacour
+
+-   yashayeav
+
+-   hanetjens 2020
 
@@
 -   plotter la MLD sur les profils verticaux de
+-   from llort 2017: This is especially true in the Southern Ocean, where the circulation has been described as equivalent barotropic (Gille et al., 2007; Killworth & Hughes, 2002).
+An equivalent barotropic circulation basically implies that the strength of the current might change with depth (weaker as depth increases) but the orientation does not change.
+Therefore, Argo and BGCArgo can be interpreted as "quasi-Lagrangian" because floats follow well ocean mesoscale circulation lasting more than 1 week; yet they cannot follow shorter and smaller features such the ones associated to submesocale circulation
 
@@
 # add levels for some field
-all_floats <- all_floats |> dplyr::mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
-                                   park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
+all_floats <- all_floats |> dplyr::mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)), park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
 
@@
 argo_map_last <- argo_map |> filter(cycle == max(cycle))
-
-# # map positions of floats
-# world_data <- map_data("world")
-# w <- list(geom_polygon(aes(long, lat, group=group), data=world_data, fill="grey70"), coord_quickmap())
-# 
-# ggplot(argo_map) + w + geom_path(aes(lon, lat, colour = WMO), size = 1) +
-#   geom_point(aes(lon, lat, colour = WMO), size = 1.5) +
-#   scale_color_brewer(palette = 'Dark2') +
-#   theme_bw() +
-#   #scale_colour_manual(values = c('#a6cee3', '#33a02c', '#b2df8a', '#1f78b4')) +
-#   coord_cartesian(xlim = c(-55, -40), ylim = c(55,65)) + geom_point(data = argo_map_first, aes(lon, lat), shape = 17, size = 3) +
-#   geom_point(data = argo_map_last, aes(lon, lat, colour = WMO), size = 3) +
-#   geom_point(data = argo_map_last, aes(lon, lat, fill = WMO), size = 3, shape = 1) +
-#   labs(x = 'longitude', y = 'latitude') +
-#   #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
-#   guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
-#                                                   shape = c(NA, NA, NA, NA))))
 
@@
 Data from BGC-ARGO float WMO 6904240, drifting at 1000 m during its 21$^{st}$ cycle.
+
 clean_cp_data <- function(data, moving_median_order){
@@
 
-# ggplot(clean_data, aes(x = dates, y = cp)) + geom_point(size= 3) + geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour), size = 2) + 
-#     scale_colour_manual(values = c('#003366','#E31B23', '#FFC325')) + 
-#     scale_shape_manual(values = c(1,1,4)) +
-#     geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', size = 1) + theme_bw() + 
-#     labs(x = 'Time', y = 'Cp (1/m)') +
-#   geom_point(data = check, aes(x = dates, y = cp, colour = colour))
-
 ggplot(tmp2, aes(x = dates, y = cp, colour = colour)) + geom_point(size = 3) +
@@
   geom_point(data = check, aes(x = dates, y = cp), shape = 1, size = 3, colour = 'black')  +
-  #https://r-graph-gallery.com/239-custom-layout-legend-ggplot2.html
-  # theme(legend.position = c(0.02,0.99), legend.justification = c('left', 'top'), 
-  #       legend.title = element_blank())
     theme(legend.position = 'none')
@@
   check_colour <- unique(tmp2$colour)
-  # if(length(check_colour) == 3){ # there is at least a negative jump
-  #   index_neg_jump <- which(tmp2$colour == 'negative jump')
-  #   tmp2 <- tmp2[-c(index_neg_jump, index_neg_jump+1),]
-  #   tmp2 <- tmp2 |> dplyr::mutate(diff_jump = cp - lag(cp)) 
-  #   even_indexes <- seq(2,nrow(tmp2),2)
-  #   tmp2 <- tmp2[even_indexes,]
-  # }else if(length(check_colour) == 2){ # only positive jumps
-  #   tmp2 <- tmp2 |> dplyr::mutate(diff_jump = cp - lag(cp)) 
-  #   even_indexes <- seq(2,nrow(tmp2),2)
-  #   tmp2 <- tmp2[even_indexes,]
-  # }else{ # No jump
-  #   tmp2 <- NULL
-  # }
   if(length(check_colour) >= 2){ # there is a least a jump (positive or negative)
@@
 Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.
-
-up <- cflux |> ggplot(aes(x = date, y = small_flux)) +
-    geom_point(data = cflux, aes(x = date, y = small_flux, colour = WMO, shape = WMO)) +
-  geom_smooth(data = cflux, aes(x = date, y = small_flux), colour = 'black') +
+There is no data after January and February at, respectively, 200 m and 500 m because the drifting was stopped at those depths due to the increase of the MLD.
+
+# add black data to have the same x axis for all drifting depths (to show that we stopped the drifting at 200 and 500 m)
+blank_data <- tibble(max_time = c(NA, NA), min_time = c(NA,NA), small_flux = c(NA,NA), big_flux = c(NA,NA), park_depth = c('200 m', '500 m'),
+                     wmo = rep(4903634,2), cycle = c(NA,NA), drifting_time = c(NA,NA), WMO = rep(as.factor(4903634),2), date = rep('2023-03-05',2))
+
+tmp <- rbind(cflux, blank_data)
+
+up <- tmp |> ggplot(aes(x = date, y = small_flux)) +
+    geom_point(data = tmp, aes(x = date, y = small_flux, colour = WMO, shape = WMO)) +
+  geom_smooth(data = tmp, aes(x = date, y = small_flux), colour = 'black') +
   scale_color_brewer(palette = 'Dark2') + 
@@
 
-down <- cflux |> ggplot(aes(x = date, y = big_flux)) +
-  geom_point(data = cflux, aes(x = date, y = big_flux, colour = WMO, shape = WMO)) +
-  geom_smooth(data = cflux, aes(x = date, y = big_flux), colour = 'black') +
+down <- tmp |> ggplot(aes(x = date, y = big_flux)) +
+  geom_point(data = tmp, aes(x = date, y = big_flux, colour = WMO, shape = WMO)) +
+  geom_smooth(data = tmp, aes(x = date, y = big_flux), colour = 'black') +
   scale_color_brewer(palette = 'Dark2') + 
@@
 In addition, the formulation of such a model implies that $A$ and $b$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt].
-@eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2022-qj].
+@eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2023-qj].
 As a result, the direct comparison of carbon fluxes derived from both OST and UVP data with @eq-guidi-article3 should be focused on the trend rather than on absolute carbon flux values.
@@
                       stat = 'summary',
-                      fun.min = function(z) {quantile(z, 0.25)},
-                      fun.max = function(z) {quantile(z, 0.75)},
+                      #fun.min = function(z) {quantile(z, 0.25)},
+                      #fun.max = function(z) {quantile(z, 0.75)},
                       fun = median, colour = '#E31B23', width = 5) +
@@
                       stat = 'summary',
-                      fun.min = function(z) {quantile(z, 0.25)},
-                      fun.max = function(z) {quantile(z, 0.75)},
+                      #fun.min = function(z) {quantile(z, 0.25)},
+                      #fun.max = function(z) {quantile(z, 0.75)},
                       fun = median, colour = '#003366', width = 5) +
@@
                       fun = median, colour = '#003366') + 
-  scale_y_continuous(trans = 'log10') + theme_bw() + labs(x = 'Month', y = TeX('$F_{total}$ (mg C m$^{-2}$ day$^{-1}$)'))
+  scale_y_continuous(trans = 'log10') + theme_bw() + labs(x = 'Month', y = TeX('$F_{total}$ (mg C m$^{-2}$ day$^{-1}$)')) +
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month')
+
+```
+
+```{r, and on the vertical?}
 
@@
 
-# i <- 4903634
-# j <- '1000 m'
-# k <- 15
+# res_slope <- data.frame()
+# res_jumps <- data.frame()
+# for (i in wmo){
+#   max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
+#   for (j in park_depth){
+#     for (k in seq(1:max_cycle)){
 #       tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
-# 
-# data <- tmp
-
-res_slope <- data.frame()
-res_jumps <- data.frame()
-for (i in wmo){
-  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
-  for (j in park_depth){
-    for (k in seq(1:max_cycle)){
-      tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
-      if(nrow(tmp) == 0){
-        next
-      }else if(nrow(tmp) < 3){ # case where there is not enough data
-        next
-      }else{
-          output <- get_jump_info(tmp)
-          res_slope <- rbind(res_slope, output$slope_info)
-          res_jumps <- rbind(res_jumps, output$jump_info)
-      }
-    }
-  }
-}
-
+#       if(nrow(tmp) == 0){
+#         next
+#       }else if(nrow(tmp) < 3){ # case where there is not enough data
+#         next
+#       }else{
+#           output <- get_jump_info(tmp)
+#           res_slope <- rbind(res_slope, output$slope_info)
+#           res_jumps <- rbind(res_jumps, output$jump_info)
+#       }
+#     }
+#   }
+# }
+
+# write data because it's long
+# vroom_write(res_slope, file = '/home/flo/dox/THESIS/DATA/ARGO/res_slope.csv')
+# vroom_write(res_jumps, file = '/home/flo/dox/THESIS/DATA/ARGO/res_jumps.csv')
+res_slope <- vroom::vroom(file = '/home/flo/dox/THESIS/DATA/ARGO/res_slope.csv')
+res_jumps <- vroom::vroom(file = '/home/flo/dox/THESIS/DATA/ARGO/res_jumps.csv')
 ```
@@
 
-ggplot(slope_vs_biovolume) + geom_point(aes(x = log(slope), y = log(weighted_group_biovolume))) + facet_wrap(~park_depth) + labs(x = 'OST slope [UNITS]', y = 'UVP biovolume [UNITS]') #+ theme(axis.title.x = element_blank())
-#ggplot(test) + geom_point(aes(x = slope, y = weighted_group_biovolume)) + facet_wrap(~park_depth)
+# ggplot(slope_vs_biovolume) + geom_point(aes(x = log(slope), y = log(weighted_group_biovolume))) + facet_wrap(~park_depth) + labs(x = 'OST slope [UNITS]', y = 'UVP biovolume [UNITS]') #+ theme(axis.title.x = element_blank())
+# #ggplot(test) + geom_point(aes(x = slope, y = weighted_group_biovolume)) + facet_wrap(~park_depth)
+ggplot(slope_vs_biovolume |> filter(slope > 0.00001)) + geom_point(aes(x = slope*0.25, y = weighted_group_biovolume)) + facet_wrap(~park_depth, scales = 'free') + labs(x = TeX('$ATN~slope~(day^{-1})$'), y = TeX('$UVP~biovolume~(mm^{3})$')) +
+  scale_x_continuous(trans = 'log10') + scale_y_continuous(trans = 'log10') + theme_bw()
 
@@
 #ggplot(test_jump) + geom_point(aes(diff_jump, biovolume))
-ggplot(diff_jump_vs_biovolume) + geom_point(aes(log(diff_jump), log(biovolume))) + facet_wrap(~park_depth) + labs(x = 'OST jump difference [UNITS]', y = 'UVP biovolume [UNITS]')
+# ggplot(diff_jump_vs_biovolume) + geom_point(aes(log(diff_jump), log(biovolume))) + facet_wrap(~park_depth) + labs(x = 'OST jump difference [UNITS]', y = 'UVP biovolume [UNITS]') +
+#   scale_x_continuous(breaks = c(-6,-4,-2,0))
+ggplot(diff_jump_vs_biovolume) + geom_point(aes(x = diff_jump*0.25, y = biovolume)) + facet_wrap(~park_depth, scales = 'free') + labs(x = TeX('$ATN~jump~difference$'), y = TeX('$UVP~biovolume~(mm^{3})$')) +
+  scale_x_continuous(trans = 'log10') + scale_y_continuous(trans = 'log10') + theme_bw()
 
@@
 
+```{r, vertical plots + sinking speed bis, fig.height=8, fig.width=8}
+# 
+# particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld), doy = yday(juld), year = year(juld))
+# 
+# particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) 
+# #particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
+# 
+# # remove buggy data from float 4903634 (negative depth for some reason)
+# particles_bis <- particles_bis |> filter(depth >= 0)
+# 
+# # rel_conc
+# particles_bis <- particles_bis |> dplyr::group_by(wmo, size) |> dplyr::mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
+# 
+# up_part <-  particles_bis |> filter(depth < 2000, size < 512, doy < 220, year == 2022) |>
+#   #group_by(size) |>
+#   ggplot() + facet_grid(wmo~size) + theme_bw() +
+#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+#   scale_y_reverse() +
+#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
+#   geom_pointrange(data = sinking_fit |> filter(size < 512, DOY < 220, standard_dev <= 30), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'black')
+# 
+# up_part
+# 
+# # down_part <- particles_bis |> filter(depth < 2000, size > 512) |>
+# #   #group_by(size) |>
+# #   ggplot() + facet_grid(wmo~size) + theme_bw() +
+# #   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+# #   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+# #   scale_y_reverse() +
+# #   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+# #   labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
+# #   geom_pointrange(data = sinking_fit |> filter(size > 512), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'white')
+# 
+# down_part <-  particles_bis |> filter(depth < 2000, size > 512, doy < 220, year == 2022) |>
+#   #group_by(size) |>
+#   ggplot() + facet_grid(wmo~size) + theme_bw() +
+#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+#   scale_y_reverse() +
+#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
+#   geom_pointrange(data = sinking_fit |> filter(size > 512, DOY < 220, standard_dev <= 30), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'black')
+# 
+# down_part
+
+```
+
 ```{r, sinking speed plot}
 
-# NOTE : pas besoin d'updater ici vu que l'on regarde par rapport au 'bloom' de fin mai/juin
-sinking_fit <- vroom('DATA/ARGO/gaussian_fit_sinking_speed_full_before_06032023.csv') |> dplyr::mutate(size = size_class |> str_remove('NP_Size_') |> as.numeric(), date = as_date(date))
-
-estimate_sinking_speed <- function(siz, data, threshold){
-
-  # above 500 m
-  tmp <- data |> filter(size == siz) |> filter(depth <= threshold)
-  model <- lm(formula = depth ~ time_max_abundance, data = tmp)
-  tmp$sinking_speed_above_threshold <- model$coefficients[2] # equivalent to slope of the fit
-  tmp2 <- tmp
-
-  # below 500 m
-  tmp <- data |> filter(size == siz) |> filter(depth > threshold)
-  model <- lm(formula = depth ~ time_max_abundance, data = tmp)
-  tmp2$sinking_speed_below_threshold <- model$coefficients[2] # equivalent to slope of the fit
-
-  return(tmp2)
-}
-
-# sinking speed for each float
-sinking_6904240 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 6904240), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
-sinking_6904241 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 6904241), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
-sinking_1902578 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 1902578), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
-sinking_4903634 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 4903634), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
-all_sinking_speed <- rbind(sinking_1902578, sinking_4903634, sinking_6904240, sinking_6904241)
-
-```
-
-```{r, vertical plots + sinking speed bis, fig.height=8, fig.width=8}
-
-particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld))
-
-particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) 
-#particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
-
-# remove buggy data from float 4903634 (negative depth for some reason)
-particles_bis <- particles_bis |> filter(depth >= 0)
-
-# rel_conc
-particles_bis <- particles_bis |> dplyr::group_by(wmo, size) |> dplyr::mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
-
-up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
+# # NOTE : pas besoin d'updater ici vu que l'on regarde par rapport au 'bloom' de fin mai/juin
+# sinking_fit <- vroom('DATA/ARGO/gaussian_fit_sinking_speed_full_before_06032023.csv') |> dplyr::mutate(size = size_class |> str_remove('NP_Size_') |> as.numeric(), date = as_date(date))
+# 
+# estimate_sinking_speed <- function(siz, data, threshold){
+#   
+#   # above 500 m
+#   tmp <- data |> filter(size == siz) |> filter(depth <= threshold)
+#   model <- lm(formula = depth ~ time_max_abundance, data = tmp)
+#   tmp$sinking_speed_above_threshold <- model$coefficients[2] # equivalent to slope of the fit
+#   tmp2 <- tmp
+#   
+#   # below 500 m
+#   tmp <- data |> filter(size == siz) |> filter(depth > threshold)
+#   model <- lm(formula = depth ~ time_max_abundance, data = tmp)
+#   tmp2$sinking_speed_below_threshold <- model$coefficients[2] # equivalent to slope of the fit
+#   
+#   return(tmp2)
+# }
+# 
+# # sinking speed for each float
+# sinking_6904240 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 6904240), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
+# sinking_6904241 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 6904241), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
+# sinking_1902578 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 1902578), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
+# sinking_4903634 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 4903634), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
+# all_sinking_speed <- rbind(sinking_1902578, sinking_4903634, sinking_6904240, sinking_6904241)
+
+```
+
+```{r, sinking speed values}
+
+# gt(all_sinking_speed)
+# 
+# all_sinking_speed |> ggplot() + geom_point(aes(y = sinking_speed_below_threshold, x = size))
+
+```
+
+```{r, check gaussian profiles}
+
+# # depth layers for Gaussian fit
+# depthL <- c(seq(40,200,20), c(250,300,400,500,600,800,1000,1250,1500,1750,2000))
+# 
+# # Gaussian formula #> but z is not the depth but the TIME
+# fgauss <- function(z, peak_conc, peak_doy, dz){
+#   peak_conc*exp(-((z-peak_doy)**2)/(dz**2))
+# }
+# 
+# return_fitted <- function(d){
+#   
+#   #timeframe near the bloom so to say
+#   ztime <- seq(160,240,1)
+#   # gaussian fit
+#   tib <- map_dfr(1:nrow(d), function(i){tibble(ztime = ztime, fgauss = fgauss(ztime, d$max_abundance[i], d$time_max_abundance[i], d$standard_dev[i]), depth = d$depth[i])})
+#   return(tib)
+# }
+# 
+# # test
+# tmp <- sinking_fit |> filter(wmo == 6904240, size == 645)
+# 
+# test <- return_fitted(tmp)
+# 
+# ggplot(test, aes(x = ztime, y = fgauss, colour = factor(depth))) + geom_point() + 
+#   scale_color_viridis_d() +
+#   theme_bw() 
+#
+
+```
+
+### TS diagram
+
+-   Pour le drift à 200, 500 et 1000 m -\> interpète ce graph -\> va checker les masses d'eau de la region...
+attention on est en PARKING DEPTH -\> mettre la couleur de la parking depth dans le truc, ça permettrait de voir si on peut comparer les processus (style ratio 500/1000m) dans les couches qui sont dans la meme masse d'eau.
+
+```{r, TS diagram}
+#ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26, park_depth != '200 m') |> dplyr::mutate(date = as_date(date))
+# ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
+#   dplyr::mutate(date = as_date(date), colour = if_else(park_depth == '200 m', '#E31B23', if_else(park_depth == '500 m', '#FFC325', 'black')))
+
+ts_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
+  dplyr::mutate(date = as_date(date))
+
+ts_data <- ts_data |> mutate(spiciness = swSpice(ts_data$psal, ts_data$temp, ts_data$depth, ts_data$lon, ts_data$lat))
+
+# # https://dankelley.github.io/oce/reference/plotTS.html
+# # https://dankelley.github.io/oce/reference/argo-class.html
+# test <- as.argo(time = ts_data$date, longitude = ts_data$lon, latitude = ts_data$lat, salinity = ts_data$psal, temperature = ts_data$temp, pressure = ts_data$depth,
+#         id = ts_data$wmo) 
+# 
+# plotTS(test, inSitu = T, type = 'p', xlab = 'Salinity [PSU]')
+# 
+# 
+# ggplot(ts_data) + geom_point(aes(x = spiciness, y = depth)) + scale_y_reverse() + facet_wrap(~wmo)
+
+a <- ggplot(ts_data |> filter(depth <= 1000)) + geom_point(aes(x=date, y=depth, colour=spiciness), size = 3, shape = 15) +
+  scale_y_reverse() +
+  #scale_color_viridis() +
+  scale_color_cmocean(name = 'balance', limits = c(-0.25,0.25), oob = scales::squish, na.value = 'grey') +
+  theme_bw() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = TeX('Spiciness (kg m$^{-3}$)'), x = 'Month', y = 'Depth (m)') +
+  facet_wrap(~wmo, ncol = 4)
+
+b <- particles_bis |> filter(depth <= 1000, size == 102) |>
   #group_by(size) |>
-  ggplot() + facet_grid(wmo~size) + theme_bw() +
-  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+  ggplot() + facet_wrap(~wmo, ncol = 4) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=2, shape=15) +
+  #scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+  scale_color_cmocean(name = 'thermal', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) +
   scale_y_reverse() +
   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
-  geom_pointrange(data = sinking_fit |> filter(size < 512), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'white')
-
-up_part
-
-down_part <- particles_bis |> filter(depth < 2000, size > 512) |>
-  #group_by(size) |>
-  ggplot() + facet_grid(wmo~size) + theme_bw() +
-  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
-  scale_y_reverse() +
-  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
-  geom_pointrange(data = sinking_fit |> filter(size > 512), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'white')
-
-down_part
-
-```
-
-```{r, sinking speed values}
-
-gt(all_sinking_speed)
-
-all_sinking_speed |> ggplot() + geom_point(aes(y = sinking_speed_below_threshold, x = size))
-
-```
-
-```{r, check gaussian profiles}
-
-# depth layers for Gaussian fit
-depthL <- c(seq(40,200,20), c(250,300,400,500,600,800,1000,1250,1500,1750,2000))
-
-# Gaussian formula #> but z is not the depth but the TIME
-fgauss <- function(z, peak_conc, peak_doy, dz){
-  peak_conc*exp(-((z-peak_doy)**2)/(dz**2))
-}
-
-return_fitted <- function(d){
-
-  #timeframe near the bloom so to say
-  ztime <- seq(160,240,1)
-  # gaussian fit
-  tib <- map_dfr(1:nrow(d), function(i){tibble(ztime = ztime, fgauss = fgauss(ztime, d$max_abundance[i], d$time_max_abundance[i], d$standard_dev[i]), depth = d$depth[i])})
-  return(tib)
-}
-
-# test
-tmp <- sinking_fit |> filter(wmo == 6904240, size == 645)
-
-test <- return_fitted(tmp)
-
-ggplot(test, aes(x = ztime, y = fgauss, colour = factor(depth))) + geom_point() + 
-  scale_color_viridis_d() +
-  theme_bw()
-
-```
-
-### TS diagram
-
--   Pour le drift à 200, 500 et 1000 m -\> interpète ce graph -\> va checker les masses d'eau de la region...
-attention on est en PARKING DEPTH -\> mettre la couleur de la parking depth dans le truc, ça permettrait de voir si on peut comparer les processus (style ratio 500/1000m) dans les couches qui sont dans la meme masse d'eau.
-
-```{r, TS diagram}
-#ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26, park_depth != '200 m') |> dplyr::mutate(date = as_date(date))
-ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
-  dplyr::mutate(date = as_date(date), colour = if_else(park_depth == '200 m', '#E31B23', if_else(park_depth == '500 m', '#FFC325', 'black')))
-
-ts_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
-  dplyr::mutate(date = as_date(date))
-
-# https://dankelley.github.io/oce/reference/plotTS.html
-# https://dankelley.github.io/oce/reference/argo-class.html
-test <- as.argo(time = ts_data$date, longitude = ts_data$lon, latitude = ts_data$lat, salinity = ts_data$psal, temperature = ts_data$temp, pressure = ts_data$depth,
-        id = ts_data$wmo)
-
-plotTS(test, inSitu = T, type = 'p', xlab = 'Salinity [PSU]')
-
-```
-
-### Processes at play
-
-J'ai enlevé le truc du spectre inversé à 500 et 1000 m car difficile de conclure -\> mais comme on est dans la meme masse d'eau, on pourrait le garder, why not
-
-```{r, monthly anomaly between drifting at 500 and 1000 m}
-#| label: fig-processes-at-play
-#| fig-cap: Ratio of monthly mean particles concentrations during drifting (1000 m over 500 m) for size classes below 1.63 mm.
-
-tmp <- uvp_data |> filter(park_depth != '200 m') |> group_by(wmo, size, month, park_depth) |>
-  summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
-  # compute concentration anomaly at 500 and 1000 m
-  dplyr::mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
-  select(wmo, size, month, diff_anomaly, div_anomaly) |> dplyr::mutate(month = month.abb[month]) |>
-  dplyr::mutate(size = as.character(size)) |>
-  dplyr::mutate(rev_div_anomaly = 1/div_anomaly) |>
-  #filter(is.finite(rev_div_anomaly)) |>
-  dplyr::mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1), div_anomaly = round(div_anomaly, 1)) |>
-  ungroup()
-
-# test en weekly
-# tmp <- uvp_data |> filter(park_depth != '200 m', month != 5) |> group_by(wmo, size, week, park_depth) |> 
-#   summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
-#   # compute concentration anomaly at 500 and 1000 m
-#   dplyr::mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
-#   select(wmo, size, week, diff_anomaly, div_anomaly) |> 
-#   dplyr::mutate(size = as.character(size)) |>
-#   dplyr::mutate(rev_div_anomaly = 1/div_anomaly) |> 
-#   #filter(is.finite(rev_div_anomaly)) |>
-#   dplyr::mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1)) |>
-#   ungroup()
-# 
-# create_data_for_heatmap <- function(float){
-#   y <- unique(tmp$size)[1:12] # above, data is way too low
-#   d <- tmp |> filter(wmo == float, size %in% y, week > 5)
-#   x <- unique(d$week)
-#   data <- expand.grid(X=x, Y=y)
-#   #data$Z <- tt$diff_anomaly
-#   data$Z <- d$rev_div_anomaly
-#   data$wmo <- float
-#   return(data)
-# }
-
-# d1 <- create_data_for_heatmap(6904241)
-# d2 <- create_data_for_heatmap(1902578)
-# d3 <- create_data_for_heatmap(4903634)
-# d4 <- create_data_for_heatmap(6904240)
-# 
-# dd <- rbind(d1, d2, d3, d4)
-
-# ggplot(dd, aes(X, Y, fill= Z)) + 
-#   geom_tile() + scale_fill_viridis() + facet_wrap(~wmo)
-
-create_data_for_heatmap <- function(float, df){
-  y <- unique(df$size)[1:12] # above, data is way too low
-  d <- df |> filter(wmo == float, size %in% y)
-  #x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')
-  x <- unique(d$month)
-  data <- expand.grid(X=x, Y=y)
-  #data$Z <- tt$diff_anomaly
-  data$Z <- d$rev_div_anomaly
-  #data$Z <- d$div_anomaly
-  data$wmo <- float
-  return(data)
-}
-
-# for some reason, there is some bug with the quarto compilation .. so I save the tibble .. and reload it 
-# d1 <- create_data_for_heatmap(6904241, tmp)
-# d2 <- create_data_for_heatmap(1902578, tmp)
-# d3 <- create_data_for_heatmap(4903634, tmp)
-# d4 <- create_data_for_heatmap(6904240, tmp)
-# dd <- rbind(d1, d2, d3, d4)
-# vroom_write(dd, 'heatmap_data.csv')
-
-# reload + redo some annoying stuff that were lost during the writing of the csv
-dd <- vroom('heatmap_data.csv') |> dplyr::mutate(Y = as.character(Y)) 
-dd$X <- factor(dd$X, levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan', 'Feb', 'Mar'))
-dd$Y <- factor(dd$Y, levels = c('102', '128', '161', '203', '256', '323', '406', '512', '645', '813', '1020', '1290'))
-dd$Y <- factor(dd$Y, levels = c('1290', '1020', '813', '645', '512', '406', '323', '256', '203', '161', '128', '102'))
-# dd$wmo <- factor(dd$wmo, levels = c(6904240,4903634,6904241,1902578))
-
-#dd <- dd |> dplyr::mutate(Z_binned = cut(Z, breaks = c(0,1,1.5,2.5,max(Z))))
-dd <- dd |> dplyr::mutate(Z_binned = cut(Z, breaks = c(0,.5,1,1.5,2,max(Z))))
-
-# ggplot(dd) +
-#   geom_tile(aes(X, Y, fill = Z), alpha = 0.75) + scale_fill_viridis() + theme_bw() + labs(x = 'Month', y = 'Size class') +
-#   facet_wrap(~wmo) +
-#   geom_text(data = dd, aes(X, Y, label = Z))
-
-ggplot(dd) + 
-  geom_tile(aes(X, Y, fill = Z_binned), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Ratio', x = 'Month', y = 'Size class') +
-  facet_wrap(~wmo) + scale_fill_viridis_d(option = 'A')
+  labs(colour = 'Particles (#/L)', x = 'Month', y = 'Depth (m)')
 

### 2023-03-10 intro.qmd
update

--- 2023-02-28 intro.qmd
@@
 Both values are used in the literature (e.g.
-10% in @Boyd2019-ez and @Iversen2022-qj and 35% in @Riebesell2009-qj, @Passow2012-zh and @DeVries2022-of).
+10% in @Boyd2019-ez and @Iversen2023-qj and 35% in @Riebesell2009-qj, @Passow2012-zh and @DeVries2022-of).
 
@@
 Sequestration times for DIC are set depending on the depth at which POC and DOC are exported.
-Figure taken from @Iversen2022-qj.](image/intro/carbon_pumps_iversen2022.jpeg){#fig-global_carbon_pumps}
+Figure taken from @Iversen2023-qj.](image/intro/carbon_pumps_iversen2022.jpeg){#fig-global_carbon_pumps}
 
@@
 macroscopic organic aggregate \> 0.5 mm) [@Alldredge1988-wm; @Turner2015-af; @Steinberg2017-ri].
-Settling velocities for such particles range from 10-100 m day$^{-1}$ up to 1 km day$^{-1}$ for large salp fecal pellets [@Steinberg2017-ri; @Iversen2022-qj].
+Settling velocities for such particles range from 10-100 m day$^{-1}$ up to 1 km day$^{-1}$ for large salp fecal pellets [@Steinberg2017-ri; @Iversen2023-qj].
 
@@
 The color-coded layers represent regions with carbon-specific degradation rate, decreasing with depth.
-Figure inspired from @Iversen2022-qj.
+Figure inspired from @Iversen2023-qj.
 
@@
 
-Based on the Martin curve, @Iversen2022-qj divided the water column in 4 layers (@fig-Martin-Curve): the upper mesopelagic layer (export depth - 300 m) where there is a rapid attenuation of the downward POC flux, the mid-mesopelagic layer (300 - 500 m) and the lower mesopelagic layer (500 - 1,000 m) where both layers show two levels of intermediate attenuation and the bathypelagic layer (1,000 m - seafloor) where the attenuation is relatively uniform.
-For each layer, @Iversen2022-qj estimated a carbon-specific degradation rate in order to better understand the processes responsible of this attenuation.
-As a result, @Iversen2022-qj suggests that the POC attenuation in the upper mesopelagic layer is the combined action of microbial degradation and zooplankton that graze on settling aggregates.
+Based on the Martin curve, @Iversen2023-qj divided the water column in 4 layers (@fig-Martin-Curve): the upper mesopelagic layer (export depth - 300 m) where there is a rapid attenuation of the downward POC flux, the mid-mesopelagic layer (300 - 500 m) and the lower mesopelagic layer (500 - 1,000 m) where both layers show two levels of intermediate attenuation and the bathypelagic layer (1,000 m - seafloor) where the attenuation is relatively uniform.
+For each layer, @Iversen2023-qj estimated a carbon-specific degradation rate in order to better understand the processes responsible of this attenuation.
+As a result, @Iversen2023-qj suggests that the POC attenuation in the upper mesopelagic layer is the combined action of microbial degradation and zooplankton that graze on settling aggregates.
 Here, zooplankton are the main responsible for the attenuation whereas in the middle and lower mesopelagic layers, microbial degradation takes over due to the decrease in zooplankton abundance.
@@
 
-The impact on the downward POC flux of zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2022-qj].
+The impact on the downward POC flux of zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2023-qj].
 On the one hand, aggregate feeding, filter-feeding, direct ingestion of fecal pellets (i.e.
@@
 
-The size of aggregates was usually thought to be a key predictor of settling velocity \[e.g.
-@Passow2006-hw\] which is coherent with @eq-sinking-speed.
+The size of aggregates was usually thought to be a key predictor of settling velocity [e.g.
+@Passow2006-hw] which is coherent with @eq-sinking-speed.
 However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor.
 Because the sinking speed determines the time available for degradation in the water column, it also directly impacts the potential for carbon sequestration.
-With all things considered, there is a thus need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities \[@Iversen2020-kb; @Giering2020-zk; @Iversen2022-qj; @Williams2022-xx\].
+With all things considered, there is a thus need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Giering2020-zk; @Iversen2023-qj; @Williams2022-xx].
 

### 2023-03-11 article3.qmd
update

@@
 It is composed of three main pathways [@Claustre2021-tx; @Boyd2019-ez; @Siegel2022-wd]: the biological gravitational pump that is the gravitational settling of particulate organic carbon (POC); the physically mediated particle injection pump that represents physical mechanisms that transport dissolved organic carbon (DOC) to depth via mixing processes; and the biologically mediated particle injection pump that characterizes the transport of POC (and DOC) by diel and seasonal migration of animals in the water column.
-Historically, the BCP was mainly studied with sediment traps [@Honjo2008-kn] and Thorium-234 radionuclide disequilibria [@Buesseler2007-ql] focused on the gravitational pump hence the contribution of DOC and migrating animals to the BCP was likely under-estimated.
-However, the advent of autonomous underwater vehicles (AUVs) equipped with a variety of physical and biogeochemical sensors have demonstrated their ability to better understand the latter [e.g.
+Historically, the BCP was mainly studied with sediment traps [@Honjo2008-kn] and Thorium-234 radionuclide disequilibria [@Buesseler2007-ql] focused on the gravitational pump hence the contribution of DOC and migrating animals to the BCP was poorly known.
+However, the advent of autonomous underwater vehicles (AUVs) equipped with a variety of physical and biogeochemical sensors have demonstrated their ability to better understand them [e.g.
 @Omand2015-mf; @DallOlmo2016-ex; @Llort2018-pa; @Lacour2019-id; @Haentjens2020-dc; @Stukel2017-lc].
@@
 In addition to the six core variables, new sensors have been developped to target specific BCP-related questions [@Claustre2020-jo; @Claustre2021-tx; @Chai2020-uq].
-In this study, we will focus on BGC-Argo floats equipped with two new sensors: the Underwater Vision Profiler (UVP6) [@Picheral2021-sx] and a transmissometer used as an optical sediment trap (OST).
-
-The UVP6 is a low power miniaturized version of the UVP5 [@Picheral2010-to].
+In this study, we will focus on BGC-Argo floats equipped with two additional sensors: the Underwater Vision Profiler (UVP6) [@Picheral2021-sx] and a transmissometer used as an optical sediment trap (OST).
+
+The UVP6 is a portable version of the UVP5 [@Picheral2010-to], specifically designed for AUVs [@Picheral2021-sx].
 It consists of an underwater camera that counts particles and images objects in, respectively, the 0.1 - 2.5 cm and 0.8 - 2.5 cm size range.
-The latest version of the UVP6 also embarks an embedded image classification algorithm that transmits the abundance of 20 classes in near-real time.
+In contrast with UVP5 deployments on cabled platforms, UVP6 deployed on long-term drifting AUVs (e.g.
+BGC-Argo floats) are rarely recovered.
+Therefore, the latest version of the UVP6 also embarks an embedded image classification algorithm that transmits the abundance of 20 classes (16 biological classes) in near-real time.
 The OST consists of a vertically mounted transmissometer whose upward-facing optical window intercepts sinking particles.
-The latter have already been used on AUVs [@Bishop2004-at; @Bishop2009-ka; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk].
-
-Two pairs of BGC-Argo floats equipped with both the UVP6 and an OST were deployed in the southern part of the Labrador Sea in May 2022 to investigate seasonal and bloom-related particle and carbon fluxes.
-The Labrador and the northern part the atlantic are charateterid by blablabla -\> lire le truc et faire la micro revie sur l'atlantique nord, son bloom et la mer du labrador.
-
-but it is the first time that BGC-Argo floats are deployed with both a UVP6 and an OST.
-
-In this work, we will
-
--   Pour l'intro je dois lire briggs 2011-2020
-
--   projet remocean Hervé
-
--   miwing convection virginie thierry?
-
--   atlantic nord, le bloom de l'atlantic nord
-
--   léo lacour
-
--   yashayeav
-
--   hanetjens 2020
-
-introduire la notion de cycle de flotteurs BGC-Argo etc (+ parking depth)
-
--   dire que le but du papier est de faire une comparaison OST UVP qui sont deux instruments qui permettent de mesurer la meme chose (le flux des particules, l'abondance de particules) et donc le flux de carbone de manière indirecte papier GRL Lacour de 2015
--   introduire le POC ici et introduire la BCP vite fait (classique)
--   dire que Chla = proxy pour phyto biomass
--   BGC-Argo floats are pseudo-lagrangian platforms
--   si on peut prouver que le diagramme ST, les massses d'eau sont les memes, alors on peut peut-être parler des ratios entre particlues à 500 et 1000 m
--   définir la méthode de calcul de la MLD
--   plotter la MLD sur les profils verticaux de
--   from llort 2017: This is especially true in the Southern Ocean, where the circulation has been described as equivalent barotropic (Gille et al., 2007; Killworth & Hughes, 2002).
-An equivalent barotropic circulation basically implies that the strength of the current might change with depth (weaker as depth increases) but the orientation does not change.
-Therefore, Argo and BGCArgo can be interpreted as "quasi-Lagrangian" because floats follow well ocean mesoscale circulation lasting more than 1 week; yet they cannot follow shorter and smaller features such the ones associated to submesocale circulation
-
-Two pairs of Biogeochemical-Argo (BGC) floats were planned to be deployed at the time of the annual spring bloom in the Labrador Sea (LS).
-The latter is characterized by two bioregions with distinct phytoplankton bloom phelonogies on both side of the 60°N parallel [@Lacour2015-da].
-The northern part is characterized by a short spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June [@Lacour2015-da].
-In this study, two pairs of floats were deployed south of the 60°N parallel with a first couple (WMO 6904240 and 4903634) deployed closely to it (58.9°N) in late May 2022.
+The latter have already been used on AUVs [@Bishop2004-at; @Bishop2009-ka; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk] and have demonstrated their usefulness to characterize the gravitational pump.
+
+For the first time, two pairs of BGC-Argo floats equipped with both the UVP6 and an OST were deployed (May 2022) in the Labrador Sea (LS) to investigate the seasonal and episodic dynamics of particle and carbon fluxes.
+The LS is a basin of the subpolar North Atlantic (NA) ocean which represents 20% of the global ocean carbon sink \[@Takahashi2009-rv\].
+The efficiency of this sink is explained by both deep winter convection that traps CO$_{2}$ from the atmosphere \[@Vage2008-qa; @Yashayaev2009-fe; @Yashayaev2017-mn\] and by the largest spring phytoplankton bloom \[@Siegel2002-hk\] that consumes CO$_{2}$ for the photosynthetic production of organic matter.
+The triggering mechanism of the NA bloom was initially explained by the 'Critical Depth Hypothesis' of @Sverdrup1953-ki that defines a critical surface mixing depth where phytoplankton growth is equal to phytoplankton losses.
+According to @Sverdrup1953-ki, a bloom is therefore triggered when the surface mixed layer depth (MLD) is shallower than this critical depth resulting in a positive phytoplankton growth rate.
+However, this hypothesis was challenged by @Behrenfeld2010-my who suggested that the initiation of the NA blooms begins in winter with a slow phytoplankton biomass increase over the following months (i.e.
+'Dilution-Recoupling' hypothesis).
+These hypotheses have been investigated with BGC-Argo floats but the debate still remains \[@Mignot2018-gm; @Yang2020-hw; @Boss2010-do\].
+For the LS, @Lacour2015-da used a climatology of satellite Chla to divide the NA into 6 bioregions with specific bloom phenologies.
+In their bioregionalization (see also their Figure 2), the LS is characterized by two bioregions with distinct bloom phenologies on both side of the 60°N parallel.
+The northern part is characterized by a short spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June.
+@Lacour2015-da concluded that the concurrent influence of surface light and mixing was the first-order mechanism controlling the onset of the spring bloom in both bioregions.
+This is coherent with the results of @Wu2008-vw and @Frajka-Williams2010-ph who found that early shoaling of the mixed layer was responsible for the early spring bloom in the northern part whereas the light controlled the bloom in the southern part.
+
+In late May 2022, two pairs of BGC-Argo floats were deployed in the southern part of the LS at the time of the spring bloom.
+While BGC-Argo floats typically drift at 1000 m for \~10 days between each ascending profile, these two pairs of floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study POC and particle fluxes derived from beam attenuance \[@Estapa2023-oz\] and UVP6 particle size distributions (PSD).
+The goals of this study are threefold: 1) Compare the UVP6 and OST as two independent measurements of downward carbon fluxes and particle abundance; 2) Explore the seasonal signal of the downward POC flux and particle abundance with Chla and MLD using BGC-Argo floats and satellite data; 3) Evaluate the new drifting configuration and how it could be improved in regards with our preliminary results.
 
@@
 
+parler aussi du fait que lesmesures en dérivés, il y a 2h en UVP et OST etc peut-etre que les mesures à 500 m ...
+ou alors on mets plus de temps à 200 m...
+voir un peu.
+faire des profils tous les 10 jours pour une telle apllication ça sert à rien étant donné qu'en plus si on a l'uvp 6 avec l ER embarquée pour éutider la vmmp il faut clairmeent avoir une résolution temporelle plus éleée, ce qui est loin parfait donc voilà.
+
+TO DO - faire le profil de Martin (flux vertical) et faire les données de HPLC pour étudier les communautés et voir si elles sont différentes entre les deux sites de déploiements.
+
 ## Conclusion

### 2023-03-12 article3.qmd
update

@@
 twilight) zone [@Martin2020-wa].
-In order to fill the observational gaps, the Biogeochemical-Argo (BGC-Argo) network aims to develop an operational global array of 1000 profiling floats measuring six core variables from the surface to 2000 m [@Claustre2020-jo]: oxygen, nitrate, pH, chlorophyll a (Chla), suspended particles ($b_{bp}$) and downwelling irradiance.
+In order to fill the observational gaps, the Biogeochemical-Argo (BGC-Argo) network aims to develop an operational global array of 1000 profiling floats measuring six core variables from the surface to 2000 m [@Claustre2020-jo]: oxygen, nitrate, pH, chlorophyll a (Chla), suspended particles and downwelling irradiance.
 In addition to the six core variables, new sensors have been developped to target specific BCP-related questions [@Claustre2020-jo; @Claustre2021-tx; @Chai2020-uq].
@@
 Therefore, the latest version of the UVP6 also embarks an embedded image classification algorithm that transmits the abundance of 20 classes (16 biological classes) in near-real time.
-The OST consists of a vertically mounted transmissometer whose upward-facing optical window intercepts sinking particles.
+The OST consists of a vertically mounted transmissometer whose upward-facing optical window intercepts sinking particles to estimate the downward POC flux.
 The latter have already been used on AUVs [@Bishop2004-at; @Bishop2009-ka; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk] and have demonstrated their usefulness to characterize the gravitational pump.
 
-For the first time, two pairs of BGC-Argo floats equipped with both the UVP6 and an OST were deployed (May 2022) in the Labrador Sea (LS) to investigate the seasonal and episodic dynamics of particle and carbon fluxes.
-The LS is a basin of the subpolar North Atlantic (NA) ocean which represents 20% of the global ocean carbon sink \[@Takahashi2009-rv\].
-The efficiency of this sink is explained by both deep winter convection that traps CO$_{2}$ from the atmosphere \[@Vage2008-qa; @Yashayaev2009-fe; @Yashayaev2017-mn\] and by the largest spring phytoplankton bloom \[@Siegel2002-hk\] that consumes CO$_{2}$ for the photosynthetic production of organic matter.
+For the first time, BGC-Argo floats equipped with both the UVP6 and an OST were deployed in the Labrador Sea (LS) to investigate the seasonal and episodic dynamics of particles and carbon fluxes.
+The LS is a basin of the subpolar North Atlantic (NA) ocean which represents 20% of the global ocean carbon sink [@Takahashi2009-rv].
+The efficiency of this sink is explained by both deep winter convection that traps CO$_{2}$ from the atmosphere [@Vage2008-qa; @Yashayaev2009-fe; @Yashayaev2017-mn] and by the largest spring phytoplankton bloom [@Siegel2002-hk] that consumes CO$_{2}$ for the photosynthetic production of organic matter.
 The triggering mechanism of the NA bloom was initially explained by the 'Critical Depth Hypothesis' of @Sverdrup1953-ki that defines a critical surface mixing depth where phytoplankton growth is equal to phytoplankton losses.
@@
 'Dilution-Recoupling' hypothesis).
-These hypotheses have been investigated with BGC-Argo floats but the debate still remains \[@Mignot2018-gm; @Yang2020-hw; @Boss2010-do\].
+These hypotheses have been investigated with BGC-Argo floats but the debate still remains [@Mignot2018-gm; @Yang2020-hw; @Boss2010-do].
 For the LS, @Lacour2015-da used a climatology of satellite Chla to divide the NA into 6 bioregions with specific bloom phenologies.
@@
 This is coherent with the results of @Wu2008-vw and @Frajka-Williams2010-ph who found that early shoaling of the mixed layer was responsible for the early spring bloom in the northern part whereas the light controlled the bloom in the southern part.
-
 In late May 2022, two pairs of BGC-Argo floats were deployed in the southern part of the LS at the time of the spring bloom.
-While BGC-Argo floats typically drift at 1000 m for \~10 days between each ascending profile, these two pairs of floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study POC and particle fluxes derived from beam attenuance \[@Estapa2023-oz\] and UVP6 particle size distributions (PSD).
-The goals of this study are threefold: 1) Compare the UVP6 and OST as two independent measurements of downward carbon fluxes and particle abundance; 2) Explore the seasonal signal of the downward POC flux and particle abundance with Chla and MLD using BGC-Argo floats and satellite data; 3) Evaluate the new drifting configuration and how it could be improved in regards with our preliminary results.
+While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study POC fluxes and particle concentrations derived from beam attenuance [@Estapa2023-oz] and UVP6 particle size distributions (PSD).
+
+The goals of this study are threefold: 1) the comparison of the UVP6 and OST as two independent measurements of downward carbon fluxes and particle abundance on BGC-Argo floats; 2) the exploration of the seasonality (bloom and non-bloom periods) of downward POC fluxes and particle abundance in the LS; 3) the evaluation of the new drifting configuration and how it could be improved to study specific BCP-related questions.
 
@@
 Two couples of BGC-Argo floats were deployed in the Labrador Sea in late May 2022.
-The first couple (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second (WMO 6904241 and 1902578) was deployed on the 30$^{th}$ at coordinates (-52.4°E, 56.8°N).
-
-All 4 floats used in this study are PROVOR Jumbo floats (NKE Instrumentation, France) equipped with a CTD SBE41-CP (Sea-Bird Scientific, USA), a UVP6-LP (Hydroptic, France), a c-Rover 660 nm beam transmissometer (WETLabs, USA) and a ECO FLBBCD (WETLabs, USA) that includes a chlorophyll-a (Chla) fluorometer (excitation at 470 nm, emission at 695 nm), a colored dissolved organic matter (CDOM) fluorometer (excitation at 370 nm, emission at 460 nm) and a 700 nm particle backscatteringmeter ($b_{bp}$).
+The first couple (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second (WMO 6904241 and 1902578) was deployed on the 29$^{th}$ at coordinates (-52.4°E, 56.8°N).
+
+All 4 floats used in this study are PROVOR Jumbo floats (NKE Instrumentation, France) equipped with a CTD SBE41-CP (Sea-Bird Scientific, USA), a UVP6-LP (Hydroptic, France), a c-Rover 660 nm beam transmissometer (WETLabs, USA) and a ECO FLBBCD (WETLabs, USA) that includes a Chla fluorometer (excitation at 470 nm, emission at 695 nm), a colored dissolved organic matter (CDOM) fluorometer (excitation at 370 nm, emission at 460 nm) and a 700 nm particle backscatteringmeter ($b_{bp}$).
 All floats but one (1902578) were equipped with a OCR-504 Multispectral Radiometer to measure the photosynthetic available radiation.
@@
 All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux.
-During these drift phases, measurements of particles abundance (14 size classes, between 0.1 mm and 2.5 mm) are taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths.
+During these drift phases, measurements of particle abundance (14 size classes, between 0.1 mm and 2.5 mm) are taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths.
 This configuration can be adjusted during the mission of the float via two-way Iridium communication.
@@
 Chla data were quality controlled for Non-Photochemical Quenching (NPQ), a photoprotective mechanism that decreases the apparent fluorescence per unit of Chla following the method implemented by @Terrats2020-ef who modified the method of @Xing2018-pn for shallow-mixing cases.
-Chla data were also divided by a slope factor of 2.15 following the recommendation of @Roesler2017-me for obtaining unbiased Chla estimates from in situ WETLabs ECO fluorometer.
+Chla data were also divided by a slope factor of 2.15 following the recommendation of @Roesler2017-me for obtaining unbiased Chla estimates from in situ WETLabs ECO fluorometers.
+Temperature and salinity data were quality controlled using @Wong2022-hz.
+The potential density anomaly was computed with the function swSigma0 of the oce R package [@Kelley_undated-sw] and the MLD was determined using the 0.03 kg m$^{-3}$ density criterion at a reference depth of 10 m [@De_Boyer_Montegut2004-hh].
 $c_{p}$ data were converted from counts to physical units (m$^{-1}$) using the following relationship
@@
 UVP6 data and in particular the averaged number of particles per size class (i.e.
-particle size distribution hereafter referred to as PSD) was divided by the number of images at each depth and the water volume imaged by the UVP (0.7L) to obtain particle concentrations (units: #/L).
-Note that in an impending update by the Coriolis data center, it should not be necessary to divide by the number of images anymore.
+PSD) was divided by the number of images at each depth and the water volume imaged by the UVP (0.7L) to obtain particle concentrations (units: #/L).
+Note that in an impending update by the Coriolis data center, it will no longer be necessary to divide by the number of images.
 
@@
 In order to compute $F_{small}$, we firstly computed the slope of each group of observations between jumps (positive and negative, see @fig-estapa-method_article).
-Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window (e.g.
+Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by factors other than the pump cleaning the optical window (e.g.
 turbulence ou zooplankton consumption of accumulated material).
@@
 
-```{r, example of a cp signal analyzed for jumps (with an old method here)}
+```{r, example of a cp signal analyzed for jumps (with an old method here), fig.width=8, fig.height=5}
 #| label: fig-estapa-method_article
@@
     #scale_shape_manual(values = c(1,1,4)) +
-    geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', size = 2) + theme_bw() + 
+    geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', linewidth = 1) + theme_bw() + 
     labs(x = 'Time', y = TeX('$c_{p}~(m^{-1})$')) +
@@
   geom_point(data = check, aes(x = dates, y = cp), shape = 1, size = 3, colour = 'black')  +
-    theme(legend.position = 'none')
+    theme(legend.position = 'none') + scale_x_datetime(breaks = c(as.POSIXct('2022-11-16'), as.POSIXct('2022-11-18')), date_labels = 'Nov %d')
 ```
@@
 
-Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: [10.48670/moi-00281](https://doi.org/10.48670/moi-00281)) with a spatial resolution of 4 km.
-Mixed layer depth (MLD) derived from satellite observations were downloaded from near-real-time weekly data from the ARMOR3D product (doi: [10.48670/moi-00052](https://doi.org/10.48670/moi-00052)).
+Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: 10.48670/moi-00281) with a spatial resolution of 4 km.
+Mixed layer depth (MLD) derived from satellite observations were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052).
 
@@
 At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$.
-For both drifting depths, the trend shows a decrease of $F_{small}$ from June to, respectively, January and February.
-Note that the drifting configuration of all 4 floats was adapted to the deepening of the MLD (Fig.
-XX) hence drifting at 200 m and 500 m was stopped at this period to avoid measurements bias due to turbulence.
+For both drifting depths, the trend shows a decrease of $F_{small}$ from June to, respectively, January and February before the drift was stopped at those depths due to the deepening of the winter MLD (@fig-mld) in order to ensure the float stability.
 At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid/end of July followed by a steady decrease towards March, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.
 
-```{r, plot Fsmall and Fbig, fig.width=8, fig.height=6}
+```{r, Fsmall and Fbig, fig.width=8, fig.height=6}
 #| label: fig-OST-flux
@@
 Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.
-There is no data after January and February at, respectively, 200 m and 500 m because the drifting was stopped at those depths due to the increase of the MLD.
+There is no data after January and February at, respectively, 200 m and 500 m because the drifting was stopped at those depths due to the deepening of the MLD.
 
@@
 up / down
-
+```
+
+```{r, MLD, fig.height=4, fig.width=8}
+#| label: fig-mld
+#| fig-cap: MLD measured by all floats from the end of May 2022 to early March 2023. Black dashed lines represent the first two drifting depths.
+
+# add MLD here
+mld_data <- vroom('/home/flo/dox/ArgoShine/npq_corrected_data.csv')
+# clean and smoothed chla data
+mld_data <- mld_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> dplyr::mutate(date = as_date(juld), WMO = factor(wmo, levels = c(6904240,4903634,6904241,1902578)))
+mld_data <- mld_data |> group_by(date, WMO, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
+
+# reverselog_trans <- function(base = exp(1)) {
+#     trans <- function(x) -log(x, base)
+#     inv <- function(x) base^(-x)
+#     trans_new(paste0("reverselog-", format(base)), trans, inv, 
+#               log_breaks(base = base), 
+#               domain = c(1e-100, Inf))
+# }
+# 
+# mld_plot <- ggplot(mld_data) + geom_line(aes(x = date, y = mld, colour = WMO)) +
+#   labs(x = 'Month', y = 'MLD (m)') + scale_y_continuous(trans = reverselog_trans(10), breaks = c(10,50,200,400,750)) +
+#   scale_color_brewer(palette = 'Dark2') + 
+#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + theme_bw()
+
+# reverselog_trans <- function(base = exp(1)) {
+#     trans <- function(x) -log(x, base)
+#     inv <- function(x) base^(-x)
+#     trans_new(paste0("reverselog-", format(base)), trans, inv, 
+#               log_breaks(base = base), 
+#               domain = c(1e-100, Inf))
+# }
+
+mld_plot <- ggplot(mld_data) + geom_line(aes(x = date, y = mld, colour = WMO)) +
+  labs(x = 'Month', y = 'MLD (m)') + scale_y_reverse() +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + theme_bw() +
+  geom_hline(yintercept = 200, colour = 'black', linetype = 'dashed') +
+  geom_hline(yintercept = 500, colour = 'black', linetype = 'dashed')
+
+mld_plot
 ```
@@
 In the upper part of the size range, @fig-uvp6-profiling-512-2500 shows a rapid emptying of the entire water column starting in November.
-In contrast with the small classes, some particles, especially above 1.5 mm, have already virtually disappeared from the water column by September.
+In contrast with small classes (referred to here as below 512 $\mu$m), some particles, especially above 1.5 mm, have almost entirely disappeared from the water column by September.
 However, the rapid increase and decrease of particles on the entire water column around July-August for floats 6904241 and 1902578 (still in the same zone at that time) could suggest an export event.
@@
 
-### Environmental conditions before and shortly after deployment
-
-All floats measured similar particle concentrations at each drifting depth and on the vertical.
+### Phenology of the spring bloom in the Labrador Sea
+
+<!-- All floats measured similar particle concentrations at each drifting depth and on the vertical during the entire time series.
 In addition, except perhaps for the upper part of the UVP6 size spectrum for floats 6904241 and 1902578, the results do not indicate a clear hint of particle export signature.
 Therefore, we suggest that the 4 BGC-Argo floats mostly measured the dynamics of the seasonal, continuous, particle signal but that they missed the main transient particle export events associated with the LS spring bloom.
-
-To estimate the time of the bloom in the deployment area, we compared MLD and surface (0 - 20 m) Chla concentrations from both satellite and in situ data (@fig-chla-mld-satellite) in a 100 km radius around each deployment.
-It can be seen that these estimates are in agreement, therefore we hypothesize that we can trust satellite data prior to the floats deployment.
-Taking this into consideration, it seems that the pair 6904240-4903634 missed a bloom that occurred in early May whereas the pair 6904241-1902578 missed a first Chla increase in late May days before their deployment.
-However, a second equivalent increase (median of 2 mg m$^{-3}$) happened in late June - early July, which could be a contributing factor to the apparent particle track observed around July and August.
+-->
+
+To estimate the time of the bloom in the deployment area, we compared MLD and surface (i.e.
+upper 20 m) Chla concentrations from both satellite and in situ data in a 100 km radius around each deployment.
+@fig-chla-mld-satellite shows that satellite and in situ Chla and MLD data concur therefore we assume that the former can be used to study environmental conditions before the floats deployment.
+Taking this into consideration, @fig-chla-mld-satellite indicates that the 6904240-4903634 pair missed a bloom that occurred in early May whereas the 6904241-1902578 pair missed a first (satellite) Chla increase at the end of May days before their deployment.
+However, a second (satellite and in situ) Chla increase reached a peak at the end of June before it decreased again towards July.
 
 ```{r, chla and MLD from satellite, fig.height=6, fig.width=6}
-#|: label: fig-chla-mld-satellite
-#| fig-cap: Comparison of MLD and Chla concentrations measured from satellites (black) and BGC-Argo floats (red) following their deployment.
-Satellite error bars represent the first and third quartiles around the median in a 50 km radius around each deployment (each pair of floats) coordinates.
-In situ Chla represent the Chla median in the upper 20 m of the water column with the upper and lower incertitude in the correction factor of @Roesler2017-me.
+#| label: fig-chla-mld-satellite
+#| fig-cap: Comparison of MLD and Chla concentrations measured from satellites (dark blue) and BGC-Argo floats (red) following their deployment (in pairs).
+Satellite error bars represent the first and third quartiles around the median in a 100 km radius around each deployment coordinates.
+In situ Chla represents the Chla median in the upper 20 m of the water column with uncertainties from the slope factor uncertainty of @Roesler2017-me.
 
@@
 d <- d |> dplyr::mutate(smoothed_chla = smooth(mean_chla)/2.15, smoothed_chla_up = smooth(mean_chla)/(2.15-0.37), smoothed_chla_down = smooth(mean_chla)/(2.15+0.37)) # 2.15 = Roesler factor in the Labrador Sea
+
+# remove NaN in smoothed chla
+d <- d |> drop_na(smoothed_chla)
 
@@
 chla_cover_deployment2 <- points_in_circle(chla_sat, lon_center = -52.3, lat_center = 56.7, radius = 10000) # 50 km around center
-chla_cover_deployment2 <- chla_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-30'))
+chla_cover_deployment2 <- chla_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-29'))
 
@@
 chla_cover <- rbind(chla_cover_deployment1, chla_cover_deployment2)
+
+# HPLC data
+# chla_hplc <- vroom('DATA/ARGO/CE22009_pigments_031122_Travail_Herve_FŔ.csv') |> filter(Station %in% c(29,9)) |> select(depth = `Depth_(m)`, chla = Total_Chlorophyll_a, station = Station) |> filter(depth < 10) |> mutate(group = if_else(station == 9, '6904240 - 4903634', '6904241 - 1902578')
 
@@
                       fun.max = function(z) {quantile(z, 0.75)},
-                      fun = median, colour = 'black', width = .5) + theme_bw() +
+                      fun = median, colour = '#003366', width = .5) + theme_bw() +
   labs(x = 'Month', y = TeX('Chla (mg m$^{-3}$)'))+ facet_wrap(~group) +
@@
                       fun.max = function(z) {quantile(z, 0.75)},
-                      fun = median, colour = 'black', width = 5) +
+                      fun = median, colour = '#003366', width = 5) +
   labs(x = 'Month', y = 'MLD (m)') + facet_wrap(~group) + theme_bw() +
@@
 
-The Chla satellite pattern for the pair 6904240-4903634 typically describes the phenology of the spring bloom in the northern part of the LS [@Lacour2015-da].
-This is not surprising given that this pair was deployed one degree southward.
-The other pair was deployed in the bioregion where the phytoplankton biomass slowly develops during May to reach its maximum in June [@Lacour2015-da].
-The Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD and that it started again in mid-June when the water column was well stratified.
-However, the Chla satellite concentration for that pair was three times lower than for the first one.
-Hence, it is possible that the bloom intensity was not sufficient to produce clear particle export signals, at least for BGC-Argo floats in their 10-day vertical sampling configuration.
-
-```{r, chla corrected for NPQ and MLD}
-# #|: label: fig-chla-bgc
-# #|: fig-cap : Chla concentrations and MLD measured by all floats in the upper 250 m for the period following the deployment.
+The Chla satellite pattern for the 6904240-4903634 pair typically describes the phenology of the spring bloom in the northern part of the LS [@Lacour2015-da].
+This is not surprising given that this pair was deployed one degree southward to the so-called 60°N limit [@Lacour2015-da].
+@fig-chla-mld-satellite shows that the satellite MLD in that region had already shoaled below 100 m since March and that the water column was stable and relatively well stratified in May which might explain the intensity of surface Chla concentration in contrast with those measured in the other region.
+The other pair was deployed in the bioregion where the phytoplankton biomass is said to slowly develops during May to reach its maximum in June [@Lacour2015-da].
+The Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD (end of May) and that it started again in mid-June when the water column was well stratified.
+<!-- However, the increase of the MLD in June could have also impacted the intensity of the bloom which might explain the three times lower than Chla concentrations.
+However, the Chla satellite concentration (before deployment) for that pair was three times lower than for the other pair and the second Chla increase (satellite and in situ) was of the same order of magnitude.
+Therefore, the second pair was deployed in a less productive area and that the inten <!--Hence, it is possible that the bloom intensity was not sufficient to produce clear particle export signals, at least for BGC-Argo floats in their 10-day vertical sampling configuration.
+-->
+
+### Unclear particle export signature
+
+Even though we suggested that the two pairs of floats were deployed in two different bloom regimes, they all measured an increase in particle concentrations in the 102 - 512 $\mu$m range (@fig-uvp6-profiling-102-406) around 800 - 1000 m as well as an increase in the abundance of big particles (@fig-uvp6-profiling-512-2500) between June and August.
+This increase raises the question of a possible particle export event.
+In the upper part of the PSD, the vertical concentration profiles highlight a bigger particle plume for the 6904241-1902578 pair than for the other one.
+However, apart from the float 6904240, all floats measured similar integrated Chla content in the upper 100 m and in the mixed layer (@fig-chla-content).
+@fig-OST-flux also shows that $F_{small}$ is similar across all floats at all parking depths and that $F_{big}$ is also similar across all floats at 1000 m with the same increase (from June to August) than for F$_{small}$.
+The same behavior is observed for UVP6 particles in the lower size range at 1000 m (@fig-uvp-1000m).
+
+```{r, mld + chla, fig.height=6}
+#| label: fig-chla-content
+#| fig-cap: Integrated Chla content in the mixed layer (up) and in the upper 100 m for each float.
+
+chla_mld <- d_chla |> drop_na(chla_0_100m)
+
+chla_plot_mld <- ggplot(chla_mld) + geom_line(aes(x = date, y = chla_0_mld, colour = as.factor(wmo)), linewidth = 1) + labs(x = 'Month', y = TeX('Integrated Chla content (mg m$^{-2}$)'), colour = 'WMO') +
+  theme_bw() + scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  #scale_colour_manual(values = c('#003366', '#E31B23'))
+  scale_color_brewer(palette = 'Dark2')
+
+chla_plot_100m <- ggplot(chla_mld) + geom_line(aes(x = date, y = chla_0_100m, colour = as.factor(wmo)), linewidth = 1) + labs(x = 'Month', y = TeX('Integrated Chla content (mg m$^{-2}$)'), colour = 'WMO') +
+  theme_bw() + scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
+  scale_color_brewer(palette = 'Dark2') + theme(legend.position = 'none')
+
+chla_plot_mld / chla_plot_100m
+```
+
+```{r, spiciness}
+#| label: fig-spice
+#| fig-cap: Spiciness (up) and particle concentrations for the 102-128 $\mu$m size class.
+
+#ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26, park_depth != '200 m') |> dplyr::mutate(date = as_date(date))
+# ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
+#   dplyr::mutate(date = as_date(date), colour = if_else(park_depth == '200 m', '#E31B23', if_else(park_depth == '500 m', '#FFC325', 'black')))
+
+ts_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
+  dplyr::mutate(date = as_date(date))
+
+ts_data <- ts_data |> mutate(spiciness = swSpice(ts_data$psal, ts_data$temp, ts_data$depth, ts_data$lon, ts_data$lat))
+
+# # https://dankelley.github.io/oce/reference/plotTS.html
+# # https://dankelley.github.io/oce/reference/argo-class.html
+# test <- as.argo(time = ts_data$date, longitude = ts_data$lon, latitude = ts_data$lat, salinity = ts_data$psal, temperature = ts_data$temp, pressure = ts_data$depth,
+#         id = ts_data$wmo) 
+# 
+# plotTS(test, inSitu = T, type = 'p', xlab = 'Salinity [PSU]')
 # 
 # 
-# # plot it
-# ggplot(d |> filter(date < '2022-08-01')) + geom_point(aes(x = date, y = pres, colour = log10(smoothed_chla))) +
-#   scale_y_reverse(limits = c(250,0)) +
-#   #scale_y_continuous(trans = 'rev_sqrt') +
-#   #scale_colour_viridis(option = 'A') +
-#   scale_colour_cmocean(name = 'algae', oob = scales::squish, limits = c(-1,1), na.value = 'grey') +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-#   labs(x = 'Time', y = 'Depth (m)', colour = TeX('Chla (mg m$^{-3}$)')) + facet_wrap(~wmo) + theme_bw() +
-#   geom_line(aes(x = date, y = mld), colour = 'black', size = 1)
-```
-
-#### Just because it needs to go somewhere
-
-```{r, mld}
-
-ggplot(d) + geom_line(aes(x = date, y = mld), colour = 'black') +
-  facet_wrap(~wmo) + labs(x = 'Month', y = 'MLD (m)') + scale_y_reverse() +
-  scale_x_date(labels = date_format("%m"), date_breaks = '1 month')
-
+# ggplot(ts_data) + geom_point(aes(x = spiciness, y = depth)) + scale_y_reverse() + facet_wrap(~wmo)
+
+a <- ggplot(ts_data |> filter(depth <= 1000)) + geom_point(aes(x=date, y=depth, colour=spiciness), size = 3, shape = 15) +
+  scale_y_reverse() +
+  #scale_color_viridis() +
+  scale_color_cmocean(name = 'balance', limits = c(-0.25,0.25), oob = scales::squish, na.value = 'grey') +
+  theme_bw() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = TeX('Spiciness (kg m$^{-3}$)'), x = 'Month', y = 'Depth (m)') +
+  facet_wrap(~wmo, ncol = 4)
+
+b <- particles_bis |> filter(depth <= 1000, size == 102) |>
+  #group_by(size) |>
+  ggplot() + facet_wrap(~wmo, ncol = 4) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=2, shape=15) +
+  #scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+  scale_color_cmocean(name = 'thermal', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) +
+  scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = 'Particles (#/L)', x = 'Month', y = 'Depth (m)')
+
+a/b
+```
+
+It is also worth noting that the progressive emptying of the water column does not start from the surface in small classes [@fig-uvp6-profiling-102-406].
+By computing the spiciness [@fig-spice], we show that this pattern does not likely result from water subduction (no spiciness anomalies) hence it would be support the hypothesis of the particle export ending.
+
+All things considered, it is not straightforward to conclude for the presence of a carbon export event.
+First, the 10-day profiling configuration hinders the characterization of particle export events that are transient processes.
+Second, the plume of big particles can be misleading because their abundance measured by the UVP6 are very low (usually \< 0.1 #/L).
+Third, all floats measured the same particle dynamics in distinct zones which suggests that if the two systems were initially different (i.e.
+different bloom dynamics), they merged back after the bloom of May.
+Therefore, we suggest that we mostly observe the latent dynamics of the system rather than clear export signatures that would likely be spotted with both higher bloom intensity and a higher profiling temporal resolution.
+
+### hplc pour les communautés pour montrer que les deux endroits sont différents (peripery vs central)
+
+mettre ça dans phenology of the bloom pour bien montrer que l'on est dans 2 systèmes différentes
+
+```{r, HPLC}
+
+pigments <- vroom('DATA/ARGO/CE22009_pigments_031122_Travail_Herve_FŔ.csv') |> filter(Station %in% c(29,9))
+
+# select interesting coumns (all QCs == 1 so OK)
+# pigments <- pigments |> select(station = Station, depth = `Depth_(m)`, Chla_c3 = Chlorophyll_c3, Chla_c1_c2_MgDVP = `Chlorophyll_c1+c2+MgDVP`, Phaeophorbid_a, Peridinin, `19'-Butanoyloxyfucoxanthin`
 ```
@@
 -   science: on voit le système latent, on a raté le bloom
--   rajouter les vitesses de sédimentations (n'importe quelle méthode fera l'affaire) -\> permet d'utiliser le graphique de l'UVP en vertical -\> hop une justification en plus
--   rajouter dans les méthodes les produits satellites qu'on utilise
--   différentes dynamiques post-bloom
--   phénologie du bloom printanier
--   qinking speed -\> trouver un truc à dire et estimer l'incertitude (en lien avec le graphique ou le DZ augmente) + expliquer la méthode vite fait (Léo Lacour, personal communication)
+
+-
+
 -   on se trouve dans un sytème producitve entre guidi 2008 et iversen (très productif)
-
-### Sinking speed
-
--   Montrer que les sinking speed observées ne sont pas compatibles avec des valeurs de bloom (càd rapides pour les grosses particules à minima)
--   rajouter dans les méthodes le calcul de cette sinking speed.
-
-```{r, vertical plots + sinking speed bis, fig.height=8, fig.width=8}
-# 
-# particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld), doy = yday(juld), year = year(juld))
-# 
-# particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) 
-# #particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
-# 
-# # remove buggy data from float 4903634 (negative depth for some reason)
-# particles_bis <- particles_bis |> filter(depth >= 0)
-# 
-# # rel_conc
-# particles_bis <- particles_bis |> dplyr::group_by(wmo, size) |> dplyr::mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
-# 
-# up_part <-  particles_bis |> filter(depth < 2000, size < 512, doy < 220, year == 2022) |>
-#   #group_by(size) |>
-#   ggplot() + facet_grid(wmo~size) + theme_bw() +
-#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
-#   scale_y_reverse() +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
-#   geom_pointrange(data = sinking_fit |> filter(size < 512, DOY < 220, standard_dev <= 30), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'black')
-# 
-# up_part
-# 
-# # down_part <- particles_bis |> filter(depth < 2000, size > 512) |>
-# #   #group_by(size) |>
-# #   ggplot() + facet_grid(wmo~size) + theme_bw() +
-# #   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-# #   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
-# #   scale_y_reverse() +
-# #   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-# #   labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
-# #   geom_pointrange(data = sinking_fit |> filter(size > 512), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'white')
-# 
-# down_part <-  particles_bis |> filter(depth < 2000, size > 512, doy < 220, year == 2022) |>
-#   #group_by(size) |>
-#   ggplot() + facet_grid(wmo~size) + theme_bw() +
-#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
-#   scale_y_reverse() +
-#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
-#   geom_pointrange(data = sinking_fit |> filter(size > 512, DOY < 220, standard_dev <= 30), aes(xmin = as_date(xmin, origin = '2022-01-01'), xmax = as_date(xmax, origin = '2022-01-01'), x = as_date(DOY, origin = '2022-01-01'), y = depth), size = .2, colour = 'black')
-# 
-# down_part
-
-```
-
-```{r, sinking speed plot}
-
-# # NOTE : pas besoin d'updater ici vu que l'on regarde par rapport au 'bloom' de fin mai/juin
-# sinking_fit <- vroom('DATA/ARGO/gaussian_fit_sinking_speed_full_before_06032023.csv') |> dplyr::mutate(size = size_class |> str_remove('NP_Size_') |> as.numeric(), date = as_date(date))
-# 
-# estimate_sinking_speed <- function(siz, data, threshold){
-#   
-#   # above 500 m
-#   tmp <- data |> filter(size == siz) |> filter(depth <= threshold)
-#   model <- lm(formula = depth ~ time_max_abundance, data = tmp)
-#   tmp$sinking_speed_above_threshold <- model$coefficients[2] # equivalent to slope of the fit
-#   tmp2 <- tmp
-#   
-#   # below 500 m
-#   tmp <- data |> filter(size == siz) |> filter(depth > threshold)
-#   model <- lm(formula = depth ~ time_max_abundance, data = tmp)
-#   tmp2$sinking_speed_below_threshold <- model$coefficients[2] # equivalent to slope of the fit
-#   
-#   return(tmp2)
-# }
-# 
-# # sinking speed for each float
-# sinking_6904240 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 6904240), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
-# sinking_6904241 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 6904241), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
-# sinking_1902578 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 1902578), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
-# sinking_4903634 <- map_dfr(unique(sinking_fit$size), estimate_sinking_speed, sinking_fit |> filter(wmo == 4903634), 400) |> select(wmo, sinking_speed_above_threshold, sinking_speed_below_threshold, size) |> distinct()
-# all_sinking_speed <- rbind(sinking_1902578, sinking_4903634, sinking_6904240, sinking_6904241)
-
-```
-
-```{r, sinking speed values}
-
-# gt(all_sinking_speed)
-# 
-# all_sinking_speed |> ggplot() + geom_point(aes(y = sinking_speed_below_threshold, x = size))
-
-```
-
-```{r, check gaussian profiles}
-
-# # depth layers for Gaussian fit
-# depthL <- c(seq(40,200,20), c(250,300,400,500,600,800,1000,1250,1500,1750,2000))
-# 
-# # Gaussian formula #> but z is not the depth but the TIME
-# fgauss <- function(z, peak_conc, peak_doy, dz){
-#   peak_conc*exp(-((z-peak_doy)**2)/(dz**2))
-# }
-# 
-# return_fitted <- function(d){
-#   
-#   #timeframe near the bloom so to say
-#   ztime <- seq(160,240,1)
-#   # gaussian fit
-#   tib <- map_dfr(1:nrow(d), function(i){tibble(ztime = ztime, fgauss = fgauss(ztime, d$max_abundance[i], d$time_max_abundance[i], d$standard_dev[i]), depth = d$depth[i])})
-#   return(tib)
-# }
-# 
-# # test
-# tmp <- sinking_fit |> filter(wmo == 6904240, size == 645)
-# 
-# test <- return_fitted(tmp)
-# 
-# ggplot(test, aes(x = ztime, y = fgauss, colour = factor(depth))) + geom_point() + 
-#   scale_color_viridis_d() +
-#   theme_bw() 
-#
-
-```
 
@@
 attention on est en PARKING DEPTH -\> mettre la couleur de la parking depth dans le truc, ça permettrait de voir si on peut comparer les processus (style ratio 500/1000m) dans les couches qui sont dans la meme masse d'eau.
-
-```{r, TS diagram}
-#ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26, park_depth != '200 m') |> dplyr::mutate(date = as_date(date))
-# ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
-#   dplyr::mutate(date = as_date(date), colour = if_else(park_depth == '200 m', '#E31B23', if_else(park_depth == '500 m', '#FFC325', 'black')))
-
-ts_data <- all_floats |> filter(PhaseName == 'NPAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
-  dplyr::mutate(date = as_date(date))
-
-ts_data <- ts_data |> mutate(spiciness = swSpice(ts_data$psal, ts_data$temp, ts_data$depth, ts_data$lon, ts_data$lat))
-
-# # https://dankelley.github.io/oce/reference/plotTS.html
-# # https://dankelley.github.io/oce/reference/argo-class.html
-# test <- as.argo(time = ts_data$date, longitude = ts_data$lon, latitude = ts_data$lat, salinity = ts_data$psal, temperature = ts_data$temp, pressure = ts_data$depth,
-#         id = ts_data$wmo) 
-# 
-# plotTS(test, inSitu = T, type = 'p', xlab = 'Salinity [PSU]')
-# 
-# 
-# ggplot(ts_data) + geom_point(aes(x = spiciness, y = depth)) + scale_y_reverse() + facet_wrap(~wmo)
-
-a <- ggplot(ts_data |> filter(depth <= 1000)) + geom_point(aes(x=date, y=depth, colour=spiciness), size = 3, shape = 15) +
-  scale_y_reverse() +
-  #scale_color_viridis() +
-  scale_color_cmocean(name = 'balance', limits = c(-0.25,0.25), oob = scales::squish, na.value = 'grey') +
-  theme_bw() +
-  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-  labs(colour = TeX('Spiciness (kg m$^{-3}$)'), x = 'Month', y = 'Depth (m)') +
-  facet_wrap(~wmo, ncol = 4)
-
-b <- particles_bis |> filter(depth <= 1000, size == 102) |>
-  #group_by(size) |>
-  ggplot() + facet_wrap(~wmo, ncol = 4) + theme_bw() +
-  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=2, shape=15) +
-  #scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
-  scale_color_cmocean(name = 'thermal', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) +
-  scale_y_reverse() +
-  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-  labs(colour = 'Particles (#/L)', x = 'Month', y = 'Depth (m)')
-
-```
 

### 2023-03-13 article3.qmd
update

@@
 
-### Phenology of the spring bloom in the Labrador Sea
+### Dynamics of the spring bloom in the Labrador Sea
 
@@
 
-The Chla satellite pattern for the 6904240-4903634 pair typically describes the phenology of the spring bloom in the northern part of the LS [@Lacour2015-da].
+The Chla satellite pattern for the 6904240-4903634 pair typically describes the dynamics of the spring bloom in the northern part of the LS [@Lacour2015-da].
 This is not surprising given that this pair was deployed one degree southward to the so-called 60°N limit [@Lacour2015-da].
-@fig-chla-mld-satellite shows that the satellite MLD in that region had already shoaled below 100 m since March and that the water column was stable and relatively well stratified in May which might explain the intensity of surface Chla concentration in contrast with those measured in the other region.
-The other pair was deployed in the bioregion where the phytoplankton biomass is said to slowly develops during May to reach its maximum in June [@Lacour2015-da].
-The Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD (end of May) and that it started again in mid-June when the water column was well stratified.
+@fig-chla-mld-satellite shows that the satellite MLD in that region had already shoaled below 100 m since March and that the water column was stable and relatively well stratified in May which might explain the intensity of surface Chla concentrations in contrast with those measured at the second deployment site.
+The two other floats were deployed in the bioregion where the phytoplankton biomass typically develops slowly in May to reach its maximum in June [@Lacour2015-da].
+However, the Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD at the end of May and that it started again in mid-June when the water column was well stratified.
 <!-- However, the increase of the MLD in June could have also impacted the intensity of the bloom which might explain the three times lower than Chla concentrations.
@@
 
+### Phytoplankton communities at deployment sites
+
+Here, we refer to the first and second deployment as, respectively, station A and station B.
+High Performance Liquid Chromatography (HPLC) analyses were performed on samples taken during the two deployments.
+The selected pigments, their taxonomic significance and their size class were taken from Table 2 in @Petit2022-de.
+Station A showed Chla concentrations (TChla pigment in @fig-hplc) two times higher than in station B in the first 100 m as well as higher concentrations of diatoms (Fucoxanthin).
+Station B showed higher concentrations of cryptophytes (Alloxanthin), dinoflagellates (Peridin) and green flagellates and prochlorophytes (TChlb) than in station A.
+However, their absolute concentrations are quite low (usually less than 0.05 mg m$^{-3}$).
+Taking that into consideration, it seems that station A was more productive than station B and was dominated by diatoms (i.e.
+microphytoplankton) whereas station B was composed of a majority of smaller phytoplankton species (mostly nano and picozooplankton).
+This supports the idea of two distinct bloom dynamics as inferred from satellite data.
+
+```{r, hplc, fig.width=8, fig.height=6}
+#| label: fig-hplc
+#| fig-cap: Pigments derived from HPLC analyses at stations A and B.
+
+hplc <- vroom('DATA/ARGO/CE22009_pigments_031122_Travail_Herve_FŔ.csv') |> filter(Station %in% c(29,9))
+
+# subset (all QCs are GOOD so OK)
+# hplc <- hplc |> select(lat = Latitude, lon = Longitude, depth = `Depth_(m)`,
+#                        Zea = Zeaxanthin, TChlb, HF_19 = `19'-Hexanoyloxyfucoxanthin`,
+#                        BF_19 = `19'-Butanoyloxyfucoxanthin`, Allo = Alloxanthin,
+#                        Fuco = Fucoxanthin, Peri = Peridinin, TChla = Total_Chlorophyll_a) |> mutate(station = c(rep('station A', 10), rep('station B', 9)))
+
+hplc <- hplc |> select(lat = Latitude, lon = Longitude, depth = `Depth_(m)`,
+                       Zeaxanthin, TChlb, `19'-Hexanoyloxyfucoxanthin`,
+                       `19'-Butanoyloxyfucoxanthin`, Alloxanthin,
+                       Fucoxanthin, Peridinin, TChla = Total_Chlorophyll_a) |> mutate(station = c(rep('station A', 10), rep('station B', 9)))
+
+# format
+# hplc <- hplc |> pivot_longer(cols = c(Zea, TChlb, HF_19, BF_19, Allo, Fuco, Peri, TChla), names_to = 'pigment', values_to = 'concentration') |> drop_na(concentration)
+
+hplc <- hplc |> pivot_longer(cols = c(Zeaxanthin, TChlb, `19'-Hexanoyloxyfucoxanthin`,
+                       `19'-Butanoyloxyfucoxanthin`, Alloxanthin,
+                       Fucoxanthin, Peridinin, TChla), names_to = 'pigment', values_to = 'concentration') |> drop_na(concentration)
+
+ggplot(hplc) + geom_path(aes(y = depth, x = concentration, colour = station)) + 
+  facet_wrap(~pigment, scales = 'free') + scale_y_reverse() +
+  labs(colour = 'Station', x = TeX('Concentration (mg m$^{-3}$)'), y = 'Depth (m)') +
+  theme_bw() + scale_color_manual(values = c('#E31B23', '#003366'))
+```
+
 ### Unclear particle export signature
@@
 This increase raises the question of a possible particle export event.
-In the upper part of the PSD, the vertical concentration profiles highlight a bigger particle plume for the 6904241-1902578 pair than for the other one.
+For classes \> 512 $\mu$m, the vertical concentration profiles highlight a bigger particle plume for the 6904241-1902578 pair than for the other one.
 However, apart from the float 6904240, all floats measured similar integrated Chla content in the upper 100 m and in the mixed layer (@fig-chla-content).
@@
 The same behavior is observed for UVP6 particles in the lower size range at 1000 m (@fig-uvp-1000m).
-
-```{r, mld + chla, fig.height=6}
+Observing similar properties across a wide geographical range around deployments suggests that we observe the latent dynamics of the system rather than episodic export events.
+
+```{r, mld + chla, fig.height=6, fig.width=8}
 #| label: fig-chla-content
-#| fig-cap: Integrated Chla content in the mixed layer (up) and in the upper 100 m for each float.
+#| fig-cap: Integrated Chla content in the mixed layer (up) and in the upper 100 m (down) for each float.
 
@@
 
-```{r, spiciness}
+```{r, spectral slope}
+
+mid_DSE <- c(0.1147968,0.1446349,0.1822286,0.2295937,0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289,1.83674934,2.31415916)
+sizebin <- c(0.02640633,0.03326989,0.04191744,0.05281267,0.06653979,0.08383488,0.10562533,0.13307958,0.16766976,0.21125066,0.26615915,0.33533952,0.422501323,0.532318310)
+
+vertical_profiles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld)) |>
+  filter(depth > 0) # remove buggy data from 4903634
+
+compute_slope <- function(i, data_spectra){
+
+  spectrum <- data_spectra[i,]
+
+  spectrum_norm <- spectrum/sizebin
+
+  # prepare data for linear regression
+  Y <- log(spectrum_norm)
+  X <- log(mid_DSE)
+
+  # check for finite value
+  h <- is.finite(Y)
+  Y <- Y[h]
+  X <- X[h]
+
+  data_slope <- tibble(X=X, Y=Y)
+
+  model <- lm(formula = Y ~ X, data = data_slope)
+  slope <- model$coefficients[2]
+
+  return(slope)
+}
+
+# compute slope on vertical profiles
+vertical_spectra <- as.matrix(data.frame(vertical_profiles[,5:18]))
+index <- seq(from = 1, to = nrow(vertical_spectra), by = 1)
+slopes <- map_dbl(index, compute_slope, data_spectra = vertical_spectra)
+vertical_profiles$spectral_slope <- slopes
+```
+
+```{r, taxo}
+
+# read vertical (NPAR) taxo data
+taxo_1902578 <- vroom::vroom('DATA/ARGO/1902578_TAXO_FromNetCDF.csv') |> mutate(wmo = 1902578, date = argoJuldToTime(juld)) |> filter(PhaseName == 'NPAR')
+taxo_4903634 <- vroom::vroom('DATA/ARGO/4903634_TAXO_FromNetCDF.csv') |> mutate(wmo = 4903634, date = argoJuldToTime(juld)) |> filter(PhaseName == 'NPAR')
+taxo_6904240 <- vroom::vroom('DATA/ARGO/6904240_TAXO_FromNetCDF.csv') |> mutate(wmo = 6904240, date = argoJuldToTime(juld)) |> filter(PhaseName == 'NPAR')
+taxo_6904241 <- vroom::vroom('DATA/ARGO/6904241_TAXO_FromNetCDF.csv') |> mutate(wmo = 6904241, date = argoJuldToTime(juld)) |> filter(PhaseName == 'NPAR')
+
+all_taxo <- rbind(taxo_1902578, taxo_4903634, taxo_6904240, taxo_6904241) |> mutate(wmo = factor(wmo, levels = c('6904240','4903634','6904241','1902578')))
+
+# test taxo -> pas forcément concluant -> à creuser..
+#ggplot(taxo_1902578 |> filter(cycle == 3)) + geom_point(aes(x = concentration, y = -depth))
+```
+
+```{r, spiciness, fig.width=6, fig.height=10}
 #| label: fig-spice
-#| fig-cap: Spiciness (up) and particle concentrations for the 102-128 $\mu$m size class.
-
-#ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26, park_depth != '200 m') |> dplyr::mutate(date = as_date(date))
-# ts_data <- all_floats |> filter(PhaseName == 'PAR') |> select(date = juld, depth = pres, temp, psal, wmo, sigma, park_depth, lon, lat) |> drop_na(temp) |> filter(sigma > 26) |> 
-#   dplyr::mutate(date = as_date(date), colour = if_else(park_depth == '200 m', '#E31B23', if_else(park_depth == '500 m', '#FFC325', 'black')))
+#| fig-cap: Potential density anomaly, spiciness, *Calanoida* concentrations (log$_{10}$ colorscale), spectral slope and particle concentrations for the 102-128 $\mu$m size class (log$_{10}$ colorscale) for each float.
 
@@
 
-# # https://dankelley.github.io/oce/reference/plotTS.html
-# # https://dankelley.github.io/oce/reference/argo-class.html
-# test <- as.argo(time = ts_data$date, longitude = ts_data$lon, latitude = ts_data$lat, salinity = ts_data$psal, temperature = ts_data$temp, pressure = ts_data$depth,
-#         id = ts_data$wmo) 
-# 
-# plotTS(test, inSitu = T, type = 'p', xlab = 'Salinity [PSU]')
-# 
-# 
-# ggplot(ts_data) + geom_point(aes(x = spiciness, y = depth)) + scale_y_reverse() + facet_wrap(~wmo)
-
-a <- ggplot(ts_data |> filter(depth <= 1000)) + geom_point(aes(x=date, y=depth, colour=spiciness), size = 3, shape = 15) +
+density <- ggplot(ts_data |> filter(depth <= 1000)) + geom_point(aes(x=date, y=depth, colour=sigma), size = 3, shape = 15) +
   scale_y_reverse() +
   #scale_color_viridis() +
-  scale_color_cmocean(name = 'balance', limits = c(-0.25,0.25), oob = scales::squish, na.value = 'grey') +
+  scale_color_cmocean(name = 'dense', limits = c(27,28), oob = scales::squish, na.value = 'grey') +
+  theme_bw() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = TeX('Sigma (kg m$^{-3}$)    '), x = 'Month', y = 'Depth (m)') +
+  facet_wrap(~wmo, ncol = 4) + theme(axis.title.x = element_blank())
+
+spiciness <- ggplot(ts_data |> filter(depth <= 1000)) + geom_point(aes(x=date, y=depth, colour=spiciness), size = 3, shape = 15) +
+  scale_y_reverse() +
+  #scale_color_viridis() +
+  scale_color_cmocean(name = 'balance', limits = c(-0.5,0.5), oob = scales::squish, na.value = 'grey') +
   theme_bw() +
@@
   labs(colour = TeX('Spiciness (kg m$^{-3}$)'), x = 'Month', y = 'Depth (m)') +
-  facet_wrap(~wmo, ncol = 4)
-
-b <- particles_bis |> filter(depth <= 1000, size == 102) |>
+  facet_wrap(~wmo, ncol = 4) + theme(axis.title.x = element_blank())
+
+part102 <- particles_bis |> filter(depth <= 1000, size == 102) |>
   #group_by(size) |>
@@
   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-  labs(colour = 'Particles (#/L)', x = 'Month', y = 'Depth (m)')
-
-a/b
-```
-
-It is also worth noting that the progressive emptying of the water column does not start from the surface in small classes [@fig-uvp6-profiling-102-406].
-By computing the spiciness [@fig-spice], we show that this pattern does not likely result from water subduction (no spiciness anomalies) hence it would be support the hypothesis of the particle export ending.
+  labs(colour = 'Particles (#/L)       ', x = 'Month', y = 'Depth (m)')
+
+taxo_plot <- all_taxo |> filter(depth <= 1000) |> mutate(date = as.Date(date)) |> filter(date < as.Date("2023-06-03")) |>
+  #group_by(size) |>
+  ggplot() + facet_wrap(~wmo, ncol = 4) + theme_bw() +
+  geom_point(aes(x=date, y=depth, colour=log10(concentration)), size=1, shape=15) +
+  scale_color_cmocean(name = 'thermal', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = 1) +
+  scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = 'Calanoida (#/L)    ', x = 'Month', y = 'Depth (m)') + theme(axis.title.x = element_blank())
+
+spectral_slope_plot <- vertical_profiles |> filter(depth <= 1000) |> mutate(wmo = factor(wmo, levels = c('6904240','4903634','6904241','1902578'))) |>
+  ggplot() + facet_wrap(~wmo, ncol = 4) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=spectral_slope), size=2, shape=15) +
+  scale_color_cmocean(name = 'thermal', limits = c(-4,-3), oob = scales::squish, na.value = 'grey', direction = 1) +
+  scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = 'Spectral slope     ', x = 'Month', y = 'Depth (m)') + theme(axis.title.x = element_blank())
+
+density / spiciness / taxo_plot / spectral_slope_plot / part102
+```
+
+Still, it is unclear why we observe a progressive emptying of the water column starting below 500 m for size classes below 323 $\mu$m [@fig-uvp6-profiling-102-406].
+This pattern could result from several causes.
+It could either be the signature of a subduction event [@Llort2018-pa] either the end of a slow sinking export (i.e latent dynamic of the system) or a sign of disaggregation processes by zooplankton (i.e.
+zooplankton feeding).
+Therefore, we looked at physical properties (density anomaly and spiciness) and well as particle characteristics (zooplankton concentration and spectral slope) to investigate this particularity (@fig-spice).
+Spiciness is a variable that allows the differentiation of water masses with distinct thermohaline properties but similar density [@Flament2002-zm], to look for the signature of subduction.
+@fig-spice shows that both the density field and the spiciness field are uniform below 500 m hence it is unlikely that a subduction event would be the reason for this decrease in particle abundance below 500 m.
+Using the embedded zooplankton classification results from the UVP6, we looked at concentrations of *Calanoida*, a species that is abundant in the NA [@Corkett1979-ha].
+@fig-spice shows that the concentration of *Calanoida* increase with depth similarly to the plume observed in @fig-uvp6-profiling-512-2500.
+However, their concentrations decrease with the advent of a poorer water column in terms of particles (see @fig-uvp6-profiling-102-406 and @fig-uvp6-profiling-512-2500).
+The last panel of that figure shows that the particle concentrations for the smallest size class had already slightly decreased when the the *Calanoida* were at their maximum (see below 500 m between August and November).
+This could correspond to zooplankton feeding where *Calanoida* would ingest small particles until food vanishes.
+The spectral slope defines the slope of the PSD where a high (in absolute values) spectral slope indicates that small particules dominate the PSD.
+The panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope at the beginning of the time series that is concurrent with *Calanoida* abundance at first order (except for float 4903634).
+Would this mean that *Calanoida* feed on small particles and eject fecal pellets (via repackaging) that in turn would be the source of big particles observed in @fig-uvp6-profiling-512-2500?
+Such an hypothesis would solve the issue brought earlier by trying to link the surface to the big particle plume but at this stage, it is very difficult conclude without more data.
 
@@
 Second, the plume of big particles can be misleading because their abundance measured by the UVP6 are very low (usually \< 0.1 #/L).
+However, the plume could be the result of aggregation (repackaging) by zooplankton feeding on small particles.
 Third, all floats measured the same particle dynamics in distinct zones which suggests that if the two systems were initially different (i.e.
 different bloom dynamics), they merged back after the bloom of May.
-Therefore, we suggest that we mostly observe the latent dynamics of the system rather than clear export signatures that would likely be spotted with both higher bloom intensity and a higher profiling temporal resolution.
-
-### hplc pour les communautés pour montrer que les deux endroits sont différents (peripery vs central)
-
-mettre ça dans phenology of the bloom pour bien montrer que l'on est dans 2 systèmes différentes
-
-```{r, HPLC}
-
-pigments <- vroom('DATA/ARGO/CE22009_pigments_031122_Travail_Herve_FŔ.csv') |> filter(Station %in% c(29,9))
-
-# select interesting coumns (all QCs == 1 so OK)
-# pigments <- pigments |> select(station = Station, depth = `Depth_(m)`, Chla_c3 = Chlorophyll_c3, Chla_c1_c2_MgDVP = `Chlorophyll_c1+c2+MgDVP`, Phaeophorbid_a, Peridinin, `19'-Butanoyloxyfucoxanthin`
-```
+Therefore, we suggest that we mostly observe the latent dynamics of the system rather than clear export signatures that would likely be spotted with both higher bloom intensities (e.g.
+the May bloom that was missed) and a higher profiling temporal resolution.
 
@@
 
+#TO OR NOT TO DO?
+
 ```
@@
 
--   science: on voit le système latent, on a raté le bloom
-
 -
@@
 -   on se trouve dans un sytème producitve entre guidi 2008 et iversen (très productif)
-
-    ### TS diagram
-
--   Pour le drift à 200, 500 et 1000 m -\> interpète ce graph -\> va checker les masses d'eau de la region...
-attention on est en PARKING DEPTH -\> mettre la couleur de la parking depth dans le truc, ça permettrait de voir si on peut comparer les processus (style ratio 500/1000m) dans les couches qui sont dans la meme masse d'eau.
 

### 2023-03-15 article3.qmd
update

@@
 # Case study in the Labrador Sea {#sec-case-study-labrador-sea}
+
+In the first chapter, I explained the different steps that led to the creation of the embedded zooplankton classification model for the UVP6.
+However, I did not specify that in spite of the technical and energetic constraints, the model was also due for April 2022 for a deployment of BGC-Argo floats in the Labrador Sea a month later.
+Once deployed, it was only a question of time before I could dig into the UVP6 and the OST data.
+In the meantime I asked myself, what could be a cool thing to do about those data?
+I already had some experience with BGC-Argo floats during my master's degree (study of the deep chlorophyll maximum in the Black Sea, see also the abstract of that paper in APPENDIX XX) and because I was a bit bored waiting for new BGC-Argo profiles, I thought it would be useful to have a visualization tool in the form of a web app to investigate incoming data and already prepare the ground for the following study.
+Here is the link to the [web app](https://fricour.shinyapps.io/CARBON_REFINE/), which is still currently being updated every night with new data (near) live from the Labrador Sea and hopefully with other BGC-Argo floats equipped with the UVP6 and the OST to investigate carbon and particle fluxes in other regions.
+This study uses preliminary data (not even one year of data) and a new BGC-Argo float configuration hence more work should be addressed to better understand the dynamics of the system.
 
@@
 continuous deposition of small particles on the upward-facing window of the transmissometer) and discontinuous (i.e.
-deposition of stochastic big particles) components of the downward POC flux, hereafter referred to as, respectively, $F_{small}$ and $F_{big}$.
+deposition of stochastic large particles) components of the downward POC flux, hereafter referred to as, respectively, $F_{small}$ and $F_{large}$.
 
@@
 standard score) of each of them.
-The detection of jumps therefore consisted in spotting outliers in the z-score distribution.
+The detection of jumps therefore consisted in spotting outliers in the slope distribution.
 For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles.
-Positive and negative jumps were defined as, respectively, a z-score above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$.
+Positive and negative jumps were defined as, respectively, a slope above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$.
 We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.
@@
 
-We used the same relationship to compute $F_{big}$ by replacing $slope_{w}$ as the sum of positive jumps divided by the drifting time.
+We used the same relationship to compute $F_{large}$ by replacing $slope_{w}$ as the sum of positive jumps divided by the drifting time.
 
@@
 
-# add big particles flux
+# add large particles flux
 rows_to_keep <- c(jump_index, jump_index-1) 
@@
 
-# big particles flux
+# large particles flux
 if(is.null(tmp3)){ # no jumps
-  big_part_poc_flux <- 0
+  large_part_poc_flux <- 0
 }else{
@@
   if(nrow(tmp4) == 0){ # no positive jumps
-    big_part_poc_flux <- 0
+    large_part_poc_flux <- 0
   }else{
@@
     delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
-    slope_big_part <- delta_y/delta_x
-    big_part_poc_flux <- 633*(slope_big_part**0.77)
+    slope_large_part <- delta_y/delta_x
+    large_part_poc_flux <- 633*(slope_large_part**0.77)
   }
@@
 
+  # IQR <- quantile(tmp$slope, probs = 0.75, na.rm=T) - quantile(tmp$slope, probs = 0.25, na.rm=T)
+  # # outliers ('spikes' in the Z score signal)
+  # spikes_down <- tmp$slope < quantile(tmp$slope, 0.25, na.rm=T) - 1.5 *IQR
+  # spikes_up <-  tmp$slope > quantile(tmp$slope, 0.75, na.rm=T) + 1.5 *IQR
+  # spikes <- as.logical(spikes_down + spikes_up)
+
   # assign spikes
@@
 
-  # add big particles flux to the party
+  # add large particles flux to the party
   rows_to_keep <- c(jump_index, jump_index-1)
@@
   if(is.null(tmp2)){ # no jump
-    big_part_poc_flux <- 0
+    large_part_poc_flux <- 0
     tmp3 <- NULL
@@
     if(nrow(tmp3) == 0){ # no positive jumps
-      big_part_poc_flux <- 0
+      large_part_poc_flux <- 0
     }else{
@@
       delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
-      slope_big_part <- delta_y/delta_x
-      big_part_poc_flux <- 633*(slope_big_part**0.77)
+      slope_large_part <- delta_y/delta_x
+      large_part_poc_flux <- 633*(slope_large_part**0.77)
     }
@@
                          'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
-                         'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1'), yaxis = list(title = 'Cp (1/m)'), xaxis = list(title = 'Time'))
+                         'POC flux (large particles): ', round(large_part_poc_flux,1), ' mg C m-2 day-1'), yaxis = list(title = 'Cp (1/m)'), xaxis = list(title = 'Time'))
 
@@
 
-    # adapt script to return big and small flux
-  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],
+    # adapt script to return large and small flux
+  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'large_flux' = large_part_poc_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],
                cycle = data$cycle[1])
@@
 
-```{r, compute Fsmall and F big}
+```{r, compute Fsmall and F large}
 
@@
                                                                 std_smallf = sd(small_flux, na.rm=T),
-                                                                min_bigf = min(big_flux, na.rm=T), max_bigf = max(big_flux, na.rm=T),
-                                                                mean_bigf = mean(big_flux, na.rm=T), median_bigf = median(big_flux, na.rm=T),
-                                                                std_bigf = sd(big_flux, na.rm=T))
+                                                                min_largef = min(large_flux, na.rm=T), max_largef = max(large_flux, na.rm=T),
+                                                                mean_largef = mean(large_flux, na.rm=T), median_largef = median(large_flux, na.rm=T),
+                                                                std_largef = sd(large_flux, na.rm=T))
 
@@
 
-$F_{small}$ and $F_{big}$ were computed at each drifting depth for all floats (@fig-OST-flux).
+$F_{small}$ and $F_{large}$ were computed at each drifting depth for all floats (@fig-OST-flux).
 At 200 m, $F_{small}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ (average $\pm$ standard deviation).
@@
 
-```{r, Fsmall and Fbig, fig.width=8, fig.height=6}
+```{r, Fsmall and Flarge, fig.width=8, fig.height=6}
 #| label: fig-OST-flux
-#| fig-cap: OST-derived small (top) and big (bottom) particles carbon flux for all floats at each drifting depth.
+#| fig-cap: OST-derived small (top) and large (bottom) particles carbon flux for all floats at each drifting depth.
 Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.
@@
 # add black data to have the same x axis for all drifting depths (to show that we stopped the drifting at 200 and 500 m)
-blank_data <- tibble(max_time = c(NA, NA), min_time = c(NA,NA), small_flux = c(NA,NA), big_flux = c(NA,NA), park_depth = c('200 m', '500 m'),
+blank_data <- tibble(max_time = c(NA, NA), min_time = c(NA,NA), small_flux = c(NA,NA), large_flux = c(NA,NA), park_depth = c('200 m', '500 m'),
                      wmo = rep(4903634,2), cycle = c(NA,NA), drifting_time = c(NA,NA), WMO = rep(as.factor(4903634),2), date = rep('2023-03-05',2))
@@
 
-down <- tmp |> ggplot(aes(x = date, y = big_flux)) +
-  geom_point(data = tmp, aes(x = date, y = big_flux, colour = WMO, shape = WMO)) +
-  geom_smooth(data = tmp, aes(x = date, y = big_flux), colour = 'black') +
+down <- tmp |> ggplot(aes(x = date, y = large_flux)) +
+  geom_point(data = tmp, aes(x = date, y = large_flux, colour = WMO, shape = WMO)) +
+  geom_smooth(data = tmp, aes(x = date, y = large_flux), colour = 'black') +
   scale_color_brewer(palette = 'Dark2') + 
@@
   facet_wrap(~park_depth, scales = 'free') +
-  labs(x = 'Month', y = TeX('$F_{big}$ (mg C m$^{-2}$ day$^{-1}$)')) +
+  labs(x = 'Month', y = TeX('$F_{large}$ (mg C m$^{-2}$ day$^{-1}$)')) +
   theme_bw() +
@@
 
-$F_{big}$ was estimated at 200 m between 0 (no deposition of big particles on the transmissometer window) and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$.
+$F_{large}$ was estimated at 200 m between 0 (no deposition of large particles on the transmissometer window) and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$.
 High fluxes (\> 100 mg C m$^{-2}$ day$^{-1}$) were measured by 3 out of 4 floats at the end of May before a sharp decrease towards July.
-At 500 m, $F_{big}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to February.
-At 1000 m, $F_{big}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards March, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$.
+At 500 m, $F_{large}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_largef, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to February.
+At 1000 m, $F_{large}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards March, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$.
 
@@
 size class 102, float 6904240).
-All floats also measured a local particle abundance maximum around 800 - 1000 m whose origin could be due to a previous particle export or remineralization of bigger particles.
+All floats also measured a local particle abundance maximum around 800 - 1000 m whose origin could be due to a previous particle export or remineralization of largeger particles.
 
@@
 
-```{r, UVP6 profiling mode without sinking speed big classes, fig.height=8, fig.width=8}
+```{r, UVP6 profiling mode without sinking speed large classes, fig.height=8, fig.width=8}
 #| label: fig-uvp6-profiling-512-2500
@@
 @fig-chla-mld-satellite shows that the satellite MLD in that region had already shoaled below 100 m since March and that the water column was stable and relatively well stratified in May which might explain the intensity of surface Chla concentrations in contrast with those measured at the second deployment site.
-The two other floats were deployed in the bioregion where the phytoplankton biomass typically develops slowly in May to reach its maximum in June [@Lacour2015-da].
+The latter is located in the bioregion where the phytoplankton biomass typically develops slowly in May to reach its maximum in June [@Lacour2015-da].
 However, the Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD at the end of May and that it started again in mid-June when the water column was well stratified.
@@
 
-Here, we refer to the first and second deployment as, respectively, station A and station B.
+Here, we refer to the first and second deployment sites as, respectively, station A and station B.
 High Performance Liquid Chromatography (HPLC) analyses were performed on samples taken during the two deployments.
@@
 microphytoplankton) whereas station B was composed of a majority of smaller phytoplankton species (mostly nano and picozooplankton).
-This supports the idea of two distinct bloom dynamics as inferred from satellite data.
+This supports the idea of two distinct bloom dynamics as inferred from float and satellite data.
 
@@
   facet_wrap(~pigment, scales = 'free') + scale_y_reverse() +
-  labs(colour = 'Station', x = TeX('Concentration (mg m$^{-3}$)'), y = 'Depth (m)') +
+  labs(colour = '', x = TeX('Concentration (mg m$^{-3}$)'), y = 'Depth (m)') +
   theme_bw() + scale_color_manual(values = c('#E31B23', '#003366'))
@@
 
-Even though we suggested that the two pairs of floats were deployed in two different bloom regimes, they all measured an increase in particle concentrations in the 102 - 512 $\mu$m range (@fig-uvp6-profiling-102-406) around 800 - 1000 m as well as an increase in the abundance of big particles (@fig-uvp6-profiling-512-2500) between June and August.
+Even though we suggested that the two pairs of floats were deployed in two different bloom regimes, they all measured an increase in particle concentrations in the 102 - 512 $\mu$m range (@fig-uvp6-profiling-102-406) around 800 - 1000 m as well as an increase in the abundance of large particles (@fig-uvp6-profiling-512-2500) between June and August.
 This increase raises the question of a possible particle export event.
-For classes \> 512 $\mu$m, the vertical concentration profiles highlight a bigger particle plume for the 6904241-1902578 pair than for the other one.
+For classes \> 512 $\mu$m, the vertical concentration profiles highlight a largeger particle plume for the 6904241-1902578 pair than for the other one.
 However, apart from the float 6904240, all floats measured similar integrated Chla content in the upper 100 m and in the mixed layer (@fig-chla-content).
-@fig-OST-flux also shows that $F_{small}$ is similar across all floats at all parking depths and that $F_{big}$ is also similar across all floats at 1000 m with the same increase (from June to August) than for F$_{small}$.
+@fig-OST-flux also shows that $F_{small}$ is similar across all floats at all parking depths and that $F_{large}$ is also similar across all floats at 1000 m with an increase from June to August as observed in F$_{small}$.
 The same behavior is observed for UVP6 particles in the lower size range at 1000 m (@fig-uvp-1000m).
@@
 # test taxo -> pas forcément concluant -> à creuser..
-#ggplot(taxo_1902578 |> filter(cycle == 3)) + geom_point(aes(x = concentration, y = -depth))
+#ggplot(taxo_6904241 |> filter(cycle == 3)) + geom_point(aes(x = conc_calanoid, y = -depth))
 ```
@@
   ggplot() + facet_wrap(~wmo, ncol = 4) + theme_bw() +
-  geom_point(aes(x=date, y=depth, colour=log10(concentration)), size=1, shape=15) +
+  geom_point(aes(x=date, y=depth, colour=log10(conc_calanoid)), size=1, shape=15) +
   scale_color_cmocean(name = 'thermal', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = 1) +
@@
 Therefore, we looked at physical properties (density anomaly and spiciness) and well as particle characteristics (zooplankton concentration and spectral slope) to investigate this particularity (@fig-spice).
-Spiciness is a variable that allows the differentiation of water masses with distinct thermohaline properties but similar density [@Flament2002-zm], to look for the signature of subduction.
+Spiciness is a variable that allows the differentiation of water masses with distinct thermohaline properties but similar density [@Flament2002-zm].
 @fig-spice shows that both the density field and the spiciness field are uniform below 500 m hence it is unlikely that a subduction event would be the reason for this decrease in particle abundance below 500 m.
 Using the embedded zooplankton classification results from the UVP6, we looked at concentrations of *Calanoida*, a species that is abundant in the NA [@Corkett1979-ha].
-@fig-spice shows that the concentration of *Calanoida* increase with depth similarly to the plume observed in @fig-uvp6-profiling-512-2500.
-However, their concentrations decrease with the advent of a poorer water column in terms of particles (see @fig-uvp6-profiling-102-406 and @fig-uvp6-profiling-512-2500).
-The last panel of that figure shows that the particle concentrations for the smallest size class had already slightly decreased when the the *Calanoida* were at their maximum (see below 500 m between August and November).
-This could correspond to zooplankton feeding where *Calanoida* would ingest small particles until food vanishes.
+@fig-spice shows that the concentration of *Calanoida* is higher from the end of May to July in the surface layer.
+There seems to be a lower patch between 250 and 500 m from August to November however it is difficult to estimate their impact on particle concentrations at this stage.
+Finally, we computed the spectral slope to look for aggregation/fragmentation processes.
 The spectral slope defines the slope of the PSD where a high (in absolute values) spectral slope indicates that small particules dominate the PSD.
-The panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope at the beginning of the time series that is concurrent with *Calanoida* abundance at first order (except for float 4903634).
-Would this mean that *Calanoida* feed on small particles and eject fecal pellets (via repackaging) that in turn would be the source of big particles observed in @fig-uvp6-profiling-512-2500?
-Such an hypothesis would solve the issue brought earlier by trying to link the surface to the big particle plume but at this stage, it is very difficult conclude without more data.
+Here, the panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope at the beginning of the time series but it is also difficult to attach it to biological processes such as zooplankton feeding without additional data.
 
@@
 First, the 10-day profiling configuration hinders the characterization of particle export events that are transient processes.
-Second, the plume of big particles can be misleading because their abundance measured by the UVP6 are very low (usually \< 0.1 #/L).
+Second, the plume of large particles can be misleading because their abundance measured by the UVP6 are very low (usually \< 0.1 #/L).
 However, the plume could be the result of aggregation (repackaging) by zooplankton feeding on small particles.
@@
 
-For the first time, BGC-Argo floats equipped with both a transmissometer and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx].
-The quantification of the POC flux using the OST method [@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr] is based on an empirical POC:ATN ratio \[REF estapa 2022\] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:
+#### Carbon flux trends and estimations
+
+For the first time, BGC-Argo floats equipped with both an OST and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx].
+The quantification of the POC flux using the OST method [@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk; @Estapa2023-oz] is based on an empirical POC:ATN ratio [@Estapa2023-oz] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:
 
@@
 
-where $C_{i}$, A and b represent, respectively, the concentration of particles (number/L) for the $i^{th}$ class of mean diameter $d_{i}$ (in mm) and 2 constants with $b$ being related to the fractal dimension ($D$) of an aggregate by $D = (b +1)/2$ [@Guidi2008-vp].
+where $C_{i}$, A and b represent, respectively, the concentration of particles (#/L) for the $i^{th}$ class of mean diameter $d_{i}$ (in mm) and 2 constants with $b$ being related to the fractal dimension ($D$) of an aggregate by $D = (b +1)/2$ [@Guidi2008-vp].
 
@@
 
-```{r, compute guidi fluxes at parking depth, fig.height=6, fig.width=6}
+```{r, compute guidi fluxes at parking depth, fig.height=6, fig.width=8}
 #| label: fig-comparison-ost-uvp
-#| fig-cap: Comparison of total carbon fluxes derived from the OST (black) and the UVP6 (red) where the error bars represent the 25$^{th}$ and 75$^{th}$ percentiles around the median.
+#| fig-cap: Comparison of total carbon fluxes (F$_{small}$ + F$_{large}$) derived from the OST (black) and the UVP6 using the A and b values of @Guidi2008-vp (red) and @Iversen2010-an (yellow) in @eq-guidi-article3.
+OST data from float 6904241 stops in November due to a transmissometer failure.
 
 compute_flux <- function(A, b, PSD){ # PSD = Particle Size Distribution
-  # for classes sizes between 0.102 mm and 1.5 mm
-  #mid_ESD <- c(0.1164512, 0.1463600, 0.1843915, 0.2325200, 0.2933257, 0.3691650, 0.4650400, 0.5860455, 0.7385533, 0.9280422, 1.1705684, 1.4795321) # With Rainer's code (paruvpy)
-  # mid_ESD <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
-  #              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
-  #              1.16170742, 1.46365963) # With Lionel's script
-  # mid_ESD <- c(0.23051195, 0.29042685,
-  #              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
-  #              1.16170742, 1.46365963) # With Lionel's script
-  #mid_ESD <- c(0.287555212,0.362129811,0.455929819,0.57466512,0.724144323,0.91063714,1.147083258,1.450068964)
+  # for classes sizes between 0.25 mm and 1.5 mm
   mid_ESD <- c(0.29042685,
@@
 # plot it
-comparison_flux_ost_uvp <- merge(parking_data, cflux) |> dplyr::mutate(total_flux = small_flux + big_flux)
+comparison_flux_ost_uvp <- merge(parking_data, cflux) |> dplyr::mutate(total_flux = small_flux + large_flux)
 
@@
 # https://stackoverflow.com/questions/41077199/how-to-use-r-ggplot-stat-summary-to-plot-median-and-quartiles
-ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = total_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
+ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = total_flux), colour = 'black') + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
   geom_errorbar(aes(x = date, y = guidi_flux),
@@
                       #fun.max = function(z) {quantile(z, 0.75)},
-                      fun = median, colour = '#003366', width = 5) +
+                      fun = median, colour = '#FFC325', width = 5) + # #003366
     geom_line(aes(x = date, y = iversen_flux),
                       stat = 'summary',
-                      fun = median, colour = '#003366') + 
+                      fun = median, colour = '#FFC325') + 
   scale_y_continuous(trans = 'log10') + theme_bw() + labs(x = 'Month', y = TeX('$F_{total}$ (mg C m$^{-2}$ day$^{-1}$)')) +
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month')
-
-```
-
-```{r, and on the vertical?}
-
-#TO OR NOT TO DO?
-
-```
-
-ici je compare finalement un système ultra productif (iversen) et un système pas produtif (lionel) -\> on se situerait entre les deux.
-Après je peux fournir els valeurs que j'ai trouvées avec le fminsearch avec la fonctiono optimize depython (21.98 et 1.18.
-Mais bon, cela assumerait que 'lon pourrait prendre l'OST comme référence, ce qui est un peu too much..)
+```
 
@@
 #| label: tbl-correlation
-#| tbl-cap: TODO
-correlations <- comparison_flux_ost_uvp |> dplyr::select(park_depth, wmo, cycle, guidi_flux, small_flux, big_flux, date, total_flux) |> dplyr::group_by(wmo, park_depth, cycle) |> dplyr::mutate(median_guidi_flux = median(guidi_flux)) |> dplyr::group_by(park_depth, wmo) |> dplyr::summarize(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_big = round(cor(median_guidi_flux, big_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))
-
-correlations <- correlations |> pivot_wider(names_from = park_depth, values_from = c(cor_small, cor_big, cor_total))
-
-gt(correlations) |> tab_spanner(label = "Small flux", columns = 2:4) |> tab_spanner(label = "Big flux", columns = 5:7) |> tab_spanner(label = 'Total flux', columns = 8:10) |>
+#| tbl-cap: Pearson correlations between UVP6 PSD-derived carbon fluxes and total, large particles and small particles carbon flux measured by the OST at 3 drifting depths.
+correlations <- comparison_flux_ost_uvp |> dplyr::select(park_depth, wmo, cycle, guidi_flux, small_flux, large_flux, date, total_flux) |> dplyr::group_by(wmo, park_depth, cycle) |> dplyr::mutate(median_guidi_flux = median(guidi_flux)) |> dplyr::group_by(park_depth, wmo) |> dplyr::summarize(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_large = round(cor(median_guidi_flux, large_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))
+
+correlations <- correlations |> pivot_wider(names_from = park_depth, values_from = c(cor_small, cor_large, cor_total))
+
+gt(correlations) |> tab_spanner(label = "Small flux", columns = 2:4) |> tab_spanner(label = "Large flux", columns = 5:7) |> tab_spanner(label = 'Total flux', columns = 8:10) |>
   cols_label(wmo = 'WMO', `cor_small_200 m` = '200 m',
@@
              `cor_small_1000 m` = '1000 m',
-             `cor_big_200 m` = '200 m',
-             `cor_big_500 m` = '500 m',
-             `cor_big_1000 m` = '1000 m',
+             `cor_large_200 m` = '200 m',
+             `cor_large_500 m` = '500 m',
+             `cor_large_1000 m` = '1000 m',
              `cor_total_200 m` = '200 m',
@@
              `cor_total_1000 m` = '1000 m')
-
-```
-
-Figure @fig-comparison-ost-uvp shows a comparison of carbon fluxes estimated with the OST ($F_{total} = F_{small} + F_{big}$) and the UVP6 (median of all UVP measurements while the float was drifting) at all drifting depths.
-Overall, carbon fluxes estimated with the coefficients ($A$, $b$) of @Guidi2008-vp, are lower than the ones measured by the OST.
+```
+
+@fig-comparison-ost-uvp shows the comparison between the total carbon flux measured by the OST ($F_{total} = F_{small} + F_{large}$) and the UVP6 (median of all UVP measurements while the float was drifting) at all drifting depths.
+Overall, carbon fluxes estimated with the $A$ and $b$ coefficients of @Guidi2008-vp are lower than the ones measured by the OST whereas the same estimation using the coefficient of @Iversen2010-an generally overestimates the OST at 200 m but is of the same order at 500 m and 1000 m.
 Without additional measurements (e.g.
-co-deployed sediment traps), it is however impossible to decipher which measurements is closest to the truth.
-Nonetheless, the trends measured by both instruments are very well correlated at all drifting depths for floats 6904240 and 6904241 with correlations comprised between `r min(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])` and `r max(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])`.
-@tbl-correlation also shows that the correlation between UVP-derived carbon fluxes and $F_{small}$ and $F_{big}$ is `r mean(as.vector(as.matrix(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,2:7])))` on average.
-We calculated lower correlations with the other two floats, especially the float 1902578 at 200 m, where correlations between the UVP-derived carbon fluxes and all fluxes (small, big and total) are close to 0.
-
-Floats equipped with both a transmissometer and a UVP6 represent an opportunity to look at the influence of parts of the size spectrum on $F_{small}$ and $F_{big}$.
-For this, we decided to build a regression model using XGBoost, a gradient boosting method \[REF\], using biovolumes measured by the UVP to predict the slope of $F_{small}$ and $F_{big}$ measured by the OST.
+co-deployed sediment traps), it is however impossible to decipher which measurement is closest to the truth.
+Nonetheless, carbon fluxes measured at 1000 m by the OST and the UVP6 with the coefficients of @Iversen2010-an are both well correlated (\> 0.6 for all floats) and of similar magnitudes while carbon fluxes using @Guidi2008-vp are one order of magnitude lower.
+In their paper, @Iversen2010-an proposed that the carbon fluxes discrepancy could be due to different biogeochemical regimes where the relationship of @Guidi2008-vp was based on sediment traps at open ocean sites with lower carbon fluxes [@Honjo2008-kn] while the study site of @Iversen2010-an was located in a highly productive system at the continental margin off Cape Blanc in Mauritania.
+The LS is productive during the spring bloom [@Lacour2015-da] therefore it is coherent to find a closer relationship with estimations based on a highly productive system.
+
+The trends of total flux measured by both instruments are very well correlated at all drifting depths for floats 6904240 and 4903634 with correlations (see @tbl-correlation) ranging from `r min(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])` to `r max(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])`.
+@tbl-correlation also shows that the correlation between UVP-derived carbon fluxes and $F_{small}$ and $F_{large}$ is `r round(mean(as.vector(as.matrix(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,2:7]))),2)` on average.
+We calculated lower correlations with the other two floats, especially the float 1902578 at 200 m, where correlations between the UVP-derived carbon fluxes and all fluxes (small, large and total) are close to 0.
+Overall, it nevertheless suggests that there is a possibility to get more accurate estimations of carbon fluxes using PSD only therefore using only the particle size as an estimator of carbon flux, contrarily to what other studies have argued \[@Iversen2020-kb; @McDonnell2010-un; @Iversen2023-qj\].
+
+#### Nature of particles seen by the OST and the UVP6
+
+F$_{small}$ represents the continuous accumulation of small particles hence it could be related to the accumulated biovolume of particles seen by the UVP6.
+Similarly, $F_{large}$ could be related to the biovolume seen by the UVP6 when the OST detects jumps in the attenuation signal.
+Consequently, we firstly computed all the subslopes (see lightblue straight lines in @fig-estapa-method_article) and all the jumps at all drifting depths as well as the total biovolume seen by the UVP6 while the float was drifting.
+Then, we computed the average biovolume seen during each subslope duration and we associated the closest total biovolume (in time) with a jump.
 
@@
 
-get_jump_info <- function(data){
-
-  # make sure that the cp signal is in chronological order
-  tmp <- data |> arrange(dates)
-
-  # despike cp data with a 7-point moving window
-  tmp$cp <- despike(tmp$cp, k = 3)
-
-  # smooth cp data with a 3-point moving median, n time(s)
-  tmp$cp <- slide(tmp$cp, fun = median, k = 3, n = 1, na.rm=T)
-
-  # compute slope between two adjacent points (except first point) # we could start after 1h to let the float stabilize
-  delta_x <- as.numeric(tmp$dates - lag(tmp$dates), units = 'days')
-  delta_y <- tmp$cp - lag(tmp$cp)
-  tmp$slope <- delta_y/delta_x
-
-  # compute a Z score (assuming a normal distribution of the slopes) on the slopes
-  tmp <- tmp |> dplyr::mutate(zscore = (slope - mean(slope, na.rm = T))/sd(slope, na.rm = T))
-
-  # spot outliers on the Z score signal
-  # interquartile range between Q25 and Q75 -> had to used that and not the despike function because slopes are often close (or equal) to 0 so it can miss clear jumps.
+# get_jump_info <- function(data){
+#   
+#   # make sure that the cp signal is in chronological order
+#   tmp <- data |> arrange(dates)
+#   
+#   # despike cp data with a 7-point moving window
+#   tmp$cp <- despike(tmp$cp, k = 3)
+#   
+#   # smooth cp data with a 3-point moving median, n time(s)
+#   tmp$cp <- slide(tmp$cp, fun = median, k = 3, n = 1, na.rm=T)
+#   
+#   # compute slope between two adjacent points (except first point) # we could start after 1h to let the float stabilize
+#   delta_x <- as.numeric(tmp$dates - lag(tmp$dates), units = 'days')
+#   delta_y <- tmp$cp - lag(tmp$cp)
+#   tmp$slope <- delta_y/delta_x
+#   
+#   # compute a Z score (assuming a normal distribution of the slopes) on the slopes
+#   tmp <- tmp |> dplyr::mutate(zscore = (slope - mean(slope, na.rm = T))/sd(slope, na.rm = T))
+#   
+#   # spot outliers on the Z score signal
+#   # interquartile range between Q25 and Q75 -> had to used that and not the despike function because slopes are often close (or equal) to 0 so it can miss clear jumps.
 Q25 and Q75 are more trustworthy in this case than the despike function of Jean-Olivier (see package castr on his github: https://github.com/jiho/castr)
-  IQR <- quantile(tmp$zscore, probs = 0.75, na.rm=T) - quantile(tmp$zscore, probs = 0.25, na.rm=T)
-  # outliers ('spikes' in the Z score signal)
-  spikes_down <- tmp$zscore < quantile(tmp$zscore, 0.25, na.rm=T) - 1.5 *IQR
-  spikes_up <-  tmp$zscore > quantile(tmp$zscore, 0.75, na.rm=T) + 1.5 *IQR
-  spikes <- as.logical(spikes_down + spikes_up)
-
-  # assign spikes
-  tmp$spikes <- spikes
-
-  # assign colour code to cp signal
-  tmp$colour <- 'base signal' # base signal = smoothed despiked cp signal
-  tmp[which(tmp$spikes == TRUE),]$colour <- 'jump'
-
-  # add group to compute the slope of each group of points, separated by a jump
-  tmp$group <- NA
-
-  # index of jumps in the array
-  jump_index <- which(tmp$colour == 'jump')
-
-  # assign group identity to each group of points, separated by a jump (= subgroup)
-  for (i in jump_index){
-    for (j in 1:nrow(tmp)){
-      if ((j < i) & (is.na(tmp$group[j]))){
-        tmp$group[j] <- paste0('group_',i)
-      }
-    }
-  }
-  tmp$group[which(is.na(tmp$group))] <- 'last_group'
-
-  # compute slope for each subgroup
-  slope_df <- tmp |> filter(colour == 'base signal', slope != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), 
-                                                                                                                  nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],
-                                                                                                                  delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),
-                                                                                                                  delta_y = (last_cp-first_cp)*0.25, slope = delta_y/delta_x) # *0.25 to convert cp to ATN
-
-  # remove negative slope from the mean slope (no physical meaning)
-  slope_df <- slope_df |> filter(slope > 0)
-
-  # remove if only one point (cannot fit a slope with one point)
-  slope_df <- slope_df |> filter(nb_points > 1)
-
-  # compute weighted average slope (to take into account the fact that some subgroups might have 2 points and a high slope vs. large group of points with a small slope)
-  mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)
-
-  # convert cp to POC using Estapa's relationship 
-  poc_flux <- 633*(mean_slope**0.77) # /!\ slope computed for ATN on y axis (delta_y *0.25 because ATN = cp*0.25) -> should be OK
-
-  # build dataframe to plot each subgroup
-  part1 <- slope_df |> dplyr::select(group, time = min_time, cp = first_cp)
-  part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
-  part_slope <- rbind(part1, part2)
-
-  # spot negative jump
-  tmp$colour[which((tmp$colour == 'jump') & (tmp$slope < 0))]  <- 'negative jump'
-
-  # add big particles flux to the party
-  rows_to_keep <- c(jump_index, jump_index-1)
-  tmp2 <- tmp[rows_to_keep,] |> select(dates, cp, slope, colour, group) |> arrange(dates)
-
-  # remove negative jumps, if any
-  check_colour <- unique(tmp2$colour)
-  if(length(check_colour) >= 2){ # there is a least a jump (positive or negative)
-    tmp2 <- tmp2 |> dplyr::mutate(diff_jump = cp - lag(cp)) 
-    even_indexes <- seq(2,nrow(tmp2),2)
-    tmp2 <- tmp2[even_indexes,]
-  }else{ # No jump
-    tmp2 <- NULL
-  }
-
-  if(is.null(tmp2)){ # no jump
-    big_part_poc_flux <- 0
-    tmp3 <- NULL
-  }else{
-    tmp3 <- tmp2 |> filter(diff_jump > 0) # remove negative
-    if(nrow(tmp3) == 0){ # no positive jumps
-      big_part_poc_flux <- 0
-    }else{
-      delta_y <- sum(tmp3$diff_jump) *0.25 # to get ATN (= cp*0.25)
-      max_time <- max(tmp$dates)
-      min_time <- min(tmp$dates)
-      delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
-      slope_big_part <- delta_y/delta_x
-      big_part_poc_flux <- 633*(slope_big_part**0.77)
-    }
-  }
-
-  # compute total drifting time
-  max_time <- max(tmp$dates)
-  min_time <- min(tmp$dates)
-  drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))
-
-  # to plot subgroups
-  part_slope_tmp <- part_slope |> dplyr::mutate(dates = time, colour = 'slope')
-
-  # adapt script to return big and small flux
-  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],
-               cycle = data$cycle[1])
-
-  # clean data to send
-  slope_df <- slope_df |> arrange(min_time) |> dplyr::mutate(min_drifting_time = df$min_time, max_drifting_time = df$max_time, small_flux = df$small_flux, big_flux = df$big_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1])
-  if(is.null(tmp3)){
-    big_jump <- tmp3
-  }else{
-     big_jump <- tmp3 |> dplyr::mutate(park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1]) 
-  }
-
-  return(list(slope_info = slope_df, jump_info = big_jump))
-}
-```
-
-TO DO HERE
+#   IQR <- quantile(tmp$zscore, probs = 0.75, na.rm=T) - quantile(tmp$zscore, probs = 0.25, na.rm=T)
+#   # outliers ('spikes' in the Z score signal)
+#   spikes_down <- tmp$zscore < quantile(tmp$zscore, 0.25, na.rm=T) - 1.5 *IQR
+#   spikes_up <-  tmp$zscore > quantile(tmp$zscore, 0.75, na.rm=T) + 1.5 *IQR
+#   spikes <- as.logical(spikes_down + spikes_up)
+#   
+#   # assign spikes
+#   tmp$spikes <- spikes
+#   
+#   # assign colour code to cp signal
+#   tmp$colour <- 'base signal' # base signal = smoothed despiked cp signal
+#   tmp[which(tmp$spikes == TRUE),]$colour <- 'jump'
+#   
+#   # add group to compute the slope of each group of points, separated by a jump
+#   tmp$group <- NA
+#   
+#   # index of jumps in the array
+#   jump_index <- which(tmp$colour == 'jump')
+#   
+#   # assign group identity to each group of points, separated by a jump (= subgroup)
+#   for (i in jump_index){
+#     for (j in 1:nrow(tmp)){
+#       if ((j < i) & (is.na(tmp$group[j]))){
+#         tmp$group[j] <- paste0('group_',i)
+#       }
+#     }
+#   }
+#   tmp$group[which(is.na(tmp$group))] <- 'last_group'
+#   
+#   # compute slope for each subgroup
+#   slope_df <- tmp |> filter(colour == 'base signal', slope != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), 
+#                                                                                                                   nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],
+#                                                                                                                   delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),
+#                                                                                                                   delta_y = (last_cp-first_cp)*0.25, slope = delta_y/delta_x) # *0.25 to convert cp to ATN
+#   
+#   # remove negative slope from the mean slope (no physical meaning)
+#   slope_df <- slope_df |> filter(slope > 0)
+#   
+#   # remove if only one point (cannot fit a slope with one point)
+#   slope_df <- slope_df |> filter(nb_points > 1)
+#   
+#   # compute weighted average slope (to take into account the fact that some subgroups might have 2 points and a high slope vs. large group of points with a small slope)
+#   mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)
+#   
+#   # convert cp to POC using Estapa's relationship 
+#   poc_flux <- 633*(mean_slope**0.77) # /!\ slope computed for ATN on y axis (delta_y *0.25 because ATN = cp*0.25) -> should be OK
+#   
+#   # build dataframe to plot each subgroup
+#   part1 <- slope_df |> dplyr::select(group, time = min_time, cp = first_cp)
+#   part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
+#   part_slope <- rbind(part1, part2)
+#   
+#   # spot negative jump
+#   tmp$colour[which((tmp$colour == 'jump') & (tmp$slope < 0))]  <- 'negative jump'
+#   
+#   # add large particles flux to the party
+#   rows_to_keep <- c(jump_index, jump_index-1)
+#   tmp2 <- tmp[rows_to_keep,] |> select(dates, cp, slope, colour, group) |> arrange(dates)
+#   
+#   # remove negative jumps, if any
+#   check_colour <- unique(tmp2$colour)
+#   if(length(check_colour) >= 2){ # there is a least a jump (positive or negative)
+#     tmp2 <- tmp2 |> dplyr::mutate(diff_jump = cp - lag(cp)) 
+#     even_indexes <- seq(2,nrow(tmp2),2)
+#     tmp2 <- tmp2[even_indexes,]
+#   }else{ # No jump
+#     tmp2 <- NULL
+#   }
+#   
+#   if(is.null(tmp2)){ # no jump
+#     large_part_poc_flux <- 0
+#     tmp3 <- NULL
+#   }else{
+#     tmp3 <- tmp2 |> filter(diff_jump > 0) # remove negative
+#     if(nrow(tmp3) == 0){ # no positive jumps
+#       large_part_poc_flux <- 0
+#     }else{
+#       delta_y <- sum(tmp3$diff_jump) *0.25 # to get ATN (= cp*0.25)
+#       max_time <- max(tmp$dates)
+#       min_time <- min(tmp$dates)
+#       delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
+#       slope_large_part <- delta_y/delta_x
+#       large_part_poc_flux <- 633*(slope_large_part**0.77)
+#     }
+#   }
+#   
+#   # compute total drifting time
+#   max_time <- max(tmp$dates)
+#   min_time <- min(tmp$dates)
+#   drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))
+#   
+#   # to plot subgroups
+#   part_slope_tmp <- part_slope |> dplyr::mutate(dates = time, colour = 'slope')
+#   
+#   # adapt script to return large and small flux
+#   df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'large_flux' = large_part_poc_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],
+#                cycle = data$cycle[1])
+#   
+#   # clean data to send
+#   slope_df <- slope_df |> arrange(min_time) |> dplyr::mutate(min_drifting_time = df$min_time, max_drifting_time = df$max_time, small_flux = df$small_flux, large_flux = df$large_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1])
+#   if(is.null(tmp3)){
+#     large_jump <- tmp3
+#   }else{
+#      large_jump <- tmp3 |> dplyr::mutate(park_depth = data$park_depth[1], wmo = data$wmo[1],cycle = data$cycle[1]) 
+#   }
+#   
+#   return(list(slope_info = slope_df, jump_info = large_jump))
+# }
+```
 
@@
 
-```{r, check total biovolume with slope value}
-
-uvp_drifting <- all_floats |> filter(PhaseName == 'PAR') |> select(juld, cycle, wmo, park_depth, all_of(lpm_class)) |> 
-  drop_na(NP_Size_102) |> dplyr::mutate(date = as_date(juld))
-
-# DSE for computing biovolumes
-#DSE <- c(0.29042685,0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779, 1.16170742, 1.46365963)
-DSE_long <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685, 0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
-                1.16170742, 1.46365963, 1.84409558, 2.32341484)
-
-# compute biovolume
-get_biovolume <- function(DSE){
-  r <- DSE/2
-  biovolume <- (4*pi/3)*r**3
-  return(biovolume)
-}
-
-# get biovolume for each DSE (equivalent to say for each size bin)
-biovolumes <- map_dbl(DSE_long, get_biovolume)  # unit = mm^3
-
-# compute biovolume and add them to the dataframe
-spectra_biovolume <- tibble(as.data.frame(t(t(uvp_drifting[,all_of(lpm_class)])*biovolumes)))
-colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('NP_Size', 'Biovolume')
-biovolumes_names <- colnames(spectra_biovolume)
-uvp_drifting <- cbind(uvp_drifting, spectra_biovolume) |> rowwise() |> dplyr::mutate(total_biovolume = sum(across(all_of(biovolumes_names))))
-
-assign_time_group <- function(d_uvp, d_ost){
-  d_uvp$group <- NA
-  for (i in 1:nrow(d_ost)){
-    low_time <- d_ost$min_time[i]
-    high_time <- d_ost$max_time[i]
-    #print(c(low_time, high_time))
-    group <- d_ost$group[i]
-    time_index <- which(d_uvp$juld >= low_time & d_uvp$juld <= high_time)
-    d_uvp$group[time_index] <- group
-  }
-  return(d_uvp)
-}
-
-# loop for it
-# start loop for all data
-wmo <- c(4903634, 6904240, 6904241, 1902578)
-park_depth <- c('200 m', '500 m', '1000 m')
+```{r, check total biovolume with slope value, fig.width=10, fig.height=6}
+#| label: fig-biovolume-slope
+#| fig-cap: Average biovolume seen by the UVP6 as a function of the attenuance slope (top) and total biovolume seen by the UVP as a function of the intensity of the jump seen by the OST (bottom) at each drifting depth.
+
+# uvp_drifting <- all_floats |> filter(PhaseName == 'PAR') |> select(juld, cycle, wmo, park_depth, all_of(lpm_class)) |> 
+#   drop_na(NP_Size_102) |> dplyr::mutate(date = as_date(juld))
+# 
+# # DSE for computing biovolumes
+# #DSE <- c(0.29042685,0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779, 1.16170742, 1.46365963)
+# DSE_long <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685, 0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
+#                 1.16170742, 1.46365963, 1.84409558, 2.32341484)
+# 
+# # compute biovolume
+# get_biovolume <- function(DSE){
+#   r <- DSE/2
+#   biovolume <- (4*pi/3)*r**3
+#   return(biovolume)
+# }
+# 
+# # get biovolume for each DSE (equivalent to say for each size bin)
+# biovolumes <- map_dbl(DSE_long, get_biovolume)  # unit = mm^3
+# 
+# # compute biovolume and add them to the dataframe
+# spectra_biovolume <- tibble(as.data.frame(t(t(uvp_drifting[,all_of(lpm_class)])*biovolumes)))
+# colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('NP_Size', 'Biovolume')
+# biovolumes_names <- colnames(spectra_biovolume)
+# uvp_drifting <- cbind(uvp_drifting, spectra_biovolume) |> rowwise() |> dplyr::mutate(total_biovolume = sum(across(all_of(biovolumes_names))))
+# 
+# assign_time_group <- function(d_uvp, d_ost){
+#   d_uvp$group <- NA
+#   for (i in 1:nrow(d_ost)){
+#     low_time <- d_ost$min_time[i]
+#     high_time <- d_ost$max_time[i]
+#     #print(c(low_time, high_time))
+#     group <- d_ost$group[i]
+#     time_index <- which(d_uvp$juld >= low_time & d_uvp$juld <= high_time)
+#     d_uvp$group[time_index] <- group
+#   }
+#   return(d_uvp)
+# }
+# 
+# # loop for it
+# # start loop for all data
+# wmo <- c(4903634, 6904240, 6904241, 1902578)
+# park_depth <- c('200 m', '500 m', '1000 m')
 
@@
 
-# ggplot(test) + geom_line(aes(x = mean_date, y = weighted_group_biovolume)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
-# ggplot(test) + geom_line(aes(x = mean_date, y = slope)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4)
-
-# check
-# test2 <- test |> group_by(wmo, cycle) |> dplyr::mutate(mean_mean_date = mean(floor_date(mean_date))) 
-# 
-# ggplot(test2) +
-#   geom_pointrange(aes(x = mean_mean_date, y = weighted_group_biovolume),
-#                       stat = 'summary',
-#                       fun.min = function(z) {quantile(z, 0.25)},
-#                       fun.max = function(z) {quantile(z, 0.75)},
-#                       fun = median, colour = 'black', size = .1) +
-#     facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
-#   
-# ggplot(test2) +
-#   geom_pointrange(aes(x = mean_mean_date, y = slope),
-#                       stat = 'summary',
-#                       fun.min = function(z) {quantile(z, 0.25)},
-#                       fun.max = function(z) {quantile(z, 0.75)},
-#                       fun = median, colour = 'black', size = .1) +
-#     facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) 
-# 
-# test2 |> group_by(wmo, park_depth) |> summarize(correl = cor(slope, weighted_group_biovolume))
-
-# ggplot(slope_vs_biovolume) + geom_point(aes(x = log(slope), y = log(weighted_group_biovolume))) + facet_wrap(~park_depth) + labs(x = 'OST slope [UNITS]', y = 'UVP biovolume [UNITS]') #+ theme(axis.title.x = element_blank())
-# #ggplot(test) + geom_point(aes(x = slope, y = weighted_group_biovolume)) + facet_wrap(~park_depth)
-ggplot(slope_vs_biovolume |> filter(slope > 0.00001)) + geom_point(aes(x = slope*0.25, y = weighted_group_biovolume)) + facet_wrap(~park_depth, scales = 'free') + labs(x = TeX('$ATN~slope~(day^{-1})$'), y = TeX('$UVP~biovolume~(mm^{3})$')) +
-  scale_x_continuous(trans = 'log10') + scale_y_continuous(trans = 'log10') + theme_bw()
+slope_vs_biovolume <- slope_vs_biovolume |> filter(slope > 0.00001) |> mutate(slope = slope*0.25, logslope = log10(slope), logweight = log10(weighted_group_biovolume))
+
+# ggplot(slope_vs_biovolume, aes(x = logslope, y = logweight, group = park_depth)) + geom_bin2d(bins = 30) + scale_fill_continuous(type = 'viridis') + facet_wrap(~park_depth, scales = 'free') + theme_bw() +
+#    stat_smooth(aes(x = logslope, y = logweight), method = "lm", se = FALSE, colour = "red") + labs(x = TeX('$ATN~slope~(day^{-1})$'), y = TeX('$UVP~biovolume~(mm^{3})$'))
+
+uvp_ost_up <- ggplot(slope_vs_biovolume, aes(x = slope, y = weighted_group_biovolume, group = park_depth)) + geom_bin2d(bins = 30) + scale_fill_continuous(type = 'viridis') + facet_wrap(~park_depth, scales = 'free') + theme_bw() +
+    labs(x = TeX('$ATN~slope~(day^{-1})$'), y = TeX('$UVP~biovolume~(mm^{3})$')) + scale_x_continuous(trans = 'log10') + scale_y_continuous(trans = 'log10')
 
 # what about jump values with uvp measurements at jump
-uvp_drifting <- uvp_drifting |> dplyr::mutate(num_date = as.numeric(juld))
-res_jumps <- res_jumps |> dplyr::mutate(num_date = as.numeric(dates))
-
-assign_uvp_to_jump <- function(d_uvp, d_jump){
-  d_jump$biovolume <- NA
-  for (i in 1:nrow(d_jump)){
-    index_jump_for_uvp <- which.min(abs(tmp_uvp$num_date - tmp_ost_jump$num_date[i]))
-    d_jump$biovolume[i] <- tmp_uvp$total_biovolume[index_jump_for_uvp]
-  }
-  return(d_jump)
-}
+# uvp_drifting <- uvp_drifting |> dplyr::mutate(num_date = as.numeric(juld))
+# res_jumps <- res_jumps |> dplyr::mutate(num_date = as.numeric(dates))
+# 
+# assign_uvp_to_jump <- function(d_uvp, d_jump){
+#   d_jump$biovolume <- NA
+#   for (i in 1:nrow(d_jump)){
+#     index_jump_for_uvp <- which.min(abs(tmp_uvp$num_date - tmp_ost_jump$num_date[i]))
+#     d_jump$biovolume[i] <- tmp_uvp$total_biovolume[index_jump_for_uvp]
+#   }
+#   return(d_jump)
+# }
 
@@
 diff_jump_vs_biovolume <- vroom::vroom('DATA/ARGO/diff_jump_vs_biovolume.csv')
-diff_jump_vs_biovolume <- diff_jump_vs_biovolume|> mutate(park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
-
-#ggplot(test_jump) + geom_point(aes(diff_jump, biovolume))
-# ggplot(diff_jump_vs_biovolume) + geom_point(aes(log(diff_jump), log(biovolume))) + facet_wrap(~park_depth) + labs(x = 'OST jump difference [UNITS]', y = 'UVP biovolume [UNITS]') +
-#   scale_x_continuous(breaks = c(-6,-4,-2,0))
-ggplot(diff_jump_vs_biovolume) + geom_point(aes(x = diff_jump*0.25, y = biovolume)) + facet_wrap(~park_depth, scales = 'free') + labs(x = TeX('$ATN~jump~difference$'), y = TeX('$UVP~biovolume~(mm^{3})$')) +
-  scale_x_continuous(trans = 'log10') + scale_y_continuous(trans = 'log10') + theme_bw()
-
-```
-
-TODO :
-
--
-
--   on se trouve dans un sytème producitve entre guidi 2008 et iversen (très productif)
-
-## Discussion2
-
-PARLER DU TRADEOFF BETWEEN SMALL FLUX AND BIG PARTICLE FLUX
-
-The latter are however subject to a tradeoff between the likely underestimation of $F_{small}$ and the detection of false jumps (without great impact on the value of $F_{big}$ because false jumps have low $c_{p}$ values) if the threshold is too small and the likely overestimation of $F_{small}$ to the detriment of $F_{big}$ with missed jumps if the threshold is too high.
-
-parler aussi du fait que lesmesures en dérivés, il y a 2h en UVP et OST etc peut-etre que les mesures à 500 m ...
-ou alors on mets plus de temps à 200 m...
-voir un peu.
-faire des profils tous les 10 jours pour une telle apllication ça sert à rien étant donné qu'en plus si on a l'uvp 6 avec l ER embarquée pour éutider la vmmp il faut clairmeent avoir une résolution temporelle plus éleée, ce qui est loin parfait donc voilà.
-
-TO DO - faire le profil de Martin (flux vertical) et faire les données de HPLC pour étudier les communautés et voir si elles sont différentes entre les deux sites de déploiements.
-
-## Conclusion
+diff_jump_vs_biovolume <- diff_jump_vs_biovolume|> mutate(park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')), diff_jump = diff_jump*0.25, log_diff_jump = log10(diff_jump), log_biovolume = log10(biovolume))
+
+# ggplot(diff_jump_vs_biovolume, aes(y = log_biovolume, x = log_diff_jump, group = park_depth)) + geom_bin2d(bins = 30) + scale_fill_continuous(type = 'viridis') + facet_wrap(~park_depth, scales = 'free') + theme_bw() +
+#    stat_smooth(aes(y = log_biovolume, x = log_diff_jump), method = "lm", se = FALSE, colour = "red") + labs(x = TeX('$ATN~jump~difference$'), y = TeX('$UVP~biovolume~(mm^{3})$'))
+
+uvp_ost_down <- ggplot(diff_jump_vs_biovolume, aes(y = biovolume, x = diff_jump, group = park_depth)) + geom_bin2d(bins = 30) + scale_fill_continuous(type = 'viridis') + facet_wrap(~park_depth, scales = 'free') + theme_bw() +
+    scale_x_continuous(trans = 'log10') + scale_y_continuous(trans = 'log10') + labs(x = TeX('$ATN~jump~difference$'), y = TeX('$UVP~biovolume~(mm^{3})$'))
+
+uvp_ost_up / uvp_ost_down
+```
+
+@fig-biovolume-slope shows that there is no clear relationship between the biovolume seen by the UVP6 and the attenuation (continuous or stochastic) measured by the OST.
+For small particles, the relationship between the ATN slopes and the average biovolume shows a weak correlation (0.30 at 500 m and 1000 m).
+For large particles, the OST misses particles that are seen by the UVP6 (see 500 m and 1000 m in the lower panel).
+It is coherent with the fact that OSTs were not designed to intercept and detect large sinking particles [@Estapa2023-oz].
+The current version of the OST has a housing above the optical window which can produce a "self-shading" effect [@Estapa2017-jr] that prevents the collection of fast sinking particles (assumed here to be large particles).
+In addition, the nature of particles is also a source of uncertainty in the attenuation signal.
+Mathematically, the attenuation efficiency ($Q_{c}$) of a spherical particle can be defined [@Mobley2022-de] as
+
+$$
+Q_{c} = Q_{a} + Q_{b}
+$$ {#eq-optics}
+
+where $Q_{a}$ and $Q_{b}$ are, respectively, the absorption and scattering efficiencies.
+Consequently, @Mobley2022-de defines the attenuance cross section ($\sigma_{c}$) as $Q_{c}$ multiplied by the geometrical cross section of that particle ($\pi r^{2}$ for a spherical particle).
+Therefore, if a particle of a given size has a small $\sigma_{c}$ (e.g.
+mostly transparent) or a high $\sigma_{c}$ (e.g.
+opaque), the attenuation will be different (lower for a smaller $\sigma_{c}$) but the size measured by the UVP6 will be the same.
+
+Therefore, if the UVP6 and the OST are well correlated at first order (@fig-comparison-ost-uvp, @tbl-correlation), it is more difficult to directly related the attenuance to the biovolume (hence the size) of a particle because of the nature of particles themselves.
+Consequently, the combined use of the UVP6 and the OST seems useful to study the latent particle dynamics but the interest of using the OST for large particles (e.g.
+during a bloom) might be more limited.
+
+## Perspectives
+
+### A complementary variable to investigate POC fluxes - $b_{bp}$
+
+I did not use $b_{bp}$ data, a proxy for suspended particle concentration, that has already been used to derive POC concentrations, POC fluxes and particle size [@Briggs2013-hj; @Briggs2011-du; @Alkire2012-gv; @Cetinic2012-lz; @DallOlmo2014-o; @Lacour2019-id; @Briggs2020-gz; @Giering2020-mc; @Rasse2017-tg].
+In @Briggs2011-du, the authors explain that when a large particle goes through the sampling volume of the $b_{bp}$ sensor, it causes a spike in the $b_{bp}$ signal.
+However, the height of this spike depends on many factors such as the cross section of the particle (or aggregate), $Q_{b}$, the sampling volume, the fraction of the particle in the sampling volume and the time the particle spends in the sampling volume.
+Using an empirical relationship between ship-based $b_{bp}$ and POC, they were able to derive POC concentrations (mg C m$^{-3}$) and then POC fluxes (mg C m$^{-2}$ day$^{-1}$) by multiplying those concentrations by a given sinking speed (m day$^{-1}$).
+
+As of today (March 15, 2023), 111361 $b_{bp}$ profiles were acquired by BGC-Argo floats.
+If a relationship between particle size measured by the UVP6 and particle size derived from $b_{bp}$ spikes (e.g.
+spike height) could be found, we could go back in time and infer information on the abundance of large particles at global scale.
+
+### Configuration adaptation to specific BCP-related processes
+
+The current configuration of all floats in the LS was a 10-day ascending profile acquisition with drifting at 200 m for one day, 500 m for 3 days and 1000 m for 5 days.
+During drifting, OST measurements were taken every 30 minutes at all depths whereas UVP measurements were taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m.
+This initial configuration was based on a compromise between science targets and energy consumption limitations for long-term deployment (years).
+However, I think this configuration could be adapted depending on the period of the year to target specific processes.
+For instance, I tried to evaluate sinking speeds from particle concentrations using a gaussian fit in different water layers to follow the gaussian peak with time and depth for each size class.
+I managed to compute realistic sinking speeds but the uncertainty was huge and potentially biased because I was just trying to fit something I saw in the data.
+Therefore, I did not use those estimations here but I think it would be possible to compute them if the float was sampling every day when the abundance of particles is at its maximum (e.g.
+during the spring bloom).
+Similarly, to address the question of the vertical migration, a 10-day temporal resolution is way to low because plankton images are rare compared to detritus so the sampling volume should be greatly increased to have a better estimation of the presence (or absence) of zooplankton.
+Finally, I have not made up my mind about the choice of the three drifting depths but maybe it would be interesting to increase the time spent at 200 m (to 3 days) and decrease the time spent at 1000 m §to 3 days instead of 5) where the probability of capturing big particles is lower than at 200 m.
+There could also be UVP and OST measurements closer in time at all drifting depths to have a better comparison of attenuation slopes/jumps and biovolumes.
+
+### A statistical method to detect jumps instead of using a threshold
+
+One fundamental difference between our jump detection method and the initial method based on an attenuance jump threshold [@Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk; @Estapa2023-oz] is that we use a statistical method to generalize the jump detection to new environments and new drifting depths.
+Initially, I used slope thresholds based on the 90$^{th}$ and 95$^{th}$ percentiles of the $c_{p}$ slope distribution.
+Therefore, I had three distinct slope thresholds for each drifting depth.
+We also thought that with time, this threshold should stabilize (i.e.
+having enough data points in the slope distribution to have stables percentiles).
+However, the rapid decrease in particle concentrations starting in November kept decreasing the threshold because the 90$^{th}$ and 95$^{th}$ percentiles were decreasing due the vanishing of large particles (@fig-uvp6-profiling-512-2500).
+One option was to stop the slope distribution in October, before large particles virtually disappeared.
+At that time, we therefore looked for a method that could derive a slope threshold for each drifting profile.
+I think it makes sense because it solved the issue of getting the best threshold that would be a unique one for all profiles and because the method is also relatively objective, easy to implement and resilient.
+By resilient, I mean that if the method is not perfect, it does not miss big jumps which are important for the correct estimation of $F_{large}$ and for data points wrongly identified as jumps (as outliers in the slope distribution), they do not contribute to $F_{large}$ because they have a small $c_{p}$ contribution and $F_{small}$ is generally not so impacted either because the contribution loss of that false jump is negligible to the slow accumulation on the transmissometer window.
+This can be checked on the web app mentonned earlier in the *drifting OST* tab.

### 2023-03-15 bcp_sequestration.qmd
update

```{=latex}
\includepdf[pages=-]{"/home/flo/dox/THESIS/Ricour2023main"} %https://tex.stackexchange.com/questions/183469/includepdf-only-shows-first-page thank you stranger
\includepdf[pages=-]{"/home/flo/dox/THESIS/Ricour2023supp"}
```

### 2023-03-16 annexes.qmd
update

@@
 # Annexes
+
+```{=latex}
+\includepdf[pages=1]{"/home/flo/dox/THESIS/bg-18-755-2021"}
+```
 
@@
 ```
-
 ```{r, poc database from traps}

### 2023-03-16 article3.qmd
update

@@
 In the first chapter, I explained the different steps that led to the creation of the embedded zooplankton classification model for the UVP6.
-However, I did not specify that in spite of the technical and energetic constraints, the model was also due for April 2022 for a deployment of BGC-Argo floats in the Labrador Sea a month later.
+However, I did not specify that in spite of the technical and energy constraints, the model was also due for April 2022 for a deployment of BGC-Argo floats in the Labrador Sea a month later.
 Once deployed, it was only a question of time before I could dig into the UVP6 and the OST data.
 In the meantime I asked myself, what could be a cool thing to do about those data?
-I already had some experience with BGC-Argo floats during my master's degree (study of the deep chlorophyll maximum in the Black Sea, see also the abstract of that paper in APPENDIX XX) and because I was a bit bored waiting for new BGC-Argo profiles, I thought it would be useful to have a visualization tool in the form of a web app to investigate incoming data and already prepare the ground for the following study.
-Here is the link to the [web app](https://fricour.shinyapps.io/CARBON_REFINE/), which is still currently being updated every night with new data (near) live from the Labrador Sea and hopefully with other BGC-Argo floats equipped with the UVP6 and the OST to investigate carbon and particle fluxes in other regions.
+I already had some experience with BGC-Argo floats during my master's degree (study of the deep chlorophyll maximum in the Black Sea, see also the abstract of that paper in the annexes) and so while I was waiting for new profiles, I thought it would be useful to have a visualization tool in the form of a web app to investigate incoming data and already prepare the ground for the following study.
+Here is the link to the [web app](https://fricour.shinyapps.io/CARBON_REFINE/), which is still currently being automatically updated every day with new data (almost) live from the Labrador Sea and hopefully with other BGC-Argo floats equipped with the UVP6 and the OST to investigate carbon and particle fluxes in other regions.
 This study uses preliminary data (not even one year of data) and a new BGC-Argo float configuration hence more work should be addressed to better understand the dynamics of the system.
@@
 The biological carbon pump (BCP) refers to a series of processes that transport biogenic carbon produced in sunlit surface waters to the ocean interior [@Claustre2021-tx].
-It is composed of three main pathways [@Claustre2021-tx; @Boyd2019-ez; @Siegel2022-wd]: the biological gravitational pump that is the gravitational settling of particulate organic carbon (POC); the physically mediated particle injection pump that represents physical mechanisms that transport dissolved organic carbon (DOC) to depth via mixing processes; and the biologically mediated particle injection pump that characterizes the transport of POC (and DOC) by diel and seasonal migration of animals in the water column.
+It is composed of three main pathways [@Claustre2021-tx; @Boyd2019-ez; @Siegel2022-wd]: the biological gravitational pump that is the gravitational settling of particulate organic carbon (POC); the physically mediated particle injection pump that represents physical mechanisms that transport dissolved organic carbon (DOC) to depth via mixing processes; and the biologically mediated particle injection pump that characterizes the transport of POC by diel and seasonal migration of animals in the water column.
 Historically, the BCP was mainly studied with sediment traps [@Honjo2008-kn] and Thorium-234 radionuclide disequilibria [@Buesseler2007-ql] focused on the gravitational pump hence the contribution of DOC and migrating animals to the BCP was poorly known.
@@
 
-Two couples of BGC-Argo floats were deployed in the Labrador Sea in late May 2022.
-The first couple (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second (WMO 6904241 and 1902578) was deployed on the 29$^{th}$ at coordinates (-52.4°E, 56.8°N).
+Two pairs of BGC-Argo floats were deployed in the Labrador Sea in late May 2022.
+The first pair (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second (WMO 6904241 and 1902578) was deployed on the 29$^{th}$ at coordinates (-52.4°E, 56.8°N).
 
@@
 Temperature and salinity data were quality controlled using @Wong2022-hz.
-The potential density anomaly was computed with the function swSigma0 of the oce R package [@Kelley_undated-sw] and the MLD was determined using the 0.03 kg m$^{-3}$ density criterion at a reference depth of 10 m [@De_Boyer_Montegut2004-hh].
+The potential density anomaly was computed with the function swSigma0 of the oce R package [@Kelley2021-sw] and the MLD was determined using the 0.03 kg m$^{-3}$ density criterion at a reference depth of 10 m [@De_Boyer_Montegut2004-hh].
 $c_{p}$ data were converted from counts to physical units (m$^{-1}$) using the following relationship
@@
 UVP6 data and in particular the averaged number of particles per size class (i.e.
-PSD) was divided by the number of images at each depth and the water volume imaged by the UVP (0.7L) to obtain particle concentrations (units: #/L).
+PSD) was divided by the number of images at each depth and the water volume imaged by the UVP (0.7 L) to obtain particle concentrations (units: #/L).
 Note that in an impending update by the Coriolis data center, it will no longer be necessary to divide by the number of images.
@@
 
-We applied a similar approach to @Estapa2013-gw and @Estapa2017-jr to derive the continuous component (i.e.
+We applied a similar approach to [@Estapa2013-gw; -@Estapa2017-jr] to derive the continuous component (i.e.
 continuous deposition of small particles on the upward-facing window of the transmissometer) and discontinuous (i.e.
@@
 
-First, the drifting $c_{p}$ signal was cleaned from spikes (hypothesized as transient particles not landing on the window or active swimmers [@Estapa2013-gw]).
+First, the drifting $c_{p}$ signal was cleaned from spikes (assumed as transient particles not landing on the window or active swimmers [@Estapa2013-gw]).
 For this, we computed a 7-point moving median and a 7-point moving median absolute deviation (MAD).
@@
 
-Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr; @Estapa2023-oz], we chose an empirical statistical approach instead of fixed thresholds to detect jumps.
-Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase and computed the z-score (i.e.
-standard score) of each of them.
-The detection of jumps therefore consisted in spotting outliers in the slope distribution.
+Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk], we chose an empirical statistical approach instead of thresholds to detect the deposition of large particles (hereafter referred to as jumps) in the $c_{p}$ signal.
+Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase.
+The detection of jumps consisted in spotting outliers in the slope distribution.
 For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles.
@@
 
-In order to compute $F_{small}$, we firstly computed the slope of each group of observations between jumps (positive and negative, see @fig-estapa-method_article).
-Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by factors other than the pump cleaning the optical window (e.g.
+In order to compute $F_{small}$, we firstly computed the slope of each segment between jumps (positive and/or negative, see @fig-estapa-method_article).
+Any segment with less than 4 observations or a negative slope was removed assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by factors other than the pump cleaning the optical window (e.g.
 turbulence ou zooplankton consumption of accumulated material).
-Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).
+Afterwards, we computed the weighted average slope ($slope_{w}$) of each segment to derive the small particles flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).
 
@@
 
-where $slope_{w}$ is multiplied by the transmissometer pathlength due to the fact that @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.
+where $slope_{w}$ is multiplied by the transmissometer pathlength because @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.
 
@@
 #| label: fig-estapa-method_article
-#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component and a discontinuous (positive jumps, red) component.
-Slopes for each group are depicted in light blue.
+#| fig-cap: Example of the jump detection method on a drifting transmissometer where the $c_{p}$ signal is divided into a continuous (dark blue) component and a discontinuous (positive jumps, red) component.
+Slopes for each segment are depicted in light blue.
 A negative jump was also detected (yellow).
@@
 # remove if only one point (cannot fit a slope with one point)
-slope_df <- slope_df |> filter(nb_points > 1)
+slope_df <- slope_df |> filter(nb_points > 3)
 
@@
 Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: 10.48670/moi-00281) with a spatial resolution of 4 km.
-Mixed layer depth (MLD) derived from satellite observations were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052).
+Mixed layer depth (MLD) derived from satellite observations were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052) with a spatial resolution of 0.25°.
 
@@
 
-  # remove if only one point (cannot fit a slope with one point)
-  slope_df <- slope_df |> filter(nb_points > 1)
+  # remove if only one point (cannot fit a slope with one point) -> switched to 3 points
+  slope_df <- slope_df |> filter(nb_points > 3)
 
@@
 At 200 m, $F_{small}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ (average $\pm$ standard deviation).
-At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$.
+At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$.
 For both drifting depths, the trend shows a decrease of $F_{small}$ from June to, respectively, January and February before the drift was stopped at those depths due to the deepening of the winter MLD (@fig-mld) in order to ensure the float stability.
-At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid/end of July followed by a steady decrease towards March, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.
+At 1000 m, all floats measured an increase of $F_{small}$ from late May to July followed by a steady decrease towards March, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.
 
@@
 #| label: fig-OST-flux
-#| fig-cap: OST-derived small (top) and large (bottom) particles carbon flux for all floats at each drifting depth.
+#| fig-cap: OST-derived small (top) and large (bottom) particle carbon fluxes for all floats at each drifting depth.
 Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.
@@
 In @fig-uvp-200m, @fig-uvp-500m and @fig-uvp-1000m, we show the daily averaged concentrations of particles in each size class for each float at their respective drifting depth.
-At 200 m, it can be observed that particle concentration in the 102 - 512 $\mu$m range are relatively uniform between June and October before decreasing sharply towards January.
+At 200 m, it can be observed that particle concentrations in the 102 - 512 $\mu$m range are relatively uniform between June and October before decreasing sharply towards January.
 Between 512 $\mu$m and 1.63 mm, concentrations are steadily decreasing from June.
@@
   scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
-  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
+  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
   scale_y_continuous(trans = 'log10') +
@@
 At 500 m, particle concentrations in the 102 - 645 $\mu$m range slowly decrease from July to September with a sharp decrease starting in October.
-Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (\< 0.5 particle/L) on the entire timeserie.
+Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (\< 0.5 particle/L) on the entire time series.
 
@@
 
-Vertical profiles of particle concentration in the lower (102 - 512 $\mu$m) and upper (0.512 - 2.5 mm) size range of the UVP6 shows similar patterns across all floats.
-In the lower part of the size spectrum, @fig-uvp6-profiling-102-406 shows the rapid emptying of the water column in November.
+Vertical profiles of particle concentrations in the lower (102 - 512 $\mu$m) and upper (0.512 - 2.5 mm) size range of the UVP6 show similar patterns across all floats.
+In the lower part of the size range, @fig-uvp6-profiling-102-406 shows the rapid emptying of the water column in November.
 This process occurred on the entire water column (e.g.
@@
 size class 102, float 6904240).
-All floats also measured a local particle abundance maximum around 800 - 1000 m whose origin could be due to a previous particle export or remineralization of largeger particles.
+All floats also measured a local particle abundance maximum around 800 - 1000 m whose origin is analyzed in the Discussion.
 
@@
 #| label: fig-uvp6-profiling-102-406
-#| fig-cap: Vertical profiles of UVP6 particle concentration for size classes in the 102 - 512 $\mu$m size range.
+#| fig-cap: Vertical profiles of UVP6 particle concentrations for size classes in the 102 - 512 $\mu$m size range.
 Concentrations were normalized for each size class.
@@
 #| label: fig-uvp6-profiling-512-2500
-#| fig-cap: Vertical profiles of UVP6 particle concentration for size classes in the 0.512 - 2.5 mm size range.
+#| fig-cap: Vertical profiles of UVP6 particle concentrations for size classes in the 0.512 - 2.5 mm size range.
 Concentrations were normalized for each size class.
@@
 microphytoplankton) whereas station B was composed of a majority of smaller phytoplankton species (mostly nano and picozooplankton).
-This supports the idea of two distinct bloom dynamics as inferred from float and satellite data.
+This supports the idea of two distinct biological dynamics as inferred from float and satellite data.
 
@@
 This increase raises the question of a possible particle export event.
-For classes \> 512 $\mu$m, the vertical concentration profiles highlight a largeger particle plume for the 6904241-1902578 pair than for the other one.
+For classes \> 512 $\mu$m, the vertical concentration profiles highlight a larger particle plume for the 6904241-1902578 pair than for the other one.
 However, apart from the float 6904240, all floats measured similar integrated Chla content in the upper 100 m and in the mixed layer (@fig-chla-content).
@@
 
-Still, it is unclear why we observe a progressive emptying of the water column starting below 500 m for size classes below 323 $\mu$m [@fig-uvp6-profiling-102-406].
+Still, it is unclear why we observe a progressive emptying of the water column starting below 500 m for size classes below 323 $\mu$m (@fig-uvp6-profiling-102-406).
 This pattern could result from several causes.
-It could either be the signature of a subduction event [@Llort2018-pa] either the end of a slow sinking export (i.e latent dynamic of the system) or a sign of disaggregation processes by zooplankton (i.e.
+It could either be the signature of a subduction event [@Llort2018-pa] either the end of a slow sinking export (i.e latent dynamic of the system) or a sign of fragmentation processes by zooplankton (i.e.
 zooplankton feeding).
@@
 Spiciness is a variable that allows the differentiation of water masses with distinct thermohaline properties but similar density [@Flament2002-zm].
-@fig-spice shows that both the density field and the spiciness field are uniform below 500 m hence it is unlikely that a subduction event would be the reason for this decrease in particle abundance below 500 m.
+@fig-spice shows that both the density and the spiciness fields are uniform below 500 m hence it is unlikely that a subduction event would be the reason for this decrease in particle abundance below 500 m.
 Using the embedded zooplankton classification results from the UVP6, we looked at concentrations of *Calanoida*, a species that is abundant in the NA [@Corkett1979-ha].
 @fig-spice shows that the concentration of *Calanoida* is higher from the end of May to July in the surface layer.
-There seems to be a lower patch between 250 and 500 m from August to November however it is difficult to estimate their impact on particle concentrations at this stage.
+There also seems to be a lower patch between 250 and 500 m from August to November however it is difficult to estimate their impact on particle concentrations at this stage.
 Finally, we computed the spectral slope to look for aggregation/fragmentation processes.
 The spectral slope defines the slope of the PSD where a high (in absolute values) spectral slope indicates that small particules dominate the PSD.
-Here, the panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope at the beginning of the time series but it is also difficult to attach it to biological processes such as zooplankton feeding without additional data.
+Here, the panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope at the beginning of the time series but it is also difficult to relate it to biological processes such as zooplankton feeding without additional data.
 
@@
 First, the 10-day profiling configuration hinders the characterization of particle export events that are transient processes.
-Second, the plume of large particles can be misleading because their abundance measured by the UVP6 are very low (usually \< 0.1 #/L).
+Second, the plume of large particles can be misleading because the abundance measured by the UVP6 are very low (usually \< 0.1 #/L).
 However, the plume could be the result of aggregation (repackaging) by zooplankton feeding on small particles.
@@
 
-For the first time, BGC-Argo floats equipped with both an OST and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx].
+For the first time, BGC-Argo floats equipped with both an OST and the UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx].
 The quantification of the POC flux using the OST method [@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk; @Estapa2023-oz] is based on an empirical POC:ATN ratio [@Estapa2023-oz] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:
@@
 #| label: tbl-correlation
-#| tbl-cap: Pearson correlations between UVP6 PSD-derived carbon fluxes and total, large particles and small particles carbon flux measured by the OST at 3 drifting depths.
+#| tbl-cap: Pearson correlations between UVP6 PSD-derived carbon fluxes and total, large and small particle carbon fluxes measured by the OST at 3 drifting depths.
 correlations <- comparison_flux_ost_uvp |> dplyr::select(park_depth, wmo, cycle, guidi_flux, small_flux, large_flux, date, total_flux) |> dplyr::group_by(wmo, park_depth, cycle) |> dplyr::mutate(median_guidi_flux = median(guidi_flux)) |> dplyr::group_by(park_depth, wmo) |> dplyr::summarize(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_large = round(cor(median_guidi_flux, large_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))
@@
 The trends of total flux measured by both instruments are very well correlated at all drifting depths for floats 6904240 and 4903634 with correlations (see @tbl-correlation) ranging from `r min(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])` to `r max(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])`.
-@tbl-correlation also shows that the correlation between UVP-derived carbon fluxes and $F_{small}$ and $F_{large}$ is `r round(mean(as.vector(as.matrix(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,2:7]))),2)` on average.
+@tbl-correlation also shows that the correlation between UVP-derived carbon fluxes and $F_{small}$ and $F_{large}$ for that pair is `r round(mean(as.vector(as.matrix(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,2:7]))),2)` on average.
 We calculated lower correlations with the other two floats, especially the float 1902578 at 200 m, where correlations between the UVP-derived carbon fluxes and all fluxes (small, large and total) are close to 0.
-Overall, it nevertheless suggests that there is a possibility to get more accurate estimations of carbon fluxes using PSD only therefore using only the particle size as an estimator of carbon flux, contrarily to what other studies have argued \[@Iversen2020-kb; @McDonnell2010-un; @Iversen2023-qj\].
+Overall, it nevertheless suggests that there is a possibility to get more accurate estimations of carbon fluxes using only the PSD hence using only the particle size as an estimator of carbon flux, contrarily to what other studies have argued [@Iversen2020-kb; @McDonnell2010-un; @Iversen2023-qj].
 
@@
 Similarly, $F_{large}$ could be related to the biovolume seen by the UVP6 when the OST detects jumps in the attenuation signal.
-Consequently, we firstly computed all the subslopes (see lightblue straight lines in @fig-estapa-method_article) and all the jumps at all drifting depths as well as the total biovolume seen by the UVP6 while the float was drifting.
-Then, we computed the average biovolume seen during each subslope duration and we associated the closest total biovolume (in time) with a jump.
+Consequently, we firstly computed the slope of each segment (see lightblue straight lines in @fig-estapa-method_article) and all the jumps at all drifting depths as well as the total biovolume seen by the UVP6 while the float was drifting.
+Then, we computed the average biovolume seen during each segment duration and we associated the closest total biovolume (in time) with a jump.
 
@@
 #| label: fig-biovolume-slope
-#| fig-cap: Average biovolume seen by the UVP6 as a function of the attenuance slope (top) and total biovolume seen by the UVP as a function of the intensity of the jump seen by the OST (bottom) at each drifting depth.
+#| fig-cap: Average biovolume seen by the UVP6 as a function of the attenuance slope (top) and total biovolume seen by the UVP6 as a function of the intensity of the jump seen by the OST (bottom) at each drifting depth.
 
@@
 @fig-biovolume-slope shows that there is no clear relationship between the biovolume seen by the UVP6 and the attenuation (continuous or stochastic) measured by the OST.
-For small particles, the relationship between the ATN slopes and the average biovolume shows a weak correlation (0.30 at 500 m and 1000 m).
+For small particles, the relationship between the ATN slope and the average biovolume shows a weak correlation (0.30 at 500 m and 1000 m).
 For large particles, the OST misses particles that are seen by the UVP6 (see 500 m and 1000 m in the lower panel).
@@
 
-Therefore, if the UVP6 and the OST are well correlated at first order (@fig-comparison-ost-uvp, @tbl-correlation), it is more difficult to directly related the attenuance to the biovolume (hence the size) of a particle because of the nature of particles themselves.
+Therefore, if the UVP6 and the OST are well correlated at first order (@fig-comparison-ost-uvp, @tbl-correlation), it is more difficult to directly link the attenuance to the biovolume (hence the size) of a particle because of the nature of particles themselves.
 Consequently, the combined use of the UVP6 and the OST seems useful to study the latent particle dynamics but the interest of using the OST for large particles (e.g.
@@
 
-I did not use $b_{bp}$ data, a proxy for suspended particle concentration, that has already been used to derive POC concentrations, POC fluxes and particle size [@Briggs2013-hj; @Briggs2011-du; @Alkire2012-gv; @Cetinic2012-lz; @DallOlmo2014-o; @Lacour2019-id; @Briggs2020-gz; @Giering2020-mc; @Rasse2017-tg].
+I did not use $b_{bp}$ data, a proxy for suspended particle concentration, that has already been used to derive POC concentrations, POC fluxes and particle size [@Briggs2013-hj; @Briggs2011-du; @Alkire2012-gv; @Cetinic2012-lz; @DallOlmo2014-ot; @Lacour2019-id; @Briggs2020-gz; @Giering2020-mc; @Rasse2017-tg].
 In @Briggs2011-du, the authors explain that when a large particle goes through the sampling volume of the $b_{bp}$ sensor, it causes a spike in the $b_{bp}$ signal.
@@
 As of today (March 15, 2023), 111361 $b_{bp}$ profiles were acquired by BGC-Argo floats.
-If a relationship between particle size measured by the UVP6 and particle size derived from $b_{bp}$ spikes (e.g.
-spike height) could be found, we could go back in time and infer information on the abundance of large particles at global scale.
+With the advent of floats equipped with both the UVP6 and a $b_{bp}$ sensor, it would be interesting to investigate the relationship between the size measured by the UVP6 and the size derived by spikes (i.e.
+spike height) in the $b_{bp}$ signal.
+Such an empirical relationship could bring information on the abundance of large particles at global scale using the huge, already available, $b_{bp}$ database.
 
@@
 Similarly, to address the question of the vertical migration, a 10-day temporal resolution is way to low because plankton images are rare compared to detritus so the sampling volume should be greatly increased to have a better estimation of the presence (or absence) of zooplankton.
-Finally, I have not made up my mind about the choice of the three drifting depths but maybe it would be interesting to increase the time spent at 200 m (to 3 days) and decrease the time spent at 1000 m §to 3 days instead of 5) where the probability of capturing big particles is lower than at 200 m.
-There could also be UVP and OST measurements closer in time at all drifting depths to have a better comparison of attenuation slopes/jumps and biovolumes.
-
-### A statistical method to detect jumps instead of using a threshold
+Finally, I have not made up my mind about the choice of the three drifting depths but maybe it would be interesting to increase the time spent at 200 m (to 3 days) and decrease the time spent at 1000 m to 3 days instead of 5) where the probability of capturing big particles is lower than at 200 m.
+There could also be UVP6 and OST measurements closer in time at all drifting depths to have a better comparison of attenuation slopes/jumps and biovolumes.
+
+```{=html}
+<!-- ### A statistical method to detect jumps instead of using a threshold
 
@@
 This can be checked on the web app mentonned earlier in the *drifting OST* tab.
+-->
+```

### 2023-03-16 intro.qmd
update

@@
 | Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
-|-----------------|------------------|-----------------|--------------------|
+|------------------------|----------------------------------|---------------------------|--------------------------------------|
 | Gravitational          | export depth - sediment          | None                      | N/A (142)                            |

### 2023-03-16 summary.qmd
update

# Summary {.unnumbered}

**Towards a new insight of the carbon transport in the global ocean**

The ocean is known to play a key role in the carbon cycle. Without it, atmospheric CO2 levels would be much higher than what they are today thanks to the presence of carbon pumps that maintain a gradient of dissolved inorganic carbon (DIC) between the surface and the deep ocean. The biological carbon pump (BCP) is primarily responsible for this gradient. It consists in a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis in sunlit surface waters and then transported to the ocean interior and possibly the sediment where it will be sequestered from the atmosphere for millions of years. The BCP was long thought as solely the gravitational settling of particulate organic carbon (POC). However, a new paradigm for the BCP has recently been defined in which physically and biologically mediated particle injection pumps have been added to the original definition. Physically mediated particle injection pumps provide a pathway to better understand the transport of dissolved organic carbon (DOC) whereas biologically mediated particle injection pumps focus on the transport of POC by vertically migrating animals, either daily or seasonally. Therefore, a better understanding of these processes could help bridge the gap between carbon leaving the surface and carbon demand in the ocean interior. To address this new paradigm, this work will benefit from the advent of recent sensors that equip a new generation of Biogeochemical-Argo floats (BGC-Argo). The first part focuses on the development of an embedded zooplankton classification model for the Underwater Vision Profiler 6 (UVP6) under strict technical and energy constraints. The second part studies particle and carbon fluxes in the Labrador Sea using BGC-Argo floats equipped for the first time with the UVP6 and an optical sediment trap (OST), providing two independent measurements of sinking particles. The last part consists in revisiting the BCP using a new framework called CONVERSE for Continuous Vertical Sequestration. With this new approach, we re-evaluate the total carbon sequestered from the atmosphere (> 100 years) by the BCP and its transport pathways on the entire water column, in contrast to the carbon sequestration typically assumed below a fixed reference depth.

**Keywords** Carbon pumps -- Underwater Vision Profiler 6 -- BGC-Argo floats -- Machine learning

```{=latex}
\newpage
```

**Vers une meilleur compréhension du transport du carbone dans l'océan global**

L'océan est connu pour jouer un rôle clé dans le cycle du carbone. Sans lui, les niveaux de CO2 atmosphérique seraient bien plus élevés qu'aujourd'hui grâce à la présence de pompes à carbone qui maintiennent un gradient de carbone inorganique dissous (DIC) entre la surface et le fond de l'océan. La pompe à carbone biologique (BCP) est principalement responsable de ce gradient. Elle consiste en une série de processus océaniques au cours desquels le carbone inorganique est converti en matière organique via la photosynthèse dans les eaux de surface, puis transporté vers l'intérieur de l'océan et éventuellement les sédiments où il sera séquestré par rapport à l'atmosphère pour des millions d'années. La BCP était longtemps considérée comme étant uniquement la déposition gravitationnelle de carbone organique particulaire (POC). Cependant, un nouveau paradigme pour la BCP a récemment été défini dans lequel des pompes d'origine physique et biologique d'injection de particules ont été ajoutées à la définition originale. Les pompes physiques d'injection de particules fournissent un moyen de mieux comprendre le transport de carbone organique dissous (DOC), tandis que les pompes biologique d'injection de particules se concentrent sur le transport de POC par des animaux migrant verticalement, quotidiennement ou saisonnièrement. Par conséquent, une meilleure compréhension de ces processus pourrait aider à combler l'écart entre le carbone quittant la surface et la demande de carbone dans l'océan profond. Pour aborder ce nouveau paradigme, ce travail bénéficiera de l'arrivée de capteurs récents équipant une nouvelle génération de flotteurs Biogéochimiques-Argo (BGC-Argo). La première partie se concentre sur le développement d'un modèle embarqué de classification de zooplancton pour l'Underwater Vision Profiler 6 (UVP6) avec des contraintes techniques et énergétiques strictes. La deuxième partie étudie les flux de particules et de carbone dans la mer du Labrador en utilisant des flotteurs BGC-Argo équipés pour la première fois de l'UVP6 et d'un piège à sédiments optique (OST), fournissant deux mesures indépendantes des particules. La dernière partie consiste à revisiter la BCP en utilisant un nouveau cadre appelé CONVERSE qui fait référence à la séquestration verticale continue du carbone dans la colonne d'eau. Avec cette nouvelle approche, nous réévaluons le carbone total séquestré par rapport à l'atmosphère (\> 100 ans) par la BCP et ses voies de transport sur toute la colonne d'eau, par opposition à la séquestration de carbone généralement supposée en-dessous d'une profondeur de référence fixe.

**Mots clés** Pompes à carbone -- Underwater Vision Profiler 6 -- Flotteurs BGC-Argo -- Machine learning

### 2023-03-16 uvp6_part.qmd
update

@@
 
-Introduire le chapitre et dire que le papier est en cours somehow ...
+Building the zooplankton classification model for the UVP6 was the first task of my PhD.
+To be honest, I knew nothing about machine learning before I started this so I spent the first three weeks reading a nice machine learning book in the corner of my team leader's office.
+To be even more honest, I basically knew nothing about zooplankton either.
+It is time to admit it.
+
+In machine learning, the no free lunch theorem [@Wolpert1996-hw] states that if you do not make any assumption about your data, then there is no model that is a priori guaranteed to work better [@Geron2019-is].
+In other words, it means that you need to get your hands dirty to find a model that works.
+The first step consisted in developing a full data processing pipeline using UVP5 images.
+The aim was to be prepared for the UVP6 image data set (not ready at the time) to train the true final classification model.
+To this end, I developed a python package called [UVPec](https://github.com/ecotaxa/uvpec) that is able to create a full classification model based on image folders with labelled particle/zooplankton images and some hyperparameters (e.g.
+learning rate).
+It really made my life easier when I got the final UVP6 data set and it saved me a lot of time.
+
+In the end, I am rather satisfied with the results.
+Classifying zooplankton is already difficult for human operators (at least, for me) so building a model to automatically classify them is one challenge but adding technical and energy constraints to it was the ice on the cake.
+Below is the first draft of a publication on the classification model.
 
@@
 library(tibble)
+library(tidyverse)
 
@@
 test <- test %>% group_by(Pressure_dbar) %>% summarize(n_sample = n(), total_conc = sum(ObjectConc5), mean_conc = total_conc/n_sample)
-cal <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip()+
+cal <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point(colour = "#005CAB") + scale_x_reverse() + coord_flip()+
   xlab('Pressure (dbar)') + ylab('Calanoida (#/L)') + geom_point(data = test[test$mean_conc>0,], 
-                                                                  aes(x = Pressure_dbar, y = mean_conc), colour = "red") +
+                                                                  aes(x = Pressure_dbar, y = mean_conc), colour = "#E31B23") +
   geom_vline(xintercept = DCM, lty = 'dotted') + theme_bw()
@@
 test <- test %>% group_by(Pressure_dbar) %>% summarize(n_sample = n(), total_conc = sum(ObjectConc14), mean_conc = total_conc/n_sample)
-det <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip() +
+det <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) +  geom_point(colour = "#005CAB") + scale_x_reverse() + coord_flip() +
   geom_point(data = test[test$mean_conc>0,], 
-             aes(x = Pressure_dbar, y = mean_conc), colour = "red") +
+             aes(x = Pressure_dbar, y = mean_conc), colour = "#E31B23") +
   xlab('Pressure (dbar)') + ylab('detritus (#/L)') + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
@@
 test <- test %>% group_by(Pressure_dbar) %>% summarize(n_sample = n(), total_conc = sum(ObjectConc17), mean_conc = total_conc/n_sample)
-puf <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip()+
+puf <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) +  geom_point(colour = "#005CAB") + scale_x_reverse() + coord_flip()+
   xlab('Pressure (dbar)') + ylab('puff (#/L)') + geom_point(data = test[test$mean_conc>0,], 
-                                                                          aes(x = Pressure_dbar, y = mean_conc), colour = "red") +
+                                                                          aes(x = Pressure_dbar, y = mean_conc), colour = "#E31B23") +
   geom_vline(xintercept = DCM, lty = 'dotted') + theme_bw()
@@
 test <- test %>% group_by(Pressure_dbar) %>% summarize(n_sample = n(), total_conc = sum(ObjectConc20), mean_conc = total_conc/n_sample)
-tuf <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip()+
+tuf <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) +  geom_point(colour = "#005CAB") + scale_x_reverse() + coord_flip()+
   xlab('Pressure (dbar)') + ylab('tuff (#/L)') + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
   geom_point(data = test[test$mean_conc>0,], 
-             aes(x = Pressure_dbar, y = mean_conc), colour = "red") +
+             aes(x = Pressure_dbar, y = mean_conc), colour = "#E31B23") +
   geom_vline(xintercept = DCM, lty = 'dotted') + theme_bw()

### 2023-03-17 annexes.qmd
update

--- 2023-02-15 annexes.qmd
@@
 
+```{=latex}
+%\includepdf[pages=1]{"/home/flo/dox/THESIS/bg-18-755-2021"}
+```
 ```{=html}

### 2023-03-17 article3.qmd
update

@@
 In the meantime I asked myself, what could be a cool thing to do about those data?
-I already had some experience with BGC-Argo floats during my master's degree (study of the deep chlorophyll maximum in the Black Sea, see also the abstract of that paper in the annexes) and so while I was waiting for new profiles, I thought it would be useful to have a visualization tool in the form of a web app to investigate incoming data and already prepare the ground for the following study.
+I already had some experience with BGC-Argo floats during my master's degree (I published a paper on the study of the deep chlorophyll maximum in the Black Sea, [DOI: 10.5194/bg-18-755-2021](https://bg.copernicus.org/articles/18/755/2021/)) and so while I was waiting for new profiles, I thought it would be useful to have a visualization tool in the form of a web app to investigate incoming data and already prepare the ground for the following study.
 Here is the link to the [web app](https://fricour.shinyapps.io/CARBON_REFINE/), which is still currently being automatically updated every day with new data (almost) live from the Labrador Sea and hopefully with other BGC-Argo floats equipped with the UVP6 and the OST to investigate carbon and particle fluxes in other regions.
@@
 
-p <- basemap(limits = c(-55,-40,55,65), bathymetry = T, lat.interval = 2.5, bathy.style = 'poly_blues')
+p <- basemap(limits = c(-55,-40,55,65), bathymetry = T, lat.interval = 2.5, bathy.style = 'poly_blues', bathy.alpha = .5)
 

### 2023-03-17 bcp_sequestration.qmd
update

@@
+# Revisiting the BCP with the CONVERSE framework {#bcp-sequestration}
+
+Initially, it was not intended that I work on this topic.
+As I will briefly explain in the general discussion, my attempts to find a global (or regional) relationship to derive POC fluxes from UVP5 size spectra failed constantly.
+Therefore, I was proposed to work on a new concept related to carbon sequestration.
+I was told I would be co-author, that it would be done quickly and that it was only about revising the carbon sequestration flux of the gravitational pump.
+
+Why revise one pump when you can revise them all?
+1 year and a half later, hundreds of emails, one promotion to first author, one negative review and one strong rebuttal, we are still on the road to publication.
+Here below is the revised version of our paper submitted in Nature Geoscience about the reassessment of the global sequestration (\> 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
+
 ```{=latex}

### 2023-03-17 observingCarbonCycle.qmd
update

@@
 They are typically deployed for several years, which requires special attention to the energy consumption of each sensor.
-In @sec-uvp-ia, I will therefore describe the methodology used to develop the zooplankton classification model for the UVP6 accounting for energy limitations.
+In @sec-uvp6, I will therefore describe the methodology used to develop the zooplankton classification model for the UVP6 accounting for energy limitations.
 2.
-Four BGC-Argo floats equipped with both a transmissometer and a UVP6 were deployed for the first time in the Labrador Sea in late May 2022 with the aim of observing POC export events due to the annual spring bloom.
-Unfortunately, it appears that the deployment occurred at the end of the spring bloom therefore we likely missed the major export events.
-In @sec-case-study-labrador-sea, I will therefore study the aftermaths of the spring bloom in terms of carbon fluxes estimated by both instruments.
-I will also compare the small (continuous) and big (stochastic) flux of particles as seen by transmissometers used as OSTs with PSDs measurements by UVP6s at three different drifting depths to investigate the relationship between the two instruments.
+Four BGC-Argo floats equipped with both an OST and the UVP6 were deployed for the first time in the Labrador Sea in late May 2022 to study the dynamics of carbon fluxes and particle size distributions.
+In this chapter, I will explore the seasonality of the later and compare how the two sensors relate as two independent measurements of particle and carbon fluxes.
 3.
 Many studies have assumed that carbon sequestration only occurs in the deep ocean (typically 1000 m).
-In @sec-to-do, we reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
+In @bcp-sequestration, we reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.

### 2023-03-18 discussion.qmd
update

@@
 
-### Classification metric
+### Trust index
 
@@
 In @sec-uvp6, I explained that XGBoost minimizes a regularized objective function $L$ (see @eq-objective-function-xgboost) during model training.
-The XGBoost library proposes two choices of objective functions for multiclass classification problems: `multi:softmax` and `multi:softprob` (https://xgboost.readthedocs.io/en/latest/parameter.html, last accessed: 03/03/2023).
+In practice, the XGBoost library proposes two choices of objective functions for multiclass classification problems: `multi:softmax` and `multi:softprob` (https://xgboost.readthedocs.io/en/latest/parameter.html, last accessed: 03/03/2023).
 Both functions are based on the softmax normalization which normalizes the classification score of each class using an exponential function (@eq-softmax).
-`multi:softmax` returns the class of highest probability whereas `multi:softprob` outputs a class probability array, that allows the computation of custom objective functions.
+`multi:softmax` returns the class of highest probability whereas `multi:softprob` outputs a class probability array, which allows the computation of custom objective functions.
 
@@
 where $x_{i}$ and $N$ are, respectively, the classification score of the $i^{th}$ class and the number of classes.
+
+::: {.blackbox data-latex=""}
+::: {.flushleft data-latex=""}
+**Information Box 4.** **Cross entropy loss**
+:::
+
+Cross entropy loss is a loss function often used in multiclass classification problems [@Geron2019-is].
+
+$$
+J({\Theta}) = -\frac{1}{m}\sum_{i=1}^{m}\sum_{k=1}^{K}y_{k}^{(i)}log(\hat{p}_{k}^{(i)})
+$$
+
+where $J$ is the loss (i.e.
+cost) function, $\Theta$ a parameter matrix, $m$ the number of training data (i.e.
+instances), $K$ the number of classes, $y_{k}^{(i)}$ the true probability that the $i^{th}$ instance belongs to class $k$ (0 or 1) and $\hat{p}_{k}^{(i)}$ the predicted probability that the instance belongs to class $k$.
+The XGBoost library can however be confusing in comparison with the paper of @Chen2016-xt and @eq-objective-function-xgboost.
+In practice, if one sets the objective function to `multi:softmax`, it means that the objective of the model is to predict the highest probability for the true class while minimizing a loss function specified by the user (in our case, cross entropy).
+:::
 
@@
 The two most energy consuming steps are the image segmentation (median execution time of 650 ms) and the loading of the classification model whose execution time was capped at 700 ms.
-This limitation imposed a direct constraint on the prediction quality of the model because it implied a limitation on the choice of values for hyperparameters of the model (learning rate, tree depth) and the number of boosting rounds.
+This limitation imposed a direct constraint on the prediction quality of the model because it implied a limitation on the choice of the model hyperparameters (learning rate, tree depth) and the number of boosting rounds.
 
@@
 In the UVP5 SD data set, the most detailed classification at a taxonomic level by a human was roughly 150 classes (referred to as $level_{0}$).
-It was therefore necessary to merge classes based on morphological and ecological attributes to decrease that number in order to get to $level_{1}$ that had 38 classes.
-From $level_{1}$, we thus experimented a lot of different prediction models with a hyperparameters gridsearch in order to find both the right set of parameters and the right number of classes.
-There were indeed some classes that had very low precision and recall because they were confused with some specific taxonomic groups.
-They were merged accordingly with those groups.
+It was therefore necessary to merge classes based on morphological and ecological attributes to decrease that number in order to get to $level_{1}$ that had 38 classes (limited at 40).
+From $level_{1}$, we thus experimented a lot of different prediction models with a hyperparameters gridsearch in order to find both the right set of hyperparameters and the right number of classes.
+Still, some $level_{1}$ classes had very low precision and recall scores because they were confused with some specific taxonomic groups hence we merged them accordingly with those groups.
 This trial-error approach was necessary to create the final aggregation level ($level_{2}$) used for the final UVP5 model.
@@
 
-The last technical limitation was the number of features (see table XX).
+The last technical limitation was the number of features (see @tbl-features).
 Before I started my PhD, 55 features (basically geometry and gray levels) were defined.
@@
 Why was this proposition refused?
-Again, those 55 selected features had been tested earlier and they were compliant with the energy balance.
-Therefore, adding more features was not a possibility (we would have needed to redo tests and energy budget that were not possible at the time) and because they were computed quickly, they were not consuming too much energy hence there was no need to remove some of them.
+Again, those 55 selected features had been tested earlier and they were compliant with the energy budget.
+Therefore, adding more features was not a possibility (we would have needed to redo tests and energy budgets that were not possible at the time) and because they were computed quickly, they were not consuming too much energy hence there was no critical need to remove some of them.
 At worse, some features would have been redundant, without impacting the classification quality.
 
-Nonetheless, when we compare the results from a featured-based model and a CNN (see SECTION XX), perhaps the results would have been closer if we would have done some feature engineering to add other kinds of features.
+Nonetheless, when we compare the results from a features-based model and a CNN (@tbl-classification-report-uvp6-xgboost-cnn), perhaps the results would have been closer if we did some feature engineering and added other kinds of features.
 
@@
 #| label: tbl-stats-uvp6
-#| tbl-cap: Mean and median concentrations and number of segmented objects whose size is above 0.6 mm(referred to as "vignettes" in french) expected from a standard vertical profile taken by a BGC-Argo float equipped with a UVP6 in different parts of the water column.
+#| tbl-cap: Mean and median concentrations and number of segmented objects whose size is above 0.6 mm (referred to as "vignettes" in french) expected from a standard vertical profile taken by a BGC-Argo float equipped with a UVP6 in different parts of the water column.
 
@@
 
-feature engineering, citer le bouquin là
+### Detritus, this ubiquitous mess
+
+We saw in the previous section that the median number of objects detected in 3.5 m$^{3}$ is roughly 300.
+In addition, it is usual to have less than 20% of living objects hence the vast majority of images are labelled as detritus.
+The latter are ubiquitous in the ocean and display a wide variety of shapes, sizes and textures (@fig-detritus).
+In @sec-improve-classif, I showed that using several classes of detritus did not improve the classification model and that it actually made it worse because the N-detritus classes showed a high level of confusion between each other.
+
+![Examples of UVP objects labelled as detritus. The scale bar is the same for all (2 mm).](image/detritus/moz_detritus.jpg){#fig-detritus}
+
+![Examples of UVP objects labelled as living material.
+The scale bar is the same for all (2 mm).
+Note that I have hidden three detritus from @fig-detritus to highlight the difficulty of classifying plankton already at the human level.](image/detritus/uvp_user_meeting/living_except3.jpg){#fig-living}
+
+We assumed that detritus form a continuum in the 55-dimensional feature space.
+Therefore, we tried several methods to break this continuum while also trying to cope with the data set class imbalance.
+For this, we tried under-sampling them to decrease their number and non random under-sampling by keeping only detritus close to K-means clustered centroids but it did not improve the classification model.
+Using K-means to cluster detritus into several classes (e.g.
+fluffy gray or dark compact) to improve the classification of biological objects was based on the strong belief of some colleagues that were convinced it would work.
+My work and the work of another PhD student that tried using several detritus classes on a CNN (not published yet) showed that it did not improve the classification model.
+However, it is true that other clustering methods could be further explored.
+On the top of my head, I think about DBSCAN [@Schubert2017-xr] and HDBSCAN [@Campello-2013-cb] which are both density-based clustering methods in contrast to K-Means with is a partitioning clustering method that needs a predefined number of clusters in input.
+Other clustering methods could obviously be explored [see @Geron2019-is, pp.
+259-260] based on their computational complexity and the data set characteristics (number of instances and features).
+
+### The future of the embedded classification
+
+#### UVPec
+
+[UVPec](https://github.com/ecotaxa/uvpec/) - short name for Underwater Vision Profiler Embedded Classifier - is a toolbox to train classification models for UVP6 images and evaluate their performance.
+This python package has been designed to be as user-friendly as possible in order to be accessible to the largest number of users.
+Basically, the only mission of the user is to create folders (40 maximum because the maximum number of classes is fixed at 40) of labelled UVP6 images and specify some hyperparameters (e.g.
+learning rate) to train a classification model.
+The evaluation of the created model provides a confusion matrix and a classification report to help the user evaluate the quality of the model as well as the cross entropy curve to have information on the optimal boosting round number (before the model overfits).
+
+#### REFINE and further results (with the examples of Hervé)
+
+#### New methods for zooplankton (CNN embarked and tranformers) for low energy stems
+
+#### UVP6 micro -\> UVP6 meeting, aller chercher la présentation de Camille sur le google drive pour les perspectives
 
 ###
-
-### Features
-
-### Bilan énergique
-
-### detritus case + imbalanced
-
-### future embedded low energy models
-
-### time series, get the UVP6 data back (intéret de les mettre sur les flotteurs)
-
-volume échantillonnée vraiment faible, les trucs petits sont importants donc il faudrait l'uVP6 micro qui est on the way

### 2023-03-18 merci.qmd
update

@@
 faire des stats du vélo (et meme mettre une carte si possible rapidement)
+
+tous ceux qui ont partagé des moments sportifs ensembles, au bloc mais surtout en vélo et qui malgré la souffrance acceptaient encore de venir rouler avec moi (Ophélie, Flavien, Thelma, Laetitia, Alexandre, Satoshi)
+
+Julien, des challenges à n'en plus finir, un peu mazo au final, à force d'aller en enfer, tu incarneras le diable en personne bientôt(#Ducasse)
+
+Les montois (Mathieu, inconditionel, Florent, les Juliens, Chris)
+
+Les co-Phd : Chloé, AMP (the best), Anais
+
+Hadrien who accepted to be my roomate in Liège
+
+alice
+
+marine of course
+
+expatrié que je suis et qui venu me voir 'ou profiter, I'll never really know)
+
+les Liégeois (Thomas, Nico, Hadri, Flop, Florent)
+
+Lucie, mon old roomate (lucie lelotte)
+
+Alex, my 15th french history roomate, I am glad you were there suring the 6 pm curfew we had and
+
+Les montois bien sûr
+
+tous xeux qui sont descendus me voir, (vous veniez pour moi ou la mer tirquoise, j'ai des doutes mtn)

### 2023-03-18 uvp6_part.qmd
update

@@
 To be even more honest, I basically knew nothing about zooplankton either.
-It is time to admit it.
+I think it is time to admit it.
 
@@
 The first step consisted in developing a full data processing pipeline using UVP5 images.
-The aim was to be prepared for the UVP6 image data set (not ready at the time) to train the true final classification model.
+The aim was to be prepared for the UVP6 image data set (not ready at that time) to train the true final classification model.
 To this end, I developed a python package called [UVPec](https://github.com/ecotaxa/uvpec) that is able to create a full classification model based on image folders with labelled particle/zooplankton images and some hyperparameters (e.g.
@@
 
-#### Ways to improve the classification
+#### Ways to improve the classification {#sec-improve-classif}
 

### 2023-03-19 discussion.qmd
update general discussion

@@
 instances), $K$ the number of classes, $y_{k}^{(i)}$ the true probability that the $i^{th}$ instance belongs to class $k$ (0 or 1) and $\hat{p}_{k}^{(i)}$ the predicted probability that the instance belongs to class $k$.
-The XGBoost library can however be confusing in comparison with the paper of @Chen2016-xt and @eq-objective-function-xgboost.
+However, the XGBoost library can be confusing in comparison with the paper of @Chen2016-xt and especially @eq-objective-function-xgboost.
 In practice, if one sets the objective function to `multi:softmax`, it means that the objective of the model is to predict the highest probability for the true class while minimizing a loss function specified by the user (in our case, cross entropy).
@@
 
-```{r, trust index, dpi = 300}
+```{r, trust index, dpi = 300, fig.height = 12, fig.width= 10}
 #| label: fig-trust-index
@@
   rasterGrob(as.raster(readJPEG(paste0(path, "/fiber_detritus.jpeg")))),
-  ncol = 3
+  ncol = 2
 )
@@
 The second critical aspect of the model was its limitation at 40 classes.
-In the UVP5 SD data set, the most detailed classification at a taxonomic level by a human was roughly 150 classes (referred to as $level_{0}$).
-It was therefore necessary to merge classes based on morphological and ecological attributes to decrease that number in order to get to $level_{1}$ that had 38 classes (limited at 40).
+In the UVP5 SD data set, the most detailed (human) classification at a taxonomic level was roughly 150 classes (referred to as $level_{0}$).
+It was therefore necessary to merge classes based on morphological and ecological attributes to decrease that number in order to get to $level_{1}$ that had 38 classes (reminder, it is currently limited at 40).
 From $level_{1}$, we thus experimented a lot of different prediction models with a hyperparameters gridsearch in order to find both the right set of hyperparameters and the right number of classes.
@@
 The last technical limitation was the number of features (see @tbl-features).
-Before I started my PhD, 55 features (basically geometry and gray levels) were defined.
+Before I started my PhD, 55 features (basically geometry, gray levels, statistical moments) were defined.
 I was not allowed (though, I asked!) to do some feature engineering, that is either doing feature selection (selecting the most useful features) either doing feature combination (combining existing features to create new ones) or creating new features by collecting new data [@Geron2019-is].
@@
 Again, those 55 selected features had been tested earlier and they were compliant with the energy budget.
-Therefore, adding more features was not a possibility (we would have needed to redo tests and energy budgets that were not possible at the time) and because they were computed quickly, they were not consuming too much energy hence there was no critical need to remove some of them.
+Therefore, adding more features was not a possibility because we would have needed to redo tests and energy budgets that were not possible at the time.
+In addition, all 55 features were computed quickly hence they were not consuming a lot of energy so it was decided to just keep them.
 At worse, some features would have been redundant, without impacting the classification quality.
 
-Nonetheless, when we compare the results from a features-based model and a CNN (@tbl-classification-report-uvp6-xgboost-cnn), perhaps the results would have been closer if we did some feature engineering and added other kinds of features.
+Nonetheless, when we compare the results from a features-based model and a CNN (@tbl-classification-report-uvp6-xgboost-cnn), perhaps the results would have been closer if we actually did some feature engineering and added other kinds of features.
 
@@
 Therefore, we tried several methods to break this continuum while also trying to cope with the data set class imbalance.
-For this, we tried under-sampling them to decrease their number and non random under-sampling by keeping only detritus close to K-means clustered centroids but it did not improve the classification model.
-Using K-means to cluster detritus into several classes (e.g.
-fluffy gray or dark compact) to improve the classification of biological objects was based on the strong belief of some colleagues that were convinced it would work.
-My work and the work of another PhD student that tried using several detritus classes on a CNN (not published yet) showed that it did not improve the classification model.
-However, it is true that other clustering methods could be further explored.
+For this, we tried under-sampling the detritus to decrease their number and non random under-sampling by keeping only detritus close to K-means clustered centroids but it did not improve the classification model.
+Using a clustering method to split the detritus into several classes (e.g.
+fluffy gray or dark compact) to improve the classification of biological objects was based on the strong belief of some colleagues that were convinced it would improve the classification.
+However, my work and the work of another PhD student that tried using several detritus classes using a CNN (not published yet) showed that it did not improve the classification model either.
+Nonetheless, it is true that other clustering methods could be further explored.
 On the top of my head, I think about DBSCAN [@Schubert2017-xr] and HDBSCAN [@Campello-2013-cb] which are both density-based clustering methods in contrast to K-Means with is a partitioning clustering method that needs a predefined number of clusters in input.
@@
 
-#### REFINE and further results (with the examples of Hervé)
+Overall, the goal of this package is to propose a simple and fast way to create new classification models for UVP6 applications on platforms that need to send data in near-real time or that cannot be easily retrieved.
 
-#### New methods for zooplankton (CNN embarked and tranformers) for low energy stems
+#### More regions to be explored
 
-#### UVP6 micro -\> UVP6 meeting, aller chercher la présentation de Camille sur le google drive pour les perspectives
+The Labrador Sea was the first of 9 key oceanic regions studied in the frame of the REFINE (Robots Explore plankton-driven Fluxes in the marine twIlight zoNE) project (see https://erc-refine.eu).
+The latter aims to study regions with distinct biogeochemical conditions and responses to climate (@fig-refine).
 
-###
+```{r, map refine}
+#| label: fig-refine
+#| fig-cap: Oceanic regions (Baffin Bay (BB), Labrador Sea (LS), Greenland Sea (GS), Guinea Dome (GD), Arabian Sea (AS), Tropical Indian Ocean (TIO), South Pacific subtropical Gyre (SPG), West Kerguelen Plateau (WKP), East Kerguelen Plateau (EKP)) investigated in the frame of REFINE.
+
+library(tmap)
+library(tibble)
+data("World")
+library(sf)
+
+regions <- tibble(Region = c('BB', 'GS', 'LS', 'GD', 'AS', 'TIO', 'SPG', 'WKP', 'EKP'),
+                  long_name = c('Baffin Bay', 'Greenland Sea', 'Labrador Sea', 'Guinea Dome',
+                                'Arabian Sea', 'Tropical Indian Ocean', 'South Pacific subtropical gyre','West Kerguelen Plateau', 'East Kerguelen Plateau'),
+                  lat = c(74, 77, 62, 9, 19, -7.7, -32, -49.6, -50),
+                  lon = c(-67.4, -3, -55, -21, 64, 54, -139, 63.6, 75.6))
+
+points_sf <- st_as_sf(regions, coords = c("lon", "lat"), crs = "+proj=longlat")
+
+# map
+tm_shape(World) +
+    tm_polygons() +
+    tm_shape(points_sf) +
+    tm_bubbles(size = .3, col = 'Region', alpha = 1, border.col = 'black', border.alpha = 1)
+
+```
+
+@fig-kerguelen present some preliminary results from the UVP6 embedded classification that highlights the difference in *Calanoida* and *Chaetognatha* concentrations between February and March 2023 in the western and eastern part of the Kerguelen Plateau.
+
+```{r, kerguelen, fig.width=8, fig.height=5}
+#| label: fig-kerguelen
+#| fig-cap: $\textit{Calanoida}$ and $\textit{Chaetognatha}$concentrations between February-March 2023 in the Kerguelen Plateau.
+
+library(vroom)
+library(tidyverse)
+library(patchwork)
+library(lubridate)
+library(viridis)
+
+# load data
+calanoids_east <- vroom('DATA/ARGO/1902593_class_number_4_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'Calanoida - East')
+
+calanoids_west <- vroom('DATA/ARGO/4903657_class_number_4_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'Calanoida - West')
+
+chaeto_east <- vroom('DATA/ARGO/1902593_class_number_5_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'Chaetognatha - East')
+chaeto_west <- vroom('DATA/ARGO/4903657_class_number_5_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'Chaetognatha - West')
+
+taxo_kerguelen <- rbind(calanoids_east, calanoids_west, chaeto_east, chaeto_west) |> 
+  mutate(group = factor(group, levels = c('Calanoida - East', 'Calanoida - West',
+                                     'Chaetognatha - East', 'Chaetognatha - West')))
+
+ggplot(taxo_kerguelen) + geom_point(aes(x = conc_object, y = depth, colour = date)) + scale_colour_viridis_c(trans = "date",labels = c("2023-02-11","2023-02-18","2023-02-25","2023-03-04","2023-03-11"), breaks = c(as.Date("2023-02-11"), as.Date("2023-02-18"), as.Date("2023-02-25"), as.Date("2023-03-04"), as.Date("2023-03-11"))) + scale_y_reverse() + theme_bw() +
+  labs(x = 'Concentration (#/L)', colour = "Date", y = 'Depth (m)') +
+  facet_wrap(~group)
+```
+
+#### New algorithms
+
+Unless we find a cheap and efficient (low power) way to send UVP6 images via Iridium communication, we are stuck with the issue of embedded classification.
+In @sec-uvp6, I showed that CNNs usually outperform features-based classifiers.
+However, the field of computer vision is constantly evolving with the arrival of new technologies that could help improve the automatic classification of zooplankton images.
+Transformers [@Vaswani2017-il], also referred to as Foundation Models in @Bommasani2021-kbare are the latest cutting-edge tool in the field of computer vision and natural language processing [e.g.
+@Li2022-qt; @Liu2021-jq].
+The structure of transformers rely on attention mechanisms able to draw global dependencies between input and output \[@Vaswani2017-il\].
+In other words, it is able to learn dependencies between distant positions in a network architecture in constrast to CNNs.
+As an example, @Li2022-qt used a vision transformer model combining a CNN and a transformer to identify apple diseases.
+In their model, the CNN is used to extract the global features of the input image while the transformer extracts local features of the disease region to help the CNN see better.
+How does it help the CNN?
+In the case of apple images with different backgrounds (different weather conditions, brightness changes, other objects in the field of view), the CNN could be "distracted" hence the transformer is able to focus the attention of the CNN on the region of interest (the disease region) and not on irrelevant features such as the weather.
+Therefore, transformers should definitely be on the to-do list of anyone working on improving automatic zooplankton classification.
+
+#### UVP6m
+
+There is a class called "other\<living" in the UVP6 classification model.
+I did not mention earlier than I had to put the copepods in this class.
+Indeed, *Calanoida* are a "large" species of copepods compared to other copepods found in the Mediterranean Sea that are smaller.
+The classification of the latter was really not good (low precision and low recall) hence we put them in the "other\<living" class, which is a shame.
+
+The good news is that the next generation of UVP is already in the making with the UVP6m (m for micro) which will be able to look at smaller objects (particle counter between 10 and 200 $\mu$m and plankton/particle imaging \> 100 $\mu$m).
+The UVP6m will be more compact than the UVP6 but the principles will be the same and both instruments will share 5 size classes for intercalibration.

### 2023-03-20 annexes.qmd
update

--- 2023-02-15 annexes.qmd
@@
 
-library(dplyr)
-library(knitr)
-library(kableExtra)
-library(purrr)
-
-# read data
-trap <- vroom::vroom('DATA/Trap_Thorium/TRAP_DATABASE.csv')
-#thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
-
-# clean a bit
-trap[trap$doi == 'https://doi.org/10.1007/s00227-018-3376-1',]$doi <- '10.1007/s00227-018-3376-1'
-trap[trap$doi == 'https://doi.org/10.1016/j.dsr.2016.08.015',]$doi <- '10.1016/j.dsr.2016.08.015'
-trap[trap$doi == 'Geochimica et Cosmochimica Acta, Vol. 66, No. 3, pp. 457–465, 2002',]$doi <- '10.1016/S0016-7037(01)00787-6'
-
-# some stats
-trap_resume <- trap |> group_by(doi) |> dplyr::summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
-  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
-
-# BASED ON : https://stackoverflow.com/questions/48506712/kable-kableextra-cells-with-hyperlinks
-# make url links
-make_link <- function(doi){
-  link <- paste0('https://doi.org/',doi)
-  return(link)
-}
-url <- map_chr(trap_resume$doi, make_link)
-
-# # put url in dataframe
-# trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, 'html', link = url)) |> select(doi_url, data_points) |>
-#   kable('html', escape = FALSE, col.names = c('Sources of data', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))
-
-trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
-  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l')
-
-# make final table
-trap_resume
+# library(dplyr)
+# library(knitr)
+# library(kableExtra)
+# library(purrr)
+# 
+# # read data
+# trap <- vroom::vroom('DATA/Trap_Thorium/TRAP_DATABASE.csv')
+# #thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
+# 
+# # clean a bit
+# trap[trap$doi == 'https://doi.org/10.1007/s00227-018-3376-1',]$doi <- '10.1007/s00227-018-3376-1'
+# trap[trap$doi == 'https://doi.org/10.1016/j.dsr.2016.08.015',]$doi <- '10.1016/j.dsr.2016.08.015'
+# trap[trap$doi == 'Geochimica et Cosmochimica Acta, Vol. 66, No. 3, pp. 457–465, 2002',]$doi <- '10.1016/S0016-7037(01)00787-6'
+# 
+# # some stats
+# trap_resume <- trap |> group_by(doi) |> dplyr::summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
+#   mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
+# 
+# # BASED ON : https://stackoverflow.com/questions/48506712/kable-kableextra-cells-with-hyperlinks
+# # make url links
+# make_link <- function(doi){
+#   link <- paste0('https://doi.org/',doi)
+#   return(link)
+# }
+# url <- map_chr(trap_resume$doi, make_link)
+# 
+# # # put url in dataframe
+# # trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, 'html', link = url)) |> select(doi_url, data_points) |>
+# #   kable('html', escape = FALSE, col.names = c('Sources of data', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))
+# 
+# trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
+#   kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l') 
+# 
+# # make final table
+# trap_resume
 ```
@@
 
-# read data
-thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
-
-# some stats
-thorium_resume <- thorium |> group_by(doi) |> summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
-  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
-thorium_resume$doi <- c('10.1594/PANGAEA.809717', '10.1029/2018GB00615', '10.1525/elementa.2020.030')
-
-# make url links
-url <- c('https://doi.org/10.1594/PANGAEA.809717', 'https://doi.org/10.1029/2018GB006158', 'https://doi.org/10.1525/elementa.2020.030')
-
-# put url in dataframe
-thorium_resume <- thorium_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
-  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))
-
-# make final table
-thorium_resume
+# # read data
+# thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
+# 
+# # some stats
+# thorium_resume <- thorium |> group_by(doi) |> summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
+#   mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
+# thorium_resume$doi <- c('10.1594/PANGAEA.809717', '10.1029/2018GB00615', '10.1525/elementa.2020.030')
+# 
+# # make url links
+# url <- c('https://doi.org/10.1594/PANGAEA.809717', 'https://doi.org/10.1029/2018GB006158', 'https://doi.org/10.1525/elementa.2020.030')
+# 
+# # put url in dataframe
+# thorium_resume <- thorium_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
+#   kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))
+# 
+# # make final table
+# thorium_resume
 ```

### 2023-03-20 discussion.qmd
update

@@
 library(sf)
+library(ggspatial)
+library(rnaturalearth)
+library(rnaturalearthdata)
+library(ggplot2)
+library(RColorBrewer)
 
@@
                   lat = c(74, 77, 62, 9, 19, -7.7, -32, -49.6, -50),
-                  lon = c(-67.4, -3, -55, -21, 64, 54, -139, 63.6, 75.6))
+                  lon = c(-67.4, -3, -55, -21, 64, 54, -130, 63.6, 75.6))
 
@@
 
-# map
+# # map
 tm_shape(World) +
@@
   tm_shape(points_sf) +
-    tm_bubbles(size = .3, col = 'Region', alpha = 1, border.col = 'black', border.alpha = 1)
+  tm_bubbles(size = .3, col = 'Region', alpha = 1, border.col = 'black', border.alpha = 1) +
+  tm_layout(legend.position = c("left", "center"), legend.frame = T)
+
+# # load world data
+# world <- ne_countries(scale = "medium", returnclass = "sf")
+# 
+# ggplot(data = world) +
+#     geom_sf() +
+#     xlab("Longitude") + ylab("Latitude") + theme_bw() +
+#     geom_point(data = regions, aes(x = lon, y = lat, colour = Region), size = 4) +
+#     scale_colour_manual(values = brewer.pal(9, "Set3")) +
+#   geom_point(data = regions, aes(x = lon, y = lat), colour = 'black', size = 4, shape = 1)
 
@@
 @fig-kerguelen present some preliminary results from the UVP6 embedded classification that highlights the difference in *Calanoida* and *Chaetognatha* concentrations between February and March 2023 in the western and eastern part of the Kerguelen Plateau.
+For these floats, the configuration was different from the ones deployed in the Labrador Sea with a higher profiling time resolution after deployment, which can be further modified based on observations.
 
@@
 #| label: fig-kerguelen
-#| fig-cap: $\textit{Calanoida}$ and $\textit{Chaetognatha}$concentrations between February-March 2023 in the Kerguelen Plateau.
+#| fig-cap: $\textit{Calanoida}$ and $\textit{Chaetognatha}$ concentrations in February-March 2023 in the Kerguelen Plateau.
 
@@
   labs(x = 'Concentration (#/L)', colour = "Date", y = 'Depth (m)') +
-  facet_wrap(~group)
+  facet_wrap(~group, scale = "free_x")
 ```
@@
 @Li2022-qt; @Liu2021-jq].
-The structure of transformers rely on attention mechanisms able to draw global dependencies between input and output \[@Vaswani2017-il\].
+The structure of transformers rely on attention mechanisms able to draw global dependencies between input and output [@Vaswani2017-il].
 In other words, it is able to learn dependencies between distant positions in a network architecture in constrast to CNNs.
@@
 In the case of apple images with different backgrounds (different weather conditions, brightness changes, other objects in the field of view), the CNN could be "distracted" hence the transformer is able to focus the attention of the CNN on the region of interest (the disease region) and not on irrelevant features such as the weather.
-Therefore, transformers should definitely be on the to-do list of anyone working on improving automatic zooplankton classification.
+The use of transformers for automatic zooplankton classification has just begun [@Kyathanahally2022-ma] but it should certainly be explored further in the near future.
 
@@
 There is a class called "other\<living" in the UVP6 classification model.
-I did not mention earlier than I had to put the copepods in this class.
+I did not mention earlier than I was obliged to place copepods (except *Calanoida*) in that class.
 Indeed, *Calanoida* are a "large" species of copepods compared to other copepods found in the Mediterranean Sea that are smaller.
-The classification of the latter was really not good (low precision and low recall) hence we put them in the "other\<living" class, which is a shame.
+The classification of the latter was not good (low precision and low recall) hence we put them in the "other\<living" class, which is a shame.
 
@@
 The UVP6m will be more compact than the UVP6 but the principles will be the same and both instruments will share 5 size classes for intercalibration.
+
+## Deriving carbon fluxes from particle size distribution
+
+Before working on the CONVERSE framework, I was trying to find a global relationship (I also tried to find a regional one, without more success) to derive carbon fluxes from UVP5 PSDs.
+The idea was to reproduce the method of @Guidi2008-vp on the newly quality controlled UVP5 PSD data set [@Kiko2022-sk].
+For this, I needed to build a global data set of POC measurements derived from sediment traps and $^{234}Th$ (see @tbl-poc-from-sediment-traps, @tbl-poc-from-thorium and [here](https://fricour.shinyapps.io/carbon_fluxes_app/) for a more interactive view).
+Here below, I briefly review three methods we explored to find a POC-PSD relationship and what could be further investigated.
+
+```{=latex}
+%\includepdf[pages=1]{"/home/flo/dox/THESIS/bg-18-755-2021"}
+```
+```{=html}
+<!--#https://github.com/quarto-dev/quarto-cli/discussions/1042-->
+<style type="text/css">
+caption, .table-caption {
+  text-align: left;
+}
+</style>
+```
+```{r, poc database from traps}
+#| label: tbl-poc-from-sediment-traps
+#| tbl-cap: Compilation of POC flux measurements from sediment traps.
+Data taken from public databases listed on PANGAEA or extracted from articles (with the help of Frédéric Le Moigne and Meike Vogt).
+The number of measurements could be different from the original publication after data cleaning and potential removal of duplicates.
+#| tbl-cap-location: bottom
+
+library(dplyr)
+library(knitr)
+library(kableExtra)
+library(purrr)
+
+# read data
+trap <- vroom::vroom('DATA/Trap_Thorium/TRAP_DATABASE.csv')
+#thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
+
+# clean a bit
+trap[trap$doi == 'https://doi.org/10.1007/s00227-018-3376-1',]$doi <- '10.1007/s00227-018-3376-1'
+trap[trap$doi == 'https://doi.org/10.1016/j.dsr.2016.08.015',]$doi <- '10.1016/j.dsr.2016.08.015'
+trap[trap$doi == 'Geochimica et Cosmochimica Acta, Vol. 66, No. 3, pp. 457–465, 2002',]$doi <- '10.1016/S0016-7037(01)00787-6'
+
+# some stats
+trap_resume <- trap |> group_by(doi) |> dplyr::summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
+  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
+
+# BASED ON : https://stackoverflow.com/questions/48506712/kable-kableextra-cells-with-hyperlinks
+# make url links
+make_link <- function(doi){
+  link <- paste0('https://doi.org/',doi)
+  return(link)
+}
+url <- map_chr(trap_resume$doi, make_link)
+
+# # put url in dataframe
+# trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, 'html', link = url)) |> select(doi_url, data_points) |>
+#   kable('html', escape = FALSE, col.names = c('Sources of data', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))
+
+trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
+  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of observations'), align = 'l')
+
+# make final table
+trap_resume
+```
+
+```{r, poc database from thorium}
+#| label: tbl-poc-from-thorium
+#| tbl-cap: Compilation of POC flux measurements derived from Thorium 234.
+Data taken from public databases listed on PANGAEA or extracted from articles (with the help of Frédéric Le Moigne).
+The number of measurements could be different from the original publication after data cleaning and potential removal of duplicates.
+#| tbl-cap-location: bottom
+
+# read data
+thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')
+
+# some stats
+thorium_resume <- thorium |> group_by(doi) |> summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
+  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
+thorium_resume$doi <- c('10.1594/PANGAEA.809717', '10.1029/2018GB00615', '10.1525/elementa.2020.030')
+
+# make url links
+url <- c('https://doi.org/10.1594/PANGAEA.809717', 'https://doi.org/10.1029/2018GB006158', 'https://doi.org/10.1525/elementa.2020.030')
+
+# put url in dataframe
+thorium_resume <- thorium_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
+  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of observations'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))
+
+# make final table
+thorium_resume
+```
+
+### The straightforward approach
+
+This approach consisted in simply applying the method of @Guidi2008-vp (see @eq-carbon-flux-finite).
+Therefore, I looked for match-ups (i.e.
+spatially and temporally close observations) between UVP5 and POC observations within a 0.2° latitude and longitude window, a 10 m vertical depth range and a 5-day window.
+@Guidi2008-vp found 118 match-ups (with UVP2/3/4 data) whereas I only found 108 match-ups with UVP5 data, located on the western and eastern parts of the Kerguelen Plateau, in the Guinea Dome and in the California Current (panel D in @fig-uvp-trap-mapping).
+As a result, we decided not to go further because we would not have enough match-ups to improve the minimization (see @eq-cost-function).
+
+```{r, uvp traps th234 map}
+#| label: fig-uvp-trap-mapping
+#| fig-cap: Spatial distribution of sediment traps (A), $^{234}Th$ (B), UVP5 (C) and their spatio-temporal match-ups (D)
+
+d <- vroom::vroom('DATA/CFLUX/uvp_trap.csv')
+guidi_matchups <- vroom::vroom('DATA/CFLUX/guidi_matchups.csv')
+
+d_sf <- st_as_sf(d, coords = c("lon", "lat"), crs = "+proj=longlat")
+
+d_uvp <- d_sf |> filter(source == 'UVP5')
+d_trap <- d_sf |> filter(source == 'Trap')
+d_th <- d_sf |> filter(source == 'Thorium')
+d_matchups <- st_as_sf(guidi_matchups, coords = c("lon", "lat"), crs = "+proj=longlat")
+
+# map
+# tm_shape(World) +
+#   tm_polygons() +
+#   tm_shape(d_sf) +
+#   tm_dots(size = .2, col = "black", alpha = 1, border.col = 'black', border.alpha = 1) +
+#   tm_layout(legend.show = F) +
+#   tm_facets(by = 'source')
+
+# up left
+ul <- tm_shape(World) +
+  tm_polygons() +
+  tm_shape(d_trap) +
+  tm_bubbles(size = .05, col = "#1B9E77", alpha = 1, border.col = 'black', border.alpha = 1) +
+  tm_layout(legend.show = F, main.title = 'A', main.title.position = 'center', main.title.size = 1)
+
+ur <- tm_shape(World) +
+  tm_polygons() +
+  tm_shape(d_th) +
+  tm_bubbles(size = .05, col = "#D95F02", alpha = 1, border.col = 'black', border.alpha = 1) +
+  tm_layout(legend.show = F, main.title = 'B', main.title.position = 'center', main.title.size = 1)
+
+bl <- tm_shape(World) +
+  tm_polygons() +
+  tm_shape(d_uvp) +
+  tm_bubbles(size = .05, col = "#7570B3", alpha = 1, border.col = 'black', border.alpha = 1) +
+  tm_layout(legend.show = F, main.title = 'C', main.title.position = 'center', main.title.size = 1)
+
+br <- tm_shape(World) +
+  tm_polygons() +
+  tm_shape(d_matchups) +
+  tm_bubbles(size = .05, col = "#E7298A", alpha = 1, border.col = 'black', border.alpha = 1) +
+  tm_layout(legend.show = F, main.title = 'D', main.title.position = 'center', main.title.size = 1)
+
+tmap_arrange(ul, ur, bl, br, ncol = 2, nrow = 2)
+
+# https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/
+# display.brewer.all()
+# brewer.pal(n = 8, name = "Dark2")
+
+```
+
+### The environmental match-ups approach
+
+The spatio-temporal approach lacks match-ups between UVP5 and sediment traps and $^{234}Th$ data.
+This is not really surprising when we look at their spatial distribution (@fig-uvp-trap-mapping) and especially their temporal distribution (@fig-sediment-trap-thorium-yearly-distribution) because the UVP5 was firstly used in 2008 while the 'historical' data reached their observation peak before 2000.
+Therefore, we thought of using environmental data to increase the number of match-ups.
+
+The idea was based on the assumption that carbon fluxes could be similar in distinct regions of the ocean if those regions had similar environmental conditions thus removing spatial (horizontal) and temporal constraints on match-ups.
+For this, I tried different combinations of environmental variables such as surface temperature, salinity, nutrients, combination of nutrients, Chla, $b_{bp}$ or water column properties such as integrated Chla content, oxygen minimum and MLD from monthly climatologies.
+Then, I used those environmental variables to build a 3D environmental space using a 3D principal component analysis (PCA) and I projected UVP5 data and trap/$^{234}Th$ in that space based on the environmental characteristics from the respective coordinates and month of each UVP5 and trap/$^{234}Th$ profiles.
+
+Once all data were projected in the environmental space, I had to find a new match-up procedure.
+Because the PCA projection conserves the euclidean distance, I firstly computed the pairwise distance between each UVP5 profile and all trap/$^{234}Th$ profiles.
+Then, I looked at the minimal euclidean distance for each UVP5 profile and I plotted the minimal distance distribution (\~8000 points in the distribution).
+From this distribution, I computed a radius of interest (ROI) corresponding to a threshold based on the 95$^{th}$ percentile (considered as a rather objective decision criterion).
+For each UVP5 profile in the environmental space, I kept the closest trap/$^{234}Th$ profiles within that radius (@fig-match-ups).
+Finally, I still needed to find the PSD-POC match-up on the vertical axis (depth).
+For this, I used a depth range proportional to the depth of the trap or $^{234}Th$ measurement following Martin's curve [@Martin1987-um] because the downward carbon flux is less attenuated with depth.
+Then, I averaged PSDs in that depth range (weighted by the water volume) to obtain a table with weighted averaged PSDs in front of POC fluxes to do the minimization procedure of @Guidi2008-vp.
+
+![Examples of match-ups in the environmental space for 4 UVP5 profiles (red dots) and their respective trap/$^{234}Th$ profiles within the ROI.](DATA/CFLUX/moz.jpeg){#fig-match-ups}
+
+dire que le choix des variables ( + le fait que ce sont des climatos -\> bof bof, ça manquait de variavbles physiques, pas assez de variables probablement) et aussi discuter du plan de phase (voir Rmarkdown de la prez waringan + revoir cette prez là pour vérifier que l'on a rien oublié) + critiquer la figure de lionel car pour moi le problème se situe déjà là, j'ai obsrvé que ça fitte tout, on prend un petit b pour u petite A et grand b pour un grand A
+
+Bottom line, I failed. I tried multiple times but it never worked
+
+### The machine learning approach
+
+on a fait avec des cimatos etc mais après y a moyen d'ajouter des meilleires données (eg alberto)
+
+-   faire un résumé de la photo du tableau sur le tableau blanc d Lionel.. pour dire la marche à suivre.. et les ouvelles données environnemntales aussi et le multioutpubut graidnet boosting

### 2023-03-20 merci.qmd
update general discussion
update

@@
-remerciement
+# Acknowledgments
 
@@
 
-anais
+merci au comité de thèse
+
+alberto d'avoir fourni du travail que je n'ai malhereusement pas eu le temps d'exploiter
+
+merci à mon bi-encadrement Liégeois and Niçois
+
+Merci à Hervé, Lionel et Jean-Olivier pour leur aide, Marc et Camille car l'UVP6 c'était un fameux truc
+
+checker le site du LOV pour n'oublier personne (sauf ceux qui sont partis evidemment).
+
+Evgeny et Arthur Capet et Charles Troupin
+
+I thank the ACRI team for the Nice AC-cooled office I had during last Summer because my landlord kicked me out for the summer (tourists airbnb \> yearly workers, business as usual), c'est aussi ça la côte d'azur.
+That's also why I wanted to finish in May because of that and because the Ducasse of Mons was the 4th of June, nothing is worth missing it.
+
+i am pretty sure they are as happy as I am it comes to an end because I really don't fit in the academic system at all.
 
@@
 
-tous ceux qui ont partagé des moments sportifs ensembles, au bloc mais surtout en vélo et qui malgré la souffrance acceptaient encore de venir rouler avec moi (Ophélie, Flavien, Thelma, Laetitia, Alexandre, Satoshi)
+tous ceux qui ont partagé des moments sportifs ensembles, au bloc mais surtout en vélo et qui malgré la souffrance acceptaient encore de venir rouler avec moi (Ophélie, Flavien, Thelma, Laetitia, Alexandre, Satoshi, Lucas)
+
+Ophélie, je me grouille et j'arrive pour grimper, Alexandre, Switzeland see what you got in the legs
 
@@
 tous xeux qui sont descendus me voir, (vous veniez pour moi ou la mer tirquoise, j'ai des doutes mtn)
+
+et le plus important, les montois de périront pas, ça va casser la baraque
+
+Famille et amis
+
+Girlfriend (because I am not easy going every day)

### 2023-03-21 discussion.qmd
update discussion

@@
 
-This lead us to define what we called a trust index, that is a metric whose score indicates the degree of confidence in the classification score.
+This led us to define what we called a trust index, that is a metric whose score indicates the degree of confidence in the classification score.
 In other words, we looked for the best error estimator (@fig-trust-index).
@@
 We compared those metrics for each class of the test set in @fig-trust-index which clearly shows that all metrics were very similar and that they captured the errors at the same speed.
-Thus, it meant that we could keep using XGBoost as we initially did, that is using `multi:softprob` as the objective function.
+Thus, it meant that we could keep using the probability returned by XGBoost a satisfactory classification score using `multi:softprob` as the objective function.
 @fig-trust-index also shows that some classes kept accumulating errors in the prediction with increasing rank (e.g.
@@
 Before I started my PhD, 55 features (basically geometry, gray levels, statistical moments) were defined.
-I was not allowed (though, I asked!) to do some feature engineering, that is either doing feature selection (selecting the most useful features) either doing feature combination (combining existing features to create new ones) or creating new features by collecting new data [@Geron2019-is].
+I was not allowed (although, I asked!) to do some feature engineering, that is either doing feature selection (selecting the most useful features) either doing feature combination (combining existing features to create new ones) or creating new features by collecting new data [@Geron2019-is].
 Why was this proposition refused?
-Again, those 55 selected features had been tested earlier and they were compliant with the energy budget.
-Therefore, adding more features was not a possibility because we would have needed to redo tests and energy budgets that were not possible at the time.
-In addition, all 55 features were computed quickly hence they were not consuming a lot of energy so it was decided to just keep them.
+Again, the 55 selected features had been tested earlier and they were compliant with the energy budget.
+Therefore, adding more features was not a possibility because it would have required retesting them and reevaluting energy budgets that were not possible at the time.
+In addition, all 55 features were computed quickly hence they were not consuming a lot of energy so it was decided to simply keep them.
 At worse, some features would have been redundant, without impacting the classification quality.
@@
 Nonetheless, it is true that other clustering methods could be further explored.
-On the top of my head, I think about DBSCAN [@Schubert2017-xr] and HDBSCAN [@Campello-2013-cb] which are both density-based clustering methods in contrast to K-Means with is a partitioning clustering method that needs a predefined number of clusters in input.
+Clustering algorithms such as DBSCAN [@Schubert2017-xr] and HDBSCAN [@Campello2013-cb] are density-based in contrast to K-Means with is a partitioning clustering algorithm that needs a predefined number of clusters in input.
 Other clustering methods could obviously be explored [see @Geron2019-is, pp.
@@
 This python package has been designed to be as user-friendly as possible in order to be accessible to the largest number of users.
-Basically, the only mission of the user is to create folders (40 maximum because the maximum number of classes is fixed at 40) of labelled UVP6 images and specify some hyperparameters (e.g.
+Basically, the user only needs to create folders (40 maximum because the maximum number of classes is fixed at 40) of labelled UVP6 images and specify some hyperparameters (e.g.
 learning rate) to train a classification model.
@@
 For these floats, the configuration was different from the ones deployed in the Labrador Sea with a higher profiling time resolution after deployment, which can be further modified based on observations.
+It is only the beginning.
 
@@
 However, the field of computer vision is constantly evolving with the arrival of new technologies that could help improve the automatic classification of zooplankton images.
-Transformers [@Vaswani2017-il], also referred to as Foundation Models in @Bommasani2021-kbare are the latest cutting-edge tool in the field of computer vision and natural language processing [e.g.
+Transformers [@Vaswani2017-il], also referred to as Foundation Models in @Bommasani2021-kb are the latest cutting-edge tool in the field of computer vision and natural language processing [e.g.
 @Li2022-qt; @Liu2021-jq].
-The structure of transformers rely on attention mechanisms able to draw global dependencies between input and output [@Vaswani2017-il].
+The structure of transformers relies on attention mechanisms able to draw global dependencies between input and output [@Vaswani2017-il].
 In other words, it is able to learn dependencies between distant positions in a network architecture in constrast to CNNs.
@@
 There is a class called "other\<living" in the UVP6 classification model.
-I did not mention earlier than I was obliged to place copepods (except *Calanoida*) in that class.
+I did not mention earlier than we were forced to place copepods (except *Calanoida*) in that class.
 Indeed, *Calanoida* are a "large" species of copepods compared to other copepods found in the Mediterranean Sea that are smaller.
-The classification of the latter was not good (low precision and low recall) hence we put them in the "other\<living" class, which is a shame.
+The classification score of the latter was low (low precision and low recall) hence we put them in the "other\<living" class, which is a shame.
 
@@
 This approach consisted in simply applying the method of @Guidi2008-vp (see @eq-carbon-flux-finite).
-Therefore, I looked for match-ups (i.e.
+Therefore, we looked for match-ups (i.e.
 spatially and temporally close observations) between UVP5 and POC observations within a 0.2° latitude and longitude window, a 10 m vertical depth range and a 5-day window.
-@Guidi2008-vp found 118 match-ups (with UVP2/3/4 data) whereas I only found 108 match-ups with UVP5 data, located on the western and eastern parts of the Kerguelen Plateau, in the Guinea Dome and in the California Current (panel D in @fig-uvp-trap-mapping).
-As a result, we decided not to go further because we would not have enough match-ups to improve the minimization (see @eq-cost-function).
+@Guidi2008-vp found 118 match-ups (with UVP2/3/4 data) whereas we only found 108 match-ups with UVP5 data, located on the western and eastern parts of the Kerguelen Plateau, in the Guinea Dome and in the California Current (panel D in @fig-uvp-trap-mapping).
+Consequently, we decided not to go further because the number of match-ups was too low to improve the minimization (see @eq-cost-function).
 
@@
 The idea was based on the assumption that carbon fluxes could be similar in distinct regions of the ocean if those regions had similar environmental conditions thus removing spatial (horizontal) and temporal constraints on match-ups.
-For this, I tried different combinations of environmental variables such as surface temperature, salinity, nutrients, combination of nutrients, Chla, $b_{bp}$ or water column properties such as integrated Chla content, oxygen minimum and MLD from monthly climatologies.
-Then, I used those environmental variables to build a 3D environmental space using a 3D principal component analysis (PCA) and I projected UVP5 data and trap/$^{234}Th$ in that space based on the environmental characteristics from the respective coordinates and month of each UVP5 and trap/$^{234}Th$ profiles.
+For this, we tried different combinations of environmental variables such as surface temperature, salinity, nutrients, combination of nutrients, Chla, $b_{bp}$ or water column properties such as integrated Chla content, oxygen minimum and MLD from monthly climatologies.
+Then, we used those environmental variables to build a 3D environmental space using a 3D principal component analysis (PCA) decomposition and we projected UVP5 data and trap/$^{234}Th$ in that space based on the environmental characteristics from the respective coordinates and month of each UVP5 and trap/$^{234}Th$ profiles.
 
-Once all data were projected in the environmental space, I had to find a new match-up procedure.
-Because the PCA projection conserves the euclidean distance, I firstly computed the pairwise distance between each UVP5 profile and all trap/$^{234}Th$ profiles.
-Then, I looked at the minimal euclidean distance for each UVP5 profile and I plotted the minimal distance distribution (\~8000 points in the distribution).
-From this distribution, I computed a radius of interest (ROI) corresponding to a threshold based on the 95$^{th}$ percentile (considered as a rather objective decision criterion).
-For each UVP5 profile in the environmental space, I kept the closest trap/$^{234}Th$ profiles within that radius (@fig-match-ups).
-Finally, I still needed to find the PSD-POC match-up on the vertical axis (depth).
-For this, I used a depth range proportional to the depth of the trap or $^{234}Th$ measurement following Martin's curve [@Martin1987-um] because the downward carbon flux is less attenuated with depth.
-Then, I averaged PSDs in that depth range (weighted by the water volume) to obtain a table with weighted averaged PSDs in front of POC fluxes to do the minimization procedure of @Guidi2008-vp.
+Once all data were projected in the environmental space, we needed a new match-up procedure.
+Because the PCA projection conserves the euclidean distance, we firstly computed the pairwise distance between each UVP5 profile and all trap/$^{234}Th$ profiles.
+Then, we looked at the minimal euclidean distance for each UVP5 profile and we plotted the minimal distance distribution (\~8000 points in the distribution).
+From this distribution, we computed a radius of interest (ROI) corresponding to a threshold based on the 95$^{th}$ percentile (considered as a rather objective decision criterion).
+For each UVP5 profile in the environmental space, we kept the closest trap/$^{234}Th$ profiles within that radius (@fig-match-ups).
+Finally, we still needed to find the PSD-POC match-ups on the vertical axis (depth).
+For this, we used a depth range proportional to the depth of the trap or $^{234}Th$ measurement following Martin's curve [@Martin1987-um] because the downward carbon flux is less attenuated with depth.
+Then, we averaged PSDs in that depth range (weighted by the water volume for each depth bin) to obtain a table with weighted averaged PSDs in front of POC fluxes to proceed with the minimization procedure of @Guidi2008-vp.
 
-![Examples of match-ups in the environmental space for 4 UVP5 profiles (red dots) and their respective trap/$^{234}Th$ profiles within the ROI.](DATA/CFLUX/moz.jpeg){#fig-match-ups}
+![Examples of trap/$^{234}Th$ match-ups (blue dots) in the geographical space that were within the ROI in the environmental space for 4 UVP5 profiles (red star).](DATA/CFLUX/moz2.jpeg){#fig-match-ups}
 
-dire que le choix des variables ( + le fait que ce sont des climatos -\> bof bof, ça manquait de variavbles physiques, pas assez de variables probablement) et aussi discuter du plan de phase (voir Rmarkdown de la prez waringan + revoir cette prez là pour vérifier que l'on a rien oublié) + critiquer la figure de lionel car pour moi le problème se situe déjà là, j'ai obsrvé que ça fitte tout, on prend un petit b pour u petite A et grand b pour un grand A
+![Residual error $\Delta F$ (for the carbon flux) as a function of $A$ and $D$ (i.e.
+$B = 2D - 1$).
+White crosses correspond to the best ($A$, $D$) values computed during the jackknife error analysis (i.e.
+1000 random subsamples of 30% of the match-ups data set).](DATA/CFLUX/jacknife.jpeg){#fig-solution-space width="550"}
 
-Bottom line, I failed. I tried multiple times but it never worked
+This method increased the number of match-ups to \~7800 however the minimization did not find a clear local minimum in the solution space (@fig-solution-space).
+Instead of having a clear local minimum, @fig-solution-space shows what looks like a 2D solution plane.
+In addition, I think the choice of $A$ and $D$ (or $B$, $B = 2D - 1$ ) is not constrained enough.
+In Figure 5 of @Guidi2008-vp, I think it is difficult to really say that there is a local minimum.
+The light gray color area (low residual error) encompasses a wide region of $A$ and $B$ values where their jackknife procedure found local minima with ($A$, $B$) pairs such as (50, 2) or (150, 5).
+While $A$ is seen as a constant with no physical meaning, the value of $B$ should really be questioned since it is related to the fractal dimension $D$ of particles and aggregates.
+Fractal dimensions should not exceed 3 (physical limit) but they have been estimated down to 1.1-1.3 [@Alldredge1988-lx].
+In addition, it is possible to obtain coherent POC fluxes by tuning the $A$ value for a fixed $B$ therefore I have my doubts about this method because I am not sure $A$ and $B$ are trully independent.
+
+Several other reasons could explain the 'failure' of this method.
+Firstly, the match-up procedure relies on a very strong initial assumption, especially when we deal with PSD measurements (i.e.
+direct measurements) and monthly climatologies.
+Second, the choice of environmental variables we used was limited and most likely insufficient to represent the dynamics of the system at surface responsible for the downward POC flux.
+Thirdly, the UVP5 data set of @Kiko2022-sk is composed of UVP5 SD and HD data hence the choice of size classes is also an issue if the minimization is done on a wide size spectrum due to instrumental differences.
+Fourthly, the model itself as well as the minimization procedure could be revised.
+In their paper, @Bisson2022-bt clearly enonciates the limitations of using the UVP to assess POC fluxes and I agree (see also @sec-BGC-sensors).
+All things considered, I think the search for a global relationship to derive POC fluxes from UVP data might try to generalize a relationship that cannot be generalized.
+Some studies did try to apply the ($A$, $B$) coefficients of @Guidi2008-vp but the results were orders of magnitude different from local ($A$, $B$) estimations [@Iversen2010-an; @Fender2019-ou] hence the question of its global applicability.
 
@@
 
-on a fait avec des cimatos etc mais après y a moyen d'ajouter des meilleires données (eg alberto)
+So far, we have seen that the match-up quest was not enough in the geographical space and not working in the environmental space but what about a method that does not need any match-ups?
+With XGBoost, it is possible to create a regression model to predict carbon fluxes using environmental variables and geographical variables (depth) as predictors (i.e.
+explanatory variables) and trap/$^{234}Th$ POC fluxes as target variables.
+Once the model is trained, it can be applied to the entire data set of PSDs (\> 1.5 million) to derive POC fluxes.
+Our model had a $R^{2}$ of 0.47 with a percentage of explained variance dominated by depth (54%), surface Chla (16%), surface temperature (13%) and the oxygen minimum (10%).
+The minimization of @Guidi2008-vp applied on all true PSD and predicted POC fluxes led to the same kind of solution space as shown in @fig-solution-space.
 
--   faire un résumé de la photo du tableau sur le tableau blanc d Lionel.. pour dire la marche à suivre.. et les ouvelles données environnemntales aussi et le multioutpubut graidnet boosting
+Instead of directly predicting carbon fluxes using environmental variable, we could focus on predicting PSD slope or PSD.
+In a recent paper, @Clements2022-oq used physical and biological variables (see their Table 1) to predict global maps of PSD slope and biovolume at the base of the EZ using UVP5 data and a random forest algorithm.
+@Clements2022-oq highlighted the importance of Chla (i.e.
+proxy of biomass), oxygen (i.e.
+proxy of respiration) and MLD as predictors in their random forest model.
+They also emphasized the importance of including phytoplankton functional groups and abundance as key controlling factors on both the PSD slope and the biovolume in future models due to their impact on the downward POC flux (i.e.
+aggregation/disaggregation processes).
+The study of @Denvil-Sommer2022-oi used both random forest and XGBoost models to predict small and large POC concentrations using geographical variables (depth, latitude and longitude), physical variables (incident light, MLD and temperature), chemical (phosphate and nitrate) and biological variables (Chla and 12 Plankton Functional Types (PFTs) such as diatoms or microzooplankton).
+@Denvil-Sommer2022-oi show that for small POC, the most important predictors are the depth level, temperature, microzooplankton and phosphate while the large POC is mostly driven by depth, temperature, MLD, microzooplankton, phaeocystis (another PFT), phosphate and Chla averaged over the MLD.
+Overall, the results of these studies suggest that it is possible to link the surface environmental and ecosystem structure to the POC distribution in the ocean interior.
+
+### Improving the PSD-POC relationship
+
+I used three methods to derive POC fluxes from UVP5 PSDs.
+The main issues I encountered were related to the number and accuracy of match-ups as well as the choice of environmental variables to link the surface ocean to the POC distribution below.
+With the deployment of new BGC-Argo floats equipped with both the UVP6 and the OST as well as the core biogeochemical variables (e.g.
+Chla, $b_{bp}$, incident light) of the BGC-Argo program, we will hopefully be able to better constraint the PSD-POC relationship however we will probably need a few more years to acquire enough profiles in different oceanic regions to better understand the key drivers of the downward POC flux.
+
+## PARTUM LOUIS LEGENDRE

### 2023-03-22 article3.qmd
update

@@
 I already had some experience with BGC-Argo floats during my master's degree (I published a paper on the study of the deep chlorophyll maximum in the Black Sea, [DOI: 10.5194/bg-18-755-2021](https://bg.copernicus.org/articles/18/755/2021/)) and so while I was waiting for new profiles, I thought it would be useful to have a visualization tool in the form of a web app to investigate incoming data and already prepare the ground for the following study.
-Here is the link to the [web app](https://fricour.shinyapps.io/CARBON_REFINE/), which is still currently being automatically updated every day with new data (almost) live from the Labrador Sea and hopefully with other BGC-Argo floats equipped with the UVP6 and the OST to investigate carbon and particle fluxes in other regions.
-This study uses preliminary data (not even one year of data) and a new BGC-Argo float configuration hence more work should be addressed to better understand the dynamics of the system.
+Here is the link to the [web app](https://fricour.shinyapps.io/CARBON_REFINE/), which is still currently being automatically updated every day with new data (almost) live from the Labrador Sea and hopefully soon with other BGC-Argo floats equipped with the UVP6 and the OST to investigate carbon and particle fluxes in other regions.
+This study uses preliminary data (not even one year of data) and a new BGC-Argo float configuration, hence more work should be addressed to better understand the dynamics of the system.
 
@@
 The biological carbon pump (BCP) refers to a series of processes that transport biogenic carbon produced in sunlit surface waters to the ocean interior [@Claustre2021-tx].
-It is composed of three main pathways [@Claustre2021-tx; @Boyd2019-ez; @Siegel2022-wd]: the biological gravitational pump that is the gravitational settling of particulate organic carbon (POC); the physically mediated particle injection pump that represents physical mechanisms that transport dissolved organic carbon (DOC) to depth via mixing processes; and the biologically mediated particle injection pump that characterizes the transport of POC by diel and seasonal migration of animals in the water column.
+It is composed of three main pathways [@Claustre2021-tx; @Boyd2019-ez; @Siegel2022-wd]: the biological gravitational pump that is the gravitational settling of particulate organic carbon (POC); the physically mediated particle injection pump that represents physical mechanisms that transport both POC and dissolved organic carbon (DOC) to depth via mixing processes; and the biologically mediated particle injection pump that characterizes the transport of POC by diel and seasonal migration of animals in the water column.
 Historically, the BCP was mainly studied with sediment traps [@Honjo2008-kn] and Thorium-234 radionuclide disequilibria [@Buesseler2007-ql] focused on the gravitational pump hence the contribution of DOC and migrating animals to the BCP was poorly known.
@@
 The UVP6 is a portable version of the UVP5 [@Picheral2010-to], specifically designed for AUVs [@Picheral2021-sx].
-It consists of an underwater camera that counts particles and images objects in, respectively, the 0.1 - 2.5 cm and 0.8 - 2.5 cm size range.
+It consists of an underwater camera that counts particles and images objects in, respectively, the 0.1 - 2.5 mm and 0.6 - 2.5 cm size range.
 In contrast with UVP5 deployments on cabled platforms, UVP6 deployed on long-term drifting AUVs (e.g.
@@
 These hypotheses have been investigated with BGC-Argo floats but the debate still remains [@Mignot2018-gm; @Yang2020-hw; @Boss2010-do].
-For the LS, @Lacour2015-da used a climatology of satellite Chla to divide the NA into 6 bioregions with specific bloom phenologies.
+Distinct NA bloom phenologies were highlighted by @Lacour2015-da who used a climatology of satellite Chla to divide the NA into 6 bioregions.
 In their bioregionalization (see also their Figure 2), the LS is characterized by two bioregions with distinct bloom phenologies on both side of the 60°N parallel.
-The northern part is characterized by a short spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June.
+The northern part is characterized by a short and potentially more intense spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June.
 @Lacour2015-da concluded that the concurrent influence of surface light and mixing was the first-order mechanism controlling the onset of the spring bloom in both bioregions.
-This is coherent with the results of @Wu2008-vw and @Frajka-Williams2010-ph who found that early shoaling of the mixed layer was responsible for the early spring bloom in the northern part whereas the light controlled the bloom in the southern part.
+This is coherent with the results of @Wu2008-vw and @Frajka-Williams2010-ph who found that the early bloom in the northern part is due to a very shallow mixed layer associated with low-salinity waters (i.e.
+haline stratification) while the southern bloom is related to the development of the seasonal thermocline (i.e.
+thermal stratification due to increased solar radiation).
 In late May 2022, two pairs of BGC-Argo floats were deployed in the southern part of the LS at the time of the spring bloom.
-While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study POC fluxes and particle concentrations derived from beam attenuance [@Estapa2023-oz] and UVP6 particle size distributions (PSD).
+While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study the downward POC flux and its attenuation derived from beam attenuance [@Estapa2023-oz] and to study particle size distributions (PSD) and concentrations with the UVP6.
 
@@
 Two pairs of BGC-Argo floats were deployed in the Labrador Sea in late May 2022.
-The first pair (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second (WMO 6904241 and 1902578) was deployed on the 29$^{th}$ at coordinates (-52.4°E, 56.8°N).
+The first pair (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second pair (WMO 6904241 and 1902578) was deployed on the 29$^{th}$ at coordinates (-52.4°E, 56.8°N).
 
 All 4 floats used in this study are PROVOR Jumbo floats (NKE Instrumentation, France) equipped with a CTD SBE41-CP (Sea-Bird Scientific, USA), a UVP6-LP (Hydroptic, France), a c-Rover 660 nm beam transmissometer (WETLabs, USA) and a ECO FLBBCD (WETLabs, USA) that includes a Chla fluorometer (excitation at 470 nm, emission at 695 nm), a colored dissolved organic matter (CDOM) fluorometer (excitation at 370 nm, emission at 460 nm) and a 700 nm particle backscatteringmeter ($b_{bp}$).
-All floats but one (1902578) were equipped with a OCR-504 Multispectral Radiometer to measure the photosynthetic available radiation.
-
-All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux.
+<!-- All floats are equipped with a OCR-504 Multispectral Radiometer (Sea-Bird Scientific, USA) except float 1902578 that is equipped with a RAMSES hyperspectral radiometer (TriOS, Germany) to measure the downwelling irradiance at three wavelengths (380, 412, 490 nm) and of the light available for photosynthesis (PAR).
+-->
+
+All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux and its attenuation.
 During these drift phases, measurements of particle abundance (14 size classes, between 0.1 mm and 2.5 mm) are taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths.
 This configuration can be adjusted during the mission of the float via two-way Iridium communication.
+Finally, the vertical orientation of the transmissometer causes the accumulation of sinking material on the optical window.
+Therefore, a pump is cleaning the latter to remove the accumulated material between each drifting phase and subsequently avoid sensor drift.
 
@@
 #| label: fig-float-position
-#| fig-cap: Route of each float since their deployment (black triangles).
+#| fig-cap: Trajectory of each float since their deployment (black triangles).
 Dots represent ascending profiles.
@@
 Chla data were quality controlled for Non-Photochemical Quenching (NPQ), a photoprotective mechanism that decreases the apparent fluorescence per unit of Chla following the method implemented by @Terrats2020-ef who modified the method of @Xing2018-pn for shallow-mixing cases.
-Chla data were also divided by a slope factor of 2.15 following the recommendation of @Roesler2017-me for obtaining unbiased Chla estimates from in situ WETLabs ECO fluorometers.
+Chla data were also divided by a slope factor of 2 following the recommendation of @Roesler2017-me for obtaining unbiased Chla estimates from in situ WETLabs ECO fluorometers.
 Temperature and salinity data were quality controlled using @Wong2022-hz.
@@
 The latter were obtained from the calibration sheets of each c-Rover.
+In practice, this calibration is not necessary to use the transmissometer in OST mode for the estimation of small and large sinking particles (see next section).
 
@@
 
-Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk], we chose an empirical statistical approach instead of thresholds to detect the deposition of large particles (hereafter referred to as jumps) in the $c_{p}$ signal.
-Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase.
-The detection of jumps consisted in spotting outliers in the slope distribution.
-For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles.
-Positive and negative jumps were defined as, respectively, a slope above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$.
-We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.
-
-In order to compute $F_{small}$, we firstly computed the slope of each segment between jumps (positive and/or negative, see @fig-estapa-method_article).
-Any segment with less than 4 observations or a negative slope was removed assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by factors other than the pump cleaning the optical window (e.g.
-turbulence ou zooplankton consumption of accumulated material).
-Afterwards, we computed the weighted average slope ($slope_{w}$) of each segment to derive the small particles flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).
-
-$$
-F_{small} = 633(0.25~slope_{w})^{0.77}
-$$ {#eq-estapa}
-
-where $slope_{w}$ is multiplied by the transmissometer pathlength because @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.
-
-We used the same relationship to compute $F_{large}$ by replacing $slope_{w}$ as the sum of positive jumps divided by the drifting time.
-
 ```{r, example of a cp signal analyzed for jumps (with an old method here), fig.width=8, fig.height=5}
@@
 A negative jump was also detected (yellow).
-Data from BGC-ARGO float WMO 6904240, drifting at 1000 m during its 21$^{st}$ cycle.
+Data from BGC-Argo float WMO 6904240, drifting at 1000 m during its 21$^{st}$ cycle.
 
@@
 
+Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk], we chose an empirical statistical approach instead of thresholds to detect the deposition of large particles (hereafter referred to as jumps) in the $c_{p}$ signal.
+Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase.
+The detection of jumps consisted in spotting outliers in the slope distribution.
+For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles.
+Positive and negative jumps were defined as, respectively, a slope above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$.
+We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.
+
+In order to compute $F_{small}$, we firstly computed the slope of each segment between jumps (positive and/or negative, see @fig-estapa-method_article).
+Any segment with less than 4 observations or a negative slope was removed assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by factors other than the pump cleaning the optical window (e.g.
+turbulence or zooplankton consumption of accumulated material).
+Afterwards, we computed the weighted average slope ($slope_{w}$) of each segment to derive the small particles flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).
+
+$$
+F_{small} = 633(0.25~slope_{w})^{0.77}
+$$ {#eq-estapa}
+
+where $slope_{w}$ is multiplied by the transmissometer pathlength because @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.
+
+We used the same relationship to compute $F_{large}$ by replacing $slope_{w}$ as the sum of positive jumps divided by the drifting time.
+
 ### Satellite data
@@
 
+```{=latex}
+\newpage
+```
 ## Results
 
+### UVP6 particle abundance
+
+#### On profiling mode
+
+Vertical profiles of particle concentrations in the lower (102 - 512 $\mu$m, see @fig-uvp6-profiling-102-406) and upper (0.512 - 2.5 mm, see @fig-uvp6-profiling-512-2500) size range of the UVP6 show similar patterns across all floats.
+In the lower part of the size range, @fig-uvp6-profiling-102-406 shows the rapid emptying of the water column in November.
+This process occurred on the entire water column (e.g.
+size class 406, float 4903634) but also started with more intensity from below the surface (e.g.
+size class 102, float 6904240).
+All floats also measured a local particle abundance maximum around 800 - 1000 m whose origin is analyzed in the Discussion.
+In the upper part of the size range, @fig-uvp6-profiling-512-2500 shows a rapid emptying of the entire water column starting in November.
+In contrast with small classes (referred to here as below 512 $\mu$m), some particles, especially above 1.5 mm, have almost entirely disappeared from the water column by September.
+However, the rapid increase and decrease of particles on the entire water column around July-August for floats 6904241 and 1902578 (still in the same zone at that time) could suggest an export event.
+
+```{r, UVP6 profiling mode without sinking speed small classes, fig.height=8, fig.width=8}
+#| label: fig-uvp6-profiling-102-406
+#| fig-cap: Vertical profiles of UVP6 particle concentrations for size classes in the 102 - 512 $\mu$m size range.
+Concentrations were normalized for each size class.
+The colorbar is in log$_{10}$ scale.
+
+particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld))
+
+particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+
+# remove buggy data from float 4903634 (negative depth for some reason)
+particles_bis <- particles_bis |> filter(depth >= 0)
+
+# rel_conc
+particles_bis <- particles_bis |> dplyr::group_by(wmo, size) |> dplyr::mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
+
+up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
+  #group_by(size) |>
+  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+  scale_y_reverse() +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
+  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
+
+up_part
+```
+
+```{r, UVP6 profiling mode without sinking speed large classes, fig.height=8, fig.width=8}
+#| label: fig-uvp6-profiling-512-2500
+#| fig-cap: Vertical profiles of UVP6 particle concentrations for size classes in the 0.512 - 2.5 mm size range.
+Concentrations were normalized for each size class.
+The colorbar is in log$_{10}$ scale.
+The gray background depicts the absence of particle in the specified size class.
+
+down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
+  #group_by(size) |>
+  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
+  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
+  scale_y_reverse() +
+  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
+  #geom_vline(xintercept = as_date('2022-01-01')) +
+  scale_x_date(labels = date_format("%m"), date_breaks = '3 month')
+
+down_part
+```
+
+```{=latex}
+\newpage
+```
+#### On drifting mode
+
+```{r, uvp data}
+
+lpm_class <- c("NP_Size_102", "NP_Size_128", "NP_Size_161",
+                         "NP_Size_203", "NP_Size_256", "NP_Size_323",
+                         "NP_Size_406", "NP_Size_512", "NP_Size_645",
+                         "NP_Size_813", "NP_Size_1020", "NP_Size_1290",
+                        "NP_Size_1630", "NP_Size_2050")
+
+uvp_data <- all_floats |> filter(PhaseName == 'PAR') |> dplyr::mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date, cycle, wmo, park_depth, all_of(lpm_class)) |>
+  drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld), WMO = factor(wmo))
+
+uvp_data <- uvp_data |> pivot_longer(cols = lpm_class, names_to = 'size', values_to = 'conc') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
+
+# compute daily mean chla for each float at each cycle and at each drifting depth
+mean_uvp_data <- uvp_data |> dplyr::group_by(cycle, WMO, size, park_depth, juld) |> dplyr::summarize(mean_conc = mean(conc, na.rm=T))
+```
+
+In this study, we focus on 14 size classes, from 102 $\mu$m to 2.5 mm.
+In @fig-uvp-200m, @fig-uvp-500m and @fig-uvp-1000m, we show the daily averaged concentrations of particles in each size class for each float at their respective drifting depth.
+At 200 m, it can be observed that particle concentrations in the 102 - 512 $\mu$m range are relatively uniform between June and October before decreasing sharply towards January.
+Between 512 $\mu$m and 1.63 mm, concentrations are steadily decreasing from June.
+Above 1.63 mm, the particle abundance is already very low (\< 0.1 particle/L) in June but keeps decreasing until September when they are practically absent (\< 0.01 particle/L).
+At 500 m, particle concentrations in the 102 - 645 $\mu$m range slowly decrease from July to September with a sharp decrease starting in October.
+Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (\< 0.5 particle/L) on the entire time series.
+At 1000 m, all 4 floats measured a clear maximum in particle abundance in the 102 - 645 $\mu$m in the second half of July.
+Above 645 $\mu$m, the same trend can still be observed with concentrations of less than 1 particle/L.
+
+```{r, uvp data at 200 m, fig.height=8, fig.width=8}
+#| label: fig-uvp-200m
+#| fig-cap: Daily averaged concentrations of particle abundance in 14 size classes measured by the UVP6 for each float at 200 m.
+Black lines and shaded areas represent, respectively, the daily mean particle abundance and the 95% confidence interval around the mean.
+Dashed vertical lines indicate the time at which the daily mean particle abundance is maximum.
+
+# add blank
+blank_data <- tibble(cycle = c(NA, NA), WMO = rep(as.factor(4903634),2), size = c(102,102), park_depth = c('200 m', '500 m'),
+                     juld = rep(as.Date('2023-03-05'),2), mean_conc = c(NA,NA))
+
+mean_uvp_data <- rbind(mean_uvp_data, blank_data)
+
+p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') #+
+  #geom_vline(data = test, aes(xintercept = date_max))
+
+tmp200m <- ggplot_build(p200m)
+
+test <- tmp200m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> dplyr::summarize(index_max = which.max(y), max_date = x[index_max]) |> dplyr::mutate(date_max = as_date(max_date, origin = "1970-01-01"))
+test$size <- unique(uvp_data$size)
+
+p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') +
+  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
+
+p200m
+```
+
+```{r, uvp data at 500 m, fig.width=8, fig.height=8}
+#| label: fig-uvp-500m
+#| fig-cap: Daily averaged concentrations of particle abundance in 14 size classes measured by the UVP6 for each float at 500 m.
+Black lines and shaded areas represent, respectively, the daily mean particle abundance and the 95% confidence interval around the mean.
+Dashed vertical lines indicate the time at which the daily mean particle abundance is maximum.
+
+p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') #+
+  #geom_vline(data = test, aes(xintercept = date_max))
+
+tmp500m <- ggplot_build(p500m)
+
+test <- tmp500m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> dplyr::summarize(index_max = which.max(y), max_date = x[index_max]) |> dplyr::mutate(date_max = as_date(max_date, origin = "1970-01-01"))
+test$size <- unique(uvp_data$size)
+
+p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') +
+  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
+
+p500m
+```
+
+```{r, uvp data at 1000 m, fig.width=8, fig.height=8}
+#| label: fig-uvp-1000m
+#| fig-cap: Daily averaged concentrations of particle abundance in 14 size classes measured by the UVP6 for each float at 1000 m.
+Black lines and shaded areas represent, respectively, the daily mean particle abundance and the 95% confidence interval around the mean.
+Dashed vertical lines indicate the time at which the daily mean particle abundance is maximum.
+#| fig-dpi: 96
+
+#{r, uvp data at 1000 m, out.width="900px", out.height="800px"}
+
+p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') #+
+  #geom_vline(data = test, aes(xintercept = date_max))
+
+tmp1000m <- ggplot_build(p1000m)
+
+test <- tmp1000m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> dplyr::summarize(index_max = which.max(y), max_date = x[index_max]) |> dplyr::mutate(date_max = as_date(max_date, origin = "1970-01-01"))
+test$size <- unique(uvp_data$size)
+
+p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
+  geom_smooth(aes(juld, mean_conc), colour = 'black') +
+  scale_color_brewer(palette = 'Dark2') + 
+  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
+  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
+  scale_y_continuous(trans = 'log10') +
+  facet_wrap(~size, scales = 'free_y') +
+  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
+
+p1000m
+```
+
+```{=latex}
+\newpage
+```
 ### OST-derived carbon fluxes
@@
 Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.
-There is no data after January and February at, respectively, 200 m and 500 m because the drifting was stopped at those depths due to the deepening of the MLD.
+There is no data after January and February at, respectively, 200 m and 500 m because the drifting was stopped at those depths due to the deepening of the MLD associated with convection and the impossibility to stabilize the floats.
 
@@
 
+$F_{large}$ was estimated at 200 m between 0 (no deposition of large particles on the transmissometer window) and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$.
+High fluxes (\> 100 mg C m$^{-2}$ day$^{-1}$) were measured by 3 out of 4 floats at the end of May before a sharp decrease towards July.
+At 500 m, $F_{large}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_largef, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to February.
+At 1000 m, $F_{large}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards March, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$.
+
 ```{r, MLD, fig.height=4, fig.width=8}
@@
 
-mld_plot <- ggplot(mld_data) + geom_line(aes(x = date, y = mld, colour = WMO)) +
+mld_plot <- ggplot(mld_data) + geom_line(aes(x = date, y = mld, colour = WMO), size = 1) +
   labs(x = 'Month', y = 'MLD (m)') + scale_y_reverse() +
@@
 
-$F_{large}$ was estimated at 200 m between 0 (no deposition of large particles on the transmissometer window) and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$.
-High fluxes (\> 100 mg C m$^{-2}$ day$^{-1}$) were measured by 3 out of 4 floats at the end of May before a sharp decrease towards July.
-At 500 m, $F_{large}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_largef, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to February.
-At 1000 m, $F_{large}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards March, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$.
-
-### UVP6 particle abundance
-
-#### On drifting mode
-
-```{r, uvp data}
-
-lpm_class <- c("NP_Size_102", "NP_Size_128", "NP_Size_161",
-                         "NP_Size_203", "NP_Size_256", "NP_Size_323",
-                         "NP_Size_406", "NP_Size_512", "NP_Size_645",
-                         "NP_Size_813", "NP_Size_1020", "NP_Size_1290",
-                        "NP_Size_1630", "NP_Size_2050")
-
-uvp_data <- all_floats |> filter(PhaseName == 'PAR') |> dplyr::mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date, cycle, wmo, park_depth, all_of(lpm_class)) |>
-  drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld), WMO = factor(wmo))
-
-uvp_data <- uvp_data |> pivot_longer(cols = lpm_class, names_to = 'size', values_to = 'conc') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
-
-# compute daily mean chla for each float at each cycle and at each drifting depth
-mean_uvp_data <- uvp_data |> dplyr::group_by(cycle, WMO, size, park_depth, juld) |> dplyr::summarize(mean_conc = mean(conc, na.rm=T))
-```
-
-In this study, we focus on 14 size classes, from 102 $\mu$m to 2.5 mm.
-In @fig-uvp-200m, @fig-uvp-500m and @fig-uvp-1000m, we show the daily averaged concentrations of particles in each size class for each float at their respective drifting depth.
-At 200 m, it can be observed that particle concentrations in the 102 - 512 $\mu$m range are relatively uniform between June and October before decreasing sharply towards January.
-Between 512 $\mu$m and 1.63 mm, concentrations are steadily decreasing from June.
-Above 1.63 mm, the particle abundance is already very low (\< 0.1 particle/L) in June but keeps decreasing until September when they are practically absent (\< 0.01 particle/L).
-
-```{r, uvp data at 200 m, fig.height=8, fig.width=8}
-#| label: fig-uvp-200m
-#| fig-cap: Daily averaged concentrations of particle abundance in 14 size classes measured by the UVP6 for each float at 200 m.
-Black lines and shaded areas represent, respectively, the daily mean particle abundance and the 95% confidence interval around the mean.
-Dashed vertical lines indicate the time at which the daily mean particle abundance is maximum.
-
-p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
-  geom_smooth(aes(juld, mean_conc), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
-  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
-  scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size, scales = 'free_y') #+
-  #geom_vline(data = test, aes(xintercept = date_max))
-
-tmp200m <- ggplot_build(p200m)
-
-test <- tmp200m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> dplyr::summarize(index_max = which.max(y), max_date = x[index_max]) |> dplyr::mutate(date_max = as_date(max_date, origin = "1970-01-01"))
-test$size <- unique(uvp_data$size)
-
-p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
-  geom_smooth(aes(juld, mean_conc), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
-  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
-  scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size, scales = 'free_y') +
-  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
-
-p200m
-```
-
-At 500 m, particle concentrations in the 102 - 645 $\mu$m range slowly decrease from July to September with a sharp decrease starting in October.
-Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (\< 0.5 particle/L) on the entire time series.
-
-```{r, uvp data at 500 m, fig.width=8, fig.height=8}
-#| label: fig-uvp-500m
-#| fig-cap: Daily averaged concentrations of particle abundance in 14 size classes measured by the UVP6 for each float at 500 m.
-Black lines and shaded areas represent, respectively, the daily mean particle abundance and the 95% confidence interval around the mean.
-Dashed vertical lines indicate the time at which the daily mean particle abundance is maximum.
-
-p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
-  geom_smooth(aes(juld, mean_conc), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
-  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
-  scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size, scales = 'free_y') #+
-  #geom_vline(data = test, aes(xintercept = date_max))
-
-tmp500m <- ggplot_build(p500m)
-
-test <- tmp500m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> dplyr::summarize(index_max = which.max(y), max_date = x[index_max]) |> dplyr::mutate(date_max = as_date(max_date, origin = "1970-01-01"))
-test$size <- unique(uvp_data$size)
-
-p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
-  geom_smooth(aes(juld, mean_conc), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
-  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
-  scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size, scales = 'free_y') +
-  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
-
-p500m
-```
-
-At 1000 m, all 4 floats measured a clear maximum in particle abundance in the 102 - 645 $\mu$m in the second half of July.
-Above 645 $\mu$m, the same trend can still be observed with concentrations of less than 1 particle/L.
-
-```{r, uvp data at 1000 m, fig.width=8, fig.height=8}
-#| label: fig-uvp-1000m
-#| fig-cap: Daily averaged concentrations of particle abundance in 14 size classes measured by the UVP6 for each float at 1000 m.
-Black lines and shaded areas represent, respectively, the daily mean particle abundance and the 95% confidence interval around the mean.
-Dashed vertical lines indicate the time at which the daily mean particle abundance is maximum.
-#| fig-dpi: 96
-
-#{r, uvp data at 1000 m, out.width="900px", out.height="800px"}
-
-p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
-  geom_smooth(aes(juld, mean_conc), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
-  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
-  scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size, scales = 'free_y') #+
-  #geom_vline(data = test, aes(xintercept = date_max))
-
-tmp1000m <- ggplot_build(p1000m)
-
-test <- tmp1000m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> dplyr::summarize(index_max = which.max(y), max_date = x[index_max]) |> dplyr::mutate(date_max = as_date(max_date, origin = "1970-01-01"))
-test$size <- unique(uvp_data$size)
-
-p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
-  geom_smooth(aes(juld, mean_conc), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
-  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
-  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
-  scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size, scales = 'free_y') +
-  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
-
-p1000m
-```
-
-#### On profiling mode
-
-Vertical profiles of particle concentrations in the lower (102 - 512 $\mu$m) and upper (0.512 - 2.5 mm) size range of the UVP6 show similar patterns across all floats.
-In the lower part of the size range, @fig-uvp6-profiling-102-406 shows the rapid emptying of the water column in November.
-This process occurred on the entire water column (e.g.
-size class 406, float 4903634) but also started with more intensity from below the surface (e.g.
-size class 102, float 6904240).
-All floats also measured a local particle abundance maximum around 800 - 1000 m whose origin is analyzed in the Discussion.
-
-```{r, UVP6 profiling mode without sinking speed small classes, fig.height=8, fig.width=8}
-#| label: fig-uvp6-profiling-102-406
-#| fig-cap: Vertical profiles of UVP6 particle concentrations for size classes in the 102 - 512 $\mu$m size range.
-Concentrations were normalized for each size class.
-The colorbar is in log$_{10}$ scale.
-
-particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld))
-
-particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> dplyr::mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
-
-# remove buggy data from float 4903634 (negative depth for some reason)
-particles_bis <- particles_bis |> filter(depth >= 0)
-
-# rel_conc
-particles_bis <- particles_bis |> dplyr::group_by(wmo, size) |> dplyr::mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))
-
-up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
-  #group_by(size) |>
-  ggplot() + facet_grid(wmo~size) + theme_bw() +
-  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
-  scale_y_reverse() +
-  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
-  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
-
-up_part
-```
-
-In the upper part of the size range, @fig-uvp6-profiling-512-2500 shows a rapid emptying of the entire water column starting in November.
-In contrast with small classes (referred to here as below 512 $\mu$m), some particles, especially above 1.5 mm, have almost entirely disappeared from the water column by September.
-However, the rapid increase and decrease of particles on the entire water column around July-August for floats 6904241 and 1902578 (still in the same zone at that time) could suggest an export event.
-
-```{r, UVP6 profiling mode without sinking speed large classes, fig.height=8, fig.width=8}
-#| label: fig-uvp6-profiling-512-2500
-#| fig-cap: Vertical profiles of UVP6 particle concentrations for size classes in the 0.512 - 2.5 mm size range.
-Concentrations were normalized for each size class.
-The colorbar is in log$_{10}$ scale.
-The gray background depicts the absence of particle in the specified size class.
-
-down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
-  #group_by(size) |>
-  ggplot() + facet_grid(wmo~size) + theme_bw() +
-  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
-  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
-  scale_y_reverse() +
-  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
-  #geom_vline(xintercept = as_date('2022-01-01')) +
-  scale_x_date(labels = date_format("%m"), date_breaks = '3 month')
-
-down_part
-```
-
+```{=latex}
+\newpage
+```
 ## Discussion
@@
 upper 20 m) Chla concentrations from both satellite and in situ data in a 100 km radius around each deployment.
-@fig-chla-mld-satellite shows that satellite and in situ Chla and MLD data concur therefore we assume that the former can be used to study environmental conditions before the floats deployment.
-Taking this into consideration, @fig-chla-mld-satellite indicates that the 6904240-4903634 pair missed a bloom that occurred in early May whereas the 6904241-1902578 pair missed a first (satellite) Chla increase at the end of May days before their deployment.
+@fig-chla-mld-satellite shows that satellite and in situ Chla and MLD data concur.
+Therefore we assume that the former can be used to study environmental conditions before the floats deployment.
+Taking this into consideration, @fig-chla-mld-satellite indicates that the 6904240-4903634 pair likely missed a bloom that occurred in early May whereas the 6904241-1902578 pair missed a first (satellite) Chla increase at the end of May days before their deployment.
 However, a second (satellite and in situ) Chla increase reached a peak at the end of June before it decreased again towards July.
 
-```{r, chla and MLD from satellite, fig.height=6, fig.width=6}
+The Chla satellite pattern for the 6904240-4903634 pair typically describes the dynamics of the spring bloom in the northern part of the LS [@Lacour2015-da].
+This is not surprising given that this pair was deployed one degree southward to the so-called 60°N limit [@Lacour2015-da].
+@fig-chla-mld-satellite shows that the satellite MLD in that region had already shoaled below 100 m since March and that the water column was stable and relatively well stratified in May which might explain the intensity of surface Chla concentrations in contrast with those measured at the second deployment site.
+The latter is located in the bioregion where the phytoplankton biomass typically develops slowly in May to reach its maximum in June [@Lacour2015-da].
+However, the Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by a mixing event at the end of May and that it started again in mid-June when the water column was well stratified.
+
+```{r, chla and MLD from satellite, fig.height=4, fig.width=7}
 #| label: fig-chla-mld-satellite
@@
 Satellite error bars represent the first and third quartiles around the median in a 100 km radius around each deployment coordinates.
-In situ Chla represents the Chla median in the upper 20 m of the water column with uncertainties from the slope factor uncertainty of @Roesler2017-me.
+In situ Chla represents the Chla median in the upper 20 m of the water column.
 
@@
 d <- d |> group_by(date, wmo, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
-d <- d |> dplyr::mutate(smoothed_chla = smooth(mean_chla)/2.15, smoothed_chla_up = smooth(mean_chla)/(2.15-0.37), smoothed_chla_down = smooth(mean_chla)/(2.15+0.37)) # 2.15 = Roesler factor in the Labrador Sea
+d <- d |> dplyr::mutate(smoothed_chla = smooth(mean_chla)/2)
 
@@
                                        chla_0_mld = integrate(smoothed_chla, pres, from = 0, to=mld),
-                                       median_chla_0_20m = median(smoothed_chla[pres<=20], na.rm=T),
-                                       median_chla_0_20m_up = median(smoothed_chla_up[pres<=20], na.rm=T),
-                                       median_chla_0_20m_down = median(smoothed_chla_down[pres<=20], na.rm=T))
+                                       median_chla_0_20m = median(smoothed_chla[pres<=20], na.rm=T))
 #
@@
   geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
-  geom_pointrange(data = d_chla |> filter(date < '2022-08-01'), aes(x = date, ymin = median_chla_0_20m_down, ymax = median_chla_0_20m_up, y = median_chla_0_20m), colour = "#E31B23", size = .25) +
+  geom_point(data = d_chla |> filter(date < '2022-08-01'), aes(x = date, y = median_chla_0_20m), colour = "#E31B23", size = 1) +
    scale_x_date(labels = date_format("%m"), date_breaks = '1 month')
@@
   geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
-  geom_point(data = d_chla |> filter(date < '2022-08-01') |> dplyr::mutate(week = week(date), DOY = yday(date)) |> group_by(week, group) |> summarize(median_mld = median(mld), mean_date = as_date(mean(DOY), origin = '2022-01-01')), aes(mean_date, median_mld), colour = "#E31B23") +
+  geom_point(data = d_chla |> filter(date < '2022-08-01') |> dplyr::mutate(week = week(date), DOY = yday(date)) |> group_by(week, group) |> summarize(median_mld = median(mld), mean_date = as_date(mean(DOY), origin = '2022-01-01')), aes(mean_date, median_mld), colour = "#E31B23", size = 1) +
    scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + scale_y_continuous(trans = 'log10')
@@
 
-The Chla satellite pattern for the 6904240-4903634 pair typically describes the dynamics of the spring bloom in the northern part of the LS [@Lacour2015-da].
-This is not surprising given that this pair was deployed one degree southward to the so-called 60°N limit [@Lacour2015-da].
-@fig-chla-mld-satellite shows that the satellite MLD in that region had already shoaled below 100 m since March and that the water column was stable and relatively well stratified in May which might explain the intensity of surface Chla concentrations in contrast with those measured at the second deployment site.
-The latter is located in the bioregion where the phytoplankton biomass typically develops slowly in May to reach its maximum in June [@Lacour2015-da].
-However, the Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD at the end of May and that it started again in mid-June when the water column was well stratified.
-<!-- However, the increase of the MLD in June could have also impacted the intensity of the bloom which might explain the three times lower than Chla concentrations.
-However, the Chla satellite concentration (before deployment) for that pair was three times lower than for the other pair and the second Chla increase (satellite and in situ) was of the same order of magnitude.
-Therefore, the second pair was deployed in a less productive area and that the inten <!--Hence, it is possible that the bloom intensity was not sufficient to produce clear particle export signals, at least for BGC-Argo floats in their 10-day vertical sampling configuration.
--->
-
 ### Phytoplankton communities at deployment sites
 
-Here, we refer to the first and second deployment sites as, respectively, station A and station B.
 High Performance Liquid Chromatography (HPLC) analyses were performed on samples taken during the two deployments.
 The selected pigments, their taxonomic significance and their size class were taken from Table 2 in @Petit2022-de.
-Station A showed Chla concentrations (TChla pigment in @fig-hplc) two times higher than in station B in the first 100 m as well as higher concentrations of diatoms (Fucoxanthin).
-Station B showed higher concentrations of cryptophytes (Alloxanthin), dinoflagellates (Peridin) and green flagellates and prochlorophytes (TChlb) than in station A.
+The pair 6904240-4903634 showed Chla concentrations (TChla pigment in @fig-hplc) two times higher than the pair 6904241-1902578 in the first 100 m as well as higher concentrations of diatoms (Fucoxanthin).
+The pair 6904241-1902578 showed higher concentrations of cryptophytes (Alloxanthin), dinoflagellates (Peridin) and green flagellates and prochlorophytes (TChlb).
 However, their absolute concentrations are quite low (usually less than 0.05 mg m$^{-3}$).
-Taking that into consideration, it seems that station A was more productive than station B and was dominated by diatoms (i.e.
-microphytoplankton) whereas station B was composed of a majority of smaller phytoplankton species (mostly nano and picozooplankton).
+Taking that into consideration, it seems that, at the time of float deployment, the northern site was located in a more productive area than the southern site and that it was dominated by diatoms (i.e.
+microphytoplankton) whereas the latter was composed of a majority of smaller phytoplankton species (mostly nano and picophytoplankton).
 This supports the idea of two distinct biological dynamics as inferred from float and satellite data.
@@
 #| label: fig-hplc
-#| fig-cap: Pigments derived from HPLC analyses at stations A and B.
+#| fig-cap: Pigments derived from HPLC analyses in the northern (floats 6904240 and 4903634) and southern (floats 6904241 and 1902578) deployment sites.
 
@@
                        `19'-Butanoyloxyfucoxanthin`, Alloxanthin,
-                       Fucoxanthin, Peridinin, TChla = Total_Chlorophyll_a) |> mutate(station = c(rep('station A', 10), rep('station B', 9)))
+                       Fucoxanthin, Peridinin, TChla = Total_Chlorophyll_a) |> mutate(station = c(rep('6904240 - 4903634', 10), rep('6904241 - 1902578', 9)))
 
@@
 
+```{=latex}
+\newpage
+```
 ```{r, spectral slope}
@@
 The spectral slope defines the slope of the PSD where a high (in absolute values) spectral slope indicates that small particules dominate the PSD.
-Here, the panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope at the beginning of the time series but it is also difficult to relate it to biological processes such as zooplankton feeding without additional data.
+Here, the panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope at the beginning of the time series but it is also difficult to relate it to biological processes such as zooplankton feeding without any additional data.
 
@@
 
+```{=latex}
+\newpage
+```
 ### Comparing OST and UVP data
@@
 Nonetheless, carbon fluxes measured at 1000 m by the OST and the UVP6 with the coefficients of @Iversen2010-an are both well correlated (\> 0.6 for all floats) and of similar magnitudes while carbon fluxes using @Guidi2008-vp are one order of magnitude lower.
-In their paper, @Iversen2010-an proposed that the carbon fluxes discrepancy could be due to different biogeochemical regimes where the relationship of @Guidi2008-vp was based on sediment traps at open ocean sites with lower carbon fluxes [@Honjo2008-kn] while the study site of @Iversen2010-an was located in a highly productive system at the continental margin off Cape Blanc in Mauritania.
+In their paper, @Iversen2010-an proposed that the carbon fluxes discrepancy could be due to different biogeochemical regimes where the relationship of @Guidi2008-vp was rather based on sediment traps at open ocean sites with lower carbon fluxes [@Honjo2008-kn] while the study site of @Iversen2010-an was located in a highly productive system at the continental margin off Cape Blanc in Mauritania.
 The LS is productive during the spring bloom [@Lacour2015-da] therefore it is coherent to find a closer relationship with estimations based on a highly productive system.
@@
 It is coherent with the fact that OSTs were not designed to intercept and detect large sinking particles [@Estapa2023-oz].
-The current version of the OST has a housing above the optical window which can produce a "self-shading" effect [@Estapa2017-jr] that prevents the collection of fast sinking particles (assumed here to be large particles).
+The current version of the OST has a housing above the optical window which, together with the float itself, can produce a "self-shading" effect [@Estapa2017-jr] that prevents the collection of fast sinking particles (assumed here to be large particles).
 In addition, the nature of particles is also a source of uncertainty in the attenuation signal.
@@
 Therefore, if the UVP6 and the OST are well correlated at first order (@fig-comparison-ost-uvp, @tbl-correlation), it is more difficult to directly link the attenuance to the biovolume (hence the size) of a particle because of the nature of particles themselves.
-Consequently, the combined use of the UVP6 and the OST seems useful to study the latent particle dynamics but the interest of using the OST for large particles (e.g.
-during a bloom) might be more limited.
+Overall, further exploration analyses are needed to better understand the particle dynamics and their composition, especially considering a broader range of trophic situations.
 
@@
 Similarly, to address the question of the vertical migration, a 10-day temporal resolution is way to low because plankton images are rare compared to detritus so the sampling volume should be greatly increased to have a better estimation of the presence (or absence) of zooplankton.
-Finally, I have not made up my mind about the choice of the three drifting depths but maybe it would be interesting to increase the time spent at 200 m (to 3 days) and decrease the time spent at 1000 m to 3 days instead of 5) where the probability of capturing big particles is lower than at 200 m.
+Finally, it would perhaps be interesting to increase the time spent at 200 m (to 3 days) and decrease the time spent at 1000 m to 3 days instead of 5) where the probability of capturing big particles is lower than at 200 m.
 There could also be UVP6 and OST measurements closer in time at all drifting depths to have a better comparison of attenuation slopes/jumps and biovolumes.

### 2023-03-22 discussion.qmd
update

@@
 
-## PARTUM LOUIS LEGENDRE
+## Continuous vertical sequestration
+
+Using the CONVERSE approach, we have estimated that the global carbon sequestration (\> 100 years) flux by the BCP is 0.9-2.6 Pg C y$^{-1}$.
+This is 2-3 and 3-6 times larger than POC fluxes at, respectively, 1000 and 2000 m.
+CONVERSE sequestration flux estimates are also up to 4 times larger than those from studies assuming that sequestration occurs below a fixed reference depth [@Honjo2008-kn, @Guidi2015-sj; @Henson2012-ck].
+Here below, I highlight some key points and ways to improve further our estimations.
+
+### Marine carbon dioxide removal strategies
+
+The goal of marine carbon dioxide removal (mCDR) strategies [@Boyd2023-xf] is the removal of excess CO$_{2}$ from the atmosphere and the sequestration of carbon away from the atmosphere for a given time period [@National_Academies_of_Sciences2022-ah].
+Several mCDR strategies have been proposed such as increasing alkalinity to enhance the chemical uptake of atmospheric CO$_{2}$ [@Hartmann2023-vk], increasing the photosynthetic uptake of atmospheric CO$_{2}$ with iron and macronutrients fertilization [@Blain2007-ws] or directly injecting CO$_{2}$ into the deep ocean [@Siegel2021-oz].
+
+We have seen with CONVERSE that carbon sequestration occurs on the whole water column and that using a fixed reference depth would likely miss the bulk of sequestration.
+Therefore, the evaluation of future mCDR strategies using biologically-driven carbon sequestration fluxes should be estimated over the whole water column.
+
+### Particle injection pumps
+
+Our study on the sequestration flux by the BCP also shows that the contribution of the vertical migration and mixing pumps contribute at least to 20-31% of the sequestration flux by the BCP.
+
+dire qu'il faut améliorer la physical et la VMP (diel et saisonnière) -\> mettre le paquet sur ARGO (et les gliders pour la physical) + papier d'Hervé de 2021.
++ préciser qye la migration saisonnire est sous-évaluée.

### 2023-03-22 index.qmd
update

@@
 
-I hope you will enjoy it and no, I won't tweet about it. -->
+I hope you will enjoy it.
 ```
+```{=html}
+<!-- This thesis started in probably the worst conditions possible but here it is, it is finally over.
+Why such a relief?
+I have my personal case which I really do not wish to anyone.
+I hope things have changed but when I reached for help at my university, I quickly understood that I had to shut my mouth.
+Wait, it is not that I have understood, I was told out loud that if I kept pushing, I was going to be laid off.
+I was not, do you want to know why?
+
+#Respect, I got so many emails about that, it is hilarious.
+Instead of tweeting, perhaps some people should listen and act.
+For what , academia is a centralized institution without safeguards.
+The same person can be faculty president, lab chief, group master and mental health hotline (true story) at the same time.
+The only difference with politics is that you do not get more money in your pocket by cumulating
+
+Here it is, the product of my resilience.
+I have to say that I am wrapping this up in a good shape, both mentally and physically.
+I am a lucky one.
+Some people are surprised that some of us do not want to pursue a career in academia.
+Just wait for it.
+
+I truly wish good luck to all current and future PhD students and I hope they can have the strength to say no when they need to. Trust me, it will be useful. -->
+```

### 2023-03-22 merci.qmd
update discussion
update

@@
 merci au comité de thèse
+
+merci antoine poteau et fabrizio depuis 2018
+
+merci a chatGPT pour la cover, elle n'est pas top mais j'avais la flemme, c'était suffisant
 
@@
 Evgeny et Arthur Capet et Charles Troupin
+
+Mathieu de m'avoir accueillu sur ton divan à Bx quand j'en avais besoin
 
@@
 Ophélie, je me grouille et j'arrive pour grimper, Alexandre, Switzeland see what you got in the legs
+
+ma maman qui a accepté de m'héberger dans une période incertaine même si elle connait mon besoin d'indépendance.
 
@@
 Girlfriend (because I am not easy going every day)
+
+Go Belgium and go Mons

### 2023-03-23 discussion.qmd
update discussion

@@
 Therefore, a significant part of my work was to test and prepare a fully functional data preparation and model training pipeline for the upcoming UVP6 data set using UVP5 images.
-In the following sections, I will describe a brief historic of what has been done prior to the acquisition and human labeling of UVP6 images.
+In the following sections, I will describe a brief historic of what has been done prior to the acquisition and human labeling of UVP6 images and discuss about the use and perspectives on the embedded classification.
 
@@
 instances), $K$ the number of classes, $y_{k}^{(i)}$ the true probability that the $i^{th}$ instance belongs to class $k$ (0 or 1) and $\hat{p}_{k}^{(i)}$ the predicted probability that the instance belongs to class $k$.
-However, the XGBoost library can be confusing in comparison with the paper of @Chen2016-xt and especially @eq-objective-function-xgboost.
-In practice, if one sets the objective function to `multi:softmax`, it means that the objective of the model is to predict the highest probability for the true class while minimizing a loss function specified by the user (in our case, cross entropy).
+In practice, if one sets the objective function to `multi:softmax`, it means that the objective of the model is to predict the highest probability for the true class while minimizing the cross entropy loss.
 :::
@@
 To this end, we trained a first classification model with classes from the UVP5 data set using a `multi:softprob` objective function.
-Then, we used the model to make a prediction on a test set that returned a matrix $(M \times N)$ of $M$ images and $N$ classes.
+Then, we used the model to make a prediction on a test set that returned a matrix $(M \times N)$ of $M$ images and $N$ probabilities ($N$ classes in the model).
 Afterwards, we computed four metrics on each row of the probability matrix: the maximum probability ($max$, equivalent to `multi:softmax`), the softmax of the probability array ($softmax$, equivalent to a double `multi:softmax`), the difference between the highest and the second highest probability ($diff_{max}$) and the difference between the highest and second highest probability of the softmax array ($diff_{softmax}$).
 We compared those metrics for each class of the test set in @fig-trust-index which clearly shows that all metrics were very similar and that they captured the errors at the same speed.
-Thus, it meant that we could keep using the probability returned by XGBoost a satisfactory classification score using `multi:softprob` as the objective function.
-@fig-trust-index also shows that some classes kept accumulating errors in the prediction with increasing rank (e.g.
-*Copepoda*, darksphere and fiber\<detritus).
-In their case, an increasing rank is not a trustworthy indicator of increasing confidence.
+This result is threefold: 1) in our case, there is no need to use a trust index in addition to the probabilities given by `multi:softprob`, 2) some classes kept accumulating errors in the prediction with increasing rank (e.g.
+*Copepoda*, darksphere and fiber\<detritus) thus an increasing rank is not always a trustworthy indicator of increasing confidence 3) it partly explains why the quantification did not improve the overall classification by either using a fixed probability threshold (i.e.
+TC) or using the probability matrix (i.e.
+SPAC).
 
@@
 The last technical limitation was the number of features (see @tbl-features).
-Before I started my PhD, 55 features (basically geometry, gray levels, statistical moments) were defined.
-I was not allowed (although, I asked!) to do some feature engineering, that is either doing feature selection (selecting the most useful features) either doing feature combination (combining existing features to create new ones) or creating new features by collecting new data [@Geron2019-is].
-Why was this proposition refused?
-Again, the 55 selected features had been tested earlier and they were compliant with the energy budget.
-Therefore, adding more features was not a possibility because it would have required retesting them and reevaluting energy budgets that were not possible at the time.
-In addition, all 55 features were computed quickly hence they were not consuming a lot of energy so it was decided to simply keep them.
-At worse, some features would have been redundant, without impacting the classification quality.
+Before I started my PhD, 55 features (basically geometry, gray levels and statistical moments) were defined and were compliant with the energy budget imposed by the float (the UVP6 is adapted for other AUVs but it was especially designed for profiling floats).
+When XGBoost results are compared with a state-of-the-art CNN, the main difference does not likely come from the classifiers themselves but most likely from the selected features.
+We could have done some feature engineering, that is either doing feature selection (selecting the most useful features) either doing feature combination (combining existing features to create new ones) or creating new features by collecting new data [@Geron2019-is].
+Therefore, texture features could have been added to our global 55 features.
+The main issue with the latter (e.g.
+histogram of oriented gradients for detecting edges, gray-level co-occurrence matrix for homogeneity and contrast) is that their processing time is proportional to the size of the image hence it was difficult to evaluate them in the energy budget.
 
-Nonetheless, when we compare the results from a features-based model and a CNN (@tbl-classification-report-uvp6-xgboost-cnn), perhaps the results would have been closer if we actually did some feature engineering and added other kinds of features.
+### Detritus, this ubiquitous mess
+
+@tbl-stats-uvp6 shows that the median number of objects detected in 3.5 m$^{3}$ is roughly 300.
+In addition, it is usual to have less than 20% of living objects hence the vast majority of images are labelled as detritus.
+The latter are ubiquitous in the ocean and display a wide variety of shapes, sizes and textures (@fig-detritus).
+In @sec-improve-classif, I showed that using several classes of detritus did not improve the classification model and that it actually made it worse because the N-detritus classes showed a high level of confusion between each other.
 
@@
 
-### Detritus, this ubiquitous mess
-
-We saw in the previous section that the median number of objects detected in 3.5 m$^{3}$ is roughly 300.
-In addition, it is usual to have less than 20% of living objects hence the vast majority of images are labelled as detritus.
-The latter are ubiquitous in the ocean and display a wide variety of shapes, sizes and textures (@fig-detritus).
-In @sec-improve-classif, I showed that using several classes of detritus did not improve the classification model and that it actually made it worse because the N-detritus classes showed a high level of confusion between each other.
-
 ![Examples of UVP objects labelled as detritus. The scale bar is the same for all (2 mm).](image/detritus/moz_detritus.jpg){#fig-detritus}
@@
 Using a clustering method to split the detritus into several classes (e.g.
-fluffy gray or dark compact) to improve the classification of biological objects was based on the strong belief of some colleagues that were convinced it would improve the classification.
+fluffy gray or dark compact) to improve the classification of biological objects was based on the strong belief that it would improve the classification.
 However, my work and the work of another PhD student that tried using several detritus classes using a CNN (not published yet) showed that it did not improve the classification model either.
@@
 
+```{=latex}
+\newpage
+```
 ### The future of the embedded classification
@@
 
-@fig-kerguelen present some preliminary results from the UVP6 embedded classification that highlights the difference in *Calanoida* and *Chaetognatha* concentrations between February and March 2023 in the western and eastern part of the Kerguelen Plateau.
-For these floats, the configuration was different from the ones deployed in the Labrador Sea with a higher profiling time resolution after deployment, which can be further modified based on observations.
-It is only the beginning.
+@fig-kerguelen present some preliminary results from the UVP6 embedded classification that highlights the difference in *Calanoida*, *Chaetognatha*, detritus, puff and tuff concentrations between February and March 2023 in the western and eastern part of the Kerguelen Plateau.
+For these floats, the configuration was different from the ones deployed in the Labrador Sea with a higher profiling time resolution after deployment.
 
-```{r, kerguelen, fig.width=8, fig.height=5}
+```{r, kerguelen, fig.width=9, fig.height=8}
 #| label: fig-kerguelen
-#| fig-cap: $\textit{Calanoida}$ and $\textit{Chaetognatha}$ concentrations in February-March 2023 in the Kerguelen Plateau.
+#| fig-cap: $\textit{Calanoida}$, $\textit{Chaetognatha}$, detritus, puff and tuff (i.e. $\textit{Trichodesmium}$) concentrations in February-March 2023 on the Kerguelen Plateau.
 
@@
   filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'Chaetognatha - East')
+
 chaeto_west <- vroom('DATA/ARGO/4903657_class_number_5_FromNetCDF.csv') |> 
@@
 
-taxo_kerguelen <- rbind(calanoids_east, calanoids_west, chaeto_east, chaeto_west) |> 
-  mutate(group = factor(group, levels = c('Calanoida - East', 'Calanoida - West',
-                                     'Chaetognatha - East', 'Chaetognatha - West')))
+detritus_east <- vroom('DATA/ARGO/1902593_class_number_13_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'detritus - East')
 
-ggplot(taxo_kerguelen) + geom_point(aes(x = conc_object, y = depth, colour = date)) + scale_colour_viridis_c(trans = "date",labels = c("2023-02-11","2023-02-18","2023-02-25","2023-03-04","2023-03-11"), breaks = c(as.Date("2023-02-11"), as.Date("2023-02-18"), as.Date("2023-02-25"), as.Date("2023-03-04"), as.Date("2023-03-11"))) + scale_y_reverse() + theme_bw() +
+detritus_west <- vroom('DATA/ARGO/4903657_class_number_13_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'detritus - West')
+
+puff_east <- vroom('DATA/ARGO/1902593_class_number_16_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'puff - East')
+
+puff_west <- vroom('DATA/ARGO/4903657_class_number_16_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'puff - West')
+
+tuff_east <- vroom('DATA/ARGO/1902593_class_number_19_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'tuff - East')
+
+tuff_west <- vroom('DATA/ARGO/4903657_class_number_19_FromNetCDF.csv') |> 
+  filter(PhaseName == 'NPAR') |> mutate(date = as.Date(date), group = 'tuff - West')
+
+# tuff_west |> filter(cycle == 9) |> ggplot(aes(y = -depth, x = conc_object))
+
+taxo_kerguelen <- rbind(calanoids_east, calanoids_west, chaeto_east, chaeto_west,
+                        detritus_east, detritus_west,
+                         puff_east, puff_west, tuff_east, tuff_west)
+
+ggplot(taxo_kerguelen) + geom_point(aes(x = conc_object, y = depth, colour = date)) + scale_colour_viridis_c(trans = "date",labels = c("2023-02-11","2023-02-18","2023-02-25","2023-03-04","2023-03-11", "2023-03-18"), breaks = c(as.Date("2023-02-11"), as.Date("2023-02-18"), as.Date("2023-02-25"), as.Date("2023-03-04"), as.Date("2023-03-11"), as.Date("2023-03-18"))) + scale_y_reverse() + theme_bw() +
   labs(x = 'Concentration (#/L)', colour = "Date", y = 'Depth (m)') +
-  facet_wrap(~group, scale = "free_x")
+  facet_wrap(~group, scale = "free_x", nrow = 5)
 ```
+
+```{=latex}
+\newpage
+```
+The observed concentrations of the 4 biological classes in @fig-kerguelen needs to be interpreted with the UVP6 classification report (@tbl-classification-report-uvp6-xgboost-cnn).
+*Calanoida* have a precision and recall of, respectively, 66% and 69% thus we potentially miss 31% of them (i.e.
+31% of *Calanoida* are non detected because they are classified in another class) but 34% of images classified as *Calanoida* are false positives and should be placed in another class.
+In that case, the observed concentrations can therefore be resumed as the real concentration.
+*Chaetognatha* have a precision and recall of, respectively, 80% and 62% therefore when a *Chaetognaha* is detected, the model is confident (20% of false positives) but it misses 38% of them so we might have more *Chaetognatha* than measured by the UVP6.
+However, One must also look at the concentration values because it is likely that the 2 points measured on the eastern part of the Keguerlen Plateau are misclassified objects.
+We could therefore conclude to the absence *Chaetognatha* on the eastern part while their presence is likely on the western part because, on the one hand, this class has a high precision and, on the other hand, they are observed on successive short-lapsed vertical profiles and between 0-500 m where they likely feed on *Calanoida*.
+
+Puff and tuff are both *Trichodesmium* which are usually observed in tropical and subtropical regions [@Westberry2006-bt] hence their presence on the Kerguelen Plateau is dubious.
+Tuff have a low precision and recall (55% and 63%) in contrast to puff (73% and 78%).
+On the eastern part, both classes have low concentrations and their position in the water column should be questioned for photosynthetic organisms.
+On the western part, they also both have low concentrations (0.25 #/L for puff is likely false) even though detritus concentrations are 5-10 times higher than on the eastern part.
+
+The first results of the embedded classification are promising.
+However, they need to be interpreted with the UVP6 classification report and contextualized (e.g.
+species geographical and temporal distribution, correlation with the presence of other species, position in the water column).
 
@@
 Transformers [@Vaswani2017-il], also referred to as Foundation Models in @Bommasani2021-kb are the latest cutting-edge tool in the field of computer vision and natural language processing [e.g.
-@Li2022-qt; @Liu2021-jq].
+@Liu2021-jq; @Li2022-qt].
 The structure of transformers relies on attention mechanisms able to draw global dependencies between input and output [@Vaswani2017-il].
@@
 
+```{=latex}
+\newpage
+```
 ## Deriving carbon fluxes from particle size distribution
@@
 CONVERSE sequestration flux estimates are also up to 4 times larger than those from studies assuming that sequestration occurs below a fixed reference depth [@Honjo2008-kn, @Guidi2015-sj; @Henson2012-ck].
-Here below, I highlight some key points and ways to improve further our estimations.
+Here below, I highlight some key points and ways to improve our estimations.
 
@@
 We have seen with CONVERSE that carbon sequestration occurs on the whole water column and that using a fixed reference depth would likely miss the bulk of sequestration.
-Therefore, the evaluation of future mCDR strategies using biologically-driven carbon sequestration fluxes should be estimated over the whole water column.
+Therefore, the evaluation of future mCDR strategies using biologically-driven carbon sequestration fluxes should be estimated over the entire water column.
 
@@
 
-Our study on the sequestration flux by the BCP also shows that the contribution of the vertical migration and mixing pumps contribute at least to 20-31% of the sequestration flux by the BCP.
+Our study on the sequestration flux by the BCP also shows that the contribution of the vertical migration and mixing pumps contribute at least to 21-34% of the sequestration flux by the BCP.
+However, those two pumps could probably be better estimated with a multi-instruments approach [@Claustre2021-tx].
+
+#### Physically mediated particle injection pump
+
+With CONVERSE, we estimate that the mixing pump contributes to 11-23% of the sequestration flux by the BCP.
+However, we were not able to distinct the contribution of the three identified pathways composing the mixing pump due to lacking information in the literature.
+This could be improved by using the observational framework developed in @Claustre2021-tx by combining satellite, ship-based, mooring and AUV data.
+BGC-Argo floats have shown that they are adapted to study the detrainment of mixed-layer particles [@DallOlmo2014-ot, @DallOlmo2016-ex] and so have gliders for the eddy subduction pump [@Stukel2017-lc; @Omand2015-mf].
+Nonetheless, the study of the large-scale subduction pump will be more challenging because it will require a larger number of measurements made at the scale of ocean basins and over annuals cycles [@Claustre2021-tx], hence the need for a multi-instruments approach to better understand it.
+
+#### Biologically mediated particle injection pump
+
+With CONVERSE, we estimate that the vertical migration pump contributes to 9-12% (diel) and 0.1-0.4% (seasonal).
+Our estimation of the seasonal migration pump is limited to the northern North Atlantic hence it is underestimated.
+The
 
@@
 + préciser qye la migration saisonnire est sous-évaluée.
++ acoustique?
+passsive acoustique?
+parler de nowicky pour la migration saisonnière en disant qu'il y a aussi les poissons qui peuvent migrer etc

### 2023-03-23 uvp6_part.qmd
update discussion

@@
 As a result, even though *Calanoida* have been shown to overwinter in deep waters (\> 1000 m) [@Krumhansl2018-dr], an isolated signal at depth does not mean a copepod was really there.
-Similarly, *Trichodesmium* should not be observed outside the tropical and subtropical regions @Westberry2006-bt therefore depending on the sampling region, some results can directly be rejected.
+Similarly, *Trichodesmium* should not be observed outside the tropical and subtropical regions [@Westberry2006-bt] therefore depending on the sampling region, some results can directly be rejected.
 Here, tuff are predicted in higher concentrations than puff which are only predicted at one depth.

### 2023-03-24 discussion.qmd
update

@@
 Therefore, a significant part of my work was to test and prepare a fully functional data preparation and model training pipeline for the upcoming UVP6 data set using UVP5 images.
-In the following sections, I will describe a brief historic of what has been done prior to the acquisition and human labeling of UVP6 images and discuss about the use and perspectives on the embedded classification.
+In the following sections, I will describe a brief history of what has been done prior to the acquisition and human labeling of UVP6 images and discuss about the use and perspectives on the embedded classification.
 
@@
 
-```{r, kerguelen, fig.width=9, fig.height=8}
+```{r, kerguelen, fig.width=8, fig.height=10}
 #| label: fig-kerguelen
@@
 In other words, it is able to learn dependencies between distant positions in a network architecture in constrast to CNNs.
-As an example, @Li2022-qt used a vision transformer model combining a CNN and a transformer to identify apple diseases.
+As an example, @Li2022-qt used a vision transformer model combining a CNN and a vision transformer to identify apple diseases.
 In their model, the CNN is used to extract the global features of the input image while the transformer extracts local features of the disease region to help the CNN see better.
@@
 
+```{=latex}
+\newpage
+```
 ### The straightforward approach
@@
 #| label: fig-uvp-trap-mapping
-#| fig-cap: Spatial distribution of sediment traps (A), $^{234}Th$ (B), UVP5 (C) and their spatio-temporal match-ups (D)
+#| fig-cap: Spatial distribution of (A) sediment traps, (B) $^{234}Th$, (C) UVP5 and (D) their spatio-temporal match-ups.
 
@@
 The idea was based on the assumption that carbon fluxes could be similar in distinct regions of the ocean if those regions had similar environmental conditions thus removing spatial (horizontal) and temporal constraints on match-ups.
-For this, we tried different combinations of environmental variables such as surface temperature, salinity, nutrients, combination of nutrients, Chla, $b_{bp}$ or water column properties such as integrated Chla content, oxygen minimum and MLD from monthly climatologies.
+For this, we tried different combinations of environmental variables such as surface temperature, salinity, nutrients, Chla, $b_{bp}$ or water column properties such as integrated Chla content, oxygen minimum and MLD from monthly climatologies.
 Then, we used those environmental variables to build a 3D environmental space using a 3D principal component analysis (PCA) decomposition and we projected UVP5 data and trap/$^{234}Th$ in that space based on the environmental characteristics from the respective coordinates and month of each UVP5 and trap/$^{234}Th$ profiles.
@@
 
-This method increased the number of match-ups to \~7800 however the minimization did not find a clear local minimum in the solution space (@fig-solution-space).
-Instead of having a clear local minimum, @fig-solution-space shows what looks like a 2D solution plane.
+This method increased the number of match-ups to \~7800 however the minimization did not find a non ambiguous minimum in the solution space (@fig-solution-space).
+Instead of having a global minimum, @fig-solution-space shows what looks like a 2D solution plane.
 In addition, I think the choice of $A$ and $D$ (or $B$, $B = 2D - 1$ ) is not constrained enough.
-In Figure 5 of @Guidi2008-vp, I think it is difficult to really say that there is a local minimum.
-The light gray color area (low residual error) encompasses a wide region of $A$ and $B$ values where their jackknife procedure found local minima with ($A$, $B$) pairs such as (50, 2) or (150, 5).
+In Figure 5 of @Guidi2008-vp, the light gray color area (low residual error) encompasses a wide region of $A$ and $B$ values where their jackknife procedure found local minima with ($A$, $B$) pairs such as (50, 2) or (150, 5).
 While $A$ is seen as a constant with no physical meaning, the value of $B$ should really be questioned since it is related to the fractal dimension $D$ of particles and aggregates.
 Fractal dimensions should not exceed 3 (physical limit) but they have been estimated down to 1.1-1.3 [@Alldredge1988-lx].
-In addition, it is possible to obtain coherent POC fluxes by tuning the $A$ value for a fixed $B$ therefore I have my doubts about this method because I am not sure $A$ and $B$ are trully independent.
+In addition, it is possible to obtain coherent POC fluxes by tuning the $A$ value for a fixed $B$ therefore I am not sure that $A$ and $B$ are truly independent.
 
@@
 Once the model is trained, it can be applied to the entire data set of PSDs (\> 1.5 million) to derive POC fluxes.
+We tried to build such a model with a few predictors (i.e.
+depth, surface temperature, Chla, $b_{bp}$ and the oxygen minimum in the water column).
 Our model had a $R^{2}$ of 0.47 with a percentage of explained variance dominated by depth (54%), surface Chla (16%), surface temperature (13%) and the oxygen minimum (10%).
-The minimization of @Guidi2008-vp applied on all true PSD and predicted POC fluxes led to the same kind of solution space as shown in @fig-solution-space.
+The minimization of @Guidi2008-vp applied on all true PSDs and predicted POC fluxes led to the same kind of solution space as shown in @fig-solution-space.
 
@@
 The study of @Denvil-Sommer2022-oi used both random forest and XGBoost models to predict small and large POC concentrations using geographical variables (depth, latitude and longitude), physical variables (incident light, MLD and temperature), chemical (phosphate and nitrate) and biological variables (Chla and 12 Plankton Functional Types (PFTs) such as diatoms or microzooplankton).
-@Denvil-Sommer2022-oi show that for small POC, the most important predictors are the depth level, temperature, microzooplankton and phosphate while the large POC is mostly driven by depth, temperature, MLD, microzooplankton, phaeocystis (another PFT), phosphate and Chla averaged over the MLD.
+@Denvil-Sommer2022-oi showed that for small POC, the most important predictors are the depth level, temperature, microzooplankton and phosphate while the large POC is mostly driven by depth, temperature, MLD, microzooplankton, phaeocystis (another PFT), phosphate and Chla averaged over the MLD.
 Overall, the results of these studies suggest that it is possible to link the surface environmental and ecosystem structure to the POC distribution in the ocean interior.
@@
 With the deployment of new BGC-Argo floats equipped with both the UVP6 and the OST as well as the core biogeochemical variables (e.g.
-Chla, $b_{bp}$, incident light) of the BGC-Argo program, we will hopefully be able to better constraint the PSD-POC relationship however we will probably need a few more years to acquire enough profiles in different oceanic regions to better understand the key drivers of the downward POC flux.
+Chla, $b_{bp}$, incident light) of the BGC-Argo program, we will hopefully be able to better constraint the PSD-POC relationship however it will probably need a few more years to acquire enough profiles in different oceanic regions to better understand the key drivers of the downward POC flux.
 
+```{=latex}
+\newpage
+```
 ## Continuous vertical sequestration
@@
 This is 2-3 and 3-6 times larger than POC fluxes at, respectively, 1000 and 2000 m.
-CONVERSE sequestration flux estimates are also up to 4 times larger than those from studies assuming that sequestration occurs below a fixed reference depth [@Honjo2008-kn, @Guidi2015-sj; @Henson2012-ck].
+CONVERSE sequestration flux estimates are also up to 4 times larger than those from studies assuming that sequestration occurs below a fixed reference depth [@Honjo2008-kn; @Guidi2015-sj; @Henson2012-ck].
 Here below, I highlight some key points and ways to improve our estimations.
@@
 
-Our study on the sequestration flux by the BCP also shows that the contribution of the vertical migration and mixing pumps contribute at least to 21-34% of the sequestration flux by the BCP.
+Our study on the sequestration flux by the BCP also shows that the contribution of the vertical migration and mixing pumps represent at least 21-34% of the sequestration flux by the BCP.
 However, those two pumps could probably be better estimated with a multi-instruments approach [@Claustre2021-tx].
@@
 This could be improved by using the observational framework developed in @Claustre2021-tx by combining satellite, ship-based, mooring and AUV data.
-BGC-Argo floats have shown that they are adapted to study the detrainment of mixed-layer particles [@DallOlmo2014-ot, @DallOlmo2016-ex] and so have gliders for the eddy subduction pump [@Stukel2017-lc; @Omand2015-mf].
-Nonetheless, the study of the large-scale subduction pump will be more challenging because it will require a larger number of measurements made at the scale of ocean basins and over annuals cycles [@Claustre2021-tx], hence the need for a multi-instruments approach to better understand it.
+BGC-Argo floats have shown that they are adapted to study the detrainment of mixed-layer particles [@DallOlmo2014-ot; @DallOlmo2016-ex] and so have gliders for the eddy subduction pump [@Stukel2017-lc; @Omand2015-mf].
+Nonetheless, the study of the large-scale subduction pump will be more challenging because it will require a larger number of measurements made at the scale of ocean basins and over annual cycles [@Claustre2021-tx], hence the need for a multi-instruments approach to better understand it.
 
@@
 With CONVERSE, we estimate that the vertical migration pump contributes to 9-12% (diel) and 0.1-0.4% (seasonal).
-Our estimation of the seasonal migration pump is limited to the northern North Atlantic hence it is underestimated.
-The
+The estimation of the latter was limited to the northern North Atlantic thus it is clearly underestimated.
+However, the estimation of vertically migrating (either daily or seasonally) zooplankton and other animals (e.g.
+fish, jellyfish) is expected to improve with robotic measurements at high temporal resolution and over the entire annual cycle [@Claustre2021-tx].
+The combination of acoustic sensors [@Reiss2021-rk; @De_Robertis2021-nw], optical sensors [@Picheral2021-sx] and vertical profiles of biogeochemical variables (e.g.
+Chla, $b_{bp}$, CDOM) have already shown promising results to detect the migration of mesopelagic organisms in the water column [@Haentjens2020-dc].
 
-dire qu'il faut améliorer la physical et la VMP (diel et saisonnière) -\> mettre le paquet sur ARGO (et les gliders pour la physical) + papier d'Hervé de 2021.
-+ préciser qye la migration saisonnire est sous-évaluée.
-+ acoustique?
-passsive acoustique?
-parler de nowicky pour la migration saisonnière en disant qu'il y a aussi les poissons qui peuvent migrer etc
+```{=latex}
+\newpage
+```
+## Contributions of this work
+
+The embedded classification for AUVs contributes to a better comprehension and quantification of the vertical migration pump.
+So far, BGC-Argo floats equipped with the UVP6 and the classification model have been deployed in some oceanic regions (e.g.
+Labrador Sea, Kerguelen Plateau) and have just started to acquire data.
+In addition to the particle (\> 0.6 mm) identification, the UVP6 also sends their mean area and gray levels which will hopefully contribute to better characterize their morphological traits and therefore better predict their sinking rates (cfr.
+size versus excess density).
+The latter are important to estimate the downward carbon flux and the sequestration time.
+
+The BCP can no longer be seen as the sole gravitational settling of particles.
+Physically and biologically mediated particle injection pumps need to be taken into consideration which is highlighted by our work that estimates their contribution to 21-34% of the total BCP sequestration flux.
+Our work also shows that the common assumption of sequestration below a fixed reference depth (typically 1000 or 2000 m) misses the bulk of sequestration flux.
+Therefore, the evaluation of mCDR efficiencies using biologically-driven carbon sequestration fluxes should be estimated over the entire water column.
+
+This work has shown the first comparison of BGC-Argo floats equipped with both the UVP6 and an OST to better quantify carbon fluxes from particle counts and beam attenuation.
+If the prediction of carbon fluxes at global scale was not a success, this work has highlighted current difficulties and ways to more forward using machine learning approaches.
+The deployment of BGC-Argo floats in the frame of REFINE with the combination of UVP6 measurements (particle counts and identification), OST beam attenuation fluxes as well as bio-optical variables (e.g.
+Chla, $b_{bp}$, incident light) will likely improve our understanding of processes that affect the attenuation of the downward carbon flux at local scale.

### 2023-03-24 index.qmd
update

--- 2023-01-22 index.qmd
@@
 
-```{=html}
-<!--  As of 2022, it has now become obvious to the general public that the Earth's climate is changing.
+As of 2022, it has now become obvious to the general public that the Earth's climate is changing.
 Massive flood events, repeated heatwaves, never-ending droughts, disappearing glaciers, biodiversity collapse, … Reports on weather anomalies were already published in the 1960's, the Intergovernmental Panel on Climate Change (IPCC) was created in 1988, the Kyoto Protocol was adopted in 1997, An Inconvenient Truth was released in 2006, the Paris Agreement was signed in 2015 and still, what has really changed?
 
-As I am writing this part, the COP27 (sponsored by Coca-Cola, no joke) is taking place with an acknowledged record number of fossil fuels lobbyists, bloody stadiums with outside air conditioning are about to be used in the desert for a month and it seems like 'leaders' of the Free World have been having hearing issues.
+As I am writing this part, the COP27 (sponsored by Coca-Cola, no joke) is taking place with an acknowledged record number of fossil fuel lobbyists, bloody stadiums with outside air conditioning are about to be used in the desert for a month and it seems like 'leaders' of the Free World have been having hearing issues.
 See for yourself:
@@
 
-*"Of course, I would love building new pipelines in your garden"* said the Free World
+*"Aaaah okay. Yes you're absolutely right, let's save the banks one more time"* said the Free World
 
-While I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured, there are no fake news in this work.
-
-I hope you will enjoy it and no, I won't tweet about it. -->
-```
+Although I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured that there is no fake news in this work. I hope you will enjoy it.

### 2023-03-24 intro.qmd
update

--- 2023-03-10 intro.qmd
@@
 @Buesseler2006-og].
-More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size distributions (PSD, also referred to in the literature as particle size spectra (PSS)) that could then be converted to carbon fluxes [e.g.
+More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size distributions (PSDs, also referred to in the literature as particle size spectra (PSS)) that could then be converted to carbon fluxes [e.g.
 @Guidi2008-vp].
@@
 
-However, there are still many uncertainties regarding the magnitude of VMP [@Archibald2019-dj; @Aumont2018-wx; @Gorgues2019-qp] and its sequestration potential [@Boyd2019-ez; @Pinti2022-dc] thus its importance for a better understanding of the BCP.
+However, there are still many uncertainties regarding the magnitude of the VMP [@Archibald2019-dj; @Aumont2018-wx; @Gorgues2019-qp] and its sequestration potential [@Boyd2019-ez; @Pinti2023-td] thus its importance for a better understanding of the BCP.
 
@@
 
-The impact on the downward POC flux of zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2023-qj].
+The impact on the downward POC flux by zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2023-qj].
 On the one hand, aggregate feeding, filter-feeding, direct ingestion of fecal pellets (i.e.
@@
 Because the sinking speed determines the time available for degradation in the water column, it also directly impacts the potential for carbon sequestration.
-With all things considered, there is a thus need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Giering2020-zk; @Iversen2023-qj; @Williams2022-xx].
+With all things considered, there is thus a need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Giering2020-zk; @Iversen2023-qj; @Williams2022-xx].
 
@@
 
-: Range of depths generally reached by POC and DOC via the six BCPs [@Claustre2021-tx] and sequestration time potential evaluated by the model of @Boyd2019-ez with an average estimation from the model of @Nowicki2022-no in brackets.
+: Range of depths generally reached by POC and DOC via the six BCPs [@Claustre2021-tx] and sequestration time potentials evaluated by the model of @Boyd2019-ez with an average estimation from the model of @Nowicki2022-no in brackets.
 None means that a pump does not transport DOC.

### 2023-03-24 merci.qmd
update

--- 2023-03-18 merci.qmd
@@
-remerciement
+# Acknowledgments
 
-thomas boniface pour les figures
+Thanks to the members of the jury for taking the time to read and evaluate this three and a half year work. I am looking forward to debate it during the defense.
 
-anais
+Cette thèse a été particulière à bien des égards, avec son lot de positifs et de négatifs mais surtout avec son lot de rebondissements qui m'ont permis d'affûter mes connaissances, ma résilience et je l'espère ma tolérance.
+Pour cela, je tiens à remercier chaleureusement toutes les personnes y ayant contribué de près ou de loin.
 
-alexbarth pour son bouquin sur le machine car j'ai passé mes 3 première semaines à faire ça vu que j'y connaissais absolument rien
+Merci à Hervé, Lionel et Jean-Olivier pour leur accueil au Laboratoire de Villefranche-sur-Mer et pour leur disponibilité ainsi que pour leurs conseils avisés, tant scientifiques que personnels.
 
-faire des stats du vélo (et meme mettre une carte si possible rapidement)
+Merci à Alex, Charles et Arthur pour leur aide et leur gentillesse.
 
-tous ceux qui ont partagé des moments sportifs ensembles, au bloc mais surtout en vélo et qui malgré la souffrance acceptaient encore de venir rouler avec moi (Ophélie, Flavien, Thelma, Laetitia, Alexandre, Satoshi)
+Merci également aux personnes que j'ai pu cotoyer à COMPLEX comme chez OMTAB depuis mon passage en stage en 2018 et les multiples allers-retours Belgique-France.
 
-Julien, des challenges à n'en plus finir, un peu mazo au final, à force d'aller en enfer, tu incarneras le diable en personne bientôt(#Ducasse)
+Merci à Antoine Mangin de m'avoir prêté un bureau dans les locaux d'ACRI sur Grasse l'été dernier pendant que mon appartement servait d'Airbnb estival pour le plus grand bonheur du portefeuille de mon proprio.
 
-Les montois (Mathieu, inconditionel, Florent, les Juliens, Chris)
+Un grand merci à mes colocs de bureau: Ophélie, Laetitia et Thelma pour votre soutien et votre bonne humeur sans faille (contrairement à moi).
 
-Les co-Phd : Chloé, AMP (the best), Anais
+Merci à AMP (elle se reconnaitra), Chloé et Anaïs pour les pauses midi sur le ponton.
+AMP, merci d'avoir accepté de vivre virtuellement avec moi afin que la sécu daigne enfin me donner une carte vitale !
 
-Hadrien who accepted to be my roomate in Liège
+Ne me demandez pas des noms de restos ou de bars sur la côte d'Azur, je n'en ai aucune idée.
+Par contre, je connais pas mal les routes du coin.
+Merci à tous ceux qui ont partagé (ou subi) des kilomètres avec moi, surtout en montée (le plat pays, qui est le mien): Flavien, Ophé, Lucas, Thelma, Laeti et Satoshi.
 
-alice
+Un grand merci à Marine et Alice pour leur bienveillance. Marine, je te souhaite le meilleur pour tes projets futurs mais fais attention à qui tu donnes tes clés quand même.
 
-marine of course
+Merci également à Alex qui a été un excellent roommate pendant les confinements, couvre-feux et ordonnances en tout genre. See you in London !
 
-expatrié que je suis et qui venu me voir 'ou profiter, I'll never really know)
+Merci aux Liégeois(es) (Thomas, Florent, Flop, Nico, Evgeny et Lucie) et en particulier Hadrien qui a accepté de m'avoir comme coloc pendant plusieurs mois (et qui y a survécu).
 
-les Liégeois (Thomas, Nico, Hadri, Flop, Florent)
+Merci aux Montois (Mathieu, Chris, Flo, Ju & Ju) pour votre soutien à distance, un mariage en 3 actes et pour vos visites sur Nice.
+Petite dédicace à Ju K qui, à force d'avoir joué avec le diable en subissant l'enfer des pourcentages, va littéralement en devenir un.
+Vivement 2024 pour casser la baraque.
 
-Lucie, mon old roomate (lucie lelotte)
+Un très grand merci à ma famille nucléaire qui ne comprend toujours pas mon job mais qui fait semblant quand je lui explique, c'est le principal.
+C'est vrai que la défense tombe mal au niveau timing donc on se voit à la Ducasse :-)
 
-Alex, my 15th french history roomate, I am glad you were there suring the 6 pm curfew we had and
-
-Les montois bien sûr
-
-tous xeux qui sont descendus me voir, (vous veniez pour moi ou la mer tirquoise, j'ai des doutes mtn)
+Enfin, un grand merci à une personne toute particulière qui en a probablement marre que je parle de la Belgique mais qui à présent est parfaite bilingue, mission accomplie?

### 2023-03-24 observingCarbonCycle.qmd
update

@@
 The former counts particles in different class sizes in order to compute the PSD while the latter extracts pixel-related and geometric features from each image to classify them (i.e.
-assign them a taxonomic group) directly on board using artificial intelligence (@sec-uvp-ia) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].
+assign them a taxonomic group) directly on board using artificial intelligence (@sec-uvp6) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].
 
@@
 Many studies have assumed that carbon sequestration only occurs in the deep ocean (typically 1000 m).
-In @bcp-sequestration, we reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
+In @sec-bcp-sequestration, we reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.

### 2023-03-24 summary.qmd
update

@@
 The ocean is known to play a key role in the carbon cycle.
-Without it, atmospheric CO2 levels would be much higher than what they are today thanks to the presence of carbon pumps that maintain a gradient of dissolved inorganic carbon (DIC) between the surface and the deep ocean.
+Without it, atmospheric CO$_{2}$ levels would be much higher than what they are today thanks to the presence of carbon pumps that maintain a gradient of dissolved inorganic carbon (DIC) between the surface and the deep ocean.
 The biological carbon pump (BCP) is primarily responsible for this gradient.
@@
 The last part consists in revisiting the BCP using a new framework called CONVERSE for Continuous Vertical Sequestration.
-With this new approach, we re-evaluate the total carbon sequestered from the atmosphere (> 100 years) by the BCP and its transport pathways on the entire water column, in contrast to the carbon sequestration typically assumed below a fixed reference depth.
+With this new approach, we re-evaluate the total carbon sequestered from the atmosphere (\> 100 years) by the BCP and its transport pathways on the entire water column, in contrast to the carbon sequestration typically assumed below a fixed reference depth.
 
@@
 ```
-
 **Vers une meilleur compréhension du transport du carbone dans l'océan global**
@@
 L'océan est connu pour jouer un rôle clé dans le cycle du carbone.
-Sans lui, les niveaux de CO2 atmosphérique seraient bien plus élevés qu'aujourd'hui grâce à la présence de pompes à carbone qui maintiennent un gradient de carbone inorganique dissous (DIC) entre la surface et le fond de l'océan.
+Sans lui, les niveaux de CO$_{2}$ atmosphérique seraient bien plus élevés qu'aujourd'hui grâce à la présence de pompes à carbone qui maintiennent un gradient de carbone inorganique dissous (DIC) entre la surface et le fond de l'océan.
 La pompe à carbone biologique (BCP) est principalement responsable de ce gradient.

### 2023-03-24 uvp6_part.qmd
update

@@
 
-kable(cf_uvp5_xgboost_cnn) |> row_spec(0, bold=T, align = 'c') |> #row_spec(1:16, align = 'c') |>
-  row_spec(row = 12:16, background = '#F2F2F2') |> add_header_above(c(" " = 1, "XGBoost" = 3, "CNN" = 3))
-  #column_spec(2, background = spec_color(cf_uvp5_xgboost$precision))
-  #row_spec(12:16, background = '#F2F2F2')
+# kable(cf_uvp5_xgboost_cnn) |> row_spec(0, bold=T, align = 'c') |> #row_spec(1:16, align = 'c') |>
+#   row_spec(row = 12:16, background = '#F2F2F2') |> add_header_above(c(" " = 1, "XGBoost" = 3, "CNN" = 3))
+#   #column_spec(2, background = spec_color(cf_uvp5_xgboost$precision))
+#   #row_spec(12:16, background = '#F2F2F2')
+
+kable(cf_uvp5_xgboost_cnn) |> row_spec(0, bold=T, align = 'c') |> add_header_above(c(" " = 1, "XGBoost" = 3, "CNN" = 3))
 ```
@@
 
-kable(cf_uvp6_xgboost_cnn) |> row_spec(0, bold=T, align = 'c') |> #row_spec(1:16, align = 'c') |>
-  row_spec(row = 21:25, background = '#F2F2F2') |> add_header_above(c(" " = 1, "XGBoost" = 3, "CNN" = 3))
+# kable(cf_uvp6_xgboost_cnn) |> row_spec(0, bold=T, align = 'c') |> #row_spec(1:16, align = 'c') |>
+#   row_spec(row = 21:25, background = '#F2F2F2') |> add_header_above(c(" " = 1, "XGBoost" = 3, "CNN" = 3))
+
+kable(cf_uvp6_xgboost_cnn) |> row_spec(0, bold=T, align = 'c') |> add_header_above(c(" " = 1, "XGBoost" = 3, "CNN" = 3))
 ```

### 2023-03-25 discussion.qmd
typo check

@@
 
-Using the CONVERSE approach, we have estimated that the global carbon sequestration (\> 100 years) flux by the BCP is 0.9-2.6 Pg C y$^{-1}$.
-This is 2-3 and 3-6 times larger than POC fluxes at, respectively, 1000 and 2000 m.
+Using the CONVERSE approach, we have estimated that the global carbon sequestration ($\geq$ 100 years) flux by the BCP is 0.9-2.6 Pg C y$^{-1}$.
+This is 2-3 and 3-6 times larger than POC fluxes at, respectively, 1,000 and 2,000 m.
 CONVERSE sequestration flux estimates are also up to 4 times larger than those from studies assuming that sequestration occurs below a fixed reference depth [@Honjo2008-kn; @Guidi2015-sj; @Henson2012-ck].
@@
 Physically and biologically mediated particle injection pumps need to be taken into consideration which is highlighted by our work that estimates their contribution to 21-34% of the total BCP sequestration flux.
-Our work also shows that the common assumption of sequestration below a fixed reference depth (typically 1000 or 2000 m) misses the bulk of sequestration flux.
+Our work also shows that the common assumption of sequestration below a fixed reference depth (typically 1,000 or 2,000 m) misses the bulk of sequestration flux.
 Therefore, the evaluation of mCDR efficiencies using biologically-driven carbon sequestration fluxes should be estimated over the entire water column.

### 2023-03-25 intro.qmd
typo check

@@
 
+![Illustration of the main factors influencing the sinking velocity following Stokes' Law and hypothetical attenuation fluxes (in % of export flux) with depth (vertical axis) for slow and fast sinking particles.
+The latter shows that slow sinking particles experience higher rates of flux attenuation due to increased duration of exposure to remineralization hence contributing less to carbon sequestration.
+Figure taken from @Williams2022-xx.](image/intro/sinking_speed.jpeg){#fig-sinking-speed}
+
 Taking this into consideration, several processes can impact the sinking speed of POC such as aggregation (e.g.
@@
 
-![Illustration of the main factors influencing the sinking velocity following Stokes' Law and hypothetical attenuation fluxes (in % of export flux) with depth (vertical axis) for slow and fast sinking particles.
-The latter shows that slow sinking particles experience higher rates of flux attenuation due to increased duration of exposure to remineralization hence contributing less to carbon sequestration.
-Figure taken from @Williams2022-xx.](image/intro/sinking_speed.jpeg){#fig-sinking-speed}
-
 ### Sequestration potential {#sec-will-it-be-sequestered}
@@
 | Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
-|-----------------|-------------------|-----------------|---------------------|
+|------------------|------------------|------------------|--------------------|
 | Gravitational          | export depth - sediment          | None                      | N/A (142)                            |

### 2023-03-25 observingCarbonCycle.qmd
typo check

@@
 
-Because measuring directly the POC export flux would be challenging, it is common to estimate the fraction of the total NPP that is exported (i.e.
+Because measuring directly the POC export flux is challenging, it is common to estimate the fraction of the total NPP that is exported (i.e.
 export efficiency) [@Henson2011-mt].
@@
 satellite and/or in situ data) or by optimizing model parameters to the observed fields [@Siegel2022-wd].
-@DeVries2017-wu combined both satellite-derived data (NPP and PSS) and in situ observations (POC fluxes, DOC concentrations and dissolved $O_{2}$ distributions) to assess the carbon export and its transfer in the mesopelagic zone using a steady state ocean circulation inverse model (i.e.
+@DeVries2017-wu combined both satellite data (NPP and PSS) and in situ observations (POC fluxes, DOC concentrations and dissolved $O_{2}$ distributions) to assess the carbon export and its transfer in the mesopelagic zone using a steady state ocean circulation inverse model (i.e.
 the circulation component of their model).
-Their carbon export estimate range between 7.3 and 8.6 Pg C year$^{-1}$, depending on the selected export depth (i.e.
+Their carbon export estimates range between 7.3 and 8.6 Pg C year$^{-1}$, depending on the selected export depth (i.e.
 respectively, 100 m and the euphotic depth).
-@Nowicki2022-no added a two-phytoplankton/two-zooplankton food-web model with a migration zooplankton component to the model of @DeVries2017-wu in order to estimate the magnitude of each export pathway of the BCP.
+@Nowicki2022-no added a two-phytoplankton/two-zooplankton food-web model with a migration zooplankton component to the model of @DeVries2017-wu in order to estimate the magnitude of each BCP export pathway.
 In all, @Nowicki2022-no estimated a carbon export of 10 Pg C year$^{-1}$.
@@
 
-![Top of the atmosphere (TOA) radiance is converted into remote sensing reflectance $R_{rs}(\lambda)$, which is the ratio of the solar radiation reflected from the surface of the ocean that penetrates the sea surface at a discrete wavelength $\lambda$.
+![Top of the atmosphere (TOA) radiance is converted into remote sensing reflectance $R_{rs}(\lambda)$, which is the ratio of the radiance leaving the water surface to the downwelling irradiance incident on the sea surface at a given wavelength $\lambda$.
 Using TOA and $R_{rs}(\lambda)$, primary products (yellow circles) can be derived (e.g.
@@
 Those primary products can be used in numerical models (e.g.
-empirical relationships) to compute secondary products (orange and red circles) such as the phytoplankton commmunity composition (PCC), particle size spectra (PSS), phytoplankton carbon concentrations (C$_{phyto})$ and the NPP.
+empirical relationships) to compute secondary products (orange and red circles) such as the phytoplankton commmunity composition (PCC), PSS, phytoplankton carbon concentrations (C$_{phyto})$ and the NPP.
 Models (light blue circles) therefore incorporate primary and secondary products as well as ancillary data (red squares) either from satellite (e.g.
@@
 
+```{=latex}
+\newpage
+```
 ## Observation tools {#sec-observation-tools}
@@
 In the 20th century, the downward POC flux was traditionally measured with sediment traps.
-Those instruments are composed of one or several collectors of sinking material whose design depends on its use and position in the water column (e.g.
+Those instruments are composed of one or several collectors of sinking material whose design depends on their use and position in the water column (e.g.
 moored traps, surface tethered traps, drifting traps) [@Honjo2008-kn].
-They were extensively used [@Honjo2008-kn] though their use has been steadily decreasing since 2004, reaching less than 500 measurements per year since 2010 (@fig-sediment-trap-thorium-yearly-distribution).
+They were extensively used [@Honjo2008-kn] but the number of measurements has been steadily decreasing since 2004, reaching less than 500 measurements per year since 2010 (@fig-sediment-trap-thorium-yearly-distribution).
 
@@
 #| label: fig-sediment-trap-thorium-yearly-distribution
-#| fig-cap: Yearly distibution of POC flux measurements by sediment traps or derived from Thorium 234 deficits. Compilation based on @tbl-poc-from-sediment-traps and @tbl-poc-from-thorium.
+#| fig-cap: Yearly distribution of POC flux measurements by sediment traps or derived from Thorium 234 deficits. Compilation based on @tbl-poc-from-sediment-traps and @tbl-poc-from-thorium.
 
@@
 To estimate the downward POC flux, one therefore needs to quantify the amount of POC collected through a given area over a known time period at a given depth.
-However, the apparent sinking flux measured with sediment traps can be biased due to hydrodynamic issues that could result in over- or under-collection of particles (especially at the surface), swimmers (zooplankton, small fish) that could enter sediment traps and affect the collected material and solubilization of collected material (e.g.
+However, the apparent sinking flux measured with sediment traps can be biased due to hydrodynamic issues that could result in over- or under-collection of particles (especially at the surface) or swimmers (zooplankton, small fish) that could enter sediment traps and affect the collected material and solubilization of collected material (e.g.
 by bacteria) [@Buesseler2007-ql].
@@
 Remote sensing data acquired with ocean color satellites have been used for a long time [@Gordon1980-dm] to study key EZ variables related to carbon export [@Claustre2021-tx] such as phytoplankton biomass [@Robinson2004-cj], phytoplankton communities [@Uitz2015-hd] and particle size distribution [@Kostadinov2009-gf; @Kostadinov2010-bu] to name a few (see also @sec-biogeochem-models and @fig-bgc-models).
-While having a wide spatial coverage, these platforms are limited to the ocean's surface and their measures can be affected by several factors such as the cloud cover or low solar elevation (e.g.
-@Marshall1994-vk).
+While having a wide spatial coverage, these platforms are limited to the ocean's surface and their measures can be affected by several factors such as the cloud cover or low solar elevation [e.g.
+@Marshall1994-vk].
 
@@
 Moorings can be equipped with many sensors (e.g.
-sediment traps or imaging systems to study the type of sinking particles (@sec-BGC-sensors)) to study carbon-related variables (see Table 1 in @Claustre2021-tx).
+sediment traps or imaging systems to study the type of sinking particles to study carbon-related variables [see Table 1 in @Claustre2021-tx]).
 They are therefore ideal platforms for long-term time series (e.g.
@@
 
-Biogeochemical Argo (BGC-Argo) profiling floats (@fig-bgc-float) are probably the most well known to address the BCP.
+Biogeochemical-Argo (BGC-Argo) profiling floats (@fig-bgc-float) are probably the most well known AUVs to study the BCP.
 They are cost-effective, able to monitor biogeochemical processes at a global scale throughout the year and under all weather conditions [@Chai2020-uq].
@@
 Note that the profiling float can be programmed (and reprogrammed during its mission via Iridium transmission) with different configurations, depending on its utilization.
-For the study of the BCP, some floats could be configured to drift at multiple parking depths to study the BGP or to shorter cycle duration to study the VMP.
+For the study of the BCP, some floats could be configured to drift at multiple parking depths to study the BGP or to shorter cycle durations to study the VMP.
 Image taken from @Claustre2020-jo.](image/ObservingCarbonCycle/argo_float_profiling.jpeg){#fig-BGC-Argo-cycle}
@@
 ![Illustration of an augmented BGC-Argo float.
-From top to bottom: GPS & Iridium (two-way communication) antenna, Oxygen optode, CTD (salinity, temperature and depth), pH sensor, radiometer (downwelling photosynthetic available radiation and downwelling irradiance at 380, 412 and 490 nm), fluorometer (Chla and colored dissolved organic matter) and scattering meter ($b_{bp}$), underwater camera UVP6 (particle size spectra and images), transmissometer (particle beam attenuation) and nitrate sensor.
+From top to bottom: GPS & Iridium (two-way communication) antenna, Oxygen optode, CTD (salinity, temperature and depth), pH sensor, radiometer (downwelling photosynthetic available radiation and downwelling irradiance at 380, 412 and 490 nm), fluorometer (Chla and colored dissolved organic matter) and scattering meter ($b_{bp}$), underwater camera UVP6 (PSD and images), transmissometer (particle beam attenuation) and nitrate sensor.
 \@Thomas Boniface.](image/ObservingCarbonCycle/float_picture1.png){#fig-bgc-float width="400"}
@@
 
-The use of transmissometers to characterize POC concentrations was shown by @Bishop1999-es who found that POC concentrations were well correlated with $c_{p}$, independently of ocean environment, season or depth.
-Subsequently, the idea of using a transmissometer as an OST was firstly applied by @Bishop2004-at where the magnitude of carbon export was quantified based on the particles accumulation on the upward-facing window of vertically oriented transmissometers on autonomous floats while they were drifting (i.e.
+The use of transmissometers to characterize POC concentrations was shown by @Bishop1999-es who found that POC concentrations were well correlated with $c_{p}$, regardless of ocean environment, season or depth.
+Subsequently, the idea of using a transmissometer as an OST was firstly applied by @Bishop2004-at where the magnitude of carbon export was quantified based on the particle accumulation on the upward-facing window of vertically oriented transmissometers mounted on autonomous floats while they were drifting (i.e.
 parked at a specific depth).
-This method was later improved by [@Estapa2013-gw] who divided the attenuation signal into a continuous component, that is the regular accumulation of small particles on the transmissometer window and a discontinuous component resulting in positive jumps likely attributed to the deposition of (rarer) large particles (see @fig-estapa-method and @sec-case-study-labrador-sea).
-[@Estapa2017-jr; -@Estapa2023-oz] further calibrated the OST carbon flux proxy using codeployed sediment traps and drifting transmissometers thus allowing the conversion of $c_{p}$ attenuation fluxes (units: m$^{-1}$d$^{-1}$) into carbon fluxes (units: mg C m$^{-2}$d$^{-1}$) using empirical POC-to-attenuance (ATN) ratio were ATN (unitless) is the product of $c_{p}$ and the transmissometer pathlength.
+This method was later improved by @Estapa2013-gw who divided the attenuation signal into a continuous component, that is the regular accumulation of small particles on the transmissometer window and a discontinuous component resulting in positive jumps likely attributed to the deposition of (rarer) large particles (see @fig-estapa-method and @sec-case-study-labrador-sea).
+An OST carbon flux proxy was further calibrated using codeployed sediment traps and drifting transmissometers [@Estapa2017-jr; @Estapa2023-oz] thus allowing the conversion of $c_{p}$ attenuation fluxes (units: m$^{-1}$d$^{-1}$) into carbon fluxes (units: mg C m$^{-2}$d$^{-1}$) using empirical POC-to-attenuance (ATN) ratio were ATN (unitless) is the product of $c_{p}$ and the transmissometer pathlength.
 
@@
 #| label: fig-estapa-method
-#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component, cleaned from spikes and negative jumps (assumed as not related to carbon flux) and a discontinuous (positive jumps, light blue) component.
+#| fig-cap: Application of the method of @Estapa2023-oz on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component, cleaned from spikes and negative jumps (assumed as not related to carbon flux) and a discontinuous (positive jumps, light blue) component.
 Data from BGC-ARGO float WMO 4903634, drifting at 1000 m during its 9$^{th}$ cycle.
@@
 
+```{=latex}
+\newpage
+```
+
 The UVP6 (@fig-UVP6) is a miniaturized version of the UVP5 [@Picheral2021-sx], an imaging system developed to characterize particles (e.g.
@@
 
-The UVP6-LP (hereafter referred to as UVP6) is thus both a particle counter (0.1 mm - 2 mm) and a particle classifier (\> 0.6 mm).
-The former counts particles in different class sizes in order to compute the PSD while the latter extracts pixel-related and geometric features from each image to classify them (i.e.
-assign them a taxonomic group) directly on board using artificial intelligence (@sec-uvp6) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].
-
 ![Illustration of the basic functioning of the UVP6.
@@
 \@Thomas Boniface.](image/ObservingCarbonCycle/UVP6_b.png){#fig-UVP6}
+
+```{=latex}
+\newpage
+```
+
+The UVP6-LP (hereafter referred to as UVP6) is thus both a particle counter (0.1 mm - 2.5 mm) and a particle classifier (\> 0.6 mm).
+The former counts particles in different size classes in order to compute the PSD while the latter extracts pixel-related and geometric features from each image to classify them (i.e.
+assign them a taxonomic group) directly on board using artificial intelligence (@sec-uvp6) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].
 
@@
 Using the PSD, one can compute the number spectrum $n_{i}$ [@Stemmann2004-jv] which can be defined as the normalized concentration of particles in a given size bin $i$.
-Knowing the mass ($m$) and sinking speed ($w$, see @eq-sinking-speed) of aggregates, one can derive the total mass flux ($F$)
+Knowing the mass ($m$) and the sinking speed ($w$, see @eq-sinking-speed) of aggregates, one can derive the total mass flux ($F$)
 
@@
 
-@Guidi2008-vp used POC fluxes derived from sediment traps and UVP PSD data to estimate the $A$ and $b$ values of @eq-carbon-flux-finite by minimizing the following cost function
+@Guidi2008-vp used POC fluxes derived from sediment traps and UVP PSD data to estimate the $A$ and $B$ values of @eq-carbon-flux-finite by minimizing the following cost function
 
 $$
-Cost = \sum_{j = 1}^{n} (log(F_{UVP,j})-log(F_{ST},j))^{2}
+Cost = \sum_{j = 1}^{n} (log(F_{ST,j})-log(F_{UVP},j))^{2}
 $$ {#eq-cost-function}
 
-where $F_{UVP,j}$, $F_{ST,j}$ and $n$ are, respectively, the flux estimate for the $j^{th}$ sediment trap measurement, the flux measured in sediment trap $j$ and the number of paired measurements (i.e.
+where $F_{ST,j}$, $F_{UVP,j}$ and $n$ are, respectively, the carbon flux measured in sediment trap $j$, the carbon flux estimate for the $j^{th}$ UVP PSD and the number of paired measurements (i.e.
 match-ups) for trap-UVP comparisons.
 
-@Guidi2008-vp found an optimal couple of parameters $(A, B) = (12.5, 3.81)$ while @Iversen2010-an followed the same procedure and found $(A, B) = (273, 4.27)$ whereas @Fender2019-ou found $(A, B) = (15.4, 1.05)$.
+@Guidi2008-vp found an optimal couple of parameters $(A, B) = (12.5, 3.81)$ while @Iversen2010-an followed the same procedure and found $(A, B) = (273.8, 4.27)$ whereas @Fender2019-ou found $(A, B) = (15.4, 1.05)$.
 @Iversen2010-an assumed that the difference in the optimal couple of parameters was due to the study site used for the minimization (productive upwelling system versus open ocean in @Guidi2008-vp).
-@Fender2019-ou suggested that a value of $B < 3$ means that either the density or sinking speed of particles is inversely correlated with particles diameter thus they suspected that the flux was dominated by particles in the lower size detection limit of the UVP they were using.
+@Fender2019-ou suggested that a value of $B < 3$ means that either the density or sinking speed of particles is inversely correlated with particle diameter thus they suspected that the flux was dominated by particles in the lower size detection limit of the UVP they were using.
 Other values for $A$ and $B$ are used in the literature [e.g.
@@
 compact fecal pellets, porous aggregates) and that the main driver of the sinking speed is currently subject to debate [@sec-sinking-speed].
-Therefore, the size-dependent model of [@Guidi2008-vp] could be improved by taking into account the composition and nature of particles and by increasing the number of paired measurements between UVP PSD-derived flux estimations and other direct POC flux estimations (e.g.
+Therefore, the size-dependent model of @Guidi2008-vp could be improved by taking into account the composition and nature of particles and by increasing the number of paired measurements between UVP PSD-derived flux estimations and other direct POC flux estimations (e.g.
 sediment traps, $^{234}Th$, OST).
@@
 
-This framework will hopefully address the key points raised so far, i.e., better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and refine carbon export and sequestration estimates [@Boyd2019-ez; @Siegel2022-wd].
+This framework will hopefully address the key points raised so far, i.e.
+better understand the underlying drivers of each component of the BCP, decouple the contribution of each pump to the observed downward carbon flux lengthily attributed to the BGP and refine carbon export and sequestration estimates [@Boyd2019-ez; @Siegel2022-wd].
 
@@
 2.
-Four BGC-Argo floats equipped with both an OST and the UVP6 were deployed for the first time in the Labrador Sea in late May 2022 to study the dynamics of carbon fluxes and particle size distributions.
-In this chapter, I will explore the seasonality of the later and compare how the two sensors relate as two independent measurements of particle and carbon fluxes.
+Four BGC-Argo floats equipped with both an OST and a UVP6 were deployed for the first time in the Labrador Sea in late May 2022 to study the dynamics of carbon fluxes and particle concentrations.
+In this chapter, I will explore the seasonality of the latter and compare how the two sensors relate as two independent measurements of particle and carbon fluxes.
 3.
-Many studies have assumed that carbon sequestration only occurs in the deep ocean (typically 1000 m).
+Many studies have assumed that carbon sequestration only occurs in the deep ocean (typically 1,000 m).
 In @sec-bcp-sequestration, we reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.

### 2023-03-25 uvp6_part.qmd
typo check

@@
 The aim was to be prepared for the UVP6 image data set (not ready at that time) to train the true final classification model.
-To this end, I developed a python package called [UVPec](https://github.com/ecotaxa/uvpec) that is able to create a full classification model based on image folders with labelled particle/zooplankton images and some hyperparameters (e.g.
+To this end, I developed a python package called [UVPec](https://github.com/ecotaxa/uvpec) that is able to create a full classification model based on image folders with labeled particle/zooplankton images and some hyperparameters (e.g.
 learning rate).
@@
 Therefore, using an embedded CNN to classify plankton images or sending images through iridium transmission to be classified in the lab would not be possible for economical (cost of iridium transmission) and technical reasons.
-Typically, a BGC-Argo float samples the water column between the surface and 1000 m, it would therefore have to stay at the surface for a time long enough to transmit all images collected after each profile, consuming energy proportionally to the size of the image set while being subject to biofouling or unpredictable hazards.
+Typically, a BGC-Argo float samples the water column between the surface and 1,000 m, it would therefore have to stay at the surface for a time long enough to transmit all images collected after each profile, consuming energy proportionally to the size of the image set while being subject to biofouling or unpredictable hazards.
 For all these reasons, it was decided to use a features-based classifier for the development of the embedded plankton recognition algorithm on board the UVP6.
@@
 ![UVP5 and UVP6 profile coverage used in this study. UVP5 profiles used in the test set are depicted by green crosses.](image/uvp6/map_of_profiles.png){#fig-map-of-profiles}
+```{=latex}
+\newpage
+```
 
@@
 From it, 55 features were extracted to describe this object in terms of shape (e.g.
-pixel area, eccentricity), gray level and mostly statistical moments (see @tbl-features).
+pixel area, eccentricity), gray levels and statistical moments (see @tbl-features).
 For technical reasons, all UVP6 images with a pixel area below 79 pixels were removed (7% of images).
@@
 #| label: tbl-features
-#| tbl-cap: List of 55 features (abbreviations used in our python package UVPec (https://github.com/ecotaxa/uvpec) and their descriptions) extracted from each object based on the object geometrical characteristics, the pixel gray values histogram and statistical moments.
+#| tbl-cap: List of 55 features (abbreviations used in our python package UVPec (https://github.com/ecotaxa/uvpec) and their descriptions) extracted from each object based on their geometry and gray levels.
 
@@
 
-Validated objects were manually classified into specific taxa belonging to well known biological groups such as Rhizaria, Copepoda or Cnidaria to name a few.
+Validated objects were manually classified into specific taxa belonging to well known biological groups such as *Rhizaria*, *Copepoda* or *Cnidaria* to name a few.
 However, the maximum number of classes (also referred to as labels or taxa) for the ER was fixed to 40 in order to limit the size of the ER model and the volume of data that needs to be transmitted.
@@
 The first level (i.e.
-all labels (\> 40) are retained) served as a baseline for a prior model, without making any hypotheses.
+all labels (\> 40) are retained) served as a baseline for a prior model, without making any assumptions.
 The next level was used as a first aggregation step based on the number of objects in each class and the purity of the class.
@@
 #| label: tbl-classification-report-uvp5-xgboost-cnn
-#| tbl-cap: Precision, recall and f1-score for each class and globally (lightgray background) in the UVP5 model (XGBoost and CNN). The abbreviation avg stands for average.
+#| tbl-cap: Precision, recall and f1-score for each class and globally in the UVP5 model (XGBoost and CNN). The abbreviation avg stands for average.
 
-labels <- c('Acantharea', 'Aulacanthidae', 'Aulosphaeridae', 'Copepoda', 'Trichodesmium', 
-            'artefact', 'bubble', 'darsphere', 'detritus', 'fiber', 'other<living', 'accuracy',
+labels <- c('Acantharia', 'Aulacanthidae', 'Aulosphaeridae', 'Copepoda', 'Trichodesmium', 
+            'artefact', 'bubble', 'darksphere', 'detritus', 'fiber', 'other<living', 'accuracy',
             'macro avg', 'weighted avg', 'living macro avg', 'living weighted avg')
@@
 
-One objective of the ER on board the UVP6 is to be used on AUVs such as BGC-Argo floats to study the vertical migration pump [@Claustre2021-tx].
-Hence, we built the test set with entire UVP5 profiles going at least to 1000 m depth, the typical parking depth of BGC-Argo floats.
+One objective of the ER for the UVP6 is to be used on AUVs such as BGC-Argo floats to study the vertical migration pump [@Claustre2021-tx].
+Hence, we built the test set with entire UVP5 profiles going at least to 1,000 m depth, the typical parking depth of BGC-Argo floats.
 Those profiles were also chosen to be representative of the global ocean (@fig-map-of-profiles).
@@
 $k$) before the model overfits.
-For this, we searched for the minimum value of the multi class log loss (also known as logistic loss or cross-entropy loss) evaluated at each boosting round.
+For this, we searched for the minimum value of the multiclass log loss (also known as cross-entropy loss) evaluated at each boosting round.
 Using that optimal $k$ value, we then trained the ER model with the entire training set.
-Therefore, finding the best model consisted in finding the couple ($\eta$, $D_{max}$) that respected the energy constraints and outputed the best global metrics.
-The latters were determined using a customary grid search (@fig-grid-search) while $\omega$ was chosen empirically (see hereafter).
+Therefore, finding the best model consisted in finding the set of hyperparameters ($k$, $\eta$, $D_{max}$) that respected the energy constraints and returned the best global metrics.
+The latter were determined using a customary grid search (@fig-grid-search) while $\omega$ was chosen empirically (see hereafter).
 
@@
 $$
-class~weight = \left(\frac{size~max}{class~size}\right)^{\omega}
+class~weight = \left(\frac{size~max}{size~class}\right)^{\omega}
 $$ {#eq-class-weight}
 
-where $size~max$ and $class~size$ are, respectively, the size of the biggest class (i.e.
+where $size~max$ and $size~class$ are, respectively, the size of the biggest class (i.e.
 detritus) and the size of each class.
@@
 We decided to select the best $\omega$ among 0 (no weights), 0.25, 0.50 and 0.75 (model very biased towards biological classes).
-In order to decipher between all options, we compared the predicted concentrations of classes for each $\omega$ with the real concentration of each class in different regions of the global ocean.
+In order to decipher between all options, we compared the predicted concentrations of classes for each $\omega$ with the real concentrations of each class in different regions of the global ocean.
 As always in machine learning, we encountered a trade off between precision and recall.
@@
 This is known as transfer learning and it was already applied with success for zooplankton classification in @Orenstein2017-ya.
-Input images have a size of \[224, 224, 3\], the size of the features extractor vector is 1792 followed by a fully connected layer (600 nodes) with a dropout of 0.4 to avoid over-fitting.
+Input images have a size of \[224, 224, 3\], the size of the features extractor vector is 1792 followed by a fully connected layer (600 nodes) with a dropout of 0.4 to avoid overfitting.
 The classification layer at the end of the network also has a dropout of 0.2.
@@
 It is also very frustrating to see that one class (largely) dominates when you are interested in the small ones and that some of them are so scarce that you need to merge them with their siblings (at the very best) on the phylogenetic tree.
-All in all, we found - after multiple trials and errors - 11 classes fulfilling a first compromise between classification performances and ecological importance: 7 biological classes (*Acantharia*, *Aulacanthidae*, *Aulosphaeridae*, *Copepoda*, *Trichodesmium*, darksphere and other\<living) and 4 non-living classes (detritus, bubble, fiber and artefact).
+Overall, we found - after multiple trials and errors - 11 classes fulfilling a first compromise between classification performances and ecological importance: 7 biological classes (*Acantharia*, *Aulacanthidae*, *Aulosphaeridae*, *Copepoda*, *Trichodesmium*, darksphere and other\<living) and 4 non-living classes (detritus, bubble, fiber and artefact).
 
-Once the number of classes is known, hyperparameters can be tuned to find the optimal couple for the classification results (i.e.
+Once the number of classes is known, hyperparameters can be tuned to find the optimal set for the classification results (i.e.
 grid search results shown in @fig-grid-search).
 In our case, the optimal model hyperparameters were 0.2 for $\eta$ and 5 for D$_{max}$ (also with only 1 class of detritus, see Discussion) to have both a good weighted living (or biological) precision and recall and a loading time (i.e.
-loading of the ER model in memory) below 700 ms for energetic reasons.
+loading of the ER model in memory) below 700 ms for energy reasons.
 In that configuration, the UVP5 model had an accuracy of 81%, a macro (i.e.
 label imbalance not taken into account) precision and recall of, respectively, 64% and 69% and a weighted living precision and recall of, respectively, 60% and 65%.
-The CV found an optimal number of 850 boosting rounds before over-fitting and the expected loading time is 382 ms.
+The CV found an optimal number of 850 boosting rounds before overfitting and the expected loading time is 382 ms.
 
 The classification report (CR) gives the precision, the recall, the f1-score (harmonic mean of the precision and recall) and global metrics on the test set.
-As it can be seen in @tbl-classification-report-uvp5-xgboost-cnn, detritus have a recall of 86% thus because detritus are abundant (more than 230.000 in the test set), it results in more than 30.000 detritus wrongly labelled and therefore disseminated in all other classes.
-The confusion matrix (CM) helps to see where those wrongly labelled images are.
+As it can be seen in @tbl-classification-report-uvp5-xgboost-cnn, detritus have a recall of 86% thus because detritus are abundant (more than 230.000 in the test set), it results in more than 30.000 detritus wrongly labeled and therefore disseminated in all other classes.
+The confusion matrix (CM) helps to see where those wrongly labeled images are.
 @fig-confusion-matrix-uvp5 shows that a lot of Copepoda are confused with detritus as well as fiber.
@@
 It was already shown in @Solow2001-jc, using plankton images from the Video Plankton Recorder, that the estimation of taxonomic abundances could be improved by computing the confusion matrix $\pmb{P}$ where $p_{i,j}$ is the probability that a taxon $i$ is classified as taxon $j$.
-If the classifier is perfect, $p_{i,j} = 1$ where $i = j$, the quantification resumes to a simple Classify and Count method (CC, see @Forman2005-eu).
+If the classifier is perfect then $p_{i,j} = 1$ where $i = j$ thus the quantification resumes to a simple Classify and Count method (CC, see @Forman2005-eu).
 However, a perfect (especially in plankton ecology) classifier does not exist.
 Therefore, quantification is a tool to better estimate the abundance of planktonic groups while perfectly knowing that the classifier is biased.
-This quantification method was successively applied in the paper of @Sosik2007-cj on a 2 month data set although it is very likely that P needs to be adapted to the sampling conditions.
+This quantification method was successively applied in the paper of @Sosik2007-cj on a 2 month data set although it is very likely that $\pmb{P}$ needs to be adapted to the sampling conditions.
 This reevaluation of $\pmb{P}$ is due to what is defined as a data set shift where part of the unseen data follows a different distribution than the one from the training data [@Orenstein2020-jl; @Moreno-Torres2012-oq]).
@@
 The aim of building an ER model was intended towards its implementation on the UVP6 for future deployments on BGC-Argo floats (or other profiling AUVs).
-Following the work done with the UVP5 data set, we applied the same method for UVP6 data.
-Thus, we started with the data aggregation step (choices of classes) where, in contrast with UVP5 data, some classes were kept (see @tbl-classification-report-uvp6-xgboost-cnn) because of their ecological role (e.g.
+Following the work done with the UVP5 data set, we applied the same method on UVP6 data.
+Thus, we started with the data aggregation step (choice of classes) where, in contrast with UVP5 data, some classes were kept (see @tbl-classification-report-uvp6-xgboost-cnn) because of their ecological role (e.g.
 small\<*Cnidaria* or tuff, an elongated cyanobacteria belonging to *Trichodesmium*) even though their classification metrics (precision and recall) were not particularly convincing.
@@
 
-```{r, confusion matrix N detritus}
+```{r, confusion matrix N detritus, fig.height=4, fig.width=6}
 #| label: fig-confusion-matrix-final-model
@@
 #| label: tbl-classification-report-uvp6-xgboost-cnn
-#| tbl-cap: Precision, recall and f1-score for each class and globally (lightgray background) of the best UVP5 model trained with XGBoost and a CNN. The abbreviation avg stands for average.
+#| tbl-cap: Precision, recall and f1-score for each class and globally of the best UVP6 model trained with XGBoost and a CNN. The abbreviation avg stands for average.
 
@@
 
+```{=latex}
+\newpage
+```
+
 #### How to use the CR in real applications?
@@
 4 BGC-Argo floats equipped with the ER were deployed in late May 2022 in the southern part of the Labrador Sea at the time of the spring bloom.
-As it can be seen in @fig-bgc-application-uvp6, *Calanoida* are more abundant around the deep chlorophyll maximum (DCM) and they would be observed in very low concentrations even below 1000 m.
+As it can be seen in @fig-bgc-application-uvp6, *Calanoida* are more abundant around the deep chlorophyll maximum (DCM) and they might be observed in very low concentrations even below 1000 m.
 The CR for *Calanoida* shows that the model has a precision of 66% and a recall of 69% for that class.
@@
 For instance, it can be seen that detritus nicely follow the DCM curve.
-In contrast with biological classes where we are looking for patterns of presence or absence, we know that detritus are ubiquitous but their successive abundances could be interesting for flux studies.
+In contrast with biological classes where we are looking for patterns of presence or absence, we know that detritus are ubiquitous but their successive abundances could be interesting for carbon flux studies.
 
@@
 #| label: fig-bgc-application-uvp6
-#| fig-cap : Concentrations of 4 classes (3 biological and detritus) estimated with data acquired from the UVP6 on board the BGC-Argo float (WMO 4903634) on the 15$^{th}$ of June 2022 in the Labrador Sea at latitude 59.27° N and longitude -50.56° E.
-Red points represent positive concentrations.
+#| fig-cap : Concentrations of 4 classes (3 biological and the detritus class) estimated with data acquired from the UVP6 on board the BGC-Argo float (WMO 4903634) on the 15$^{th}$ of June 2022 in the Labrador Sea.
+Red points represent positive concentrations while blue points represent the absence of the specified class.
 Dotted lines represent the depth of the deep chlorophyll maximum.
@@
 chla <- ggplot(tmp, aes(x = pres, y = mean_conc)) + geom_point() + scale_x_reverse() + coord_flip() +
-  xlab('Pressure (dbar)') + ylab(expression('Chlorophyll a (mu/L)')) + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) 
+  xlab('Depth (m)') + ylab(expression('Chlorophyll a (mu/L)')) + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) 
 DCM <- tmp$pres[which.max(tmp$mean_conc)]
@@
 cal <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) + geom_point(colour = "#005CAB") + scale_x_reverse() + coord_flip()+
-  xlab('Pressure (dbar)') + ylab('Calanoida (#/L)') + geom_point(data = test[test$mean_conc>0,], 
+  xlab('Depth (m)') + ylab('Calanoida (#/L)') + geom_point(data = test[test$mean_conc>0,], 
                                                                   aes(x = Pressure_dbar, y = mean_conc), colour = "#E31B23") +
@@
              aes(x = Pressure_dbar, y = mean_conc), colour = "#E31B23") +
-  xlab('Pressure (dbar)') + ylab('detritus (#/L)') + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
+  xlab('Depth (m)') + ylab('detritus (#/L)') + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
   geom_vline(xintercept = DCM, lty = 'dotted') + theme_bw()
@@
 puf <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) +  geom_point(colour = "#005CAB") + scale_x_reverse() + coord_flip()+
-  xlab('Pressure (dbar)') + ylab('puff (#/L)') + geom_point(data = test[test$mean_conc>0,], 
+  xlab('Depth (m)') + ylab('puff (#/L)') + geom_point(data = test[test$mean_conc>0,], 
                                                                           aes(x = Pressure_dbar, y = mean_conc), colour = "#E31B23") +
@@
 tuf <- ggplot(test, aes(x = Pressure_dbar, y = mean_conc)) +  geom_point(colour = "#005CAB") + scale_x_reverse() + coord_flip()+
-  xlab('Pressure (dbar)') + ylab('tuff (#/L)') + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
+  xlab('Depth (m)') + ylab('tuff (#/L)') + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
   geom_point(data = test[test$mean_conc>0,], 
@@
 
+```{=latex}
+\newpage
+```
+
 #### Prospects
@@
 However, it should be noted that not all biological classes are better classified (e.g.
-*Actinopeterygii*, *Appendicularia*, *Chaetognatha*, *Creseis*).
+*Actinopterygii*, *Appendicularia*, *Chaetognatha*, *Creseis*).
 One can also observe that the difference between XGBoost and the CNN trained on UVP5 data on biological classes is less significant than for UVP6 data.
-This is likely explained by the fact that we were able to keep more (small) classes for the UVP6.
-Nevertheless, previous tests on UVP5 data with more classes increased the gap between XGBoost and CNN weighted results therefore the latter recognizes better small classes.
+This is likely explained by the fact that we were able to keep more (smaller) classes for the UVP6.
+Nevertheless, previous tests on UVP5 data with more classes increased the gap between XGBoost and CNN weighted results therefore the latter better recognizes small classes.
 If energy and hardware constraints limit the performances of our current ER model, it is not forbidden to think about light-weight CNNs for future applications on AUVs to better understand the dynamics and the behavior of zooplankton.
@@
 It was clear at the beginning that the use of powerful energy consuming CNNs had to be disregarded and that we needed to build a satisfactory features-based model instead.
-To this aim, we firstly used XGBoost on UVP5 data, a more spatially and temporally representative data set than the recent UVP6 data set to create the data processing pipeline (classes aggregation, hyperparameter tuning, model training, see also @fig-uvp-pipeline) and investigate ways to improve classification scores (quantification, detritus subgroups, label weights).
+To this aim, we firstly used XGBoost on UVP5 data, a more spatially and temporally representative data set than the recent UVP6 data set to create the data processing pipeline (class aggregation, hyperparameter tuning, model training, see also @fig-uvp-pipeline) and investigate ways to improve classification scores (quantification, detritus subgroups, label weights).
 In order to see how the model could be improved, we compared the XGBoost model with results from a CNN not limited by energy.

### 2023-03-26 article3.qmd
finalize typo check

@@
 @Omand2015-mf; @DallOlmo2016-ex; @Llort2018-pa; @Lacour2019-id; @Haentjens2020-dc; @Stukel2017-lc].
-Still, we are missing processes to close the carbon budget [@Giering2014-tj; @Boyd2019-ez] and unkowns remain in the under-sampled mesopelagic (i.e.
+Still, we are missing processes to close the carbon budget [@Giering2014-tj; @Boyd2019-ez] and unknowns remain in the under-sampled mesopelagic (i.e.
 twilight) zone [@Martin2020-wa].
 In order to fill the observational gaps, the Biogeochemical-Argo (BGC-Argo) network aims to develop an operational global array of 1000 profiling floats measuring six core variables from the surface to 2000 m [@Claustre2020-jo]: oxygen, nitrate, pH, chlorophyll a (Chla), suspended particles and downwelling irradiance.
-In addition to the six core variables, new sensors have been developped to target specific BCP-related questions [@Claustre2020-jo; @Claustre2021-tx; @Chai2020-uq].
+In addition to the six core variables, new sensors have been developed to target specific BCP-related questions [@Claustre2020-jo; @Claustre2021-tx; @Chai2020-uq].
 In this study, we will focus on BGC-Argo floats equipped with two additional sensors: the Underwater Vision Profiler (UVP6) [@Picheral2021-sx] and a transmissometer used as an optical sediment trap (OST).
@@
 The UVP6 is a portable version of the UVP5 [@Picheral2010-to], specifically designed for AUVs [@Picheral2021-sx].
-It consists of an underwater camera that counts particles and images objects in, respectively, the 0.1 - 2.5 mm and 0.6 - 2.5 cm size range.
+It consists of an underwater camera that counts particles and takes images of objects in, respectively, the 0.1 - 2.5 mm and 0.6 - 2.5 cm size range.
 In contrast with UVP5 deployments on cabled platforms, UVP6 deployed on long-term drifting AUVs (e.g.
@@
 According to @Sverdrup1953-ki, a bloom is therefore triggered when the surface mixed layer depth (MLD) is shallower than this critical depth resulting in a positive phytoplankton growth rate.
-However, this hypothesis was challenged by @Behrenfeld2010-my who suggested that the initiation of the NA blooms begins in winter with a slow phytoplankton biomass increase over the following months (i.e.
+However, this hypothesis was challenged by @Behrenfeld2010-my who suggested that the initiation of the NA bloom begins in winter with a slow phytoplankton biomass increase over the following months (i.e.
 'Dilution-Recoupling' hypothesis).
@@
 The northern part is characterized by a short and potentially more intense spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June.
-@Lacour2015-da concluded that the concurrent influence of surface light and mixing was the first-order mechanism controlling the onset of the spring bloom in both bioregions.
+@Lacour2015-da concluded that the concurrent influence of surface light and mixing was the first order mechanism controlling the onset of the spring bloom in both bioregions.
 This is coherent with the results of @Wu2008-vw and @Frajka-Williams2010-ph who found that the early bloom in the northern part is due to a very shallow mixed layer associated with low-salinity waters (i.e.
@@
 In late May 2022, two pairs of BGC-Argo floats were deployed in the southern part of the LS at the time of the spring bloom.
-While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study the downward POC flux and its attenuation derived from beam attenuance [@Estapa2023-oz] and to study particle size distributions (PSD) and concentrations with the UVP6.
-
-The goals of this study are threefold: 1) the comparison of the UVP6 and OST as two independent measurements of downward carbon fluxes and particle abundance on BGC-Argo floats; 2) the exploration of the seasonality (bloom and non-bloom periods) of downward POC fluxes and particle abundance in the LS; 3) the evaluation of the new drifting configuration and how it could be improved to study specific BCP-related questions.
+While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study the downward POC flux and its attenuation derived from beam attenuance [@Estapa2023-oz] and to study particle size distributions (PSDs) and concentrations with the UVP6.
+
+The goals of this study are threefold: 1) the comparison of the UVP6 and the OST as two independent measurements of downward carbon fluxes and particle abundance on BGC-Argo floats; 2) the exploration of the seasonality (bloom and non-bloom periods) of downward POC fluxes and particle abundance in the LS; 3) the evaluation of the new drifting configuration and how it could be improved to study specific BCP-related questions.
 
@@
 turbulence or zooplankton consumption of accumulated material).
-Afterwards, we computed the weighted average slope ($slope_{w}$) of each segment to derive the small particles flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).
+Afterwards, we computed the weighted average slope ($slope_{w}$) of each segment to derive the small particle flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).
 
@@
 Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: 10.48670/moi-00281) with a spatial resolution of 4 km.
-Mixed layer depth (MLD) derived from satellite observations were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052) with a spatial resolution of 0.25°.
+MLD derived from satellite observations were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052) with a spatial resolution of 0.25°.
 
@@
 In contrast with small classes (referred to here as below 512 $\mu$m), some particles, especially above 1.5 mm, have almost entirely disappeared from the water column by September.
-However, the rapid increase and decrease of particles on the entire water column around July-August for floats 6904241 and 1902578 (still in the same zone at that time) could suggest an export event.
+However, the rapid increase and decrease of particles on the entire water column around July-August for floats 6904241 and 1902578 (still in the same geographical zone at that time) could suggest an export event.
 
@@
 At 500 m, $F_{large}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_largef, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to February.
-At 1000 m, $F_{large}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards March, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$.
+At 1000 m, $F_{large}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards March, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ and an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$.
 
@@
 This pattern could result from several causes.
-It could either be the signature of a subduction event [@Llort2018-pa] either the end of a slow sinking export (i.e latent dynamic of the system) or a sign of fragmentation processes by zooplankton (i.e.
+It could either be the signature of a subduction event [@Llort2018-pa] either the end of a slow sinking export (i.e latent dynamics of the system) or a sign of fragmentation processes by zooplankton (i.e.
 zooplankton feeding).
@@
 ```
-### Comparing OST and UVP data
+### Comparing OST and UVP6 data
 
@@
 
-For the first time, BGC-Argo floats equipped with both an OST and the UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx].
+For the first time, BGC-Argo floats equipped with both an OST and the UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instruments approach [@Claustre2021-tx].
 The quantification of the POC flux using the OST method [@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk; @Estapa2023-oz] is based on an empirical POC:ATN ratio [@Estapa2023-oz] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:
@@
 $$
-F = \sum_{i}^{N}C_{i}Ad_{i}^{b}
+F = \sum_{i}^{N}C_{i}Ad_{i}^{B}
 $$ {#eq-guidi-article3}
 
-where $C_{i}$, A and b represent, respectively, the concentration of particles (#/L) for the $i^{th}$ class of mean diameter $d_{i}$ (in mm) and 2 constants with $b$ being related to the fractal dimension ($D$) of an aggregate by $D = (b +1)/2$ [@Guidi2008-vp].
-
-The use of @eq-guidi-article3 to derive carbon fluxes from PSD was firstly implemented by @Guidi2008-vp who used a minimization procedure between sediment traps data and estimated fluxes using PSD from several previous versions of the UVP to find the optimal value of $A$ and $b$ (i.e.
+where $C_{i}$, A and B represent, respectively, the concentration of particles (#/L) for the $i^{th}$ class of mean diameter $d_{i}$ (in mm) and 2 constants with $B$ being related to the fractal dimension ($D$) of an aggregate by $D = (B +1)/2$ [@Guidi2008-vp].
+
+The use of @eq-guidi-article3 to derive carbon fluxes from PSD was firstly implemented by @Guidi2008-vp who used a minimization procedure between sediment traps data and estimated fluxes using PSD from several previous versions of the UVP to find the optimal value of $A$ and $B$ (i.e.
 12.5 and 3.81).
 @Iversen2010-an and @Fender2019-ou followed the same procedure and concluded that the couple of parameters found by [@Guidi2008-vp] could not be used globally at the expense of greatly over- or underestimating POC fluxes, depending on the sampling region.
-In addition, the formulation of such a model implies that $A$ and $b$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt].
+In addition, the formulation of such a model implies that $A$ and $B$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt].
 @eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2023-qj].
@@
 #| label: fig-comparison-ost-uvp
-#| fig-cap: Comparison of total carbon fluxes (F$_{small}$ + F$_{large}$) derived from the OST (black) and the UVP6 using the A and b values of @Guidi2008-vp (red) and @Iversen2010-an (yellow) in @eq-guidi-article3.
+#| fig-cap: Comparison of total carbon fluxes (F$_{small}$ + F$_{large}$) derived from the OST (black) and the UVP6 using the $A$ and $B$ values of @Guidi2008-vp (red) and @Iversen2010-an (yellow) in @eq-guidi-article3.
 OST data from float 6904241 stops in November due to a transmissometer failure.
@@
 @fig-comparison-ost-uvp shows the comparison between the total carbon flux measured by the OST ($F_{total} = F_{small} + F_{large}$) and the UVP6 (median of all UVP measurements while the float was drifting) at all drifting depths.
-Overall, carbon fluxes estimated with the $A$ and $b$ coefficients of @Guidi2008-vp are lower than the ones measured by the OST whereas the same estimation using the coefficient of @Iversen2010-an generally overestimates the OST at 200 m but is of the same order at 500 m and 1000 m.
+Overall, carbon fluxes estimated with the $A$ and $B$ coefficients of @Guidi2008-vp are lower than the ones measured by the OST whereas the same estimation using the coefficient of @Iversen2010-an generally overestimates the OST at 200 m but is of the same order at 500 m and 1000 m.
 Without additional measurements (e.g.
@@
 Similarly, to address the question of the vertical migration, a 10-day temporal resolution is way to low because plankton images are rare compared to detritus so the sampling volume should be greatly increased to have a better estimation of the presence (or absence) of zooplankton.
-Finally, it would perhaps be interesting to increase the time spent at 200 m (to 3 days) and decrease the time spent at 1000 m to 3 days instead of 5) where the probability of capturing big particles is lower than at 200 m.
+Finally, it would perhaps be interesting to increase the time spent at 200 m (to 3 days) and decrease the time spent at 1000 m (to 3 days instead of 5) where the probability of capturing big particles is lower than at 200 m.
 There could also be UVP6 and OST measurements closer in time at all drifting depths to have a better comparison of attenuation slopes/jumps and biovolumes.

### 2023-03-26 bcp_sequestration.qmd
update
finalize typo check

@@
-# Revisiting the BCP with the CONVERSE framework {#bcp-sequestration}
+# Revisiting the BCP with the CONVERSE framework {#sec-bcp-sequestration}
 
@@
 1 year and a half later, hundreds of emails, one promotion to first author, one negative review and one strong rebuttal, we are still on the road to publication.
-Here below is the revised version of our paper submitted in Nature Geoscience about the reassessment of the global sequestration (\> 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
+Here below is the revised version of our paper submitted in Nature Geoscience about the reassessment of the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
 

### 2023-03-26 discussion.qmd
finalize typo check

@@
 Both functions are based on the softmax normalization which normalizes the classification score of each class using an exponential function (@eq-softmax).
-`multi:softmax` returns the class of highest probability whereas `multi:softprob` outputs a class probability array, which allows the computation of custom objective functions.
+`multi:softmax` returns the class of highest probability whereas `multi:softprob` returns a class probability array, which allows the computation of custom objective functions.
 
@@
 #| label: fig-trust-index
-#| fig-cap: Percentage of error captured as a function of rank for four metrics (max, softmax, diff$_{max}$ and diff$_{softmax}$) on six UVP5 classes (4 biological, 2 non biological).
+#| fig-cap: Percentage of error captured as a function of rank for four metrics (max, softmax, diff$_{max}$ and diff$_{softmax}$) on six UVP5 classes (4 biological, 2 non-biological).
 The number of objects predicted in each class is provided in the title of each subplot.
@@
 It was therefore necessary to merge classes based on morphological and ecological attributes to decrease that number in order to get to $level_{1}$ that had 38 classes (reminder, it is currently limited at 40).
-From $level_{1}$, we thus experimented a lot of different prediction models with a hyperparameters gridsearch in order to find both the right set of hyperparameters and the right number of classes.
+From $level_{1}$, we thus experimented a lot of different prediction models with a hyperparameters grid search in order to find both the right set of hyperparameters and the right number of classes.
 Still, some $level_{1}$ classes had very low precision and recall scores because they were confused with some specific taxonomic groups hence we merged them accordingly with those groups.
@@
 @tbl-stats-uvp6 shows that the median number of objects detected in 3.5 m$^{3}$ is roughly 300.
-In addition, it is usual to have less than 20% of living objects hence the vast majority of images are labelled as detritus.
+In addition, it is usual to have less than 20% of living objects hence the vast majority of images are labeled as detritus.
 The latter are ubiquitous in the ocean and display a wide variety of shapes, sizes and textures (@fig-detritus).
-In @sec-improve-classif, I showed that using several classes of detritus did not improve the classification model and that it actually made it worse because the N-detritus classes showed a high level of confusion between each other.
+In @sec-improve-classif, I showed that using several classes of detritus did not improve the classification model and that it actually made it worse because the different detritus classes showed a high level of confusion between each other.
 
@@
 
-![Examples of UVP objects labelled as detritus. The scale bar is the same for all (2 mm).](image/detritus/moz_detritus.jpg){#fig-detritus}
+![Examples of UVP objects labeled as detritus. The scale bar is the same for all (2 mm).](image/detritus/moz_detritus.jpg){#fig-detritus}
 
-![Examples of UVP objects labelled as living material.
+![Examples of UVP objects labeled as living material.
 The scale bar is the same for all (2 mm).
@@
 This python package has been designed to be as user-friendly as possible in order to be accessible to the largest number of users.
-Basically, the user only needs to create folders (40 maximum because the maximum number of classes is fixed at 40) of labelled UVP6 images and specify some hyperparameters (e.g.
+Basically, the user only needs to create folders (40 maximum because the maximum number of classes is fixed at 40) of labeled UVP6 images and specify some hyperparameters (e.g.
 learning rate) to train a classification model.
@@
 31% of *Calanoida* are non detected because they are classified in another class) but 34% of images classified as *Calanoida* are false positives and should be placed in another class.
-In that case, the observed concentrations can therefore be resumed as the real concentration.
-*Chaetognatha* have a precision and recall of, respectively, 80% and 62% therefore when a *Chaetognaha* is detected, the model is confident (20% of false positives) but it misses 38% of them so we might have more *Chaetognatha* than measured by the UVP6.
-However, One must also look at the concentration values because it is likely that the 2 points measured on the eastern part of the Keguerlen Plateau are misclassified objects.
-We could therefore conclude to the absence *Chaetognatha* on the eastern part while their presence is likely on the western part because, on the one hand, this class has a high precision and, on the other hand, they are observed on successive short-lapsed vertical profiles and between 0-500 m where they likely feed on *Calanoida*.
+In that case, the observed concentrations can therefore be assumed as the real concentration.
+*Chaetognatha* have a precision and recall of, respectively, 80% and 62% therefore when a *Chaetognatha* is detected, the model is confident (20% of false positives) but it misses 38% of them so we might have more *Chaetognatha* than measured by the UVP6.
+However, One must also look at the concentration values because it is likely that the 2 points measured on the eastern part of the Kerguelen Plateau are misclassified objects.
+We could therefore conclude to the absence of *Chaetognatha* on the eastern part while their presence is likely on the western part because, on the one hand, this class has a high precision and, on the other hand, they are observed on successive short-lapsed vertical profiles and between 0-500 m where they likely feed on *Calanoida*.
 
@@
 However, the field of computer vision is constantly evolving with the arrival of new technologies that could help improve the automatic classification of zooplankton images.
-Transformers [@Vaswani2017-il], also referred to as Foundation Models in @Bommasani2021-kb are the latest cutting-edge tool in the field of computer vision and natural language processing [e.g.
+Transformers [@Vaswani2017-il], also referred to as Foundation Models in @Bommasani2021-kb are the latest cutting-edge tools in the field of computer vision and natural language processing [e.g.
 @Liu2021-jq; @Li2022-qt].
@@
 There is a class called "other\<living" in the UVP6 classification model.
-I did not mention earlier than we were forced to place copepods (except *Calanoida*) in that class.
+I did not mention earlier that we were forced to place copepods (except *Calanoida*) in that class.
 Indeed, *Calanoida* are a "large" species of copepods compared to other copepods found in the Mediterranean Sea that are smaller.
@@
 
-The good news is that the next generation of UVP is already in the making with the UVP6m (m for micro) which will be able to look at smaller objects (particle counter between 10 and 200 $\mu$m and plankton/particle imaging \> 100 $\mu$m).
+The good news is that the next generation of UVPs is already in the making with the UVP6m (m for micro) which will be able to look at smaller objects (particle counter between 10 and 200 $\mu$m and plankton/particle imaging \> 100 $\mu$m).
 The UVP6m will be more compact than the UVP6 but the principles will be the same and both instruments will share 5 size classes for intercalibration.
@@
 Before working on the CONVERSE framework, I was trying to find a global relationship (I also tried to find a regional one, without more success) to derive carbon fluxes from UVP5 PSDs.
-The idea was to reproduce the method of @Guidi2008-vp on the newly quality controlled UVP5 PSD data set [@Kiko2022-sk].
+The idea was to reproduce the method of @Guidi2008-vp on the newly quality controlled UVP5 PSD data set of @Kiko2022-sk.
 For this, I needed to build a global data set of POC measurements derived from sediment traps and $^{234}Th$ (see @tbl-poc-from-sediment-traps, @tbl-poc-from-thorium and [here](https://fricour.shinyapps.io/carbon_fluxes_app/) for a more interactive view).
@@
 
-Instead of directly predicting carbon fluxes using environmental variable, we could focus on predicting PSD slope or PSD.
+Instead of directly predicting carbon fluxes using environmental variables, we could focus on predicting PSD slopes or PSDs.
 In a recent paper, @Clements2022-oq used physical and biological variables (see their Table 1) to predict global maps of PSD slope and biovolume at the base of the EZ using UVP5 data and a random forest algorithm.
@@
 aggregation/disaggregation processes).
-The study of @Denvil-Sommer2022-oi used both random forest and XGBoost models to predict small and large POC concentrations using geographical variables (depth, latitude and longitude), physical variables (incident light, MLD and temperature), chemical (phosphate and nitrate) and biological variables (Chla and 12 Plankton Functional Types (PFTs) such as diatoms or microzooplankton).
+The study of @Denvil-Sommer2022-oi used both random forest and XGBoost models to predict small and large POC concentrations using geographical (depth, latitude and longitude), physical (incident light, MLD and temperature), chemical (phosphate and nitrate) and biological variables (Chla and 12 Plankton Functional Types (PFTs) such as diatoms or microzooplankton).
 @Denvil-Sommer2022-oi showed that for small POC, the most important predictors are the depth level, temperature, microzooplankton and phosphate while the large POC is mostly driven by depth, temperature, MLD, microzooplankton, phaeocystis (another PFT), phosphate and Chla averaged over the MLD.
@@
 
-The embedded classification for AUVs contributes to a better comprehension and quantification of the vertical migration pump.
+The embedded classification for AUVs will contribute to a better comprehension and quantification of the vertical migration pump.
 So far, BGC-Argo floats equipped with the UVP6 and the classification model have been deployed in some oceanic regions (e.g.
@@
 In addition to the particle (\> 0.6 mm) identification, the UVP6 also sends their mean area and gray levels which will hopefully contribute to better characterize their morphological traits and therefore better predict their sinking rates (cfr.
-size versus excess density).
+size versus excess density debate).
 The latter are important to estimate the downward carbon flux and the sequestration time.

### 2023-03-26 index.qmd
finalize typo check

@@
-# Foreword {.unnumbered}
+# Summary {.unnumbered}
+
+**Towards a new insight of the carbon transport in the global ocean**
+
+The ocean is known to play a key role in the carbon cycle.
+Without it, atmospheric CO$_{2}$ levels would be much higher than what they are today thanks to the presence of carbon pumps that maintain a gradient of dissolved inorganic carbon (DIC) between the surface and the deep ocean.
+The biological carbon pump (BCP) is primarily responsible for this gradient.
+It consists in a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis in sunlit surface waters and then transported to the ocean interior and possibly the sediment where it will be sequestered from the atmosphere for millions of years.
+The BCP was long thought as solely the gravitational settling of particulate organic carbon (POC).
+However, a new paradigm for the BCP has recently been defined in which physically and biologically mediated particle injection pumps have been added to the original definition.
+Physically mediated particle injection pumps provide a pathway to better understand the transport of dissolved organic carbon (DOC) whereas biologically mediated particle injection pumps focus on the transport of POC by vertically migrating animals, either daily or seasonally.
+Therefore, a better understanding of these processes could help bridge the gap between carbon leaving the surface and carbon demand in the ocean interior.
+To address this new paradigm, this work will benefit from the advent of recent sensors that equip a new generation of Biogeochemical-Argo floats (BGC-Argo).
+The first part focuses on the development of an embedded zooplankton classification model for the Underwater Vision Profiler 6 (UVP6) under strict technical and energy constraints.
+The second part studies particle and carbon fluxes in the Labrador Sea using BGC-Argo floats equipped for the first time with the UVP6 and an optical sediment trap (OST), providing two independent measurements of sinking particles.
+The last part consists in revisiting the BCP using a new framework called CONVERSE for Continuous Vertical Sequestration.
+With this new approach, we re-evaluate the total carbon sequestered from the atmosphere (\> 100 years) by the BCP and its transport pathways on the entire water column, in contrast to the carbon sequestration typically assumed below a fixed reference depth.
+
+**Keywords** Carbon pumps -- Underwater Vision Profiler 6 -- BGC-Argo floats -- Machine learning
+
+```{=latex}
+\newpage
+```
+**Vers une meilleur compréhension du transport du carbone dans l'océan global**
+
+L'océan est connu pour jouer un rôle clé dans le cycle du carbone.
+Sans lui, les niveaux de CO$_{2}$ atmosphérique seraient bien plus élevés qu'aujourd'hui grâce à la présence de pompes à carbone qui maintiennent un gradient de carbone inorganique dissous (DIC) entre la surface et le fond de l'océan.
+La pompe à carbone biologique (BCP) est principalement responsable de ce gradient.
+Elle consiste en une série de processus océaniques au cours desquels le carbone inorganique est converti en matière organique via la photosynthèse dans les eaux de surface, puis transporté vers l'intérieur de l'océan et éventuellement les sédiments où il sera séquestré par rapport à l'atmosphère pour des millions d'années.
+La BCP était longtemps considérée comme étant uniquement la déposition gravitationnelle de carbone organique particulaire (POC).
+Cependant, un nouveau paradigme pour la BCP a récemment été défini dans lequel des pompes d'origine physique et biologique d'injection de particules ont été ajoutées à la définition originale.
+Les pompes physiques d'injection de particules fournissent un moyen de mieux comprendre le transport de carbone organique dissous (DOC), tandis que les pompes biologique d'injection de particules se concentrent sur le transport de POC par des animaux migrant verticalement, quotidiennement ou saisonnièrement.
+Par conséquent, une meilleure compréhension de ces processus pourrait aider à combler l'écart entre le carbone quittant la surface et la demande de carbone dans l'océan profond.
+Pour aborder ce nouveau paradigme, ce travail bénéficiera de l'arrivée de capteurs récents équipant une nouvelle génération de flotteurs Biogéochimiques-Argo (BGC-Argo).
+La première partie se concentre sur le développement d'un modèle embarqué de classification de zooplancton pour l'Underwater Vision Profiler 6 (UVP6) avec des contraintes techniques et énergétiques strictes.
+La deuxième partie étudie les flux de particules et de carbone dans la mer du Labrador en utilisant des flotteurs BGC-Argo équipés pour la première fois de l'UVP6 et d'un piège à sédiments optique (OST), fournissant deux mesures indépendantes des particules.
+La dernière partie consiste à revisiter la BCP en utilisant un nouveau cadre appelé CONVERSE qui fait référence à la séquestration verticale continue du carbone dans la colonne d'eau.
+Avec cette nouvelle approche, nous réévaluons le carbone total séquestré par rapport à l'atmosphère (\> 100 ans) par la BCP et ses voies de transport sur toute la colonne d'eau, par opposition à la séquestration de carbone généralement supposée en-dessous d'une profondeur de référence fixe.
+
+**Mots clés** Pompes à carbone -- Underwater Vision Profiler 6 -- Flotteurs BGC-Argo -- Machine learning
+
+```{=latex}
+\newpage
+```
+
+<!-- # Foreword {.unnumbered}
 
 As of 2022, it has now become obvious to the general public that the Earth's climate is changing.
-Massive flood events, repeated heatwaves, never-ending droughts, disappearing glaciers, biodiversity collapse, … Reports on weather anomalies were already published in the 1960's, the Intergovernmental Panel on Climate Change (IPCC) was created in 1988, the Kyoto Protocol was adopted in 1997, An Inconvenient Truth was released in 2006, the Paris Agreement was signed in 2015 and still, what has really changed?
+Massive flood events, repeated heatwaves, never-ending droughts, disappearing glaciers, biodiversity collapse, ...
+Reports on weather anomalies were already published in the 1960's, the Intergovernmental Panel on Climate Change (IPCC) was created in 1988, the Kyoto Protocol was adopted in 1997, An Inconvenient Truth was released in 2006, the Paris Agreement was signed in 2015 and still, what has really changed?
 
@@
 
-Although I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured that there is no fake news in this work. I hope you will enjoy it.
+Although I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured that there is no fake news in this work. I hope you will enjoy it. -->

### 2023-03-26 merci.qmd
typo check
finalize typo check

@@
-# Acknowledgments
+# Acknowledgments {.unnumbered}
 
-Thanks to the members of the jury for taking the time to read and evaluate this three and a half year work. I am looking forward to debate it during the defense.
+I would like to thank the members of the jury for taking the time to read and evaluate this three and a half year work. I look forward to discussing it during the defense.
 
-Cette thèse a été particulière à bien des égards, avec son lot de positifs et de négatifs mais surtout avec son lot de rebondissements qui m'ont permis d'affûter mes connaissances, ma résilience et je l'espère ma tolérance.
+Merci aux membres du comité de thèse pour leur temps et leurs retours sur l'évolution du projet.
+
+Cette thèse a été particulière à bien des égards, avec son lot de points positifs et négatifs mais surtout avec son lot de rebondissements qui m'ont permis d'affûter mes connaissances, ma résilience et je l'espère ma tolérance.
 Pour cela, je tiens à remercier chaleureusement toutes les personnes y ayant contribué de près ou de loin.
@@
 
-Merci également aux personnes que j'ai pu cotoyer à COMPLEX comme chez OMTAB depuis mon passage en stage en 2018 et les multiples allers-retours Belgique-France.
+Merci également aux personnes que j'ai pu côtoyer à COMPLEX comme chez OMTAB depuis mon passage en stage en 2018 et les multiples allers-retours Belgique-France.
 
@@
 Petite dédicace à Ju K qui, à force d'avoir joué avec le diable en subissant l'enfer des pourcentages, va littéralement en devenir un.
-Vivement 2024 pour casser la baraque.
+Vivement 2024 pour casser la baraque!
 

### 2023-03-27 uvp6_part.qmd
Revue du paragraphe de préambule

@@
 Building the zooplankton classification model for the UVP6 was the first task of my PhD.
-To be honest, I knew nothing about machine learning before I started this so I spent the first three weeks reading a nice machine learning book in the corner of my team leader's office.
-To be even more honest, I basically knew nothing about zooplankton either.
-I think it is time to admit it.
+I knew nothing about machine learning before I started, so I spent the first three weeks reading a machine learning book to get familiar with the concepts.
+To be even more honest, I also knew nothing about zooplankton either, so there was something else to learn.
 
-In machine learning, the no free lunch theorem [@Wolpert1996-hw] states that if you do not make any assumption about your data, then there is no model that is a priori guaranteed to work better [@Geron2019-is].
-In other words, it means that you need to get your hands dirty to find a model that works.
-The first step consisted in developing a full data processing pipeline using UVP5 images.
-The aim was to be prepared for the UVP6 image data set (not ready at that time) to train the true final classification model.
+In machine learning, the "no free lunch theorem" [@Wolpert1996-hw] states that if you do not make any assumption about your data, then there is no model that is a priori guaranteed to work better [@Geron2019-is].
+In other words, it means that you need to get to know your data and your hands dirty to find a model that works.
+The first step consisted in developing a full data processing pipeline for UVP5 images.
+The aim was to be prepared for the UVP6 image data set (which was not ready at that time) to train the true, final, classification model.
 To this end, I developed a python package called [UVPec](https://github.com/ecotaxa/uvpec) that is able to create a full classification model based on image folders with labeled particle/zooplankton images and some hyperparameters (e.g.
 learning rate).
-It really made my life easier when I got the final UVP6 data set and it saved me a lot of time.
+It made my life easier when I got the final UVP6 data set and should serve other future users in the same way.
 
-In the end, I am rather satisfied with the results.
-Classifying zooplankton is already difficult for human operators (at least, for me) so building a model to automatically classify them is one challenge but adding technical and energy constraints to it was the ice on the cake.
+Classifying zooplankton on UVP images is difficult for human operators so building a model to automatically classify them is already a challenge, but adding technical and energy constraints to it was the icing on the cake.
 Below is the first draft of a publication on the classification model.

### 2023-03-27 uvp6_part.qmd
Abstract

@@
 Millions of images of oceanic particles have been acquired with ship-based imaging instruments.
-The task of labeling images is time consuming but necessary to train classification models using artificial intelligence algorithms.
-Autonomous Underwater Vehicles (AUVs) such as Biogeochemical-Argo floats (BGC-Argo) have revolutionized our understanding of oceanic systems but they are flawed by their battery life and by the time they can spend at the ocean surface to transmit their data.
-In this paper, we show how we developed the first zooplankton classification model for AUVs using the Underwater Vision Profiler 6 (UVP6).
-To this end, we used a features-based classifier called eXtreme Gradient Boosting (XGBoost) adapted to specific energy constraints imposed by profiling floats and we compared its performance with a state-of-the-art convolutional neural network as if energy was not a limiting factor.
-In addition, we show how we dealt with a highly imbalanced data set towards marine snow (\> 80% of detritus) and what were the compromises to improve the classification scores of biological classes.
-Finally, we show the classification report (precision and recall for each label) of the final classification model and use it to analyse some preliminary results from BGC-Argo floats deployed in the Labrador Sea.
+The identification of objects on images is time consuming for human operators but can be accelerated using artificial intelligence algorithms.
+Autonomous Underwater Vehicles (AUVs) such as Biogeochemical-Argo floats (BGC-Argo) have revolutionized our understanding of oceanic systems but have limited battery life and must spend as little time as possible at the ocean surface to transmit their data, which means that images cannot be transmitted.
+In this paper, we describe an image classification model embedded in the Underwater Vision Profiler 6 (UVP6) that is appropriate to be run on the limited power and bandwidth of AUVs.
+To this end, we computed simple global features of images of objects of interest and gradient boosted trees as a classifier.
+We compared its performance with a state-of-the-art convolutional neural network, which is much more energy demanding.
+In addition, we show how we dealt with class imbalance (\> 80% of images are marine snow) and what were the compromises to improve the classification scores of biological classes.
+Finally, we show the classification report (precision and recall for each label) of the final classification model and use it to analyze preliminary results from BGC-Argo floats deployed in the Labrador Sea.
 

### 2023-03-27 uvp6_part.qmd
Intro

@@
 
-Inside the ocean, a cohort of diverse organisms known as zooplankton thrive.
-Those tiny beasts are so important for our oceans [@Steinberg2017-ri] that we decided to take pictures of them [@Picheral2010-to; @Gorsky2010-ca; @Cowen2008-wq; @Benfield2007-lt; @Lombard2019-ld].
-Nonetheless, the amount of images taken by those imaging instruments has become so huge that it is now a massive work to classify them by hand.
+Inside the ocean, zooplankton constitutes a cohort of very diverse organisms.
+They are important enough for the functioning of our oceans [@Steinberg2017-ri] to motivate the development of a large number of instruments dedicated to taking pictures of them [@Picheral2010-to; @Gorsky2010-ca; @Cowen2008-wq; @Benfield2007-lt; @Lombard2019-ld].
+The amount of images taken by those instruments has quickly become so large that it is now unfeasible to classify them manually, through the observation of human operators.
 Fortunately, powerful machine learning techniques have been implemented to tackle the issue of automatic plankton image classification [@Luo2018-oo].
@@
 However, it remains a fierce challenge.
-Within a single taxon, one can find organisms with very different sizes and shapes, organisms with significant morphological differences acquired during their ontogenetic development, partial organisms (e.g.
-siphonophores) or organisms with various orientations, hindering the automatic classification process [@Benfield2007-lt].
-In addition to this global heterogeneity, the great majority of images are not composed of zooplankton.
-Instead, non-living particles (mostly marine snow) steal the show, producing a highly imbalanced data set towards the latter.
+Within a single taxon, individuals with very different sizes and shapes coexist, due to significant morphological changes during ontogeny, imaging of portions of organisms only (e.g.
+siphonophores) or organisms taken from various angles.
+This variability hinders the performance of automatic classification [@Benfield2007-lt].
+In addition to this global heterogeneity, for in situ instruments, the large majority of images are not of zooplankton: non-living particles (mostly marine snow) largely dominate, resulting in highly imbalanced data sets.
 
-The discussion around automatic plankton classification is a key topic in oceanography and the development of convolutional neural networks (CNNs) within that frame has already lead to promising results [@Luo2018-oo; @Cui2018-nt; @Dai2017-ex; @Campbell2020-cq; @Schroder2019-um; @Bochinski2019-af; @Lee2016-jl; @Cheng2019-ar; @Guo2021-xi; @Ellen2019-nh; @Orenstein2017-ya].
-Nevertheless, real-time image processing done at sea is far less mentioned to our knowledge.
-Usually, images are brought back to the lab and classified with CNNs which, in comparison with features-based classifiers, that is machine learning algorithms using a set of features extracted from the image (e.g.
-area, eccentricity, gray levels) have outperformed the latter for plankton recognition [@Ellen2019-nh].
-Though, it comes at a cost.
-CNNs are energy consuming hence their implementation for embedded applications can be limited by power consumption [@Biswas2018-uw; @Khabbazan2019-oi; @Oh2017-kj].
+Automatic plankton classification is an active research topic.
+Within this context, the application of Convolutional Neural Networks (CNNs) has already lead to promising results [@Luo2018-oo; @Cui2018-nt; @Dai2017-ex; @Campbell2020-cq; @Schroder2019-um; @Bochinski2019-af; @Lee2016-jl; @Cheng2019-ar; @Guo2021-xi; @Ellen2019-nh; @Orenstein2017-ya].
+Indeed, CNNs have outperformed the previous features-based classifiers (i.e.
+machine learning algorithms using a set of features extracted from the image such as area, eccentricity, gray levels, etc.) [@Ellen2019-nh].
+Nevertheless, real-time image processing, performed at sea, in the instrument, has not been tackled, to our knowledge.
+Usually, images are brought back to the lab and classified with classic machine learning classifiers and, increasingly, CNNs.
+But CNNs are energy consuming, even just for inference (i.e.
+processing and classification of new images, without retraining), and their implementation for embedded applications can be limited by power consumption [@Biswas2018-uw; @Khabbazan2019-oi; @Oh2017-kj].
 
@@
 The lifetime of floats is limited by their batteries hence by the power consumed by its sensors.
-Therefore, using an embedded CNN to classify plankton images or sending images through iridium transmission to be classified in the lab would not be possible for economical (cost of iridium transmission) and technical reasons.
-Typically, a BGC-Argo float samples the water column between the surface and 1,000 m, it would therefore have to stay at the surface for a time long enough to transmit all images collected after each profile, consuming energy proportionally to the size of the image set while being subject to biofouling or unpredictable hazards.
-For all these reasons, it was decided to use a features-based classifier for the development of the embedded plankton recognition algorithm on board the UVP6.
+The floats are not recovered and their data has to be sent through satellite communication.
+Therefore, using an embedded CNN to classify plankton images onboard or sending images through the satellite link are both impossible with currently available hardware.
+[JOI: ici il manque une justif de pq c'est poas possible de mettre un CNN dans le flotteur; il y a qq chiffres dans l'étude de FOTONOWER; tu peux aussi parler des cartes dédiées mais qui sont encore largement trop consommatrices => il manque un chiffre sur la conso énergétique en W attendue d'un capteur; un ordre de grandeur suffit].
+Typically, BGC-Argo floats sample the water column between 1,000 m [JOI: c'est pas 2000?] and the surface; in this range, the UVP6 is expected to take images of hundreds of planktonic objects and, given the bandwidth of the current connection, it would require the float to stay [JOI: n minutes/hours/days] to transmit them [JOI: pareil, on avait des stats et même des estimations de temps, il faut donner les chiffres], consuming energy proportionally to the amount of data sent and being subject to biofouling or unpredictable hazards.
+For these reasons, it was decided to use a simpler features-based classifier for the development of the embedded plankton recognition algorithm on board the UVP6.
 
-In the literature, such algorithms (e.g.
-Decision Tree (DT), Random Forest (RF), Support Vector Machine (SVM)) have already been used for plankton recognition (see Table 1 of @Gonzalez2017-tr for an exhaustive list).
+Such algorithms (e.g.
+Decision Tree (DT), Random Forest (RF), Support Vector Machine (SVM)) have already been used for plankton classification (see Table 1 of @Gonzalez2017-tr for an exhaustive list at that time).
 For instance, @Blaschko2005-ep obtained an accuracy of 71% using a SVM on a set of 982 images among 13 classes using a 10-fold cross-validation (CV), @Sosik2007-cj obtained a global accuracy of 88% using a SVM on a test set of 3300 images covering 22 classes, @Bell2008-yo obtained 82% of accuracy using a 10-fold CV RF algorithm on 53 classes with 10 to 30 images per class.
 However, it should be noted that this accuracy dropped to 63% after the removal of non-biological classes such as detritus and fiber.
-For bigger training sets, which will be the case in this study, @Orenstein2017-ya trained a RF model on approximately 650.000 images across 70 classes for an accuracy of 91% on the test set (2.75 million images).
-Though, one class in that test set contained 1.6 million images and if the classifier had put every object in that class, the accuracy would have been around 60% so caution must be applied when interpreting performance metrics.
+Using a larger data set or 650.000 images, similar to the one used in this study, @Orenstein2017-ya trained a RF model over 70 classes that yielded an accuracy of 91% on the test set (2.75 million images).
+Though, one class in that test set contained 1.6 million images and a dummy classifier that would have put every object in that class would have given an accuracy of about 60%; so caution must be applied when interpreting this global performance metric.
 @Faillettaz2016-wr trained a RF model on 6000 images across 14 classes, reaching an accuracy of 56% on a test set of 1.5 million objects, discarding those below a certain probability threshold.
 This allowed to improve the global accuracy by 16% but resulted in the loss of 72% of objects in the predicted data set.
+[JOI: Il y a une review par Irisson et al qui fait un petit bilan de tout ça ;-)]
 
-It is quite difficult to compare all methods due to the diversity of imaging instruments hence the variety of image formats (e.g.
-resolution), the number of classes (trade off between the model performance and the number of classes as illustrated in @Fernandes2009-hw, the training set (e.g.
-number of images per class, diversity, balance), the scientific objective (e.g.
-precision-recall trade off), the type and number of features and the algorithm used.
-In many cases, RF seemed to be the best classifier [@Gorsky2010-ca; @Faillettaz2016-wr] but in embedded applications, energy constraints need to be taken into account.
-Therefore, the model needs to be small and the classification should be easy and fast.
-Gradient Tree Boosting (GTB) seems to offer this compromise when compared to RF (model size) and SVM (computation complexity).
-For these reasons, we chose XGBoost [@Chen2016-xt], an optimized implementation of GTB, also known to be one of the best assets to win Kaggle's machine learning competitions.
+It is quite difficult to compare all methods due to the diversity of imaging instruments resulting in a variety of image resolutions, numbers of classes (while there is a trade off between model performance and the number of classes as illustrated in @Fernandes2009-hw), training sets (e.g.
+number of images per class, diversity, balance), scientific objectives (e.g.
+precision-recall trade off), types and numbers of features used and classification algorithms.
+In many cases, RF seemed to be the best classifier [e.g.
+@Gorsky2010-ca] [JOI: je crois que c'est plutôt grosjean].
+For embedded applications, the number of parameters in the model need to be small for its memory footprint to be reduced and the computation of new classifications should be fast.
+Gradient Boosted Trees (GBT) seem to offer this compromise when compared to RF (model size) and SVM (computation complexity).
+For these reasons, we chose to classify images using the XGBoost library [@Chen2016-xt], an optimized implementation of GBT, known to be high performing in machine learning competitions (on the Kaggle website for example).
 
-In this paper, we present the results of an embedded zooplankton recognition algorithm using XGBoost on UVP5 data.
-We show how we built the embedded recognition (ER) model under strict technical constraints, what were our compromises and what techniques we used to improve the classification.
+In this paper, we present the results of an embedded zooplankton classification algorithm using XGBoost on UVP5 data.
+We show how we built the embedded classification (EC) [JOI: UVPec = UVP embedded classification => soit cohérent; et puis recognition c'est pas trop ça en Anglais] model under strict technical constraints, what were our compromises and what techniques we used to improve the classification.
 We also compare the performances of our features-based algorithm with a state-of-the-art neural network as if energy was not a limiting factor.
-Finally, we apply the method and follow the recommendations learnt from the preliminary UVP5 model to the UVP6 data set to build the ER model that will be included in the UVP6 for AUVs.
+Finally, we apply the method and follow the recommendations learnt from the preliminary UVP5 model to the UVP6 data set to build the EC model that will be included in the UVP6 for AUVs.
 
@@
 
-GTB generates a sequence of trees, each one trying to cope with the errors made by their predecessors.
-XGBoost is designed to be an optimized and regularized version of GTB.
+GBT generates a sequence of trees, each one trying to cope with the errors made by their predecessors.
+XGBoost is designed to be an optimized and regularized version of GBT.
 To learn from the data and derive a model, XGBoost first minimizes a regularized objective function $L$

### 2023-03-27 index.qmd
update

--- 2023-03-24 index.qmd
@@
-# Foreword {.unnumbered}
+# Introduction
 
-As of 2022, it has now become obvious to the general public that the Earth's climate is changing.
-Massive flood events, repeated heatwaves, never-ending droughts, disappearing glaciers, biodiversity collapse, … Reports on weather anomalies were already published in the 1960's, the Intergovernmental Panel on Climate Change (IPCC) was created in 1988, the Kyoto Protocol was adopted in 1997, An Inconvenient Truth was released in 2006, the Paris Agreement was signed in 2015 and still, what has really changed?
+## The global carbon cycle {#sec-global_carbon_cycle}
 
-As I am writing this part, the COP27 (sponsored by Coca-Cola, no joke) is taking place with an acknowledged record number of fossil fuel lobbyists, bloody stadiums with outside air conditioning are about to be used in the desert for a month and it seems like 'leaders' of the Free World have been having hearing issues.
-See for yourself:
+The global carbon cycle (@fig-global_carbon_cycle) consists of a series of exchanges of carbon in various organic and inorganic forms among four major reservoirs: the lithosphere, the hydrosphere, the atmosphere and the biosphere.
+These exchanges occur at different time scales, such that the carbon cycle can be roughly divided into a slow and a fast component.
+The former operates on geological time scales where carbon is hidden from the surface for millions of years (e.g.
+in hydrocarbon deposits such as petroleum or natural gas) before a part of it is released through volcanic activity while the latter operates on shorter time scales, from minutes to years (e.g.
+photosynthesis, air-sea fluxes) up to decades and millenia (e.g.
+ocean ventilation, changing ocean buffering capacity) [@Bertrand2021-nh; @Canadell2021-ds].
 
-*"I am drowning"* said Pakistan
+99% of the Earth's carbon is stored in rocks, holding about 100,000 times as much carbon as the atmosphere while the ocean contains about 60 times as much carbon as the atmosphere [@Sarmiento2006-ec; @Bertrand2021-nh; @DeVries2022-of] with roughly 37,000 Pg (i.e.
+1 Pg = 10$^{15}$ g = 1 Gt) of dissolved inorganic carbon (DIC) and 700 Pg of dissolved organic carbon (DOC) [@Friedlingstein2022-th].
 
-*"I'm sorry, what?"* said the Free World
+![Illustration of the global carbon cycle (numbers represent a global average for the decade 2012-2021).
+Arrows and circles depict, respectively, carbon fluxes in Pg C year$^{-1}$ and carbon content in Pg C.
+Figure taken from @Friedlingstein2022-th.](image/intro/global_carbon_cycle.png){#fig-global_carbon_cycle}
 
-*"I am starving"* said Ethiopia
+The global carbon cycle had been at close equilibrium for millions of years before the industrial revolution started in the 18th century.
+Since then, human have been burning fossil fuels at an unprecedented rate, releasing huge amounts of carbon dioxide ($CO_{2}$) in the atmosphere that should have been discharged slowly through the slow carbon cycle [@Bertrand2021-nh].
+These emissions, resulting from fossil emissions but also from land use (e.g.
+deforestation, land management) have been estimated around 700 Pg C since 1750 [@Canadell2021-ds].
 
-*"Can't hear you, louder please!"* said the Free World
+Thankfully, more than half of this anthropogenic $CO_{2}$ has been absorbed by terrestrial ecosystems and by the ocean [@Friedlingstein2022-th], forming major carbon sinks.
 
-*"We are suffocating"* said India
+## How is the ocean a $CO_{2}$ sink? {#sec-ocean_co2_sink}
 
-*"Aaaah okay. Yes you're absolutely right, let's save the banks one more time"* said the Free World
+The ocean is known to play a key role in the fast carbon cycle due to its size and rapid exchange of $CO_{2}$ with the atmosphere, controlling atmospheric $CO_{2}$ levels on time scales of decades to millennia [@Sarmiento2006-ec; @DeVries2022-of].
+Currently, the ocean absorbs \~2.5 Pg C year$^{-1}$, which represents about 25% of annual anthropogenic $CO_{2}$ emissions [@Canadell2021-ds; @Friedlingstein2022-th].
+In order to understand how the ocean absorbs the latter, one must take a look at seawater carbonate chemistry.
 
-Although I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured that there is no fake news in this work. I hope you will enjoy it.
+### Inorganic carbonate chemistry {#sec-inorganic_carbonate_chemistry}
+
+The intensity of the exchange of $CO_{2}$ at the air-sea interface is driven by the difference between the partial pressure of $CO_{2}$ in the atmosphere ($pCO_{2}^{atm}$) and in the surface ocean ($pCO_{2}^{surface}$).
+
+$$
+\Delta pCO_{2} = pCO_{2}^{surface} - pCO_{2}^{atm}
+$$ {#eq-deltaCO2}
+
+When $\Delta pCO_{2} > 0$, surface waters are supersaturated in $CO_{2}$ resulting in a net flux of $CO_{2}$ from the ocean to the atmosphere.
+On the contrary, surface waters are undersaturated with a net $CO_{2}$ flux in the opposite direction [@Sarmiento2006-ec].
+
+When a molecule of $CO_{2~(gas)}$ dissolves in seawater, it first gets hydrated to form $CO_{2~(aq)}$ then it reacts with water to form carbonic acid ($H_{2}CO_{3}$).
+
+$$
+CO_{2~(gas)} + H_2{O} \rightleftharpoons CO_{2~(aq)} + H_{2}O \rightleftharpoons H_{2}CO_{3}
+$$ {#eq-carbonic_acid}
+
+$H_{2}CO_{3}$ is a strong acid that quickly dissociates in two steps to form bicarbonate ($HCO_{3}^{-}$) and carbonate ($CO_{3}^{2-}$) where each reaction releases one free proton ($H^{+}$).
+
+$$
+H_{2}CO_{3} \rightleftarrows H^{+} + HCO_{3}^{-}
+$$ {#eq-bicarbonate}
+
+$$
+HCO_{3}^{-} \rightleftarrows H^{+} + CO_{3}^{2-}
+$$ {#eq-carbonate}
+
+The abundance of each carbonate species is governed by the equilibrium between those three relationships [@Sarmiento2006-ec].
+According to @Gattuso2011-di, at a typical surface pH of 8.2, the relative abundances of $CO_{2}$ (encompassing $CO_{2~(aq)}$ and $H_{2}CO_{3}$ because it is difficult to distinguish them analytically), $HCO_{3}^{-}$ and $CO_{3}^{2-}$ are, respectively, 0.5%, 89% and 10.5%.
+For convenience, the sum of all dissolved carbonate species will be hereafter referred to as DIC
+
+$$
+DIC = [CO_{2}] + [HCO_{3}^{-}] + [CO_{3}^{2-}]
+$$ {#eq-DIC}
+
+where brackets denote concentrations.
+
+### Distribution of DIC in the water column {#sec-DIC_distribution}
+
+It can be shown [e.g.
+@DeVries2022-of] that if we were to consider the ocean as a homogeneous reservoir of inorganic carbon, $pCO_{2}^{atm}$ would be much higher than it is now (roughly 416 parts per million (ppm) as of October 2022 according to the Mauna Loa Observatory).
+This observation is explained by the presence of a strong vertical gradient of DIC in the ocean, with surface DIC concentrations more than 200 $\mu$mol kg$^{-1}$ lower than at 1,000 m as it can be seen in @fig-DIC_plot.
+
+```{r, DIC plot}
+#| label: fig-DIC_plot
+#| fig-cap: Mean DIC distribution in the global ocean.
+Data taken from the Merged Master File from GLODAPv2, available at https://www.glodap.info/index.php/merged-and-adjusted-data-product-v2-2022/, last accessed on 26/10/2022.
+
+library(ggplot2)
+library(tidyverse)
+library(ggtext)
+library(latex2exp)
+
+# CODE BELOW IS *OKAY* but takes time to compile so I am just loading the last working dataframe to plot the DIC distribution.
+
+# DIC_data <- vroom::vroom('/home/flo/Downloads/GLODAPv2_2022_all/GLODAPv2.2022_Merged_Master_File.csv')
+# DIC_data <- DIC_data |> select(depth= G2depth, DIC = G2tco2, G2tco2f, G2tco2qc)
+# # clean a bit
+# DIC_data <- DIC_data |> filter(G2tco2qc == 1, G2tco2f == 2)
+# DIC_data <- DIC_data |> filter(depth < 4500)
+# # bin it
+# # brks <- c(0,10,20,30,40,50,60,70,80,90,100,150,200,250,300,350,400,450,500,600,700,800,900,1000,1500,2000,2500,3000,3500,4000,4500)
+# # big bin
+# brks <- c(0,20,40,60,80,100,150,200,250,300,350,400,450,500,600,700,800,900,1000,1500,2000,2500,3000,4500)
+# # smaller bins
+# brks <- c(0,20,40,60,80,100,150,200,250,300,350,400,450,500,600,700,800,900,1250,1500,1750, 2000,2250, 2500, 2750,3000,3250,3500, 3750, 4000, 4250, 4500)
+# ints <- findInterval(DIC_data$depth, brks, all.inside = T)
+# mids <- (brks[ints] + brks[ints + 1]) / 2
+# DIC_data$mids <- mids
+# DIC_data <- DIC_data |> dplyr::group_by(mids) |> dplyr::summarise(mean_DIC = mean(DIC, na.rm=T), nb_points = n())
+
+# save data
+#vroom::vroom_write(DIC_data, '/home/flo/dox/THESIS/DATA/GLODAP/DIC_distribution.csv')
+
+# load data 
+DIC_data <- vroom::vroom('./DATA/GLODAP/DIC_distribution.csv')
+
+# https://stackoverflow.com/questions/12514612/how-to-annotate-ggplot-with-latex
+ggplot(DIC_data, aes(y = mids, x = mean_DIC)) + geom_path(size = 1) + scale_y_reverse() + ylab('Depth (m)') + xlab(TeX('DIC ($\\mu$mol kg$^{-1}$)')) + theme_bw()
+
+```
+
+This gradient is mostly due to processes often referred to as carbon pumps, operating against continuous ocean mixing processes that want to homogenize DIC in the water column [@Sarmiento2006-ec; @Claustre2021-tx].
+They were originally defined by @Volk1985-oc as "processes that deplete the ocean surface of DIC relative to the deep-water DIC".
+
+### Several carbon pumps at work {#sec-carbon_pumps}
+
+#### Solubility pump {#sec-solubility-pump}
+
+From @eq-carbonic_acid, one can derive that p$CO_{2}^{water~parcel}$ is related to the dissolved $CO_{2}$ concentration by the following relation
+
+$$
+pCO_{2}^{water~parcel} = \frac{[CO_{2}]}{K_{0}}  
+$$ {#eq-henry_law}
+
+where $K_{0}$ is the solubility coefficient of $CO_{2}$ in seawater.
+
+$K_{0}$ is a function of temperature and salinity with a strong dependence on the former [@Sarmiento2006-ec] such that the solubility of $CO_{2}$ increases with decreasing temperatures.
+Consequently, waters that cool as they are transported to higher latitudes by the thermohaline circulation will be undersaturated in $CO_{2}$ once they reach the surface.
+Thus, there will be a net exchange of carbon dioxide from the atmosphere to the ocean, enriching surface waters in DIC before they dive back to the abyss at deep water formation zones in North Atlantic polar regions (e.g.
+the Labrador Sea) and along the Antarctic shelf in the Southern Ocean [@DeVries2022-of].
+
+The solubility pump is responsible for the great majority of all the anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed DIC gradient according to @Sarmiento2006-ec or 35% according to @Toggweiler2003-ye.
+Both values are used in the literature (e.g.
+10% in @Boyd2019-ez and @Iversen2023-qj and 35% in @Riebesell2009-qj, @Passow2012-zh and @DeVries2022-of).
+
+#### Carbonate pump {#sec-carbonate-pump}
+
+Calcifying organisms (mostly coccolithophores, foraminifera and pteropods) produce particulate inorganic carbon (PIC) by precipitating biogenic $CaCO_{3}$ [@Sarmiento2006-ec]
+
+$$
+Ca^{2+} + CO_{3}^{2-} \rightleftharpoons CaCO_{3}
+$$ {#eq-calcium_carbonate}
+
+to form calcite structures such as calcite platelets (coccoliths), calcite shells or spines [@Neukermans2023-vl].
+
+Combining @eq-calcium_carbonate, @eq-bicarbonate and @eq-carbonate, one can derive another formulation of calcification
+
+$$
+Ca^{2+} + 2HCO_{3}^{-} \rightleftharpoons CaCO_{3} + CO_{2} + H_{2}O
+$$ {#eq-calcification}
+
+This calcification occurs in surface waters, releasing $CO_{2}$ in the surrounding waters and subsequently to the atmosphere (0.26-1.19 Pg C year$^{-1}$ according to @Neukermans2023-vl).
+That is why this pump is sometimes also referred to as the carbonate counter pump [@Zeebe2012-zk].
+However, the dissolution of $CaCO_{3}$ that mainly occurs at depth increases the DIC concentration in contrast with the calcification (see @eq-calcification).
+In all, @Sarmiento2006-ec estimated that the carbonate pump contributes to 20% of the observed DIC gradient.
+
+#### Biological pump {#sec-biological-pump}
+
+Also known as the soft-tissue pump [@Volk1985-oc], the biological carbon pump (BCP), is defined by @Ipcc2019-qc as "a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis (@eq-photosynthesis) in sunlit surface water and then transported to the ocean interior, and possibly the sediment, resulting in the storage of carbon".
+
+$$
+CO_{2} + H_{2}O \rightarrow CH_{2}O + O_{2} 
+$$ {#eq-photosynthesis}
+
+The carbon flux leaving the surface layer for the ocean interior is referred to as the export flux [@Siegel2022-wd].
+Historically, the export depth was usually set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should now use the base of the euphotic zone (EZ, where there is enough light for net primary production (NPP)) as the export depth to clearly differentiate the layer where particulate organic carbon (POC) is produced (i.e.
+via photosynthesis) and where POC is consumed (i.e.
+remineralized).
+This remineralization (or respiration) consists in the conversion of organic matter back to inorganic forms (i.e.
+nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd] such as zooplankton or fish.
+
+Those two biological activities significantly enhance the DIC gradient such that the BCP would thus be responsible for the remaining 70% of the observed DIC gradient [@Sarmiento2006-ec].
+Without it, there would be no production of organic matter in the EZ (consuming DIC) and no remineralization in the ocean interior (producing DIC) resulting in a weaker DIC gradient.
+In fact, @DeVries2022-of calculated that such a decrease in the DIC gradient would reduce the ocean carbon inventory by roughly 10% and increase atmospheric $CO_{2}$ levels by a factor 2.
+Similarly, @Kwon2009-ir showed that a deepening of the remineralization depth (i.e.
+where 63% of sinking carbon is respired) would decrease atmospheric $CO_{2}$ concentrations due to the redistribution of respired carbon from intermediate waters to bottom waters (i.e.
+increased DIC gradient).
+
+The sinking carbon that is not remineralized is said to be sequestered.
+The sequestration (storage) of carbon is usually defined as the long-term removal of $CO_{2}$ (or other forms of carbon) from the atmosphere on climatically significant time scales [@Boyd2019-ez; @Ipcc2019-qc; @Siegel2022-wd].
+While this definition seems logical, it does not specify precisely what is meant by "climatically significant time scales".
+In what follows, I will use the 100-year time horizon to refer to carbon sequestration (unless specified otherwise) as used by other BCP studies [e.g.
+@Lampitt2008-kq; @Passow2012-zh].
+
+The amount of carbon that is sequestered by the BCP mainly depends on its strength and its efficiency.
+In this manuscript, the strength of the BCP is defined as "the magnitude of the POC flux at a reference export depth" [@Claustre2021-tx] and its efficiency (also known as transfer efficiency) is defined as the fraction of POC flux transported to some reference depth below the export depth [@Buesseler2020-sn; @Claustre2021-tx].
+
+::: {.blackbox data-latex=""}
+::: {.flushleft data-latex=""}
+**Information Box 1.** **BCP and blue carbon**
+:::
+
+The BCP is a natural-based solution for carbon sequestration by definition (@eq-photosynthesis).
+However, natural-based solutions for carbon sequestration are generally aligned with the concept of 'blue carbon' defined by @Ipcc2019-qc as "all biologically-driven carbon fluxes and storage in marine systems that are amenable to management" which usually encompasses only coastal ecosystems [@Ipcc2019-qc; @Lovelock2019-dr].
+Coastal blue carbon ecosystems (e.g.
+mangroves, tidal marshes, seagrass meadows) are thought to store in their living biomass (years to decades) and sediment (millenia) more than 30 Pg C with an estimated absorption of \~0.3 Pg C year$^{-1}$ [@Macreadie2021-rb] though those numbers should be taken with caution because there are major uncertainties in the distribution of seagrass meadows and tidal marshes [@Macreadie2021-rb] as well as in carbon burial rates [@Williamson2022-ud].
+In spite of that, the current expectation of additional carbon removal by coastal blue carbon by 2100 should be around 25-30 Pg C [@Gattuso2021-cr; @Williamson2022-ud], that is the equivalent of 2.5-3 years of current anthropogenic emissions [@Friedlingstein2022-th].
+:::
+
+Estimates of global carbon export range between 5 and 12 Pg C year$^{-1}$ [@Laws2000-jl; @Dunne2007-mi; @Henson2011-mt; @Siegel2014-wg; @DeVries2017-wu; @Nowicki2022-no] with only 0.02-0.4 Pg C year$^{-1}$ reaching the sediment [@Dunne2007-mi; @Hayes2021-gg; @Nowicki2022-no] where carbon will be sequestered for millions of years.
+Almost all of the exported organic carbon will therefore be remineralized and sequestered away from the atmosphere on time scales ranging from months to millennia.
+[@Siegel2022-wd].
+
+Nonetheless, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess atmospheric $CO_{2}$, released by anthropogenic activities.
+First, as explained in @Sarmiento2006-ec, atmospheric $CO_{2}$ levels were stable for several thousand years before the industrial revolution started, therefore it is assumed that the BCP was operating at steady state.
+Second, photosynthesis is mostly controlled by light and nutrients availability, not by $CO_{2}$.
+
+However, climate change is most likely going to impact the biological (e.g.
+shift in phytoplankton communities), physical (e.g.
+temperature) and chemical (e.g.
+nutrients availability, acidification) drivers of the BCP [@Riebesell2009-qj; @Passow2012-zh; @Bopp2013-jx] whose long-term carbon sequestration capacity remains uncertain [@Passow2012-zh; @Turner2015-af; @DeVries2022-of; @Henson2022-hg ; @Wilson2022-bf].
+
+#### Microbial pump {#sec-microbial-pump}
+
+Unlike the three previous carbon pumps, the microbial carbon pump (MCP) was only introduced recently by @Jiao2010-ax as "a conceptual framework to address the role of microbial generation of recalcitrant dissolved organic matter[^intro-1] and relevant carbon storage".
+In simple terms, it is a suite of marine microbial mechanisms that produce longer-lived DOC such as semi-refractory, refractory and ultra-refractory DOC that have an average lifetime of respectively 205, 16,000 and 40,000 years [@Legendre2015-jc].
+
+[^intro-1]: i.e. dissolved organic matter resistant to microbial utilization that can persist in the ocean interior for thousands of years.
+
+The term 'pump' is attributed to the MCP because it maintains a gradient between DOC$_{<100~years}$ (14 Pg C) and DOC$_{>100~years}$ (non-bioavailable DOC, 636 Pg C) where 100 years corresponds to the 100-year time horizon that I use to refer to carbon sequestration.
+
+In contrast with the other pumps which are depth-dependent, the MCP sequesters carbon by biochemically transforming DOC to non-bioavailable DOC.
+This process occurs at all depths, making the MCP depth-independent [@Legendre2015-jc].
+
+In order to fully understand the fate of DOC and DIC in the ocean, one would need to integrate both classic carbon pumps and the MCP.
+However, in this manuscript, I will only focus on the BCP, which was still considered as a single process until recently.
+
+## A new paradigm for the BCP {#sec-a-new-paradigm-for-the-bcp}
+
+In the frame of the BCP, global carbon budgets have generally been imbalanced such that the exported POC was up to 2-3 orders of magnitude less abundant that it would need to be to sustain heterotrophic respiration in the mesopelagic zone and beyond [@Burd2010-hc; @Turner2015-af].
+This imbalance raises one question: are we currently underestimating the export carbon flux or are we overestimating the heterotrophic respiration?
+
+The last part of the question is out of the scope of this manuscript but it should be said that a lot of uncertainties orbit around those carbon deficits, namely parameters used to estimate the metabolic activity of deep ocean organisms as well as deep food web dynamics and unquantified carbon sources such as the contribution of slowly settling particles and better estimation of POC fluxes [@Burd2010-hc; @Martin2020-wa].
+
+Regarding the latter, @Boyd2019-ez published a paper in which they review the BCP by adding new biological and physical mechanisms[^intro-2] that inject suspended and sinking particles to depth (@fig-global_carbon_pumps).
+A better understanding of those mechanisms (see next subsections) could help to close the carbon budget.
+
+[^intro-2]: More specifically, as stated by @Claustre2021-tx "the word 'biological' in BCP refers to the transfer of biogenic material to depth, mediated by either biological or physical processes."
+
+![Illustration of the three main mechanisms through which the BCP exports carbon to the deep ocean.
+1) Production of POC and transport to depth by gravitational settling.
+2) Vertical migration of zooplankton that feed at the surface and metabolize the ingested carbon at depth (e.g.
+production of fecal pellets).
+3) Subduction of DOC and suspended POC via mixing processes.
+Sequestration times for DIC are set depending on the depth at which POC and DOC are exported.
+Figure taken from @Iversen2023-qj.](image/intro/carbon_pumps_iversen2022.jpeg){#fig-global_carbon_pumps}
+
+### Biological gravitational pump {#sec-BGP}
+
+The most well known pump is the gravitational carbon pump (BGP).
+It consists in the gravitational settling (i.e.
+sinking) of POC[^intro-3] to the deep ocean and it is the main contributor (60-70%) of the export flux [@Boyd2019-ez; @Nowicki2022-no].
+Historically, the downward POC flux was directly measured using sediment traps [@Martin1987-um; @Honjo2008-kn] and indirectly measured using thorium-234 ($^{234}Th$) deficits as a POC flux proxy [e.g.
+@Buesseler2006-og].
+More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size distributions (PSDs, also referred to in the literature as particle size spectra (PSS)) that could then be converted to carbon fluxes [e.g.
+@Guidi2008-vp].
+
+[^intro-3]: There obviously exists a gravitational settling of particulate inorganic carbon but in this manuscript, I focus on the BCP which was defined earlier as only focused on organic carbon as opposed to the carbonate pump.
+
+### Physically mediated particle injection pump {#sec-MixingPump}
+
+The physically mediated particle injection pump (also referred to as the mixing pump in the literature) consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez].
+Contrarily to the BGP, all three pumps also transport DOC to depth where it is injected along with suspended (i.e.
+neutrally buoyant) and slow-sinking POC below the EZ on different spatial and time scales.
+
+The mixed layer pump can be defined as a 1D mechanism through which organic carbon is exported below the EZ when there is a deepening of the mixed layer followed by the formation of a new shallower mixed layer, isolating the exported carbon from the surface [@DallOlmo2016-ex].
+This pump can occur at different time scales but it is mainly related to the building of the summer stratification in mid and high latitude regions where seasonal variations of the mixed layer are the strongest [@DallOlmo2016-ex; @Boyd2019-ez].
+
+The other two pumps are both 2D spatial mechanisms (i.e.
+they transport POC and DOC vertically and laterally).
+The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on time scales of days across horizontal spatial scales of 1-10 km [@Omand2015-mf; @Resplandy2019-nx].
+The large-scale (100-1000 km) subduction pump corresponds to the slow advective transport of carbon associated with the meridional overturning circulation and Ekman pumping [@Levy2013-ro; @Claustre2021-tx; @Siegel2022-wd].
+
+In all, the mixing pump would represent at least 20% of the export flux according to @Nowicki2022-no although large uncertainties still remain (see Figure 2C in the paper of @Boyd2019-ez).
+
+### Biologically mediated particle injection pump {#sec-MigrantPump}
+
+The biologically mediated particle injection pump is usually referred to as the vertical migration pump (VMP) [@Boyd2019-ez].
+It consists on the vertical migration of zooplankton or fish [@Steinberg2017-ri; @Saba2021-ew] that directly inject POC (e.g.
+fecal pellets) below the EZ at the depth of their migration.
+This pump can act on daily as well as on seasonal time scales.
+In the case of the diel (i.e.
+daily) migration pump, animals feed at the surface at night and then migrate to depth during daytime to avoid visual predation [@Siegel2022-wd], metabolizing injected food at depth[^intro-4], by-passing the EZ directly [@Steinberg2017-ri].
+In the case of the seasonal migration pump, carbon-rich lipids accumulated in the EZ during summer are catabolized by hibernating copepods at depth, between 600 m and 1,400 m, where they shunt carbon below the permanent pycnocline [@Jonasdottir2015-ob].
+This pump is also referred to as the seasonal lipid pump [@Jonasdottir2015-ob].
+
+[^intro-4]: They can also die at depth during the day (or their overwintering in the case of the seasonal migration pump) which also contributes directly to the export of carbon at depth.
+
+In his model, @Nowicki2022-no estimates that the VMP accounts for approximately 10% of the total export flux, albeit large uncertainties remain on the number of migrating animals and their migration depth.
+The VMP would thus export less organic carbon than the mixing pump but its sequestration potential would be higher (@tbl-PumpDepthRange).
+In contrast to the mixing pump that mostly acts on DOC, migrating animals release rapid sinking fecal pellets that would reach depths such that subsequent remineralized DIC would be hidden from the surface for longer time periods (@tbl-PumpDepthRange).
+As a result, it is estimated that the diel migration pump could potentially sequester 60% as much carbon as the BGP [@Boyd2019-ez].
+
+However, there are still many uncertainties regarding the magnitude of the VMP [@Archibald2019-dj; @Aumont2018-wx; @Gorgues2019-qp] and its sequestration potential [@Boyd2019-ez; @Pinti2023-td] thus its importance for a better understanding of the BCP.
+
+## The fate of carbon {#sec-the-fate-of-carbon}
+
+### Biological processes impacting the downward POC flux {#sec-biological-processes-at-play}
+
+Sinking particles are part of a pool of highly diversified materials combining fecal pellets, phytodetritus from sinking phytoplankton, dead zooplankton and organic aggregates referred to as marine snow (i.e.
+macroscopic organic aggregate \> 0.5 mm) [@Alldredge1988-wm; @Turner2015-af; @Steinberg2017-ri].
+Settling velocities for such particles range from 10-100 m day$^{-1}$ up to 1 km day$^{-1}$ for large salp fecal pellets [@Steinberg2017-ri; @Iversen2023-qj].
+
+Marine snow is paramount for both the BCP (see @sec-a-new-paradigm-for-the-bcp) and the support of marine life within the mesopelagic domain [@Burd2010-hc].
+The remineralization of organic matter by heterotrophic organisms releases $CO_{2}$ in the water column that will be released back to the atmosphere in a time frame depending both on its geographical location and depth (see @sec-will-it-be-sequestered).
+Therefore, understanding how biological processes interact with the downward POC flux and where the bulk of the remineralization occurs is important to better understand the efficiency of the BCP.
+
+The Martin curve [@Martin1987-um] is probably the most well known POC flux attenuation curve (@eq-Martin-Curve and @fig-Martin-Curve). It is a power-law function described as follow
+
+$$
+F_{z} = F_{z_{export}} \left(\frac{z}{z_{export}}\right)^{-b}
+$$ {#eq-Martin-Curve}
+
+where $F_{z}$, $z_{export}$ and $b$ are, respectively, the estimated POC flux at a depth $z$, the depth of the export flux and the remineralization factor.
+The latter is adimensional and often set at 0.86 [@Martin1987-um] although it is well known that the magnitude of flux attenuation varies regionally (and temporally) due to the variability in food web structures beneath the EZ [@Passow2012-zh].
+Geographically variable [@Henson2012-ck] and regionalized [@Guidi2015-sj] estimations of $b$ were proposed to account for non-uniform remineralization at global scale.
+
+```{r, Martin Curve}
+#| label: fig-Martin-Curve
+#| fig-cap: Typical attenuation curve of the POC flux obtained with @eq-Martin-Curve.
+The color-coded layers represent regions with carbon-specific degradation rate, decreasing with depth.
+Figure inspired from @Iversen2023-qj.
+
+library(scales)
+
+z_export <- 100
+F_export <- 1
+b <- 0.86
+
+F_martin <- function(z){
+  flux <- F_export * ((z/z_export)**(-b))
+  return(flux)
+}
+
+# test <- F_martin(100)
+depth <- seq(100,3000,10)
+fluxes <- F_martin(depth)
+df <- data.frame(z = depth, flux = fluxes)
+opacity <- 1
+
+ggplot(df, aes(x = fluxes, y = depth)) + 
+  annotate("rect", ymin = 100, ymax = 300, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#edf8b1") +
+  annotate("rect", ymin = 300, ymax = 500, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#7fcdbb") + 
+  annotate("rect", ymin = 500, ymax = 1000, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#1d91c0") +
+  annotate("rect", ymin = 1000, ymax = 3000, xmin = -Inf, xmax = Inf, alpha = opacity, fill = "#225ea8") +
+  scale_y_continuous(trans = c('reverse')) + geom_path(size = 1.5) +
+  theme_bw() +
+  labs(x = 'POC flux (mg C m<sup>-2</sup> day<sup>-1</sup>)', y = 'Depth (m)') + theme(
+    axis.title.x = element_markdown(),
+    axis.title.y = element_markdown()) 
+```
+
+Based on the Martin curve, @Iversen2023-qj divided the water column in 4 layers (@fig-Martin-Curve): the upper mesopelagic layer (export depth - 300 m) where there is a rapid attenuation of the downward POC flux, the mid-mesopelagic layer (300 - 500 m) and the lower mesopelagic layer (500 - 1,000 m) where both layers show two levels of intermediate attenuation and the bathypelagic layer (1,000 m - seafloor) where the attenuation is relatively uniform.
+For each layer, @Iversen2023-qj estimated a carbon-specific degradation rate in order to better understand the processes responsible of this attenuation.
+As a result, @Iversen2023-qj suggests that the POC attenuation in the upper mesopelagic layer is the combined action of microbial degradation and zooplankton that graze on settling aggregates.
+Here, zooplankton are the main responsible for the attenuation whereas in the middle and lower mesopelagic layers, microbial degradation takes over due to the decrease in zooplankton abundance.
+Finally, the relatively constant degradation occurring below 1,000 m in the bathypelagic zone could be explained by the increasing pressure and decreasing temperatures on microbial communities.
+
+The impact on the downward POC flux by zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2023-qj].
+On the one hand, aggregate feeding, filter-feeding, direct ingestion of fecal pellets (i.e.
+coprophagy) and fragmentation of large aggregates to small slower-sinking particles will likely decrease the downward POC flux.
+On the other hand, the ingestion of slow-sinking cells to defecation of large and compact (i.e.
+fast-settling) fecal pellets will increase it.
+The remaining organic matter (\~15% according to @Turner2015-af) will then be degraded by microbial communities.
+In the end, only a small fraction of the exported carbon will reach the sediment where it will be stored for millions of years.
+
+### Physical properties influencing the downward POC flux {#sec-sinking-speed}
+
+The settling speed $w$ is the most important physical parameter controlling the remineralization depth (@fig-sinking-speed) of aggregates and fecal pellets [@Laurenceau-Cornec2015-pm].
+Using Stokes' Law for a spherical particle at low Reynolds number, the sinking speed can be expressed as
+
+$$
+w = \frac{g}{18\mu} (\rho_{p}-\rho_{w})d^{2}
+$$ {#eq-sinking-speed}
+
+where $g$ is the gravitational acceleration, $\mu$ is the dynamic viscosity of seawater, $\rho_{p}$ and $\rho_{w}$ are the particle and seawater densities and $d$ is the particle diameter.
+If one assumes that changes in dynamic viscosity are negligible [@De_La_Rocha2007-yi] compared to changes from the remaining parameters (i.e.
+this assumption is reasonable based on the fact that $\mu$ can roughly vary by a factor 2 [@Sharqawy2010-ig]), then $w$ should mostly be proportional to the excess density $\Delta\rho$ (i.e.
+$\rho_{p} -\rho_{w}$) and $d$.
+It is however a simplification of reality because most aggregates are not spherical and have fractal geometries [@Logan1990-gy; @Logan1995-id; @Laurenceau-Cornec2020-zc].
+
+![Illustration of the main factors influencing the sinking velocity following Stokes' Law and hypothetical attenuation fluxes (in % of export flux) with depth (vertical axis) for slow and fast sinking particles.
+The latter shows that slow sinking particles experience higher rates of flux attenuation due to increased duration of exposure to remineralization hence contributing less to carbon sequestration.
+Figure taken from @Williams2022-xx.](image/intro/sinking_speed.jpeg){#fig-sinking-speed}
+
+Taking this into consideration, several processes can impact the sinking speed of POC such as aggregation (e.g.
+physical coagulation, incorporation of small particles into fecal pellets (i.e.
+repackaging)), disaggregation or fragmentation (either physically (e.g.
+fluid shear) or biologically by microbial degradation and zooplankton feeding), solubilization to DOC by bacteria and mineral ballasting [@Turner2015-af].
+
+The driving factors of the sinking speed are nevertheless subject to caution.
+It would make sense that the quadratic dependence of $w$ to the aggregate size would favor $d$ as the main driver.
+However, aggregates are far from being perfect theoretical objects for @eq-sinking-speed and the excess density as the main driver is plebiscited by some authors such as @Iversen2020-kb.
+
+To the best of my knowledge, the most emblematic case of discussions around this question is the mineral ballast hypothesis [@De_La_Rocha2007-yi].
+Mineral ballasting consists in the inclusion of biominerals such as opal (i.e.
+silicate) and calcite or lithogenic material such as dust that are thought to increase $\Delta\rho$ on the one hand and protect them from microbial degradation on the other hand [@Armstrong2001-rz; @Francois2002-uc; @Klaas2002-kn].
+However, several studies argued otherwise.
+@Passow2004-re stipulated that biominerals could instead be collected (i.e.
+scavenged) and transported downwards by sinking aggregates.
+@Ploug2008-qk did not measure significant differences in microbial degradation of non ballasted vs.
+ballasted material.
+Finally, @Passow2006-hw developed an experiment to investigate the relationship between the incorporation of biominerals into aggregates and their sinking speed.
+They showed that this relationship was not straigthforward due to the fact that the inclusion of biominerals fragmented aggregates into smaller but denser particles which could either decrease (smaller size) or increase (denser particles) their sinking speed.
+Other studies on the ballasting effect have been published since but even 20 years after its proposition, no global ballasting mechanism has been found [@Neukermans2023-vl].
+
+The size of aggregates was usually thought to be a key predictor of settling velocity [e.g.
+@Passow2006-hw] which is coherent with @eq-sinking-speed.
+However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor.
+Because the sinking speed determines the time available for degradation in the water column, it also directly impacts the potential for carbon sequestration.
+With all things considered, there is thus a need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Giering2020-zk; @Iversen2023-qj; @Williams2022-xx].
+
+### Sequestration potential {#sec-will-it-be-sequestered}
+
+Once injected in the ocean interior (@tbl-PumpDepthRange), the POC flux originating from the five particle injection pumps is merged back with the BGP.
+How long is this carbon going to be sequestered from the atmosphere is one of the key questions of carbon cycle studies [@Claustre2021-tx].
+
+```{=html}
+<!--#https://github.com/quarto-dev/quarto-cli/discussions/1042-->
+<style type="text/css">
+caption, .table-caption {
+  text-align: left;
+}
+</style>
+```
+---
+tbl-cap-location: top
+---
+
+| Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
+|------------------|------------------|------------------|-------------------|
+| Gravitational          | export depth - sediment          | None                      | N/A (142)                            |
+| Mixed layer            | 200 - 1,000 m                    | 200 - 1,000 m             | 25 - 100 (54$^{b}$)                  |
+| Eddy subduction        | 150 - 400 m                      | 150 - 400 m               | up to 150 (N/A)                      |
+| Large-scale subduction | 200 - 1,000 m                    | 200 - 1,000 m             | 25 - 100 (54$^{b}$)                  |
+| Diel migration         | 200 - 600 m                      | None                      | up to 250 (150$^{c}$)                |
+| Seasonal migration     | 600 - 1,400 m$^{a}$              | None                      | up to 500 (N/A)                      |
+
+: Range of depths generally reached by POC and DOC via the six BCPs [@Claustre2021-tx] and sequestration time potentials evaluated by the model of @Boyd2019-ez with an average estimation from the model of @Nowicki2022-no in brackets.
+None means that a pump does not transport DOC.
+N/A means that the pump was not evaluated in the model.
+(a) 1,400 m corresponds to the overwintering depth of *Calanus finmarchicus* as observed by @Jonasdottir2015-ob.
+(b) 54 years depicts the mean sequestration time of the mixing pump, without the contribution of the eddy subduction pump (c) Migration of mesopelagic fish was neglected in the model.
+{#tbl-PumpDepthRange}
+
+```{r, test table}
+# THERE IS NO CURRENT SOLUTION TO THIS ISSUE WITH PDF RENDERING UNFORTUNATELY, SEE https://github.com/rstudio/gt/issues/119
+# library(tibble)
+# library(gt)
+# library(gtsummary)
+# sequestration_table <- tibble('Pump' = c('Gravitational', 'Mixed layer', 'Eddy subduction', 'Large-scale subduction', 'Diel migration', 'Seasonal migration'),
+#                               'POC export/injection depth range' = c('export depth - sediment', '200 - 1000 m', '150 - 400 m', '200 - 1000 m', '200 - 600 m', '600 - 1400 m*'),
+#                               'DOC injection depth range' = c('None', '200 - 1000 m', '150 - 400 m', '200 - 1000 m', 'None', 'None'),
+#                               'Sequestration time potential (years)' = c('N/A (142)', '25 - 100 (54)**', 'up to 150 (N/A)', '25 - 100 (54)**', 'up to 250 (150)***', 'up to 500 (N/A)'))
+# 
+# sequestration_table |> tbl_summary() |> tab_source_note(source_note = md('1,400 m corresponds to the overwintering depth of *Calanus finmarchicus* as observed by (@Jonasdottir2015-ob)')) |> 
+#   #cols_width(starts_with('Pump') ~ px(100), starts_with('seq') ~ px(100))
+#   as_kable_extra()
+
+```
+
+A great majority of studies related to carbon sequestration by the BCP usually assumed that the part of the exported carbon reaching 1,000 m was sequestered [e.g.
+@Lampitt2008-kq; @Passow2012-zh].
+However, @Siegel2021-oz revises this common assumption by proposing that carbon sequestration can occur on the entire water column, with varying efficiencies (i.e.
+sequestration times).
+To come to this conclusion, @Siegel2021-oz used an ocean inverse circulation model to assess the time over which injected $CO_{2}$ (direct injections at different depths) remains sequestered from the atmosphere.
+For this, they computed a factor $f_{\tau}$ corresponding to the fraction of injected $CO_{2}$ that will be stored away from the atmosphere for a given time $\tau$.
+Their results show that $f_{\tau}$ increases with depth and varies geographically.
+They also show that for a time horizon of 100 years (i.e.
+$\tau = 100$) and an injection depth at 1,000 m, there is a non negligible leakage of $CO_{2}$ back to the atmosphere with strong regional disparities (e.g.
+high leakage in the Northwest Atlantic compared to virtually no leakage in the Central Eastern Pacific).
+
+So far, we have seen that significant incertitudes remain in the estimations of carbon export as well as the sequestration time by the different export pathways of the BCP.
+To address this issue, biogeochemical models and observation tools are continuously being improved to further our understanding of the BCP (see @sec-modelling-observing-BCP).

### 2023-03-28 uvp6_part.qmd
Very quick review of the methods

@@
 
-The UVP6 is the miniaturized version of the UVP5 [@Picheral2021-sx].
-It is smaller, it uses less energy and it has been specifically developed to be mounted on AUVs such as BGC-Argo floats with the aim of providing data of similar quality than from the UVP5.
+The UVP6 is the new version of the UVP5 [@Picheral2021-sx].
+It was made smaller and more energy efficient specifically to be mounted on AUVs such as BGC-Argo floats, while keeping the aim of providing data of quality similar to that from the UVP5.
 There are 2 versions: the LP (low power) and the HF (high frequency).
@@
 
-UVP5 Standard Definition (SD) profiles have been collected since 2008 at a global scale (@fig-map-of-profiles) whereas the UVP6 has been deployed very recently in specific locations and thus lacks spatial coverage to be fully representative.
-For this reason, UVP5 data will be used to create a preliminary ER model and explore ways of improvement.
-Ultimately, we will train the ER model for the UVP6 by applying the lessons learnt from the preliminary UVP5 model.
+UVP5 Standard Definition (SD) profiles have been collected since 2008 at a global scale (@fig-map-of-profiles) whereas the UVP6 has only been deployed very recently, in few locations and the resulting dataset lacks spatial coverage to be fully representative.
+Yet, the images of the two instruments have a similar aspect and they target the same size range.
+For this reason, UVP5 data will be used to create a preliminary EC model and explore ways for improvement.
+Ultimately, we will train the EC model on the growing UVP6 data, by applying the lessons learnt from the preliminary UVP5 model.
 
@@
 
-The full UVP5 SD data set is composed of 3.578.901 validated images whereas the full UVP6 LP data set has 634.459 validated images.
+The UVP5 SD data set is composed of 3.578.901 validated images whereas the full UVP6 LP data set has 634.459 validated images.
 Each image was segmented to find the main object (i.e.
@@
 Validated objects were manually classified into specific taxa belonging to well known biological groups such as *Rhizaria*, *Copepoda* or *Cnidaria* to name a few.
-However, the maximum number of classes (also referred to as labels or taxa) for the ER was fixed to 40 in order to limit the size of the ER model and the volume of data that needs to be transmitted.
-We chose to aggregate classes iteratively to find the best number of classes for the ER.
+However, the maximum number of classes (also referred to as labels or taxa) for the EC was fixed to 40 because that is the maximum that the frame of data transmitted by satellite allows.
+We chose to aggregate classes iteratively to find the best number of classes for the EC.
 The first level (i.e.
@@
 The validation set is usually used to assess the quality of a model by tuning hyperparameters while avoiding overfitting (stopping the training before the model is too acquainted with the data).
-Ideally, the test set is not to be used during any step except at the very end where it is used to assess the global quality of the final model.
+The test set is not to be used during any step except at the very end where it is used to assess the global quality of the final model.
 In our case, both data sets (UVP5 and UVP6) are very imbalanced towards few non-biological classes (mostly detritus).
@@
 
-One objective of the ER for the UVP6 is to be used on AUVs such as BGC-Argo floats to study the vertical migration pump [@Claustre2021-tx].
+One objective of the EC for the UVP6 is to be used on AUVs such as BGC-Argo floats to study the vertical migration pump [@Claustre2021-tx].
 Hence, we built the test set with entire UVP5 profiles going at least to 1,000 m depth, the typical parking depth of BGC-Argo floats.
@@
 
-GBT generates a sequence of trees, each one trying to cope with the errors made by their predecessors.
-XGBoost is designed to be an optimized and regularized version of GBT.
+GBT generates a sequence of trees, each one trying to cope with the residual error from its predecessors.
+XGBoost is an optimized and regularized version of GBT.
 To learn from the data and derive a model, XGBoost first minimizes a regularized objective function $L$
@@
 
-To build the ER model, the first step was a threefold cross-validation using XGBoost parameterized with a set of given hyperparameters (i.e.
+To build the EC model, the first step was a threefold cross-validation using XGBoost parameterized with a set of given hyperparameters (i.e.
 learning rate ($\eta$), maximum depth of a tree ($D_{max}$) and weight sensitivity ($\omega$)).
@@
 For this, we searched for the minimum value of the multiclass log loss (also known as cross-entropy loss) evaluated at each boosting round.
-Using that optimal $k$ value, we then trained the ER model with the entire training set.
+Using that optimal $k$ value, we then trained the EC model with the entire training set.
 Therefore, finding the best model consisted in finding the set of hyperparameters ($k$, $\eta$, $D_{max}$) that respected the energy constraints and returned the best global metrics.
@@
 To cope with our unbalanced data set (more than 70% of detritus in the UVP5 data set), several options were available.
-We could artificially increase the number of biological data with the well known SMOTE oversampling technique [@Chawla2002-kg; @Fernandez2018-ft], decrease the number of detritus (random undersampling, keep only detritus close to unsupervised cluster centroids, etc.) or add a weight to each label during training to nudge the model towards under-represented classes (i.e.
+We could artificially increase the number of biological data with the common Synthetic Minority Oversampling Technique (SMOTE) [@Chawla2002-kg; @Fernandez2018-ft], decrease the number of detritus (random undersampling, keep only detritus close to unsupervised cluster centroids, etc.) or add a weight to each label during training to nudge the model towards under-represented classes (i.e.
 biological classes).
-In order to compare a features-based classifier with a state-of-the-art CNN, we chose the latter and we defined the weight of a class as follows:
+After testing the various techniques, we chose the weight-based one, which also has the advantage or being directly comparable to the same implemntation in CNNs, and we defined the weight of a class as follows:
 
@@
 
-where $size~max$ and $size~class$ are, respectively, the size of the biggest class (i.e.
+where $size~max$ and $size~class$ are, respectively, the size (number of examples) of the most numerous class (i.e.
 detritus) and the size of each class.
@@
 We decided to select the best $\omega$ among 0 (no weights), 0.25, 0.50 and 0.75 (model very biased towards biological classes).
-In order to decipher between all options, we compared the predicted concentrations of classes for each $\omega$ with the real concentrations of each class in different regions of the global ocean.
+In order to choose between all options, we compared the predicted concentrations of classes for each $\omega$ with the real concentrations of each class in different regions of the global ocean.
 As always in machine learning, we encountered a trade off between precision and recall.
@@
 
-In order to compare our features-based model with a state-of-the-art neural network not limited by energy constraints, we used a MobileNetV2 (depth multiplier 1.40) pre-trained on ImageNet.
+In order to compare our features-based model with a state-of-the-art approach, not limited by energy constraints, we classified the datasets with a Convolutional Neural Network (MobileNetV2 , depth multiplier 1.40), pre-trained on ImageNet.
 This is known as transfer learning and it was already applied with success for zooplankton classification in @Orenstein2017-ya.
@@
 The classification layer at the end of the network also has a dropout of 0.2.
-The CNN was trained with a decaying learning rate, a categorical cross entropy loss function and weights were defined with the same weight sensitivity of 0.25.
+The CNN was trained with a decaying learning rate to facilitate a stable convergence towards a minimum of the loss.
+Finally, we used settings comparable to the GBT: a categorical cross entropy loss function and weights defined with the same weight sensitivity of 0.25.
 The CNN was trained with a sufficient number of epochs such that its training reached a validation plateau (i.e.
@@
 In our case, the optimal model hyperparameters were 0.2 for $\eta$ and 5 for D$_{max}$ (also with only 1 class of detritus, see Discussion) to have both a good weighted living (or biological) precision and recall and a loading time (i.e.
-loading of the ER model in memory) below 700 ms for energy reasons.
+loading of the EC model in memory) below 700 ms for energy reasons.
 In that configuration, the UVP5 model had an accuracy of 81%, a macro (i.e.
@@
 This reevaluation of $\pmb{P}$ is due to what is defined as a data set shift where part of the unseen data follows a different distribution than the one from the training data [@Orenstein2020-jl; @Moreno-Torres2012-oq]).
-It is especially true for ER systems deployed on AUVs where planktonic communities change throughout the year.
+It is especially true for EC systems deployed on AUVs where planktonic communities change throughout the year.
 Thus, we tested different quantification methods (CC, adjusted count (AC), threshold count (TC), probabilistic adjusted count (PAC) and scaled probability adjusted count (SPAC), see @Gonzalez2017-tr for a review on those quantification methods) assuming that the estimation of $\pmb{P}$ was accurate enough (\> 3M images in our training set) and that the global coverage of UVP5 profiles in the training set was sufficient to cover various planktonic distributions from the test set.
@@
 The SPAC showed the best estimations whereas TC drastically underestimated them (85% of data were lost using a probability threshold of 0.9).
-However, the aim of the ER is to be deployed on AUVs sampling vertical profiles in the water column therefore we applied the SPAC on 10.000 profile-like subsets and we summed all taxonomic abundances.
+However, the aim of the EC is to be deployed on AUVs sampling vertical profiles in the water column therefore we applied the SPAC on 10.000 profile-like subsets and we summed all taxonomic abundances.
 In that configuration, the SPAC overestimated the real total abundance by 1 to 15% depending on the projected (high or low) number of objects in each profile.
@@
 @fig-grid-search clearly shows that using more than 1 class of detritus drastically decreases the global accuracy of the model, decreases the weighted living recall and only increases by 1% the weighted living precision.
-In addition, the confusion matrix (@fig-confusion-matrix-uvp6-ndetritus) shows that it does not improve the ER model because there is a high confusion between subgroups of detritus.
+In addition, the confusion matrix (@fig-confusion-matrix-uvp6-ndetritus) shows that it does not improve the EC model because there is a high confusion between subgroups of detritus.
 Essentially, it seems that detritus form a continuum in the features space and that more research should be done on the choice of features, clustering algorithms and related methods to better differentiate them in separate entities, which is not the purpose of this study.
@@
 
-The aim of building an ER model was intended towards its implementation on the UVP6 for future deployments on BGC-Argo floats (or other profiling AUVs).
+The aim of building an EC model was intended towards its implementation on the UVP6 for future deployments on BGC-Argo floats (or other profiling AUVs).
 Following the work done with the UVP5 data set, we applied the same method on UVP6 data.
@@
 Therefore, we assume that those classes are well recognized by human classifiers so that their parent *Rhizaria* is virtually free from them.
-Afterwards, we executed a grid search to find the optimal couple $\eta$ and $D_{max}$ ($\omega$ and the number of classes of detritus were fixed based on UVP5 results) and trained the final ER model whose CM and CR are reported in @fig-confusion-matrix-final-model and @tbl-classification-report-uvp6-xgboost-cnn.
+Afterwards, we executed a grid search to find the optimal couple $\eta$ and $D_{max}$ ($\omega$ and the number of classes of detritus were fixed based on UVP5 results) and trained the final EC model whose CM and CR are reported in @fig-confusion-matrix-final-model and @tbl-classification-report-uvp6-xgboost-cnn.
 
@@
 
-The specifications of the final UVP6 ER model are the following: $\eta$: 0.2, $D_{max}$: 3, $\omega$: 0.25, number of boosting rounds: 400 for a total loading time of 73 ms.
+The specifications of the final UVP6 EC model are the following: $\eta$: 0.2, $D_{max}$: 3, $\omega$: 0.25, number of boosting rounds: 400 for a total loading time of 73 ms.
 Note that the CV initially found an optimal number of boosting rounds of 555 for a loading time of 101 ms).
@@
 
-4 BGC-Argo floats equipped with the ER were deployed in late May 2022 in the southern part of the Labrador Sea at the time of the spring bloom.
+4 BGC-Argo floats equipped with the EC were deployed in late May 2022 in the southern part of the Labrador Sea at the time of the spring bloom.
 As it can be seen in @fig-bgc-application-uvp6, *Calanoida* are more abundant around the deep chlorophyll maximum (DCM) and they might be observed in very low concentrations even below 1000 m.
@@
 
-The ER of zooplankton on board AUVs is just at its beginning.
+The EC of zooplankton on board AUVs is just at its beginning.
 As expected, the CNN trained on UVP6 data (see @tbl-classification-report-uvp6-xgboost-cnn) outperforms the results of our features-based classifier.
@@
 Nevertheless, previous tests on UVP5 data with more classes increased the gap between XGBoost and CNN weighted results therefore the latter better recognizes small classes.
-If energy and hardware constraints limit the performances of our current ER model, it is not forbidden to think about light-weight CNNs for future applications on AUVs to better understand the dynamics and the behavior of zooplankton.
+If energy and hardware constraints limit the performances of our current EC model, it is not forbidden to think about light-weight CNNs for future applications on AUVs to better understand the dynamics and the behavior of zooplankton.
 
-![Scheme of the UVP6 pipeline with the processing done in the lab to create the ER model and the processing time (in ms) of the main steps when the UVP6 is deployed on board an AUV.
+![Scheme of the UVP6 pipeline with the processing done in the lab to create the EC model and the processing time (in ms) of the main steps when the UVP6 is deployed on board an AUV.
 kpx means kilo pixel.](image/uvp6/fig_pipeline_UVP6.jpg){#fig-uvp-pipeline}
@@
 In order to see how the model could be improved, we compared the XGBoost model with results from a CNN not limited by energy.
-Afterwards, we applied the data processing pipeline and the recommendations from the UVP5 data set analysis on the UVP6 data set to create the final ER model.
+Afterwards, we applied the data processing pipeline and the recommendations from the UVP5 data set analysis on the UVP6 data set to create the final EC model.
 This model contains 20 labels, 16 biological and 4 non-living classes, with a total loading time of 73 ms.

### 2023-03-28 article3.qmd
update

@@
 
-Two pairs of BGC-Argo floats were deployed in the Labrador Sea in late May 2022.
+Two pairs of BGC-Argo floats were deployed in the Labrador Sea in late May 2022 (@fig-float-position).
 The first pair (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second pair (WMO 6904241 and 1902578) was deployed on the 29$^{th}$ at coordinates (-52.4°E, 56.8°N).
@@
 
-```{r, map of 4 floats, fig.height=6, fig.width=6}
+```{r, map of 4 floats, fig.height=5, fig.width=6}
 #| label: fig-float-position
@@
   annotation_north_arrow(location = "tr", which_north = "true") 
+```
+```{=latex}
+\newpage
 ```

### 2023-03-28 uvp6_part.qmd
update

--- 2023-03-27 uvp6_part.qmd
@@
 Building the zooplankton classification model for the UVP6 was the first task of my PhD.
-I knew nothing about machine learning before I started, so I spent the first three weeks reading a machine learning book to get familiar with the concepts.
-To be even more honest, I also knew nothing about zooplankton either, so there was something else to learn.
+To be honest, I knew nothing about machine learning before I started this so I spent the first three weeks reading a nice machine learning book in the corner of my team leader's office.
+To be even more honest, I basically knew nothing about zooplankton either.
+I think it is time to admit it.
 
-In machine learning, the "no free lunch theorem" [@Wolpert1996-hw] states that if you do not make any assumption about your data, then there is no model that is a priori guaranteed to work better [@Geron2019-is].
-In other words, it means that you need to get to know your data and your hands dirty to find a model that works.
-The first step consisted in developing a full data processing pipeline for UVP5 images.
-The aim was to be prepared for the UVP6 image data set (which was not ready at that time) to train the true, final, classification model.
+In machine learning, the no free lunch theorem [@Wolpert1996-hw] states that if you do not make any assumption about your data, then there is no model that is a priori guaranteed to work better [@Geron2019-is].
+In other words, it means that you need to get your hands dirty to find a model that works.
+The first step consisted in developing a full data processing pipeline using UVP5 images.
+The aim was to be prepared for the UVP6 image data set (not ready at that time) to train the true final classification model.
 To this end, I developed a python package called [UVPec](https://github.com/ecotaxa/uvpec) that is able to create a full classification model based on image folders with labeled particle/zooplankton images and some hyperparameters (e.g.
 learning rate).
-It made my life easier when I got the final UVP6 data set and should serve other future users in the same way.
+It really made my life easier when I got the final UVP6 data set and it saved me a lot of time.
 
-Classifying zooplankton on UVP images is difficult for human operators so building a model to automatically classify them is already a challenge, but adding technical and energy constraints to it was the icing on the cake.
+In the end, I am rather satisfied with the results.
+Classifying zooplankton is already difficult for human operators (at least, for me) so building a model to automatically classify them is one challenge but adding technical and energy constraints to it was the ice on the cake.
 Below is the first draft of a publication on the classification model.
@@
 Millions of images of oceanic particles have been acquired with ship-based imaging instruments.
-The identification of objects on images is time consuming for human operators but can be accelerated using artificial intelligence algorithms.
-Autonomous Underwater Vehicles (AUVs) such as Biogeochemical-Argo floats (BGC-Argo) have revolutionized our understanding of oceanic systems but have limited battery life and must spend as little time as possible at the ocean surface to transmit their data, which means that images cannot be transmitted.
-In this paper, we describe an image classification model embedded in the Underwater Vision Profiler 6 (UVP6) that is appropriate to be run on the limited power and bandwidth of AUVs.
-To this end, we computed simple global features of images of objects of interest and gradient boosted trees as a classifier.
-We compared its performance with a state-of-the-art convolutional neural network, which is much more energy demanding.
-In addition, we show how we dealt with class imbalance (\> 80% of images are marine snow) and what were the compromises to improve the classification scores of biological classes.
-Finally, we show the classification report (precision and recall for each label) of the final classification model and use it to analyze preliminary results from BGC-Argo floats deployed in the Labrador Sea.
+The task of labeling images is time consuming but necessary to train classification models using artificial intelligence algorithms.
+Autonomous Underwater Vehicles (AUVs) such as Biogeochemical-Argo floats (BGC-Argo) have revolutionized our understanding of oceanic systems but they are flawed by their battery life and by the time they can spend at the ocean surface to transmit their data.
+In this paper, we show how we developed the first zooplankton classification model for AUVs using the Underwater Vision Profiler 6 (UVP6).
+To this end, we used a features-based classifier called eXtreme Gradient Boosting (XGBoost) adapted to specific energy constraints imposed by profiling floats and we compared its performance with a state-of-the-art convolutional neural network as if energy was not a limiting factor.
+In addition, we show how we dealt with a highly imbalanced data set towards marine snow (\> 80% of detritus) and what were the compromises to improve the classification scores of biological classes.
+Finally, we show the classification report (precision and recall for each label) of the final classification model and use it to analyse some preliminary results from BGC-Argo floats deployed in the Labrador Sea.
 
@@
 Therefore, the model needs to be small and the classification should be easy and fast.
-Gradient Tree Boosting (GTB) seems to offer this compromise when compared to RF (model size) and SVM (computation complexity).
-For these reasons, we chose XGBoost [@Chen2016-xt], an optimized implementation of GTB, also known to be one of the best assets to win Kaggle's machine learning competitions.
+Gradient Boosting Tree (GBT) seems to offer this compromise when compared to RF (model size) and SVM (computation complexity).
+For these reasons, we chose XGBoost [@Chen2016-xt], an optimized implementation of GBT, also known to be one of the best assets to win Kaggle's machine learning competitions.
 
@@
 ![UVP5 and UVP6 profile coverage used in this study. UVP5 profiles used in the test set are depicted by green crosses.](image/uvp6/map_of_profiles.png){#fig-map-of-profiles}
+
 ```{=latex}
@@
 ```
-
 #### Data preparation
@@
 ```
-
 #### How to use the CR in real applications?
@@
 ```
-
 #### Prospects

### 2023-07-25 article3.qmd
corrections following reviews

@@
 In late May 2022, two pairs of BGC-Argo floats were deployed in the southern part of the LS at the time of the spring bloom.
-While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study the downward POC flux and its attenuation derived from beam attenuance [@Estapa2023-oz] and to study particle size distributions (PSDs) and concentrations with the UVP6.
+While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study the downward POC flux and its attenuation derived from beam attenuance [@Estapa2013-gw; @Estapa2019-nk] and to study particle size distributions (PSDs) and concentrations with the UVP6.
 
@@
 Finally, the vertical orientation of the transmissometer causes the accumulation of sinking material on the optical window.
-Therefore, a pump is cleaning the latter to remove the accumulated material between each drifting phase and subsequently avoid sensor drift.
+Therefore, a dedicated pump is used to clean the latter to remove the accumulated material between each drifting phase and subsequently avoid sensor drift.
 
@@
 # read floats data
-float_6904240 <- vroom('/home/flo/dox/ArgoShine/6904240_FromNetCDF.csv')
-float_6904241 <- vroom('/home/flo/dox/ArgoShine/6904241_FromNetCDF.csv')
-float_4903634 <- vroom('/home/flo/dox/ArgoShine/4903634_FromNetCDF.csv')
-float_1902578 <- vroom('/home/flo/dox/ArgoShine/1902578_FromNetCDF.csv')
+float_6904240 <- vroom('/home/flo/dox/ArgoShine/SAVE_FOR_MANUSCRIPT/6904240_FromNetCDF.csv')
+float_6904241 <- vroom('/home/flo/dox/ArgoShine/SAVE_FOR_MANUSCRIPT/6904241_FromNetCDF.csv')
+float_4903634 <- vroom('/home/flo/dox/ArgoShine/SAVE_FOR_MANUSCRIPT/4903634_FromNetCDF.csv')
+float_1902578 <- vroom('/home/flo/dox/ArgoShine/SAVE_FOR_MANUSCRIPT/1902578_FromNetCDF.csv')
+
+# float_6904240 <- vroom('/home/flo/dox/ArgoShine/6904240_FromNetCDF.csv')
+# float_6904241 <- vroom('/home/flo/dox/ArgoShine/6904241_FromNetCDF.csv')
+# float_4903634 <- vroom('/home/flo/dox/ArgoShine/4903634_FromNetCDF.csv')
+# float_1902578 <- vroom('/home/flo/dox/ArgoShine/1902578_FromNetCDF.csv')
 
@@
   geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 2) +
-  scale_color_brewer(palette = 'Dark2') + theme_bw() +
+  #scale_color_brewer(palette = 'Dark2') + 
+  theme_bw() +
+  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
   geom_point(data = transform_coord(argo_map_first), aes(lon, lat), shape = 17, size = 3) +
@@
 ```
+
 ```{=latex}
@@
 ```
-
 ### Data processing
@@
 Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: 10.48670/moi-00281) with a spatial resolution of 4 km.
-MLD derived from satellite observations were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052) with a spatial resolution of 0.25°.
+MLD data were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052) with a spatial resolution of 0.25°.
 
@@
 
+lpm_class <- c("NP_Size_102", "NP_Size_128", "NP_Size_161",
+                         "NP_Size_203", "NP_Size_256", "NP_Size_323",
+                         "NP_Size_406", "NP_Size_512", "NP_Size_645",
+                         "NP_Size_813", "NP_Size_1020", "NP_Size_1290",
+                        "NP_Size_1630", "NP_Size_2050")
+
 particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> dplyr::mutate(juld = as_date(juld))
@@
 
+particles_bis <- particles_bis |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)))
+
+facet_names_up <- c(`102` = '102 - 128 µm',
+                 `128` = '128 - 161 µm',
+                 `161` = '161 - 203 µm',
+                 `203` = '203 - 256 µm',
+                 `256` = '256 - 323 µm',
+                 `323` = '323 - 406 µm',
+                 `406` = '406 - 512 µm',
+                 `6904240` = '6904240',
+                 `6904241` = '6904241',
+                 `4903634` = '4903634',
+                 `1902578` = '1902578')
+
 up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
   #group_by(size) |>
-  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  ggplot() + facet_grid(wmo~size, labeller = as_labeller(facet_names_up)) + theme_bw() +
   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
@@
 
+facet_names_down <- c(`512` = '512 - 645 µm',
+                 `645` = '645 - 813 µm',
+                 `813` = '0.81 - 1.02 mm',
+                 `1020` = '1.02 - 1.29 mm',
+                 `1290` = '1.29 - 1.63 mm',
+                 `1630` = '1.63 - 2.05 mm',
+                 `2050` = '2.05 - 2.50 mm',
+                 `6904240` = '6904240',
+                 `6904241` = '6904241',
+                 `4903634` = '4903634',
+                 `1902578` = '1902578')
+
 down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
   #group_by(size) |>
-  ggplot() + facet_grid(wmo~size) + theme_bw() +
+  ggplot() + facet_grid(wmo~size, labeller = as_labeller(facet_names_down)) + theme_bw() +
   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
@@
 ```{r, uvp data}
-
-lpm_class <- c("NP_Size_102", "NP_Size_128", "NP_Size_161",
-                         "NP_Size_203", "NP_Size_256", "NP_Size_323",
-                         "NP_Size_406", "NP_Size_512", "NP_Size_645",
-                         "NP_Size_813", "NP_Size_1020", "NP_Size_1290",
-                        "NP_Size_1630", "NP_Size_2050")
 
@@
 
+facet_all <- c(`102` = '102 - 128 µm',
+                 `128` = '128 - 161 µm',
+                 `161` = '161 - 203 µm',
+                 `203` = '203 - 256 µm',
+                 `256` = '256 - 323 µm',
+                 `323` = '323 - 406 µm',
+                 `406` = '406 - 512 µm',
+                 `512` = '512 - 645 µm',
+                 `645` = '645 - 813 µm',
+                 `813` = '0.81 - 1.02 mm',
+                 `1020` = '1.02 - 1.29 mm',
+                 `1290` = '1.29 - 1.63 mm',
+                 `1630` = '1.63 - 2.05 mm',
+                 `2050` = '2.05 - 2.50 mm')
+
 # add blank
@@
   geom_smooth(aes(juld, mean_conc), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
+  #scale_color_brewer(palette = 'Dark2') + 
+  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
   scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
@@
   scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size, scales = 'free_y') #+
+  facet_wrap(~size, scales = 'free_y', labeller = as_labeller(facet_all)) #+
   #geom_vline(data = test, aes(xintercept = date_max))
@@
   scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size, scales = 'free_y') +
+  facet_wrap(~size, scales = 'free_y', labeller = as_labeller(facet_all)) +
   geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')
@@
   scale_y_continuous(trans = 'log10') +
-  facet_wrap(~size, scales = 'free_y') #+
+  facet_wrap(~size, scales = 'free_y', labeller = as_labeller(facet_all)) 
   #geom_vline(data = test, aes(xintercept = date_max))
@@
   geom_smooth(data = tmp, aes(x = date, y = small_flux), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
+  #scale_color_brewer(palette = 'Dark2') + 
+  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
@@
   geom_smooth(data = tmp, aes(x = date, y = large_flux), colour = 'black') +
-  scale_color_brewer(palette = 'Dark2') + 
+  #scale_color_brewer(palette = 'Dark2') + 
+  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
@@
 mld_data <- vroom('/home/flo/dox/ArgoShine/npq_corrected_data.csv')
+#filter date until March
+mld_data <- mld_data[mld_data$juld < "2023-03-05",]
 # clean and smoothed chla data
@@
   labs(x = 'Month', y = 'MLD (m)') + scale_y_reverse() +
-  scale_color_brewer(palette = 'Dark2') + 
+  #scale_color_brewer(palette = 'Dark2') + 
+  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + theme_bw() +
@@
 npq_data <- vroom('/home/flo/dox/ArgoShine/npq_corrected_data.csv')
+
+#filter date until March
+npq_data <- npq_data[npq_data$juld < "2023-03-05",]
 
@@
   #scale_colour_manual(values = c('#003366', '#E31B23'))
-  scale_color_brewer(palette = 'Dark2')
+  #scale_color_brewer(palette = 'Dark2') 
+  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77'))
 
@@
   theme_bw() + scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
-  scale_color_brewer(palette = 'Dark2') + theme(legend.position = 'none')
+  #scale_color_brewer(palette = 'Dark2') + 
+  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
+  theme(legend.position = 'none')
 
@@
 This pattern could result from several causes.
-It could either be the signature of a subduction event [@Llort2018-pa] either the end of a slow sinking export (i.e latent dynamics of the system) or a sign of fragmentation processes by zooplankton (i.e.
+It could either be the signature of a subduction event [@Llort2018-pa] either the end of a slow sinking export (i.e seasonal dynamics of the system) or a sign of fragmentation processes by zooplankton (i.e.
 zooplankton feeding).
@@
 @fig-spice shows that the concentration of *Calanoida* is higher from the end of May to July in the surface layer.
-There also seems to be a lower patch between 250 and 500 m from August to November however it is difficult to estimate their impact on particle concentrations at this stage.
+There also seems to be a lower patch between 250 and 500 m from August to November however it is difficult to estimate their impact on particle concentrations at this stage because more data are needed.
 Finally, we computed the spectral slope to look for aggregation/fragmentation processes.
 The spectral slope defines the slope of the PSD where a high (in absolute values) spectral slope indicates that small particules dominate the PSD.
-Here, the panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope at the beginning of the time series but it is also difficult to relate it to biological processes such as zooplankton feeding without any additional data.
+Here, the panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope (i.e.
+large particles dominate) at the beginning of the time series but it is also difficult to relate it to biological processes such as zooplankton feeding without any additional data.
 
@@
 #| label: tbl-correlation
-#| tbl-cap: Pearson correlations between UVP6 PSD-derived carbon fluxes and total, large and small particle carbon fluxes measured by the OST at 3 drifting depths.
+#| tbl-cap: Pearson correlations (all statistically significant with a p-value < 0.01) between UVP6 PSD-derived carbon fluxes and total, large and small particle carbon fluxes measured by the OST at 3 drifting depths.
 correlations <- comparison_flux_ost_uvp |> dplyr::select(park_depth, wmo, cycle, guidi_flux, small_flux, large_flux, date, total_flux) |> dplyr::group_by(wmo, park_depth, cycle) |> dplyr::mutate(median_guidi_flux = median(guidi_flux)) |> dplyr::group_by(park_depth, wmo) |> dplyr::summarize(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_large = round(cor(median_guidi_flux, large_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))
@@
 The current version of the OST has a housing above the optical window which, together with the float itself, can produce a "self-shading" effect [@Estapa2017-jr] that prevents the collection of fast sinking particles (assumed here to be large particles).
-In addition, the nature of particles is also a source of uncertainty in the attenuation signal.
+In addition, the nature of particles is also a source of uncertainty in the optical attenuation signal.
 Mathematically, the attenuation efficiency ($Q_{c}$) of a spherical particle can be defined [@Mobley2022-de] as

### 2023-07-25 discussion.qmd
corrections following reviews

@@
 
-[UVPec](https://github.com/ecotaxa/uvpec/) - short name for Underwater Vision Profiler Embedded Classifier - is a toolbox to train classification models for UVP6 images and evaluate their performance.
+[UVPec](https://github.com/ecotaxa/uvpec/) (https://github.com/ecotaxa/uvpec/) - short name for Underwater Vision Profiler Embedded Classifier - is a toolbox to train classification models for UVP6 images and evaluate their performance.
 This python package has been designed to be as user-friendly as possible in order to be accessible to the largest number of users.
@@
 In Figure 5 of @Guidi2008-vp, the light gray color area (low residual error) encompasses a wide region of $A$ and $B$ values where their jackknife procedure found local minima with ($A$, $B$) pairs such as (50, 2) or (150, 5).
-While $A$ is seen as a constant with no physical meaning, the value of $B$ should really be questioned since it is related to the fractal dimension $D$ of particles and aggregates.
+A is the single-particle carbon or mass flux for particles of size d$_{0}$ (i.e.
+a reference size bin) however it is mostly the values of $B$ that should really be questioned since it is related to the fractal dimension $D$ of particles and aggregates.
 Fractal dimensions should not exceed 3 (physical limit) but they have been estimated down to 1.1-1.3 [@Alldredge1988-lx].

### 2023-07-25 index.qmd
corrections following reviews

@@
 Since then, human have been burning fossil fuels at an unprecedented rate, releasing huge amounts of carbon dioxide ($CO_{2}$) in the atmosphere that should have been discharged slowly through the slow carbon cycle [@Bertrand2021-nh].
-These emissions, resulting from fossil emissions but also from land use (e.g.
+These emissions, resulting from the burning of fossil fuels but also from land use change (e.g.
 deforestation, land management) have been estimated around 700 Pg C since 1750 [@Canadell2021-ds].
@@
 It can be shown [e.g.
-@DeVries2022-of] that if we were to consider the ocean as a homogeneous reservoir of inorganic carbon, $pCO_{2}^{atm}$ would be much higher than it is now (roughly 416 parts per million (ppm) as of October 2022 according to the Mauna Loa Observatory).
-This observation is explained by the presence of a strong vertical gradient of DIC in the ocean, with surface DIC concentrations more than 200 $\mu$mol kg$^{-1}$ lower than at 1,000 m as it can be seen in @fig-DIC_plot.
+@DeVries2022-of] that if the ocean was a homogeneous reservoir of inorganic carbon, $pCO_{2}^{atm}$ would be much higher than it is now (roughly 416 parts per million (ppm) as of October 2022 according to the Mauna Loa Observatory).
+This is explained by the presence of a strong vertical gradient of DIC in the ocean, with deep DIC concentrations more than 200 $\mu$mol kg$^{-1}$ higher than at 1,000 m as it can be seen in @fig-DIC_plot.
 
@@
 
-This gradient is mostly due to processes often referred to as carbon pumps, operating against continuous ocean mixing processes that want to homogenize DIC in the water column [@Sarmiento2006-ec; @Claustre2021-tx].
+This gradient is mostly due to processes often referred to as carbon pumps, operating against continuous ocean mixing processes that tend to homogenize DIC in the water column [@Sarmiento2006-ec; @Claustre2021-tx].
 They were originally defined by @Volk1985-oc as "processes that deplete the ocean surface of DIC relative to the deep-water DIC".
@@
 
-$K_{0}$ is a function of temperature and salinity with a strong dependence on the former [@Sarmiento2006-ec] such that the solubility of $CO_{2}$ increases with decreasing temperatures.
+$K_{0}$ is a function of temperature and salinity with a strong dependence on the former [@Sarmiento2006-ec] such that the solubility of $CO_{2}$ increases with decreasing temperature.
 Consequently, waters that cool as they are transported to higher latitudes by the thermohaline circulation will be undersaturated in $CO_{2}$ once they reach the surface.
@@
 
-The solubility pump is responsible for the great majority of all the anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed DIC gradient according to @Sarmiento2006-ec or 35% according to @Toggweiler2003-ye.
-Both values are used in the literature (e.g.
-10% in @Boyd2019-ez and @Iversen2023-qj and 35% in @Riebesell2009-qj, @Passow2012-zh and @DeVries2022-of).
+The solubility pump is responsible for the great majority of all the anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed vertical DIC gradient according to @Sarmiento2006-ec.
 
@@
 The carbon flux leaving the surface layer for the ocean interior is referred to as the export flux [@Siegel2022-wd].
-Historically, the export depth was usually set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should now use the base of the euphotic zone (EZ, where there is enough light for net primary production (NPP)) as the export depth to clearly differentiate the layer where particulate organic carbon (POC) is produced (i.e.
+For convenience, the export depth is often set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should now use the base of the euphotic zone (EZ, where there is enough light for net primary production (NPP)) as the export depth to clearly differentiate the layer where particulate organic carbon (POC) is produced (i.e.
 via photosynthesis) and where POC is consumed (i.e.
@@
 This remineralization (or respiration) consists in the conversion of organic matter back to inorganic forms (i.e.
-nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd] such as zooplankton or fish.
+nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd] such as zooplankton and fish.
 
-Those two biological activities significantly enhance the DIC gradient such that the BCP would thus be responsible for the remaining 70% of the observed DIC gradient [@Sarmiento2006-ec].
-Without it, there would be no production of organic matter in the EZ (consuming DIC) and no remineralization in the ocean interior (producing DIC) resulting in a weaker DIC gradient.
-In fact, @DeVries2022-of calculated that such a decrease in the DIC gradient would reduce the ocean carbon inventory by roughly 10% and increase atmospheric $CO_{2}$ levels by a factor 2.
+Photosynthesis and respiration significantly enhance the DIC gradient such that the BCP would thus be responsible for the remaining 70% of the observed DIC gradient [@Sarmiento2006-ec].
+In an ocean without biology, there would be no production of organic matter in the EZ (consuming DIC) and no remineralization in the ocean interior (producing DIC) resulting in a weaker DIC gradient.
+In fact, @DeVries2022-of calculated that without the present DIC gradient, the ocean carbon inventory would be roughly 10% smaller and the atmospheric $CO_{2}$ levels twice as high.
 Similarly, @Kwon2009-ir showed that a deepening of the remineralization depth (i.e.
@@
 While this definition seems logical, it does not specify precisely what is meant by "climatically significant time scales".
-In what follows, I will use the 100-year time horizon to refer to carbon sequestration (unless specified otherwise) as used by other BCP studies [e.g.
-@Lampitt2008-kq; @Passow2012-zh].
+In what follows, I will use the 100-year time horizon to refer to carbon sequestration (unless specified otherwise) as used by some BCP studies [e.g.
+@Lampitt2008-kq; @Passow2012-zh; @National_Academies_of_Sciences2022-ah].
 
-The amount of carbon that is sequestered by the BCP mainly depends on its strength and its efficiency.
+The amount of carbon sequestered by the BCP mainly depends on its strength and its efficiency.
 In this manuscript, the strength of the BCP is defined as "the magnitude of the POC flux at a reference export depth" [@Claustre2021-tx] and its efficiency (also known as transfer efficiency) is defined as the fraction of POC flux transported to some reference depth below the export depth [@Buesseler2020-sn; @Claustre2021-tx].
@@
 
-The BCP is a natural-based solution for carbon sequestration by definition (@eq-photosynthesis).
+The BCP is a natural-based solution for carbon sequestration (@eq-photosynthesis).
 However, natural-based solutions for carbon sequestration are generally aligned with the concept of 'blue carbon' defined by @Ipcc2019-qc as "all biologically-driven carbon fluxes and storage in marine systems that are amenable to management" which usually encompasses only coastal ecosystems [@Ipcc2019-qc; @Lovelock2019-dr].
@@
 
-Nonetheless, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess atmospheric $CO_{2}$, released by anthropogenic activities.
+Nonetheless, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess atmospheric $CO_{2}$ released by anthropogenic activities.
 First, as explained in @Sarmiento2006-ec, atmospheric $CO_{2}$ levels were stable for several thousand years before the industrial revolution started, therefore it is assumed that the BCP was operating at steady state.
@@
 
-Unlike the three previous carbon pumps, the microbial carbon pump (MCP) was only introduced recently by @Jiao2010-ax as "a conceptual framework to address the role of microbial generation of recalcitrant dissolved organic matter[^intro-1] and relevant carbon storage".
+Unlike the three previous carbon pumps, the microbial carbon pump (MCP) was only introduced recently by @Jiao2010-ax as "a conceptual framework to address the role of microbial generation of recalcitrant dissolved organic matter[^index-1] and relevant carbon storage".
 In simple terms, it is a suite of marine microbial mechanisms that produce longer-lived DOC such as semi-refractory, refractory and ultra-refractory DOC that have an average lifetime of respectively 205, 16,000 and 40,000 years [@Legendre2015-jc].
 
-[^intro-1]: i.e. dissolved organic matter resistant to microbial utilization that can persist in the ocean interior for thousands of years.
+[^index-1]: i.e. dissolved organic matter resistant to microbial utilization that can persist in the ocean interior for thousands of years.
 
@@
 In order to fully understand the fate of DOC and DIC in the ocean, one would need to integrate both classic carbon pumps and the MCP.
-However, in this manuscript, I will only focus on the BCP, which was still considered as a single process until recently.
+However, in this manuscript, I will only focus on the BCP, which was considered as a single process until recently.
 
@@
 
-In the frame of the BCP, global carbon budgets have generally been imbalanced such that the exported POC was up to 2-3 orders of magnitude less abundant that it would need to be to sustain heterotrophic respiration in the mesopelagic zone and beyond [@Burd2010-hc; @Turner2015-af].
+In the frame of the BCP, global carbon budgets have generally been imbalanced such that the exported POC was up to 2-3 orders of magnitude less abundant than it would need to be to sustain heterotrophic respiration in the mesopelagic zone and beyond [@Burd2010-hc; @Turner2015-af].
 This imbalance raises one question: are we currently underestimating the export carbon flux or are we overestimating the heterotrophic respiration?
 
-The last part of the question is out of the scope of this manuscript but it should be said that a lot of uncertainties orbit around those carbon deficits, namely parameters used to estimate the metabolic activity of deep ocean organisms as well as deep food web dynamics and unquantified carbon sources such as the contribution of slowly settling particles and better estimation of POC fluxes [@Burd2010-hc; @Martin2020-wa].
+The last part of the question is out of the scope of this manuscript but it should be said that a lot of uncertainties surround those carbon deficits, namely parameters used to estimate the metabolic activity of deep ocean organisms as well as deep food web dynamics and unquantified carbon sources such as the contribution of slowly settling particles and better estimation of POC fluxes [@Burd2010-hc; @Martin2020-wa].
 
-Regarding the latter, @Boyd2019-ez published a paper in which they review the BCP by adding new biological and physical mechanisms[^intro-2] that inject suspended and sinking particles to depth (@fig-global_carbon_pumps).
+Regarding the latter, @Boyd2019-ez published a paper in which they review the BCP by adding new biological and physical mechanisms[^index-2] that inject suspended and sinking particles to depth (@fig-global_carbon_pumps).
 A better understanding of those mechanisms (see next subsections) could help to close the carbon budget.
 
-[^intro-2]: More specifically, as stated by @Claustre2021-tx "the word 'biological' in BCP refers to the transfer of biogenic material to depth, mediated by either biological or physical processes."
+[^index-2]: More specifically, as stated by @Claustre2021-tx "the word 'biological' in BCP refers to the transfer of biogenic material to depth, mediated by either biological or physical processes."
 
-![Illustration of the three main mechanisms through which the BCP exports carbon to the deep ocean.
+![Illustration of the three main mechanisms through which the BCP exports carbon downwards.
 1) Production of POC and transport to depth by gravitational settling.
@@
 
-The most well known pump is the gravitational carbon pump (BGP).
+The best known biological pump is the gravitational carbon pump (BGP).
 It consists in the gravitational settling (i.e.
-sinking) of POC[^intro-3] to the deep ocean and it is the main contributor (60-70%) of the export flux [@Boyd2019-ez; @Nowicki2022-no].
+sinking) of POC[^index-3] to the deep ocean and it is the main contributor (60-70%) of the export flux [@Boyd2019-ez; @Nowicki2022-no].
 Historically, the downward POC flux was directly measured using sediment traps [@Martin1987-um; @Honjo2008-kn] and indirectly measured using thorium-234 ($^{234}Th$) deficits as a POC flux proxy [e.g.
@@
 
-[^intro-3]: There obviously exists a gravitational settling of particulate inorganic carbon but in this manuscript, I focus on the BCP which was defined earlier as only focused on organic carbon as opposed to the carbonate pump.
+[^index-3]: There obviously exists a gravitational settling of particulate inorganic carbon but in this manuscript, I focus on the BCP which was defined earlier as only focused on organic carbon as opposed to the carbonate pump.
 
@@
 
-In all, the mixing pump would represent at least 20% of the export flux according to @Nowicki2022-no although large uncertainties still remain (see Figure 2C in the paper of @Boyd2019-ez).
+In all, the mixing pump would represent at least 20% of the export flux by the BCP according to @Nowicki2022-no although large uncertainties still remain (see Figure 2C in the paper of @Boyd2019-ez).
 
@@
 In the case of the diel (i.e.
-daily) migration pump, animals feed at the surface at night and then migrate to depth during daytime to avoid visual predation [@Siegel2022-wd], metabolizing injected food at depth[^intro-4], by-passing the EZ directly [@Steinberg2017-ri].
+daily) migration pump, animals feed at the surface at night and then migrate to depth during daytime to avoid visual predation [@Siegel2022-wd], metabolizing ingested food at depth[^index-4], by-passing the EZ directly [@Steinberg2017-ri].
 In the case of the seasonal migration pump, carbon-rich lipids accumulated in the EZ during summer are catabolized by hibernating copepods at depth, between 600 m and 1,400 m, where they shunt carbon below the permanent pycnocline [@Jonasdottir2015-ob].
@@
 
-[^intro-4]: They can also die at depth during the day (or their overwintering in the case of the seasonal migration pump) which also contributes directly to the export of carbon at depth.
+[^index-4]: They can also die at depth during the day (or their overwintering in the case of the seasonal migration pump) which also contributes directly to the export of carbon at depth.
 
-In his model, @Nowicki2022-no estimates that the VMP accounts for approximately 10% of the total export flux, albeit large uncertainties remain on the number of migrating animals and their migration depth.
+In his model, @Nowicki2022-no estimates that the VMP accounts for approximately 10% of the total export flux by the BCP, albeit large uncertainties remain on the number of migrating animals and their migration depth.
 The VMP would thus export less organic carbon than the mixing pump but its sequestration potential would be higher (@tbl-PumpDepthRange).
-In contrast to the mixing pump that mostly acts on DOC, migrating animals release rapid sinking fecal pellets that would reach depths such that subsequent remineralized DIC would be hidden from the surface for longer time periods (@tbl-PumpDepthRange).
+In contrast to the mixing pump that mostly acts on DOC, migrating animals release rapid sinking fecal pellets that would reach depths such that subsequent remineralized DIC would be removed from the surface for longer time periods (@tbl-PumpDepthRange).
 As a result, it is estimated that the diel migration pump could potentially sequester 60% as much carbon as the BGP [@Boyd2019-ez].
@@
 
-The Martin curve [@Martin1987-um] is probably the most well known POC flux attenuation curve (@eq-Martin-Curve and @fig-Martin-Curve). It is a power-law function described as follow
+The Martin curve [@Martin1987-um] is probably the best known POC flux attenuation curve (@eq-Martin-Curve and @fig-Martin-Curve). It is a power-law function described as follow
 
@@
 ![Illustration of the main factors influencing the sinking velocity following Stokes' Law and hypothetical attenuation fluxes (in % of export flux) with depth (vertical axis) for slow and fast sinking particles.
-The latter shows that slow sinking particles experience higher rates of flux attenuation due to increased duration of exposure to remineralization hence contributing less to carbon sequestration.
+The latter shows that slow sinking particles experience higher rates of flux attenuation due to increased duration of exposure to remineralization hence contributing less to deep carbon sequestration.
 Figure taken from @Williams2022-xx.](image/intro/sinking_speed.jpeg){#fig-sinking-speed}
@@
 It would make sense that the quadratic dependence of $w$ to the aggregate size would favor $d$ as the main driver.
-However, aggregates are far from being perfect theoretical objects for @eq-sinking-speed and the excess density as the main driver is plebiscited by some authors such as @Iversen2020-kb.
+However, aggregates are far from being perfect theoretical objects for @eq-sinking-speed and the excess density as the main driver is disputed by some authors such as @Iversen2020-kb.
 
@@
 | Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
-|------------------|------------------|------------------|-------------------|
+|------------------------|----------------------------------|---------------------------|--------------------------------------|
 | Gravitational          | export depth - sediment          | None                      | N/A (142)                            |
@@
 
-A great majority of studies related to carbon sequestration by the BCP usually assumed that the part of the exported carbon reaching 1,000 m was sequestered [e.g.
+A great majority of studies related to carbon sequestration by the BCP usually assumed that only the fraction of the exported carbon reaching 1,000 m was sequestered [e.g.
 @Lampitt2008-kq; @Passow2012-zh].
-However, @Siegel2021-oz revises this common assumption by proposing that carbon sequestration can occur on the entire water column, with varying efficiencies (i.e.
+However, @Siegel2021-oz revised this common assumption by proposing that carbon sequestration could occur on the entire water column, with varying efficiencies (i.e.
 sequestration times).
@@
 Their results show that $f_{\tau}$ increases with depth and varies geographically.
-They also show that for a time horizon of 100 years (i.e.
+They also showed that for a time horizon of 100 years (i.e.
 $\tau = 100$) and an injection depth at 1,000 m, there is a non negligible leakage of $CO_{2}$ back to the atmosphere with strong regional disparities (e.g.
@@
 
-So far, we have seen that significant incertitudes remain in the estimations of carbon export as well as the sequestration time by the different export pathways of the BCP.
+So far, we have seen that significant uncertainties remain in the estimation of carbon export as well as the sequestration time by the different export pathways of the BCP.
 To address this issue, biogeochemical models and observation tools are continuously being improved to further our understanding of the BCP (see @sec-modelling-observing-BCP).

### 2023-07-25 observingCarbonCycle.qmd
corrections following reviews

@@
 
-From the data we feed models with to uncertainties caused by missing or poorly understood processes, current BCP models still disagree on the estimation of carbon export both currently and in response to climate change [@Henson2022-hg].
-However, the ongoing increase in deployments of autonomous platforms and the development of new sensors is providing new opportunities to enhance the parametrization of key biogeochemical processes to better estimate carbon export and sequestration.
+Current BCP models still disagree on the estimation of the actual carbon export and its response to climate change [@Henson2022-hg].
+However, the ongoing increase in deployments of autonomous platforms and the development of new sensors is providing new opportunities to enhance the parametrization of key biogeochemical processes to reduce uncertainties and better estimate carbon export and sequestration.
 
@@
 
-In all, POC fluxes derived from sediment traps and $^{234}Th$ have their own uncertainties and errors thus their estimations of downward POC fluxes should not be seen as ground truth.
+In all, POC fluxes derived from sediment traps and $^{234}Th$ have their own uncertainties and errors thus their estimates of downward POC fluxes should not be seen as ground truth.
 In addition, they are labor-intensive and require ship support.
-Their relative contribution to the estimation of POC fluxes is thus expected to keep decreasing (@fig-sediment-trap-thorium-yearly-distribution) with the advent of alternative new methods promoted by an increasing use of automated underwater vehicles (AUVs).
+Their relative contribution to the estimation of POC fluxes is thus expected to keep decreasing (@fig-sediment-trap-thorium-yearly-distribution) with the advent of alternative methods promoted by an increasing use of autonomous underwater vehicles (AUVs).
+However, the AUV-derived estimates have also their own caveats (see @sec-BGC-sensors) therefore traditional methods to measure POC fluxes will still be needed to calibrate those from autonomous vehicles.
 
@@
 
-It is admitted that ship-based measurements are potentially the most precise.
+It is admitted that ship-based or fixed-point (e.g.
+moorings) measurements are potentially the most precise.
 However, they lack spatiotemporal coverage and they are usually not cost-effective.
@@
 
-In order to fill the scarcity of data from research vessels and moorings as well as coping with surface-constrained satellite measurements, AUVs have been developed to take in situ measurements in all weathers and at all times, only being dependent on their battery life and ocean currents (for drifting AUVs).
+In order to fill the scarcity of data from research vessels and moorings as well as coping with surface-constrained satellite measurements, AUVs have been developed to make in situ measurements in all weather conditions and seasons and at all times, only being dependent on their battery life and ocean currents (for drifting AUVs).
 
-Biogeochemical-Argo (BGC-Argo) profiling floats (@fig-bgc-float) are probably the most well known AUVs to study the BCP.
+Biogeochemical-Argo (BGC-Argo) profiling floats (@fig-bgc-float) are probably the best known platforms to study the BCP.
 They are cost-effective, able to monitor biogeochemical processes at a global scale throughout the year and under all weather conditions [@Chai2020-uq].
@@
 
-Gliders are also effective AUVs to study the BCP [e.g.
+Gliders are also effective platforms to study the BCP [e.g.
 @Omand2015-mf].
@@
 
-The BGC-Argo array has already demonstrated its ability to understand some aspects of the BCP [e.g.
+The BGC-Argo array has already demonstrated its ability to investigate some aspects of the BCP [e.g.
 @Lacour2017-ro; @Lacour2019-id; @Briggs2020-gz; @Yang2021-vg; @Llort2018-pa; @Wang2022-hn].
 However, the current array may lack some key data to further our understanding of the BCP.
-Therefore, other types of sensors could be added to help fill this gap.
+Therefore, other types of sensors are being added to help fill this gap.
 
@@
 #| label: fig-estapa-method
-#| fig-cap: Application of the method of @Estapa2023-oz on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component, cleaned from spikes and negative jumps (assumed as not related to carbon flux) and a discontinuous (positive jumps, light blue) component.
+#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component, cleaned from spikes and negative jumps (assumed as not related to carbon flux) and a discontinuous (positive jumps, light blue) component.
 Data from BGC-ARGO float WMO 4903634, drifting at 1000 m during its 9$^{th}$ cycle.
@@
 ```
-
 The UVP6 (@fig-UVP6) is a miniaturized version of the UVP5 [@Picheral2021-sx], an imaging system developed to characterize particles (e.g.
 size, gray level, taxonomic group).
-Two versions of the UVP6 are available: the UVP6-HF (high frequency) and the UVP6-LP (low power) which has been specifically designed for AUVs that have strict energy requirements (e.g.
+Two versions of the UVP6 are available: the UVP6-HF (high frequency), and the UVP6-LP (low power) which has been specifically designed for AUVs that have strict energy requirements (e.g.
 autonomous floats).
@@
 ```
-
-The UVP6-LP (hereafter referred to as UVP6) is thus both a particle counter (0.1 mm - 2.5 mm) and a particle classifier (\> 0.6 mm).
+The UVP6-LP (hereafter referred to as UVP6) is thus both a particle counter for particles in the 0.1 - 2.5 mm size range and a particle classifier (\> 0.6 mm).
 The former counts particles in different size classes in order to compute the PSD while the latter extracts pixel-related and geometric features from each image to classify them (i.e.
@@
 
-The deployment of autonomous platforms equipped with a UVP6 and the zooplankton recognition algorithm for the study of the BCP is just beginning (first deployment in May 2022 in the Labrador Sea, see @sec-case-study-labrador-sea).
-On the contrary, the use of PSD in carbon flux studies has already been thoroughly explored [@Guidi2008-vp; @Iversen2010-an; @Jouandet2011-bt; @McDonnell2012-ic; @Roullier2014-wf; @Guidi2015-sj; @Guidi2016-nj; @Ramondenc2016-nu; @Kiko2017-qy; @Fender2019-ou; @Giering2020-mc; @Cael2021-qn; @Bisson2022-bt].
+The deployment of autonomous platforms equipped with a UVP6 and a zooplankton classification algorithm for the study of the BCP is just beginning (first deployment in May 2022 in the Labrador Sea, see @sec-case-study-labrador-sea).
+On the contrary, PSDs in carbon flux studies have been extensively used [@Guidi2008-vp; @Iversen2010-an; @Jouandet2011-bt; @McDonnell2012-ic; @Roullier2014-wf; @Guidi2015-sj; @Guidi2016-nj; @Ramondenc2016-nu; @Kiko2017-qy; @Fender2019-ou; @Giering2020-mc; @Cael2021-qn; @Bisson2022-bt].
 
@@
 The assumption that all particles of a given size class have the same carbon content and settling rate questions the formulation of the model itself (@eq-carbon-flux-finite).
-It has been shown that the pool of sinking particles is varied (e.g.
+It has been shown [@Durkin2021-nq] that the pool of sinking particles is varied (e.g.
 compact fecal pellets, porous aggregates) and that the main driver of the sinking speed is currently subject to debate [@sec-sinking-speed].
-Therefore, the size-dependent model of @Guidi2008-vp could be improved by taking into account the composition and nature of particles and by increasing the number of paired measurements between UVP PSD-derived flux estimations and other direct POC flux estimations (e.g.
+Therefore, the size-dependent model of @Guidi2008-vp could be improved by taking into account the composition and nature of particles [@Durkin2021-nq] and by increasing the number of paired measurements between UVP PSD-derived flux estimates and other direct POC flux estimates (e.g.
 sediment traps, $^{234}Th$, OST).
@@
 1.
-BGC-Argo floats equipped with a UVP6 and the zooplankton recognition algorithm are able to classify particles larger than 0.6 mm.
+BGC-Argo floats equipped with a UVP6 and a zooplankton classification algorithm are able to classify particles larger than 0.6 mm.
 However, these AUVs are limited in terms of their battery life.
-They are typically deployed for several years, which requires special attention to the energy consumption of each sensor.
-In @sec-uvp6, I will therefore describe the methodology used to develop the zooplankton classification model for the UVP6 accounting for energy limitations.
+They are typically deployed for several years, which requires to minimize the energy consumption of each sensor.
+In @sec-uvp6, I will therefore describe the methodology used to develop a zooplankton classification model for the UVP6 accounting for energy limitations.
 2.

### 2023-07-25 uvp6_part.qmd
corrections following reviews

@@
 
-```{r, confusion matrix N detritus, fig.height=4, fig.width=6}
+```{r, confusion matrix UVP6, fig.height=4, fig.width=6}
 #| label: fig-confusion-matrix-final-model

### 2023-07-26 article3.qmd
final corrections for SU and Uliège

@@
 PSD) was divided by the number of images at each depth and the water volume imaged by the UVP (0.7 L) to obtain particle concentrations (units: #/L).
-Note that in an impending update by the Coriolis data center, it will no longer be necessary to divide by the number of images.
 
@@
 
-### Derivation of carbon fluxes from OST
+### Derivation of carbon fluxes from the OST
 
@@
 
-```{r, UVP6 profiling mode without sinking speed small classes, fig.height=8, fig.width=8}
+```{r, UVP6 profiling mode without sinking speed small classes, fig.height=8, fig.width=10}
 #| label: fig-uvp6-profiling-102-406
@@
 
-```{r, UVP6 profiling mode without sinking speed large classes, fig.height=8, fig.width=8}
+```{r, UVP6 profiling mode without sinking speed large classes, fig.height=8, fig.width=10}
 #| label: fig-uvp6-profiling-512-2500
@@
 The same behavior is observed for UVP6 particles in the lower size range at 1000 m (@fig-uvp-1000m).
-Observing similar properties across a wide geographical range around deployments suggests that we observe the latent dynamics of the system rather than episodic export events.
+Observing similar properties across a wide geographical range around deployments suggests that we observe the latent (i.e.
+seasonal) dynamics of the system rather than episodic export events.
 
@@
 different bloom dynamics), they merged back after the bloom of May.
-Therefore, we suggest that we mostly observe the latent dynamics of the system rather than clear export signatures that would likely be spotted with both higher bloom intensities (e.g.
+Therefore, we suggest that we mostly observe the seasonal dynamics of the system rather than clear export signatures that would likely be spotted with both higher bloom intensities (e.g.
 the May bloom that was missed) and a higher profiling temporal resolution.
@@
 
+```{=latex}
+\newpage
+```
+
 ```{r, correlations between OST and UVP carbon fluxes}
@@
 during the spring bloom).
-Similarly, to address the question of the vertical migration, a 10-day temporal resolution is way to low because plankton images are rare compared to detritus so the sampling volume should be greatly increased to have a better estimation of the presence (or absence) of zooplankton.
+Similarly, to address the question of the vertical migration, a 10-day temporal resolution is way too low because plankton images are rare compared to detritus so the sampling volume should be greatly increased to have a better estimation of the presence (or absence) of zooplankton.
 Finally, it would perhaps be interesting to increase the time spent at 200 m (to 3 days) and decrease the time spent at 1000 m (to 3 days instead of 5) where the probability of capturing big particles is lower than at 200 m.

### 2023-07-26 discussion.qmd
final corrections for SU and Uliège

@@
 This result is threefold: 1) in our case, there is no need to use a trust index in addition to the probabilities given by `multi:softprob`, 2) some classes kept accumulating errors in the prediction with increasing rank (e.g.
-*Copepoda*, darksphere and fiber\<detritus) thus an increasing rank is not always a trustworthy indicator of increasing confidence 3) it partly explains why the quantification did not improve the overall classification by either using a fixed probability threshold (i.e.
+*Copepoda*, darksphere and fiber\<detritus) thus an increasing rank is not always a trustworthy indicator of increasing confidence, 3) it partly explains why the quantification did not improve the overall classification by either using a fixed probability threshold (i.e.
 TC) or using the probability matrix (i.e.
@@
 This work has shown the first comparison of BGC-Argo floats equipped with both the UVP6 and an OST to better quantify carbon fluxes from particle counts and beam attenuation.
-If the prediction of carbon fluxes at global scale was not a success, this work has highlighted current difficulties and ways to more forward using machine learning approaches.
+If the prediction of carbon fluxes at global scale was not a success, this work has highlighted current difficulties and ways to move forward using machine learning approaches.
 The deployment of BGC-Argo floats in the frame of REFINE with the combination of UVP6 measurements (particle counts and identification), OST beam attenuation fluxes as well as bio-optical variables (e.g.

### 2023-07-26 front_page.tex
final corrections for SU and Uliège

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{titlesec}
\usepackage{geometry} % Added for adjusting the page margins

\titleformat{\chapter}[display]
{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{20pt}{\Huge}

% Added for adjusting the page margins
\geometry{
    left=3cm,
    right=3cm,
    top=1cm,
    bottom=3cm
}

\pagenumbering{gobble}

\begin{document}

\begin{titlepage}
    \begin{center}
        \includegraphics[width=0.3\textwidth]{FRS-FNR_S2-PANT_UK_CS.png} % Insert your university logo here
        \vspace{1cm}

        \LARGE \textbf{University of Liège}\\
        \LARGE \textbf{Sorbonne University}\\

        \vspace{1.5cm}
        %{\huge \textbf{Doctoral Thesis}}\\[0.5cm]

        \hrulefill \\
        \vspace{0.5cm}
        {\huge \textbf{Towards a new insight of the carbon transport in the global ocean}} \\
        \hrulefill

        \vspace{1.5cm}
        {\Large \textbf{Florian Ricour}}

        \vspace{1cm}

        Dissertation presented in fulfillment for the degree of Doctor of Sciences

        \vfill

        \begin{flushleft}
            \large
            \textbf{Supervisors}\\
            \vspace{0.1cm}
            Hervé Claustre, director\\
            Marilaure Grégoire, director\\
            Alexander Barth, co-director\\
        \end{flushleft}

        \begin{flushleft}
            \large
            \textbf{Jury Members}\\
            \vspace{0.1cm}
            \begin{tabular}{lll}
                Pr. Stephanie Henson & University of Southampton & Reviewer\\
                Dr. Margaret Estapa & University of Maine & Reviewer \\
                Dr. Griet Neukermans & University of Ghent & Examiner \\
                Dr. Ivona Cetini\'{c} & Morgan State University & Examiner\\
                Pr. Lars Stemmann & Sorbonne University & Examiner\\
                Dr. Lionel Guidi & Sorbonne University & Guest\\
                Dr. Hervé Claustre & Sorbonne University & PhD Director\\
                Pr. Marilaure Grégoire & University of Liège & PhD Director\\
                Dr. Alexander Barth & University of Liège & PhD Co-director\\
            \end{tabular}
        \end{flushleft}

        \vspace{1cm}
        {\large May 2023} % Replace with the month and year of submission

    \end{center}
\end{titlepage}

\newpage
\thispagestyle{empty}
\null

\newpage
\thispagestyle{empty}
\null

\section*{Foreword}

As of 2022, it has now become obvious to the general public that the Earth's climate is changing. Massive flood events, repeated heatwaves, never-ending droughts, disappearing glaciers, biodiversity collapse, … Reports on weather anomalies were already published in the 1960's, the Intergovernmental Panel on Climate Change (IPCC) was created in 1988, the Kyoto Protocol was adopted in 1997, An Inconvenient Truth was released in 2006, the Paris Agreement was signed in 2015 and still, what has really changed?\\
\\
As I am writing this part, the COP27 (sponsored by Coca-Cola, no joke) is taking place with an acknowledged record number of fossil fuel lobbyists, bloody stadiums with outside air conditioning are about to be used in the desert for a month and it seems like 'leaders' of the Free World have been having hearing issues. See for yourself:\\
\\
\textit{"I am drowning"} said Pakistan\\
\\
\textit{"I'm sorry, what?"} said the Free World\\
\\
\textit{"I am starving"} said Ethiopia\\
\\
\textit{"Can't hear you, louder please!"} said the Free World\\
\\
\textit{"We are suffocating"} said India\\
\\
\textit{"Aaaah okay. Yes you're absolutely right, let's save the banks one more time"} said the Free World\\
\\
Although I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured that there is no fake news in this work. I hope you will enjoy it.

\newpage
\thispagestyle{empty}
\null

\newpage
\thispagestyle{empty}
\null

\section*{Acknowledgements}

I would like to thank the members of the jury for taking the time to read and evaluate this three and a half year work. I look forward to discussing it during the defense.\\
\\
Merci aux membres du comité de thèse pour leur temps et leurs retours sur l'évolution du projet.\\
\\
Cette thèse a été particulière à bien des égards, avec son lot de points positifs et négatifs mais surtout avec son lot de rebondissements qui m'ont permis d'affûter mes connaissances, ma résilience et je l'espère ma tolérance. Pour cela, je tiens à remercier chaleureusement toutes les personnes y ayant contribué de près ou de loin.\\
\\
Merci à Hervé, Lionel et Jean-Olivier pour leur accueil au Laboratoire de Villefranche-sur-Mer et pour leur disponibilité ainsi que pour leurs conseils avisés, tant scientifiques que personnels.\\
\\
Merci à Alex, Charles et Arthur pour leur aide et leur gentillesse.\\
\\
Merci également aux personnes que j'ai pu côtoyer à COMPLEX comme chez OMTAB depuis mon passage en stage en 2018 et les multiples allers-retours Belgique-France.\\
\\
Merci à Antoine Mangin de m'avoir prêté un bureau dans les locaux d'ACRI sur Grasse l'été dernier pendant que mon appartement servait d'Airbnb estival pour le plus grand bonheur du portefeuille de mon proprio.\\
\\
Un grand merci à mes colocs de bureau: Ophélie, Laetitia et Thelma pour votre soutien et votre bonne humeur sans faille (contrairement à moi).\\
\\
Merci à AMP (elle se reconnaitra), Chloé et Anaïs pour les pauses midi sur le ponton. AMP, merci d'avoir accepté de vivre virtuellement avec moi afin que la sécu daigne enfin me donner une carte vitale !\\
\\
Ne me demandez pas des noms de restos ou de bars sur la côte d'Azur, je n'en ai aucune idée. Par contre, je connais pas mal les routes du coin. Merci à tous ceux qui ont partagé (ou subi) des kilomètres à mes côtés, surtout en montée (le plat pays, qui est le mien): Flavien, Ophé, Lucas, Thelma, Laeti et Satoshi.\\
\\
Un grand merci à Marine et Alice pour leur bienveillance. Marine, je te souhaite le meilleur pour tes projets futurs mais fais attention à qui tu donnes tes clés quand même.\\
\\
Merci également à Alex qui a été un excellent roommate pendant les confinements, couvre-feux et ordonnances en tout genre. See you in London !\\
\\
Merci aux Liégeois(es) (Thomas, Florent, Flop, Nico, Evgeny et Lucie) et en particulier Hadrien qui a accepté de m'avoir comme coloc pendant plusieurs mois (et qui y a survécu).\\
\\
Merci aux Montois (Mathieu, Chris, Flo, Ju \& Ju) pour votre soutien à distance, un mariage en 3 actes et pour vos visites sur Nice. Petite dédicace à Ju K qui, à force d'avoir joué avec le diable en subissant l'enfer des pourcentages, va littéralement en devenir un. Vivement 2024 pour casser la baraque !\\
\\
Un très grand merci à ma famille nucléaire qui ne comprend toujours pas mon job mais qui fait semblant quand je lui explique, c'est le principal. C'est vrai que la défense tombe mal au niveau timing donc on se voit à la Ducasse :-)\\
\\
Enfin, un grand merci à une personne toute particulière qui en a probablement marre que je parle de la Belgique mais qui à présent est parfaite bilingue, mission accomplie?

\newpage
\thispagestyle{empty}
\null

\newpage
\thispagestyle{empty}
\null

\section*{Summary}

\textbf{Towards a new insight of the carbon transport in the global ocean}\\
\\
The ocean is known to play a key role in the carbon cycle. Without it, atmospheric CO$_{2}$ levels would be much higher than they are today thanks to the presence of carbon pumps that maintain a gradient of dissolved inorganic carbon (DIC) between the surface and the deep ocean. The biological carbon pump (BCP) is primarily responsible for this gradient. It consists in a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis in sunlit surface waters and then transported to the ocean interior and possibly the sediment where it will be sequestered from the atmosphere for millions of years. The BCP was long thought as solely the gravitational settling of particulate organic carbon (POC). However, a new paradigm for the BCP has recently been defined in which physically and biologically mediated particle injection pumps have been added to the original definition. Physically mediated particle injection pumps provide a pathway to better understand the transport of dissolved organic carbon (DOC) whereas biologically mediated particle injection pumps focus on the transport of POC by vertically migrating animals, either daily or seasonally. Therefore, a better understanding of these processes could help bridge the gap between carbon leaving the surface and carbon demand in the ocean interior. To address this new paradigm, this work will benefit from the advent of recent sensors that equip a new generation of Biogeochemical-Argo floats (BGC-Argo). The first part focuses on the development of an embedded zooplankton classification model for the Underwater Vision Profiler 6 (UVP6) under strict technical and energy constraints. The second part studies particle and carbon fluxes in the Labrador Sea using BGC-Argo floats equipped for the first time with the UVP6 and an optical sediment trap (OST), providing two independent measurements of sinking particles. The last part consists in revisiting the BCP using a new framework called CONVERSE for Continuous Vertical Sequestration. With this new approach, we re-evaluate the total carbon sequestered from the atmosphere ($\geq$ 100 years) by the BCP and its transport pathways on the entire water column, in contrast to the carbon sequestration typically assumed below a fixed reference depth.\\
\\
\textbf{Keywords} Carbon pumps -- Underwater Vision Profiler 6 -- BGC-Argo floats -- Machine learning

\newpage

\section*{Résumé}

\textbf{Vers une meilleure compréhension du transport du carbone dans l'océan global}\\
\\
L'océan est connu pour jouer un rôle clé dans le cycle du carbone. Sans lui, les niveaux de CO$_{2}$ atmosphérique seraient bien plus élevés qu'aujourd'hui grâce à la présence de pompes à carbone qui maintiennent un gradient de carbone inorganique dissous (DIC) entre la surface et le fond de l'océan. La pompe à carbone biologique (BCP) est principalement responsable de ce gradient. Elle consiste en une série de processus océaniques au cours desquels le carbone inorganique est converti en matière organique via la photosynthèse dans les eaux de surface, puis transporté vers l'intérieur de l'océan et éventuellement les sédiments où il sera séquestré par rapport à l'atmosphère pour des millions d'années. La BCP était longtemps considérée comme étant uniquement la déposition gravitationnelle de carbone organique particulaire (POC). Cependant, un nouveau paradigme pour la BCP a récemment été défini dans lequel des pompes d'origine physique et biologique d'injection de particules ont été ajoutées à la définition originale. Les pompes physiques d'injection de particules fournissent un moyen de mieux comprendre le transport de carbone organique dissous (DOC), tandis que les pompes biologique d'injection de particules se concentrent sur le transport de POC par des animaux migrant verticalement, quotidiennement ou saisonnièrement. Par conséquent, une meilleure compréhension de ces processus pourrait aider à combler l'écart entre le carbone quittant la surface et la demande de carbone dans l'océan profond. Pour aborder ce nouveau paradigme, ce travail bénéficiera de l'arrivée de capteurs récents équipant une nouvelle génération de flotteurs Biogéochimiques-Argo (BGC-Argo). La première partie se concentre sur le développement d'un modèle embarqué de classification de zooplancton pour l'Underwater Vision Profiler 6 (UVP6) avec des contraintes techniques et énergétiques strictes. La deuxième partie étudie les flux de particules et de carbone dans la mer du Labrador en utilisant des flotteurs BGC-Argo équipés pour la première fois de l'UVP6 et d'un piège à sédiments optique (OST), fournissant deux mesures indépendantes des particules. La dernière partie consiste à revisiter la BCP en utilisant un nouveau cadre appelé CONVERSE qui fait référence à la séquestration verticale continue du carbone dans la colonne d'eau. Avec cette nouvelle approche, nous réévaluons le carbone total séquestré par rapport à l'atmosphère ($\geq$ 100 ans) par la BCP et ses voies de transport sur toute la colonne d'eau, par opposition à la séquestration de carbone généralement supposée en-dessous d'une profondeur de référence fixe.\\
\\
\textbf{Mots clés} Pompes à carbone -- Underwater Vision Profiler 6 -- Flotteurs BGC-Argo -- Machine learning

\newpage
\thispagestyle{empty}
\null

\end{document}

### 2023-07-26 index.qmd
final corrections for SU and Uliège

@@
 These emissions, resulting from the burning of fossil fuels but also from land use change (e.g.
-deforestation, land management) have been estimated around 700 Pg C since 1750 [@Canadell2021-ds].
+deforestation, land management), have been estimated around 700 Pg C since 1750 [@Canadell2021-ds].
 
@@
 
-The solubility pump is responsible for the great majority of all the anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed vertical DIC gradient according to @Sarmiento2006-ec.
+The solubility pump is responsible for the great majority of all anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed vertical DIC gradient according to @Sarmiento2006-ec.
 
@@
 | Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
-|------------------------|----------------------------------|---------------------------|--------------------------------------|
+|-----------------|-------------------|-----------------|---------------------|
 | Gravitational          | export depth - sediment          | None                      | N/A (142)                            |

### 2023-07-26 observingCarbonCycle.qmd
final corrections for SU and Uliège

@@
 To estimate the downward POC flux, one therefore needs to quantify the amount of POC collected through a given area over a known time period at a given depth.
-However, the apparent sinking flux measured with sediment traps can be biased due to hydrodynamic issues that could result in over- or under-collection of particles (especially at the surface) or swimmers (zooplankton, small fish) that could enter sediment traps and affect the collected material and solubilization of collected material (e.g.
+However, the apparent sinking flux measured with sediment traps can be biased due to hydrodynamic issues that could result in over- or under-collection of particles (especially at the surface) or swimmers (zooplankton, small fish) that could enter sediment traps and affect the collected material or solubilization of collected material (e.g.
 by bacteria) [@Buesseler2007-ql].
@@
 #| label: fig-estapa-method
-#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component, cleaned from spikes and negative jumps (assumed as not related to carbon flux) and a discontinuous (positive jumps, light blue) component.
+#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component, cleaned from spikes and negative jumps (assumed as not related to the carbon flux) and a discontinuous (positive jumps, light blue) component.
 Data from BGC-ARGO float WMO 4903634, drifting at 1000 m during its 9$^{th}$ cycle.
@@
 It has been shown [@Durkin2021-nq] that the pool of sinking particles is varied (e.g.
-compact fecal pellets, porous aggregates) and that the main driver of the sinking speed is currently subject to debate [@sec-sinking-speed].
+compact fecal pellets, porous aggregates) and that the main driver of the sinking speed is currently subject to debate (@sec-sinking-speed).
 Therefore, the size-dependent model of @Guidi2008-vp could be improved by taking into account the composition and nature of particles [@Durkin2021-nq] and by increasing the number of paired measurements between UVP PSD-derived flux estimates and other direct POC flux estimates (e.g.
@@
 BGC-Argo floats equipped with a UVP6 and a zooplankton classification algorithm are able to classify particles larger than 0.6 mm.
-However, these AUVs are limited in terms of their battery life.
+However, these AUVs are limited by their battery life.
 They are typically deployed for several years, which requires to minimize the energy consumption of each sensor.

### 2023-07-26 uvp6_part.qmd
final corrections for SU and Uliège

@@
 In the end, I am rather satisfied with the results.
-Classifying zooplankton is already difficult for human operators (at least, for me) so building a model to automatically classify them is one challenge but adding technical and energy constraints to it was the ice on the cake.
+Classifying zooplankton is already difficult for human operators (at least, for me) so building a model to automatically classify them is one challenge but adding technical and energy constraints to it was the icing on the cake.
 Below is the first draft of a publication on the classification model.
 
-## Embedded recognition of zooplankton using the Underwater Vision Profiler 6 for use on autonomous platforms
+## Embedded classification of zooplankton using the Underwater Vision Profiler 6 for use on autonomous platforms
 
@@
 Usually, images are brought back to the lab and classified with CNNs which, in comparison with features-based classifiers, that is machine learning algorithms using a set of features extracted from the image (e.g.
-area, eccentricity, gray levels) have outperformed the latter for plankton recognition [@Ellen2019-nh].
+area, eccentricity, gray levels) have outperformed the latter for plankton classification [@Ellen2019-nh].
 Though, it comes at a cost.
@@
 Typically, a BGC-Argo float samples the water column between the surface and 1,000 m, it would therefore have to stay at the surface for a time long enough to transmit all images collected after each profile, consuming energy proportionally to the size of the image set while being subject to biofouling or unpredictable hazards.
-For all these reasons, it was decided to use a features-based classifier for the development of the embedded plankton recognition algorithm on board the UVP6.
+For all these reasons, it was decided to use a features-based classifier for the development of the embedded plankton classification algorithm on board the UVP6.
 
 In the literature, such algorithms (e.g.
-Decision Tree (DT), Random Forest (RF), Support Vector Machine (SVM)) have already been used for plankton recognition (see Table 1 of @Gonzalez2017-tr for an exhaustive list).
+Decision Tree (DT), Random Forest (RF), Support Vector Machine (SVM)) have already been used for plankton classification (see Table 1 of @Gonzalez2017-tr for an exhaustive list).
 For instance, @Blaschko2005-ep obtained an accuracy of 71% using a SVM on a set of 982 images among 13 classes using a 10-fold cross-validation (CV), @Sosik2007-cj obtained a global accuracy of 88% using a SVM on a test set of 3300 images covering 22 classes, @Bell2008-yo obtained 82% of accuracy using a 10-fold CV RF algorithm on 53 classes with 10 to 30 images per class.
@@
 Gradient Boosting Tree (GBT) seems to offer this compromise when compared to RF (model size) and SVM (computation complexity).
-For these reasons, we chose XGBoost [@Chen2016-xt], an optimized implementation of GBT, also known to be one of the best assets to win Kaggle's machine learning competitions.
+For these reasons, we chose XGBoost [@Chen2016-xt], an optimized implementation of GBT, also known to be one of the best assets to win [Kaggle](https://www.kaggle.com/)'s machine learning competitions.
 
-In this paper, we present the results of an embedded zooplankton recognition algorithm using XGBoost on UVP5 data.
-We show how we built the embedded recognition (ER) model under strict technical constraints, what were our compromises and what techniques we used to improve the classification.
+In this paper, we present the results of an embedded zooplankton classification algorithm using XGBoost on UVP5 data.
+We show how we built the embedded classification (EC) model under strict technical constraints, what were our compromises and what techniques we used to improve the classification.
 We also compare the performances of our features-based algorithm with a state-of-the-art neural network as if energy was not a limiting factor.
-Finally, we apply the method and follow the recommendations learnt from the preliminary UVP5 model to the UVP6 data set to build the ER model that will be included in the UVP6 for AUVs.
+Finally, we apply the method and follow the recommendations learnt from the preliminary UVP5 model to the UVP6 data set to build the EC model that will be included in the UVP6 for AUVs.
 
@@
 UVP5 Standard Definition (SD) profiles have been collected since 2008 at a global scale (@fig-map-of-profiles) whereas the UVP6 has been deployed very recently in specific locations and thus lacks spatial coverage to be fully representative.
-For this reason, UVP5 data will be used to create a preliminary ER model and explore ways of improvement.
-Ultimately, we will train the ER model for the UVP6 by applying the lessons learnt from the preliminary UVP5 model.
+For this reason, UVP5 data will be used to create a preliminary EC model and explore ways of improvement.
+Ultimately, we will train the EC model for the UVP6 by applying the lessons learnt from the preliminary UVP5 model.
 
@@
 
-The full UVP5 SD data set is composed of 3.578.901 validated images whereas the full UVP6 LP data set has 634.459 validated images.
+The full UVP5 SD data set is composed of 3.578.901 hand-annotated (i.e.
+labeled) images whereas the full UVP6 LP data set has 634.459 labeled images.
 Each image was segmented to find the main object (i.e.
@@
 pixel area, eccentricity), gray levels and statistical moments (see @tbl-features).
-For technical reasons, all UVP6 images with a pixel area below 79 pixels were removed (7% of images).
+All UVP6 images with a pixel area below 79 pixels were removed (7% of images) to fit with the size bins used in [EcoPart](https://ecopart.obs-vlfr.fr/), that is starting at 645 $\mu$m.
 
@@
 
-Validated objects were manually classified into specific taxa belonging to well known biological groups such as *Rhizaria*, *Copepoda* or *Cnidaria* to name a few.
-However, the maximum number of classes (also referred to as labels or taxa) for the ER was fixed to 40 in order to limit the size of the ER model and the volume of data that needs to be transmitted.
-We chose to aggregate classes iteratively to find the best number of classes for the ER.
+All objects were manually classified into specific taxa belonging to well known biological groups such as *Rhizaria*, *Copepoda* or *Cnidaria* to name a few.
+However, the maximum number of classes (also referred to as labels or taxa) for the EC was fixed to 40 in order to limit the size of the EC model and the volume of data that needs to be transmitted.
+We chose to aggregate classes iteratively to find the best number of classes for the EC.
 The first level (i.e.
@@
 The next level was used as a first aggregation step based on the number of objects in each class and the purity of the class.
-Below a given threshold, rare classes were merged with their most look-alike taxa and parents on the phylogenetic tree were removed when their children were abundant (i.e.
+Below a given threshold, rare classes were manually merged with their most look-alike taxa and parents on the phylogenetic tree were removed when their children were abundant (i.e.
 parents could not be classified to a deeper taxonomic level, likely causing confusion with their children during training).
@@
 Thus, classes were merged only if they were confused amongst the same biological group on the phylogenetic tree or if they had a similar ecological role.
-This involves that some classes end up with a variety of organisms that do not necessarily show similar morphological traits (e.g.
+This involves that some classes ended up with a variety of organisms that do not necessarily show similar morphological traits (e.g.
 other\<living).
@@
 
-To build the ER model, the first step was a threefold cross-validation using XGBoost parameterized with a set of given hyperparameters (i.e.
+To build the EC model, the first step consisted in a threefold cross-validation using XGBoost parameterized with a set of given hyperparameters (i.e.
 learning rate ($\eta$), maximum depth of a tree ($D_{max}$) and weight sensitivity ($\omega$)).
@@
 To cope with our unbalanced data set (more than 70% of detritus in the UVP5 data set), several options were available.
-We could artificially increase the number of biological data with the well known SMOTE oversampling technique [@Chawla2002-kg; @Fernandez2018-ft], decrease the number of detritus (random undersampling, keep only detritus close to unsupervised cluster centroids, etc.) or add a weight to each label during training to nudge the model towards under-represented classes (i.e.
+We could artificially increase the number of biological data with an oversampling technique such as SMOTE [@Chawla2002-kg; @Fernandez2018-ft] in order to oversample the minority classes, decrease the number of detritus (random undersampling, keep only detritus close to unsupervised cluster centroids, etc.) or add a weight to each label during training to nudge the model towards under-represented classes (i.e.
 biological classes).
@@
 
-In order to compare our features-based model with a state-of-the-art neural network not limited by energy constraints, we used a MobileNetV2 (depth multiplier 1.40) pre-trained on ImageNet.
+In order to compare our features-based model with a state-of-the-art neural network not limited by energy constraints, we used a MobileNetV2 (CNN designed to perform well on mobile devices) pre-trained on ImageNet, an image database with millions of hand-annotated images.
 This is known as transfer learning and it was already applied with success for zooplankton classification in @Orenstein2017-ya.
-Input images have a size of \[224, 224, 3\], the size of the features extractor vector is 1792 followed by a fully connected layer (600 nodes) with a dropout of 0.4 to avoid overfitting.
-The classification layer at the end of the network also has a dropout of 0.2.
+The CNN was fed with images of size 224 x 224 pixels, the size of the features extractor vector was 1792 and it was followed by a fully connected layer (600 nodes) with a dropout of 0.4 to avoid overfitting.
+The classification layer at the end of the network also had a dropout of 0.2.
 The CNN was trained with a decaying learning rate, a categorical cross entropy loss function and weights were defined with the same weight sensitivity of 0.25.
@@
 
-It is well known that the task of labeling images in supervised machine learning tasks is time consuming and most probably painful when you have millions of images to label.
+It is well known that the task of labeling images in supervised machine learning tasks is time consuming and most probably painful when you have millions of images to annotate.
 It is also very frustrating to see that one class (largely) dominates when you are interested in the small ones and that some of them are so scarce that you need to merge them with their siblings (at the very best) on the phylogenetic tree.
-Overall, we found - after multiple trials and errors - 11 classes fulfilling a first compromise between classification performances and ecological importance: 7 biological classes (*Acantharia*, *Aulacanthidae*, *Aulosphaeridae*, *Copepoda*, *Trichodesmium*, darksphere and other\<living) and 4 non-living classes (detritus, bubble, fiber and artefact).
+Overall, we found - after multiple trials and errors - 11 classes fulfilling a first compromise between classification performances and ecological importance: 7 biological classes (*Acantharia*, *Aulacanthidae*, *Aulosphaeridae*, *Copepoda*, *Trichodesmium*, darksphere (round black object, supposed to be part of the living) and other\<living) and 4 non-living classes (detritus, bubble, fiber and artefact).
 
@@
 The classification report (CR) gives the precision, the recall, the f1-score (harmonic mean of the precision and recall) and global metrics on the test set.
-As it can be seen in @tbl-classification-report-uvp5-xgboost-cnn, detritus have a recall of 86% thus because detritus are abundant (more than 230.000 in the test set), it results in more than 30.000 detritus wrongly labeled and therefore disseminated in all other classes.
-The confusion matrix (CM) helps to see where those wrongly labeled images are.
+As it can be seen in @tbl-classification-report-uvp5-xgboost-cnn, detritus have a recall of 86% thus because detritus are abundant (more than 230.000 in the test set), it results in more than 30.000 detritus wrongly predicted as such and therefore disseminated in all other classes.
+The confusion matrix (CM) helps to see where those false negatives have been predicted.
 @fig-confusion-matrix-uvp5 shows that a lot of Copepoda are confused with detritus as well as fiber.
@@
 The on board classification of plankton is definitely not an easy task.
-In order to improve the classification scores of biological classes, we tried several adjustments such as quantifying planktonic samples, splitting the detritus class into several subgroups and assigning weights to classes proportional to their size, which was the only option that enhanced the recognition of scarce classes (see Methods).
+In order to improve the classification scores of biological classes, we tried several adjustments such as quantifying planktonic samples, splitting the detritus class into several subgroups and assigning weights to classes proportional to their size, which was the only option that enhanced the classification of scarce classes (see Methods).
 
@@
 The SPAC showed the best estimations whereas TC drastically underestimated them (85% of data were lost using a probability threshold of 0.9).
-However, the aim of the ER is to be deployed on AUVs sampling vertical profiles in the water column therefore we applied the SPAC on 10.000 profile-like subsets and we summed all taxonomic abundances.
+However, the aim of the EC is to be deployed on AUVs sampling vertical profiles in the water column therefore we applied the SPAC on 10.000 profile-like subsets and we summed all taxonomic abundances.
 In that configuration, the SPAC overestimated the real total abundance by 1 to 15% depending on the projected (high or low) number of objects in each profile.
-These percentages were literally exploding for small classes in both projections (also with PAC) therefore quantification was discarded for use on autonomous profilers at this stage.
+These percentages were exponentially increasing for small classes in both projections (also with PAC) therefore quantification was discarded for use on autonomous profilers at this stage.
 
 Increasing the number of detritus classes was also considered in the hope of clustering them in the features space.
-Detritus display a variety of size, shape and gray levels so it must be possible to cluster them and therefore improve the classification score.
+Detritus display a variety of size, shape and gray levels so it must be possible to cluster them [@Durkin2021-nq; @Trudnowska2021-qn] and therefore improve the classification score.
 For this, we applied a classic K-Means algorithm [@Hartigan1979-oz] on detritus cleaned from likely outliers using an Isolation Forest Algorithm [@Liu2012-xz] and we executed a grid search as described earlier.
 @fig-grid-search clearly shows that using more than 1 class of detritus drastically decreases the global accuracy of the model, decreases the weighted living recall and only increases by 1% the weighted living precision.
-In addition, the confusion matrix (@fig-confusion-matrix-uvp6-ndetritus) shows that it does not improve the ER model because there is a high confusion between subgroups of detritus.
+In addition, the confusion matrix (@fig-confusion-matrix-uvp6-ndetritus) shows that it does not improve the EC model because there is a high confusion between subgroups of detritus.
 Essentially, it seems that detritus form a continuum in the features space and that more research should be done on the choice of features, clustering algorithms and related methods to better differentiate them in separate entities, which is not the purpose of this study.
@@
 As a result, even though *Calanoida* have been shown to overwinter in deep waters (\> 1000 m) [@Krumhansl2018-dr], an isolated signal at depth does not mean a copepod was really there.
-Similarly, *Trichodesmium* should not be observed outside the tropical and subtropical regions [@Westberry2006-bt] therefore depending on the sampling region, some results can directly be rejected.
-Here, tuff are predicted in higher concentrations than puff which are only predicted at one depth.
-Again, if one looks at the CR for these 2 classes, one can see that the precision and recall for tuff are, respectively, 18% and 15% below those of puff.
+Similarly, *Trichodesmium* are usually observed in the tropical and subtropical regions [@Westberry2006-bt] therefore depending on the sampling region, the presence/absence of specific organisms should be discussed.
 Finally, the interpretation of the detritus class is more straightforward because they have excellent metrics in the CR.
@@
 
-The ER of zooplankton on board AUVs is just at its beginning.
+The EC of zooplankton on board AUVs is just at its beginning.
 As expected, the CNN trained on UVP6 data (see @tbl-classification-report-uvp6-xgboost-cnn) outperforms the results of our features-based classifier.
-However, it should be noted that not all biological classes are better classified (e.g.
+However, it should be noted that not all biological classes are better classified with the CNN (e.g.
 *Actinopterygii*, *Appendicularia*, *Chaetognatha*, *Creseis*).
@@
 This is likely explained by the fact that we were able to keep more (smaller) classes for the UVP6.
-Nevertheless, previous tests on UVP5 data with more classes increased the gap between XGBoost and CNN weighted results therefore the latter better recognizes small classes.
-If energy and hardware constraints limit the performances of our current ER model, it is not forbidden to think about light-weight CNNs for future applications on AUVs to better understand the dynamics and the behavior of zooplankton.
+Nevertheless, previous tests on UVP5 data with more classes increased the gap between XGBoost and CNN weighted results therefore the latter better classifies small classes.
+If energy and hardware constraints limit the performances of our current EC model, it is not forbidden to think about light-weight CNNs for future applications on AUVs to better understand the dynamics and the behavior of zooplankton.
 
-![Scheme of the UVP6 pipeline with the processing done in the lab to create the ER model and the processing time (in ms) of the main steps when the UVP6 is deployed on board an AUV.
-kpx means kilo pixel.](image/uvp6/fig_pipeline_UVP6.jpg){#fig-uvp-pipeline}
+![Scheme of the UVP6 pipeline with the processing done in the lab to create the EC model and the processing time (in ms) of the main steps when the UVP6 is deployed on board an AUV.
+kpx, id and p means, respectively, kilo pixel, the numeric id of a given taxonomic class and the prediction probability returned by the classification model for that class.
+](image/uvp6/fig_pipeline_UVP6.jpg){#fig-uvp-pipeline}
 

### 2023-07-26 before-body.tex
clean repo

--- 2023-07-26 front_page.tex
@@
 }
-
-\pagenumbering{gobble}
 
@@
 
-\newpage
-\thispagestyle{empty}
-\null
-
-\section*{Foreword}
-
-As of 2022, it has now become obvious to the general public that the Earth's climate is changing.
-Massive flood events, repeated heatwaves, never-ending droughts, disappearing glaciers, biodiversity collapse, … Reports on weather anomalies were already published in the 1960's, the Intergovernmental Panel on Climate Change (IPCC) was created in 1988, the Kyoto Protocol was adopted in 1997, An Inconvenient Truth was released in 2006, the Paris Agreement was signed in 2015 and still, what has really changed?\\
-\\
-As I am writing this part, the COP27 (sponsored by Coca-Cola, no joke) is taking place with an acknowledged record number of fossil fuel lobbyists, bloody stadiums with outside air conditioning are about to be used in the desert for a month and it seems like 'leaders' of the Free World have been having hearing issues.
-See for yourself:\\
-\\
-\textit{"I am drowning"} said Pakistan\\
-\\
-\textit{"I'm sorry, what?"} said the Free World\\
-\\
-\textit{"I am starving"} said Ethiopia\\
-\\
-\textit{"Can't hear you, louder please!"} said the Free World\\
-\\
-\textit{"We are suffocating"} said India\\
-\\
-\textit{"Aaaah okay. Yes you're absolutely right, let's save the banks one more time"} said the Free World\\
-\\
-Although I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured that there is no fake news in this work. I hope you will enjoy it.
-
-\newpage
-\thispagestyle{empty}
-\null
-
-\newpage
-\thispagestyle{empty}
-\null
-
-\section*{Acknowledgements}
-
-I would like to thank the members of the jury for taking the time to read and evaluate this three and a half year work. I look forward to discussing it during the defense.\\
-\\
-Merci aux membres du comité de thèse pour leur temps et leurs retours sur l'évolution du projet.\\
-\\
-Cette thèse a été particulière à bien des égards, avec son lot de points positifs et négatifs mais surtout avec son lot de rebondissements qui m'ont permis d'affûter mes connaissances, ma résilience et je l'espère ma tolérance.
-Pour cela, je tiens à remercier chaleureusement toutes les personnes y ayant contribué de près ou de loin.\\
-\\
-Merci à Hervé, Lionel et Jean-Olivier pour leur accueil au Laboratoire de Villefranche-sur-Mer et pour leur disponibilité ainsi que pour leurs conseils avisés, tant scientifiques que personnels.\\
-\\
-Merci à Alex, Charles et Arthur pour leur aide et leur gentillesse.\\
-\\
-Merci également aux personnes que j'ai pu côtoyer à COMPLEX comme chez OMTAB depuis mon passage en stage en 2018 et les multiples allers-retours Belgique-France.\\
-\\
-Merci à Antoine Mangin de m'avoir prêté un bureau dans les locaux d'ACRI sur Grasse l'été dernier pendant que mon appartement servait d'Airbnb estival pour le plus grand bonheur du portefeuille de mon proprio.\\
-\\
-Un grand merci à mes colocs de bureau: Ophélie, Laetitia et Thelma pour votre soutien et votre bonne humeur sans faille (contrairement à moi).\\
-\\
-Merci à AMP (elle se reconnaitra), Chloé et Anaïs pour les pauses midi sur le ponton.
-AMP, merci d'avoir accepté de vivre virtuellement avec moi afin que la sécu daigne enfin me donner une carte vitale !\\
-\\
-Ne me demandez pas des noms de restos ou de bars sur la côte d'Azur, je n'en ai aucune idée.
-Par contre, je connais pas mal les routes du coin.
-Merci à tous ceux qui ont partagé (ou subi) des kilomètres à mes côtés, surtout en montée (le plat pays, qui est le mien): Flavien, Ophé, Lucas, Thelma, Laeti et Satoshi.\\
-\\
-Un grand merci à Marine et Alice pour leur bienveillance. Marine, je te souhaite le meilleur pour tes projets futurs mais fais attention à qui tu donnes tes clés quand même.\\
-\\
-Merci également à Alex qui a été un excellent roommate pendant les confinements, couvre-feux et ordonnances en tout genre. See you in London !\\
-\\
-Merci aux Liégeois(es) (Thomas, Florent, Flop, Nico, Evgeny et Lucie) et en particulier Hadrien qui a accepté de m'avoir comme coloc pendant plusieurs mois (et qui y a survécu).\\
-\\
-Merci aux Montois (Mathieu, Chris, Flo, Ju \& Ju) pour votre soutien à distance, un mariage en 3 actes et pour vos visites sur Nice.
-Petite dédicace à Ju K qui, à force d'avoir joué avec le diable en subissant l'enfer des pourcentages, va littéralement en devenir un.
-Vivement 2024 pour casser la baraque !\\
-\\
-Un très grand merci à ma famille nucléaire qui ne comprend toujours pas mon job mais qui fait semblant quand je lui explique, c'est le principal.
-C'est vrai que la défense tombe mal au niveau timing donc on se voit à la Ducasse :-)\\
-\\
-Enfin, un grand merci à une personne toute particulière qui en a probablement marre que je parle de la Belgique mais qui à présent est parfaite bilingue, mission accomplie?
-
-\newpage
-\thispagestyle{empty}
-\null
-
-\newpage
-\thispagestyle{empty}
-\null
-
-\section*{Summary}
-
-\textbf{Towards a new insight of the carbon transport in the global ocean}\\
-\\
-The ocean is known to play a key role in the carbon cycle.
-Without it, atmospheric CO$_{2}$ levels would be much higher than they are today thanks to the presence of carbon pumps that maintain a gradient of dissolved inorganic carbon (DIC) between the surface and the deep ocean.
-The biological carbon pump (BCP) is primarily responsible for this gradient.
-It consists in a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis in sunlit surface waters and then transported to the ocean interior and possibly the sediment where it will be sequestered from the atmosphere for millions of years.
-The BCP was long thought as solely the gravitational settling of particulate organic carbon (POC).
-However, a new paradigm for the BCP has recently been defined in which physically and biologically mediated particle injection pumps have been added to the original definition.
-Physically mediated particle injection pumps provide a pathway to better understand the transport of dissolved organic carbon (DOC) whereas biologically mediated particle injection pumps focus on the transport of POC by vertically migrating animals, either daily or seasonally.
-Therefore, a better understanding of these processes could help bridge the gap between carbon leaving the surface and carbon demand in the ocean interior.
-To address this new paradigm, this work will benefit from the advent of recent sensors that equip a new generation of Biogeochemical-Argo floats (BGC-Argo).
-The first part focuses on the development of an embedded zooplankton classification model for the Underwater Vision Profiler 6 (UVP6) under strict technical and energy constraints.
-The second part studies particle and carbon fluxes in the Labrador Sea using BGC-Argo floats equipped for the first time with the UVP6 and an optical sediment trap (OST), providing two independent measurements of sinking particles.
-The last part consists in revisiting the BCP using a new framework called CONVERSE for Continuous Vertical Sequestration.
-With this new approach, we re-evaluate the total carbon sequestered from the atmosphere ($\geq$ 100 years) by the BCP and its transport pathways on the entire water column, in contrast to the carbon sequestration typically assumed below a fixed reference depth.\\
-\\
-\textbf{Keywords} Carbon pumps -- Underwater Vision Profiler 6 -- BGC-Argo floats -- Machine learning
-
-\newpage
-
-\section*{Résumé}
-
-\textbf{Vers une meilleure compréhension du transport du carbone dans l'océan global}\\
-\\
-L'océan est connu pour jouer un rôle clé dans le cycle du carbone.
-Sans lui, les niveaux de CO$_{2}$ atmosphérique seraient bien plus élevés qu'aujourd'hui grâce à la présence de pompes à carbone qui maintiennent un gradient de carbone inorganique dissous (DIC) entre la surface et le fond de l'océan.
-La pompe à carbone biologique (BCP) est principalement responsable de ce gradient.
-Elle consiste en une série de processus océaniques au cours desquels le carbone inorganique est converti en matière organique via la photosynthèse dans les eaux de surface, puis transporté vers l'intérieur de l'océan et éventuellement les sédiments où il sera séquestré par rapport à l'atmosphère pour des millions d'années.
-La BCP était longtemps considérée comme étant uniquement la déposition gravitationnelle de carbone organique particulaire (POC).
-Cependant, un nouveau paradigme pour la BCP a récemment été défini dans lequel des pompes d'origine physique et biologique d'injection de particules ont été ajoutées à la définition originale.
-Les pompes physiques d'injection de particules fournissent un moyen de mieux comprendre le transport de carbone organique dissous (DOC), tandis que les pompes biologique d'injection de particules se concentrent sur le transport de POC par des animaux migrant verticalement, quotidiennement ou saisonnièrement.
-Par conséquent, une meilleure compréhension de ces processus pourrait aider à combler l'écart entre le carbone quittant la surface et la demande de carbone dans l'océan profond.
-Pour aborder ce nouveau paradigme, ce travail bénéficiera de l'arrivée de capteurs récents équipant une nouvelle génération de flotteurs Biogéochimiques-Argo (BGC-Argo).
-La première partie se concentre sur le développement d'un modèle embarqué de classification de zooplancton pour l'Underwater Vision Profiler 6 (UVP6) avec des contraintes techniques et énergétiques strictes.
-La deuxième partie étudie les flux de particules et de carbone dans la mer du Labrador en utilisant des flotteurs BGC-Argo équipés pour la première fois de l'UVP6 et d'un piège à sédiments optique (OST), fournissant deux mesures indépendantes des particules.
-La dernière partie consiste à revisiter la BCP en utilisant un nouveau cadre appelé CONVERSE qui fait référence à la séquestration verticale continue du carbone dans la colonne d'eau.
-Avec cette nouvelle approche, nous réévaluons le carbone total séquestré par rapport à l'atmosphère ($\geq$ 100 ans) par la BCP et ses voies de transport sur toute la colonne d'eau, par opposition à la séquestration de carbone généralement supposée en-dessous d'une profondeur de référence fixe.\\
-\\
-\textbf{Mots clés} Pompes à carbone -- Underwater Vision Profiler 6 -- Flotteurs BGC-Argo -- Machine learning
-
-\newpage
-\thispagestyle{empty}
-\null
-
 \end{document}
